### 2020-12-21 paper/main.tex
Added directory for paper draft tex files

\documentclass[sigconf]{acmart}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

\newcommand\myworries[1]{\textcolor{red}{#1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Inverting Colors to save my eye balls
%\usepackage{xcolor}
%\pagecolor[rgb]{0.1,0.12,0.15}
%\color[rgb]{.9,.9,.95}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\citestyle{acmauthoryear}

\begin{document}

\title{Shape Matching Element Method}

%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.
\begin{teaserfigure}
  \includegraphics[width=\textwidth]{castle.png}
  \caption{tmp}
  \Description{}
  \label{fig:teaser}
\end{teaserfigure}

\maketitle

\section{Introduction}

% \textbf{x}(\textbf{X}) = \hat{A}M(\textbf{X}) + \bar{\textbf{x}}$ \\
% $\hat{A} = \sum_i^{|S|}w_i(\textbf{X})A_i$

We have $\mathbf{x} = (x_0, x_1)$ and $\mathbf{X} = (X_0, X_1) $ as deformed and undeformed positions, respectively. \\

\textbf{<insert Exposition about FEM >} \\
In our case, the function we are trying to evaluate is deformation with respect to the center of mass. In VEM, instead of reconstructing values of our function using nodal values, we seek to find a a set of polynomial coefficients that map some arbitrary undeformed position to a polynomial approximation of its deformed position. This set of polynomial coefficients is referred to as the projection operator, $\mathbf{\Pi}$. The key point here is that $\mathbf{\Pi}$ is constructed strictly using positions defined on the boundary, avoiding the need for explicit representation of the interior. \\

\subsection{Shape Matching: }
In shape matching, authors seek a polynomial transformation, $\mathbf{\Pi}$ that minimizes $\sum_i ||\mathbf{\Pi}\mathbf{M(X)} - \mathbf{p_i}||^2$ where $\mathbf{M(X)}$ is the monomial basis of the $i$-th boundary vertex and $\mathbf{p_i}$ is current displacement from the center of mass for the $i$-th point. We argue this paper can be interpreted as a sort of virtual element method. \\

\subsection{Reconstructing deformed positions}
A general form for the monomial basis takes the following form <insert here>. For example, in 2D with quadratic deformation we have:
\begin{equation}
    \mathbf{M(X)} =\begin{bmatrix}
    X_0 - \bar{X}_0  \\ 
    X_1 - \bar{X}_1  \\ 
    (X_0 - \bar{X}_0)^2  \\ 
    (X_1 - \bar{X}_1)^2  \\ 
    (X_0 - \bar{X}_0)(X_1 - \bar{X}_1) 
    \end{bmatrix} 
\end{equation}

We write the estimated deformed position $\mathbf{x}$ of some arbitrary undeformed position $\mathbf{X}$ as
\begin{equation}
     \mathbf{x}(\mathbf{X}) = \hat{\mathbf{\Pi}}\mathbf{M}(\mathbf{X}) + \bar{\mathbf{x}} 
\end{equation}
where $\mathbf{\hat{\Pi}}$ is the blended projection operator for this particular point:
\begin{equation}
    \hat{\mathbf{\Pi}}\mathbf{(X)} = \sum_i^{|S|}w_i(\mathbf{X})\mathbf{\Pi}_i 
\end{equation}
The bar symbol indicates the center of mass. For example, $\bar{\mathbf{x}}$ is the deformed object's center of mass, which we compute as the mean of the boundary values 
\begin{equation}
    \bar{\mathbf{x}} = \frac{1}{n}\sum_i^n x_i 
\end{equation}
Here $w_i$ is some weighting function for the $i$-th shape. There are many possible forms for this weighting function. Right now we compute these weights by forming the following optimization: \\
<insert here>
In order to form the equations of motion, we need to re-express the above equation in terms of the nodal values, $\mathbf{q}$. To arrive it these we note that each $\mathbf{\Pi}_i$ is a function of $\mathbf{q}$. 
\begin{equation}
    \mathbf{\Pi}_i(\mathbf{q}) = \mathbf{P(E_i q)M(E_i q_0)^{\dagger}}
\end{equation}
\myworries{I'm not a fan of what I wrote here. The MLS-MPM paper presents a similar idea when talking about EFG, and I like their notation a bit better}
This is the shape matching problem's least squares solution where $M(q_0)^{\dagger}$ is the Moore-Penrose pseudo-inverse of the monomial vectors stacked row-wise (add example equation). The matrix $\mathbf{E}_i$ selects the degrees of freedom from $\mathbf{q}$ associated with the $i$-th shape. The function $\mathbf{P(q)}$ evaluates the displacement of these values from the current center of mass. We observe that the above equation is linear in $\mathbf{q}$ so we may rewrite the equation for $\mathbf{x}$ as 
\begin{equation}
    \mathbf{x(X)} = \mathbf{J_{\hat{\Pi}}(X)q}
\end{equation}

\subsection{Simulation on NURBs models}
We used the D-NURBS model (ref) for simulating on NURBs surfaces, giving an approach to simulating directly on the degrees of freedom of parametric NURBs models. The equation for a NURBs curve takes the form:
\begin{equation}
    \mathbf{x}(u) = \frac{\sum_{i=1}^{n} \mathbf{p}_i w_i B_{i,k}(u)}{ w_i B_{i,k}(u)}
\end{equation}
where $u$ is the parametric coordinate, $\mathbf{p}_i$ is the $i$-th control point with $n$ total control points. $B_{i,k}(u)$ is the B-spline basis function for the $i-th$ control point with degree $k-1$. NURBs curves represent a generalization of B-splines with the addition of the weight $w_i$ for each control point. In our case, we primarily work with NURBs surfaces, which are a generalization of the tensor-product of two B-splines:
\begin{equation}
    \mathbf{x}(u,v) = \frac{\sum_{i=1}^{n}\sum_{j=1}^{m}  \mathbf{p}_{i,j} w_{i,j} B_{i,k}(u)B_{j,l}(v)}{ w_{i,j} B_{i,k}(u)B_{j,l}(v)}
\end{equation}
where we now have two parameters $u,v$, and $m\cdot n$ control points. In D-NURBs, we see that $\mathbf{x}(u,v)$ is linear in the control points, $\mathbf{p}$, so we can rewrite the above equation as
\begin{equation}
    \mathbf{x}(u,v,\mathbf{p}) = \mathbf{J}(u,v)\mathbf{p}.
\end{equation}
Here $\mathbf{J}$ is the Jacobian matrix that maps the control points to some world space position on the surface for some $u,v$ pair. <elaborate on the form of J>. Now with these generalized coordinates, we may perform dynamics directly on the degrees of freedom of our NURBs models.

\subsection{Generalized Coordinates}
With D-NURBs simulation layered on the VEM simulation, we have two Jacobians: $\mathbf{J}(u,v)$ mapping controls points, $\mathbf{p}$ onto surface positions, and then $\mathbf{J_{\hat{\Pi}}(X)}$ which maps surface positions $\mathbf{q}$ to their deformed positions. The Jacobian $\mathbf{J}(u,v)$  may be precomputed, and each point sampled on a surface of a NURBs has an associated $\mathbf{J}(u,v)$. With these two Jacobians we may express a single $u,v$ pair's deformed position as
\begin{equation}
    \mathbf{x}(u,v,\mathbf{p}) = \mathbf{J_{\hat{\Pi}}(X)}\mathbf{J}(u,v)\mathbf{p}
\end{equation}
and the undeformed position
\begin{equation}
    \mathbf{X}(u,v) = \mathbf{J}(u,v)\mathbf{p_0}
\end{equation}
where $\mathbf{p_0}$ is the initial values for the control points.

\subsection{Deformation Gradient}
\begin{equation}
        \mathbf{F} = \frac{\partial \mathbf{x}}{\partial \mathbf{X}}
               = \frac{\partial \hat{\mathbf{\Pi}}\mathbf{M(X)}}{\partial \mathbf{X}}
               = \hat{\mathbf{\Pi}} \frac{\partial \mathbf{M(X)}}{\partial \mathbf{X}} 
\end{equation}
The $\frac{\partial \mathbf{M(X)}}{\partial \mathbf{X}}$ terms may be precomputed and using the previous $\mathbf{M(X)}$ example, we have
\begin{equation}
     \frac{\partial \mathbf{M(X)}}{\partial \mathbf{X}} = \begin{bmatrix}
     1&0 \\ 
     0&1 \\ 
     2(X_0 - \bar{X}_0)&0 \\ 
     0&2(X_1 - \bar{X}_1) \\ 
     (X_1 - \bar{X}_1)&(X_0 - \bar{X}_0) 
    \end{bmatrix}
\end{equation}

In computing the forces and the stiffness matrix, we require the gradient of the deformation gradient with respect to the configuration,  $\frac{\partial \mathbf{F}}{\partial \mathbf{q}}$. Usually the form of this gradient is simple. For example, in linear tetrahedral FEM each quadrature point's deformation map only involves the four surrounding nodal values, simplifying this gradient. In our case, the presence of the projection operator $\hat{\mathbf{\Pi}}$ means that each quadrature point is computed using the nodal values of the entire model.

\begin{equation}
    \frac{\partial \mathbf{F}}{\partial \mathbf{q}} = \frac{\partial \mathbf{P}}{\partial q}\mathbf{M(q_0)}^\dagger  \frac{\partial \mathbf{M(X)}}{\partial \mathbf{X}} 
\end{equation}

Naively handled, this gives us a $\frac{\partial \mathbf{F}}{\partial \mathbf{q}}$ of size $3 x |q|$. However, many of the columns of have values all near zero, allowing us to prune many columns for each quadrature point. Currently, I'm computing the infinity norm of each column and prune all except some fixed number of columns. \\

\subsection{Variational Mechanics}
We have now provided all the necessary ingredient to produce equations of motion for deformable bodies, which we derive via Euler-Lagrange equation. We begin with the Lagrangian
\begin{equation}
    L=T-V
\end{equation}
where $T$ is the kinetic energy, and $V$ is the potential energy. From the Euler-Lagrange equation we have:
\begin{equation}
    \frac{d}{dt} \frac{\partial L}{\partial \dot{\mathbf{p}}} + \frac{\partial L}{\partial {\mathbf{p}}} = 0
\end{equation}
\begin{equation}
    V =  \int_\Omega \psi(\mathbf{F(X)} dX
\end{equation}
then discretizing this over $m$ quadrature points, we have
\begin{equation}
    V =  \sum_i^m \psi(\mathbf{F(X_i)} A_i
\end{equation}
\begin{equation}
    \frac{\partial L}{\partial \mathbf{q}} = \sum_i^m \frac{\partial V_i}{\partial \mathbf{q}} = \sum_i^m \frac{\partial \psi(\mathbf{F_i)}}{\partial \mathbf{F}}\frac{\partial \mathbf{F_i}}{\partial \mathbf{q}}A_i
\end{equation}
Next we have the equation of the kinetic energy
\begin{equation}
    T = \frac{1}{2}\int_{ \Omega} \rho \dot{\mathbf{x}}^T\dot{\mathbf{x}} dV
\end{equation}
which we rewrite in terms of our control points configuration
\begin{equation}
    T = \frac{1}{2} \int_{\partial \Omega} \dot{\mathbf{p}}^T \mathbf{J}(u,v)^T \mathbf{M} \mathbf{J}(u,v) \dot{\mathbf{p}} du dv
\end{equation}
\myworries{change notation to not use M for the monomial basis}
where
\begin{equation}
    \mathbf{M} = \int_\Omega \rho \mathbf{J_{\hat{\Pi}}(X)}^T \mathbf{J_{\hat{\Pi}}(X)}
\end{equation}
From the potential energy (which uses dF/dq) and this potential energy, we discretize in time and right now I'm using linearly implicit backward euler

\subsection{Stability term}

\subsection{Projection Operator Weighting}
\begin{equation}
\begin{aligned}
\mathbf{w}(\mathbf{X}) = \min_{\mathbf{w}} \quad & \mathbf{w}^T \Theta(\mathbf{X}) \mathbf{w}    \\
\textrm{s.t.} \quad & 0 \leq w_i \leq 1                     \\
                    &   \sum_i^n w_i = 1                      \\
\end{aligned}
\end{equation}

where
\begin{equation}
     \Theta(\mathbf{X}) = \text{diag}\left( \frac{1}{\theta_1},\frac{1}{\theta_2},\dots,\frac{1}{\theta_N}\right)
\end{equation}

\section{Related Work}
\subsection{Shape Matching Related Papers}
\begin{enumerate}
    \item \textit{Shape Matching} \cite{10.1145/1073204.1073216}: Meshless simulation by fitting a polynomial describing the shapes deformation using only the nodal values.
    \item \textit{Lattice Shape Matching} \cite{10.1145/1275808.1276480}: Voxelize model to construct a lattice of cubes. Use these lattice cubes to construct overlapping shape matching regions (just like clustered shape matching?). The original mesh is deformed using trilinear interpolation of lattice vertex positions.
    \item \textit{Robust Real-Time Deformation of Incompressible Surface Meshes} \cite{10.1145/2019406.2019438}: Shape matching on trimeshes with overlapping regions (clustered shape matching). Adds an additional volume preservation constraint. Position based dynamics approach to satisfying volume preservation.
    \item \textit{Shape-Up: Shaping Discrete Geometry with Projections} \cite{10.1111/j.1467-8659.2012.03171.x}: Shape constraints by least squares fitting (like in shape matching). They have some "proximity function" indicating distance to least-squares fit, then uses projection operators to minimize proximity function (pretty much just shape matching). 
    \item \textit{Shape Matching with Oriented Particles} \cite{10.1145/1964921.1964987}: More general form for shape matching, permitting wider range of motion. Also they use shape matching projection operators for skinning, much like we do. For each skinning point, they specify weights with up to 4 particles (each with their own projection operator)
    \item \textit{Fast Adaptive Shape Matching Deformations} \cite{10.5555/1632592.1632606}: Essentially same thing as Lattice Shape Matching, but instead they use an octree instead of a basic voxel grid for shape matching. It's not super significant to mention this paper, but it does make clear that much of the followup work after shape matching never didn't emphasize it's utility as a meshless boundary only method. They kept converting it to a mesh-based method!
    \item \textit{A Geometric Deformation Model for Stable Cloth Simulation} \cite{10.2312/PE/vriphys/vriphys08/039-046} Shape matching for cloth simulation.
\end{enumerate}

\subsection{Other Meshless Methods in Graphics}
\begin{enumerate}
    \item \textit{Point Based Animation }\cite{10.1145/1028523.1028542}: Purely particle based, MLS to approximate derivatives. Appears to be among the earliest meshless methods in graphics based on continuum mechanics. 
    \item \textit{Position Based Dynamics} \cite{MULLER2007109}: Operates directly on particle positions by forming set of constraints and solving for particle positions that satisfy these constraints. Meshless
    \item \textit{Projective Dynamics} \cite{10.1145/2601097.2601116}: Similar to position based dynamics but solves the constraints implicitly by minimizing energy potentials. Mesh-based. Global solver unlike PBD which satisfies constraints locally using Gauss-Seidel.
\end{enumerate}

\subsection{Virtual Element and other Element Methods}
\begin{enumerate}
    \item \textit{Mimetic Finite Differences} \cite{10.1142/S0218202505000832} \cite{10.1016/j.jcp.2013.07.031} (they double dipped!): Considered a close relative to VEM and framed as the predecessor to VEM (in VEM papers). I still haven't read on MFD yet. From PolyDDF: "extension of finite volume and finite difference techniques to polygons that first discretizes a prime operator (typically, the gradient or the divergence) via a boundary integral, and then derives other operators by mimicking continuous structural properties."
    \item \textit{Basic principles of Virtual Element Method} \cite{10.1142/S0218202512500492} Original VEM paper
    \item \textit{The Hitchhiker's Guide to Virtual Element Method} \cite{10.1142/S021820251440003X}: More understandable versions of original VEM paper.
    \item \textit{Discrete Differential Operators on Polygonal Meshes} \cite{10.1145/3386569.3392389}: Extends VEM to do discrete differential geometry on arbitrary polygonal meshes.
    \item \textit{FLexible Simulation of Deformable Models using Discontinuous Galerkin FEM} \cite{10.5555/1632592.1632608}: Uses ordinary hexahedral elements, but uses a cut cell-based approach to support arbitrary polyhedra on the surface. 
    \item \textit{Generalizing the finite element method: Diffuse approximation and diffuse elements} \cite{Nayroles1992GeneralizingTF} Predecessor to Element Free Galerkin. FEM interpolation replaced with a local Moving Least Square interpolation.
    \item \textit{Element-free Galerkin methods} \cite{https://doi.org/10.1002/nme.1620370205}: similar to DEM, but more accurate gradients (not sure of all the differences). In MLS methods, they solve least squares for each particle in the domain, weighting nearby particles with a Gaussian-like density function. In contrast to us, we only compute least squares fitting on the boundary, and then precompute some weighting for each particle to the projection operators on the boundaries.
    \item \textit{Unified Simulation of Elastic Rods, Shells, and Solids} \cite{10.1145/1778765.1778776}: Propose Generalized moving least squares (GMLS) to resolve limitation of MLS shape functions that require many particles in the support of a point (that are not coplonar).  
\end{enumerate}
\subsection{Physics based Skinning}
\begin{enumerate}
    \item \textit{Skinning Siggraph Course} \cite{10.1145/2614028.2615427}: The linear weighting of polynomials at the exterior is very similar to skinning.
    \item \textit{Linear Subspace Design for Real-Time Shape Deformation} \cite{10.1145/2766952}: Linear deformation subspace that uses linear blend skinning and generalized barycentric coordinates. Similarly, we have "handles" but they are represented by entire NURBS patches, and our coordinates are the output of a polynomial, whereas barycentric coordinates are linear (I only skimmed this paper, not sure if this description is fair).
    \item \textit{Complementary dynamics} \cite{Zhang_2020}: Physics based skinning, orthogonality constraint can be seen as similar to our stability term (conformity term, error term, whatever it's called :))
    \item \textit{Physically-Based Character Skinning} \cite{10.2312/PE.vriphys.vriphys13.025-034}: Linear blend skinning with multiple layers of skin simulated via oriented particle shape matching and position based dynamics to enforce distance constraints (avoiding unwanted intersections).
\end{enumerate}
\subsection{Isogeometric Analysis}
\begin{enumerate}
    \item \textit{Isogeometric Analysis Book} \cite{10.1002/9780470749081.ch7}: The book everyone references when they write Isogeometric analysis in their papers.
    \item \textit{Dynamic NURBS} \cite{10.1145/176579.176580}: Outline of how to simulate on NURBS with the control points as the degrees of freedom. Method used in our work.
    \item \textit{XCAD} \cite{10.1145/3355089.3356576}: Optimize CAD models. CAD embedded in hexahedral mesh, complex integration strategy. Uncut hexahedral elements simulated ordinarily, cut elements use XFEM that add additional DOF to account for new element shapes.
    \item \textit{Development of a quadratic finite element formulation based on the XFEM and NURBS} \cite{https://doi.org/10.1002/nme.3120}: XFEM to handle curve surface integration of NURBs patches. Complex subdividing of "X-Elements" to produce cut cells along NURBS surfaces.
    \item \textit{A NURBS enhanced extended finite element approach for unfitted CAD analysis} \cite{10.1007/s00466-013-0854-7}: Pretty much the same as the quadratic, but allows higher-order approximation and better handling of interface.
    \item \textit{A NURBS-based interface-enriched generalized finite element method for problems with complex discontinuous gradient fields} \cite{https://doi.org/10.1002/nme.4852}: Similar to other XFEM NURBS approaches. Uses NURBS-based enrichment functions with cut cells. Additional DOFs added to handle discontinuities.
    \item \textit{A NURBS-based generalized finite element scheme for 3D simulation of heterogeneous materials} \cite{SAFDARI2016373}: Similar to previous "NIGFEM" paper above, but now in 3D.
    \item \textit{Swept Volume Parameterization for Isogeometric} \cite{10.1007/978-3-642-03596-8_2} To provide volumetric simulation of NURBS they introduce a new NURBS volume parameterization (B-Spline Volumes ... jesus christ)
    \item \textit{A finite volume method on NURBS geometries and its application in isogeometric fluid–structure interaction} \cite{HEINRICH20121645}: Combines NURBS paramaterization with finite volume method (requiring a mesh for the volume). 
    \item \textit{NURBS-Enhanced Finite Element Method (NEFEM)} \cite{10.1002/nme.2311}: Similar to the above example. They run an order FVM simulation and deal with the interface in a complicated manner (could this be considered XFEM?).
\end{enumerate}

\subsection{Quadrature}
\begin{enumerate}
    \item \textit{A new method for meshless integration in 2D and 3D Galerkin meshfree methods} \cite{KHOSRAVIFARD201030}: Strategy we use for integrating over CAD model volumes. Raycast along single dimension, find intersections, generate quadrature points in the intervals inside the object.
    \item Adaptive image-based intersection \cite{DBLP:journals/tog/WangFP12}: related to our meshless integration strategy in that we could use this to account for errors in the above approach (increase ray density where we estimate high error to be)
    \item Efficient and accurate numerical quadrature for immersed boundary methods \cite{10.1186/s40323-015-0031-y}: Finite Cell Method. "Immerses" a shape in a set of cells (mesh!) and computes quadrature over this. To handle curved surface they use an octree to subdivide to the desired level of accuracy.
    \item Higher-Order Finite Elements for Embedded Simulation \cite{10.1145/3414685.3417853}: Another Finite Cell method like the above, but with a new quadrature generation method (the ones with the circles in the triangles)
    \item \textit{Highly accurate surface and volume integration on implicit domains by means of moment-fitting} \cite{https://doi.org/10.1002/nme.4569} and \cite{https://doi.org/10.1002/nme.5343}: XCAD paper extends upon this method \myworries{still need to read these}
\end{enumerate}

\subsection{Misc}
\begin{enumerate}
    \item FEM simulation of 3D deformable solids \cite{10.1145/2343483.2343501}
    \item Fusion 360 Gallery \cite{willis2020fusion} : source of some models
\end{enumerate}
%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{references}

Dump of some equations
\begin{align*}
    \mathbf{M(X)} = (X_0 - \bar{X}_0, X_1 - \bar{X}_1)\\
    \bar{\mathbf{x}} = \frac{1}{n}\sum_i^n x_i \\
    \bar{\mathbf{X}} = \frac{1}{n}\sum_i^n X_i \\
\end{align*}

\end{document}

### 2020-12-21 matlab/nurbs_utils/nurbs_plot.m

@@ -1,4 +1,4 @@
-function nurbs = plot_nurbs(nurbs)
+function nurbs = nurbs_plot(nurbs)
     cm=jet(numel(nurbs));
     for ii = 1:numel(nurbs)
         xi = reshape(nurbs{ii}.x0, 3, nurbs{ii}.subd(1), nurbs{ii}.subd(2));
@@ -12,7 +12,7 @@ function nurbs = plot_nurbs(nurbs)
     lighting gouraud;
     %     shading interp
     
-%     lightangle(gca,72,-4)
+    lightangle(gca,0, 20)
 %         zlim([-70 150]);
     xlabel('x');
 end

### 2020-12-21 matlab/deprecated/silly_experiments/test_nurbs_cube_animation.m
Added script testing out visualization of VEM

@@ -1,0 +1,359 @@
+function test_nurbs_cube_animation
+    % Simulation parameters
+    dt = 0.005;      	% timestep
+    C = 0.5 * 17000;   	% Lame parameter 1
+    D = 0.5 * 150000;    % Lame parameter 2
+    gravity = -450;
+    k_error = 100000;
+    order = 2;
+    rho = 10;
+    save_output = 1;
+    save_obj = 0;
+    obj_res = 18;
+    
+    plot_mode = 2;
+    plot_modes = [repelem(0,100) repelem(1,600) repelem(2,500)];
+    patch_step = [repelem(0,100) repelem(1,100) repelem(2,100) repelem(3,100) repelem(4,100) repelem(5,100) repelem(6,100)];
+    % TODO --- use volume of quadrature points ... vol = volume(V,T);
+
+    % Read in NURBs 
+    fig=figure(1);
+    clf;
+    
+    
+    % Some files I test on
+    iges_file = 'rounded_cube.iges';
+
+    % Resolution indicates how many point samples we will take on each
+    % e.g. 6 means we have 6 samples in both the U & V coordinates, so
+    %      a total of 36 samples across the NURBs patch.
+    resolution = 6;
+    part=nurbs_from_iges(iges_file, resolution,0);
+    part=nurbs_plot(part);
+    
+    % Raycasting quadrature as described nowhere yet :)
+    V = raycast_quadrature(part, [6 6], 5)';
+    % plot3(V(1,:),V(2,:),V(3,:),'.','Color','r','MarkerSize',20);
+
+    x0 = zeros(3,0);
+    q_size = 0;
+    J_size = 0;
+    
+    % build global position vectors
+    for i=1:numel(part)
+        idx1=q_size+1;
+    	q_size = q_size + size(part{i}.p,1);
+        J_size = J_size + size(part{i}.J_flat,2) * size(part{i}.J_flat,3);
+        idx2=q_size;
+        
+        % indices into global configuration vector
+        part{i}.idx1=idx1;
+        part{i}.idx2=idx2;
+    end
+    
+    q = zeros(q_size,1);
+    J = zeros(J_size,q_size);
+    J_idx = [0 0];
+    for i=1:numel(part)
+        q(part{i}.idx1:part{i}.idx2,1) = part{i}.p;
+        Ji = part{i}.J_flat(:,:)';
+        
[…]

### 2020-12-21 matlab/quadrature/raycast_quadrature.m
Raycast quadrature return volume

@@ -1,4 +1,4 @@
-function X = raycast_quadrature(parts, ray_samples, samples)
+function [X, vol] = raycast_quadrature(parts, ray_samples, samples)
 % Note: assuming nurbs_plot has been called so that we can use the
 % plots data in part.plt
 
@@ -23,7 +23,11 @@ zrange = [OT.BinBoundaries(1,3) OT.BinBoundaries(1,6)];
                   linspace(zrange(1), zrange(2), ray_samples(2)));
 UV = [U(:) V(:)];      
 
+dA = ((yrange(2)-yrange(1))/ray_samples(1)) * ...
+     ((zrange(2)-zrange(1))/ray_samples(2));
+
 X = [];
+vol = [];
 for i = 1:size(UV,1)
     % Intersect along x-axis
     [dist, ~] = ray_intersect(xrange, UV(i,:), verts, face_bins, OT);
@@ -41,16 +45,20 @@ for i = 1:size(UV,1)
                 x_samples = linspace(dist(1,j),dist(2,j),samples);
                 pnts = xrange(2) - x_samples;
                 pnts = [pnts' UV_rep];
+                vol_j = (dist(2,j)-dist(1,j)) * dA;
                 X = [X; pnts]; % inefficient, yea
+                vol = [vol; repelem(vol_j, samples,1)];
             end
         else
             UV_rep = repelem(UV(i,:),numel(dist),1);
             pnts = xrange(2) - dist;
             pnts = [pnts UV_rep];
             X = [X; pnts];
+            vol = [vol; dA];
         end
         
     end    
 end
+X = X';
 
 end

### 2020-12-21 vem_nurbs.m

@@ -1,17 +1,15 @@
 function vem_nurbs
     % Simulation parameters
-    dt = 0.01;      	% timestep
-    C = 0.5 * 1700;   	% Lame parameter 1
-    D = 0.5 * 15000;    % Lame parameter 2
-    gravity = -100;
-    k_error = 100000;
-    order = 1;
-    rho = 1;
-    save_output = 0;
-    save_obj = 0;
-    obj_res = 18;
-
-    % TODO --- use volume of quadrature points ... vol = volume(V,T);
+    dt = 0.01;          % timestep
+    C = 0.5 * 17000;    % Lame parameter 1
+    D = 0.5 * 150000;   % Lame parameter 2
+    gravity = -500;     % gravity force (direction is -z direction)
+    k_error = 100000;   % stiffness for stability term
+    order = 1;          % (1 or 2) linear or quadratic deformation
+    rho = 1;            % per point density (currently constant)
+    save_output = 0;    % (0 or 1) whether to output images of simulation
+    save_obj = 0;       % (0 or 1) whether to output obj files
+    obj_res = 18;       % the amount of subdivision for the output obj
 
     % Read in NURBs 
     fig=figure(1);
@@ -19,26 +17,30 @@ function vem_nurbs
     
     % Some files I test on
     iges_file = 'rounded_cube.iges';
+    % iges_file = 'starship_nose.iges';
     % iges_file = 'puft_simple.iges';
     % iges_file = 'castle_simple.iges';
     % iges_file = 'rocket_with_nose.iges';
+    % iges_file = 'mug.iges';
     % iges_file = 'rocket.iges'; % this rocket doesn't have the nosecone
     
     % Resolution indicates how many point samples we will take on each
     % e.g. 6 means we have 6 samples in both the U & V coordinates, so
     %      a total of 36 samples across the NURBs patch.
-    resolution = 6;
-    part=nurbs_from_iges(iges_file, resolution,0);
+    resolution = 7;
+    
+    % NOTE: if you a warning saying "Matrix is singular", this means the
+    %       resolution is too low. I will fix this so that resolution is
+    %       set automatically on monday :)
+
[…]

### 2020-12-21 vem_sim_2d.m

@@ -8,7 +8,7 @@ function vem_sim_2d
     gravity = -10;      % gravity strength
     k_error = 100000;   % Stiffness for VEM stability term
     order = 1;          % (1 or 2) linear or quadratic deformation
-    rho = 1;            % mass
+    rho = 1;            % density
     save_output = 0;    % (0 or 1) whether to output images of simulation
     
     % Load a basic square mesh

### 2020-12-28 paper/main.tex
Raycast quadrature return volume
Added mock figures as well as pointform description of methods

@@
 \section{Introduction}
+
+A pointform description of the general idea
+\begin{itemize}
+	\item A D-NURBS approach to simulation. Use a Jacobian to map from UV positions in a NURBS parametric space to world space positions on the surface. 
+	\item \textbf{Shape Matching}: From the sample world space positions on a single NURBS surface, we solve a least squares problem to fit a \textit{projection operator} to the surface.
+This projection operator maps monomials in the undeformed space, to the least squares estimate of the deformed positions for the given monomials.
+	\item With DNURBS and Shape Matching VEM we have two mappings: a UV to undeformed world space with our NURBS Jacobian, then an undeformed to deformed mapping with the projection operators.
+This lets us form a set of generalized coordinates in terms of our control points of the NURBS surface (section 1.4).
+	\item For an arbitrary position in space (not necessarily on the surface), we can build a projection operator unique to this point via a weighted sum of the projection operators of the projection operators for each NURBS (eq 5).
+From this we can reconstruct an estimate of the deformed position from the monomial basis defined using its undeformed position.
+	\item For any point in space, we need to compute a set of weights to the projection operators defined on the surfaces with the following criteria: the weights sum to 1, are nonzero, and depend on the distances to the surfaces (See section 1.8).
+	\item With this mapping for undeformed to deformed positions, this yields a simple definition for the deformation gradient (eq 12).
+	\item Armed with our deformation gradient, we next define the Lagrangian, but this basic form isn't sufficient.
+We augment our kinetic and potential energies with a stability term (ref VEM).
+This accounts for cases in which the position of the nodal values don't match their estimated positions defined by the projection operator mapping.
+	\item \textbf{Raycasting Quadrature}: To compute the volume integrals we make use of raycasting quadrature approach (ref) where we uniform sample a YZ grid and shoot rays in the X direction.
+We then use the points of intersection to define integrations bounds for 1D integrals along these rays, which we integrate numerically.
+	\item With our augmented lagrangian, we use the Euler-Lagrange equation to arrive at equations of motion.
+Our generalized coordinates give us direct updates on the control points of the NURBS.
+Currently we are using linearly-implicit Euler for time integration.
+	\item \textbf{Handling trimmed NURBS}: For this case we update the original control points, but we must only sample points within the boundary.
+To compute these points we use an approach similar to the raycasting quadrature but in 2D.
+In the paremtric domain for a T-NURBS we sample uniforly in the V direction and shoot rays along the U axis.
+At each region of intersection, we sample more values along the ray.
+The final result is a set of UV coordinates that respect the boundary defined by NURBS curves.
+\end{itemize}
 
@@
 \subsection{Simulation on NURBs models}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=\linewidth]{figures/nurbs_surface}
+  \caption{The NURBs model uses control points, shown in grey, as the original degrees of freedom.}
+  \Description{A simple NURBS surface in $R^3$.}
+\end{figure}
+
 We used the D-NURBS model (ref) for simulating on NURBs surfaces, giving an approach to simulating directly on the degrees of freedom of parametric NURBs models.
@@
 \subsection{Stability term}
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=\linewidth]{figures/error_energy}
+  \caption{A figure showing the spring energy error for each sampled point on the NURBS curve. I should change the colors of the lines to show higher energies where the error is higher.}
+  \Description{Showing an example error of a NURBS curve.}
+\end{figure}
 
 \subsection{Projection Operator Weighting}
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=\linewidth]{figures/distance_weighting}
+  \caption{A 2D example showing how $\theta$ is computed for $\Pi_B$.
+$d_1$ shows the distance to the closest point on $\Pi_B$ to $\mathbf{x}$ and $d_2$ shows the distance to the surface hit by the ray in the opposite direction.}
+  \Description{Showing a 2D NURBS curve with a quadrature point in the center.
+A ray is cast to the nearest point on a shape, showing the distance between the quadrature point and the nearest point as well as the distance to the "back surface".}
+\end{figure}
+
 \begin{equation}
@@
 \end{equation}
+In computing the theta values, we require that they a approach 1 at the target surface and fall off to zero as a point becomes closer to other surfaces.
+To compute this for some surface, we shoot a ray from the point to the nearest point on the surface.
+Then in the opposite direction we find the nearest intersection to any other surfaces.
+We compute the theta with the following linear function
+
+\begin{equation}
+theta = \max (1.0 - \frac{d_1}{\min (d_2, D)}, 0.0)
+\end{equation}
+where $D$ is a fixed cut-off distance.
+
+\subsection{Integrating the volume}
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=\linewidth]{figures/raycasting_quadrature}
+  \caption{An example raycasting quadrature on a 2D NURBS curve.}
+  \Description{Shows a 2D NURBs curve with horizontal rays intersecting the curve. }
+\end{figure}
 
@@
 \begin{enumerate}
+	\item TRACKS: Toward Directable Thin Shells \cite{10.1145/1276377.1276439}: Petrov-Galerkin test functions for weak-form constraints that handles artifacts due to pointwise constraints.
     \item FEM simulation of 3D deformable solids \cite{10.1145/2343483.2343501}

### 2021-01-07 paper/main.tex
restructure, point form intro notes

--- 2020-12-21 paper/main.tex
@@
 
-\section{Introduction}
+\input{sections/intro}
+\input{sections/related}
+\input{sections/results}
+\input{sections/conclusion}
 
-% \textbf{x}(\textbf{X}) = \hat{A}M(\textbf{X}) + \bar{\textbf{x}}$ \\
-% $\hat{A} = \sum_i^{|S|}w_i(\textbf{X})A_i$
-
-We have $\mathbf{x} = (x_0, x_1)$ and $\mathbf{X} = (X_0, X_1) $ as deformed and undeformed positions, respectively. \\
-
-\textbf{<insert Exposition about FEM >} \\
-In our case, the function we are trying to evaluate is deformation with respect to the center of mass.
-In VEM, instead of reconstructing values of our function using nodal values, we seek to find a a set of polynomial coefficients that map some arbitrary undeformed position to a polynomial approximation of its deformed position.
-This set of polynomial coefficients is referred to as the projection operator, $\mathbf{\Pi}$.
-The key point here is that $\mathbf{\Pi}$ is constructed strictly using positions defined on the boundary, avoiding the need for explicit representation of the interior.
-\\
-
-\subsection{Shape Matching: }
-In shape matching, authors seek a polynomial transformation, $\mathbf{\Pi}$ that minimizes $\sum_i ||\mathbf{\Pi}\mathbf{M(X)} - \mathbf{p_i}||^2$ where $\mathbf{M(X)}$ is the monomial basis of the $i$-th boundary vertex and $\mathbf{p_i}$ is current displacement from the center of mass for the $i$-th point.
-We argue this paper can be interpreted as a sort of virtual element method.
-\\
-
-\subsection{Reconstructing deformed positions}
-A general form for the monomial basis takes the following form <insert here>. For example, in 2D with quadratic deformation we have:
-\begin{equation}
-    \mathbf{M(X)} =\begin{bmatrix}
-    X_0 - \bar{X}_0  \\ 
-    X_1 - \bar{X}_1  \\ 
-    (X_0 - \bar{X}_0)^2  \\ 
-    (X_1 - \bar{X}_1)^2  \\ 
-    (X_0 - \bar{X}_0)(X_1 - \bar{X}_1) 
-    \end{bmatrix} 
-\end{equation}
-
-We write the estimated deformed position $\mathbf{x}$ of some arbitrary undeformed position $\mathbf{X}$ as
-\begin{equation}
-     \mathbf{x}(\mathbf{X}) = \hat{\mathbf{\Pi}}\mathbf{M}(\mathbf{X}) + \bar{\mathbf{x}} 
-\end{equation}
-where $\mathbf{\hat{\Pi}}$ is the blended projection operator for this particular point:
-\begin{equation}
-    \hat{\mathbf{\Pi}}\mathbf{(X)} = \sum_i^{|S|}w_i(\mathbf{X})\mathbf{\Pi}_i 
-\end{equation}
-The bar symbol indicates the center of mass. For example, $\bar{\mathbf{x}}$ is the deformed object's center of mass, which we compute as the mean of the boundary values 
-\begin{equation}
-    \bar{\mathbf{x}} = \frac{1}{n}\sum_i^n x_i 
-\end{equation}
-Here $w_i$ is some weighting function for the $i$-th shape. There are many possible forms for this weighting function. Right now we compute these weights by forming the following optimization: \\
-<insert here>
-In order to form the equations of motion, we need to re-express the above equation in terms of the nodal values, $\mathbf{q}$.
-To arrive it these we note that each $\mathbf{\Pi}_i$ is a function of $\mathbf{q}$.
-\begin{equation}
-    \mathbf{\Pi}_i(\mathbf{q}) = \mathbf{P(E_i q)M(E_i q_0)^{\dagger}}
-\end{equation}
-\myworries{I'm not a fan of what I wrote here. The MLS-MPM paper presents a similar idea when talking about EFG, and I like their notation a bit better}
-This is the shape matching problem's least squares solution where $M(q_0)^{\dagger}$ is the Moore-Penrose pseudo-inverse of the monomial vectors stacked row-wise (add example equation).
-The matrix $\mathbf{E}_i$ selects the degrees of freedom from $\mathbf{q}$ associated with the $i$-th shape.
-The function $\mathbf{P(q)}$ evaluates the displacement of these values from the current center of mass.
-We observe that the above equation is linear in $\mathbf{q}$ so we may rewrite the equation for $\mathbf{x}$ as
-\begin{equation}
-    \mathbf{x(X)} = \mathbf{J_{\hat{\Pi}}(X)q}
-\end{equation}
-
-\subsection{Simulation on NURBs models}
-We used the D-NURBS model (ref) for simulating on NURBs surfaces, giving an approach to simulating directly on the degrees of freedom of parametric NURBs models.
-The equation for a NURBs curve takes the form:
-\begin{equation}
-    \mathbf{x}(u) = \frac{\sum_{i=1}^{n} \mathbf{p}_i w_i B_{i,k}(u)}{ w_i B_{i,k}(u)}
-\end{equation}
-where $u$ is the parametric coordinate, $\mathbf{p}_i$ is the $i$-th control point with $n$ total control points.
-$B_{i,k}(u)$ is the B-spline basis function for the $i-th$ control point with degree $k-1$.
-NURBs curves represent a generalization of B-splines with the addition of the weight $w_i$ for each control point.
-In our case, we primarily work with NURBs surfaces, which are a generalization of the tensor-product of two B-splines:
-\begin{equation}
-    \mathbf{x}(u,v) = \frac{\sum_{i=1}^{n}\sum_{j=1}^{m}  \mathbf{p}_{i,j} w_{i,j} B_{i,k}(u)B_{j,l}(v)}{ w_{i,j} B_{i,k}(u)B_{j,l}(v)}
-\end{equation}
-where we now have two parameters $u,v$, and $m\cdot n$ control points. In D-NURBs, we see that $\mathbf{x}(u,v)$ is linear in the control points, $\mathbf{p}$, so we can rewrite the above equation as
-\begin{equation}
-    \mathbf{x}(u,v,\mathbf{p}) = \mathbf{J}(u,v)\mathbf{p}.
-\end{equation}
-Here $\mathbf{J}$ is the Jacobian matrix that maps the control points to some world space position on the surface for some $u,v$ pair.
-<elaborate on the form of J>.
-Now with these generalized coordinates, we may perform dynamics directly on the degrees of freedom of our NURBs models.
-
-\subsection{Generalized Coordinates}
-With D-NURBs simulation layered on the VEM simulation, we have two Jacobians: $\mathbf{J}(u,v)$ mapping controls points, $\mathbf{p}$ onto surface positions, and then $\mathbf{J_{\hat{\Pi}}(X)}$ which maps surface positions $\mathbf{q}$ to their deformed positions.
-The Jacobian $\mathbf{J}(u,v)$  may be precomputed, and each point sampled on a surface of a NURBs has an associated $\mathbf{J}(u,v)$.
-With these two Jacobians we may express a single $u,v$ pair's deformed position as
-\begin{equation}
-    \mathbf{x}(u,v,\mathbf{p}) = \mathbf{J_{\hat{\Pi}}(X)}\mathbf{J}(u,v)\mathbf{p}
-\end{equation}
-and the undeformed position
-\begin{equation}
-    \mathbf{X}(u,v) = \mathbf{J}(u,v)\mathbf{p_0}
-\end{equation}
-where $\mathbf{p_0}$ is the initial values for the control points.
-
-\subsection{Deformation Gradient}
-\begin{equation}
-        \mathbf{F} = \frac{\partial \mathbf{x}}{\partial \mathbf{X}}
-               = \frac{\partial \hat{\mathbf{\Pi}}\mathbf{M(X)}}{\partial \mathbf{X}}
-               = \hat{\mathbf{\Pi}} \frac{\partial \mathbf{M(X)}}{\partial \mathbf{X}} 
-\end{equation}
-The $\frac{\partial \mathbf{M(X)}}{\partial \mathbf{X}}$ terms may be precomputed and using the previous $\mathbf{M(X)}$ example, we have
-\begin{equation}
-     \frac{\partial \mathbf{M(X)}}{\partial \mathbf{X}} = \begin{bmatrix}
-     1&0 \\ 
-     0&1 \\ 
-     2(X_0 - \bar{X}_0)&0 \\ 
-     0&2(X_1 - \bar{X}_1) \\ 
-     (X_1 - \bar{X}_1)&(X_0 - \bar{X}_0) 
-    \end{bmatrix}
-\end{equation}
-
-In computing the forces and the stiffness matrix, we require the gradient of the deformation gradient with respect to the configuration,  $\frac{\partial \mathbf{F}}{\partial \mathbf{q}}$.
-Usually the form of this gradient is simple.
-For example, in linear tetrahedral FEM each quadrature point's deformation map only involves the four surrounding nodal values, simplifying this gradient.
-In our case, the presence of the projection operator $\hat{\mathbf{\Pi}}$ means that each quadrature point is computed using the nodal values of the entire model.
-
-\begin{equation}
-    \frac{\partial \mathbf{F}}{\partial \mathbf{q}} = \frac{\partial \mathbf{P}}{\partial q}\mathbf{M(q_0)}^\dagger  \frac{\partial \mathbf{M(X)}}{\partial \mathbf{X}} 
-\end{equation}
-
-Naively handled, this gives us a $\frac{\partial \mathbf{F}}{\partial \mathbf{q}}$ of size $3 x |q|$.
-However, many of the columns of have values all near zero, allowing us to prune many columns for each quadrature point.
-Currently, I'm computing the infinity norm of each column and prune all except some fixed number of columns.
-\\
-
-\subsection{Variational Mechanics}
-We have now provided all the necessary ingredient to produce equations of motion for deformable bodies, which we derive via Euler-Lagrange equation. We begin with the Lagrangian
-\begin{equation}
-    L=T-V
-\end{equation}
-where $T$ is the kinetic energy, and $V$ is the potential energy. From the Euler-Lagrange equation we have:
-\begin{equation}
-    \frac{d}{dt} \frac{\partial L}{\partial \dot{\mathbf{p}}} + \frac{\partial L}{\partial {\mathbf{p}}} = 0
-\end{equation}
-\begin{equation}
-    V =  \int_\Omega \psi(\mathbf{F(X)} dX
-\end{equation}
-then discretizing this over $m$ quadrature points, we have
-\begin{equation}
-    V =  \sum_i^m \psi(\mathbf{F(X_i)} A_i
-\end{equation}
-\begin{equation}
-    \frac{\partial L}{\partial \mathbf{q}} = \sum_i^m \frac{\partial V_i}{\partial \mathbf{q}} = \sum_i^m \frac{\partial \psi(\mathbf{F_i)}}{\partial \mathbf{F}}\frac{\partial \mathbf{F_i}}{\partial \mathbf{q}}A_i
-\end{equation}
-Next we have the equation of the kinetic energy
-\begin{equation}
-    T = \frac{1}{2}\int_{ \Omega} \rho \dot{\mathbf{x}}^T\dot{\mathbf{x}} dV
-\end{equation}
-which we rewrite in terms of our control points configuration
-\begin{equation}
-    T = \frac{1}{2} \int_{\partial \Omega} \dot{\mathbf{p}}^T \mathbf{J}(u,v)^T \mathbf{M} \mathbf{J}(u,v) \dot{\mathbf{p}} du dv
-\end{equation}
-\myworries{change notation to not use M for the monomial basis}
-where
-\begin{equation}
-    \mathbf{M} = \int_\Omega \rho \mathbf{J_{\hat{\Pi}}(X)}^T \mathbf{J_{\hat{\Pi}}(X)}
-\end{equation}
-From the potential energy (which uses dF/dq) and this potential energy, we discretize in time and right now I'm using linearly implicit backward euler
-
-\subsection{Stability term}
-
-\subsection{Projection Operator Weighting}
-\begin{equation}
-\begin{aligned}
-\mathbf{w}(\mathbf{X}) = \min_{\mathbf{w}} \quad & \mathbf{w}^T \Theta(\mathbf{X}) \mathbf{w}    \\
-\textrm{s.t.} \quad & 0 \leq w_i \leq 1                     \\
-                    &   \sum_i^n w_i = 1                      \\
-\end{aligned}
-\end{equation}
-
-where
-\begin{equation}
-     \Theta(\mathbf{X}) = \text{diag}\left( \frac{1}{\theta_1},\frac{1}{\theta_2},\dots,\frac{1}{\theta_N}\right)
-\end{equation}
-
-\section{Related Work}
-\subsection{Shape Matching Related Papers}
-\begin{enumerate}
-    \item \textit{Shape Matching} \cite{10.1145/1073204.1073216}: Meshless simulation by fitting a polynomial describing the shapes deformation using only the nodal values.
-    \item \textit{Lattice Shape Matching} \cite{10.1145/1275808.1276480}: Voxelize model to construct a lattice of cubes.
-Use these lattice cubes to construct overlapping shape matching regions (just like clustered shape matching?).
-The original mesh is deformed using trilinear interpolation of lattice vertex positions.
-    \item \textit{Robust Real-Time Deformation of Incompressible Surface Meshes} \cite{10.1145/2019406.2019438}: Shape matching on trimeshes with overlapping regions (clustered shape matching).
-Adds an additional volume preservation constraint.
-Position based dynamics approach to satisfying volume preservation.
-    \item \textit{Shape-Up: Shaping Discrete Geometry with Projections} \cite{10.1111/j.1467-8659.2012.03171.x}: Shape constraints by least squares fitting (like in shape matching).
-They have some "proximity function" indicating distance to least-squares fit, then uses projection operators to minimize proximity function (pretty much just shape matching).
-    \item \textit{Shape Matching with Oriented Particles} \cite{10.1145/1964921.1964987}: More general form for shape matching, permitting wider range of motion.
-Also they use shape matching projection operators for skinning, much like we do.
-For each skinning point, they specify weights with up to 4 particles (each with their own projection operator)
-    \item \textit{Fast Adaptive Shape Matching Deformations} \cite{10.5555/1632592.1632606}: Essentially same thing as Lattice Shape Matching, but instead they use an octree instead of a basic voxel grid for shape matching.
-It's not super significant to mention this paper, but it does make clear that much of the followup work after shape matching never didn't emphasize it's utility as a meshless boundary only method.
-They kept converting it to a mesh-based method!
-    \item \textit{A Geometric Deformation Model for Stable Cloth Simulation} \cite{10.2312/PE/vriphys/vriphys08/039-046} Shape matching for cloth simulation.
-\end{enumerate}
-
-\subsection{Other Meshless Methods in Graphics}
-\begin{enumerate}
-    \item \textit{Point Based Animation }\cite{10.1145/1028523.1028542}: Purely particle based, MLS to approximate derivatives.
-Appears to be among the earliest meshless methods in graphics based on continuum mechanics.
-    \item \textit{Position Based Dynamics} \cite{MULLER2007109}: Operates directly on particle positions by forming set of constraints and solving for particle positions that satisfy these constraints.
-Meshless
-    \item \textit{Projective Dynamics} \cite{10.1145/2601097.2601116}: Similar to position based dynamics but solves the constraints implicitly by minimizing energy potentials.
-Mesh-based.
-Global solver unlike PBD which satisfies constraints locally using Gauss-Seidel.
-\end{enumerate}
-
-\subsection{Virtual Element and other Element Methods}
-\begin{enumerate}
-    \item \textit{Mimetic Finite Differences} \cite{10.1142/S0218202505000832} \cite{10.1016/j.jcp.2013.07.031} (they double dipped!): Considered a close relative to VEM and framed as the predecessor to VEM (in VEM papers).
-I still haven't read on MFD yet.
-From PolyDDF: "extension of finite volume and finite difference techniques to polygons that first discretizes a prime operator (typically, the gradient or the divergence) via a boundary integral, and then derives other operators by mimicking continuous structural properties."
-    \item \textit{Basic principles of Virtual Element Method} \cite{10.1142/S0218202512500492} Original VEM paper
-    \item \textit{The Hitchhiker's Guide to Virtual Element Method} \cite{10.1142/S021820251440003X}: More understandable versions of original VEM paper.
-    \item \textit{Discrete Differential Operators on Polygonal Meshes} \cite{10.1145/3386569.3392389}: Extends VEM to do discrete differential geometry on arbitrary polygonal meshes.
-    \item \textit{FLexible Simulation of Deformable Models using Discontinuous Galerkin FEM} \cite{10.5555/1632592.1632608}: Uses ordinary hexahedral elements, but uses a cut cell-based approach to support arbitrary polyhedra on the surface.
-    \item \textit{Generalizing the finite element method: Diffuse approximation and diffuse elements} \cite{Nayroles1992GeneralizingTF} Predecessor to Element Free Galerkin.
-FEM interpolation replaced with a local Moving Least Square interpolation.
-    \item \textit{Element-free Galerkin methods} \cite{https://doi.org/10.1002/nme.1620370205}: similar to DEM, but more accurate gradients (not sure of all the differences).
-In MLS methods, they solve least squares for each particle in the domain, weighting nearby particles with a Gaussian-like density function.
-In contrast to us, we only compute least squares fitting on the boundary, and then precompute some weighting for each particle to the projection operators on the boundaries.
-    \item \textit{Unified Simulation of Elastic Rods, Shells, and Solids} \cite{10.1145/1778765.1778776}: Propose Generalized moving least squares (GMLS) to resolve limitation of MLS shape functions that require many particles in the support of a point (that are not coplonar).
-\end{enumerate}
-\subsection{Physics based Skinning}
-\begin{enumerate}
-    \item \textit{Skinning Siggraph Course} \cite{10.1145/2614028.2615427}: The linear weighting of polynomials at the exterior is very similar to skinning.
-    \item \textit{Linear Subspace Design for Real-Time Shape Deformation} \cite{10.1145/2766952}: Linear deformation subspace that uses linear blend skinning and generalized barycentric coordinates.
-Similarly, we have "handles" but they are represented by entire NURBS patches, and our coordinates are the output of a polynomial, whereas barycentric coordinates are linear (I only skimmed this paper, not sure if this description is fair).
-    \item \textit{Complementary dynamics} \cite{Zhang_2020}: Physics based skinning, orthogonality constraint can be seen as similar to our stability term (conformity term, error term, whatever it's called :))
-    \item \textit{Physically-Based Character Skinning} \cite{10.2312/PE.vriphys.vriphys13.025-034}: Linear blend skinning with multiple layers of skin simulated via oriented particle shape matching and position based dynamics to enforce distance constraints (avoiding unwanted intersections).
-\end{enumerate}
-\subsection{Isogeometric Analysis}
-\begin{enumerate}
-    \item \textit{Isogeometric Analysis Book} \cite{10.1002/9780470749081.ch7}: The book everyone references when they write Isogeometric analysis in their papers.
-    \item \textit{Dynamic NURBS} \cite{10.1145/176579.176580}: Outline of how to simulate on NURBS with the control points as the degrees of freedom. Method used in our work.
-    \item \textit{XCAD} \cite{10.1145/3355089.3356576}: Optimize CAD models.
-CAD embedded in hexahedral mesh, complex integration strategy.
-Uncut hexahedral elements simulated ordinarily, cut elements use XFEM that add additional DOF to account for new element shapes.
-    \item \textit{Development of a quadratic finite element formulation based on the XFEM and NURBS} \cite{https://doi.org/10.1002/nme.3120}: XFEM to handle curve surface integration of NURBs patches.
-Complex subdividing of "X-Elements" to produce cut cells along NURBS surfaces.
-    \item \textit{A NURBS enhanced extended finite element approach for unfitted CAD analysis} \cite{10.1007/s00466-013-0854-7}: Pretty much the same as the quadratic, but allows higher-order approximation and better handling of interface.
-    \item \textit{A NURBS-based interface-enriched generalized finite element method for problems with complex discontinuous gradient fields} \cite{https://doi.org/10.1002/nme.4852}: Similar to other XFEM NURBS approaches.
-Uses NURBS-based enrichment functions with cut cells.
-Additional DOFs added to handle discontinuities.
-    \item \textit{A NURBS-based generalized finite element scheme for 3D simulation of heterogeneous materials} \cite{SAFDARI2016373}: Similar to previous "NIGFEM" paper above, but now in 3D.
-    \item \textit{Swept Volume Parameterization for Isogeometric} \cite{10.1007/978-3-642-03596-8_2} To provide volumetric simulation of NURBS they introduce a new NURBS volume parameterization (B-Spline Volumes ...
-jesus christ)
-    \item \textit{A finite volume method on NURBS geometries and its application in isogeometric fluid–structure interaction} \cite{HEINRICH20121645}: Combines NURBS paramaterization with finite volume method (requiring a mesh for the volume).
-    \item \textit{NURBS-Enhanced Finite Element Method (NEFEM)} \cite{10.1002/nme.2311}: Similar to the above example.
-They run an order FVM simulation and deal with the interface in a complicated manner (could this be considered XFEM?).
-\end{enumerate}
-
-\subsection{Quadrature}
-\begin{enumerate}
-    \item \textit{A new method for meshless integration in 2D and 3D Galerkin meshfree methods} \cite{KHOSRAVIFARD201030}: Strategy we use for integrating over CAD model volumes.
-Raycast along single dimension, find intersections, generate quadrature points in the intervals inside the object.
-    \item Adaptive image-based intersection \cite{DBLP:journals/tog/WangFP12}: related to our meshless integration strategy in that we could use this to account for errors in the above approach (increase ray density where we estimate high error to be)
-    \item Efficient and accurate numerical quadrature for immersed boundary methods \cite{10.1186/s40323-015-0031-y}: Finite Cell Method.
-"Immerses" a shape in a set of cells (mesh!) and computes quadrature over this.
-To handle curved surface they use an octree to subdivide to the desired level of accuracy.
-    \item Higher-Order Finite Elements for Embedded Simulation \cite{10.1145/3414685.3417853}: Another Finite Cell method like the above, but with a new quadrature generation method (the ones with the circles in the triangles)
-    \item \textit{Highly accurate surface and volume integration on implicit domains by means of moment-fitting} \cite{https://doi.org/10.1002/nme.4569} and \cite{https://doi.org/10.1002/nme.5343}: XCAD paper extends upon this method \myworries{still need to read these}
-\end{enumerate}
-
-\subsection{Misc}
-\begin{enumerate}
-    \item FEM simulation of 3D deformable solids \cite{10.1145/2343483.2343501}
-    \item Fusion 360 Gallery \cite{willis2020fusion} : source of some models
-\end{enumerate}
 %%

### 2021-01-07 paper/sections/intro.tex
restructure, point form intro notes

\section{Introduction}

% Modelling -> Animation -> Rendering
% Use physics for animation which requires conversion from modelling to simulation representation
% for volumetric objects this conversion invovles identififyin and meshing the interior of the modelled object and that's hard
% especially difficult if object surface is described by higher-order boundary representations. 
% Ideally we would have a more direct method of translating our surface model into a volumetric, simulatable form
%
% NURBs  are an oft-used higher order modelling primitive popular for CAD and Computer Graphics for which nice modelling tools exist
% Rhinocerous 3D and Fusion 360.
% Despite lots of work performing physics-based animation on NURBS models is hard
% robustly uilding a volumetric polygonal mesh that simulataneously maintains a corresonpondence to the model surface is an open-research problem
% Isogeometric approaches which directly use the NURBs primitives as simulation variables require additional background grids, complex cutcell algorithms
% and often must rely on soft boundary conditions.
%
% Want to say there is an advantage to avoiding constructing volumetric representations because it costs time and adds complexity
% that;s our goal to provide an alternative, the first truly meshless approach to IGA which maintains a tightt mapping to the boundary
% In this paper we present a new finite element method for isogeometric, physics-based animation of NURBs models.
% Crucially our method is boundary only, and requires no volumetric meshes (thus avoiding their inherent complications), grids or cutcells of any kind
% Our method does not need an explicit labelling of the inside/outide of the simulated model
% Does not require nurbs patches in an object to be explicitly joined
% Input is just the raw NURBs model with no additional information or annotations
% The resulting simulation algorithm is compatiable with nonlinear continuum mechanics material models as well as 
% standard methods for contact resolution. 
%
% crucially because we broadly target animation not necessarily simulation for engineering or manufacturing we don't require that models
% satisfy rigourous geometric requirements common for these applications
%
% Our approach relies on the recently developed Virtual Element Method for solving PDEs on domains tiled with arbitrary polygons.
% By establsihing the connection between VEM and the well-known shape matching simulation algorithm we derive equations of motion sfor 
% an arbitrary NURBS model using Lagrangian Mechanics. We avoid using volumetric data structures for integration by demonstrating that 
% raycasting provides a fast and reliable alternative. In combination, our contributions amount to the first algorithm for 
% isogeometric elastodynamics of volumetric structures that is truly meshless (introducing no auxilillary volumetric data structures) 
% which serves to close the considerable gap betweeen modelling and physics-based animation. 
%
% Desiderata for a sucessful algorithm for PBA 
% No volumetric mesh creation
% Support for Nonlinear Constitutive Models
% Support for standard Collision 
% DOFS on the Boundary (useful for boundary conditions, contact)
% Automatic determination of model volume
% Non-manufacturable models 
% Code Available

%limitations
% sometimes you want some volumetric dofs to better resolve internal behaviour

### 2021-01-07 paper/sections/methods.tex
restructure, point form intro notes

A pointform description of the general idea
\begin{itemize}
	\item A D-NURBS approach to simulation. Use a Jacobian to map from UV positions in a NURBS parametric space to world space positions on the surface. 
	\item \textbf{Shape Matching}: From the sample world space positions on a single NURBS surface, we solve a least squares problem to fit a \textit{projection operator} to the surface. This projection operator maps monomials in the undeformed space, to the least squares estimate of the deformed positions for the given monomials.
	\item With DNURBS and Shape Matching VEM we have two mappings: a UV to undeformed world space with our NURBS Jacobian, then an undeformed to deformed mapping with the projection operators. This lets us form a set of generalized coordinates in terms of our control points of the NURBS surface (section 1.4).
	\item For an arbitrary position in space (not necessarily on the surface), we can build a projection operator unique to this point via a weighted sum of the projection operators of the projection operators for each NURBS (eq 5). From this we can reconstruct an estimate of the deformed position from the monomial basis defined using its undeformed position.
	\item For any point in space, we need to compute a set of weights to the projection operators defined on the surfaces with the following criteria: the weights sum to 1, are nonzero, and depend on the distances to the surfaces (See section 1.8).
	\item With this mapping for undeformed to deformed positions, this yields a simple definition for the deformation gradient (eq 12).
	\item Armed with our deformation gradient, we next define the Lagrangian, but this basic form isn't sufficient. We augment our kinetic and potential energies with a stability term (ref VEM). This accounts for cases in which the position of the nodal values don't match their estimated positions defined by the projection operator mapping.
	\item \textbf{Raycasting Quadrature}: To compute the volume integrals we make use of raycasting quadrature approach (ref) where we uniform sample a YZ grid and shoot rays in the X direction. We then use the points of intersection to define integrations bounds for 1D integrals along these rays, which we integrate numerically.
	\item With our augmented lagrangian, we use the Euler-Lagrange equation to arrive at equations of motion. Our generalized coordinates give us direct updates on the control points of the NURBS. Currently we are using linearly-implicit Euler for time integration.
	\item \textbf{Handling trimmed NURBS}: For this case we update the original control points, but we must only sample points within the boundary. To compute these points we use an approach similar to the raycasting quadrature but in 2D. In the paremtric domain for a T-NURBS we sample uniforly in the V direction and shoot rays along the U axis. At each region of intersection, we sample more values along the ray. The final result is a set of UV coordinates that respect the boundary defined by NURBS curves.
\end{itemize}

% \textbf{x}(\textbf{X}) = \hat{A}M(\textbf{X}) + \bar{\textbf{x}}$ \\
% $\hat{A} = \sum_i^{|S|}w_i(\textbf{X})A_i$

We have $\mathbf{x} = (x_0, x_1)$ and $\mathbf{X} = (X_0, X_1) $ as deformed and undeformed positions, respectively. \\

\textbf{<insert Exposition about FEM >} \\
In our case, the function we are trying to evaluate is deformation with respect to the center of mass. In VEM, instead of reconstructing values of our function using nodal values, we seek to find a a set of polynomial coefficients that map some arbitrary undeformed position to a polynomial approximation of its deformed position. This set of polynomial coefficients is referred to as the projection operator, $\mathbf{\Pi}$. The key point here is that $\mathbf{\Pi}$ is constructed strictly using positions defined on the boundary, avoiding the need for explicit representation of the interior. \\

\subsection{Shape Matching: }
In shape matching, authors seek a polynomial transformation, $\mathbf{\Pi}$ that minimizes $\sum_i ||\mathbf{\Pi}\mathbf{M(X)} - \mathbf{p_i}||^2$ where $\mathbf{M(X)}$ is the monomial basis of the $i$-th boundary vertex and $\mathbf{p_i}$ is current displacement from the center of mass for the $i$-th point. We argue this paper can be interpreted as a sort of virtual element method. \\

\subsection{Reconstructing deformed positions}
A general form for the monomial basis takes the following form <insert here>. For example, in 2D with quadratic deformation we have:
\begin{equation}
    \mathbf{M(X)} =\begin{bmatrix}
    X_0 - \bar{X}_0  \\ 
    X_1 - \bar{X}_1  \\ 
    (X_0 - \bar{X}_0)^2  \\ 
    (X_1 - \bar{X}_1)^2  \\ 
    (X_0 - \bar{X}_0)(X_1 - \bar{X}_1) 
    \end{bmatrix} 
\end{equation}

We write the estimated deformed position $\mathbf{x}$ of some arbitrary undeformed position $\mathbf{X}$ as
\begin{equation}
     \mathbf{x}(\mathbf{X}) = \hat{\mathbf{\Pi}}\mathbf{M}(\mathbf{X}) + \bar{\mathbf{x}} 
\end{equation}
where $\mathbf{\hat{\Pi}}$ is the blended projection operator for this particular point:
\begin{equation}
    \hat{\mathbf{\Pi}}\mathbf{(X)} = \sum_i^{|S|}w_i(\mathbf{X})\mathbf{\Pi}_i 
\end{equation}
The bar symbol indicates the center of mass. For example, $\bar{\mathbf{x}}$ is the deformed object's center of mass, which we compute as the mean of the boundary values 
\begin{equation}
    \bar{\mathbf{x}} = \frac{1}{n}\sum_i^n x_i 
\end{equation}
Here $w_i$ is some weighting function for the $i$-th shape. There are many possible forms for this weighting function. Right now we compute these weights by forming the following optimization: \\
<insert here>
In order to form the equations of motion, we need to re-express the above equation in terms of the nodal values, $\mathbf{q}$. To arrive it these we note that each $\mathbf{\Pi}_i$ is a function of $\mathbf{q}$. 
\begin{equation}
    \mathbf{\Pi}_i(\mathbf{q}) = \mathbf{P(E_i q)M(E_i q_0)^{\dagger}}
\end{equation}
\myworries{I'm not a fan of what I wrote here. The MLS-MPM paper presents a similar idea when talking about EFG, and I like their notation a bit better}
This is the shape matching problem's least squares solution where $M(q_0)^{\dagger}$ is the Moore-Penrose pseudo-inverse of the monomial vectors stacked row-wise (add example equation). The matrix $\mathbf{E}_i$ selects the degrees of freedom from $\mathbf{q}$ associated with the $i$-th shape. The function $\mathbf{P(q)}$ evaluates the displacement of these values from the current center of mass. We observe that the above equation is linear in $\mathbf{q}$ so we may rewrite the equation for $\mathbf{x}$ as 
\begin{equation}
    \mathbf{x(X)} = \mathbf{J_{\hat{\Pi}}(X)q}
\end{equation}

\subsection{Simulation on NURBs models}

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{figures/nurbs_surface}
  \caption{The NURBs model uses control points, shown in grey, as the original degrees of freedom.}
  \Description{A simple NURBS surface in $R^3$.}
\end{figure}

We used the D-NURBS model (ref) for simulating on NURBs surfaces, giving an approach to simulating directly on the degrees of freedom of parametric NURBs models. The equation for a NURBs curve takes the form:
\begin{equation}
    \mathbf{x}(u) = \frac{\sum_{i=1}^{n} \mathbf{p}_i w_i B_{i,k}(u)}{ w_i B_{i,k}(u)}
\end{equation}
where $u$ is the parametric coordinate, $\mathbf{p}_i$ is the $i$-th control point with $n$ total control points. $B_{i,k}(u)$ is the B-spline basis function for the $i-th$ control point with degree $k-1$. NURBs curves represent a generalization of B-splines with the addition of the weight $w_i$ for each control point. In our case, we primarily work with NURBs surfaces, which are a generalization of the tensor-product of two B-splines:
\begin{equation}
    \mathbf{x}(u,v) = \frac{\sum_{i=1}^{n}\sum_{j=1}^{m}  \mathbf{p}_{i,j} w_{i,j} B_{i,k}(u)B_{j,l}(v)}{ w_{i,j} B_{i,k}(u)B_{j,l}(v)}
\end{equation}
where we now have two parameters $u,v$, and $m\cdot n$ control points. In D-NURBs, we see that $\mathbf{x}(u,v)$ is linear in the control points, $\mathbf{p}$, so we can rewrite the above equation as
\begin{equation}
    \mathbf{x}(u,v,\mathbf{p}) = \mathbf{J}(u,v)\mathbf{p}.
\end{equation}
Here $\mathbf{J}$ is the Jacobian matrix that maps the control points to some world space position on the surface for some $u,v$ pair. <elaborate on the form of J>. Now with these generalized coordinates, we may perform dynamics directly on the degrees of freedom of our NURBs models.

\subsection{Generalized Coordinates}
With D-NURBs simulation layered on the VEM simulation, we have two Jacobians: $\mathbf{J}(u,v)$ mapping controls points, $\mathbf{p}$ onto surface positions, and then $\mathbf{J_{\hat{\Pi}}(X)}$ which maps surface positions $\mathbf{q}$ to their deformed positions. The Jacobian $\mathbf{J}(u,v)$  may be precomputed, and each point sampled on a surface of a NURBs has an associated $\mathbf{J}(u,v)$. With these two Jacobians we may express a single $u,v$ pair's deformed position as
\begin{equation}
    \mathbf{x}(u,v,\mathbf{p}) = \mathbf{J_{\hat{\Pi}}(X)}\mathbf{J}(u,v)\mathbf{p}
\end{equation}
and the undeformed position
\begin{equation}
    \mathbf{X}(u,v) = \mathbf{J}(u,v)\mathbf{p_0}
\end{equation}
where $\mathbf{p_0}$ is the initial values for the control points.

\subsection{Deformation Gradient}
\begin{equation}
        \mathbf{F} = \frac{\partial \mathbf{x}}{\partial \mathbf{X}}
               = \frac{\partial \hat{\mathbf{\Pi}}\mathbf{M(X)}}{\partial \mathbf{X}}
               = \hat{\mathbf{\Pi}} \frac{\partial \mathbf{M(X)}}{\partial \mathbf{X}} 
\end{equation}
The $\frac{\partial \mathbf{M(X)}}{\partial \mathbf{X}}$ terms may be precomputed and using the previous $\mathbf{M(X)}$ example, we have
\begin{equation}
     \frac{\partial \mathbf{M(X)}}{\partial \mathbf{X}} = \begin{bmatrix}
     1&0 \\ 
     0&1 \\ 
     2(X_0 - \bar{X}_0)&0 \\ 
     0&2(X_1 - \bar{X}_1) \\ 
     (X_1 - \bar{X}_1)&(X_0 - \bar{X}_0) 
    \end{bmatrix}
\end{equation}

In computing the forces and the stiffness matrix, we require the gradient of the deformation gradient with respect to the configuration,  $\frac{\partial \mathbf{F}}{\partial \mathbf{q}}$. Usually the form of this gradient is simple. For example, in linear tetrahedral FEM each quadrature point's deformation map only involves the four surrounding nodal values, simplifying this gradient. In our case, the presence of the projection operator $\hat{\mathbf{\Pi}}$ means that each quadrature point is computed using the nodal values of the entire model.

\begin{equation}
    \frac{\partial \mathbf{F}}{\partial \mathbf{q}} = \frac{\partial \mathbf{P}}{\partial q}\mathbf{M(q_0)}^\dagger  \frac{\partial \mathbf{M(X)}}{\partial \mathbf{X}} 
\end{equation}

Naively handled, this gives us a $\frac{\partial \mathbf{F}}{\partial \mathbf{q}}$ of size $3 x |q|$. However, many of the columns of have values all near zero, allowing us to prune many columns for each quadrature point. Currently, I'm computing the infinity norm of each column and prune all except some fixed number of columns. \\

\subsection{Variational Mechanics}
We have now provided all the necessary ingredient to produce equations of motion for deformable bodies, which we derive via Euler-Lagrange equation. We begin with the Lagrangian
\begin{equation}
    L=T-V
\end{equation}
where $T$ is the kinetic energy, and $V$ is the potential energy. From the Euler-Lagrange equation we have:
\begin{equation}
    \frac{d}{dt} \frac{\partial L}{\partial \dot{\mathbf{p}}} + \frac{\partial L}{\partial {\mathbf{p}}} = 0
\end{equation}
\begin{equation}
    V =  \int_\Omega \psi(\mathbf{F(X)} dX
\end{equation}
then discretizing this over $m$ quadrature points, we have
\begin{equation}
    V =  \sum_i^m \psi(\mathbf{F(X_i)} A_i
\end{equation}
\begin{equation}
    \frac{\partial L}{\partial \mathbf{q}} = \sum_i^m \frac{\partial V_i}{\partial \mathbf{q}} = \sum_i^m \frac{\partial \psi(\mathbf{F_i)}}{\partial \mathbf{F}}\frac{\partial \mathbf{F_i}}{\partial \mathbf{q}}A_i
\end{equation}
Next we have the equation of the kinetic energy
\begin{equation}
    T = \frac{1}{2}\int_{ \Omega} \rho \dot{\mathbf{x}}^T\dot{\mathbf{x}} dV
\end{equation}
which we rewrite in terms of our control points configuration
\begin{equation}
    T = \frac{1}{2} \int_{\partial \Omega} \dot{\mathbf{p}}^T \mathbf{J}(u,v)^T \mathbf{M} \mathbf{J}(u,v) \dot{\mathbf{p}} du dv
\end{equation}
\myworries{change notation to not use M for the monomial basis}
where
\begin{equation}
    \mathbf{M} = \int_\Omega \rho \mathbf{J_{\hat{\Pi}}(X)}^T \mathbf{J_{\hat{\Pi}}(X)}
\end{equation}
From the potential energy (which uses dF/dq) and this potential energy, we discretize in time and right now I'm using linearly implicit backward euler

\subsection{Stability term}
\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{figures/error_energy}
  \caption{A figure showing the spring energy error for each sampled point on the NURBS curve. I should change the colors of the lines to show higher energies where the error is higher.}
  \Description{Showing an example error of a NURBS curve.}
\end{figure}

\subsection{Projection Operator Weighting}
\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{figures/distance_weighting}
  \caption{A 2D example showing how $\theta$ is computed for $\Pi_B$. $d_1$ shows the distance to the closest point on $\Pi_B$ to $\mathbf{x}$ and $d_2$ shows the distance to the surface hit by the ray in the opposite direction.}
  \Description{Showing a 2D NURBS curve with a quadrature point in the center. A ray is cast to the nearest point on a shape, showing the distance between the quadrature point and the nearest point as well as the distance to the "back surface".}
\end{figure}

\begin{equation}
\begin{aligned}
\mathbf{w}(\mathbf{X}) = \min_{\mathbf{w}} \quad & \mathbf{w}^T \Theta(\mathbf{X}) \mathbf{w}    \\
\textrm{s.t.} \quad & 0 \leq w_i \leq 1                     \\
                    &   \sum_i^n w_i = 1                      \\
\end{aligned}
\end{equation}

where
\begin{equation}
     \Theta(\mathbf{X}) = \text{diag}\left( \frac{1}{\theta_1},\frac{1}{\theta_2},\dots,\frac{1}{\theta_N}\right)
\end{equation}
In computing the theta values, we require that they a approach 1 at the target surface and fall off to zero as a point becomes closer to other surfaces. To compute this for some surface, we shoot a ray from the point to the nearest point on the surface.  Then in the opposite direction we find the nearest intersection to any other surfaces. We compute the theta with the following linear function

\begin{equation}
theta = \max (1.0 - \frac{d_1}{\min (d_2, D)}, 0.0)
\end{equation}
where $D$ is a fixed cut-off distance.

\subsection{Integrating the volume}
\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{figures/raycasting_quadrature}
  \caption{An example raycasting quadrature on a 2D NURBS curve.}
  \Description{Shows a 2D NURBs curve with horizontal rays intersecting the curve. }
\end{figure}

### 2021-01-07 paper/sections/related.tex
restructure, point form intro notes

\section{Related Work}
\subsection{Shape Matching Related Papers}
\begin{enumerate}
    \item \textit{Shape Matching} \cite{10.1145/1073204.1073216}: Meshless simulation by fitting a polynomial describing the shapes deformation using only the nodal values.
    \item \textit{Lattice Shape Matching} \cite{10.1145/1275808.1276480}: Voxelize model to construct a lattice of cubes. Use these lattice cubes to construct overlapping shape matching regions (just like clustered shape matching?). The original mesh is deformed using trilinear interpolation of lattice vertex positions.
    \item \textit{Robust Real-Time Deformation of Incompressible Surface Meshes} \cite{10.1145/2019406.2019438}: Shape matching on trimeshes with overlapping regions (clustered shape matching). Adds an additional volume preservation constraint. Position based dynamics approach to satisfying volume preservation.
    \item \textit{Shape-Up: Shaping Discrete Geometry with Projections} \cite{10.1111/j.1467-8659.2012.03171.x}: Shape constraints by least squares fitting (like in shape matching). They have some "proximity function" indicating distance to least-squares fit, then uses projection operators to minimize proximity function (pretty much just shape matching). 
    \item \textit{Shape Matching with Oriented Particles} \cite{10.1145/1964921.1964987}: More general form for shape matching, permitting wider range of motion. Also they use shape matching projection operators for skinning, much like we do. For each skinning point, they specify weights with up to 4 particles (each with their own projection operator)
    \item \textit{Fast Adaptive Shape Matching Deformations} \cite{10.5555/1632592.1632606}: Essentially same thing as Lattice Shape Matching, but instead they use an octree instead of a basic voxel grid for shape matching. It's not super significant to mention this paper, but it does make clear that much of the followup work after shape matching never didn't emphasize it's utility as a meshless boundary only method. They kept converting it to a mesh-based method!
    \item \textit{A Geometric Deformation Model for Stable Cloth Simulation} \cite{10.2312/PE/vriphys/vriphys08/039-046} Shape matching for cloth simulation.
\end{enumerate}

\subsection{Other Meshless Methods in Graphics}
\begin{enumerate}
    \item \textit{Point Based Animation }\cite{10.1145/1028523.1028542}: Purely particle based, MLS to approximate derivatives. Appears to be among the earliest meshless methods in graphics based on continuum mechanics. 
    \item \textit{Position Based Dynamics} \cite{MULLER2007109}: Operates directly on particle positions by forming set of constraints and solving for particle positions that satisfy these constraints. Meshless
    \item \textit{Projective Dynamics} \cite{10.1145/2601097.2601116}: Similar to position based dynamics but solves the constraints implicitly by minimizing energy potentials. Mesh-based. Global solver unlike PBD which satisfies constraints locally using Gauss-Seidel.
\end{enumerate}

\subsection{Virtual Element and other Element Methods}
\begin{enumerate}
    \item \textit{Mimetic Finite Differences} \cite{10.1142/S0218202505000832} \cite{10.1016/j.jcp.2013.07.031} (they double dipped!): Considered a close relative to VEM and framed as the predecessor to VEM (in VEM papers). I still haven't read on MFD yet. From PolyDDF: "extension of finite volume and finite difference techniques to polygons that first discretizes a prime operator (typically, the gradient or the divergence) via a boundary integral, and then derives other operators by mimicking continuous structural properties."
    \item \textit{Basic principles of Virtual Element Method} \cite{10.1142/S0218202512500492} Original VEM paper
    \item \textit{The Hitchhiker's Guide to Virtual Element Method} \cite{10.1142/S021820251440003X}: More understandable versions of original VEM paper.
    \item \textit{Discrete Differential Operators on Polygonal Meshes} \cite{10.1145/3386569.3392389}: Extends VEM to do discrete differential geometry on arbitrary polygonal meshes.
    \item \textit{FLexible Simulation of Deformable Models using Discontinuous Galerkin FEM} \cite{10.5555/1632592.1632608}: Uses ordinary hexahedral elements, but uses a cut cell-based approach to support arbitrary polyhedra on the surface. 
    \item \textit{Generalizing the finite element method: Diffuse approximation and diffuse elements} \cite{Nayroles1992GeneralizingTF} Predecessor to Element Free Galerkin. FEM interpolation replaced with a local Moving Least Square interpolation.
    \item \textit{Element-free Galerkin methods} \cite{https://doi.org/10.1002/nme.1620370205}: similar to DEM, but more accurate gradients (not sure of all the differences). In MLS methods, they solve least squares for each particle in the domain, weighting nearby particles with a Gaussian-like density function. In contrast to us, we only compute least squares fitting on the boundary, and then precompute some weighting for each particle to the projection operators on the boundaries.
    \item \textit{Unified Simulation of Elastic Rods, Shells, and Solids} \cite{10.1145/1778765.1778776}: Propose Generalized moving least squares (GMLS) to resolve limitation of MLS shape functions that require many particles in the support of a point (that are not coplonar).  
\end{enumerate}
\subsection{Physics based Skinning}
\begin{enumerate}
    \item \textit{Skinning Siggraph Course} \cite{10.1145/2614028.2615427}: The linear weighting of polynomials at the exterior is very similar to skinning.
    \item \textit{Linear Subspace Design for Real-Time Shape Deformation} \cite{10.1145/2766952}: Linear deformation subspace that uses linear blend skinning and generalized barycentric coordinates. Similarly, we have "handles" but they are represented by entire NURBS patches, and our coordinates are the output of a polynomial, whereas barycentric coordinates are linear (I only skimmed this paper, not sure if this description is fair).
    \item \textit{Complementary dynamics} \cite{Zhang_2020}: Physics based skinning, orthogonality constraint can be seen as similar to our stability term (conformity term, error term, whatever it's called :))
    \item \textit{Physically-Based Character Skinning} \cite{10.2312/PE.vriphys.vriphys13.025-034}: Linear blend skinning with multiple layers of skin simulated via oriented particle shape matching and position based dynamics to enforce distance constraints (avoiding unwanted intersections).
\end{enumerate}
\subsection{Isogeometric Analysis}
\begin{enumerate}
    \item \textit{Isogeometric Analysis Book} \cite{10.1002/9780470749081.ch7}: The book everyone references when they write Isogeometric analysis in their papers.
    \item \textit{Dynamic NURBS} \cite{10.1145/176579.176580}: Outline of how to simulate on NURBS with the control points as the degrees of freedom. Method used in our work.
    \item \textit{XCAD} \cite{10.1145/3355089.3356576}: Optimize CAD models. CAD embedded in hexahedral mesh, complex integration strategy. Uncut hexahedral elements simulated ordinarily, cut elements use XFEM that add additional DOF to account for new element shapes.
    \item \textit{Development of a quadratic finite element formulation based on the XFEM and NURBS} \cite{https://doi.org/10.1002/nme.3120}: XFEM to handle curve surface integration of NURBs patches. Complex subdividing of "X-Elements" to produce cut cells along NURBS surfaces.
    \item \textit{A NURBS enhanced extended finite element approach for unfitted CAD analysis} \cite{10.1007/s00466-013-0854-7}: Pretty much the same as the quadratic, but allows higher-order approximation and better handling of interface.
    \item \textit{A NURBS-based interface-enriched generalized finite element method for problems with complex discontinuous gradient fields} \cite{https://doi.org/10.1002/nme.4852}: Similar to other XFEM NURBS approaches. Uses NURBS-based enrichment functions with cut cells. Additional DOFs added to handle discontinuities.
    \item \textit{A NURBS-based generalized finite element scheme for 3D simulation of heterogeneous materials} \cite{SAFDARI2016373}: Similar to previous "NIGFEM" paper above, but now in 3D.
    \item \textit{Swept Volume Parameterization for Isogeometric} \cite{10.1007/978-3-642-03596-8_2} To provide volumetric simulation of NURBS they introduce a new NURBS volume parameterization (B-Spline Volumes ... jesus christ)
    \item \textit{A finite volume method on NURBS geometries and its application in isogeometric fluid–structure interaction} \cite{HEINRICH20121645}: Combines NURBS paramaterization with finite volume method (requiring a mesh for the volume). 
    \item \textit{NURBS-Enhanced Finite Element Method (NEFEM)} \cite{10.1002/nme.2311}: Similar to the above example. They run an order FVM simulation and deal with the interface in a complicated manner (could this be considered XFEM?).
\end{enumerate}

\subsection{Quadrature}
\begin{enumerate}
    \item \textit{A new method for meshless integration in 2D and 3D Galerkin meshfree methods} \cite{KHOSRAVIFARD201030}: Strategy we use for integrating over CAD model volumes. Raycast along single dimension, find intersections, generate quadrature points in the intervals inside the object.
    \item Adaptive image-based intersection \cite{DBLP:journals/tog/WangFP12}: related to our meshless integration strategy in that we could use this to account for errors in the above approach (increase ray density where we estimate high error to be)
    \item Efficient and accurate numerical quadrature for immersed boundary methods \cite{10.1186/s40323-015-0031-y}: Finite Cell Method. "Immerses" a shape in a set of cells (mesh!) and computes quadrature over this. To handle curved surface they use an octree to subdivide to the desired level of accuracy.
    \item Higher-Order Finite Elements for Embedded Simulation \cite{10.1145/3414685.3417853}: Another Finite Cell method like the above, but with a new quadrature generation method (the ones with the circles in the triangles)
    \item \textit{Highly accurate surface and volume integration on implicit domains by means of moment-fitting} \cite{https://doi.org/10.1002/nme.4569} and \cite{https://doi.org/10.1002/nme.5343}: XCAD paper extends upon this method \myworries{still need to read these}
\end{enumerate}

\subsection{Misc}
\begin{enumerate}
	\item TRACKS: Toward Directable Thin Shells \cite{10.1145/1276377.1276439}: Petrov-Galerkin test functions for weak-form constraints that handles artifacts due to pointwise constraints.
    \item FEM simulation of 3D deformable solids \cite{10.1145/2343483.2343501}
    \item Fusion 360 Gallery \cite{willis2020fusion} : source of some models
\end{enumerate}

### 2021-01-07 examples/vem_sim_2d_test_degenerate.m
added example script and updated shape matrix computation

@@ -1,0 +1,225 @@
+function vem_sim_2d_test_degenerate
+    % Example 2D VEM simulation.
+
+    % Simulation parameters
+    dt = 0.01;      	% timestep
+    C = 0.5 * 10;   	% Lame parameter 1
+    D = 0.5 * 150;   	% Lame parameter 2
+    gravity = -100;     % gravity strength
+    k_error = 100000;   % Stiffness for VEM stability term
+    order = 2;          % (1 or 2) linear or quadratic deformation
+    rho = 1;            % density
+	d = 2;              % dimension (2 or 3)
+    save_output = 0;    % (0 or 1) whether to output images of simulation
+    
+    % The number of elements in the monomial basis.
+    % -- example: Draft equation 1 is second order with k == 5
+    k = basis_size(d, order);
+
+    % Note: this is just on a simple box for testing purposes.
+    
+    % Shape matching modes:
+    mode = 'default';               % global solve with regular inverse
+    mode = 'global_pinv';           % global solve with pseudoinverse
+    mode = 'global_svd_truncated';  % global solve with truncated svd inv
+    
+    % number of points along left, bottom, right, and top edges,
+    % respectively
+    npts = [2 2 2 3]; 
+    
+    % Generate points along each line
+    left = [repelem(0, npts(1)); linspace(0, 2, npts(1))];
+    bot = [linspace(0, 2, npts(3)); repelem(0, npts(3))];
+    right = [repelem(2, npts(2)); linspace(0, 2, npts(2))];
+    top = [linspace(0, 2, npts(4)); repelem(2, npts(4))];
+    
+    % Global position vector
+    x0 = [left bot right top];
+    
+    % Creating "shapes. Each shape is a vector of indices into x.
+    idx = cumsum(npts);
+    E = {[1:idx(1)], ...        % left
+         [idx(1)+1:idx(2)], ... % bottom
+         [idx(2)+1:idx(3)], ... % right
+         [idx(3)+1:idx(4)]};    % top
+     
+    % You can also merge lines so that only one degenerate shape remains:
+    E = {[E{1} E{2} E{3}], E{4}};
+%     E = {[E{1} E{2} E{3} E{4}]};
+    
+    
+    num_verts = 10;
+    [X,Y] = meshgrid(linspace(0,2,num_verts), linspace(0,2,num_verts));
+    V = [X(:) Y(:)];
+
[…]

### 2021-01-07 matlab/compute_shape_matrices.m

@@ -1,55 +1,97 @@
-function [B,L] = compute_shape_matrices(x0, x0_com, E, order)
+function L = compute_shape_matrices(x0, x0_com, E, order, mode)
+    if nargin < 5
+        mode = 'default';
+    end
+    fprintf('Computing shape matrices in mode: %s', mode);
     d = size(x0,1); % dimension
 
     % Compute number of monomials
     k = basis_size(d, order);
-
-    Q = cell(numel(E),1);
-    B = cell(numel(E),1);
-
-    M_rows = 0;
+    
+    % Monomial basis vectors
+    M = cell(numel(E),1);
+    
+    % Computing size of A matrix
+    A_rows = 0;
+    row_ranges = cell(numel(E),1);
     for i=1:numel(E)
         x = x0(:,E{i});
-        Q{i} = monomial_basis(x, x0_com, order);
-        M_rows = M_rows + numel(E{i});
+        M{i} = monomial_basis(x, x0_com, order);
+        row_ranges{i} = A_rows+1:A_rows + d*numel(E{i});
+        A_rows = A_rows + numel(E{i});
     end
-    
-    M_cols = d*k*numel(E) + d;
-    M = zeros(d * M_rows, M_cols);
+    A_cols = d*k*numel(E) + d;
+    A = zeros(d * A_rows, A_cols);
 
     col_idx = 0;
     row_idx = 0;
+    % Forming global system matrix ([A S] matrix) 
     for i=1:numel(E)
         for j=1:numel(E{i})
             for n=1:d
                 col_start = col_idx + (n-1)*k + 1;
                 col_end   = col_idx + n*k;
-                M(row_idx+n, col_start:col_end) = Q{i}(:,j);
+                A(row_idx+n, col_start:col_end) = M{i}(:,j);
             end
-            M(row_idx+1:row_idx+d, end-d+1:end) = eye(d);
+            A(row_idx+1:row_idx+d, end-d+1:end) = eye(d); % Identity term
             row_idx = row_idx + d;
         end
         col_idx = col_idx + k * d;
     end
-
-    L = M'*M;
-    L(end-d+1:end,1:end-d) = 0;
-    L = L\M';
     
-%     L = inv(L);
-%     LM = M';
+    
+    fprintf('A matrix: nrows=%d ncols=%d rank=%d\n', ...
+        size(A,1), size(A,2), rank(A));
 
-    for i=1:numel(E)
-        % Monomial basis of DOF for the i-th shape.
-        x = x0(:,E{i});
[…]

### 2021-01-07 paper/sections/intro.tex
intro draft

@@
 \section{Introduction}
+
+The consumption of geometric surface models by physics-based animation algorithms is fraught with difficulty. 
+For volumetric objects, this process often involves identifying and discretizing the interior of the modelled object, 
+typically either as a tetrahedral or hexahedral mesh.%
+This procedure is both expensive and difficult, especially if the surface model is constructed from higher-order boundary
+representations, or if the volumetric discretization is required to be conforming or feature aligned. 
+Removing explicit volumetric discretization from the physics-based-animation pipeline can avoid these difficulties and 
+also provide a more unified modelling and simulation experience.
+
+NURBS (Non-uniform Rational B Splines) are a popular higher-order modelling primitive which 
+are used for computer-aided design (CAD), computational fabrication and computer animation. 
+NURBS primitives were the first geometric representation used for physics-based animation, yet,
+despite over three decades of research, animation of NURBS objects remains a challenge.
+
+Isogeometric Analysis (IGA) is a physics simulation methodology that uses the control variables of the NURBS model
+as the degrees-of-freedom (DOFS) of the simulation itself. 
+Unfortunately IGA approaches for volumetric objects still require background volumetric structures, typically
+regular grids that make satisfaction of boundary conditions difficult (which makes collision resolution difficult)
+or more complicated cut-cell grids which introduce non-trivial root finding problems into the mix. 
+Crucially, these simulation schemes typically assume models arise from engineering applications and meet tight geometric
+criteria such as that the mesh is manufacturable. 
+These inputs are much cleaner than those produced by a typical modeller.
+
+% In this paper we present a new finite element method for isogeometric, physics-based animation of NURBs models.
+% Crucially our method is boundary only, and requires no volumetric meshes (thus avoiding their inherent complications), grids or cutcells of any kind
+% Our method does not need an explicit labelling of the inside/outide of the simulated model
+% Does not require nurbs patches in an object to be explicitly joined
+% Input is just the raw NURBs model with no additional information or annotations
+% The resulting simulation algorithm is compatiable with nonlinear continuum mechanics material models as well as 
+% standard methods for contact resolution.
+
+We present the first truly meshless (no volumetric discretization generated) algorithm for direct, nonlinear elastodynamic simulation of NURBs models. 
+Our nonlinear elastodynamic simulation scheme requires only a boundary description of the object (we do not strictly require a solid model) and
+approriate physical parameters.  
+Because we explicitly use the NURBS boundary representation in the simulation its is straightforward to handle
+Dirichilet, Neumann boundary conditions and to apply contact resolution. 
+Crucially because we broadly target animation not necessarily simulation for engineering or manufacturing we don't require that models
+satisfy the rigourous geometric requirements common for these applications
+
+Our approach is an extension of the recently developed Virtual Element Method (VEM) for solving PDEs on domains tiled with arbitrary polygons.
+We establish a connection between VEM and the well-known shape matching simulation algorithm  which enables us 
+to derive equations of motion for an arbitrary NURBS model using Lagrangian Mechanics. 
+Importantly, we show how to replace volumetric data structures for integration with ray casting approaches which enables
+our meshless approach to isogeometric elastodynamic simulation of volumetric structures.
 

### 2021-01-08 examples/vem_sim_2d_test_degenerate.m
Added local shape matching with pinv for rank deficient systems

@@ -19,13 +19,15 @@ function vem_sim_2d_test_degenerate
     % Note: this is just on a simple box for testing purposes.
     
     % Shape matching modes:
-    mode = 'default';               % global solve with regular inverse
-    mode = 'global_pinv';           % global solve with pseudoinverse
-    mode = 'global_svd_truncated';  % global solve with truncated svd inv
+%     mode = 'global';               % global solve with regular inverse
+%     mode = 'global_pinv';           % global solve with pseudoinverse
+%     mode = 'global_svd_truncated';  % global solve with truncated svd inv
+%     mode = 'local';
+    mode = 'local_pinv';
     
-    % number of points along left, bottom, right, and top edges,
+    % num ber of points along left, bottom, right, and top edges,
     % respectively
-    npts = [2 2 2 3]; 
+    npts = [3 3 3 3]; 
     
     % Generate points along each line
     left = [repelem(0, npts(1)); linspace(0, 2, npts(1))];

### 2021-01-08 matlab/compute_shape_matrices.m

@@ -1,8 +1,8 @@
 function L = compute_shape_matrices(x0, x0_com, E, order, mode)
     if nargin < 5
-        mode = 'default';
+        mode = 'global';
     end
-    fprintf('Computing shape matrices in mode: %s', mode);
+    fprintf('Computing shape matrices in mode: %s \n', mode);
     d = size(x0,1); % dimension
 
     % Compute number of monomials
@@ -17,7 +17,7 @@ function L = compute_shape_matrices(x0, x0_com, E, order, mode)
     for i=1:numel(E)
         x = x0(:,E{i});
         M{i} = monomial_basis(x, x0_com, order);
-        row_ranges{i} = A_rows+1:A_rows + d*numel(E{i});
+        row_ranges{i} = d*A_rows+1:d*A_rows + d*numel(E{i});
         A_rows = A_rows + numel(E{i});
     end
     A_cols = d*k*numel(E) + d;
@@ -45,8 +45,8 @@ function L = compute_shape_matrices(x0, x0_com, E, order, mode)
 
     ATA = A'*A;
     ATA(end-d+1:end,1:end-d) = 0; % applying center of mass constraint
-
-    if strcmp(mode, 'default')
+    ATAinv = inv(ATA);
+    if strcmp(mode, 'global')
         L = ATA \ A';
         
     elseif strcmp(mode, 'global_pinv')
@@ -65,33 +65,50 @@ function L = compute_shape_matrices(x0, x0_com, E, order, mode)
         S = 1./ S(1:t);
         truncated_inv = V(:,1:t) * diag(S) * U(:,1:t)';
         L = truncated_inv * A';
+    elseif startsWith(mode,'local')
+        n = size(x0,2);
+        L = zeros(d*(k*numel(E)+1), numel(x0));
+
+        % Setting constant term rows
+        for i=1:d
+            L(end-d+i,i:d:end) = 1/n;
+        end
+
+    	for i=1:numel(E)
+            m = numel(E{i});
+
+            % Selecting subset of b values for this shape.
+            Sbi = zeros(d*m, numel(x0));
+            Sbi(:, row_ranges{i}) = eye(d*m);
+
+            % 'S' matrix (stacked identities)
+            Si = repmat(eye(d),m,1);
+
+            % Matrix that computes the mean of the nodal values.
+            Ti = repmat(eye(d)*(1/n),1,size(x0,2));
+
+            col_range = d*k*(i-1)+1:d*k*i;
+
+            % Getting the local system for this shape.
[…]

### 2021-01-08 paper/sections/related.tex
done intro draft

@@
 \section{Related Work}
+
+\begin{table*}[h]
+    \caption{Algorithms for simulation of NURBS models. 
+        %A $*$ indicates support for non-manifold meshes via preprocessing using Zhou et al.~\shortcite{Zhou:2016:MASG}. 
+    ``Not Done'' denotes features that are not demonstrated in publication but are theoretically achievable. ``?'' denotes that the feature was not demonstrated or discussed in literature.
+    \emph{Meshless} indicates whether the algorithm requires a volumetric mesh, 
+    \emph{Boundary}  indicates whether the algorithm uses boundary degrees-of-freedom directly in the simulation
+    \emph{Relaxed} indicates the algorithm supports relaxed modelling requirements and does not requiring fabrication quality inputs,
+    \emph{Code} describes whether or not there is a publicly available implementation of the method.}
+    \label{tbl:compare}
+    \begin{center}
+    \begin{tabular}{l c c c c c c c c}
+     \textbf{Algorithm} & \textbf{Meshless} & \textbf{Boundary} & \textbf{Relaxed} & \textbf{Code} \\
+     \hline 
+     \rowcolor[HTML]{DAE8FC} 
+     \textbf{X-CAD~\cite{}} & No (Cutcells) & No & No & No \\  
+     \textbf{Ours~\cite{}} & Yes & Yes & Yes & Planned \\  
+     \hline
+    \end{tabular}
+    \end{center}
+
+    \end{table*}
 \subsection{Shape Matching Related Papers}

### 2021-01-08 examples/vem_sim_2d_test_degenerate.m
local deficient inverse test with svd

@@ -24,6 +24,7 @@ function vem_sim_2d_test_degenerate
 %     mode = 'global_svd_truncated';  % global solve with truncated svd inv
 %     mode = 'local';
     mode = 'local_pinv';
+    mode = 'local_svd_truncated';
     
     % num ber of points along left, bottom, right, and top edges,
     % respectively
@@ -47,6 +48,7 @@ function vem_sim_2d_test_degenerate
      
     % You can also merge lines so that only one degenerate shape remains:
     E = {[E{1} E{2} E{3}], E{4}};
+%     E = {[E{1} E{2}], E{3}, E{4}};
 %     E = {[E{1} E{2} E{3} E{4}]};
     
     
@@ -55,7 +57,8 @@ function vem_sim_2d_test_degenerate
     V = [X(:) Y(:)];
 
     % min_I = find(x0(2,:) == max(x0(2,:))); % pin top side
-    min_I = find(x0(1,:) == min(x0(1,:))); % pin left side
+    % min_I = find(x0(1,:) == min(x0(1,:))); % pin left side
+    min_I = find(x0(2,:) == min(x0(2,:))); % pin left side
     % min_I = [1]; % Pinning the top left corner.
 
     % Initial deformed positions and velocities
@@ -219,7 +222,7 @@ function vem_sim_2d_test_degenerate
         drawnow
         
         if save_output
-            fn=sprintf('output_png\\100k_%03d.png',ii)
+            fn=sprintf('output\\img\\deficient_%03d.png',ii)
             saveas(fig,fn);
         end
         ii=ii+1;

### 2021-01-08 matlab/compute_shape_matrices.m

@@ -45,7 +45,7 @@ function L = compute_shape_matrices(x0, x0_com, E, order, mode)
 
     ATA = A'*A;
     ATA(end-d+1:end,1:end-d) = 0; % applying center of mass constraint
-    ATAinv = inv(ATA);
+
     if strcmp(mode, 'global')
         L = ATA \ A';
         
@@ -105,6 +105,18 @@ function L = compute_shape_matrices(x0, x0_com, E, order, mode)
                 else
                     ATA_inv = inv(Ai'*Ai);
                 end
+            elseif strcmp(mode, 'local_svd_truncated')
+                epsilon = 1e-12;
+                [U, S, V] = svd(Ai'*Ai);
+                S = diag(S);
+                t = find(abs(S) < epsilon, 1, 'first') - 1; %truncation point
+
+                if isempty(t) % full rank
+                    t = numel(S);
+                end
+
+                S = 1./ S(1:t);
+                ATA_inv = V(:,1:t) * diag(S) * U(:,1:t)';
             end
             Li = ATA_inv * Ai' * (Sbi - Si*Ti);
             L(col_range,:) = Li;

### 2021-01-10 examples/quadratic_beam.m
hierarchical fitting coefficients. Added beam example

@@ -1,0 +1,30 @@
+function quadratic_beam
+
+function pinned_ids = pin_function(x)
+    x_max = 1e-2;
+    y_min = 0.3;
+    y_max = 0.7;
+    z_min = 0.3;
+    z_max = 0.7;
+    z_min = -100.005;
+    z_max = 100.5;
+    pinned_ids = find(x(1,:) < x_max & x(2,:) > y_min & x(2,:) < y_max ...
+                    & x(3,:) > z_min & x(3,:) < z_max);
+end
+
+iges_file = 'beam.iges';
+
+part = nurbs_from_iges(iges_file);
+
+options.order = 2;
+options.pin_function = @pin_function;
+options.gravity = -10;
+options.lambda = 75;
+options.mu = 150;
+options.sample_interior = 0; 
+options.rho = 1;
+options.distance_cutoff = 1;
+
+vem_simulate_nurbs(part, options);
+
+end

### 2021-01-10 matlab/compute_shape_matrices.m

@@ -1,6 +1,6 @@
 function L = compute_shape_matrices(x0, x0_com, E, order, mode)
     if nargin < 5
-        mode = 'global';
+        mode = 'hierarchical';
     end
     fprintf('Computing shape matrices in mode: %s \n', mode);
     d = size(x0,1); % dimension
@@ -45,7 +45,70 @@ function L = compute_shape_matrices(x0, x0_com, E, order, mode)
 
     ATA = A'*A;
     ATA(end-d+1:end,1:end-d) = 0; % applying center of mass constraint
-    if strcmp(mode, 'global')
+
+    if strcmp(mode, 'hierarchical')
+        n = size(x0,2);
+
+        L = zeros(d*(k*numel(E)+1), numel(x0));
+
+        % Setting constant term rows
+        for i=1:d
+            L(end-d+i,i:d:end) = 1/n;
+        end
+
+        I = eye(numel(x0));
+
+        % 'S' matrix (stacked identities)
+        S = repmat(eye(d),n,1);
+
+        % Matrix that computes the mean of the nodal values.
+        T = repmat(eye(d)*(1/n),1,n);
+
+        k_linear = nchoosek(d+1-1,1);
+        linear_cols = zeros(k_linear*d*numel(E),1);
+        % Move this to function so it supports arbitrary orders
+        for i = 1:numel(E)
+            for j = 1:d
+                range_b = k_linear*d*(i-1) + k_linear*(j-1) + 1;
+                range_e = k_linear*d*(i-1) + k_linear*j;
+                i_range = range_b:range_e;
+
+                % (+1 should be + k_linear for order 2 or the cumulative
+                % sum of the previous k values for higher orders
+                col_start = k*d*(i-1)+ k*(j-1) + 1;
+                col_end = k*d*(i-1)+ k*(j-1) + k_linear;
+                col_range = col_start : col_end;
+                linear_cols(i_range) = col_range;
+            end
+        end
+
+        A_lin = A(:,linear_cols);
+        L_lin = (A_lin'*A_lin) \ (A_lin' * (I - S*T));
+        L(linear_cols,:) = L_lin;
+
+        if order == 2
+            k_quad = nchoosek(d+2-1,2);
+            quad_cols = zeros(k_quad*numel(E),1);
+
+            for i = 1:numel(E)
+                for j = 1:d
+                    range_b = k_quad*d*(i-1) + k_quad*(j-1) + 1;
+                    range_e = k_quad*d*(i-1) + k_quad*j;
[…]

### 2021-01-10 matlab/deprecated/silly_experiments/vem_2d_test_matching.m

@@ -1,6 +1,6 @@
 function vem_2d_test_matching
     % Simulation parameters
-    order = 2;          % (1 or 2) linear or quadratic deformation
+    order = 1;          % (1 or 2) linear or quadratic deformation
 	d = 2;              % dimension (2 or 3)
     save_output = 0;    % (0 or 1) whether to output images of simulation
     
@@ -8,19 +8,20 @@ function vem_2d_test_matching
     k = basis_size(1, order);
     
     % Shape matching modes:
-    mode = 'global';               % global solve with regular inverse
+%     mode = 'global';               % global solve with regular inverse
 %     mode = 'global_pinv';           % global solve with pseudoinverse
 %     mode = 'global_svd_truncated';  % global solve with truncated svd inv
 %     mode = 'local';
-%     mode = 'local_pinv';
+    mode = 'local_pinv';
 %     mode = 'local_svd_truncated';
     
     offset = 0;
     shape_func = @(a,n) ((a - offset).^n);
     
-    range = [-1 1];
-    samples = [10 10];
-    n = 2;
+    range = [0 1];
+    samples = [20 30];
+    n = 1;
+    n_end = 5;
     x = linspace(range(1), range(2), samples(1));
     y = shape_func(x,n);
     
@@ -34,12 +35,29 @@ function vem_2d_test_matching
     
     M = monomial_basis(x0, x0_com, order);
     
-    y0 = c(1:k)' * M + x0_com + c(end);
+    y0 = c(1:k)' * M + c(end);
     
     clf;
     plot(x,y);
     hold on;
     plot(x0,y0,'.','MarkerSize',20);
     axis equal;
+    
+    deformations=100;
+    n_vals = linspace(n,n_end, deformations);
+    
+    
+    for i=1:deformations
+        clf;
+        y_new = shape_func(x, n_vals(i));
+        c = L * y_new(:);
+        y_pred = c(1:k)' * M + c(end);
+        plot(x,y_new);
+        hold on;
+        plot(x0,y_pred,'.','MarkerSize',20);
+        axis equal;
+        legend('Deformed Positions', 'Predicted positions','Location','northwest');
+        drawnow;
+    end
        
-kend
+end

### 2021-01-10 matlab/deprecated/silly_experiments/vem_sim_2d_test_lowrank.m

@@ -3,11 +3,11 @@ function vem_sim_2d_test_degenerate
 
     % Simulation parameters
     dt = 0.01;      	% timestep
-    C = 0.5 * 100;   	% Lame parameter 1
-    D = 0.5 * 150;   	% Lame parameter 2
-    gravity = -100;     % gravity strength
+    C = 0.5 * 5;   	% Lame parameter 1
+    D = 0.5 * 10;   	% Lame parameter 2
+    gravity = -500;     % gravity strength
     k_error = 100000;   % Stiffness for VEM stability term
-    order = 2;          % (1 or 2) linear or quadratic deformation
+    order = 1;          % (1 or 2) linear or quadratic deformation
     rho = 1;            % density
 	d = 2;              % dimension (2 or 3)
     save_output = 0;    % (0 or 1) whether to output images of simulation
@@ -19,19 +19,20 @@ function vem_sim_2d_test_degenerate
     % Note: this is just on a simple box for testing purposes.
     
     % Shape matching modes:
-    mode = 'global';               % global solve with regular inverse
+%     mode = 'global';               % global solve with regular inverse
 %     mode = 'global_pinv';           % global solve with pseudoinverse
 %     mode = 'global_svd_truncated';  % global solve with truncated svd inv
 %     mode = 'local';
 %     mode = 'local_pinv';
 %     mode = 'local_svd_truncated';
+    mode = 'hierarchical';
     
     % num ber of points along left, bottom, right, and top edges,
     % respectively
-    npts = [3 3 3 3]; 
+    npts = [8 8 8 8]; 
     
     l1 = 1; % height
-    l2 = 4; % width
+    l2 = 2; % width
     
     % Generate points along each line
     left = [repelem(0, npts(1)); linspace(0, l1, npts(1))];
@@ -50,20 +51,23 @@ function vem_sim_2d_test_degenerate
          [idx(3)+1:idx(4)]};    % top
      
     % You can also merge lines so that only one degenerate shape remains:
+%     x0 = [left top right bot];
 %     E = {[E{1} E{2} E{3}], E{4}};
-%     E = {[E{1} E{4} E{3}], E{2}};
[…]

### 2021-01-10 matlab/vem_simulate_nurbs.m

@@ -68,7 +68,7 @@ function vem_simulate_nurbs(parts, varargin)
     x0_com = mean(x0,2);
         
     % Shape Matrices
-    [~,L] = compute_shape_matrices(x0, x0_com, E, config.order);
+    L = compute_shape_matrices(x0, x0_com, E, config.order);
     
     % Build Monomial bases for all quadrature points
     Y = monomial_basis_matrix(V, x0_com, config.order, k);

### 2021-01-10 vem_nurbs.m

@@ -13,7 +13,7 @@ function vem_nurbs
     k_error = 100000;   % stiffness for stability term
     order = 2;          % (1 or 2) linear or quadratic deformation
     rho = .1;           % per point density (currently constant)
-    save_output = 1;    % (0 or 1) whether to output images of simulation
+    save_output = 0;    % (0 or 1) whether to output images of simulation
     save_obj = 0;       % (0 or 1) whether to output obj files
     d = 3;              % dimension (2 or 3)
 
@@ -75,7 +75,7 @@ function vem_nurbs
     % Shape Matrices
     %E=cell(1);
     %E{1}=1:size(x0,2);
-    [~,L] = compute_shape_matrices(x0, x0_com, E, order);
+    L = compute_shape_matrices(x0, x0_com, E, order);
     
     % Build Monomial bases for all quadrature points
     Y = monomial_basis_matrix(V, x0_com, order, k);

### 2021-01-10 examples/quadratic_beam.m
update beam parameters

@@ -19,11 +19,11 @@ part = nurbs_from_iges(iges_file);
 options.order = 2;
 options.pin_function = @pin_function;
 options.gravity = -10;
-options.lambda = 75;
-options.mu = 150;
+options.lambda = 100;
+options.mu = 1500;
 options.sample_interior = 0; 
 options.rho = 1;
-options.distance_cutoff = 1;
+options.distance_cutoff = 0.2;
 
 vem_simulate_nurbs(part, options);
 

### 2021-01-10 paper/sections/intro.tex

@@
 Importantly, we show how to replace volumetric data structures for integration with ray casting approaches which enables
-our meshless approach to isogeometric elastodynamic simulation of volumetric structures.
+our meshless approach to isogeometric elastodynamic physically-based animation of volumetric structures.
 

### 2021-01-10 examples/quadratic_beam.m
merging with main

@@ -24,6 +24,7 @@ options.mu = 1500;
 options.sample_interior = 0; 
 options.rho = 1;
 options.distance_cutoff = 0.2;
+options.fitting_mode = 'hierarchical';
 
 vem_simulate_nurbs(part, options);
 

### 2021-01-10 matlab/compute_shape_matrices.m

@@ -1,6 +1,6 @@
 function L = compute_shape_matrices(x0, x0_com, E, order, mode)
     if nargin < 5
-        mode = 'hierarchical';
+        mode = 'global';
     end
     fprintf('Computing shape matrices in mode: %s \n', mode);
     d = size(x0,1); % dimension

### 2021-01-10 matlab/vem_simulate_nurbs.m

@@ -15,6 +15,7 @@ function vem_simulate_nurbs(parts, varargin)
     addParameter(p, 'sample_interior', 1);
     addParameter(p, 'distance_cutoff', 20);
     addParameter(p, 'enable_secondary_rays', true);
+    addParameter(p, 'fitting_mode', 'global');
     parse(p,varargin{:});
     config = p.Results;
     
@@ -69,7 +70,7 @@ function vem_simulate_nurbs(parts, varargin)
     x0_com = mean(x0,2);
         
     % Shape Matrices
-    L = compute_shape_matrices(x0, x0_com, E, config.order);
+    L = compute_shape_matrices(x0, x0_com, E, config.order, config.fitting_mode);
     
     % Build Monomial bases for all quadrature points
     Y = monomial_basis_matrix(V, x0_com, config.order, k);

### 2021-01-10 vem_nurbs.m

@@ -75,7 +75,7 @@ function vem_nurbs
     % Shape Matrices
     %E=cell(1);
     %E{1}=1:size(x0,2);
-    L = compute_shape_matrices(x0, x0_com, E, order);
+    L = compute_shape_matrices(x0, x0_com, E, order, 'global');
     
     % Build Monomial bases for all quadrature points
     Y = monomial_basis_matrix(V, x0_com, order, k);

### 2021-01-10 vem_nurbs_with_collision.m

@@ -76,7 +76,7 @@ function vem_nurbs
     % Shape Matrices
     %E=cell(1);
     %E{1}=1:size(x0,2);
-    [~,L] = compute_shape_matrices(x0, x0_com, E, order);
+    L = compute_shape_matrices(x0, x0_com, E, order);
     
     % Build Monomial bases for all quadrature points
     Y = monomial_basis_matrix(V, x0_com, order, k);

### 2021-01-12 matlab/cantileaver_sim.m
add code and assets for fem beam sim

@@ -1,0 +1,72 @@
+%cantileaver simulation
+[V,T,F] = readMESH('vem_beam_tet2.mesh');
+
+
+%setup simulation variables
+   
+%tet volumes
+vol = volume(V,T);
+
+%triangle gradients 
+dphidX = linear_tetmesh_dphi_dX(V,T);
+
+%Mass Matrix
+rho =  100.0*ones(size(T,1),1);
+M = linear_tetmesh_mass_matrix(V,T, rho, vol);
+
+%material properties 
+YM = 5e5; %in Pascals
+pr =  0.45;
+[lambda, mu] = emu_to_lame(YM*ones(size(T,1),1), pr*ones(size(T,1),1));
+
+%boundary conditions are fun
+min_I = find(V(:,1) == min(V(:,1)));
+P = fixed_point_constraint_matrix(V, sort(min_I));
+
+%external forces
+gravity = [0 -10 0]'; % We are on krypton 
+gravity = P*M*repmat(gravity, size(V,1),1);
+
+%project mass matrix down
+M = P*M*P';
+
+%warm up the display
+[~,p] = nice_plot(V,F);
+
+%simulation setup and loop
+energy_func = @(a,b, c, d, e) linear_tetmesh_neohookean_q(a,b,c,d,e,[0.5*mu, 0.5*lambda]);
+gradient_func = @(a,b, c, d, e) linear_tetmesh_neohookean_dq(a,b,c,d,e,[0.5*mu, 0.5*lambda]);
+hessian_func = @(a,b, c, d, e) linear_tetmesh_neohookean_dq2(a,b,c,d,e,[0.5*mu, 0.5*lambda], 'fixed');
+
+dt = 0.1;
+
+qt = reshape(V', 3*size(V,1),1);
+
+%setup all thw variables I need using bc projection matrix
+b = qt - P'*P*qt;
+qt = P*qt;
+vt = 0*qt;
+
+
+disp('Press SPACEBAR to CONTINUE');
+pause 
+
+h = text(0.5*(max(V(:,1)) + min(V(:,1))), 0.9*max(V(:,2)), num2str(0));
+h.FontSize = 28;
+    
+for ti=1:500
+    
+     h.String = num2str(ti);
+
+     g =  -M*vt + ...
+           dt*P*gradient_func(V,T, P'*qt+b, dphidX, vol) + ...
+           - dt*gravity;
+       
[…]

### 2021-01-13 paper/sections/results.tex
add figures of the distance weight

%
\begin{figure}
  \centering
  \includegraphics[width=3.33in]{figures/distance_weight_puft.pdf}
  \caption{Our distance weights decay smoothly from 1.0 (yellow) to 0.0 (green) when moving away from its closest surface. Here we visualize the distribution of the distance weights (with cutoff distance 5.0) corresponding to each part.   
  }
  \label{fig:distance_weight_puft}
  \vspace{-5pt}
\end{figure} 
%
%
\begin{figure}
  \centering
  \includegraphics[width=3.33in]{figures/plot_distance_weight.pdf}
  \caption{We plot the distance weight with respect to the primary and total ray length. Here the cutoff distance is 50.
  }
  \label{fig:plot_distance_weight}
  \vspace{-5pt}
\end{figure} 
%

### 2021-01-13 figures/render_puft_distance_weight.m

@@ -1,0 +1,59 @@
+clear;
+iges_file = 'puft_no_collar.iges';
+parts = nurbs_from_iges(iges_file);
+[V,F,N,C] = triangulate_iges(parts);
+
+save_path = ['../output/'];
+mkdir(save_path); 
+clf;
+figure(1);
+t = tsurf(F,V,'CData', C, 'VertexNormals', N, 'EdgeColor', 'none',...
+          fsoft, 'FaceAlpha', 0.9, ...
+          'AmbientStrength', 0.8, 'SpecularStrength', 0.2, 'DiffuseStrength', 0.4);
+hold on
+  CM = cbrewer('RdYlBu',20);
+  colormap(CM);
+shading(gca,'interp')
+axis equal
+set(gca,'Visible','off');
+bg_color = [1.0 1.0 1.0];
+set(gcf,'Color',bg_color);
+camlight;
+camproj('persp');
+view(0, 25);
+l = light('Position',[4 -4 10],'Style','infinite');
+add_shadow(t,l,'Color',bg_color*0.9,'BackgroundColor',bg_color,'Fade','infinite');
+apply_ambient_occlusion(t,'AddLights',false,'SoftLighting',false,'Factor',0.75);
+hold off
+
+print(1, [save_path iges_file(1:end-5) '.png'], '-dpng', '-r400');
+
+%% triangulate the nurbs parts
+function [verts,faces,normals,colors] = triangulate_iges(parts)
+  faces=[];
+  verts=[];
+  normals=[];
+  colors=[];
+ 
+  perm = randperm(numel(parts));
+  for ii=1:numel(parts)
+      if isfield(parts{ii}, 'hires_T')
+          F = parts{ii}.hires_T;
+          V = parts{ii}.hires_x0';
+          N = nurbs_normals(parts{ii}.srf.nurbs, parts{ii}.hires_UV, parts{ii}.p);
+      else
+          F = parts{ii}.T;
+          V = parts{ii}.x0';
+          N = nurbs_normals(parts{ii}.srf.nurbs, parts{ii}.UV, parts{ii}.p);
+      end
+      F = F + size(verts,1);
+      faces=[faces; F];
+      verts=[verts; V];     
+      normals=[normals; N];
+      colors = [colors; repmat(perm(ii), size(V,1), 1)];
+  end
+  % check and remove degenerate faces
+  dblA = doublearea(verts,faces);
+  deg_fid = find(dblA <= 0);
+  faces(deg_fid, :) = [];
+end

### 2021-01-13 figures/slice_distance_weight.m

@@ -1,14 +1,20 @@
 clear;
-iges_file = 'puft_simple.iges';
+iges_file = 'puft_no_collar.iges';
 parts = nurbs_from_iges(iges_file);
-[V, F] = triangulate_iges(parts);
+[V, F, ~] = triangulate_iges(parts);
+
+iges_file_wn = 'puft_no_collar.iges'; % for winding number
+parts_wn = nurbs_from_iges(iges_file_wn);
+[V_wn, F_wn] = triangulate_iges(parts_wn);
 
 % define the boundary of the cross section plane to be a rectangle 
-U_sqr = [min(V(:,1))-5 min(V(:,3))-5;...
-         min(V(:,1))-5 max(V(:,3))+5;...
-         max(V(:,1))+5 max(V(:,3))+5;...
-         max(V(:,1))+5 min(V(:,3))-5];
+U_sqr = [min(V(:,1))-1 min(V(:,3))-1;...
+         min(V(:,1))-1 max(V(:,3))+1;...
+         max(V(:,1))+1 max(V(:,3))+1;...
+         max(V(:,1))+1 min(V(:,3))-1];
 E_sqr = [1 2; 2 3; 3 4; 4 1];
+V_mean = mean(V);
+hole_pos = [V_mean(:,1) V_mean(:,3)];
 writePOLY_triangle('./test_sqr.poly', U_sqr, E_sqr, []);
 [TV_sqr,TF_sqr,TN_sqr] = triangle('./test_sqr.poly');
 
@@ -16,34 +22,45 @@ tsurf(TF_sqr, TV_sqr);
 
 alpha = 5.0; % cutoff distance
 
-step = 50; % number of the slicing steps 
+step = 3; % number of the slicing steps 
 TF_list = cell(step,1);
 X_list = cell(step,1);
 w_list = cell(step,1);
 
 ii = 1;
-for d = linspace(min(V(:,2))+0.2,max(V(:,2)-0.2),step)
+for d = linspace(min(V(:,2))+1.0,max(V(:,2)-1.0),step)
   plane = [0 1 0 -d]; % define the plane
   try
     [U,E,J] = slice_triangles(V,F,plane); % find the cross section of the nurbs
-    U_seg = [U(:,1) U(:,3)];
-    U = [U_seg; U_sqr];
+    U = [U(:,1) U(:,3)];
+    U = [U; U_sqr];
[…]

### 2021-01-16 matlab/cantileaver_sim.m
cantileaver simulation code

@@ -1,5 +1,5 @@
 %cantileaver simulation
-[V,T,F] = readMESH('vem_beam_tet2.mesh');
+[V,T,F] = readMESH('vem_beam_tet0.mesh');
 
 
 %setup simulation variables
@@ -15,7 +15,7 @@ rho =  100.0*ones(size(T,1),1);
 M = linear_tetmesh_mass_matrix(V,T, rho, vol);
 
 %material properties 
-YM = 5e5; %in Pascals
+YM = 1e5; %in Pascals
 pr =  0.45;
 [lambda, mu] = emu_to_lame(YM*ones(size(T,1),1), pr*ones(size(T,1),1));
 

### 2021-01-16 matlab/compute_shape_matrices.m

@@ -25,6 +25,7 @@ function L = compute_shape_matrices(x0, x0_com, E, order, mode)
 
     col_idx = 0;
     row_idx = 0;
+    
     % Forming global system matrix ([A S] matrix) 
     for i=1:numel(E)
         for j=1:numel(E{i})

### 2021-01-16 paper/sections/methods.tex

@@
-A pointform description of the general idea
-\begin{itemize}
-	\item A D-NURBS approach to simulation. Use a Jacobian to map from UV positions in a NURBS parametric space to world space positions on the surface. 
-	\item \textbf{Shape Matching}: From the sample world space positions on a single NURBS surface, we solve a least squares problem to fit a \textit{projection operator} to the surface.
+%A pointform description of the general idea
+%\begin{itemize}
+%	\item A D-NURBS approach to simulation. Use a Jacobian to map from UV positions in a NURBS parametric space to world space positions on the surface. 
+%	\item \textbf{Shape Matching}: From the sample world space positions on a single NURBS surface, we solve a least squares problem to fit a \textit{projection operator} to the surface.
 This projection operator maps monomials in the undeformed space, to the least squares estimate of the deformed positions for the given monomials.
-	\item With DNURBS and Shape Matching VEM we have two mappings: a UV to undeformed world space with our NURBS Jacobian, then an undeformed to deformed mapping with the projection operators.
+%	\item With DNURBS and Shape Matching VEM we have two mappings: a UV to undeformed world space with our NURBS Jacobian, then an undeformed to deformed mapping with the projection operators.
 This lets us form a set of generalized coordinates in terms of our control points of the NURBS surface (section 1.4).
-	\item For an arbitrary position in space (not necessarily on the surface), we can build a projection operator unique to this point via a weighted sum of the projection operators of the projection operators for each NURBS (eq 5).
+%	\item For an arbitrary position in space (not necessarily on the surface), we can build a projection operator unique to this point via a weighted sum of the projection operators of the projection operators for each NURBS (eq 5).
 From this we can reconstruct an estimate of the deformed position from the monomial basis defined using its undeformed position.
-	\item For any point in space, we need to compute a set of weights to the projection operators defined on the surfaces with the following criteria: the weights sum to 1, are nonzero, and depend on the distances to the surfaces (See section 1.8).
-	\item With this mapping for undeformed to deformed positions, this yields a simple definition for the deformation gradient (eq 12).
-	\item Armed with our deformation gradient, we next define the Lagrangian, but this basic form isn't sufficient.
+%	\item For any point in space, we need to compute a set of weights to the projection operators defined on the surfaces with the following criteria: the weights sum to 1, are nonzero, and depend on the distances to the surfaces (See section 1.8).
+%	\item With this mapping for undeformed to deformed positions, this yields a simple definition for the deformation gradient (eq 12).
+%	\item Armed with our deformation gradient, we next define the Lagrangian, but this basic form isn't sufficient.
 We augment our kinetic and potential energies with a stability term (ref VEM).
 This accounts for cases in which the position of the nodal values don't match their estimated positions defined by the projection operator mapping.
-	\item \textbf{Raycasting Quadrature}: To compute the volume integrals we make use of raycasting quadrature approach (ref) where we uniform sample a YZ grid and shoot rays in the X direction.
+%	\item \textbf{Raycasting Quadrature}: To compute the volume integrals we make use of raycasting quadrature approach (ref) where we uniform sample a YZ grid and shoot rays in the X direction.
 We then use the points of intersection to define integrations bounds for 1D integrals along these rays, which we integrate numerically.
-	\item With our augmented lagrangian, we use the Euler-Lagrange equation to arrive at equations of motion.
+%	\item With our augmented lagrangian, we use the Euler-Lagrange equation to arrive at equations of motion.
 Our generalized coordinates give us direct updates on the control points of the NURBS.
 Currently we are using linearly-implicit Euler for time integration.
-	\item \textbf{Handling trimmed NURBS}: For this case we update the original control points, but we must only sample points within the boundary.
+%	\item \textbf{Handling trimmed NURBS}: For this case we update the original control points, but we must only sample points within the boundary.
 To compute these points we use an approach similar to the raycasting quadrature but in 2D.
@@
 The final result is a set of UV coordinates that respect the boundary defined by NURBS curves.
-\end{itemize}
+%\end{itemize}
 
-% \textbf{x}(\textbf{X}) = \hat{A}M(\textbf{X}) + \bar{\textbf{x}}$ \\
-% $\hat{A} = \sum_i^{|S|}w_i(\textbf{X})A_i$
+\section{Methods}
+Brief overview. We have some control points. We have UV samples on the NURBS parametric domain. We have a NURBS jacobian UV to points on surface in R3.
+Given a volumetric model composed of one or more NURBS surfaces, the surface of each surface is reconstructed using a set of control points $\textbf{p}$ and their associated weights $w$.
+Following D-NURBS \cite{10.1145/176579.176580}, we take our degrees of freedom (DOF) to be control points of the surfaces and produce displacements directly on these control points at each simulation step.
+We note that we used a simplified version of D-NURBS in which the weights are held constant throughout the simulation and find this not to be limiting \myworries{(feel like we'll need better justification than saying just this?)}.
+We represent the degrees of freedom for the $i$-th NURBS surface with $m$ control points as $\mathbf{q}_i = \left[ \mathbf{p}_1^T, \dots, \mathbf{p}_m^T \right]^T \in \mathbb{R}^{3m}$.
+From this definition, we write the DOF for the entire model as $\mathbf{q} = \left[ \mathbf{q}_i^T, \dots, \mathbf{q}_n^T \right]^T$ where $n$ is the number of NURBS surfaces for the model.
 
-We have $\mathbf{x} = (x_0, x_1)$ and $\mathbf{X} = (X_0, X_1) $ as deformed and undeformed positions, respectively. \\
+To evaluate the volumetric integrals in solid mechanics models, we require material points $\mathbf{X} \in \mathbb{R}^{3m}$ in the undeformed space that serve as integration points.
+Our shape matching element method enables us to evaluate the deformed positions of the material points strictly using positions defined on the boundary.
+Like in VEM, we fit a polynomial to each boundary primitive (NURBS in this case) describing its deformation.
+The fitting of these polynomials closely follows that of the Shape Matching \cite{10.1145/1073204.1073216} approach where we perform a least squares polynomial fit to the deformed boundary points.
+These boundary points are represented by sampled positions on each NURBS surface.
+Since each surface is associated with a single polynomial, we can construct a unique polynomial for some arbitrary point by blending these polynomials.
+Thus given some material point $\mathbf{X}_i$ we find it's corresponding deformed position $\mathbf{x}_i$ via a weighted combination of the polynomials.
+This permits us to build a general deformation map which we use to construct our equations of motion.
 
-\textbf{<insert Exposition about FEM >} \\
-In our case, the function we are trying to evaluate is deformation with respect to the center of mass.
+In the following sections, we will first review the simplified D-NURBS model used to form our generalized coordinates.
+We then present our shape matching algorithm for fitting polynomials to surfaces ....
+In section ...
+we next describe how to blend these surface polynomials to construct a general deformation map for material points.
+Section ...
+describes our strategy for selecting integration points in the volume through a raycasting approach.
+Finally, we formulate our equations of motions from a Lagrangian formulation.
+
+\section{D-NURB Model}
+\section{Shape Matching}
+\section{Blending Polynomials}
+\section{Raycasting Quadrature}
+\section{Dynamic Simulation}
+\subsection{Generalized Inertia}
+\subsection{Generalized Forces}
+\subsection{Equations of Motion}
+This is most notable in cases where the degree of the polynomial on a NURBS is higher than the degree of the B-spline surface.
+
+%
+%
+%% \textbf{x}(\textbf{X}) = \hat{A}M(\textbf{X}) + \bar{\textbf{x}}$ \\
+%% $\hat{A} = \sum_i^{|S|}w_i(\textbf{X})A_i$
+%
+%We have $\mathbf{x} = (x_0, x_1)$ and $\mathbf{X} = (X_0, X_1) $ as deformed and undeformed positions, respectively. \\
+%
+%\textbf{<insert Exposition about FEM >} \\
+%In our case, the function we are trying to evaluate is deformation with respect to the center of mass.
 In VEM, instead of reconstructing values of our function using nodal values, we seek to find a a set of polynomial coefficients that map some arbitrary undeformed position to a polynomial approximation of its deformed position.
@@
 \\
-
-\subsection{Shape Matching: }
-In shape matching, authors seek a polynomial transformation, $\mathbf{\Pi}$ that minimizes $\sum_i ||\mathbf{\Pi}\mathbf{M(X)} - \mathbf{p_i}||^2$ where $\mathbf{M(X)}$ is the monomial basis of the $i$-th boundary vertex and $\mathbf{p_i}$ is current displacement from the center of mass for the $i$-th point.
+%
+%\subsection{Shape Matching: }
+%In shape matching, authors seek a polynomial transformation, $\mathbf{\Pi}$ that minimizes $\sum_i ||\mathbf{\Pi}\mathbf{M(X)} - \mathbf{p_i}||^2$ where $\mathbf{M(X)}$ is the monomial basis of the $i$-th boundary vertex and $\mathbf{p_i}$ is current displacement from the center of mass for the $i$-th point.
 We argue this paper can be interpreted as a sort of virtual element method.
 \\
-
-\subsection{Reconstructing deformed positions}
-A general form for the monomial basis takes the following form <insert here>. For example, in 2D with quadratic deformation we have:
-\begin{equation}
-    \mathbf{M(X)} =\begin{bmatrix}
-    X_0 - \bar{X}_0  \\ 
-    X_1 - \bar{X}_1  \\ 
-    (X_0 - \bar{X}_0)^2  \\ 
-    (X_1 - \bar{X}_1)^2  \\ 
-    (X_0 - \bar{X}_0)(X_1 - \bar{X}_1) 
-    \end{bmatrix} 
-\end{equation}
-
-We write the estimated deformed position $\mathbf{x}$ of some arbitrary undeformed position $\mathbf{X}$ as
-\begin{equation}
-     \mathbf{x}(\mathbf{X}) = \hat{\mathbf{\Pi}}\mathbf{M}(\mathbf{X}) + \bar{\mathbf{x}} 
-\end{equation}
-where $\mathbf{\hat{\Pi}}$ is the blended projection operator for this particular point:
-\begin{equation}
-    \hat{\mathbf{\Pi}}\mathbf{(X)} = \sum_i^{|S|}w_i(\mathbf{X})\mathbf{\Pi}_i 
-\end{equation}
-The bar symbol indicates the center of mass. For example, $\bar{\mathbf{x}}$ is the deformed object's center of mass, which we compute as the mean of the boundary values 
-\begin{equation}
-    \bar{\mathbf{x}} = \frac{1}{n}\sum_i^n x_i 
-\end{equation}
-Here $w_i$ is some weighting function for the $i$-th shape. There are many possible forms for this weighting function. Right now we compute these weights by forming the following optimization: \\
-<insert here>
-In order to form the equations of motion, we need to re-express the above equation in terms of the nodal values, $\mathbf{q}$.
+%
+%\subsection{Reconstructing deformed positions}
+%A general form for the monomial basis takes the following form <insert here>. For example, in 2D with quadratic deformation we have:
+%\begin{equation}
+%    \mathbf{M(X)} =\begin{bmatrix}
+%    X_0 - \bar{X}_0  \\ 
+%    X_1 - \bar{X}_1  \\ 
+%    (X_0 - \bar{X}_0)^2  \\ 
+%    (X_1 - \bar{X}_1)^2  \\ 
+%    (X_0 - \bar{X}_0)(X_1 - \bar{X}_1) 
+%    \end{bmatrix} 
+%\end{equation}
+%
+%    
+%We write the estimated deformed position $\mathbf{x}$ of some arbitrary undeformed position $\mathbf{X}$ as
+%\begin{equation}
+%     \mathbf{x}(\mathbf{X}) = \hat{\mathbf{\Pi}}\mathbf{M}(\mathbf{X}) + \bar{\mathbf{x}} 
+%\end{equation}
+%where $\mathbf{\hat{\Pi}}$ is the blended projection operator for this particular point:
+%\begin{equation}
+%    \hat{\mathbf{\Pi}}\mathbf{(X)} = \sum_i^{|S|}w_i(\mathbf{X})\mathbf{\Pi}_i 
+%\end{equation}
+%The bar symbol indicates the center of mass. For example, $\bar{\mathbf{x}}$ is the deformed object's center of mass, which we compute as the mean of the boundary values 
+%\begin{equation}
+%    \bar{\mathbf{x}} = \frac{1}{n}\sum_i^n x_i 
+%\end{equation}
+%Here $w_i$ is some weighting function for the $i$-th shape. There are many possible forms for this weighting function. Right now we compute these weights by forming the following optimization: \\
+%<insert here>
+%In order to form the equations of motion, we need to re-express the above equation in terms of the nodal values, $\mathbf{q}$.
 To arrive it these we note that each $\mathbf{\Pi}_i$ is a function of $\mathbf{q}$.
-\begin{equation}
-    \mathbf{\Pi}_i(\mathbf{q}) = \mathbf{P(E_i q)M(E_i q_0)^{\dagger}}
-\end{equation}
-\myworries{I'm not a fan of what I wrote here. The MLS-MPM paper presents a similar idea when talking about EFG, and I like their notation a bit better}
-This is the shape matching problem's least squares solution where $M(q_0)^{\dagger}$ is the Moore-Penrose pseudo-inverse of the monomial vectors stacked row-wise (add example equation).
+%\begin{equation}
+%    \mathbf{\Pi}_i(\mathbf{q}) = \mathbf{P(E_i q)M(E_i q_0)^{\dagger}}
+%\end{equation}
+%\myworries{I'm not a fan of what I wrote here. The MLS-MPM paper presents a similar idea when talking about EFG, and I like their notation a bit better}
+%This is the shape matching problem's least squares solution where $M(q_0)^{\dagger}$ is the Moore-Penrose pseudo-inverse of the monomial vectors stacked row-wise (add example equation).
 The matrix $\mathbf{E}_i$ selects the degrees of freedom from $\mathbf{q}$ associated with the $i$-th shape.
@@
 We observe that the above equation is linear in $\mathbf{q}$ so we may rewrite the equation for $\mathbf{x}$ as
-\begin{equation}
-    \mathbf{x(X)} = \mathbf{J_{\hat{\Pi}}(X)q}
-\end{equation}
-
-\subsection{Simulation on NURBs models}
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=\linewidth]{figures/nurbs_surface}
-  \caption{The NURBs model uses control points, shown in grey, as the original degrees of freedom.}
-  \Description{A simple NURBS surface in $R^3$.}
-\end{figure}
-
-We used the D-NURBS model (ref) for simulating on NURBs surfaces, giving an approach to simulating directly on the degrees of freedom of parametric NURBs models.
+%\begin{equation}
+%    \mathbf{x(X)} = \mathbf{J_{\hat{\Pi}}(X)q}
+%\end{equation}
+%
+%\subsection{Simulation on NURBs models}
+%
+%\begin{figure}[h]
+%  \centering
+%  \includegraphics[width=\linewidth]{figures/nurbs_surface}
+%  \caption{The NURBs model uses control points, shown in grey, as the original degrees of freedom.}
+%  \Description{A simple NURBS surface in $R^3$.}
+%\end{figure}
+%
+%We used the D-NURBS model (ref) for simulating on NURBs surfaces, giving an approach to simulating directly on the degrees of freedom of parametric NURBs models.
 The equation for a NURBs curve takes the form:
-\begin{equation}
-    \mathbf{x}(u) = \frac{\sum_{i=1}^{n} \mathbf{p}_i w_i B_{i,k}(u)}{ w_i B_{i,k}(u)}
-\end{equation}
-where $u$ is the parametric coordinate, $\mathbf{p}_i$ is the $i$-th control point with $n$ total control points.
+%\begin{equation}
+%    \mathbf{x}(u) = \frac{\sum_{i=1}^{n} \mathbf{p}_i w_i B_{i,k}(u)}{ w_i B_{i,k}(u)}
+%\end{equation}
+%where $u$ is the parametric coordinate, $\mathbf{p}_i$ is the $i$-th control point with $n$ total control points.
 $B_{i,k}(u)$ is the B-spline basis function for the $i-th$ control point with degree $k-1$.
@@
 In our case, we primarily work with NURBs surfaces, which are a generalization of the tensor-product of two B-splines:
-\begin{equation}
-    \mathbf{x}(u,v) = \frac{\sum_{i=1}^{n}\sum_{j=1}^{m}  \mathbf{p}_{i,j} w_{i,j} B_{i,k}(u)B_{j,l}(v)}{ w_{i,j} B_{i,k}(u)B_{j,l}(v)}
-\end{equation}
-where we now have two parameters $u,v$, and $m\cdot n$ control points. In D-NURBs, we see that $\mathbf{x}(u,v)$ is linear in the control points, $\mathbf{p}$, so we can rewrite the above equation as
-\begin{equation}
-    \mathbf{x}(u,v,\mathbf{p}) = \mathbf{J}(u,v)\mathbf{p}.
-\end{equation}
-Here $\mathbf{J}$ is the Jacobian matrix that maps the control points to some world space position on the surface for some $u,v$ pair.
+%\begin{equation}
+%    \mathbf{x}(u,v) = \frac{\sum_{i=1}^{n}\sum_{j=1}^{m}  \mathbf{p}_{i,j} w_{i,j} B_{i,k}(u)B_{j,l}(v)}{ w_{i,j} B_{i,k}(u)B_{j,l}(v)}
+%\end{equation}
+%where we now have two parameters $u,v$, and $m\cdot n$ control points. In D-NURBs, we see that $\mathbf{x}(u,v)$ is linear in the control points, $\mathbf{p}$, so we can rewrite the above equation as
+%\begin{equation}
+%    \mathbf{x}(u,v,\mathbf{p}) = \mathbf{J}(u,v)\mathbf{p}.
+%\end{equation}
+%Here $\mathbf{J}$ is the Jacobian matrix that maps the control points to some world space position on the surface for some $u,v$ pair.
 <elaborate on the form of J>.
 Now with these generalized coordinates, we may perform dynamics directly on the degrees of freedom of our NURBs models.
-
-\subsection{Generalized Coordinates}
-With D-NURBs simulation layered on the VEM simulation, we have two Jacobians: $\mathbf{J}(u,v)$ mapping controls points, $\mathbf{p}$ onto surface positions, and then $\mathbf{J_{\hat{\Pi}}(X)}$ which maps surface positions $\mathbf{q}$ to their deformed positions.
+%
+%\subsection{Generalized Coordinates}
+%With D-NURBs simulation layered on the VEM simulation, we have two Jacobians: $\mathbf{J}(u,v)$ mapping controls points, $\mathbf{p}$ onto surface positions, and then $\mathbf{J_{\hat{\Pi}}(X)}$ which maps surface positions $\mathbf{q}$ to their deformed positions.
 The Jacobian $\mathbf{J}(u,v)$  may be precomputed, and each point sampled on a surface of a NURBs has an associated $\mathbf{J}(u,v)$.
 With these two Jacobians we may express a single $u,v$ pair's deformed position as
-\begin{equation}
-    \mathbf{x}(u,v,\mathbf{p}) = \mathbf{J_{\hat{\Pi}}(X)}\mathbf{J}(u,v)\mathbf{p}
-\end{equation}
-and the undeformed position
-\begin{equation}
-    \mathbf{X}(u,v) = \mathbf{J}(u,v)\mathbf{p_0}
-\end{equation}
-where $\mathbf{p_0}$ is the initial values for the control points.
-
-\subsection{Deformation Gradient}
-\begin{equation}
-        \mathbf{F} = \frac{\partial \mathbf{x}}{\partial \mathbf{X}}
-               = \frac{\partial \hat{\mathbf{\Pi}}\mathbf{M(X)}}{\partial \mathbf{X}}
-               = \hat{\mathbf{\Pi}} \frac{\partial \mathbf{M(X)}}{\partial \mathbf{X}} 
-\end{equation}
-The $\frac{\partial \mathbf{M(X)}}{\partial \mathbf{X}}$ terms may be precomputed and using the previous $\mathbf{M(X)}$ example, we have
-\begin{equation}
-     \frac{\partial \mathbf{M(X)}}{\partial \mathbf{X}} = \begin{bmatrix}
-     1&0 \\ 
-     0&1 \\ 
-     2(X_0 - \bar{X}_0)&0 \\ 
-     0&2(X_1 - \bar{X}_1) \\ 
-     (X_1 - \bar{X}_1)&(X_0 - \bar{X}_0) 
-    \end{bmatrix}
-\end{equation}
-
-In computing the forces and the stiffness matrix, we require the gradient of the deformation gradient with respect to the configuration,  $\frac{\partial \mathbf{F}}{\partial \mathbf{q}}$.
+%\begin{equation}
+%    \mathbf{x}(u,v,\mathbf{p}) = \mathbf{J_{\hat{\Pi}}(X)}\mathbf{J}(u,v)\mathbf{p}
+%\end{equation}
+%and the undeformed position
+%\begin{equation}
+%    \mathbf{X}(u,v) = \mathbf{J}(u,v)\mathbf{p_0}
+%\end{equation}
+%where $\mathbf{p_0}$ is the initial values for the control points.
+%
+%\subsection{Deformation Gradient}
+%\begin{equation}
+%        \mathbf{F} = \frac{\partial \mathbf{x}}{\partial \mathbf{X}}
+%               = \frac{\partial \hat{\mathbf{\Pi}}\mathbf{M(X)}}{\partial \mathbf{X}}
+%               = \hat{\mathbf{\Pi}} \frac{\partial \mathbf{M(X)}}{\partial \mathbf{X}} 
+%\end{equation}
+%The $\frac{\partial \mathbf{M(X)}}{\partial \mathbf{X}}$ terms may be precomputed and using the previous $\mathbf{M(X)}$ example, we have
+%\begin{equation}
+%     \frac{\partial \mathbf{M(X)}}{\partial \mathbf{X}} = \begin{bmatrix}
+%     1&0 \\ 
+%     0&1 \\ 
+%     2(X_0 - \bar{X}_0)&0 \\ 
+%     0&2(X_1 - \bar{X}_1) \\ 
+%     (X_1 - \bar{X}_1)&(X_0 - \bar{X}_0) 
+%    \end{bmatrix}
+%\end{equation}
+%
+%In computing the forces and the stiffness matrix, we require the gradient of the deformation gradient with respect to the configuration,  $\frac{\partial \mathbf{F}}{\partial \mathbf{q}}$.
 Usually the form of this gradient is simple.
@@
 In our case, the presence of the projection operator $\hat{\mathbf{\Pi}}$ means that each quadrature point is computed using the nodal values of the entire model.
-
-\begin{equation}
-    \frac{\partial \mathbf{F}}{\partial \mathbf{q}} = \frac{\partial \mathbf{P}}{\partial q}\mathbf{M(q_0)}^\dagger  \frac{\partial \mathbf{M(X)}}{\partial \mathbf{X}} 
-\end{equation}
-
-Naively handled, this gives us a $\frac{\partial \mathbf{F}}{\partial \mathbf{q}}$ of size $3 x |q|$.
+%
+%\begin{equation}
+%    \frac{\partial \mathbf{F}}{\partial \mathbf{q}} = \frac{\partial \mathbf{P}}{\partial q}\mathbf{M(q_0)}^\dagger  \frac{\partial \mathbf{M(X)}}{\partial \mathbf{X}} 
+%\end{equation}
+%
+%
+%Naively handled, this gives us a $\frac{\partial \mathbf{F}}{\partial \mathbf{q}}$ of size $3 x |q|$.
 However, many of the columns of have values all near zero, allowing us to prune many columns for each quadrature point.
@@
 \\
-
-\subsection{Variational Mechanics}
-We have now provided all the necessary ingredient to produce equations of motion for deformable bodies, which we derive via Euler-Lagrange equation. We begin with the Lagrangian
-\begin{equation}
-    L=T-V
-\end{equation}
-where $T$ is the kinetic energy, and $V$ is the potential energy. From the Euler-Lagrange equation we have:
-\begin{equation}
-    \frac{d}{dt} \frac{\partial L}{\partial \dot{\mathbf{p}}} + \frac{\partial L}{\partial {\mathbf{p}}} = 0
-\end{equation}
-\begin{equation}
-    V =  \int_\Omega \psi(\mathbf{F(X)} dX
-\end{equation}
-then discretizing this over $m$ quadrature points, we have
-\begin{equation}
-    V =  \sum_i^m \psi(\mathbf{F(X_i)} A_i
-\end{equation}
-\begin{equation}
-    \frac{\partial L}{\partial \mathbf{q}} = \sum_i^m \frac{\partial V_i}{\partial \mathbf{q}} = \sum_i^m \frac{\partial \psi(\mathbf{F_i)}}{\partial \mathbf{F}}\frac{\partial \mathbf{F_i}}{\partial \mathbf{q}}A_i
-\end{equation}
-Next we have the equation of the kinetic energy
-\begin{equation}
-    T = \frac{1}{2}\int_{ \Omega} \rho \dot{\mathbf{x}}^T\dot{\mathbf{x}} dV
-\end{equation}
-which we rewrite in terms of our control points configuration
-\begin{equation}
-    T = \frac{1}{2} \int_{\partial \Omega} \dot{\mathbf{p}}^T \mathbf{J}(u,v)^T \mathbf{M} \mathbf{J}(u,v) \dot{\mathbf{p}} du dv
-\end{equation}
-\myworries{change notation to not use M for the monomial basis}
-where
-\begin{equation}
-    \mathbf{M} = \int_\Omega \rho \mathbf{J_{\hat{\Pi}}(X)}^T \mathbf{J_{\hat{\Pi}}(X)}
-\end{equation}
-From the potential energy (which uses dF/dq) and this potential energy, we discretize in time and right now I'm using linearly implicit backward euler
-
-\subsection{Stability term}
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=\linewidth]{figures/error_energy}
-  \caption{A figure showing the spring energy error for each sampled point on the NURBS curve. I should change the colors of the lines to show higher energies where the error is higher.}
-  \Description{Showing an example error of a NURBS curve.}
-\end{figure}
-
-\subsection{Projection Operator Weighting}
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=\linewidth]{figures/distance_weighting}
-  \caption{A 2D example showing how $\theta$ is computed for $\Pi_B$.
+%
+%
+%\subsection{Variational Mechanics}
+%We have now provided all the necessary ingredient to produce equations of motion for deformable bodies, which we derive via Euler-Lagrange equation. We begin with the Lagrangian
+%\begin{equation}
+%    L=T-V
+%\end{equation}
+%where $T$ is the kinetic energy, and $V$ is the potential energy. From the Euler-Lagrange equation we have:
+%\begin{equation}
+%    \frac{d}{dt} \frac{\partial L}{\partial \dot{\mathbf{p}}} + \frac{\partial L}{\partial {\mathbf{p}}} = 0
+%\end{equation}
+%\begin{equation}
+%    V =  \int_\Omega \psi(\mathbf{F(X)} dX
+%\end{equation}
+%then discretizing this over $m$ quadrature points, we have
+%\begin{equation}
+%    V =  \sum_i^m \psi(\mathbf{F(X_i)} A_i
+%\end{equation}
+%\begin{equation}
+%    \frac{\partial L}{\partial \mathbf{q}} = \sum_i^m \frac{\partial V_i}{\partial \mathbf{q}} = \sum_i^m \frac{\partial \psi(\mathbf{F_i)}}{\partial \mathbf{F}}\frac{\partial \mathbf{F_i}}{\partial \mathbf{q}}A_i
+%\end{equation}
+%Next we have the equation of the kinetic energy
+%\begin{equation}
+%    T = \frac{1}{2}\int_{ \Omega} \rho \dot{\mathbf{x}}^T\dot{\mathbf{x}} dV
+%\end{equation}
+%which we rewrite in terms of our control points configuration
+%\begin{equation}
+%    T = \frac{1}{2} \int_{\partial \Omega} \dot{\mathbf{p}}^T \mathbf{J}(u,v)^T \mathbf{M} \mathbf{J}(u,v) \dot{\mathbf{p}} du dv
+%\end{equation}
+%\myworries{change notation to not use M for the monomial basis}
+%where
+%\begin{equation}
+%    \mathbf{M} = \int_\Omega \rho \mathbf{J_{\hat{\Pi}}(X)}^T \mathbf{J_{\hat{\Pi}}(X)}
+%\end{equation}
+%From the potential energy (which uses dF/dq) and this potential energy, we discretize in time and right now I'm using linearly implicit backward euler
+%
+%\subsection{Stability term}
+%\begin{figure}[h]
+%  \centering
+%  \includegraphics[width=\linewidth]{figures/error_energy}
+%  \caption{A figure showing the spring energy error for each sampled point on the NURBS curve. I should change the colors of the lines to show higher energies where the error is higher.}
+%  \Description{Showing an example error of a NURBS curve.}
+%\end{figure}
+%
+%\subsection{Projection Operator Weighting}
+%\begin{figure}[h]
+%  \centering
+%  \includegraphics[width=\linewidth]{figures/distance_weighting}
+%  \caption{A 2D example showing how $\theta$ is computed for $\Pi_B$.
 $d_1$ shows the distance to the closest point on $\Pi_B$ to $\mathbf{x}$ and $d_2$ shows the distance to the surface hit by the ray in the opposite direction.}
-  \Description{Showing a 2D NURBS curve with a quadrature point in the center.
+%  \Description{Showing a 2D NURBS curve with a quadrature point in the center.
 A ray is cast to the nearest point on a shape, showing the distance between the quadrature point and the nearest point as well as the distance to the "back surface".}
-\end{figure}
-
-\begin{equation}
-\begin{aligned}
-\mathbf{w}(\mathbf{X}) = \min_{\mathbf{w}} \quad & \mathbf{w}^T \Theta(\mathbf{X}) \mathbf{w}    \\
-\textrm{s.t.} \quad & 0 \leq w_i \leq 1                     \\
-                    &   \sum_i^n w_i = 1                      \\
-\end{aligned}
-\end{equation}
-
-where
-\begin{equation}
-     \Theta(\mathbf{X}) = \text{diag}\left( \frac{1}{\theta_1},\frac{1}{\theta_2},\dots,\frac{1}{\theta_N}\right)
-\end{equation}
-In computing the theta values, we require that they a approach 1 at the target surface and fall off to zero as a point becomes closer to other surfaces.
+%\end{figure}
+%
+%\begin{equation}
+%\begin{aligned}
+%\mathbf{w}(\mathbf{X}) = \min_{\mathbf{w}} \quad & \mathbf{w}^T \Theta(\mathbf{X}) \mathbf{w}    \\
+%\textrm{s.t.} \quad & 0 \leq w_i \leq 1                     \\
+%                    &   \sum_i^n w_i = 1                      \\
+%\end{aligned}
+%\end{equation}
+%
+%where
+%\begin{equation}
+%     \Theta(\mathbf{X}) = \text{diag}\left( \frac{1}{\theta_1},\frac{1}{\theta_2},\dots,\frac{1}{\theta_N}\right)
+%\end{equation}
+%In computing the theta values, we require that they a approach 1 at the target surface and fall off to zero as a point becomes closer to other surfaces.
 To compute this for some surface, we shoot a ray from the point to the nearest point on the surface.
@@
 We compute the theta with the following linear function
-
-\begin{equation}
-theta = \max (1.0 - \frac{d_1}{\min (d_2, D)}, 0.0)
-\end{equation}
-where $D$ is a fixed cut-off distance.
-
-\subsection{Integrating the volume}
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=\linewidth]{figures/raycasting_quadrature}
-  \caption{An example raycasting quadrature on a 2D NURBS curve.}
-  \Description{Shows a 2D NURBs curve with horizontal rays intersecting the curve. }
-\end{figure}
+%
+%\begin{equation}
+%theta = \max (1.0 - \frac{d_1}{\min (d_2, D)}, 0.0)
+%\end{equation}
+%where $D$ is a fixed cut-off distance.
+%
+%
+%\subsection{Integrating the volume}
+%\begin{figure}[h]
+%  \centering
+%  \includegraphics[width=\linewidth]{figures/raycasting_quadrature}
+%  \caption{An example raycasting quadrature on a 2D NURBS curve.}
+%  \Description{Shows a 2D NURBs curve with horizontal rays intersecting the curve. }
+%\end{figure}

### 2021-01-17 paper/sections/related.tex
related work draft

--- 2021-01-07 paper/sections/related.tex
@@
 \section{Related Work}
-\subsection{Shape Matching Related Papers}
-\begin{enumerate}
-    \item \textit{Shape Matching} \cite{10.1145/1073204.1073216}: Meshless simulation by fitting a polynomial describing the shapes deformation using only the nodal values.
-    \item \textit{Lattice Shape Matching} \cite{10.1145/1275808.1276480}: Voxelize model to construct a lattice of cubes.
+
+%% Set the scene
+Geometric modeling is a necessary precursor to physics-based animation, however differing geometric representations for modelling and simulation often require
+time-consuming, complex geometry processing pipelines. 
+For instance, the popular tetrahedral finite element approach for simulating solid elastodynamics requires robust algorithms for converting input surface geometry into 
+a volumetric tetrahedral mesh. This is a difficult problem and while significant progress has been~\cite{10.1145/2343483.2343501} made,
+even the most robust volumetric methods~\cite{Hu:2018:TMW:3197517.3201353} can be time consuming, fail to maintain correspondence between the input model and output simulation mesh,
+and don't work directly on curved surface representations such as NURBS.
+
+%% Outline desiderata for success
+%- no meshing of any kind, no tet, hex cut cells, nothing
+%- simulation can be directly imported and edited in modelling software
+%- supports standard continuum mechanics constitutive models
+%- supports different time integrators (can show a few maybe linearly implicit, fully implicit and RK4 ?)
+
+For many physically-based animation tasks, it would be desirable to bypass volumetric meshing entirely and directly simulate the geometric model.
+An ideal approach would avoid meshing of any kind (no volumetric meshes or cutcells), support continuum mechanics type constitutive models 
+and energies that have become standard in physics-based animation pipelines, be compatible with a wide range of time integration schemes and
+ensure that simulation output can be edited in the same modelling software used to create the input (important for post-processing). 
+Finally, our method should put only moderate constraints on input model quality to facilitate ease-of-use.
+
+\emph{Isogeometric Analysis}~\cite{10.1002/9780470749081.ch7} endeavors to perform simulation directly on the NURBS output from 
+Computer-Aided Design (CAD) software. Initial attempts used NURBS surfaces to represent the mid-surface of thin objects~\cite{10.1145/176579.176580}.
+Initial attempts at volumetric simulation relied on volumetric NURBS~\cite{10.1007/978-3-642-03596-8_2} but were limited to a narrow class of geometries.
+Finite volume methods are applicable to more general geometries~\cite{HEINRICH20121645,10.1002/nme.2311} but require a volumetric mesh be generated.
+Modern approaches are constructed around the extended finite element method which enriches the standard finite element basis with discontinuous basis functions
+to enable improved boundary handling~\cite{https://doi.org/10.1002/nme.3120,10.1007/s00466-013-0854-7,https://doi.org/10.1002/nme.4852,SAFDARI2016373,10.1145/3355089.3356576}. 
+These methods typically start with an easy-to-generate structured volumetric mesh (tetrahedral or hexahedral), ``cutting'' the NURBS geomtric model against
+it to enable boundary handling (such a mesh is called a \emph{cutcell} mesh). 
+Like volumetric meshing, this cutting operation can be difficult and our ideal method would avoid it if possible. 
+Some cutcell algorithms assume engineering/manufacturing quality input, which puts tight requirements on input models quality~\cite{10.1145/3355089.3356576}.
+Finally these approaches require additional mechanisms to ensure that simulation results lie inside the shape space of the input model's primitives,
+increasing complexity.
+
+\emph{Shape Matching} is a meshless approach to physics-based animation~\cite{10.1145/1073204.1073216,10.1145/2019406.2019438} and geometry processing~\cite{10.1111/j.1467-8659.2012.03171.x}
+built around shape registration. The algorithm has been extended from volumetric triangle mesh input to cloth~\cite{10.2312/PE/vriphys/vriphys08/039-046} to particles~\cite{10.1145/1964921.1964987}
+and even to visual geometry for video games~\cite{10.1145/2994258.2994260}. 
+Shape Matching is fast~\cite{10.1145/1275808.1276480,10.5555/1632592.1632606} and meshless, but state-of-the-art methods require 
+additional modeling input to position simulation primitives ~\cite{0.1145/1964921.1964987} or limit simulation primitives to be 
+collections of convex polytopes~\cite{10.1145/2994258.2994260}.
+Finally, Shape Matching is tightly coupled to the position-based dynamics~\cite{muller2007position}
+time integration methodology. 
+While incredibly performant, this approach is incompatible with the constitutive models that are popular for physics-based animation as well as 
+other time integration schemes. 
+The popular Projective Dynamics algorithm~\cite{10.1145/2601097.2601116} enables a more flexible Shape Matching implementation but is still limited to a subset of
+constitutive models for elastic solids.
+
+To alleviate these restrictions we can to turn to other meshless methods popular in computer graphics and 
+engineering~\cite{10.1145/1028523.1028542,https://doi.org/10.1002/fld.1650200824,10.1145/1778765.1778776,10.1145/2010324.1964968, 10.1145/1944846.1944855}.
+These support more advanced constitutive models and integration schemes but often require background integration meshes and lose the direct connection with
+modelling geometry. 
+Like cutcell-based, Isogeometric Analysis approaches, additional constraints must be added to ensure that simulation output can be represented by the input model.
+Given this, we conclude that there is no existing algorithm that meets our desiderata for success.
+
+Our algorithm takes the Shape Matching approach as inspiration, but rather than follow the PBD formalism, we interpret Shape Matching as a 
+Virtual Element Method (VEM)~\cite{10.1142/S0218202512500492,10.1142/S021820251440003X}. Virtual Elements are an extension of 
+mimetic finite differences~\cite{10.1142/S0218202505000832,10.1016/j.jcp.2013.07.031} to weak-form variational problems. 
+VEM relaxed the mesh generation requirements of finite element method by enabling the solution of partial differential equations
+on domains partitioned with arbitrary polytopes~\cite{10.1145/3386569.3392389}. 
+The solution inside each polytope is approximated using a polynomial function of a specified order.
+VEM typically assumes that the boundary of the the problem domain is described by a piecewise linear complex
+which makes its standard formulation incompatible with our NURBS based input geometry.
+
+\subsection*{Contributions} 
+In this paper we develop a new Shape Matching-based, Virtual Element Method which is directly compatible 
+with NURBS input geometry rather than piecewise linear surfaces. 
+Furthermore, we improve the expressivity of the VEM basis by using a shape blending approach inspired by
+algorithms for skinning~\cite{skinningcourse:2014}.
+Our method is entirely meshless (requiring no volumetric meshes or cutcells) and is compatible with standard constitutive models and time integrators.
+It guarantees that simulation output is directly consumable by the input modeling software and can ingest models which include large gaps,
+intersections and disconnected primitives without additional user input. 
+These features mean that our algorithm is the first truly meshless approach for the simulation of NURBS models for physics-based animation.
+
+%\subsection{Virtual Element and other Element Methods}
+%\begin{enumerate}
+    %\item \textit{Mimetic Finite Differences} \cite{10.1142/S0218202505000832} \cite{10.1016/j.jcp.2013.07.031} (they double dipped!): Considered a close relative to VEM and framed as the predecessor to VEM (in VEM papers).
+I still haven't read on MFD yet.
+From PolyDDF: "extension of finite volume and finite difference techniques to polygons that first discretizes a prime operator (typically, the gradient or the divergence) via a boundary integral, and then derives other operators by mimicking continuous structural properties."
+    %\item \textit{Basic principles of Virtual Element Method} \cite{10.1142/S0218202512500492} Original VEM paper
+    %\item \textit{The Hitchhiker's Guide to Virtual Element Method} \cite{10.1142/S021820251440003X}: More understandable versions of original VEM paper.
+    %\item \textit{Discrete Differential Operators on Polygonal Meshes} \cite{10.1145/3386569.3392389}: Extends VEM to do discrete differential geometry on arbitrary polygonal meshes.
+    %Don't think we need to cite this \item \textit{FLexible Simulation of Deformable Models using Discontinuous Galerkin FEM} \cite{10.5555/1632592.1632608}: Uses ordinary hexahedral elements, but uses a cut cell-based approach to support arbitrary polyhedra on the surface.
+    %Later ?\item \textit{Generalizing the finite element method: Diffuse approximation and diffuse elements} \cite{Nayroles1992GeneralizingTF} Predecessor to Element Free Galerkin.
+FEM interpolation replaced with a local Moving Least Square interpolation.
+    %Later ?\item \textit{Element-free Galerkin methods} \cite{https://doi.org/10.1002/nme.1620370205}: similar to DEM, but more accurate gradients (not sure of all the differences).
+In MLS methods, they solve least squares for each particle in the domain, weighting nearby particles with a Gaussian-like density function.
+In contrast to us, we only compute least squares fitting on the boundary, and then precompute some weighting for each particle to the projection operators on the boundaries.
+    %\item \textit{Unified Simulation of Elastic Rods, Shells, and Solids} \cite{10.1145/1778765.1778776}: Propose Generalized moving least squares (GMLS) to resolve limitation of MLS shape functions that require many particles in the support of a point (that are not coplonar).
+%\end{enumerate}
+
+%\subsection{Shape Matching Related Papers}
+%\begin{enumerate}
+%    \item \textit{Shape Matching} \cite{10.1145/1073204.1073216}: Meshless simulation by fitting a polynomial describing the shapes deformation using only the nodal values.
+    %\item \textit{Lattice Shape Matching} \cite{10.1145/1275808.1276480}: Voxelize model to construct a lattice of cubes.
 Use these lattice cubes to construct overlapping shape matching regions (just like clustered shape matching?).
 The original mesh is deformed using trilinear interpolation of lattice vertex positions.
-    \item \textit{Robust Real-Time Deformation of Incompressible Surface Meshes} \cite{10.1145/2019406.2019438}: Shape matching on trimeshes with overlapping regions (clustered shape matching).
+    %\item \textit{Robust Real-Time Deformation of Incompressible Surface Meshes} \cite{10.1145/2019406.2019438}: Shape matching on trimeshes with overlapping regions (clustered shape matching).
 Adds an additional volume preservation constraint.
 Position based dynamics approach to satisfying volume preservation.
-    \item \textit{Shape-Up: Shaping Discrete Geometry with Projections} \cite{10.1111/j.1467-8659.2012.03171.x}: Shape constraints by least squares fitting (like in shape matching).
+ %   \item \textit{Shape-Up: Shaping Discrete Geometry with Projections} \cite{10.1111/j.1467-8659.2012.03171.x}: Shape constraints by least squares fitting (like in shape matching).
 They have some "proximity function" indicating distance to least-squares fit, then uses projection operators to minimize proximity function (pretty much just shape matching).
-    \item \textit{Shape Matching with Oriented Particles} \cite{10.1145/1964921.1964987}: More general form for shape matching, permitting wider range of motion.
+    %\item \textit{Shape Matching with Oriented Particles} \cite{10.1145/1964921.1964987}: More general form for shape matching, permitting wider range of motion.
 Also they use shape matching projection operators for skinning, much like we do.
 For each skinning point, they specify weights with up to 4 particles (each with their own projection operator)
-    \item \textit{Fast Adaptive Shape Matching Deformations} \cite{10.5555/1632592.1632606}: Essentially same thing as Lattice Shape Matching, but instead they use an octree instead of a basic voxel grid for shape matching.
+    %\item \textit{Fast Adaptive Shape Matching Deformations} \cite{10.5555/1632592.1632606}: Essentially same thing as Lattice Shape Matching, but instead they use an octree instead of a basic voxel grid for shape matching.
 It's not super significant to mention this paper, but it does make clear that much of the followup work after shape matching never didn't emphasize it's utility as a meshless boundary only method.
 They kept converting it to a mesh-based method!
-    \item \textit{A Geometric Deformation Model for Stable Cloth Simulation} \cite{10.2312/PE/vriphys/vriphys08/039-046} Shape matching for cloth simulation.
-\end{enumerate}
+    %\item \textit{A Geometric Deformation Model for Stable Cloth Simulation} \cite{10.2312/PE/vriphys/vriphys08/039-046} Shape matching for cloth simulation.
+%\end{enumerate}
+%% start with iso geometric analysis (XCAD here)
+%\subsection{Isogeometric Analysis}
+%\begin{enumerate}
+    %\item \textit{Isogeometric Analysis Book} \cite{10.1002/9780470749081.ch7}: The book everyone references when they write Isogeometric analysis in their papers.
+    %\item \textit{Dynamic NURBS} \cite{10.1145/176579.176580}: Outline of how to simulate on NURBS with the control points as the degrees of freedom. Method used in our work.
+    %\item \textit{XCAD} \cite{10.1145/3355089.3356576}: Optimize CAD models.
+CAD embedded in hexahedral mesh, complex integration strategy.
+Uncut hexahedral elements simulated ordinarily, cut elements use XFEM that add additional DOF to account for new element shapes.
+    %\item \textit{Development of a quadratic finite element formulation based on the XFEM and NURBS} \cite{https://doi.org/10.1002/nme.3120}: XFEM to handle curve surface integration of NURBs patches.
+Complex subdividing of "X-Elements" to produce cut cells along NURBS surfaces.
+    %\item \textit{A NURBS enhanced extended finite element approach for unfitted CAD analysis} \cite{10.1007/s00466-013-0854-7}: Pretty much the same as the quadratic, but allows higher-order approximation and better handling of interface.
+    %\item \textit{A NURBS-based interface-enriched generalized finite element method for problems with complex discontinuous gradient fields} \cite{https://doi.org/10.1002/nme.4852}: Similar to other XFEM NURBS approaches.
+Uses NURBS-based enrichment functions with cut cells.
+Additional DOFs added to handle discontinuities.
+    %\item \textit{A NURBS-based generalized finite element scheme for 3D simulation of heterogeneous materials} \cite{SAFDARI2016373}: Similar to previous "NIGFEM" paper above, but now in 3D.
+    %\item \textit{Swept Volume Parameterization for Isogeometric} \cite{10.1007/978-3-642-03596-8_2} To provide volumetric simulation of NURBS they introduce a new NURBS volume parameterization (B-Spline Volumes ...
+jesus christ)
+    %\item \textit{A finite volume method on NURBS geometries and its application in isogeometric fluid–structure interaction} \cite{HEINRICH20121645}: Combines NURBS paramaterization with finite volume method (requiring a mesh for the volume).
+    %\item \textit{NURBS-Enhanced Finite Element Method (NEFEM)} \cite{10.1002/nme.2311}: Similar to the above example.
+They run an order FVM simulation and deal with the interface in a complicated manner (could this be considered XFEM?).
+%\end{enumerate}
 
-\subsection{Other Meshless Methods in Graphics}
-\begin{enumerate}
-    \item \textit{Point Based Animation }\cite{10.1145/1028523.1028542}: Purely particle based, MLS to approximate derivatives.
+%% start with standard fem stuff
+%% start with iso geometric analysis (XCAD here)
+%% talk about shape matching and simulating visual geometry but not constitutive models
+%% talk about how sparse mesh free methods fixes these things but still loses connection to modelling geometry
+%% - ^^^ throw rig space physics in here talk about constraints talk about finding method that does this by construction but is still meshes.
+%% talk about VEM has a potential solution 
+%% limitation is that it looks at piecewise linear stuff, one polynomial per element, limits expressivity
+
+%\subsection{Other Meshless Methods in Graphics}
+%\begin{enumerate}
+    %\item \textit{Point Based Animation }\cite{10.1145/1028523.1028542}: Purely particle based, MLS to approximate derivatives.
 Appears to be among the earliest meshless methods in graphics based on continuum mechanics.
-    \item \textit{Position Based Dynamics} \cite{MULLER2007109}: Operates directly on particle positions by forming set of constraints and solving for particle positions that satisfy these constraints.
+    %\item \textit{Position Based Dynamics} \cite{MULLER2007109}: Operates directly on particle positions by forming set of constraints and solving for particle positions that satisfy these constraints.
 Meshless
-    \item \textit{Projective Dynamics} \cite{10.1145/2601097.2601116}: Similar to position based dynamics but solves the constraints implicitly by minimizing energy potentials.
+    %\item \textit{Projective Dynamics} \cite{10.1145/2601097.2601116}: Similar to position based dynamics but solves the constraints implicitly by minimizing energy potentials.
 Mesh-based.
 Global solver unlike PBD which satisfies constraints locally using Gauss-Seidel.
-\end{enumerate}
+%\end{enumerate}
 
-\subsection{Virtual Element and other Element Methods}
-\begin{enumerate}
-    \item \textit{Mimetic Finite Differences} \cite{10.1142/S0218202505000832} \cite{10.1016/j.jcp.2013.07.031} (they double dipped!): Considered a close relative to VEM and framed as the predecessor to VEM (in VEM papers).
-I still haven't read on MFD yet.
-From PolyDDF: "extension of finite volume and finite difference techniques to polygons that first discretizes a prime operator (typically, the gradient or the divergence) via a boundary integral, and then derives other operators by mimicking continuous structural properties."
-    \item \textit{Basic principles of Virtual Element Method} \cite{10.1142/S0218202512500492} Original VEM paper
-    \item \textit{The Hitchhiker's Guide to Virtual Element Method} \cite{10.1142/S021820251440003X}: More understandable versions of original VEM paper.
-    \item \textit{Discrete Differential Operators on Polygonal Meshes} \cite{10.1145/3386569.3392389}: Extends VEM to do discrete differential geometry on arbitrary polygonal meshes.
-    \item \textit{FLexible Simulation of Deformable Models using Discontinuous Galerkin FEM} \cite{10.5555/1632592.1632608}: Uses ordinary hexahedral elements, but uses a cut cell-based approach to support arbitrary polyhedra on the surface.
-    \item \textit{Generalizing the finite element method: Diffuse approximation and diffuse elements} \cite{Nayroles1992GeneralizingTF} Predecessor to Element Free Galerkin.
-FEM interpolation replaced with a local Moving Least Square interpolation.
-    \item \textit{Element-free Galerkin methods} \cite{https://doi.org/10.1002/nme.1620370205}: similar to DEM, but more accurate gradients (not sure of all the differences).
-In MLS methods, they solve least squares for each particle in the domain, weighting nearby particles with a Gaussian-like density function.
-In contrast to us, we only compute least squares fitting on the boundary, and then precompute some weighting for each particle to the projection operators on the boundaries.
-    \item \textit{Unified Simulation of Elastic Rods, Shells, and Solids} \cite{10.1145/1778765.1778776}: Propose Generalized moving least squares (GMLS) to resolve limitation of MLS shape functions that require many particles in the support of a point (that are not coplonar).
-\end{enumerate}
-\subsection{Physics based Skinning}
-\begin{enumerate}
-    \item \textit{Skinning Siggraph Course} \cite{10.1145/2614028.2615427}: The linear weighting of polynomials at the exterior is very similar to skinning.
-    \item \textit{Linear Subspace Design for Real-Time Shape Deformation} \cite{10.1145/2766952}: Linear deformation subspace that uses linear blend skinning and generalized barycentric coordinates.
+%%% Might need later in methods section
+%\subsection{Physics based Skinning}
+%\begin{enumerate}
+%    \item \textit{Skinning Siggraph Course} \cite{10.1145/2614028.2615427}: The linear weighting of polynomials at the exterior is very similar to skinning.
+%    \item \textit{Linear Subspace Design for Real-Time Shape Deformation} \cite{10.1145/2766952}: Linear deformation subspace that uses linear blend skinning and generalized barycentric coordinates.
 Similarly, we have "handles" but they are represented by entire NURBS patches, and our coordinates are the output of a polynomial, whereas barycentric coordinates are linear (I only skimmed this paper, not sure if this description is fair).
-    \item \textit{Complementary dynamics} \cite{Zhang_2020}: Physics based skinning, orthogonality constraint can be seen as similar to our stability term (conformity term, error term, whatever it's called :))
-    \item \textit{Physically-Based Character Skinning} \cite{10.2312/PE.vriphys.vriphys13.025-034}: Linear blend skinning with multiple layers of skin simulated via oriented particle shape matching and position based dynamics to enforce distance constraints (avoiding unwanted intersections).
-\end{enumerate}
-\subsection{Isogeometric Analysis}
-\begin{enumerate}
-    \item \textit{Isogeometric Analysis Book} \cite{10.1002/9780470749081.ch7}: The book everyone references when they write Isogeometric analysis in their papers.
-    \item \textit{Dynamic NURBS} \cite{10.1145/176579.176580}: Outline of how to simulate on NURBS with the control points as the degrees of freedom. Method used in our work.
-    \item \textit{XCAD} \cite{10.1145/3355089.3356576}: Optimize CAD models.
-CAD embedded in hexahedral mesh, complex integration strategy.
-Uncut hexahedral elements simulated ordinarily, cut elements use XFEM that add additional DOF to account for new element shapes.
-    \item \textit{Development of a quadratic finite element formulation based on the XFEM and NURBS} \cite{https://doi.org/10.1002/nme.3120}: XFEM to handle curve surface integration of NURBs patches.
-Complex subdividing of "X-Elements" to produce cut cells along NURBS surfaces.
-    \item \textit{A NURBS enhanced extended finite element approach for unfitted CAD analysis} \cite{10.1007/s00466-013-0854-7}: Pretty much the same as the quadratic, but allows higher-order approximation and better handling of interface.
-    \item \textit{A NURBS-based interface-enriched generalized finite element method for problems with complex discontinuous gradient fields} \cite{https://doi.org/10.1002/nme.4852}: Similar to other XFEM NURBS approaches.
-Uses NURBS-based enrichment functions with cut cells.
-Additional DOFs added to handle discontinuities.
-    \item \textit{A NURBS-based generalized finite element scheme for 3D simulation of heterogeneous materials} \cite{SAFDARI2016373}: Similar to previous "NIGFEM" paper above, but now in 3D.
-    \item \textit{Swept Volume Parameterization for Isogeometric} \cite{10.1007/978-3-642-03596-8_2} To provide volumetric simulation of NURBS they introduce a new NURBS volume parameterization (B-Spline Volumes ...
-jesus christ)
-    \item \textit{A finite volume method on NURBS geometries and its application in isogeometric fluid–structure interaction} \cite{HEINRICH20121645}: Combines NURBS paramaterization with finite volume method (requiring a mesh for the volume).
-    \item \textit{NURBS-Enhanced Finite Element Method (NEFEM)} \cite{10.1002/nme.2311}: Similar to the above example.
-They run an order FVM simulation and deal with the interface in a complicated manner (could this be considered XFEM?).
-\end{enumerate}
+%    \item \textit{Complementary dynamics} \cite{Zhang_2020}: Physics based skinning, orthogonality constraint can be seen as similar to our stability term (conformity term, error term, whatever it's called :))
+%    \item \textit{Physically-Based Character Skinning} \cite{10.2312/PE.vriphys.vriphys13.025-034}: Linear blend skinning with multiple layers of skin simulated via oriented particle shape matching and position based dynamics to enforce distance constraints (avoiding unwanted intersections).
+%\end{enumerate}
 
-\subsection{Quadrature}
-\begin{enumerate}
-    \item \textit{A new method for meshless integration in 2D and 3D Galerkin meshfree methods} \cite{KHOSRAVIFARD201030}: Strategy we use for integrating over CAD model volumes.
+%Might need later in methods
+%\subsection{Quadrature}
+%\begin{enumerate}
+%    \item \textit{A new method for meshless integration in 2D and 3D Galerkin meshfree methods} \cite{KHOSRAVIFARD201030}: Strategy we use for integrating over CAD model volumes.
 Raycast along single dimension, find intersections, generate quadrature points in the intervals inside the object.
-    \item Adaptive image-based intersection \cite{DBLP:journals/tog/WangFP12}: related to our meshless integration strategy in that we could use this to account for errors in the above approach (increase ray density where we estimate high error to be)
-    \item Efficient and accurate numerical quadrature for immersed boundary methods \cite{10.1186/s40323-015-0031-y}: Finite Cell Method.
+%    \item Adaptive image-based intersection \cite{DBLP:journals/tog/WangFP12}: related to our meshless integration strategy in that we could use this to account for errors in the above approach (increase ray density where we estimate high error to be)
+%    \item Efficient and accurate numerical quadrature for immersed boundary methods \cite{10.1186/s40323-015-0031-y}: Finite Cell Method.
 "Immerses" a shape in a set of cells (mesh!) and computes quadrature over this.
 To handle curved surface they use an octree to subdivide to the desired level of accuracy.
-    \item Higher-Order Finite Elements for Embedded Simulation \cite{10.1145/3414685.3417853}: Another Finite Cell method like the above, but with a new quadrature generation method (the ones with the circles in the triangles)
-    \item \textit{Highly accurate surface and volume integration on implicit domains by means of moment-fitting} \cite{https://doi.org/10.1002/nme.4569} and \cite{https://doi.org/10.1002/nme.5343}: XCAD paper extends upon this method \myworries{still need to read these}
-\end{enumerate}
+%    \item Higher-Order Finite Elements for Embedded Simulation \cite{10.1145/3414685.3417853}: Another Finite Cell method like the above, but with a new quadrature generation method (the ones with the circles in the triangles)
+%    \item \textit{Highly accurate surface and volume integration on implicit domains by means of moment-fitting} \cite{https://doi.org/10.1002/nme.4569} and \cite{https://doi.org/10.1002/nme.5343}: XCAD paper extends upon this method \myworries{still need to read these}
+%\end{enumerate}
 
-\subsection{Misc}
-\begin{enumerate}
-	\item TRACKS: Toward Directable Thin Shells \cite{10.1145/1276377.1276439}: Petrov-Galerkin test functions for weak-form constraints that handles artifacts due to pointwise constraints.
-    \item FEM simulation of 3D deformable solids \cite{10.1145/2343483.2343501}
-    \item Fusion 360 Gallery \cite{willis2020fusion} : source of some models
-\end{enumerate}
+%Might need later
+%\subsection{Misc}
+%\begin{enumerate}
+%	\item TRACKS: Toward Directable Thin Shells \cite{10.1145/1276377.1276439}: Petrov-Galerkin test functions for weak-form constraints that handles artifacts due to pointwise constraints.
+    %\item Fusion 360 Gallery \cite{willis2020fusion} : source of some models
+%\end{enumerate}

### 2021-01-17 paper/sections/methods.tex

@@
 
-\section{D-NURB Model}
+\section{D-NURBS Model}
+We provide a brief overview of the D-NURBS formulation presented in \cite{10.1145/176579.176580}. Our models in our method are composed of a collection of NURBS surfaces.
+
+\subsection{D-NURBS Formulation}
+A point on a NURBS surface is typically expressed in the following form
+\begin{equation}
+\label{eqn:nurbs_srf}
+    \mathbf{x}(u,v) = \frac{\sum_{i=1}^{n}\sum_{j=1}^{m}  \mathbf{p}_{i,j} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
+    {\sum_{i=1}^{n}\sum_{j=1}^{m} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
+    \text{,}
+\end{equation}
+where where have a total of $nm$ control points $\mathbf{p}$ and weights $w$.
+Using the B-spline basis functions and the weighted control points, this formula gives a map from our parametric coordinates $u,v$ to their corresponding position on the surface in $\mathbb{R}^{3m}$.
+$B_{i,k}(u)$ is the B-spline basis function for the $i-th$ control point with degree $k-1$.
+
+We would like our generalized coordinates to be the control points of the model so that the dynamics operates directly on the original representation.
+Therefore, we would like to rewrite equation \ref{eqn:nurbs_srf} in the form
+\begin{equation}
+    \mathbf{x}(u,v) = \mathbf{J}(u,v)\mathbf{q}
+    \text{,}
+\end{equation}
+where $\mathbf{q} = \left[ \mathbf{p}_{1,1}^T, \dots, \mathbf{p}_{i,j}^T, \dots, \mathbf{p}_{n,m}^T \right]^T \in \mathbb{R}^{3nm}$ is the set of $nm$ control points \myworries{in the overview I just write $n$ control points, but here i'm writing $nm$} for a single NURBS surface arranged as a column vector.
+$\mathbf{J} \in  \mathbb{R}^{3 x 3nm}$ is the NURBS jacobian that maps $(u,v)$ coordinates to world positions in $\mathbb{R}^{3}$.
+
+The NURBS jacobian, $\mathbf{J}$, is a matrix composed of horizontally concatenated $3x3$ blocks where the $(i,j)$-th block is $\frac{\partial \mathbf{x}}{\partial \mathbf{p}_{i,j}}$ for the $(i,j)$-th control point.
+$\frac{\partial \mathbf{x}}{\partial \mathbf{p}_{i,j}}$ is a diagonal $3x3$ matrix where each diagonal entry $N_{i,j}(\mathbf{u})$ takes the form
+
+\begin{equation}
+\label{eqn:jacobian_diagonal}
+    N_{i,j}(u,v)
+    = \frac{w_{i,j} B_{i,k}(u)B_{j,l}(v)}{\sum_{i=1}^{n}\sum_{j=1}^{m} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
+    \text{.}
+\end{equation}
+Thus the NURBS jacobian for a single $(u,v)$ pair is written as
+\begin{equation}
+\label{eqn:uv_jacobian}
+    \mathbf{J}(u,v)
+    \left[ \frac{\partial \mathbf{x}}{\partial \mathbf{p}_{1,1}} \dots
+           \frac{\partial \mathbf{x}}{\partial \mathbf{p}_{i,j}} \dots 
+           \frac{\partial \mathbf{x}}{\partial \mathbf{p}_{n,m}}
+    \right]
+    \text{.}
+\end{equation}
+
+This formulation is a simplification of the full D-NURBS described in \cite{10.1145/176579.176580} the weights of the NURBS are fixed throughout the simulation.
+Permitting the dynamics to modify weights would likely improve the expressiveness of the kinematics, but we find fixing the weights affords sufficiently expressive results and improves performance.
+The performance improvement is a result of having fixed $N_{i,j}(u,v)$ entries over the course of the simulation.
+The full D-NURBS model requires rebuilding the jacobian and consequently the mass matrices at each time step.
+
+\subsection{Generalized Coordinates for a NURBS model}
+The above formulation describes developing the map from a single $(u,v)$ coordinate to its corresponding position on the surface.
+In our simulation we represent the entire boundary of the NURBS model, so we require samples along each surface.
+We emphasize that we only require an initial selection of $(u,v)$ coordinates across each surface and the dynamics of the simulation modifies the set of control points $\mathbf{q}$ to modify the world space maps $\mathbf{x}(u,v)=\mathbf{J}(u,v)\mathbf{q}$.
+
+Let's say for a NURBS surface on the model the DOF column vector is $\mathbf{q}_i$, which is the subset of control points in the generalized coordinates for the $i$-th surface.
+For a given surface with $m$ pairs of $\mathbf{u}=(u,v)$ coordinates sampled in parametric space, we write the jacobian for the $i$-th surface as \myworries{(should I use something like $\hat{\mathbf{J}}$ instead?)}
+\begin{equation}
+\label{eqn:surface_jacobian}
+    \mathbf{J}_i =
+    \left[ \mathbf{J}(\mathbf{u})_1^T \dots \mathbf{J}(\mathbf{u})_m^T \right]^T
+    \text{.}
+\end{equation}
+Then if $\mathbf{x}_i$ is the set of $m$ world space positions for the $i$-th surface, we may write $\mathbf{x}_i = \mathbf{J}_i \mathbf{q}_i$.
+Finally, given the full set of generalized coordinates $\mathbf{q} = \left[ \mathbf{q}_i^T, \dots, \mathbf{q}_n^T \right]^T$ with $n$ surfaces, we may write jacobian $\mathbf{J}$ for the entire model as a block diagonal matrix:
+\begin{equation}
+\mathbf{J} = \left( \begin{array}{ccccc}
+\mathbf{J}_1 &  &  &  &  \\
+ & \ddots &  &  &  \\
+ &  & \mathbf{J}_i & &  \\
+ &  &  & \ddots &  \\
+ &  &  &  & \mathbf{J}_n \\
+\end{array} \right)
+\end{equation}
+With this equation, the full set of world positions of the NURBS model may be written as $\mathbf{x} = \mathbf{J}\mathbf{q} \in \mathbb{R}^{3N}$ where $N$ is the total number of $(u,v)$ coordinates.
+
 \section{Shape Matching}
@@
 \subsection{Equations of Motion}
-This is most notable in cases where the degree of the polynomial on a NURBS is higher than the degree of the B-spline surface.
+%This is most notable in cases where the degree of the polynomial on a NURBS is higher than the degree of the B-spline surface.
 
 %
+
+% \textbf{x}(\textbf{X}) = \hat{A}M(\textbf{X}) + \bar{\textbf{x}}$ \\
+% $\hat{A} = \sum_i^{|S|}w_i(\textbf{X})A_i$
 %
-%% \textbf{x}(\textbf{X}) = \hat{A}M(\textbf{X}) + \bar{\textbf{x}}$ \\
-%% $\hat{A} = \sum_i^{|S|}w_i(\textbf{X})A_i$
 %
-%We have $\mathbf{x} = (x_0, x_1)$ and $\mathbf{X} = (X_0, X_1) $ as deformed and undeformed positions, respectively. \\
-%
-%\textbf{<insert Exposition about FEM >} \\
 %In our case, the function we are trying to evaluate is deformation with respect to the center of mass.
@@
 %
-%\subsection{Simulation on NURBs models}
-%
-%\begin{figure}[h]
-%  \centering
-%  \includegraphics[width=\linewidth]{figures/nurbs_surface}
-%  \caption{The NURBs model uses control points, shown in grey, as the original degrees of freedom.}
-%  \Description{A simple NURBS surface in $R^3$.}
-%\end{figure}
-%
-%We used the D-NURBS model (ref) for simulating on NURBs surfaces, giving an approach to simulating directly on the degrees of freedom of parametric NURBs models.
-The equation for a NURBs curve takes the form:
-%\begin{equation}
-%    \mathbf{x}(u) = \frac{\sum_{i=1}^{n} \mathbf{p}_i w_i B_{i,k}(u)}{ w_i B_{i,k}(u)}
-%\end{equation}
-%where $u$ is the parametric coordinate, $\mathbf{p}_i$ is the $i$-th control point with $n$ total control points.
-$B_{i,k}(u)$ is the B-spline basis function for the $i-th$ control point with degree $k-1$.
-NURBs curves represent a generalization of B-splines with the addition of the weight $w_i$ for each control point.
-In our case, we primarily work with NURBs surfaces, which are a generalization of the tensor-product of two B-splines:
-%\begin{equation}
-%    \mathbf{x}(u,v) = \frac{\sum_{i=1}^{n}\sum_{j=1}^{m}  \mathbf{p}_{i,j} w_{i,j} B_{i,k}(u)B_{j,l}(v)}{ w_{i,j} B_{i,k}(u)B_{j,l}(v)}
-%\end{equation}
-%where we now have two parameters $u,v$, and $m\cdot n$ control points. In D-NURBs, we see that $\mathbf{x}(u,v)$ is linear in the control points, $\mathbf{p}$, so we can rewrite the above equation as
-%\begin{equation}
-%    \mathbf{x}(u,v,\mathbf{p}) = \mathbf{J}(u,v)\mathbf{p}.
-%\end{equation}
-%Here $\mathbf{J}$ is the Jacobian matrix that maps the control points to some world space position on the surface for some $u,v$ pair.
-<elaborate on the form of J>.
-Now with these generalized coordinates, we may perform dynamics directly on the degrees of freedom of our NURBs models.
 %

### 2021-01-17 paper/sections/related.tex
correct some mistakes in related work

@@
 time-consuming, complex geometry processing pipelines. 
-For instance, the popular tetrahedral finite element approach for simulating solid elastodynamics requires robust algorithms for converting input surface geometry into 
-a volumetric tetrahedral mesh. This is a difficult problem and while significant progress has been~\cite{10.1145/2343483.2343501} made,
+For instance, the popular tetrahedral finite element approach for simulating solid elastodynamics~\cite{10.1145/2343483.2343501} requires robust algorithms for converting input surface geometry into 
+a volumetric tetrahedral mesh. This is a difficult problem and while significant progress has been made,
 even the most robust volumetric methods~\cite{Hu:2018:TMW:3197517.3201353} can be time consuming, fail to maintain correspondence between the input model and output simulation mesh,
@@
 Computer-Aided Design (CAD) software. Initial attempts used NURBS surfaces to represent the mid-surface of thin objects~\cite{10.1145/176579.176580}.
-Initial attempts at volumetric simulation relied on volumetric NURBS~\cite{10.1007/978-3-642-03596-8_2} but were limited to a narrow class of geometries.
+Volumetric simulations relied on volumetric NURBS~\cite{10.1007/978-3-642-03596-8_2} but were limited to a narrow class of geometries.
 Finite volume methods are applicable to more general geometries~\cite{HEINRICH20121645,10.1002/nme.2311} but require a volumetric mesh be generated.
@@
 it to enable boundary handling (such a mesh is called a \emph{cutcell} mesh). 
-Like volumetric meshing, this cutting operation can be difficult and our ideal method would avoid it if possible. 
-Some cutcell algorithms assume engineering/manufacturing quality input, which puts tight requirements on input models quality~\cite{10.1145/3355089.3356576}.
+Like volumetric meshing, this cutting operation can be difficult and our ideal method would avoid it. 
+Some cutcell algorithms assume engineering/manufacturing quality input, which puts tight requirements on input models~\cite{10.1145/3355089.3356576}.
 Finally these approaches require additional mechanisms to ensure that simulation results lie inside the shape space of the input model's primitives,
@@
 Shape Matching is fast~\cite{10.1145/1275808.1276480,10.5555/1632592.1632606} and meshless, but state-of-the-art methods require 
-additional modeling input to position simulation primitives ~\cite{0.1145/1964921.1964987} or limit simulation primitives to be 
+additional modeling input to position simulation primitives~\cite{10.1145/1964921.1964987} or limit simulation primitives to be 
 collections of convex polytopes~\cite{10.1145/2994258.2994260}.
-Finally, Shape Matching is tightly coupled to the position-based dynamics~\cite{muller2007position}
+Finally, Shape Matching is tightly coupled to the position-based dynamics (PBD)~\cite{muller2007position}
 time integration methodology. 
-While incredibly performant, this approach is incompatible with the constitutive models that are popular for physics-based animation as well as 
+While incredibly performant, this approach is incompatible with the constitutive models that are popular for physics-based animation, as well as 
 other time integration schemes. 
@@
 mimetic finite differences~\cite{10.1142/S0218202505000832,10.1016/j.jcp.2013.07.031} to weak-form variational problems. 
-VEM relaxed the mesh generation requirements of finite element method by enabling the solution of partial differential equations
+VEM relaxes the mesh generation requirements of the finite element method by enabling the solution of partial differential equations
 on domains partitioned with arbitrary polytopes~\cite{10.1145/3386569.3392389}. 
@@
 intersections and disconnected primitives without additional user input. 
-These features mean that our algorithm is the first truly meshless approach for the simulation of NURBS models for physics-based animation.
+These features mean that our algorithm is the first truly meshless approach for the direct simulation of NURBS models for physics-based animation.
 

### 2021-01-17 paper/sections/methods.tex
Filled out most of shape matching part of methods

@@
 \section{Methods}
-Brief overview. We have some control points. We have UV samples on the NURBS parametric domain. We have a NURBS jacobian UV to points on surface in R3.
 Given a volumetric model composed of one or more NURBS surfaces, the surface of each surface is reconstructed using a set of control points $\textbf{p}$ and their associated weights $w$.
@@
 \label{eqn:uv_jacobian}
-    \mathbf{J}(u,v)
+    \mathbf{J}(u,v) =
     \left[ \frac{\partial \mathbf{x}}{\partial \mathbf{p}_{1,1}} \dots
@@
 \begin{equation}
-\mathbf{J} = \left( \begin{array}{ccccc}
+\mathbf{J} = \left[ \begin{array}{ccccc}
 \mathbf{J}_1 &  &  &  &  \\
@@
  &  &  &  & \mathbf{J}_n \\
-\end{array} \right)
+\end{array} \right]
+\text{.}
 \end{equation}
@@
 \section{Shape Matching}
+The key feature of our method is its unique deformation map $\phi(\mathbf{X}) = \mathbf{x}$ mapping some arbitrary undeformed material point $\mathbf{X}$ to its corresponding deformed position in space $\mathbf{x}$.
+We develop this similarly to the VEM approach in that we say the function we intend to reproduce may be some highly nonlinear one, but it may be approximated accurately by a polynomial.
+Thus we seek some $\phi$ function where  $\phi^* \approx \phi$ where $\phi^*$ is the true solution.
+In $1$-D this means we wish to find a function of the form $\phi(X) = c_1 + c_2X + c_3X^2 + \dots + c_nX^{n-1}$ where each $c_i$ is an unknown polynomial coefficient and the $X^i$ terms are elements of the polynomial basis.
+
+\subsection{The Shape Matching Element}
+In typical FEM in graphics, we make use of tetrahedral or triangular elements with a piecewise-linear function space.
+In such a case our functions may be evaluated as a weighted combination of the nodal values of an element: $f(\mathbf x)=\sum_{i=0}^n f_i \phi_i(\mathbf{x})$ where $n$ is the number of nodal values and $\phi_i$ is the i-th barycentric coordinates.
+Thus in this case we have the simple deformation map $\mathbf{x(X)} = \sum_{i=0}^n \mathbf{x}_i \phi_i(\mathbf{X})$.
+However, our method contrasts this typical approach in that our elements are more general and take the form of any arbitrary boundary primitive, such as NURBS, and also that the function space in non-linear.
+
+In our algorithm, a single shape matching element is represented by a NURBS surface, a center of mass, and polynomial describing its current deformation.
+We construct these polynomials via a shape matching strategy.
+With the addition of the center of mass, we describe the shapes deformation about this point and thus the monomial basis in the case of a linear polynomial is $\left[ \mathbf 1^T (\mathbf{X - \bar{X}})^T \right]$ where $\mathbf 1^T$ is a constant $1$ column vector and $\mathbf{\bar{X}}$ is the center of mass for the element.
+To simplify the following formulas, we collect the non-constant terms of the monomial basis into a single vector $\mathbf{m}(\mathbf{X}-\mathbf{\bar{X}})$.
+For example if $\mathbf{X} = (X,Y,Z)$, $\mathbf{\bar{X}} = (\bar X, \bar Y, \bar Z)$, and $\mathbf{q}=\mathbf{X}-\mathbf{\bar{X}}$ with a quadratic polynomial we would have
+\begin{equation}
+m(\mathbf{q)} = (
+					q_1, q_2, q_3,
+                   	q_1^2, q_2^2, q_3^3,
+                   	q_1 q_2, q_2 q_3
+                ) \in \mathbb{R}^k
+                \text{,}
+\end{equation}
+where $k$ is the size of the monomial basis. Next we arrange this basis into a matrix of the form
+\begin{equation}
+\mathbf{M}(\mathbf{q}) = \left[ \begin{array}{ccc}
+\mathbf{m}(\mathbf{q}) &  &  \\
+ & \mathbf{m}(\mathbf{q})  &  \\
+ &  & \mathbf{m}(\mathbf{q})  \\
+\end{array} \right] \in \mathbb{R}^{3x3k}
+\text{.}
+\end{equation}
+With this definition of the monomial basis, we may reconstruct the deformed position $\mathbf{x}$ for some material point $\mathbf{X}$ with
+\begin{equation}
+\mathbf{x}(\mathbf{X}) = \left[\mathbf{M}(\mathbf{X-\bar{X}}) \quad \mathbf{I} \right]
+\left[ \begin{array}{c}
+\mathbf{c} \\
+\mathbf{t} \\
+\end{array} \right]
+\text{,}
+\end{equation}
+where $\mathbf{c} \in \mathbb{R}^{3k}$ is the set of polynomial coefficients for the non-constant terms, $\mathbf{t} \in \mathbb{R}^3$ is the constant term coefficients, and $\mathbf{I} \in \mathbb{R}^{3x3}$ is an identity matrix.
+The reason for separating out the constant term becomes more evident when we examine case where we have multiple elements.
+
+The remaining step in defining a single element is to describe how to solve for the polynomial coefficients.
+For a single element, we have $n$ material points $(\mathbf{X}_1 \dots \mathbf{X}_n)$ and their corresponding deformed positions $(\mathbf{x}^*_1 \dots \mathbf{x}^*_n)$.
+As these elements are defined on a boundary, these are both known throughout the simulation.
+Therefore we may form a fitting problem to solve for the polynomial coefficients that minimize the error $\sum_i^n ||\mathbf{x}_i^* - \mathbf{X}_i ||^2$.
+We minimize this error in a least squares sense where we first form the system
+\begin{equation}
+\left[\begin{array}{c}
+\mathbf{M}(\mathbf{X_1-\bar{X}}) \quad \mathbf{I} \\
+\vdots \\
+\mathbf{M}(\mathbf{X_i-\bar{X}}) \quad \mathbf{I} \\
+\vdots \\
+\mathbf{M}(\mathbf{X_n-\bar{X}}) \quad \mathbf{I} \\
+\end{array} \right]
+\left[ \begin{array}{c}
+\mathbf{c} \\
+\mathbf{t} \\
+\end{array} \right] = 
+\left[ \begin{array}{c}
+\mathbf{x}_1^* \\
+\vdots \\
+\mathbf{x}_i^* \\
+\vdots \\
+\mathbf{x}_n^*
+\end{array} \right]
+\text{,}
+\end{equation}
+and then in the more compact form
+\begin{equation}
+\left[
+\mathbf{A} \quad \mathbf{S}
+\right]
+\left[ \begin{array}{c}
+\mathbf{c} \\
+\mathbf{t} \\
+\end{array} \right] = \mathbf{b}
+\text{,}
+\end{equation}
+where $\mathbf{A}$ is the block containing the monomial matrices of the system, and $\mathbf{S}$ is the block of identity matrices.
+Next, to reduce the degrees of freedom of this system we let the deformed center of mass (the constant term coefficients), $t$, be equal to the mean of the deformed positions influenced by this center of mass.
+In the case of one element we have $t = \frac{1}{n}\sum_i^n x_i^*$.
+We describe the details on center of mass selection in section \myworries{insert here}.
+In matrix form we have
+\begin{equation}
+t = 
+\underbrace{
+\left[\begin{array}{ccccccc}
+\frac{1}{n} & & \cdots \frac{1}{n} & & \\
+& \frac{1}{n} & \cdots & \frac{1}{n} & \\
+& & \frac{1}{n} \cdots & & \frac{1}{n} \\
+\end{array} \right]}_{T} 
+\mathbf{b}
+\text{.}
+\end{equation} (\myworries{spacing on this looks bad....}). Finally, this leaves us with the system
+\begin{equation}
+\mathbf{A}\mathbf{c}=(\mathbf{I}-\mathbf{ST})\mathbf{b}
+\text{.}
+\end{equation}
+Like in the original Shape Matching \cite{10.1145/1073204.1073216}, we solve for the coefficients by minimizing the error in a least squares sense, resulting in
+\begin{equation}
+\begin{split}
+\mathbf{c} & = (\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T(\mathbf{I-ST})\mathbf{b} \\
+           & = \mathbf{L}\mathbf{b}
+\end{split}
+\text{.}
+\end{equation}
+This $\mathbf{L}$ matrix thus gives us our file ingredient in the element, coefficients defining the polynomial.
+We emphasize that $\mathbf{L}$ is constant throughout the simulation, therefore as $\mathbf{b}$ varies in the simulation, $\mathbf{Lb}$ gives us a new set of polynomial coefficients as $\mathbf{b}$ changes.
+
+\subsection{Shape Matching with Multiple Elements}
+\subsection{Hierarchical fitting}
+A problem may emerge in the shape matching process where the shape matching system may be rank deficient.
+
 \section{Blending Polynomials}

### 2021-01-17 paper/sections/methods.tex
finished draft of shape matching part of paper. I really couldnt be a slower writer....

@@
 \section{D-NURBS Model}
+\myworries{TODO: Most of these matrices should be paranthesis matrices}
 We provide a brief overview of the D-NURBS formulation presented in \cite{10.1145/176579.176580}. Our models in our method are composed of a collection of NURBS surfaces.
@@
 \begin{equation}
+\label{eqn:fitting_one}
 \left[
@@
 \begin{equation}
+\label{eqn:t_solution}
 t = 
@@
 \begin{equation}
+\label{eqn:c_one_element}
 \begin{split}
@@
 \subsection{Shape Matching with Multiple Elements}
+In a typical simulation we will have multiple shape matching elements where the boundary of each element is defined by a different NURBS surface.
+Therefore for the entire model each shape has a boundary, a polynomial describing its deformation, as well as an associated center of mass.
+These centers of mass may not be distinct, meaning that a single center of mass may be shared among multiple surfaces.
+This sharing of center of masses is typical and in some cases required (\myworries{reference center of mass section that describes how for planar surfaces we need to share COM to get a COM out of plane}).
+This leads to a new definition of the center of mass where it is the mean of all the boundary points associated with it.
+
+In defining the shape matching fitting problem, we will form a system similar to equation \ref{eqn:fitting_one}.
+In this case where we have $n$ shape elements our set of unknown coefficients is $\hat{\mathbf{c}} = \left[\mathbf{c}_1^T \dots \mathbf{c}_i^T \dots \mathbf{c}_n^T \right]^T \in \mathbb{R}^{3kn}$ where $\mathbf{c}_i$ corresponds to the unknown polynomial coefficients for the $i$-th shape element.
+The set of boundary positions is similarly written as $\hat{\mathbf{b}}=\left[\mathbf{b}_1^T \dots \mathbf{b}_i^T \dots \mathbf{b}_n^T \right]^T \in \mathbb{R}^{3N}$ where $N$ is the total number of boundary points across all elements.
+If we have $m$ centers of mass we write the set of polynomial constant terms as $\hat{\mathbf{t}} = \left[\mathbf{t}_1^T \dots \mathbf{t}_i^T \dots \mathbf{t}_m^T \right]^T \in \mathbb{R}^{3m}$.
+With these terms, we write the polynomial fitting system for the entire model as
+\begin{equation}
+\begin{pmatrix}
+ \mathbf{A}_1 &      &			  &			&				&\mathbf{S}_1\\  
+ &            \ddots &            &  		&				&\vdots\\ 
+ &            &      \mathbf{A}_i &   		&				&\mathbf{S}_i\\
+ &            &      &            \ddots	&				&\vdots\\
+ &            &      &            &			\mathbf{A}_n	&\mathbf{S}_n
+\end{pmatrix}
+\begin{bmatrix}
+ \hat{\mathbf{c}}  \\ 
+ \hat{\mathbf{t}}   
+\end{bmatrix}
+=
+\hat{\mathbf{b}}
+\text{,}
+\end{equation}
+where $\mathbf{A}_i$ is the monomial basis matrix for the $i$-th shape element, and $\mathbf{S}_i$ is a selection matrix that extracts the $j$-th center of mass which is associated with the $i$-th shape.
+We simplify this as
+\begin{equation}
+\begin{pmatrix}
+\hat{\mathbf{A}} & \hat{\mathbf{S}}
+\end{pmatrix}
+\begin{bmatrix}
+ \hat{\mathbf{c}}  \\ 
+ \hat{\mathbf{t}}   
+\end{bmatrix}
+=
+\hat{\mathbf{b}}
+\text{.}
+\end{equation}
+Like in the single element case, we set the $\hat{\mathbf{t}}$ to be equal to the mean of a set of boundary positions, but in this case each $\mathbf{t}_j$ is equal to the mean of only the boundary positions associated with the $j$-th center of mass.
+The definition of the values of $\hat{\mathbf{t}}$ is similar to \ref{eqn:t_solution}:
+\begin{equation}
+\hat{\mathbf{t}} =
+\underbrace{
+\begin{bmatrix}
+\mathbf{T}_1 \\ 
+\vdots       \\
+\mathbf{T_m}
+\end{bmatrix}}_{\hat{\mathbf{T}}}
+\text{,}
+\end{equation}
+
+where the entries each $\mathbf{T}_i \in \mathbb{R}^{3N}$ take a similar form, but are only nonzero for columns corresponding to associated boundary vertices.
+
+Finally, to solve for the full set of non-constant term coefficients, $\hat{\mathbf{c}}$, we solve in the same manner as equation \ref{eqn:c_one_element} where we have
+\begin{equation}
+\label{eqn:c_multiple_elements}
+\begin{split}
+\hat{\mathbf{c}} & = (\hat{\mathbf{A}}^T\hat{\mathbf{A}})^{-1}\hat{\mathbf{A}}^T(\mathbf{I-\hat{S}\hat{T})\hat{\mathbf{b}}} \\
+           & = \mathbf{L}\hat{\mathbf{b}}
+\end{split}
+\text{.}
+\end{equation}
+
 \subsection{Hierarchical fitting}
-A problem may emerge in the shape matching process where the shape matching system may be rank deficient.
+A problem may emerge in the shape matching process where the system may be underdetermined.
+To account for this we solve for the coefficients in a hierarchical manner where we initially solve for the linear terms and at each step solve the coefficients for the next higher order of deformation.
+The produces a set of smaller systems which are guaranteed to be determined \myworries{probably need to justify this more}.
+The intuition behind this can be described by the \textit{hierarchical ordering principle}, which argues that the behavior of a system is typically dominated by lower order effects \myworries{cite something here.}.
+To describe our fitting strategy, we begin with the following shape matching system for a single element
+\begin{equation}
+\label{eqn:fitting_one}
+\left[
+\mathbf{A} \quad \mathbf{S}
+\right]
+\left[ \begin{array}{c}
+\mathbf{c} \\
+\mathbf{t} \\
+\end{array} \right] = \mathbf{b}
+\text{.}
+\end{equation}
+We will show the hierarchical fitting on a polynomial with a set of coefficients for the linear terms, $\mathbf{c}_{\text{lin}}$, and the coefficients for the quadratic terms $\mathbf{c}_{\text{quad}}$.
+Similarly, we also have $\mathbf{A}_{\text{lin}}$ and $\mathbf{A}_{\text{quad}}$ which are the blocks of $\mathbf{A}$ corresponding to the columns for the linear and quadratic monomial terms, respectively.
+First we construct a system for the linear terms:
+\begin{equation}
+\label{eqn:fitting_linear}
+\left[
+\mathbf{A}_{\text{lin}} \quad \mathbf{S}
+\right]
+\left[ \begin{array}{c}
+\mathbf{c}_{\text{lin}} \\
+\mathbf{t} \\
+\end{array} \right] = \mathbf{b}
+\text{,}
+\end{equation}
+which leads to the following expression for the linear coefficients
+\begin{equation}
+\label{eqn:c_linear}
+\begin{split}
+\mathbf{c}_{\text{lin}} & = (\mathbf{A}_{\text{lin}}^T\mathbf{A}_{\text{lin}})^{-1}\mathbf{A}_{\text{lin}}^T(\mathbf{I-ST})\mathbf{b} \\
+           & = \mathbf{L}_{\text{lin}}\mathbf{b}
+\end{split}
+\text{.}
+\end{equation}
+where $\mathbf{T}$ is defined in equation \ref{eqn:t_solution}. For the quadratic system, we subtract linear term estimate of the solution and solve for the coefficients that fit this residual:
+\begin{equation}
+\label{eqn:fitting_quadratic}
+\left[
+\mathbf{A}_{\text{quad}} \quad \mathbf{S}
+\right]
+\left[ \begin{array}{c}
+\mathbf{c}_{\text{quad}} \\
+\mathbf{t} \\
+\end{array} \right] = \mathbf{b} - \mathbf{A}_{\text{lin}}\mathbf{c}_{\text{lin}}
+\text{.}
+\end{equation}
+Using the solution for the linear coefficients in \ref{eqn:c_linear}, we arrive at the solution for the quadratic coefficients:
+\begin{equation}
+\label{eqn:c_quad}
+\begin{split}
+\mathbf{c}_{\text{quad}} & = (\mathbf{A}_{\text{quad}}^T\mathbf{A}_{\text{quad}})^{-1}\mathbf{A}_{\text{quad}}^T((\mathbf{I-ST})\mathbf{b} - \mathbf{A}_{\text{lin}}\mathbf{c}_{\text{lin}}) \\
+		   & = (\mathbf{A}_{\text{quad}}^T\mathbf{A}_{\text{quad}})^{-1}\mathbf{A}_{\text{quad}}^T(\mathbf{I-ST} - \mathbf{A}_{\text{lin}}\mathbf{L}_{\text{lin}})\mathbf{b} \\
+           & = \mathbf{L}_{\text{quad}}\mathbf{b}
+\end{split}
+\text{.}
+%\mathbf{c}_{\text{quad}} & = (\mathbf{A}_{\text{quad}}^T\mathbf{A}_{\text{quad}})^{-1}\mathbf{A}_{\text{quad}}^T((\mathbf{I-ST})\mathbf{b} - \mathbf{A}_{\text{lin}}\mathbf{c}_{\text{lin}})
+\end{equation}.
+The final form for the full $\mathbf{L}$ matrix is thus given by
+\begin{equation}
+\mathbf{L} =
+\begin{bmatrix}
+\mathbf{L}_{\text{lin}} \\
+\mathbf{L}_{\text{quad}} \\
+\mathbf{T}
+\end{bmatrix}
+\end{equation}
+
+The recursive nature of this process permits a more general description for arbitrary orders, which we write in algorithm \myworries{TODO}.
 

### 2021-01-17 paper/sections/methods.tex
draft of the blending weights section

@@
 \text{.}
-%\mathbf{c}_{\text{quad}} & = (\mathbf{A}_{\text{quad}}^T\mathbf{A}_{\text{quad}})^{-1}\mathbf{A}_{\text{quad}}^T((\mathbf{I-ST})\mathbf{b} - \mathbf{A}_{\text{lin}}\mathbf{c}_{\text{lin}})
-\end{equation}.
+\end{equation}
 The final form for the full $\mathbf{L}$ matrix is thus given by
@@
 
-\section{Blending Polynomials}
+\section{The Deformation Map}
+At this point we have shown how to fit a polynomial describing the deformation of each surface.
+With these polynomial coefficients coupled with their associated centers of mass, we can construct an estimate of some arbitrary material point $\mathbf{X}$ using the deformation described by the surface.
+This lends itself to a general definition for the deformed position of a point as a weighted sum of the polynomials, which yields a unique polynomial for the point.
+Thus we arrive at a definition for the deformed position, $\mathbf{x}$, of a material point $\mathbf{X}$ in the undeformed space:
+\begin{equation}
+	\mathbf{x}(\mathbf{X}) = \sum_i^n w_i(\mathbf{X}) \left[ \mathbf{M}(\mathbf{X-\bar{X}}_j)\mathbf{c}_i + \mathbf{t}_j \right]
+	\text{,}
+\end{equation}
+where we have $n$ surfaces, and $w_i$ and $\mathbf{c}_i$ are the weights and the coefficients for the $i$-th surface, respectively.
+The terms $\mathbf{\bar{X}}_j$ and $\mathbf{t}_j$ the undeformed and deformed centers of mass, respectively, that are associated with the $i$-th surface.
+
+\subsection{Computing the Blending Weights}
+All that remains for the deformation map is to compute blending weights.
+Our choice of the weighting function is guided by the simple idea that that for a given material point its weight should be highest to whichever surfaces it is closest.
+This idea lends itself to a simple distance-based weighting method.
+The first step in this process is to compute \textit{distance weights}.
+For a given point $\mathbf{X}$ and a set of $n$ we wish to compute a set of distance weights to each surface $\mathbf{\theta}(\mathbf{X}) = \left(\theta_1, \dots, \theta_i, \dots, \theta_n \right)$.
+
+For the $i$-th surface, we first find the distance, $d_{\text{primary}}$, from the point $\mathbf{X}$ to the closest point, $\mathbf{p}$, on the surface.
+Next we wish to satisfy the condition that the distance weight is highest on the surface and it decays towards zero as it approaches other surfaces.
+To achieve this, we shoot a ray from $\mathbf{p}$ towards $\mathbf{X}$ and check for intersections between this ray and the other $n-1$ surfaces.
+If this ray hits another surface, we let $d_{\text{total}}$ be the distance from $\mathbf{p}$ to the closest ray hit $\mathbf{q}$.
+These two distances enable us to write a simple linear function to compute the \textit{distance weight}:
+\begin{equation}
+\theta_i = \max (1.0 - \frac{d_{\text{primary}}}{\min (d_{\text{total}}, D)}, 0.0)
+\text{,}
+\end{equation}
+where $D$ is a fixed cutoff distance parameter.
+We see that this distance weight is the result of interpolating along this ray from $\mathbf{p}$ to $\mathbf{q}$ (or a fixed point that depends on $D$).
+Importantly, this gives a result where the distance weight to surface $i$ is equal to $1$ if it lies on $i$ and $0$ if it lies on another surface.
+More elaborate schemes may be developed for computing distances, such as using the Boundary Element Method, but we find this simple linear function to perform well for our purposes.
+
+Now that we have a set of distances weights $\mathbf{theta}$, we seek to form a final set of \textit{blending weights} $\mathbf{w}(\mathbf{X}) =  \left( w_1, \dots, w_i, \dots, w_n \right)$ to blend the polynomials of the $n$ shape elements.
+In constructing these blending weights, we require that they adhere to a set of conditions.
+We require that (1) each weight is non-negative, (2) the sum of the weights is equal to 1 (\textit{partition of unity}), (3) and that if a point lies on a surface $i$, it will have a weight of 1 to $i$ and consequently zero to the others.
+To produce these blending weights we form a quadratic program that satisfies these constraints:
+\begin{equation}
+\begin{aligned}
+\mathbf{w}(\mathbf{X}) = \min_{\mathbf{w}} \quad & \mathbf{w}^T \Theta(\mathbf{X}) \mathbf{w}    \\
+\textrm{s.t.} \quad & 0 \leq w_i \leq 1                     \\
+                    &   \sum_i^n w_i = 1                      \\
+\end{aligned}
+\end{equation}
+
+where
+\begin{equation}
+     \Theta(\mathbf{X}) = \text{diag}\left( \frac{1}{\theta_1},\frac{1}{\theta_2},\dots,\frac{1}{\theta_n}\right)
+\end{equation}
+We see that through the use of linear constraints we satisfy conditions (1) and (2).
+Satisfying the third condition (3) is achieved through using the distance weights.
+Intuitively, we see that for some small $\theta_i$, $\frac{1}{\theta_i}$ will be very large, encouraging the quadratic program to assign a small weight for $w_i$.
+Conversely, we see the opposite effect for large values of $\theta_i$.
+
+\section{Selecting Centers of Mass}
+\myworries{TODO. waiting to implement this before I start writing how we do it}
 \section{Raycasting Quadrature}
@@
 
-%
-
 % \textbf{x}(\textbf{X}) = \hat{A}M(\textbf{X}) + \bar{\textbf{x}}$ \\
 % $\hat{A} = \sum_i^{|S|}w_i(\textbf{X})A_i$
-%
-%
-%In our case, the function we are trying to evaluate is deformation with respect to the center of mass.
-In VEM, instead of reconstructing values of our function using nodal values, we seek to find a a set of polynomial coefficients that map some arbitrary undeformed position to a polynomial approximation of its deformed position.
-This set of polynomial coefficients is referred to as the projection operator, $\mathbf{\Pi}$.
-The key point here is that $\mathbf{\Pi}$ is constructed strictly using positions defined on the boundary, avoiding the need for explicit representation of the interior.
-\\
-%
-%\subsection{Shape Matching: }
-%In shape matching, authors seek a polynomial transformation, $\mathbf{\Pi}$ that minimizes $\sum_i ||\mathbf{\Pi}\mathbf{M(X)} - \mathbf{p_i}||^2$ where $\mathbf{M(X)}$ is the monomial basis of the $i$-th boundary vertex and $\mathbf{p_i}$ is current displacement from the center of mass for the $i$-th point.
-We argue this paper can be interpreted as a sort of virtual element method.
-\\
-%
-%\subsection{Reconstructing deformed positions}
-%A general form for the monomial basis takes the following form <insert here>. For example, in 2D with quadratic deformation we have:
-%\begin{equation}
-%    \mathbf{M(X)} =\begin{bmatrix}
-%    X_0 - \bar{X}_0  \\ 
-%    X_1 - \bar{X}_1  \\ 
-%    (X_0 - \bar{X}_0)^2  \\ 
-%    (X_1 - \bar{X}_1)^2  \\ 
-%    (X_0 - \bar{X}_0)(X_1 - \bar{X}_1) 
-%    \end{bmatrix} 
-%\end{equation}
-%
-%    
-%We write the estimated deformed position $\mathbf{x}$ of some arbitrary undeformed position $\mathbf{X}$ as
-%\begin{equation}
-%     \mathbf{x}(\mathbf{X}) = \hat{\mathbf{\Pi}}\mathbf{M}(\mathbf{X}) + \bar{\mathbf{x}} 
-%\end{equation}
-%where $\mathbf{\hat{\Pi}}$ is the blended projection operator for this particular point:
-%\begin{equation}
-%    \hat{\mathbf{\Pi}}\mathbf{(X)} = \sum_i^{|S|}w_i(\mathbf{X})\mathbf{\Pi}_i 
-%\end{equation}
-%The bar symbol indicates the center of mass. For example, $\bar{\mathbf{x}}$ is the deformed object's center of mass, which we compute as the mean of the boundary values 
-%\begin{equation}
-%    \bar{\mathbf{x}} = \frac{1}{n}\sum_i^n x_i 
-%\end{equation}
-%Here $w_i$ is some weighting function for the $i$-th shape. There are many possible forms for this weighting function. Right now we compute these weights by forming the following optimization: \\
-%<insert here>
-%In order to form the equations of motion, we need to re-express the above equation in terms of the nodal values, $\mathbf{q}$.
-To arrive it these we note that each $\mathbf{\Pi}_i$ is a function of $\mathbf{q}$.
-%\begin{equation}
-%    \mathbf{\Pi}_i(\mathbf{q}) = \mathbf{P(E_i q)M(E_i q_0)^{\dagger}}
-%\end{equation}
-%\myworries{I'm not a fan of what I wrote here. The MLS-MPM paper presents a similar idea when talking about EFG, and I like their notation a bit better}
-%This is the shape matching problem's least squares solution where $M(q_0)^{\dagger}$ is the Moore-Penrose pseudo-inverse of the monomial vectors stacked row-wise (add example equation).
-The matrix $\mathbf{E}_i$ selects the degrees of freedom from $\mathbf{q}$ associated with the $i$-th shape.
-The function $\mathbf{P(q)}$ evaluates the displacement of these values from the current center of mass.
-We observe that the above equation is linear in $\mathbf{q}$ so we may rewrite the equation for $\mathbf{x}$ as
-%\begin{equation}
-%    \mathbf{x(X)} = \mathbf{J_{\hat{\Pi}}(X)q}
-%\end{equation}
-%
-%
-%\subsection{Generalized Coordinates}
-%With D-NURBs simulation layered on the VEM simulation, we have two Jacobians: $\mathbf{J}(u,v)$ mapping controls points, $\mathbf{p}$ onto surface positions, and then $\mathbf{J_{\hat{\Pi}}(X)}$ which maps surface positions $\mathbf{q}$ to their deformed positions.
-The Jacobian $\mathbf{J}(u,v)$  may be precomputed, and each point sampled on a surface of a NURBs has an associated $\mathbf{J}(u,v)$.
-With these two Jacobians we may express a single $u,v$ pair's deformed position as
-%\begin{equation}
-%    \mathbf{x}(u,v,\mathbf{p}) = \mathbf{J_{\hat{\Pi}}(X)}\mathbf{J}(u,v)\mathbf{p}
-%\end{equation}
-%and the undeformed position
-%\begin{equation}
-%    \mathbf{X}(u,v) = \mathbf{J}(u,v)\mathbf{p_0}
-%\end{equation}
-%where $\mathbf{p_0}$ is the initial values for the control points.
 %
@@
 %
-%\begin{equation}
-%\begin{aligned}
-%\mathbf{w}(\mathbf{X}) = \min_{\mathbf{w}} \quad & \mathbf{w}^T \Theta(\mathbf{X}) \mathbf{w}    \\
-%\textrm{s.t.} \quad & 0 \leq w_i \leq 1                     \\
-%                    &   \sum_i^n w_i = 1                      \\
-%\end{aligned}
-%\end{equation}
-%
-%where
-%\begin{equation}
-%     \Theta(\mathbf{X}) = \text{diag}\left( \frac{1}{\theta_1},\frac{1}{\theta_2},\dots,\frac{1}{\theta_N}\right)
-%\end{equation}
-%In computing the theta values, we require that they a approach 1 at the target surface and fall off to zero as a point becomes closer to other surfaces.
-To compute this for some surface, we shoot a ray from the point to the nearest point on the surface.
-Then in the opposite direction we find the nearest intersection to any other surfaces.
-We compute the theta with the following linear function
-%
-%\begin{equation}
-%theta = \max (1.0 - \frac{d_1}{\min (d_2, D)}, 0.0)
-%\end{equation}
-%where $D$ is a fixed cut-off distance.
-%
 %

### 2021-01-17 paper/sections/methods.tex
added very rough draft for the raycasting quadrature section

@@
 \section{Raycasting Quadrature}
+The last step before we may discuss the dynamic simulation is to describe our strategy for generating material points $\mathbf{X}$ in the undeformed model.
+Our volume integration strategy is based on \cite{KHOSRAVIFARD201030} which we briefly describe in the following section
+\subsection{Review of the Meshless Method for Domain Integral Evaluation in \cite{KHOSRAVIFARD201030}}
+This paper addresses the integration of volume integrals of the form $\mathbf{I} = \int_\Omega f(\mathbf{x}) d\Omega$ where $\mathbf{x} = (x_1,x_2,x_3)$.
+The authors show that through application of the Green-Gauss theorem, this integral may be rewritten as 2D integral over a $yz$ plane where at each position along the plane we evaluate a line integral along the $x$-axis.
+To evaluate these line integrals, a raycasting procedure is performed to find a set of intersection along the line.
+These intersection define integration domains in which we sample 3D integration points.
+As a result of integrating over the 2D plane and raycasting at each 2D point, we produce an integration rule over the entire domain that can converge to the true integral value.
+
+In our approach we modify the raycasting procedure by allowing integration over intersecting NURBS surfaces.
+It is common in NURBS modeling to produce models in which many surfaces overlap.
+To account for this we augment the raycasting intersection step to use an approach common in Constructive Solid Geometry (CSG).
+For the set of intersection intervals, we instead take the union over all intervals to yield a new set of intervals handling intersecting solids.
+We find that this extension serves as a simple and practical approach to support inexact models (\myworries{not sure what to right here}).
+We note that this solution is not perfect and making this intersection step more robust serves as future work.
+
 \section{Dynamic Simulation}

### 2021-01-17 paper/sections/methods.tex
Started dynamics method

@@
 \begin{equation}
-\label{eqn:fitting_one}
+\label{eqn:fitting_one_copy}
 \left[
@@
 \begin{equation}
+\label{eqn:deformation_map}
 	\mathbf{x}(\mathbf{X}) = \sum_i^n w_i(\mathbf{X}) \left[ \mathbf{M}(\mathbf{X-\bar{X}}_j)\mathbf{c}_i + \mathbf{t}_j \right]
@@
 \section{Dynamic Simulation}
+With our definition of the deformation map as well as a strategy for integrating the volume, we now have the necessary ingredient to perform dynamic simulation.
+We follow the Variational Mechanics approach that defines a pair of kinetic and potential energies, and we apply the Euler-Lagrange equation to develop our equations of motion.
+
+Before we define these energies we first would like to simplify the form of the deformation map presented in equation \ref{eqn:deformation_map}. We may write $\mathbf{x}(\mathbf{X})$ in matrix form as
+\begin{equation}
+\mathbf{x}(\mathbf{X}) = \underbrace{\left[w_1\mathbf{M}(\mathbf{X - \bar{X}}_j) \dots w_n\mathbf{M}(\mathbf{X - \bar{X}}_j) \middle| \sum_i^{n_1}w_1^i\mathbf{I} \dots \sum_i^{n_m}w_m^i\mathbf{I} \right]}_{\mathbf{Y}(\mathbf{X})}
+\begin{bmatrix}
+\mathbf{\hat{c}} \\
+\mathbf{\hat{t}}
+\end{bmatrix}
+\text{,}
+\end{equation}
+\myworries{I'm not sure how to write this matrix form for x(X), especially on finding the best notation for the center of mass indexing. So this above equation is temporary.}
+where we have $n$ shape elements, $m$ centers of mass. We simplify the above equation and use the following form
+\begin{equation}
+\label{eqn:x_simple_form}
+\mathbf{x}(\mathbf{X}) = \mathbf{Y}(\mathbf{X})\mathbf{L}\hat{\mathbf{b}}
+\text{,}
+\end{equation}
+where we make use of equation \ref{eqn:c_multiple_elements} to rewrite this in terms of the boundary points $\mathbf{\hat{b}}$.
+As one final step, we rewrite this again in terms of the control points of our NURBS surfaces:
+\begin{equation}
+\label{eqn:compact_x_map}
+\mathbf{x}(\mathbf{X}) = \mathbf{Y}(\mathbf{X})\mathbf{LJq}
+\text{,}
+\end{equation}
+where we use the fact that $\mathbf{\hat{b}} = \mathbf{Jq}$.
+\myworries{not a fan of this, perhaps we change $\hat{b}$ to something else...} This gives us an expression for the deformed position in terms of our generalized coordinates, $\mathbf{q}$.
+
 \subsection{Generalized Inertia}
+The kinetic energy for a model may be expressed as
+\begin{equation}
+T = \frac{1}{2}\int_{ \Omega} \rho \dot{\mathbf{x}}^T\dot{\mathbf{x}} d\Omega
+\text{,}
+\end{equation}
+where $\Omega$ is the 3D integration domain, $\rho$ is the density, and $\dot{\mathbf{x}}$ is the velocity for some point in the domain.
+Using the deformation map defined in \ref{eqn:compact_x_map}, we may rewrite this kinetic energy as
+\begin{equation}
+\begin{split}
+T & = \frac{1}{2}\int_{ \Omega} \rho \dot{\mathbf{q}}^T \mathbf{J}^T\mathbf{Y(X)}^T\mathbf{Y(X)}\ d\Omega \\
+  & = \frac{1}{2} \dot{\mathbf{q}}^T \mathbf{J}^T \left[ \int_{ \Omega} \rho \mathbf{Y(X)}^T\mathbf{Y(X)} d\Omega \right] \mathbf{J}\dot{\mathbf{q}} \\
+  & = \frac{1}{2} \dot{\mathbf{q}}^T \mathbf{M} \dot{\mathbf{q}}
+\end{split}
+\text{.}
+\end{equation}
+The second line is a result of the fact that all terms except for $\mathbf{Y(X)}$ are spatially constant, allowing us to move these outside the volume integral, which enables fast assembly of the mass matrix, $\mathbf{M}$.
+\subsection{Error Potential}
 \subsection{Generalized Forces}
 \subsection{Equations of Motion}
+\subsection{Time Integration}
 %This is most notable in cases where the degree of the polynomial on a NURBS is higher than the degree of the B-spline surface.
 
-% \textbf{x}(\textbf{X}) = \hat{A}M(\textbf{X}) + \bar{\textbf{x}}$ \\
-% $\hat{A} = \sum_i^{|S|}w_i(\textbf{X})A_i$
-%
-%\subsection{Deformation Gradient}
-%\begin{equation}
-%        \mathbf{F} = \frac{\partial \mathbf{x}}{\partial \mathbf{X}}
-%               = \frac{\partial \hat{\mathbf{\Pi}}\mathbf{M(X)}}{\partial \mathbf{X}}
-%               = \hat{\mathbf{\Pi}} \frac{\partial \mathbf{M(X)}}{\partial \mathbf{X}} 
-%\end{equation}
-%The $\frac{\partial \mathbf{M(X)}}{\partial \mathbf{X}}$ terms may be precomputed and using the previous $\mathbf{M(X)}$ example, we have
-%\begin{equation}
-%     \frac{\partial \mathbf{M(X)}}{\partial \mathbf{X}} = \begin{bmatrix}
-%     1&0 \\ 
-%     0&1 \\ 
-%     2(X_0 - \bar{X}_0)&0 \\ 
-%     0&2(X_1 - \bar{X}_1) \\ 
-%     (X_1 - \bar{X}_1)&(X_0 - \bar{X}_0) 
-%    \end{bmatrix}
-%\end{equation}
-%
-%In computing the forces and the stiffness matrix, we require the gradient of the deformation gradient with respect to the configuration,  $\frac{\partial \mathbf{F}}{\partial \mathbf{q}}$.
+\subsection{Deformation Gradient}
+\begin{equation}
+        \mathbf{F} = \frac{\partial \mathbf{x}}{\partial \mathbf{X}}
+               = \frac{\partial \hat{\mathbf{\Pi}}\mathbf{M(X)}}{\partial \mathbf{X}}
+               = \hat{\mathbf{\Pi}} \frac{\partial \mathbf{M(X)}}{\partial \mathbf{X}}
+\end{equation}
+
+In computing the forces and the stiffness matrix, we require the gradient of the deformation gradient with respect to the configuration,  $\frac{\partial \mathbf{F}}{\partial \mathbf{q}}$.
 Usually the form of this gradient is simple.
@@
 In our case, the presence of the projection operator $\hat{\mathbf{\Pi}}$ means that each quadrature point is computed using the nodal values of the entire model.
-%
-%\begin{equation}
-%    \frac{\partial \mathbf{F}}{\partial \mathbf{q}} = \frac{\partial \mathbf{P}}{\partial q}\mathbf{M(q_0)}^\dagger  \frac{\partial \mathbf{M(X)}}{\partial \mathbf{X}} 
-%\end{equation}
-%
-%
-%Naively handled, this gives us a $\frac{\partial \mathbf{F}}{\partial \mathbf{q}}$ of size $3 x |q|$.
+
+\begin{equation}
+    \frac{\partial \mathbf{F}}{\partial \mathbf{q}} = \frac{\partial \mathbf{P}}{\partial q}\mathbf{M(q_0)}^\dagger  \frac{\partial \mathbf{M(X)}}{\partial \mathbf{X}} 
+\end{equation}
+
+Naively handled, this gives us a $\frac{\partial \mathbf{F}}{\partial \mathbf{q}}$ of size $3 x |q|$.
 However, many of the columns of have values all near zero, allowing us to prune many columns for each quadrature point.
@@
 \\
-%
-%
-%\subsection{Variational Mechanics}
-%We have now provided all the necessary ingredient to produce equations of motion for deformable bodies, which we derive via Euler-Lagrange equation. We begin with the Lagrangian
-%\begin{equation}
-%    L=T-V
-%\end{equation}
-%where $T$ is the kinetic energy, and $V$ is the potential energy. From the Euler-Lagrange equation we have:
-%\begin{equation}
-%    \frac{d}{dt} \frac{\partial L}{\partial \dot{\mathbf{p}}} + \frac{\partial L}{\partial {\mathbf{p}}} = 0
-%\end{equation}
-%\begin{equation}
-%    V =  \int_\Omega \psi(\mathbf{F(X)} dX
-%\end{equation}
-%then discretizing this over $m$ quadrature points, we have
-%\begin{equation}
-%    V =  \sum_i^m \psi(\mathbf{F(X_i)} A_i
-%\end{equation}
-%\begin{equation}
-%    \frac{\partial L}{\partial \mathbf{q}} = \sum_i^m \frac{\partial V_i}{\partial \mathbf{q}} = \sum_i^m \frac{\partial \psi(\mathbf{F_i)}}{\partial \mathbf{F}}\frac{\partial \mathbf{F_i}}{\partial \mathbf{q}}A_i
-%\end{equation}
-%Next we have the equation of the kinetic energy
-%\begin{equation}
-%    T = \frac{1}{2}\int_{ \Omega} \rho \dot{\mathbf{x}}^T\dot{\mathbf{x}} dV
-%\end{equation}
-%which we rewrite in terms of our control points configuration
-%\begin{equation}
-%    T = \frac{1}{2} \int_{\partial \Omega} \dot{\mathbf{p}}^T \mathbf{J}(u,v)^T \mathbf{M} \mathbf{J}(u,v) \dot{\mathbf{p}} du dv
-%\end{equation}
-%\myworries{change notation to not use M for the monomial basis}
-%where
-%\begin{equation}
-%    \mathbf{M} = \int_\Omega \rho \mathbf{J_{\hat{\Pi}}(X)}^T \mathbf{J_{\hat{\Pi}}(X)}
-%\end{equation}
-%From the potential energy (which uses dF/dq) and this potential energy, we discretize in time and right now I'm using linearly implicit backward euler
-%
-%\subsection{Stability term}
-%\begin{figure}[h]
-%  \centering
-%  \includegraphics[width=\linewidth]{figures/error_energy}
-%  \caption{A figure showing the spring energy error for each sampled point on the NURBS curve. I should change the colors of the lines to show higher energies where the error is higher.}
-%  \Description{Showing an example error of a NURBS curve.}
-%\end{figure}
-%
-%\subsection{Projection Operator Weighting}
-%\begin{figure}[h]
-%  \centering
-%  \includegraphics[width=\linewidth]{figures/distance_weighting}
-%  \caption{A 2D example showing how $\theta$ is computed for $\Pi_B$.
-$d_1$ shows the distance to the closest point on $\Pi_B$ to $\mathbf{x}$ and $d_2$ shows the distance to the surface hit by the ray in the opposite direction.}
-%  \Description{Showing a 2D NURBS curve with a quadrature point in the center.
-A ray is cast to the nearest point on a shape, showing the distance between the quadrature point and the nearest point as well as the distance to the "back surface".}
-%\end{figure}
-%
-%
-%\subsection{Integrating the volume}
-%\begin{figure}[h]
-%  \centering
-%  \includegraphics[width=\linewidth]{figures/raycasting_quadrature}
-%  \caption{An example raycasting quadrature on a 2D NURBS curve.}
-%  \Description{Shows a 2D NURBs curve with horizontal rays intersecting the curve. }
-%\end{figure}
+
+\subsection{Variational Mechanics}
+We have now provided all the necessary ingredient to produce equations of motion for deformable bodies, which we derive via Euler-Lagrange equation. We begin with the Lagrangian
+\begin{equation}
+    L=T-V
+\end{equation}
+where $T$ is the kinetic energy, and $V$ is the potential energy. From the Euler-Lagrange equation we have:
+\begin{equation}
+    \frac{d}{dt} \frac{\partial L}{\partial \dot{\mathbf{p}}} + \frac{\partial L}{\partial {\mathbf{p}}} = 0
+\end{equation}
+\begin{equation}
+    V =  \int_\Omega \psi(\mathbf{F(X)} dX
+\end{equation}
+then discretizing this over $m$ quadrature points, we have
+\begin{equation}
+    V =  \sum_i^m \psi(\mathbf{F(X_i)} A_i
+\end{equation}
+\begin{equation}
+    \frac{\partial L}{\partial \mathbf{q}} = \sum_i^m \frac{\partial V_i}{\partial \mathbf{q}} = \sum_i^m \frac{\partial \psi(\mathbf{F_i)}}{\partial \mathbf{F}}\frac{\partial \mathbf{F_i}}{\partial \mathbf{q}}A_i
+\end{equation}
+Next we have the equation of the kinetic energy
+\begin{equation}
+    T = \frac{1}{2}\int_{ \Omega} \rho \dot{\mathbf{x}}^T\dot{\mathbf{x}} dV
+\end{equation}
+which we rewrite in terms of our control points configuration
+\begin{equation}
+    T = \frac{1}{2} \int_{\partial \Omega} \dot{\mathbf{p}}^T \mathbf{J}(u,v)^T \mathbf{M} \mathbf{J}(u,v) \dot{\mathbf{p}} du dv
+\end{equation}
+\myworries{change notation to not use M for the monomial basis}
+where
+\begin{equation}
+    \mathbf{M} = \int_\Omega \rho \mathbf{J_{\hat{\Pi}}(X)}^T \mathbf{J_{\hat{\Pi}}(X)}
+\end{equation}
+From the potential energy (which uses dF/dq) and this potential energy, we discretize in time and right now I'm using linearly implicit backward euler

### 2021-01-17 paper/sections/methods.tex
Most of methods written out, except center of mass section and time integration part

@@
 The second line is a result of the fact that all terms except for $\mathbf{Y(X)}$ are spatially constant, allowing us to move these outside the volume integral, which enables fast assembly of the mass matrix, $\mathbf{M}$.
-\subsection{Error Potential}
+\subsection{Error Energy}
+Over the course of the simulation we may see the true boundary values deviate from their positions estimated by the polynomial.
+This is most notable in cases where the degree of the polynomial on a NURBS is higher than the degree of the B-spline surface.
+To address this we augment our kinetic energy with a term to account for this error.
+The error for a boundary point $\mathbf{x}_i^*$ that is known to lie on the limit surface of the NURBS can be written as
+\begin{equation}
+\begin{split}
+\epsilon_i & = ||\mathbf{x}(\mathbf{X}_i) - \mathbf{x}_i^*|| \\
+		   & = ||(\mathbf{Y}(\mathbf{X}_i)\mathbf{L} - \mathbf{S_i})\mathbf{\hat{b}}||
+\end{split}
+\text{,}
+\end{equation}
+where $\mathbf{S_i}$ extracts $\mathbf{x}_i^*$ from $\mathbf{\hat{b}}$.
+We modify this error further to be expressed in terms of the control points.
+We note that $\mathbf{x}^*(u,v)$ yields the position on the NURBS surface whereas $\mathbf{x(X)}$ is the deformed value estimated by the polynomial fitting.
+Furthermore $\mathbf{X}$ may be written as a function of the parametric coordinates $\mathbf{X}(u,v)$ as its position on the surface is produced using the initial set of control points representing the undeformed model.
+This leads to an energy potential over the NURBS surface in parametric space
+\begin{equation}
+T_E = \iint (\mathbf{x}(\mathbf{X}(u,v)) - \mathbf{x}^*(u,v))^T(\mathbf{x}(\mathbf{X}(u,v)) - \mathbf{x}^*(u,v)) du dv
+\text{.}
+\end{equation}
+The discrete form of this written in terms of the control points may be written as
+\begin{equation}
+\begin{split}
+T_E & = \sum_i^N  \dot{\mathbf{q}}^T\mathbf{J}^T(\mathbf{Y}(\mathbf{X}_i)\mathbf{L} - \mathbf{S_i})^T(\mathbf{Y}(\mathbf{X}_i)\mathbf{L} - \mathbf{S_i})\mathbf{J}\dot{\mathbf{q}} \\
+  & = \dot{\mathbf{q}}^T\mathbf{J}^T \left[ \sum_i^N  (\mathbf{Y}(\mathbf{X}_i)\mathbf{L} - \mathbf{S_i})^T(\mathbf{Y}(\mathbf{X}_i)\mathbf{L} - \mathbf{S_i}) \right] \mathbf{J}\dot{\mathbf{q}} \\
+  & = \dot{\mathbf{q}}^T \mathbf{M}_E \dot{\mathbf{q}}
+\end{split}
+\text{,}
+\end{equation}
+where $N$ is the total number of boundary points and $\mathbf{M}_E$ is our \textit{error mass matrix}.
 \subsection{Generalized Forces}
-\subsection{Equations of Motion}
-\subsection{Time Integration}
-%This is most notable in cases where the degree of the polynomial on a NURBS is higher than the degree of the B-spline surface.
+We write our potential energy as 
+\begin{equation}
+    V = \int_\Omega \psi(\mathbf{F(X)}) d\Omega
+\text{.}
+\end{equation}
+To evaluate this energy (as well as its derivatives), we require a definition for the deformation gradient, $\mathbf{F(X)}$:
+\begin{equation}
+        \mathbf{F(X)} = \frac{\partial \mathbf{x}}{\partial \mathbf{X}}
+               = \frac{\partial \mathbf{Y(X)}}{\partial \mathbf{X}}\mathbf{L}\hat{\mathbf{b}}
+\text{.}
+\end{equation}
+In the expression for the deformed position $\mathbf{x(X)}$ we see that the dependence on $\mathbf{X}$ only exists in $\mathbf{Y(X)}$, making $\mathbf{L}$ and $\hat{\mathbf{b}}$ constant terms when compute $\mathbf{F}$.
+Full details on how to compute $\mathbf{F}$ and its derivative $\frac{\partial \mathbf{F}}{\partial \mathbf{q}}$ may be found in the appendix.
+(\myworries{I just wrote this because I'm too tired to right the d etails right now :p}.
+Now that we have an expression for the deformation gradient, we discretize our potential energy as follows
+\begin{equation}
+    V =  \sum_i^m \psi(\mathbf{F(X}_i) A_i
+\text{,}
+\end{equation}
+where $m$ is the number of integration points and $A_i$ is the volume for the $i$-th integration point.
 
-\subsection{Deformation Gradient}
+Like in the case of the kinetic energy, we also introduce an error force potential, which takes the simple form
 \begin{equation}
-        \mathbf{F} = \frac{\partial \mathbf{x}}{\partial \mathbf{X}}
-               = \frac{\partial \hat{\mathbf{\Pi}}\mathbf{M(X)}}{\partial \mathbf{X}}
-               = \hat{\mathbf{\Pi}} \frac{\partial \mathbf{M(X)}}{\partial \mathbf{X}}
+V_E = \mathbf{q}^T \mathbf{M}_E \mathbf{q}
 \end{equation}
 
-In computing the forces and the stiffness matrix, we require the gradient of the deformation gradient with respect to the configuration,  $\frac{\partial \mathbf{F}}{\partial \mathbf{q}}$.
-Usually the form of this gradient is simple.
-For example, in linear tetrahedral FEM each quadrature point's deformation map only involves the four surrounding nodal values, simplifying this gradient.
-In our case, the presence of the projection operator $\hat{\mathbf{\Pi}}$ means that each quadrature point is computed using the nodal values of the entire model.
-
-\begin{equation}
-    \frac{\partial \mathbf{F}}{\partial \mathbf{q}} = \frac{\partial \mathbf{P}}{\partial q}\mathbf{M(q_0)}^\dagger  \frac{\partial \mathbf{M(X)}}{\partial \mathbf{X}} 
-\end{equation}
-
-Naively handled, this gives us a $\frac{\partial \mathbf{F}}{\partial \mathbf{q}}$ of size $3 x |q|$.
-However, many of the columns of have values all near zero, allowing us to prune many columns for each quadrature point.
-Currently, I'm computing the infinity norm of each column and prune all except some fixed number of columns.
-\\
-
-\subsection{Variational Mechanics}
+\subsection{Time Integration}
 We have now provided all the necessary ingredient to produce equations of motion for deformable bodies, which we derive via Euler-Lagrange equation. We begin with the Lagrangian
 \begin{equation}
-    L=T-V
+    L=(T+T_E)-(V + V_E)
 \end{equation}
-where $T$ is the kinetic energy, and $V$ is the potential energy. From the Euler-Lagrange equation we have:
+Then from the Euler-Lagrange equation we have:
 \begin{equation}
@@
 \end{equation}
-\begin{equation}
-    V =  \int_\Omega \psi(\mathbf{F(X)} dX
-\end{equation}
-then discretizing this over $m$ quadrature points, we have
-\begin{equation}
-    V =  \sum_i^m \psi(\mathbf{F(X_i)} A_i
-\end{equation}
-\begin{equation}
-    \frac{\partial L}{\partial \mathbf{q}} = \sum_i^m \frac{\partial V_i}{\partial \mathbf{q}} = \sum_i^m \frac{\partial \psi(\mathbf{F_i)}}{\partial \mathbf{F}}\frac{\partial \mathbf{F_i}}{\partial \mathbf{q}}A_i
-\end{equation}
-Next we have the equation of the kinetic energy
-\begin{equation}
-    T = \frac{1}{2}\int_{ \Omega} \rho \dot{\mathbf{x}}^T\dot{\mathbf{x}} dV
-\end{equation}
-which we rewrite in terms of our control points configuration
-\begin{equation}
-    T = \frac{1}{2} \int_{\partial \Omega} \dot{\mathbf{p}}^T \mathbf{J}(u,v)^T \mathbf{M} \mathbf{J}(u,v) \dot{\mathbf{p}} du dv
-\end{equation}
-\myworries{change notation to not use M for the monomial basis}
-where
-\begin{equation}
-    \mathbf{M} = \int_\Omega \rho \mathbf{J_{\hat{\Pi}}(X)}^T \mathbf{J_{\hat{\Pi}}(X)}
-\end{equation}
-From the potential energy (which uses dF/dq) and this potential energy, we discretize in time and right now I'm using linearly implicit backward euler
+\myworries{todo: not totally sure what to write here since we're experimenting with both linearly implicit backward euler as well as Newton's method soon}
+
+%\begin{equation}
+%    \frac{\partial L}{\partial \mathbf{q}} = \sum_i^m \frac{\partial V_i}{\partial \mathbf{q}} = \sum_i^m \frac{\partial \psi(\mathbf{F_i)}}{\partial \mathbf{F}}\frac{\partial \mathbf{F_i}}{\partial \mathbf{q}}A_i
+%\end{equation}
+%\

### 2021-01-18 paper/sections/results.tex
figures

@@
 %
+\section{Results}
+
+\dave{we should show the nurbs model in rhino/fusion for every example, maybe an exploded view as well}
+
+%% Results that relate to preprocessing
+% weights are good
 \begin{figure}
@@
 \end{figure}
-%
+
+%kinematic model is expressive
+\begin{figure}
+  \includegraphics[width=\columnwidth]{example-image-a}
+  \caption{Take an intesting geometry (maybe that geometry processing cacture), deform it, do our shape matching and then reconstruct the deformed pose.
+Do this for a bunch of poses (show's kinematic model is good).}
+  \label{fig:deform}
+\end{figure}
+
+%% Results that relate to simulation output
+%correctness
+\begin{figure}
+  \includegraphics[width=\columnwidth]{example-image-a}
+  \caption{2D examples (left) rigidly transform boundary and show that solving the static problem gives us rigid motion on the interior (right) isotropically scale boundary and show solving static problem gives us constant strain on the interior}
+  \label{fig:patchtest}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{example-image-a}
+  \caption{(left) high res FEM cantileavered bar (right) several simulations using more and more surface patches plus center of mass. Need to show beam is just  unattached patches}
+  \label{fig:convergence}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{example-image-a}
+  \caption{X shape and show arms of X move independently}
+  \label{fig:independence}
+\end{figure}
+
+% relaxed modeling  constraints
+\begin{figure}
+  \includegraphics[width=\columnwidth]{example-image-a}
+  \caption{Coffee mug with overlapping handle (top) row of models with increasing overlap (middle) single weight image showing behaviour in the overlapping region, (bottom) simulation result for each one}
+  \label{fig:badmodels}
+\end{figure}
+
+%different material parameters / material models 
+\begin{figure}
+  \includegraphics[width=\columnwidth]{example-image-a}
+  \caption{Big image simulating the rocket with all different combinations of things}
+  \label{fig:materials}
+\end{figure}
+
+%different material parameters / material models 
+\begin{figure}
+  \includegraphics[width=\columnwidth]{example-image-a}
+  \caption{Big image simulating the rocket with all different combinations of things}
+  \label{fig:materials}
+\end{figure}
+
+%multiple materials
+\begin{figure}
+  \includegraphics[width=\textwidth]{example-image-a}
+  \caption{Avocado to show heterogenous materials}
+  \label{fig:avocado}
+\end{figure}
+
+%show stoppers
+\begin{figure*}[htp]
+  \includegraphics[width=\textwidth,height=3in]{example-image-a}
+  \caption{sequence of frames from staypuft simulation}
+  \label{fig:staypuft}
+\end{figure*}
+
+\begin{figure*}[htp]
+  \includegraphics[width=\textwidth,height=3in]{example-image-a}
+  \caption{sequence of frames from F1 car simulation}
+  \label{fig:f1}
+\end{figure*}

### 2021-01-19 examples/linear_mug_implicit.m
most things except linear_trimmed_mug seem to work

@@ -21,7 +21,7 @@ options.lambda = lambda;
 options.mu = mu;
 options.sample_interior = 0; % only sample on boundary
 options.distance_cutoff = 1;
-vem_simulate_nurbs_fmincon_param(part, options);  %%%%% verify hierarchical works!!!!!
+vem_simulate_nurbs_newtons(part, options);  %%%%% verify hierarchical works!!!!!
                                                         
  
 end

### 2021-01-19 examples/linear_rocket.m

@@ -19,9 +19,9 @@ options.pin_function = @pin_function;
 options.lambda = lambda;
 options.mu = mu;
 options.distance_cutoff = 10;
-options.k_stability = YM*1e5;
+options.k_stability = YM*1e6;
 
-vem_simulate_nurbs_implicit(part, options);
-% vem_simulate_nurbs(part, options);
+% vem_simulate_nurbs_implicit(part, options);
+vem_simulate_nurbs(part, options);
 
 end

### 2021-01-19 examples/linear_rounded_cube.m

@@ -20,7 +20,7 @@ options.pin_function = @pin_function;
 options.lambda = lambda;
 options.mu = mu;
 
-vem_simulate_nurbs_fmincon_param(part, options);
-% vem_simulate_nurbs_implicit(part, options);
+vem_simulate_nurbs_newtons(part, options);
+% vem_simulate_nurbs(part, options);
 
 end

### 2021-01-19 examples/linear_trimmed_cylinder.m

@@ -21,6 +21,6 @@ options.lambda = lambda;
 options.mu = mu;
 options.pin_function = @pin_function;
 
-vem_simulate_nurbs(part, options);
+vem_simulate_nurbs_newtons(part, options);
   
 end

### 2021-01-19 matlab/nurbs_utils/nurbs_assemble_coords.m

@@ -46,8 +46,6 @@ function [J,hires_J,q,E,x] = nurbs_assemble_coords(parts)
     hires_J = sparse(hires_J);
     E=cell(numel(parts),1);
     
-    E_size = 0;
-    
     for i=1:numel(parts)
         idx1=size(x,2)+1;
         x = [x parts{i}.x0];

### 2021-01-19 paper/sections/methods_ty.tex
backup Tys method section

--- 2021-01-17 paper/sections/methods.tex
@@
 \section{Methods}
+\dave{We should probably put some pseudo code for the preprocessing and the runtime algorithms right up fron, along with a quick textual walk through of the method.}
 Given a volumetric model composed of one or more NURBS surfaces, the surface of each surface is reconstructed using a set of control points $\textbf{p}$ and their associated weights $w$.
@@
 For example if $\mathbf{X} = (X,Y,Z)$, $\mathbf{\bar{X}} = (\bar X, \bar Y, \bar Z)$, and $\mathbf{q}=\mathbf{X}-\mathbf{\bar{X}}$ with a quadratic polynomial we would have
+
+\begin{figure}
+    \includegraphics[width=\columnwidth]{example-image-a}
+    \caption{Figure showing shape matching to NURBS}
+    \label{fig:shapematching}
+\end{figure}
+
 \begin{equation}
@@
 
+\begin{figure}
+    \includegraphics[width=\columnwidth]{example-image-a}
+    \caption{Figure showing how weight calculation works (left) ray casting plus result (right) after correcting for partition of unity}
+    \label{fig:weightcompute}
+\end{figure}
+
 Now that we have a set of distances weights $\mathbf{theta}$, we seek to form a final set of \textit{blending weights} $\mathbf{w}(\mathbf{X}) =  \left( w_1, \dots, w_i, \dots, w_n \right)$ to blend the polynomials of the $n$ shape elements.
@@
 We note that this solution is not perfect and making this intersection step more robust serves as future work.
+
+\begin{figure}
+    \includegraphics[width=\columnwidth]{example-image-a}
+    \caption{Figure showing how raycasting quadrature works (top), convergence plot for volume integration (bottom}
+    \label{fig:raycasting}
+\end{figure}
 

### 2021-01-19 paper/sections/methods.tex
methods overview (commented out Tys methods)

--- 2021-01-17 paper/sections/methods.tex
@@
 
+\section{Overview}
+The input to the Shape Matching Element Method (SEM)  is a NURBS \emph{boundary representation} of a volumetric object, along with a set of physical parameters 
+(density, constitutive model, model parameters etc). The output is an elastodynamic simulation. 
+SEM consists of preprocessing and runtime simulation phases~\refalg{sem} and at no point do we need to generate a volumetric mesh of the input. 
+In the \emph{preprocessing} stage we use raycasting to find quadrature points and weights for volumetric integration, as well as to compute part blending parameters. 
+Finally we construct local shape matching operators and construct the mass matrix for our problem.
+
+At runtime we use standard time integration schemes to time step the system (we show linearly implicit, fully implicit, runge-kutta four). 
+These require us to evaluate the energy, gradient and Hessian (implicit only) which we do at the previously computed quadrature points. 
+For collisions we use penalty forces and apply them as external forces during integration.
+
+\begin{algorithm}[h]
+    \label{alg:sem}
+    \SetAlgoLined
+    \KwResult{Write here the result }
+     initialization\;
+     \While{While condition}{
+      instructions\;
+      \eIf{condition}{
+       instructions1\;
+       instructions2\;
+       }{
+       instructions3\;
+      }
+     }
+     \caption{SEM Algorithm}
+    \end{algorithm}
 \section{Methods}
-Given a volumetric model composed of one or more NURBS surfaces, the surface of each surface is reconstructed using a set of control points $\textbf{p}$ and their associated weights $w$.
+
+% \section{Methods}
+% \dave{We should probably put some pseudo code for the preprocessing and the runtime algorithms right up fron, along with a quick textual walk through of the method.}
+% Given a volumetric model composed of one or more NURBS surfaces, the surface of each surface is reconstructed using a set of control points $\textbf{p}$ and their associated weights $w$.
 Following D-NURBS \cite{10.1145/176579.176580}, we take our degrees of freedom (DOF) to be control points of the surfaces and produce displacements directly on these control points at each simulation step.
@@
 
-To evaluate the volumetric integrals in solid mechanics models, we require material points $\mathbf{X} \in \mathbb{R}^{3m}$ in the undeformed space that serve as integration points.
+% To evaluate the volumetric integrals in solid mechanics models, we require material points $\mathbf{X} \in \mathbb{R}^{3m}$ in the undeformed space that serve as integration points.
 Our shape matching element method enables us to evaluate the deformed positions of the material points strictly using positions defined on the boundary.
@@
 
-In the following sections, we will first review the simplified D-NURBS model used to form our generalized coordinates.
+% In the following sections, we will first review the simplified D-NURBS model used to form our generalized coordinates.
 We then present our shape matching algorithm for fitting polynomials to surfaces ....
@@
 
-\section{D-NURBS Model}
-\myworries{TODO: Most of these matrices should be paranthesis matrices}
-We provide a brief overview of the D-NURBS formulation presented in \cite{10.1145/176579.176580}. Our models in our method are composed of a collection of NURBS surfaces.
+% \section{D-NURBS Model}
+% \myworries{TODO: Most of these matrices should be paranthesis matrices}
+% We provide a brief overview of the D-NURBS formulation presented in \cite{10.1145/176579.176580}. Our models in our method are composed of a collection of NURBS surfaces.
 
-\subsection{D-NURBS Formulation}
-A point on a NURBS surface is typically expressed in the following form
-\begin{equation}
-\label{eqn:nurbs_srf}
-    \mathbf{x}(u,v) = \frac{\sum_{i=1}^{n}\sum_{j=1}^{m}  \mathbf{p}_{i,j} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
-    {\sum_{i=1}^{n}\sum_{j=1}^{m} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
-    \text{,}
-\end{equation}
-where where have a total of $nm$ control points $\mathbf{p}$ and weights $w$.
+% \subsection{D-NURBS Formulation}
+% A point on a NURBS surface is typically expressed in the following form
+% \begin{equation}
+% \label{eqn:nurbs_srf}
+%     \mathbf{x}(u,v) = \frac{\sum_{i=1}^{n}\sum_{j=1}^{m}  \mathbf{p}_{i,j} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
+%     {\sum_{i=1}^{n}\sum_{j=1}^{m} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
+%     \text{,}
+% \end{equation}
+% where where have a total of $nm$ control points $\mathbf{p}$ and weights $w$.
 Using the B-spline basis functions and the weighted control points, this formula gives a map from our parametric coordinates $u,v$ to their corresponding position on the surface in $\mathbb{R}^{3m}$.
@@
 
-We would like our generalized coordinates to be the control points of the model so that the dynamics operates directly on the original representation.
+% We would like our generalized coordinates to be the control points of the model so that the dynamics operates directly on the original representation.
 Therefore, we would like to rewrite equation \ref{eqn:nurbs_srf} in the form
-\begin{equation}
-    \mathbf{x}(u,v) = \mathbf{J}(u,v)\mathbf{q}
-    \text{,}
-\end{equation}
-where $\mathbf{q} = \left[ \mathbf{p}_{1,1}^T, \dots, \mathbf{p}_{i,j}^T, \dots, \mathbf{p}_{n,m}^T \right]^T \in \mathbb{R}^{3nm}$ is the set of $nm$ control points \myworries{in the overview I just write $n$ control points, but here i'm writing $nm$} for a single NURBS surface arranged as a column vector.
+% \begin{equation}
+%     \mathbf{x}(u,v) = \mathbf{J}(u,v)\mathbf{q}
+%     \text{,}
+% \end{equation}
+% where $\mathbf{q} = \left[ \mathbf{p}_{1,1}^T, \dots, \mathbf{p}_{i,j}^T, \dots, \mathbf{p}_{n,m}^T \right]^T \in \mathbb{R}^{3nm}$ is the set of $nm$ control points \myworries{in the overview I just write $n$ control points, but here i'm writing $nm$} for a single NURBS surface arranged as a column vector.
 $\mathbf{J} \in  \mathbb{R}^{3 x 3nm}$ is the NURBS jacobian that maps $(u,v)$ coordinates to world positions in $\mathbb{R}^{3}$.
 
-The NURBS jacobian, $\mathbf{J}$, is a matrix composed of horizontally concatenated $3x3$ blocks where the $(i,j)$-th block is $\frac{\partial \mathbf{x}}{\partial \mathbf{p}_{i,j}}$ for the $(i,j)$-th control point.
+% The NURBS jacobian, $\mathbf{J}$, is a matrix composed of horizontally concatenated $3x3$ blocks where the $(i,j)$-th block is $\frac{\partial \mathbf{x}}{\partial \mathbf{p}_{i,j}}$ for the $(i,j)$-th control point.
 $\frac{\partial \mathbf{x}}{\partial \mathbf{p}_{i,j}}$ is a diagonal $3x3$ matrix where each diagonal entry $N_{i,j}(\mathbf{u})$ takes the form
 
-\begin{equation}
-\label{eqn:jacobian_diagonal}
-    N_{i,j}(u,v)
-    = \frac{w_{i,j} B_{i,k}(u)B_{j,l}(v)}{\sum_{i=1}^{n}\sum_{j=1}^{m} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
-    \text{.}
-\end{equation}
-Thus the NURBS jacobian for a single $(u,v)$ pair is written as
-\begin{equation}
-\label{eqn:uv_jacobian}
-    \mathbf{J}(u,v) =
-    \left[ \frac{\partial \mathbf{x}}{\partial \mathbf{p}_{1,1}} \dots
-           \frac{\partial \mathbf{x}}{\partial \mathbf{p}_{i,j}} \dots 
-           \frac{\partial \mathbf{x}}{\partial \mathbf{p}_{n,m}}
-    \right]
-    \text{.}
-\end{equation}
+% \begin{equation}
+% \label{eqn:jacobian_diagonal}
+%     N_{i,j}(u,v)
+%     = \frac{w_{i,j} B_{i,k}(u)B_{j,l}(v)}{\sum_{i=1}^{n}\sum_{j=1}^{m} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
+%     \text{.}
+% \end{equation}
+% Thus the NURBS jacobian for a single $(u,v)$ pair is written as
+% \begin{equation}
+% \label{eqn:uv_jacobian}
+%     \mathbf{J}(u,v) =
+%     \left[ \frac{\partial \mathbf{x}}{\partial \mathbf{p}_{1,1}} \dots
+%            \frac{\partial \mathbf{x}}{\partial \mathbf{p}_{i,j}} \dots 
+%            \frac{\partial \mathbf{x}}{\partial \mathbf{p}_{n,m}}
+%     \right]
+%     \text{.}
+% \end{equation}
 
-This formulation is a simplification of the full D-NURBS described in \cite{10.1145/176579.176580} the weights of the NURBS are fixed throughout the simulation.
+% This formulation is a simplification of the full D-NURBS described in \cite{10.1145/176579.176580} the weights of the NURBS are fixed throughout the simulation.
 Permitting the dynamics to modify weights would likely improve the expressiveness of the kinematics, but we find fixing the weights affords sufficiently expressive results and improves performance.
@@
 
-\subsection{Generalized Coordinates for a NURBS model}
-The above formulation describes developing the map from a single $(u,v)$ coordinate to its corresponding position on the surface.
+% \subsection{Generalized Coordinates for a NURBS model}
+% The above formulation describes developing the map from a single $(u,v)$ coordinate to its corresponding position on the surface.
 In our simulation we represent the entire boundary of the NURBS model, so we require samples along each surface.
@@
 
-Let's say for a NURBS surface on the model the DOF column vector is $\mathbf{q}_i$, which is the subset of control points in the generalized coordinates for the $i$-th surface.
+% Let's say for a NURBS surface on the model the DOF column vector is $\mathbf{q}_i$, which is the subset of control points in the generalized coordinates for the $i$-th surface.
 For a given surface with $m$ pairs of $\mathbf{u}=(u,v)$ coordinates sampled in parametric space, we write the jacobian for the $i$-th surface as \myworries{(should I use something like $\hat{\mathbf{J}}$ instead?)}
-\begin{equation}
-\label{eqn:surface_jacobian}
-    \mathbf{J}_i =
-    \left[ \mathbf{J}(\mathbf{u})_1^T \dots \mathbf{J}(\mathbf{u})_m^T \right]^T
-    \text{.}
-\end{equation}
-Then if $\mathbf{x}_i$ is the set of $m$ world space positions for the $i$-th surface, we may write $\mathbf{x}_i = \mathbf{J}_i \mathbf{q}_i$.
+% \begin{equation}
+% \label{eqn:surface_jacobian}
+%     \mathbf{J}_i =
+%     \left[ \mathbf{J}(\mathbf{u})_1^T \dots \mathbf{J}(\mathbf{u})_m^T \right]^T
+%     \text{.}
+% \end{equation}
+% Then if $\mathbf{x}_i$ is the set of $m$ world space positions for the $i$-th surface, we may write $\mathbf{x}_i = \mathbf{J}_i \mathbf{q}_i$.
 Finally, given the full set of generalized coordinates $\mathbf{q} = \left[ \mathbf{q}_i^T, \dots, \mathbf{q}_n^T \right]^T$ with $n$ surfaces, we may write jacobian $\mathbf{J}$ for the entire model as a block diagonal matrix:
-\begin{equation}
-\mathbf{J} = \left[ \begin{array}{ccccc}
-\mathbf{J}_1 &  &  &  &  \\
- & \ddots &  &  &  \\
- &  & \mathbf{J}_i & &  \\
- &  &  & \ddots &  \\
- &  &  &  & \mathbf{J}_n \\
-\end{array} \right]
-\text{.}
-\end{equation}
-With this equation, the full set of world positions of the NURBS model may be written as $\mathbf{x} = \mathbf{J}\mathbf{q} \in \mathbb{R}^{3N}$ where $N$ is the total number of $(u,v)$ coordinates.
+% \begin{equation}
+% \mathbf{J} = \left[ \begin{array}{ccccc}
+% \mathbf{J}_1 &  &  &  &  \\
+%  & \ddots &  &  &  \\
+%  &  & \mathbf{J}_i & &  \\
+%  &  &  & \ddots &  \\
+%  &  &  &  & \mathbf{J}_n \\
+% \end{array} \right]
+% \text{.}
+% \end{equation}
+% With this equation, the full set of world positions of the NURBS model may be written as $\mathbf{x} = \mathbf{J}\mathbf{q} \in \mathbb{R}^{3N}$ where $N$ is the total number of $(u,v)$ coordinates.
 
-\section{Shape Matching}
-The key feature of our method is its unique deformation map $\phi(\mathbf{X}) = \mathbf{x}$ mapping some arbitrary undeformed material point $\mathbf{X}$ to its corresponding deformed position in space $\mathbf{x}$.
+% \section{Shape Matching}
+% The key feature of our method is its unique deformation map $\phi(\mathbf{X}) = \mathbf{x}$ mapping some arbitrary undeformed material point $\mathbf{X}$ to its corresponding deformed position in space $\mathbf{x}$.
 We develop this similarly to the VEM approach in that we say the function we intend to reproduce may be some highly nonlinear one, but it may be approximated accurately by a polynomial.
@@
 
-\subsection{The Shape Matching Element}
-In typical FEM in graphics, we make use of tetrahedral or triangular elements with a piecewise-linear function space.
+% \subsection{The Shape Matching Element}
+% In typical FEM in graphics, we make use of tetrahedral or triangular elements with a piecewise-linear function space.
 In such a case our functions may be evaluated as a weighted combination of the nodal values of an element: $f(\mathbf x)=\sum_{i=0}^n f_i \phi_i(\mathbf{x})$ where $n$ is the number of nodal values and $\phi_i$ is the i-th barycentric coordinates.
@@
 
-In our algorithm, a single shape matching element is represented by a NURBS surface, a center of mass, and polynomial describing its current deformation.
+% In our algorithm, a single shape matching element is represented by a NURBS surface, a center of mass, and polynomial describing its current deformation.
 We construct these polynomials via a shape matching strategy.
@@
 For example if $\mathbf{X} = (X,Y,Z)$, $\mathbf{\bar{X}} = (\bar X, \bar Y, \bar Z)$, and $\mathbf{q}=\mathbf{X}-\mathbf{\bar{X}}$ with a quadratic polynomial we would have
-\begin{equation}
-m(\mathbf{q)} = (
-					q_1, q_2, q_3,
-                   	q_1^2, q_2^2, q_3^3,
-                   	q_1 q_2, q_2 q_3
-                ) \in \mathbb{R}^k
-                \text{,}
-\end{equation}
-where $k$ is the size of the monomial basis. Next we arrange this basis into a matrix of the form
-\begin{equation}
-\mathbf{M}(\mathbf{q}) = \left[ \begin{array}{ccc}
-\mathbf{m}(\mathbf{q}) &  &  \\
- & \mathbf{m}(\mathbf{q})  &  \\
- &  & \mathbf{m}(\mathbf{q})  \\
-\end{array} \right] \in \mathbb{R}^{3x3k}
-\text{.}
-\end{equation}
-With this definition of the monomial basis, we may reconstruct the deformed position $\mathbf{x}$ for some material point $\mathbf{X}$ with
-\begin{equation}
-\mathbf{x}(\mathbf{X}) = \left[\mathbf{M}(\mathbf{X-\bar{X}}) \quad \mathbf{I} \right]
-\left[ \begin{array}{c}
-\mathbf{c} \\
-\mathbf{t} \\
-\end{array} \right]
-\text{,}
-\end{equation}
-where $\mathbf{c} \in \mathbb{R}^{3k}$ is the set of polynomial coefficients for the non-constant terms, $\mathbf{t} \in \mathbb{R}^3$ is the constant term coefficients, and $\mathbf{I} \in \mathbb{R}^{3x3}$ is an identity matrix.
+
+% \begin{figure}
+%     \includegraphics[width=\columnwidth]{example-image-a}
+%     \caption{Figure showing shape matching to NURBS}
+%     \label{fig:shapematching}
+% \end{figure}
+
+% \begin{equation}
+% m(\mathbf{q)} = (
+% 					q_1, q_2, q_3,
+%                    	q_1^2, q_2^2, q_3^3,
+%                    	q_1 q_2, q_2 q_3
+%                 ) \in \mathbb{R}^k
+%                 \text{,}
+% \end{equation}
+% where $k$ is the size of the monomial basis. Next we arrange this basis into a matrix of the form
+% \begin{equation}
+% \mathbf{M}(\mathbf{q}) = \left[ \begin{array}{ccc}
+% \mathbf{m}(\mathbf{q}) &  &  \\
+%  & \mathbf{m}(\mathbf{q})  &  \\
+%  &  & \mathbf{m}(\mathbf{q})  \\
+% \end{array} \right] \in \mathbb{R}^{3x3k}
+% \text{.}
+% \end{equation}
+% With this definition of the monomial basis, we may reconstruct the deformed position $\mathbf{x}$ for some material point $\mathbf{X}$ with
+% \begin{equation}
+% \mathbf{x}(\mathbf{X}) = \left[\mathbf{M}(\mathbf{X-\bar{X}}) \quad \mathbf{I} \right]
+% \left[ \begin{array}{c}
+% \mathbf{c} \\
+% \mathbf{t} \\
+% \end{array} \right]
+% \text{,}
+% \end{equation}
+% where $\mathbf{c} \in \mathbb{R}^{3k}$ is the set of polynomial coefficients for the non-constant terms, $\mathbf{t} \in \mathbb{R}^3$ is the constant term coefficients, and $\mathbf{I} \in \mathbb{R}^{3x3}$ is an identity matrix.
 The reason for separating out the constant term becomes more evident when we examine case where we have multiple elements.
 
-The remaining step in defining a single element is to describe how to solve for the polynomial coefficients.
+% The remaining step in defining a single element is to describe how to solve for the polynomial coefficients.
 For a single element, we have $n$ material points $(\mathbf{X}_1 \dots \mathbf{X}_n)$ and their corresponding deformed positions $(\mathbf{x}^*_1 \dots \mathbf{x}^*_n)$.
@@
 We minimize this error in a least squares sense where we first form the system
-\begin{equation}
-\left[\begin{array}{c}
-\mathbf{M}(\mathbf{X_1-\bar{X}}) \quad \mathbf{I} \\
-\vdots \\
-\mathbf{M}(\mathbf{X_i-\bar{X}}) \quad \mathbf{I} \\
-\vdots \\
-\mathbf{M}(\mathbf{X_n-\bar{X}}) \quad \mathbf{I} \\
-\end{array} \right]
-\left[ \begin{array}{c}
-\mathbf{c} \\
-\mathbf{t} \\
-\end{array} \right] = 
-\left[ \begin{array}{c}
-\mathbf{x}_1^* \\
-\vdots \\
-\mathbf{x}_i^* \\
-\vdots \\
-\mathbf{x}_n^*
-\end{array} \right]
-\text{,}
-\end{equation}
-and then in the more compact form
-\begin{equation}
-\label{eqn:fitting_one}
-\left[
-\mathbf{A} \quad \mathbf{S}
-\right]
-\left[ \begin{array}{c}
-\mathbf{c} \\
-\mathbf{t} \\
-\end{array} \right] = \mathbf{b}
-\text{,}
-\end{equation}
-where $\mathbf{A}$ is the block containing the monomial matrices of the system, and $\mathbf{S}$ is the block of identity matrices.
+% \begin{equation}
+% \left[\begin{array}{c}
+% \mathbf{M}(\mathbf{X_1-\bar{X}}) \quad \mathbf{I} \\
+% \vdots \\
+% \mathbf{M}(\mathbf{X_i-\bar{X}}) \quad \mathbf{I} \\
+% \vdots \\
+% \mathbf{M}(\mathbf{X_n-\bar{X}}) \quad \mathbf{I} \\
+% \end{array} \right]
+% \left[ \begin{array}{c}
+% \mathbf{c} \\
+% \mathbf{t} \\
+% \end{array} \right] = 
+% \left[ \begin{array}{c}
+% \mathbf{x}_1^* \\
+% \vdots \\
+% \mathbf{x}_i^* \\
+% \vdots \\
+% \mathbf{x}_n^*
+% \end{array} \right]
+% \text{,}
+% \end{equation}
+% and then in the more compact form
+% \begin{equation}
+% \label{eqn:fitting_one}
+% \left[
+% \mathbf{A} \quad \mathbf{S}
+% \right]
+% \left[ \begin{array}{c}
+% \mathbf{c} \\
+% \mathbf{t} \\
+% \end{array} \right] = \mathbf{b}
+% \text{,}
+% \end{equation}
+% where $\mathbf{A}$ is the block containing the monomial matrices of the system, and $\mathbf{S}$ is the block of identity matrices.
 Next, to reduce the degrees of freedom of this system we let the deformed center of mass (the constant term coefficients), $t$, be equal to the mean of the deformed positions influenced by this center of mass.
@@
 In matrix form we have
-\begin{equation}
-\label{eqn:t_solution}
-t = 
-\underbrace{
-\left[\begin{array}{ccccccc}
-\frac{1}{n} & & \cdots \frac{1}{n} & & \\
-& \frac{1}{n} & \cdots & \frac{1}{n} & \\
-& & \frac{1}{n} \cdots & & \frac{1}{n} \\
-\end{array} \right]}_{T} 
-\mathbf{b}
-\text{.}
-\end{equation} (\myworries{spacing on this looks bad....}). Finally, this leaves us with the system
-\begin{equation}
-\mathbf{A}\mathbf{c}=(\mathbf{I}-\mathbf{ST})\mathbf{b}
-\text{.}
-\end{equation}
-Like in the original Shape Matching \cite{10.1145/1073204.1073216}, we solve for the coefficients by minimizing the error in a least squares sense, resulting in
-\begin{equation}
-\label{eqn:c_one_element}
-\begin{split}
-\mathbf{c} & = (\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T(\mathbf{I-ST})\mathbf{b} \\
-           & = \mathbf{L}\mathbf{b}
-\end{split}
-\text{.}
-\end{equation}
-This $\mathbf{L}$ matrix thus gives us our file ingredient in the element, coefficients defining the polynomial.
+% \begin{equation}
+% \label{eqn:t_solution}
+% t = 
+% \underbrace{
+% \left[\begin{array}{ccccccc}
+% \frac{1}{n} & & \cdots \frac{1}{n} & & \\
+% & \frac{1}{n} & \cdots & \frac{1}{n} & \\
+% & & \frac{1}{n} \cdots & & \frac{1}{n} \\
+% \end{array} \right]}_{T} 
+% \mathbf{b}
+% \text{.}
+% \end{equation} (\myworries{spacing on this looks bad....}). Finally, this leaves us with the system
+% \begin{equation}
+% \mathbf{A}\mathbf{c}=(\mathbf{I}-\mathbf{ST})\mathbf{b}
+% \text{.}
+% \end{equation}
+% Like in the original Shape Matching \cite{10.1145/1073204.1073216}, we solve for the coefficients by minimizing the error in a least squares sense, resulting in
+% \begin{equation}
+% \label{eqn:c_one_element}
+% \begin{split}
+% \mathbf{c} & = (\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T(\mathbf{I-ST})\mathbf{b} \\
+%            & = \mathbf{L}\mathbf{b}
+% \end{split}
+% \text{.}
+% \end{equation}
+% This $\mathbf{L}$ matrix thus gives us our file ingredient in the element, coefficients defining the polynomial.
 We emphasize that $\mathbf{L}$ is constant throughout the simulation, therefore as $\mathbf{b}$ varies in the simulation, $\mathbf{Lb}$ gives us a new set of polynomial coefficients as $\mathbf{b}$ changes.
 
-\subsection{Shape Matching with Multiple Elements}
-In a typical simulation we will have multiple shape matching elements where the boundary of each element is defined by a different NURBS surface.
+% \subsection{Shape Matching with Multiple Elements}
+% In a typical simulation we will have multiple shape matching elements where the boundary of each element is defined by a different NURBS surface.
 Therefore for the entire model each shape has a boundary, a polynomial describing its deformation, as well as an associated center of mass.
@@
 
-In defining the shape matching fitting problem, we will form a system similar to equation \ref{eqn:fitting_one}.
+% In defining the shape matching fitting problem, we will form a system similar to equation \ref{eqn:fitting_one}.
 In this case where we have $n$ shape elements our set of unknown coefficients is $\hat{\mathbf{c}} = \left[\mathbf{c}_1^T \dots \mathbf{c}_i^T \dots \mathbf{c}_n^T \right]^T \in \mathbb{R}^{3kn}$ where $\mathbf{c}_i$ corresponds to the unknown polynomial coefficients for the $i$-th shape element.
@@
 With these terms, we write the polynomial fitting system for the entire model as
-\begin{equation}
-\begin{pmatrix}
- \mathbf{A}_1 &      &			  &			&				&\mathbf{S}_1\\  
- &            \ddots &            &  		&				&\vdots\\ 
- &            &      \mathbf{A}_i &   		&				&\mathbf{S}_i\\
- &            &      &            \ddots	&				&\vdots\\
- &            &      &            &			\mathbf{A}_n	&\mathbf{S}_n
-\end{pmatrix}
-\begin{bmatrix}
- \hat{\mathbf{c}}  \\ 
- \hat{\mathbf{t}}   
-\end{bmatrix}
-=
-\hat{\mathbf{b}}
-\text{,}
-\end{equation}
-where $\mathbf{A}_i$ is the monomial basis matrix for the $i$-th shape element, and $\mathbf{S}_i$ is a selection matrix that extracts the $j$-th center of mass which is associated with the $i$-th shape.
+% \begin{equation}
+% \begin{pmatrix}
+%  \mathbf{A}_1 &      &			  &			&				&\mathbf{S}_1\\  
+%  &            \ddots &            &  		&				&\vdots\\ 
+%  &            &      \mathbf{A}_i &   		&				&\mathbf{S}_i\\
+%  &            &      &            \ddots	&				&\vdots\\
+%  &            &      &            &			\mathbf{A}_n	&\mathbf{S}_n
+% \end{pmatrix}
+% \begin{bmatrix}
+%  \hat{\mathbf{c}}  \\ 
+%  \hat{\mathbf{t}}   
+% \end{bmatrix}
+% =
+% \hat{\mathbf{b}}
+% \text{,}
+% \end{equation}
+% where $\mathbf{A}_i$ is the monomial basis matrix for the $i$-th shape element, and $\mathbf{S}_i$ is a selection matrix that extracts the $j$-th center of mass which is associated with the $i$-th shape.
 We simplify this as
-\begin{equation}
-\begin{pmatrix}
-\hat{\mathbf{A}} & \hat{\mathbf{S}}
-\end{pmatrix}
-\begin{bmatrix}
- \hat{\mathbf{c}}  \\ 
- \hat{\mathbf{t}}   
-\end{bmatrix}
-=
-\hat{\mathbf{b}}
-\text{.}
-\end{equation}
-Like in the single element case, we set the $\hat{\mathbf{t}}$ to be equal to the mean of a set of boundary positions, but in this case each $\mathbf{t}_j$ is equal to the mean of only the boundary positions associated with the $j$-th center of mass.
+% \begin{equation}
+% \begin{pmatrix}
+% \hat{\mathbf{A}} & \hat{\mathbf{S}}
+% \end{pmatrix}
+% \begin{bmatrix}
+%  \hat{\mathbf{c}}  \\ 
+%  \hat{\mathbf{t}}   
+% \end{bmatrix}
+% =
+% \hat{\mathbf{b}}
+% \text{.}
+% \end{equation}
+% Like in the single element case, we set the $\hat{\mathbf{t}}$ to be equal to the mean of a set of boundary positions, but in this case each $\mathbf{t}_j$ is equal to the mean of only the boundary positions associated with the $j$-th center of mass.
 The definition of the values of $\hat{\mathbf{t}}$ is similar to \ref{eqn:t_solution}:
-\begin{equation}
-\hat{\mathbf{t}} =
-\underbrace{
-\begin{bmatrix}
-\mathbf{T}_1 \\ 
-\vdots       \\
-\mathbf{T_m}
-\end{bmatrix}}_{\hat{\mathbf{T}}}
-\text{,}
-\end{equation}
+% \begin{equation}
+% \hat{\mathbf{t}} =
+% \underbrace{
+% \begin{bmatrix}
+% \mathbf{T}_1 \\ 
+% \vdots       \\
+% \mathbf{T_m}
+% \end{bmatrix}}_{\hat{\mathbf{T}}}
+% \text{,}
+% \end{equation}
 
-where the entries each $\mathbf{T}_i \in \mathbb{R}^{3N}$ take a similar form, but are only nonzero for columns corresponding to associated boundary vertices.
+% where the entries each $\mathbf{T}_i \in \mathbb{R}^{3N}$ take a similar form, but are only nonzero for columns corresponding to associated boundary vertices.
 
-Finally, to solve for the full set of non-constant term coefficients, $\hat{\mathbf{c}}$, we solve in the same manner as equation \ref{eqn:c_one_element} where we have
-\begin{equation}
-\label{eqn:c_multiple_elements}
-\begin{split}
-\hat{\mathbf{c}} & = (\hat{\mathbf{A}}^T\hat{\mathbf{A}})^{-1}\hat{\mathbf{A}}^T(\mathbf{I-\hat{S}\hat{T})\hat{\mathbf{b}}} \\
-           & = \mathbf{L}\hat{\mathbf{b}}
-\end{split}
-\text{.}
-\end{equation}
+% Finally, to solve for the full set of non-constant term coefficients, $\hat{\mathbf{c}}$, we solve in the same manner as equation \ref{eqn:c_one_element} where we have
+% \begin{equation}
+% \label{eqn:c_multiple_elements}
+% \begin{split}
+% \hat{\mathbf{c}} & = (\hat{\mathbf{A}}^T\hat{\mathbf{A}})^{-1}\hat{\mathbf{A}}^T(\mathbf{I-\hat{S}\hat{T})\hat{\mathbf{b}}} \\
+%            & = \mathbf{L}\hat{\mathbf{b}}
+% \end{split}
+% \text{.}
+% \end{equation}
 
-\subsection{Hierarchical fitting}
-A problem may emerge in the shape matching process where the system may be underdetermined.
+% \subsection{Hierarchical fitting}
+% A problem may emerge in the shape matching process where the system may be underdetermined.
 To account for this we solve for the coefficients in a hierarchical manner where we initially solve for the linear terms and at each step solve the coefficients for the next higher order of deformation.
@@
 To describe our fitting strategy, we begin with the following shape matching system for a single element
-\begin{equation}
-\label{eqn:fitting_one}
-\left[
-\mathbf{A} \quad \mathbf{S}
-\right]
-\left[ \begin{array}{c}
-\mathbf{c} \\
-\mathbf{t} \\
-\end{array} \right] = \mathbf{b}
-\text{.}
-\end{equation}
-We will show the hierarchical fitting on a polynomial with a set of coefficients for the linear terms, $\mathbf{c}_{\text{lin}}$, and the coefficients for the quadratic terms $\mathbf{c}_{\text{quad}}$.
+% \begin{equation}
+% \label{eqn:fitting_one_copy}
+% \left[
+% \mathbf{A} \quad \mathbf{S}
+% \right]
+% \left[ \begin{array}{c}
+% \mathbf{c} \\
+% \mathbf{t} \\
+% \end{array} \right] = \mathbf{b}
+% \text{.}
+% \end{equation}
+% We will show the hierarchical fitting on a polynomial with a set of coefficients for the linear terms, $\mathbf{c}_{\text{lin}}$, and the coefficients for the quadratic terms $\mathbf{c}_{\text{quad}}$.
 Similarly, we also have $\mathbf{A}_{\text{lin}}$ and $\mathbf{A}_{\text{quad}}$ which are the blocks of $\mathbf{A}$ corresponding to the columns for the linear and quadratic monomial terms, respectively.
 First we construct a system for the linear terms:
-\begin{equation}
-\label{eqn:fitting_linear}
-\left[
-\mathbf{A}_{\text{lin}} \quad \mathbf{S}
-\right]
-\left[ \begin{array}{c}
-\mathbf{c}_{\text{lin}} \\
-\mathbf{t} \\
-\end{array} \right] = \mathbf{b}
-\text{,}
-\end{equation}
-which leads to the following expression for the linear coefficients
-\begin{equation}
-\label{eqn:c_linear}
-\begin{split}
-\mathbf{c}_{\text{lin}} & = (\mathbf{A}_{\text{lin}}^T\mathbf{A}_{\text{lin}})^{-1}\mathbf{A}_{\text{lin}}^T(\mathbf{I-ST})\mathbf{b} \\
-           & = \mathbf{L}_{\text{lin}}\mathbf{b}
-\end{split}
-\text{.}
-\end{equation}
-where $\mathbf{T}$ is defined in equation \ref{eqn:t_solution}. For the quadratic system, we subtract linear term estimate of the solution and solve for the coefficients that fit this residual:
-\begin{equation}
-\label{eqn:fitting_quadratic}
-\left[
-\mathbf{A}_{\text{quad}} \quad \mathbf{S}
-\right]
-\left[ \begin{array}{c}
-\mathbf{c}_{\text{quad}} \\
-\mathbf{t} \\
-\end{array} \right] = \mathbf{b} - \mathbf{A}_{\text{lin}}\mathbf{c}_{\text{lin}}
-\text{.}
-\end{equation}
-Using the solution for the linear coefficients in \ref{eqn:c_linear}, we arrive at the solution for the quadratic coefficients:
-\begin{equation}
-\label{eqn:c_quad}
-\begin{split}
-\mathbf{c}_{\text{quad}} & = (\mathbf{A}_{\text{quad}}^T\mathbf{A}_{\text{quad}})^{-1}\mathbf{A}_{\text{quad}}^T((\mathbf{I-ST})\mathbf{b} - \mathbf{A}_{\text{lin}}\mathbf{c}_{\text{lin}}) \\
-		   & = (\mathbf{A}_{\text{quad}}^T\mathbf{A}_{\text{quad}})^{-1}\mathbf{A}_{\text{quad}}^T(\mathbf{I-ST} - \mathbf{A}_{\text{lin}}\mathbf{L}_{\text{lin}})\mathbf{b} \\
-           & = \mathbf{L}_{\text{quad}}\mathbf{b}
-\end{split}
-\text{.}
-\end{equation}
-The final form for the full $\mathbf{L}$ matrix is thus given by
-\begin{equation}
-\mathbf{L} =
-\begin{bmatrix}
-\mathbf{L}_{\text{lin}} \\
-\mathbf{L}_{\text{quad}} \\
-\mathbf{T}
-\end{bmatrix}
-\end{equation}
+% \begin{equation}
+% \label{eqn:fitting_linear}
+% \left[
+% \mathbf{A}_{\text{lin}} \quad \mathbf{S}
+% \right]
+% \left[ \begin{array}{c}
+% \mathbf{c}_{\text{lin}} \\
+% \mathbf{t} \\
+% \end{array} \right] = \mathbf{b}
+% \text{,}
+% \end{equation}
+% which leads to the following expression for the linear coefficients
+% \begin{equation}
+% \label{eqn:c_linear}
+% \begin{split}
+% \mathbf{c}_{\text{lin}} & = (\mathbf{A}_{\text{lin}}^T\mathbf{A}_{\text{lin}})^{-1}\mathbf{A}_{\text{lin}}^T(\mathbf{I-ST})\mathbf{b} \\
+%            & = \mathbf{L}_{\text{lin}}\mathbf{b}
+% \end{split}
+% \text{.}
+% \end{equation}
+% where $\mathbf{T}$ is defined in equation \ref{eqn:t_solution}. For the quadratic system, we subtract linear term estimate of the solution and solve for the coefficients that fit this residual:
+% \begin{equation}
+% \label{eqn:fitting_quadratic}
+% \left[
+% \mathbf{A}_{\text{quad}} \quad \mathbf{S}
+% \right]
+% \left[ \begin{array}{c}
+% \mathbf{c}_{\text{quad}} \\
+% \mathbf{t} \\
+% \end{array} \right] = \mathbf{b} - \mathbf{A}_{\text{lin}}\mathbf{c}_{\text{lin}}
+% \text{.}
+% \end{equation}
+% Using the solution for the linear coefficients in \ref{eqn:c_linear}, we arrive at the solution for the quadratic coefficients:
+% \begin{equation}
+% \label{eqn:c_quad}
+% \begin{split}
+% \mathbf{c}_{\text{quad}} & = (\mathbf{A}_{\text{quad}}^T\mathbf{A}_{\text{quad}})^{-1}\mathbf{A}_{\text{quad}}^T((\mathbf{I-ST})\mathbf{b} - \mathbf{A}_{\text{lin}}\mathbf{c}_{\text{lin}}) \\
+% 		   & = (\mathbf{A}_{\text{quad}}^T\mathbf{A}_{\text{quad}})^{-1}\mathbf{A}_{\text{quad}}^T(\mathbf{I-ST} - \mathbf{A}_{\text{lin}}\mathbf{L}_{\text{lin}})\mathbf{b} \\
+%            & = \mathbf{L}_{\text{quad}}\mathbf{b}
+% \end{split}
+% \text{.}
+% \end{equation}
+% The final form for the full $\mathbf{L}$ matrix is thus given by
+% \begin{equation}
+% \mathbf{L} =
+% \begin{bmatrix}
+% \mathbf{L}_{\text{lin}} \\
+% \mathbf{L}_{\text{quad}} \\
+% \mathbf{T}
+% \end{bmatrix}
+% \end{equation}
 
-The recursive nature of this process permits a more general description for arbitrary orders, which we write in algorithm \myworries{TODO}.
+% The recursive nature of this process permits a more general description for arbitrary orders, which we write in algorithm \myworries{TODO}.
 
-\section{The Deformation Map}
-At this point we have shown how to fit a polynomial describing the deformation of each surface.
+% \section{The Deformation Map}
+% At this point we have shown how to fit a polynomial describing the deformation of each surface.
 With these polynomial coefficients coupled with their associated centers of mass, we can construct an estimate of some arbitrary material point $\mathbf{X}$ using the deformation described by the surface.
@@
 Thus we arrive at a definition for the deformed position, $\mathbf{x}$, of a material point $\mathbf{X}$ in the undeformed space:
-\begin{equation}
-	\mathbf{x}(\mathbf{X}) = \sum_i^n w_i(\mathbf{X}) \left[ \mathbf{M}(\mathbf{X-\bar{X}}_j)\mathbf{c}_i + \mathbf{t}_j \right]
-	\text{,}
-\end{equation}
-where we have $n$ surfaces, and $w_i$ and $\mathbf{c}_i$ are the weights and the coefficients for the $i$-th surface, respectively.
+% \begin{equation}
+% \label{eqn:deformation_map}
+% 	\mathbf{x}(\mathbf{X}) = \sum_i^n w_i(\mathbf{X}) \left[ \mathbf{M}(\mathbf{X-\bar{X}}_j)\mathbf{c}_i + \mathbf{t}_j \right]
+% 	\text{,}
+% \end{equation}
+% where we have $n$ surfaces, and $w_i$ and $\mathbf{c}_i$ are the weights and the coefficients for the $i$-th surface, respectively.
 The terms $\mathbf{\bar{X}}_j$ and $\mathbf{t}_j$ the undeformed and deformed centers of mass, respectively, that are associated with the $i$-th surface.
 
-\subsection{Computing the Blending Weights}
-All that remains for the deformation map is to compute blending weights.
+% \subsection{Computing the Blending Weights}
+% All that remains for the deformation map is to compute blending weights.
 Our choice of the weighting function is guided by the simple idea that that for a given material point its weight should be highest to whichever surfaces it is closest.
@@
 
-For the $i$-th surface, we first find the distance, $d_{\text{primary}}$, from the point $\mathbf{X}$ to the closest point, $\mathbf{p}$, on the surface.
+% For the $i$-th surface, we first find the distance, $d_{\text{primary}}$, from the point $\mathbf{X}$ to the closest point, $\mathbf{p}$, on the surface.
 Next we wish to satisfy the condition that the distance weight is highest on the surface and it decays towards zero as it approaches other surfaces.
@@
 These two distances enable us to write a simple linear function to compute the \textit{distance weight}:
-\begin{equation}
-\theta_i = \max (1.0 - \frac{d_{\text{primary}}}{\min (d_{\text{total}}, D)}, 0.0)
-\text{,}
-\end{equation}
-where $D$ is a fixed cutoff distance parameter.
+% \begin{equation}
+% \theta_i = \max (1.0 - \frac{d_{\text{primary}}}{\min (d_{\text{total}}, D)}, 0.0)
+% \text{,}
+% \end{equation}
+% where $D$ is a fixed cutoff distance parameter.
 We see that this distance weight is the result of interpolating along this ray from $\mathbf{p}$ to $\mathbf{q}$ (or a fixed point that depends on $D$).
@@
 
-Now that we have a set of distances weights $\mathbf{theta}$, we seek to form a final set of \textit{blending weights} $\mathbf{w}(\mathbf{X}) =  \left( w_1, \dots, w_i, \dots, w_n \right)$ to blend the polynomials of the $n$ shape elements.
+% \begin{figure}
+%     \includegraphics[width=\columnwidth]{example-image-a}
+%     \caption{Figure showing how weight calculation works (left) ray casting plus result (right) after correcting for partition of unity}
+%     \label{fig:weightcompute}
+% \end{figure}
+
+% Now that we have a set of distances weights $\mathbf{theta}$, we seek to form a final set of \textit{blending weights} $\mathbf{w}(\mathbf{X}) =  \left( w_1, \dots, w_i, \dots, w_n \right)$ to blend the polynomials of the $n$ shape elements.
 In constructing these blending weights, we require that they adhere to a set of conditions.
@@
 To produce these blending weights we form a quadratic program that satisfies these constraints:
-\begin{equation}
-\begin{aligned}
-\mathbf{w}(\mathbf{X}) = \min_{\mathbf{w}} \quad & \mathbf{w}^T \Theta(\mathbf{X}) \mathbf{w}    \\
-\textrm{s.t.} \quad & 0 \leq w_i \leq 1                     \\
-                    &   \sum_i^n w_i = 1                      \\
-\end{aligned}
-\end{equation}
+% \begin{equation}
+% \begin{aligned}
+% \mathbf{w}(\mathbf{X}) = \min_{\mathbf{w}} \quad & \mathbf{w}^T \Theta(\mathbf{X}) \mathbf{w}    \\
+% \textrm{s.t.} \quad & 0 \leq w_i \leq 1                     \\
+%                     &   \sum_i^n w_i = 1                      \\
+% \end{aligned}
+% \end{equation}
 
-where
-\begin{equation}
-     \Theta(\mathbf{X}) = \text{diag}\left( \frac{1}{\theta_1},\frac{1}{\theta_2},\dots,\frac{1}{\theta_n}\right)
-\end{equation}
-We see that through the use of linear constraints we satisfy conditions (1) and (2).
+% where
+% \begin{equation}
+%      \Theta(\mathbf{X}) = \text{diag}\left( \frac{1}{\theta_1},\frac{1}{\theta_2},\dots,\frac{1}{\theta_n}\right)
+% \end{equation}
+% We see that through the use of linear constraints we satisfy conditions (1) and (2).
 Satisfying the third condition (3) is achieved through using the distance weights.
@@
 
-\section{Selecting Centers of Mass}
-\myworries{TODO. waiting to implement this before I start writing how we do it}
-\section{Raycasting Quadrature}
-The last step before we may discuss the dynamic simulation is to describe our strategy for generating material points $\mathbf{X}$ in the undeformed model.
+% \section{Selecting Centers of Mass}
+% \myworries{TODO. waiting to implement this before I start writing how we do it}
+% \section{Raycasting Quadrature}
+% The last step before we may discuss the dynamic simulation is to describe our strategy for generating material points $\mathbf{X}$ in the undeformed model.
 Our volume integration strategy is based on \cite{KHOSRAVIFARD201030} which we briefly describe in the following section
-\subsection{Review of the Meshless Method for Domain Integral Evaluation in \cite{KHOSRAVIFARD201030}}
-This paper addresses the integration of volume integrals of the form $\mathbf{I} = \int_\Omega f(\mathbf{x}) d\Omega$ where $\mathbf{x} = (x_1,x_2,x_3)$.
+% \subsection{Review of the Meshless Method for Domain Integral Evaluation in \cite{KHOSRAVIFARD201030}}
+% This paper addresses the integration of volume integrals of the form $\mathbf{I} = \int_\Omega f(\mathbf{x}) d\Omega$ where $\mathbf{x} = (x_1,x_2,x_3)$.
 The authors show that through application of the Green-Gauss theorem, this integral may be rewritten as 2D integral over a $yz$ plane where at each position along the plane we evaluate a line integral along the $x$-axis.
@@
 
-In our approach we modify the raycasting procedure by allowing integration over intersecting NURBS surfaces.
+% In our approach we modify the raycasting procedure by allowing integration over intersecting NURBS surfaces.
 It is common in NURBS modeling to produce models in which many surfaces overlap.
@@
 
-\section{Dynamic Simulation}
-\subsection{Generalized Inertia}
-\subsection{Generalized Forces}
-\subsection{Equations of Motion}
-%This is most notable in cases where the degree of the polynomial on a NURBS is higher than the degree of the B-spline surface.
+% \begin{figure}
+%     \includegraphics[width=\columnwidth]{example-image-a}
+%     \caption{Figure showing how raycasting quadrature works (top), convergence plot for volume integration (bottom}
+%     \label{fig:raycasting}
+% \end{figure}
 
-% \textbf{x}(\textbf{X}) = \hat{A}M(\textbf{X}) + \bar{\textbf{x}}$ \\
-% $\hat{A} = \sum_i^{|S|}w_i(\textbf{X})A_i$
-%
-%\subsection{Deformation Gradient}
-%\begin{equation}
-%        \mathbf{F} = \frac{\partial \mathbf{x}}{\partial \mathbf{X}}
-%               = \frac{\partial \hat{\mathbf{\Pi}}\mathbf{M(X)}}{\partial \mathbf{X}}
-%               = \hat{\mathbf{\Pi}} \frac{\partial \mathbf{M(X)}}{\partial \mathbf{X}} 
-%\end{equation}
-%The $\frac{\partial \mathbf{M(X)}}{\partial \mathbf{X}}$ terms may be precomputed and using the previous $\mathbf{M(X)}$ example, we have
-%\begin{equation}
-%     \frac{\partial \mathbf{M(X)}}{\partial \mathbf{X}} = \begin{bmatrix}
-%     1&0 \\ 
-%     0&1 \\ 
-%     2(X_0 - \bar{X}_0)&0 \\ 
-%     0&2(X_1 - \bar{X}_1) \\ 
-%     (X_1 - \bar{X}_1)&(X_0 - \bar{X}_0) 
+% \section{Dynamic Simulation}
+% With our definition of the deformation map as well as a strategy for integrating the volume, we now have the necessary ingredient to perform dynamic simulation.
+We follow the Variational Mechanics approach that defines a pair of kinetic and potential energies, and we apply the Euler-Lagrange equation to develop our equations of motion.
+
+% Before we define these energies we first would like to simplify the form of the deformation map presented in equation \ref{eqn:deformation_map}.
+We may write $\mathbf{x}(\mathbf{X})$ in matrix form as
+% \begin{equation}
+% \mathbf{x}(\mathbf{X}) = \underbrace{\left[w_1\mathbf{M}(\mathbf{X - \bar{X}}_j) \dots w_n\mathbf{M}(\mathbf{X - \bar{X}}_j) \middle| \sum_i^{n_1}w_1^i\mathbf{I} \dots \sum_i^{n_m}w_m^i\mathbf{I} \right]}_{\mathbf{Y}(\mathbf{X})}
+% \begin{bmatrix}
+% \mathbf{\hat{c}} \\
+% \mathbf{\hat{t}}
 % \end{bmatrix}
-%\end{equation}
-%
-%In computing the forces and the stiffness matrix, we require the gradient of the deformation gradient with respect to the configuration,  $\frac{\partial \mathbf{F}}{\partial \mathbf{q}}$.
-Usually the form of this gradient is simple.
-For example, in linear tetrahedral FEM each quadrature point's deformation map only involves the four surrounding nodal values, simplifying this gradient.
-In our case, the presence of the projection operator $\hat{\mathbf{\Pi}}$ means that each quadrature point is computed using the nodal values of the entire model.
-%
-%\begin{equation}
-%    \frac{\partial \mathbf{F}}{\partial \mathbf{q}} = \frac{\partial \mathbf{P}}{\partial q}\mathbf{M(q_0)}^\dagger  \frac{\partial \mathbf{M(X)}}{\partial \mathbf{X}} 
-%\end{equation}
-%
-%
-%Naively handled, this gives us a $\frac{\partial \mathbf{F}}{\partial \mathbf{q}}$ of size $3 x |q|$.
-However, many of the columns of have values all near zero, allowing us to prune many columns for each quadrature point.
-Currently, I'm computing the infinity norm of each column and prune all except some fixed number of columns.
-\\
-%
-%
-%\subsection{Variational Mechanics}
-%We have now provided all the necessary ingredient to produce equations of motion for deformable bodies, which we derive via Euler-Lagrange equation. We begin with the Lagrangian
-%\begin{equation}
-%    L=T-V
-%\end{equation}
-%where $T$ is the kinetic energy, and $V$ is the potential energy. From the Euler-Lagrange equation we have:
-%\begin{equation}
+% \text{,}
+% \end{equation}
+% \myworries{I'm not sure how to write this matrix form for x(X), especially on finding the best notation for the center of mass indexing. So this above equation is temporary.}
+% where we have $n$ shape elements, $m$ centers of mass. We simplify the above equation and use the following form
+% \begin{equation}
+% \label{eqn:x_simple_form}
+% \mathbf{x}(\mathbf{X}) = \mathbf{Y}(\mathbf{X})\mathbf{L}\hat{\mathbf{b}}
+% \text{,}
+% \end{equation}
+% where we make use of equation \ref{eqn:c_multiple_elements} to rewrite this in terms of the boundary points $\mathbf{\hat{b}}$.
+As one final step, we rewrite this again in terms of the control points of our NURBS surfaces:
+% \begin{equation}
+% \label{eqn:compact_x_map}
+% \mathbf{x}(\mathbf{X}) = \mathbf{Y}(\mathbf{X})\mathbf{LJq}
+% \text{,}
+% \end{equation}
+% where we use the fact that $\mathbf{\hat{b}} = \mathbf{Jq}$.
+\myworries{not a fan of this, perhaps we change $\hat{b}$ to something else...} This gives us an expression for the deformed position in terms of our generalized coordinates, $\mathbf{q}$.
+
+% \subsection{Generalized Inertia}
+% The kinetic energy for a model may be expressed as
+% \begin{equation}
+% T = \frac{1}{2}\int_{ \Omega} \rho \dot{\mathbf{x}}^T\dot{\mathbf{x}} d\Omega
+% \text{,}
+% \end{equation}
+% where $\Omega$ is the 3D integration domain, $\rho$ is the density, and $\dot{\mathbf{x}}$ is the velocity for some point in the domain.
+Using the deformation map defined in \ref{eqn:compact_x_map}, we may rewrite this kinetic energy as
+% \begin{equation}
+% \begin{split}
+% T & = \frac{1}{2}\int_{ \Omega} \rho \dot{\mathbf{q}}^T \mathbf{J}^T\mathbf{Y(X)}^T\mathbf{Y(X)}\ d\Omega \\
+%   & = \frac{1}{2} \dot{\mathbf{q}}^T \mathbf{J}^T \left[ \int_{ \Omega} \rho \mathbf{Y(X)}^T\mathbf{Y(X)} d\Omega \right] \mathbf{J}\dot{\mathbf{q}} \\
+%   & = \frac{1}{2} \dot{\mathbf{q}}^T \mathbf{M} \dot{\mathbf{q}}
+% \end{split}
+% \text{.}
+% \end{equation}
+% The second line is a result of the fact that all terms except for $\mathbf{Y(X)}$ are spatially constant, allowing us to move these outside the volume integral, which enables fast assembly of the mass matrix, $\mathbf{M}$.
+% \subsection{Error Energy}
+% Over the course of the simulation we may see the true boundary values deviate from their positions estimated by the polynomial.
+This is most notable in cases where the degree of the polynomial on a NURBS is higher than the degree of the B-spline surface.
+To address this we augment our kinetic energy with a term to account for this error.
+The error for a boundary point $\mathbf{x}_i^*$ that is known to lie on the limit surface of the NURBS can be written as
+% \begin{equation}
+% \begin{split}
+% \epsilon_i & = ||\mathbf{x}(\mathbf{X}_i) - \mathbf{x}_i^*|| \\
+% 		   & = ||(\mathbf{Y}(\mathbf{X}_i)\mathbf{L} - \mathbf{S_i})\mathbf{\hat{b}}||
+% \end{split}
+% \text{,}
+% \end{equation}
+% where $\mathbf{S_i}$ extracts $\mathbf{x}_i^*$ from $\mathbf{\hat{b}}$.
+We modify this error further to be expressed in terms of the control points.
+We note that $\mathbf{x}^*(u,v)$ yields the position on the NURBS surface whereas $\mathbf{x(X)}$ is the deformed value estimated by the polynomial fitting.
+Furthermore $\mathbf{X}$ may be written as a function of the parametric coordinates $\mathbf{X}(u,v)$ as its position on the surface is produced using the initial set of control points representing the undeformed model.
+This leads to an energy potential over the NURBS surface in parametric space
+% \begin{equation}
+% T_E = \iint (\mathbf{x}(\mathbf{X}(u,v)) - \mathbf{x}^*(u,v))^T(\mathbf{x}(\mathbf{X}(u,v)) - \mathbf{x}^*(u,v)) du dv
+% \text{.}
+% \end{equation}
+% The discrete form of this written in terms of the control points may be written as
+% \begin{equation}
+% \begin{split}
+% T_E & = \sum_i^N  \dot{\mathbf{q}}^T\mathbf{J}^T(\mathbf{Y}(\mathbf{X}_i)\mathbf{L} - \mathbf{S_i})^T(\mathbf{Y}(\mathbf{X}_i)\mathbf{L} - \mathbf{S_i})\mathbf{J}\dot{\mathbf{q}} \\
+%   & = \dot{\mathbf{q}}^T\mathbf{J}^T \left[ \sum_i^N  (\mathbf{Y}(\mathbf{X}_i)\mathbf{L} - \mathbf{S_i})^T(\mathbf{Y}(\mathbf{X}_i)\mathbf{L} - \mathbf{S_i}) \right] \mathbf{J}\dot{\mathbf{q}} \\
+%   & = \dot{\mathbf{q}}^T \mathbf{M}_E \dot{\mathbf{q}}
+% \end{split}
+% \text{,}
+% \end{equation}
+% where $N$ is the total number of boundary points and $\mathbf{M}_E$ is our \textit{error mass matrix}.
+% \subsection{Generalized Forces}
+% We write our potential energy as 
+% \begin{equation}
+%     V = \int_\Omega \psi(\mathbf{F(X)}) d\Omega
+% \text{.}
+% \end{equation}
+% To evaluate this energy (as well as its derivatives), we require a definition for the deformation gradient, $\mathbf{F(X)}$:
+% \begin{equation}
+%         \mathbf{F(X)} = \frac{\partial \mathbf{x}}{\partial \mathbf{X}}
+%                = \frac{\partial \mathbf{Y(X)}}{\partial \mathbf{X}}\mathbf{L}\hat{\mathbf{b}}
+% \text{.}
+% \end{equation}
+% In the expression for the deformed position $\mathbf{x(X)}$ we see that the dependence on $\mathbf{X}$ only exists in $\mathbf{Y(X)}$, making $\mathbf{L}$ and $\hat{\mathbf{b}}$ constant terms when compute $\mathbf{F}$.
+Full details on how to compute $\mathbf{F}$ and its derivative $\frac{\partial \mathbf{F}}{\partial \mathbf{q}}$ may be found in the appendix.
+(\myworries{I just wrote this because I'm too tired to right the d etails right now :p}.
+Now that we have an expression for the deformation gradient, we discretize our potential energy as follows
+% \begin{equation}
+%     V =  \sum_i^m \psi(\mathbf{F(X}_i) A_i
+% \text{,}
+% \end{equation}
+% where $m$ is the number of integration points and $A_i$ is the volume for the $i$-th integration point.
+
+% Like in the case of the kinetic energy, we also introduce an error force potential, which takes the simple form
+% \begin{equation}
+% V_E = \mathbf{q}^T \mathbf{M}_E \mathbf{q}
+% \end{equation}
+
+% \subsection{Time Integration}
+% We have now provided all the necessary ingredient to produce equations of motion for deformable bodies, which we derive via Euler-Lagrange equation. We begin with the Lagrangian
+% \begin{equation}
+%     L=(T+T_E)-(V + V_E)
+% \end{equation}
+% Then from the Euler-Lagrange equation we have:
+% \begin{equation}
 %     \frac{d}{dt} \frac{\partial L}{\partial \dot{\mathbf{p}}} + \frac{\partial L}{\partial {\mathbf{p}}} = 0
-%\end{equation}
-%\begin{equation}
-%    V =  \int_\Omega \psi(\mathbf{F(X)} dX
-%\end{equation}
-%then discretizing this over $m$ quadrature points, we have
-%\begin{equation}
-%    V =  \sum_i^m \psi(\mathbf{F(X_i)} A_i
-%\end{equation}
+% \end{equation}
+% \myworries{todo: not totally sure what to write here since we're experimenting with both linearly implicit backward euler as well as Newton's method soon}
+
 %\begin{equation}
@@
 %\end{equation}
-%Next we have the equation of the kinetic energy
-%\begin{equation}
-%    T = \frac{1}{2}\int_{ \Omega} \rho \dot{\mathbf{x}}^T\dot{\mathbf{x}} dV
-%\end{equation}
-%which we rewrite in terms of our control points configuration
-%\begin{equation}
-%    T = \frac{1}{2} \int_{\partial \Omega} \dot{\mathbf{p}}^T \mathbf{J}(u,v)^T \mathbf{M} \mathbf{J}(u,v) \dot{\mathbf{p}} du dv
-%\end{equation}
-%\myworries{change notation to not use M for the monomial basis}
-%where
-%\begin{equation}
-%    \mathbf{M} = \int_\Omega \rho \mathbf{J_{\hat{\Pi}}(X)}^T \mathbf{J_{\hat{\Pi}}(X)}
-%\end{equation}
-%From the potential energy (which uses dF/dq) and this potential energy, we discretize in time and right now I'm using linearly implicit backward euler
-%
-%\subsection{Stability term}
-%\begin{figure}[h]
-%  \centering
-%  \includegraphics[width=\linewidth]{figures/error_energy}
-%  \caption{A figure showing the spring energy error for each sampled point on the NURBS curve. I should change the colors of the lines to show higher energies where the error is higher.}
-%  \Description{Showing an example error of a NURBS curve.}
-%\end{figure}
-%
-%\subsection{Projection Operator Weighting}
-%\begin{figure}[h]
-%  \centering
-%  \includegraphics[width=\linewidth]{figures/distance_weighting}
-%  \caption{A 2D example showing how $\theta$ is computed for $\Pi_B$.
-$d_1$ shows the distance to the closest point on $\Pi_B$ to $\mathbf{x}$ and $d_2$ shows the distance to the surface hit by the ray in the opposite direction.}
-%  \Description{Showing a 2D NURBS curve with a quadrature point in the center.
-A ray is cast to the nearest point on a shape, showing the distance between the quadrature point and the nearest point as well as the distance to the "back surface".}
-%\end{figure}
-%
-%
-%\subsection{Integrating the volume}
-%\begin{figure}[h]
-%  \centering
-%  \includegraphics[width=\linewidth]{figures/raycasting_quadrature}
-%  \caption{An example raycasting quadrature on a 2D NURBS curve.}
-%  \Description{Shows a 2D NURBs curve with horizontal rays intersecting the curve. }
-%\end{figure}
+%\

### 2021-01-19 paper/sections/methods.tex
ok remixed the methods so equations are where I think they should go, now need to make it make sense again

--- 2021-01-17 paper/sections/methods.tex
@@
 
+\section{Overview}
+The input to the Shape Matching Element Method (SEM)  is a NURBS \emph{boundary representation} of a volumetric object, along with a set of physical parameters 
+(density, constitutive model, model parameters etc). The output is an elastodynamic simulation. 
+SEM consists of preprocessing and runtime simulation phases~\refalg{sem} and at no point do we need to generate a volumetric mesh of the input. 
+In the \emph{preprocessing} stage we use raycasting to find quadrature points and weights for volumetric integration, as well as to compute part blending parameters. 
+Finally we construct local shape matching operators and construct the mass matrix for our problem.
+
+At runtime we use standard time integration schemes to time step the system (we show linearly implicit, fully implicit, runge-kutta four). 
+These require us to evaluate the energy, gradient and Hessian (implicit only) which we do at the previously computed quadrature points. 
+For collisions we use penalty forces and apply them as external forces during integration.
+
+\begin{algorithm}[h]
+    \label{alg:sem}
+    \SetAlgoLined
+    \KwResult{Write here the result }
+     initialization\;
+     \While{While condition}{
+      instructions\;
+      \eIf{condition}{
+       instructions1\;
+       instructions2\;
+       }{
+       instructions3\;
+      }
+     }
+     \caption{SEM Algorithm}
+    \end{algorithm}
+
 \section{Methods}
-Given a volumetric model composed of one or more NURBS surfaces, the surface of each surface is reconstructed using a set of control points $\textbf{p}$ and their associated weights $w$.
-Following D-NURBS \cite{10.1145/176579.176580}, we take our degrees of freedom (DOF) to be control points of the surfaces and produce displacements directly on these control points at each simulation step.
-We note that we used a simplified version of D-NURBS in which the weights are held constant throughout the simulation and find this not to be limiting \myworries{(feel like we'll need better justification than saying just this?)}.
-We represent the degrees of freedom for the $i$-th NURBS surface with $m$ control points as $\mathbf{q}_i = \left[ \mathbf{p}_1^T, \dots, \mathbf{p}_m^T \right]^T \in \mathbb{R}^{3m}$.
-From this definition, we write the DOF for the entire model as $\mathbf{q} = \left[ \mathbf{q}_i^T, \dots, \mathbf{q}_n^T \right]^T$ where $n$ is the number of NURBS surfaces for the model.
 
-To evaluate the volumetric integrals in solid mechanics models, we require material points $\mathbf{X} \in \mathbb{R}^{3m}$ in the undeformed space that serve as integration points.
-Our shape matching element method enables us to evaluate the deformed positions of the material points strictly using positions defined on the boundary.
-Like in VEM, we fit a polynomial to each boundary primitive (NURBS in this case) describing its deformation.
-The fitting of these polynomials closely follows that of the Shape Matching \cite{10.1145/1073204.1073216} approach where we perform a least squares polynomial fit to the deformed boundary points.
-These boundary points are represented by sampled positions on each NURBS surface.
-Since each surface is associated with a single polynomial, we can construct a unique polynomial for some arbitrary point by blending these polynomials.
-Thus given some material point $\mathbf{X}_i$ we find it's corresponding deformed position $\mathbf{x}_i$ via a weighted combination of the polynomials.
-This permits us to build a general deformation map which we use to construct our equations of motion.
+\dave{explain NURBS but not D-NURBS because that's Terzopolous's specific algorithm just explain the NURBS parameterization}
 
-In the following sections, we will first review the simplified D-NURBS model used to form our generalized coordinates.
-We then present our shape matching algorithm for fitting polynomials to surfaces ....
-In section ...
-we next describe how to blend these surface polynomials to construct a general deformation map for material points.
-Section ...
-describes our strategy for selecting integration points in the volume through a raycasting approach.
-Finally, we formulate our equations of motions from a Lagrangian formulation.
-
-\section{D-NURBS Model}
-\myworries{TODO: Most of these matrices should be paranthesis matrices}
-We provide a brief overview of the D-NURBS formulation presented in \cite{10.1145/176579.176580}. Our models in our method are composed of a collection of NURBS surfaces.
-
-\subsection{D-NURBS Formulation}
 A point on a NURBS surface is typically expressed in the following form
@@
 
-This formulation is a simplification of the full D-NURBS described in \cite{10.1145/176579.176580} the weights of the NURBS are fixed throughout the simulation.
-Permitting the dynamics to modify weights would likely improve the expressiveness of the kinematics, but we find fixing the weights affords sufficiently expressive results and improves performance.
-The performance improvement is a result of having fixed $N_{i,j}(u,v)$ entries over the course of the simulation.
-The full D-NURBS model requires rebuilding the jacobian and consequently the mass matrices at each time step.
-
-\subsection{Generalized Coordinates for a NURBS model}
-The above formulation describes developing the map from a single $(u,v)$ coordinate to its corresponding position on the surface.
-In our simulation we represent the entire boundary of the NURBS model, so we require samples along each surface.
-We emphasize that we only require an initial selection of $(u,v)$ coordinates across each surface and the dynamics of the simulation modifies the set of control points $\mathbf{q}$ to modify the world space maps $\mathbf{x}(u,v)=\mathbf{J}(u,v)\mathbf{q}$.
-
-Let's say for a NURBS surface on the model the DOF column vector is $\mathbf{q}_i$, which is the subset of control points in the generalized coordinates for the $i$-th surface.
-For a given surface with $m$ pairs of $\mathbf{u}=(u,v)$ coordinates sampled in parametric space, we write the jacobian for the $i$-th surface as \myworries{(should I use something like $\hat{\mathbf{J}}$ instead?)}
-\begin{equation}
-\label{eqn:surface_jacobian}
-    \mathbf{J}_i =
-    \left[ \mathbf{J}(\mathbf{u})_1^T \dots \mathbf{J}(\mathbf{u})_m^T \right]^T
-    \text{.}
-\end{equation}
-Then if $\mathbf{x}_i$ is the set of $m$ world space positions for the $i$-th surface, we may write $\mathbf{x}_i = \mathbf{J}_i \mathbf{q}_i$.
-Finally, given the full set of generalized coordinates $\mathbf{q} = \left[ \mathbf{q}_i^T, \dots, \mathbf{q}_n^T \right]^T$ with $n$ surfaces, we may write jacobian $\mathbf{J}$ for the entire model as a block diagonal matrix:
-\begin{equation}
-\mathbf{J} = \left[ \begin{array}{ccccc}
-\mathbf{J}_1 &  &  &  &  \\
- & \ddots &  &  &  \\
- &  & \mathbf{J}_i & &  \\
- &  &  & \ddots &  \\
- &  &  &  & \mathbf{J}_n \\
-\end{array} \right]
-\text{.}
-\end{equation}
 With this equation, the full set of world positions of the NURBS model may be written as $\mathbf{x} = \mathbf{J}\mathbf{q} \in \mathbb{R}^{3N}$ where $N$ is the total number of $(u,v)$ coordinates.
 
-\section{Shape Matching}
-The key feature of our method is its unique deformation map $\phi(\mathbf{X}) = \mathbf{x}$ mapping some arbitrary undeformed material point $\mathbf{X}$ to its corresponding deformed position in space $\mathbf{x}$.
-We develop this similarly to the VEM approach in that we say the function we intend to reproduce may be some highly nonlinear one, but it may be approximated accurately by a polynomial.
-Thus we seek some $\phi$ function where  $\phi^* \approx \phi$ where $\phi^*$ is the true solution.
-In $1$-D this means we wish to find a function of the form $\phi(X) = c_1 + c_2X + c_3X^2 + \dots + c_nX^{n-1}$ where each $c_i$ is an unknown polynomial coefficient and the $X^i$ terms are elements of the polynomial basis.
+\begin{figure}
+    \includegraphics[width=\columnwidth]{example-image-a}
+    \caption{Figure showing input model NURBS parts and shape matching to each part}
+    \label{fig:shapematching}
+\end{figure}
 
-\subsection{The Shape Matching Element}
+\subsection{Shape Matching}
+\dave{Shape Matching to a single NURBS part}
 In typical FEM in graphics, we make use of tetrahedral or triangular elements with a piecewise-linear function space.
@@
 The reason for separating out the constant term becomes more evident when we examine case where we have multiple elements.
+
+\dave{here explain sampling on NURBS surface to get material points}
+The above formulation describes developing the map from a single $(u,v)$ coordinate to its corresponding position on the surface.
+In our simulation we represent the entire boundary of the NURBS model, so we require samples along each surface.
+We emphasize that we only require an initial selection of $(u,v)$ coordinates across each surface and the dynamics of the simulation modifies the set of control points $\mathbf{q}$ to modify the world space maps $\mathbf{x}(u,v)=\mathbf{J}(u,v)\mathbf{q}$.
+
+Let's say for a NURBS surface on the model the DOF column vector is $\mathbf{q}_i$, which is the subset of control points in the generalized coordinates for the $i$-th surface.
+For a given surface with $m$ pairs of $\mathbf{u}=(u,v)$ coordinates sampled in parametric space, we write the jacobian for the $i$-th surface as \myworries{(should I use something like $\hat{\mathbf{J}}$ instead?)}
+\begin{equation}
+\label{eqn:surface_jacobian}
+    \mathbf{J}_i =
+    \left[ \mathbf{J}(\mathbf{u})_1^T \dots \mathbf{J}(\mathbf{u})_m^T \right]^T
+    \text{.}
+\end{equation}
+Then if $\mathbf{x}_i$ is the set of $m$ world space positions for the $i$-th surface, we may write $\mathbf{x}_i = \mathbf{J}_i \mathbf{q}_i$.
+Finally, given the full set of generalized coordinates $\mathbf{q} = \left[ \mathbf{q}_i^T, \dots, \mathbf{q}_n^T \right]^T$ with $n$ surfaces, we may write jacobian $\mathbf{J}$ for the entire model as a block diagonal matrix:
+\begin{equation}
+\mathbf{J} = \left[ \begin{array}{ccccc}
+\mathbf{J}_1 &  &  &  &  \\
+ & \ddots &  &  &  \\
+ &  & \mathbf{J}_i & &  \\
+ &  &  & \ddots &  \\
+ &  &  &  & \mathbf{J}_n \\
+\end{array} \right]
+\text{.}
+\end{equation}
 
@@
 
-\subsection{Shape Matching with Multiple Elements}
+\subsection{Shape Matching Element Method}
+\dave{Multiple Parts}
+\dave{each polynomial/NURBS tuple is a shape matching element}
+\dave{Shape Matching Element assembles global kinematics using blendeded formultion}
  In a typical simulation we will have multiple shape matching elements where the boundary of each element is defined by a different NURBS surface.
@@
 
-\subsection{Hierarchical fitting}
+\paragraph*{Hierarchical fitting}
 A problem may emerge in the shape matching process where the system may be underdetermined.
@@
 
-\section{The Deformation Map}
-At this point we have shown how to fit a polynomial describing the deformation of each surface.
-With these polynomial coefficients coupled with their associated centers of mass, we can construct an estimate of some arbitrary material point $\mathbf{X}$ using the deformation described by the surface.
-This lends itself to a general definition for the deformed position of a point as a weighted sum of the polynomials, which yields a unique polynomial for the point.
-Thus we arrive at a definition for the deformed position, $\mathbf{x}$, of a material point $\mathbf{X}$ in the undeformed space:
-\begin{equation}
-\label{eqn:deformation_map}
-	\mathbf{x}(\mathbf{X}) = \sum_i^n w_i(\mathbf{X}) \left[ \mathbf{M}(\mathbf{X-\bar{X}}_j)\mathbf{c}_i + \mathbf{t}_j \right]
-	\text{,}
-\end{equation}
-where we have $n$ surfaces, and $w_i$ and $\mathbf{c}_i$ are the weights and the coefficients for the $i$-th surface, respectively.
-The terms $\mathbf{\bar{X}}_j$ and $\mathbf{t}_j$ the undeformed and deformed centers of mass, respectively, that are associated with the $i$-th surface.
-
-\subsection{Computing the Blending Weights}
+\subsection{Blending Weights}
+\dave{Desireable Properties}
+\dave{Construction via post normalization}
+\dave{only need blending weights at quadrature points and surface samples so given sparse set of quadrature points compute blending weights using raycasting}
+\dave{things to explain (1) raycasting works for overlapping geometric, seperated geometry (2) weights glue objects together}
+\dave{maybe but weight figure here so reviewers immediately know that it looks pretty good}
 All that remains for the deformation map is to compute blending weights.
@@
 More elaborate schemes may be developed for computing distances, such as using the Boundary Element Method, but we find this simple linear function to perform well for our purposes.
+
+\begin{figure}
+    \includegraphics[width=\columnwidth]{example-image-a}
+    \caption{Figure showing how weight calculation works (left) ray casting plus result (right) after correcting for partition of unity}
+    \label{fig:weightcompute}
+\end{figure}
 
@@
 
-\section{Selecting Centers of Mass}
-\myworries{TODO. waiting to implement this before I start writing how we do it}
-\section{Raycasting Quadrature}
+\subsection{Multiple Part Projection}
+\dave{explain why you need to share constant terms (fit polynomial around a point in space), explain clustering and how to pick, forward reference to quadrature point}
+\subsection{Quadrature Points}
+\dave{quadrature via raycasting}
 The last step before we may discuss the dynamic simulation is to describe our strategy for generating material points $\mathbf{X}$ in the undeformed model.
 Our volume integration strategy is based on \cite{KHOSRAVIFARD201030} which we briefly describe in the following section
-\subsection{Review of the Meshless Method for Domain Integral Evaluation in \cite{KHOSRAVIFARD201030}}
+
 This paper addresses the integration of volume integrals of the form $\mathbf{I} = \int_\Omega f(\mathbf{x}) d\Omega$ where $\mathbf{x} = (x_1,x_2,x_3)$.
@@
 
-\section{Dynamic Simulation}
-With our definition of the deformation map as well as a strategy for integrating the volume, we now have the necessary ingredient to perform dynamic simulation.
-We follow the Variational Mechanics approach that defines a pair of kinetic and potential energies, and we apply the Euler-Lagrange equation to develop our equations of motion.
+\begin{figure}
+    \includegraphics[width=\columnwidth]{example-image-a}
+    \caption{Figure showing how raycasting quadrature works (top), convergence plot for volume integration (bottom}
+    \label{fig:raycasting}
+\end{figure}
 
-Before we define these energies we first would like to simplify the form of the deformation map presented in equation \ref{eqn:deformation_map}. We may write $\mathbf{x}(\mathbf{X})$ in matrix form as
-\begin{equation}
-\mathbf{x}(\mathbf{X}) = \underbrace{\left[w_1\mathbf{M}(\mathbf{X - \bar{X}}_j) \dots w_n\mathbf{M}(\mathbf{X - \bar{X}}_j) \middle| \sum_i^{n_1}w_1^i\mathbf{I} \dots \sum_i^{n_m}w_m^i\mathbf{I} \right]}_{\mathbf{Y}(\mathbf{X})}
-\begin{bmatrix}
-\mathbf{\hat{c}} \\
-\mathbf{\hat{t}}
-\end{bmatrix}
-\text{,}
-\end{equation}
-\myworries{I'm not sure how to write this matrix form for x(X), especially on finding the best notation for the center of mass indexing. So this above equation is temporary.}
-where we have $n$ shape elements, $m$ centers of mass. We simplify the above equation and use the following form
-\begin{equation}
-\label{eqn:x_simple_form}
-\mathbf{x}(\mathbf{X}) = \mathbf{Y}(\mathbf{X})\mathbf{L}\hat{\mathbf{b}}
-\text{,}
-\end{equation}
-where we make use of equation \ref{eqn:c_multiple_elements} to rewrite this in terms of the boundary points $\mathbf{\hat{b}}$.
-As one final step, we rewrite this again in terms of the control points of our NURBS surfaces:
-\begin{equation}
-\label{eqn:compact_x_map}
-\mathbf{x}(\mathbf{X}) = \mathbf{Y}(\mathbf{X})\mathbf{LJq}
-\text{,}
-\end{equation}
-where we use the fact that $\mathbf{\hat{b}} = \mathbf{Jq}$.
-\myworries{not a fan of this, perhaps we change $\hat{b}$ to something else...} This gives us an expression for the deformed position in terms of our generalized coordinates, $\mathbf{q}$.
-
-\subsection{Generalized Inertia}
+\subsection{Mass Matrix}
 The kinetic energy for a model may be expressed as
@@
 where $N$ is the total number of boundary points and $\mathbf{M}_E$ is our \textit{error mass matrix}.
-\subsection{Generalized Forces}
-We write our potential energy as 
-\begin{equation}
-    V = \int_\Omega \psi(\mathbf{F(X)}) d\Omega
-\text{.}
-\end{equation}
-To evaluate this energy (as well as its derivatives), we require a definition for the deformation gradient, $\mathbf{F(X)}$:
-\begin{equation}
-        \mathbf{F(X)} = \frac{\partial \mathbf{x}}{\partial \mathbf{X}}
-               = \frac{\partial \mathbf{Y(X)}}{\partial \mathbf{X}}\mathbf{L}\hat{\mathbf{b}}
-\text{.}
-\end{equation}
-In the expression for the deformed position $\mathbf{x(X)}$ we see that the dependence on $\mathbf{X}$ only exists in $\mathbf{Y(X)}$, making $\mathbf{L}$ and $\hat{\mathbf{b}}$ constant terms when compute $\mathbf{F}$.
+% \subsection{Generalized Forces}
+% We write our potential energy as 
+% \begin{equation}
+%     V = \int_\Omega \psi(\mathbf{F(X)}) d\Omega
+% \text{.}
+% \end{equation}
+% To evaluate this energy (as well as its derivatives), we require a definition for the deformation gradient, $\mathbf{F(X)}$:
+% \begin{equation}
+%         \mathbf{F(X)} = \frac{\partial \mathbf{x}}{\partial \mathbf{X}}
+%                = \frac{\partial \mathbf{Y(X)}}{\partial \mathbf{X}}\mathbf{L}\hat{\mathbf{b}}
+% \text{.}
+% \end{equation}
+% In the expression for the deformed position $\mathbf{x(X)}$ we see that the dependence on $\mathbf{X}$ only exists in $\mathbf{Y(X)}$, making $\mathbf{L}$ and $\hat{\mathbf{b}}$ constant terms when compute $\mathbf{F}$.
 Full details on how to compute $\mathbf{F}$ and its derivative $\frac{\partial \mathbf{F}}{\partial \mathbf{q}}$ may be found in the appendix.
@@
 Now that we have an expression for the deformation gradient, we discretize our potential energy as follows
+% \begin{equation}
+%     V =  \sum_i^m \psi(\mathbf{F(X}_i) A_i
+% \text{,}
+% \end{equation}
+% where $m$ is the number of integration points and $A_i$ is the volume for the $i$-th integration point.
+
+% Like in the case of the kinetic energy, we also introduce an error force potential, which takes the simple form
+% \begin{equation}
+% V_E = \mathbf{q}^T \mathbf{M}_E \mathbf{q}
+% \end{equation}
+\subsection{Deformation Gradient}
+At this point we have shown how to fit a polynomial describing the deformation of each surface.
+With these polynomial coefficients coupled with their associated centers of mass, we can construct an estimate of some arbitrary material point $\mathbf{X}$ using the deformation described by the surface.
+This lends itself to a general definition for the deformed position of a point as a weighted sum of the polynomials, which yields a unique polynomial for the point.
+Thus we arrive at a definition for the deformed position, $\mathbf{x}$, of a material point $\mathbf{X}$ in the undeformed space:
 \begin{equation}
-    V =  \sum_i^m \psi(\mathbf{F(X}_i) A_i
+\label{eqn:deformation_map}
+	\mathbf{x}(\mathbf{X}) = \sum_i^n w_i(\mathbf{X}) \left[ \mathbf{M}(\mathbf{X-\bar{X}}_j)\mathbf{c}_i + \mathbf{t}_j \right]
 	\text{,}
 \end{equation}
-where $m$ is the number of integration points and $A_i$ is the volume for the $i$-th integration point.
+where we have $n$ surfaces, and $w_i$ and $\mathbf{c}_i$ are the weights and the coefficients for the $i$-th surface, respectively.
+The terms $\mathbf{\bar{X}}_j$ and $\mathbf{t}_j$ the undeformed and deformed centers of mass, respectively, that are associated with the $i$-th surface.
 
-Like in the case of the kinetic energy, we also introduce an error force potential, which takes the simple form
+\subsection{External Forces}
+\subsection{Time Integration}
+With our definition of the deformation map as well as a strategy for integrating the volume, we now have the necessary ingredient to perform dynamic simulation.
+We follow the Variational Mechanics approach that defines a pair of kinetic and potential energies, and we apply the Euler-Lagrange equation to develop our equations of motion.
+
+Before we define these energies we first would like to simplify the form of the deformation map presented in equation \ref{eqn:deformation_map}. We may write $\mathbf{x}(\mathbf{X})$ in matrix form as
 \begin{equation}
-V_E = \mathbf{q}^T \mathbf{M}_E \mathbf{q}
+\mathbf{x}(\mathbf{X}) = \underbrace{\left[w_1\mathbf{M}(\mathbf{X - \bar{X}}_j) \dots w_n\mathbf{M}(\mathbf{X - \bar{X}}_j) \middle| \sum_i^{n_1}w_1^i\mathbf{I} \dots \sum_i^{n_m}w_m^i\mathbf{I} \right]}_{\mathbf{Y}(\mathbf{X})}
+\begin{bmatrix}
+\mathbf{\hat{c}} \\
+\mathbf{\hat{t}}
+\end{bmatrix}
+\text{,}
 \end{equation}
+\myworries{I'm not sure how to write this matrix form for x(X), especially on finding the best notation for the center of mass indexing. So this above equation is temporary.}
+where we have $n$ shape elements, $m$ centers of mass. We simplify the above equation and use the following form
+\begin{equation}
+\label{eqn:x_simple_form}
+\mathbf{x}(\mathbf{X}) = \mathbf{Y}(\mathbf{X})\mathbf{L}\hat{\mathbf{b}}
+\text{,}
+\end{equation}
+where we make use of equation \ref{eqn:c_multiple_elements} to rewrite this in terms of the boundary points $\mathbf{\hat{b}}$.
+As one final step, we rewrite this again in terms of the control points of our NURBS surfaces:
+\begin{equation}
+\label{eqn:compact_x_map}
+\mathbf{x}(\mathbf{X}) = \mathbf{Y}(\mathbf{X})\mathbf{LJq}
+\text{,}
+\end{equation}
+where we use the fact that $\mathbf{\hat{b}} = \mathbf{Jq}$.
+\myworries{not a fan of this, perhaps we change $\hat{b}$ to something else...} This gives us an expression for the deformed position in terms of our generalized coordinates, $\mathbf{q}$.
 
-\subsection{Time Integration}
+\dave{discrete Equations of motion with quadratured integrals, reference/details on how we handle collisions}
+
 We have now provided all the necessary ingredient to produce equations of motion for deformable bodies, which we derive via Euler-Lagrange equation. We begin with the Lagrangian
@@
 
-%\begin{equation}
-%    \frac{\partial L}{\partial \mathbf{q}} = \sum_i^m \frac{\partial V_i}{\partial \mathbf{q}} = \sum_i^m \frac{\partial \psi(\mathbf{F_i)}}{\partial \mathbf{F}}\frac{\partial \mathbf{F_i}}{\partial \mathbf{q}}A_i
-%\end{equation}
-%\
+\begin{equation}
+   \frac{\partial L}{\partial \mathbf{q}} = \sum_i^m \frac{\partial V_i}{\partial \mathbf{q}} = \sum_i^m \frac{\partial \psi(\mathbf{F_i)}}{\partial \mathbf{F}}\frac{\partial \mathbf{F_i}}{\partial \mathbf{q}}A_i
+\end{equation}
+\
+
+%%% ------- OLD STUFF BELOW --------
+% \section{Methods}
+% \dave{We should probably put some pseudo code for the preprocessing and the runtime algorithms right up fron, along with a quick textual walk through of the method.}
+% Given a volumetric model composed of one or more NURBS surfaces, the surface of each surface is reconstructed using a set of control points $\textbf{p}$ and their associated weights $w$.
+Following D-NURBS \cite{10.1145/176579.176580}, we take our degrees of freedom (DOF) to be control points of the surfaces and produce displacements directly on these control points at each simulation step.
+We note that we used a simplified version of D-NURBS in which the weights are held constant throughout the simulation and find this not to be limiting \myworries{(feel like we'll need better justification than saying just this?)}.
+We represent the degrees of freedom for the $i$-th NURBS surface with $m$ control points as $\mathbf{q}_i = \left[ \mathbf{p}_1^T, \dots, \mathbf{p}_m^T \right]^T \in \mathbb{R}^{3m}$.
+From this definition, we write the DOF for the entire model as $\mathbf{q} = \left[ \mathbf{q}_i^T, \dots, \mathbf{q}_n^T \right]^T$ where $n$ is the number of NURBS surfaces for the model.
+
+% To evaluate the volumetric integrals in solid mechanics models, we require material points $\mathbf{X} \in \mathbb{R}^{3m}$ in the undeformed space that serve as integration points.
+Our shape matching element method enables us to evaluate the deformed positions of the material points strictly using positions defined on the boundary.
+Like in VEM, we fit a polynomial to each boundary primitive (NURBS in this case) describing its deformation.
+The fitting of these polynomials closely follows that of the Shape Matching \cite{10.1145/1073204.1073216} approach where we perform a least squares polynomial fit to the deformed boundary points.
+These boundary points are represented by sampled positions on each NURBS surface.
+Since each surface is associated with a single polynomial, we can construct a unique polynomial for some arbitrary point by blending these polynomials.
+Thus given some material point $\mathbf{X}_i$ we find it's corresponding deformed position $\mathbf{x}_i$ via a weighted combination of the polynomials.
+This permits us to build a general deformation map which we use to construct our equations of motion.
+
+% In the following sections, we will first review the simplified D-NURBS model used to form our generalized coordinates.
+We then present our shape matching algorithm for fitting polynomials to surfaces ....
+In section ...
+we next describe how to blend these surface polynomials to construct a general deformation map for material points.
+Section ...
+describes our strategy for selecting integration points in the volume through a raycasting approach.
+Finally, we formulate our equations of motions from a Lagrangian formulation.
+
+% \section{D-NURBS Model}
+% \myworries{TODO: Most of these matrices should be paranthesis matrices}
+% We provide a brief overview of the D-NURBS formulation presented in \cite{10.1145/176579.176580}. Our models in our method are composed of a collection of NURBS surfaces.
+
+% \subsection{D-NURBS Formulation}
+% A point on a NURBS surface is typically expressed in the following form
+% \begin{equation}
+% \label{eqn:nurbs_srf}
+%     \mathbf{x}(u,v) = \frac{\sum_{i=1}^{n}\sum_{j=1}^{m}  \mathbf{p}_{i,j} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
+%     {\sum_{i=1}^{n}\sum_{j=1}^{m} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
+%     \text{,}
+% \end{equation}
+% where where have a total of $nm$ control points $\mathbf{p}$ and weights $w$.
+Using the B-spline basis functions and the weighted control points, this formula gives a map from our parametric coordinates $u,v$ to their corresponding position on the surface in $\mathbb{R}^{3m}$.
+$B_{i,k}(u)$ is the B-spline basis function for the $i-th$ control point with degree $k-1$.
+
+% We would like our generalized coordinates to be the control points of the model so that the dynamics operates directly on the original representation.
+Therefore, we would like to rewrite equation \ref{eqn:nurbs_srf} in the form
+% \begin{equation}
+%     \mathbf{x}(u,v) = \mathbf{J}(u,v)\mathbf{q}
+%     \text{,}
+% \end{equation}
+% where $\mathbf{q} = \left[ \mathbf{p}_{1,1}^T, \dots, \mathbf{p}_{i,j}^T, \dots, \mathbf{p}_{n,m}^T \right]^T \in \mathbb{R}^{3nm}$ is the set of $nm$ control points \myworries{in the overview I just write $n$ control points, but here i'm writing $nm$} for a single NURBS surface arranged as a column vector.
+$\mathbf{J} \in  \mathbb{R}^{3 x 3nm}$ is the NURBS jacobian that maps $(u,v)$ coordinates to world positions in $\mathbb{R}^{3}$.
+
+% The NURBS jacobian, $\mathbf{J}$, is a matrix composed of horizontally concatenated $3x3$ blocks where the $(i,j)$-th block is $\frac{\partial \mathbf{x}}{\partial \mathbf{p}_{i,j}}$ for the $(i,j)$-th control point.
+$\frac{\partial \mathbf{x}}{\partial \mathbf{p}_{i,j}}$ is a diagonal $3x3$ matrix where each diagonal entry $N_{i,j}(\mathbf{u})$ takes the form
+
+% \begin{equation}
+% \label{eqn:jacobian_diagonal}
+%     N_{i,j}(u,v)
+%     = \frac{w_{i,j} B_{i,k}(u)B_{j,l}(v)}{\sum_{i=1}^{n}\sum_{j=1}^{m} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
+%     \text{.}
+% \end{equation}
+% Thus the NURBS jacobian for a single $(u,v)$ pair is written as
+% \begin{equation}
+% \label{eqn:uv_jacobian}
+%     \mathbf{J}(u,v) =
+%     \left[ \frac{\partial \mathbf{x}}{\partial \mathbf{p}_{1,1}} \dots
+%            \frac{\partial \mathbf{x}}{\partial \mathbf{p}_{i,j}} \dots 
+%            \frac{\partial \mathbf{x}}{\partial \mathbf{p}_{n,m}}
+%     \right]
+%     \text{.}
+% \end{equation}
+
+% This formulation is a simplification of the full D-NURBS described in \cite{10.1145/176579.176580} the weights of the NURBS are fixed throughout the simulation.
+Permitting the dynamics to modify weights would likely improve the expressiveness of the kinematics, but we find fixing the weights affords sufficiently expressive results and improves performance.
+The performance improvement is a result of having fixed $N_{i,j}(u,v)$ entries over the course of the simulation.
+The full D-NURBS model requires rebuilding the jacobian and consequently the mass matrices at each time step.
+
+% \subsection{Generalized Coordinates for a NURBS model}
+% The above formulation describes developing the map from a single $(u,v)$ coordinate to its corresponding position on the surface.
+In our simulation we represent the entire boundary of the NURBS model, so we require samples along each surface.
+We emphasize that we only require an initial selection of $(u,v)$ coordinates across each surface and the dynamics of the simulation modifies the set of control points $\mathbf{q}$ to modify the world space maps $\mathbf{x}(u,v)=\mathbf{J}(u,v)\mathbf{q}$.
+
+% Let's say for a NURBS surface on the model the DOF column vector is $\mathbf{q}_i$, which is the subset of control points in the generalized coordinates for the $i$-th surface.
+For a given surface with $m$ pairs of $\mathbf{u}=(u,v)$ coordinates sampled in parametric space, we write the jacobian for the $i$-th surface as \myworries{(should I use something like $\hat{\mathbf{J}}$ instead?)}
+% \begin{equation}
+% \label{eqn:surface_jacobian}
+%     \mathbf{J}_i =
+%     \left[ \mathbf{J}(\mathbf{u})_1^T \dots \mathbf{J}(\mathbf{u})_m^T \right]^T
+%     \text{.}
+% \end{equation}
+% Then if $\mathbf{x}_i$ is the set of $m$ world space positions for the $i$-th surface, we may write $\mathbf{x}_i = \mathbf{J}_i \mathbf{q}_i$.
+Finally, given the full set of generalized coordinates $\mathbf{q} = \left[ \mathbf{q}_i^T, \dots, \mathbf{q}_n^T \right]^T$ with $n$ surfaces, we may write jacobian $\mathbf{J}$ for the entire model as a block diagonal matrix:
+% \begin{equation}
+% \mathbf{J} = \left[ \begin{array}{ccccc}
+% \mathbf{J}_1 &  &  &  &  \\
+%  & \ddots &  &  &  \\
+%  &  & \mathbf{J}_i & &  \\
+%  &  &  & \ddots &  \\
+%  &  &  &  & \mathbf{J}_n \\
+% \end{array} \right]
+% \text{.}
+% \end{equation}
+% With this equation, the full set of world positions of the NURBS model may be written as $\mathbf{x} = \mathbf{J}\mathbf{q} \in \mathbb{R}^{3N}$ where $N$ is the total number of $(u,v)$ coordinates.
+
+% \subsection{Shape Matching with Multiple Elements}
+
+% \subsection{Computing the Blending Weights}
+
+% \subsection{Generalized Inertia}
+
+% \subsection{Time Integration}
+%

### 2021-01-20 cpp/include/optimization_gradient_based.h
make L sparse by default to fit the mex input

@@ -100,12 +100,6 @@ namespace sim {
                 exit(1);
             }
 
-            // if(!search_direction_newton_sparse(tmp_d, tmp_g, tmp_H, solver))
-            // {
-            //     std::cout<<"Computing search direction failed in newtons method \n";
-            //     exit(1);
-            // }
-
             //line search 
             SolverExitStatus ls_status = linesearch_backtracking_bisection(x, tmp_d, f, tmp_g, 
                                                                            max_iterations_ls, 

### 2021-01-20 cpp/include/vem3dmesh_neohookean_dq.h

@@ -18,7 +18,7 @@ void vem3dmesh_neohookean_dq(Eigen::VectorXx<DerivedRet> &g,
 	const std::vector<Eigen::MatrixXd, Eigen::aligned_allocator<Eigen::MatrixXd> > & dF_dc,
 	const std::vector<Eigen::SparseMatrix<double>, Eigen::aligned_allocator<Eigen::SparseMatrix<double>> > & dF_dc_S,
 	const Eigen::MatrixXd &ME,
-	const Eigen::MatrixXd &L,
+	const Eigen::SparseMatrix<double> &L,
 	int k, int n, int d, int x0_coms_size, double k_stability);
 }
 

### 2021-01-20 cpp/include/vem3dmesh_polynomial_coefficients.h

@@ -7,10 +7,10 @@
 
 namespace sim {
 
-template<typename DerivedRet, typename DerivedX, typename DerivedL>
+template<typename DerivedRet, typename DerivedX>
 void vem3dmesh_polynomial_coefficients(Eigen::VectorXx<DerivedRet> &c,
 	const Eigen::MatrixXx<DerivedX> &x,
-	const Eigen::MatrixXx<DerivedL> &L,
+	const Eigen::SparseMatrix<double> &L,
 	const std::vector<Eigen::VectorXi, Eigen::aligned_allocator<Eigen::VectorXi> > & E);
 }
 

### 2021-01-20 cpp/include/vem3dmesh_simulate_one_step.h

@@ -29,7 +29,7 @@ void vem3dmesh_simulate_one_step(Eigen::VectorXx<DerivedRet> &qdot_new,
 	const std::vector<Eigen::VectorXi, Eigen::aligned_allocator<Eigen::VectorXi> > & E,
 	const Eigen::MatrixXd &M,
 	const Eigen::MatrixXd &ME,
-	const Eigen::MatrixXd &L,
+	const Eigen::SparseMatrix<double> &L,
 	const Eigen::SparseMatrix<double> &P,
 	const Eigen::SparseMatrix<double> &J,
 	int k, int n, int d, int x0_coms_size, double k_stability, double dt);

### 2021-01-20 cpp/src/vem3dmesh_neohookean_dq.cpp

@@ -15,7 +15,7 @@ void sim::vem3dmesh_neohookean_dq(Eigen::VectorXx<DerivedRet> &g,
 	const std::vector<Eigen::MatrixXd, Eigen::aligned_allocator<Eigen::MatrixXd> > & dF_dc,
 	const std::vector<Eigen::SparseMatrix<double>, Eigen::aligned_allocator<Eigen::SparseMatrix<double>> > & dF_dc_S,
 	const Eigen::MatrixXd &ME,
-	const Eigen::MatrixXd &L,
+	const Eigen::SparseMatrix<double> &L,
 	int k, int n, int d, int x0_coms_size, double k_stability) {
 
 	// std::cout << "hello "<< std::endl;

### 2021-01-20 cpp/src/vem3dmesh_polynomial_coefficients.cpp

@@ -6,26 +6,34 @@
 #include <omp.h>
 #endif
 
-template<typename DerivedRet, typename DerivedX, typename DerivedL>
+template<typename DerivedRet, typename DerivedX>
 void sim::vem3dmesh_polynomial_coefficients(Eigen::VectorXx<DerivedRet> &c,
 	const Eigen::MatrixXx<DerivedX> &x,
-	const Eigen::MatrixXx<DerivedL> &L,
-	const std::vector<Eigen::VectorXi, Eigen::aligned_allocator<Eigen::VectorXi> > & E) {
+	const Eigen::SparseMatrix<double> &L,
+	const std::vector<Eigen::VectorXi, Eigen::aligned_allocator<Eigen::VectorXi> > & E){
 
+		// std::cout << "start polynomial_coefficients" << std::endl;
 		Eigen::MatrixXd B(x.rows(), x.cols());
 		B.setZero();
 		int col_start = 0;
 		Eigen::MatrixXd x_slice;
+		// std::cout << "1 polynomial_coefficients" << std::endl;
 		for (int i = 0; i < E.size(); i++) {
 			igl::slice(x, E[i], 2, x_slice);
+			// std::cout << "2 polynomial_coefficients" << std::endl;
 			B.block(0, col_start, 3, E[i].rows()) = x_slice;
+			// std::cout << "3 polynomial_coefficients" << std::endl;
 
 			col_start += E[i].rows();
+			// std::cout << "4 polynomial_coefficients" << std::endl;
 		}
+		// std::cout << "5 polynomial_coefficients" << std::endl;
 		Eigen::VectorXd b = Eigen::Map<const Eigen::VectorXd>(B.data(), B.size());
+		// std::cout << "6 polynomial_coefficients" << std::endl;
 		
 		// Solve for polynomial coefficients (projection operators).
 		c = L * b;
+		// std::cout << "pass polynomial_coefficients" << std::endl;
 }
 
 #include <iostream>

### 2021-01-20 cpp/src/vem3dmesh_simulate_one_step.cpp

@@ -20,7 +20,7 @@ void sim::vem3dmesh_simulate_one_step(Eigen::VectorXx<DerivedRet> &qdot_new,
 	const std::vector<Eigen::VectorXi, Eigen::aligned_allocator<Eigen::VectorXi> > & E,
 	const Eigen::MatrixXd &M,
 	const Eigen::MatrixXd &ME,
-	const Eigen::MatrixXd &L,
+	const Eigen::SparseMatrix<double> &L,
 	const Eigen::SparseMatrix<double> &P,
 	const Eigen::SparseMatrix<double> &J,
 	int k, int n, int d, int x0_coms_size, double k_stability, double dt) {
@@ -115,7 +115,7 @@ void sim::vem3dmesh_simulate_one_step(Eigen::VectorXx<DerivedRet> &qdot_new,
 	Eigen::VectorXd tmp_d(qdot.rows());
 
 	qdot_new = qdot;
-	newtons_method_bisection(qdot_new, value, gradient, hessian, tmp_g, tmp_H, tmp_d, 1e-3, 5, 5);
+	newtons_method_bisection(qdot_new, value, gradient, hessian, tmp_g, tmp_H, tmp_d, 1e-3, 1, 5);
 }
 
 #include <iostream>

### 2021-01-20 examples/linear_mug.m

@@ -21,6 +21,7 @@ options.lambda = lambda;
 options.mu = mu;
 options.sample_interior = 0; % only sample on boundary
 options.distance_cutoff = 1;
+options.fitting_mode = 'hierarchical';
 vem_simulate_nurbs(part, options);  %%%%% verify hierarchical works!!!!!
                                                         
  

### 2021-01-20 examples/linear_mug_implicit.m

@@ -21,6 +21,7 @@ options.lambda = lambda;
 options.mu = mu;
 options.sample_interior = 0; % only sample on boundary
 options.distance_cutoff = 1;
+% options.fitting_mode = 'hierarchical';
 vem_simulate_nurbs_newtons(part, options);  %%%%% verify hierarchical works!!!!!
                                                         
  

### 2021-01-20 matlab/mex/vem3dmesh_neohookean_dq.cpp

@@ -75,8 +75,10 @@ void mexFunction(int nlhs, mxArray *plhs[],
     /* variable declarations here */
     Eigen::VectorXd g;
 
-    Eigen::MatrixXd x, params, ME, L;
+    Eigen::MatrixXd x, params, ME;
     Eigen::VectorXd c, volumes; 
+
+		Eigen::SparseMatrix<double> L;
 
 		std::vector<Eigen::MatrixXd, Eigen::aligned_allocator<Eigen::MatrixXd> > dF_dc;
 		std::vector<Eigen::SparseMatrix<double>, Eigen::aligned_allocator<Eigen::SparseMatrix<double>>> dF_dc_S;
@@ -88,7 +90,7 @@ void mexFunction(int nlhs, mxArray *plhs[],
 		parse_cell(prhs, 4, dF_dc);
 		parse_cell_sparse(prhs, 5, dF_dc_S);
 		igl::matlab::parse_rhs_double(prhs+6, ME);
-		igl::matlab::parse_rhs_double(prhs+7, L);
+		igl::matlab::parse_rhs(prhs+7, L);
 
 		int k = (int)*mxGetPr(prhs[8]);
 		int n = (int)*mxGetPr(prhs[9]);

### 2021-01-20 matlab/mex/vem3dmesh_polynomial_coefficients.cpp

@@ -74,15 +74,24 @@ void mexFunction(int nlhs, mxArray *plhs[],
                  int nrhs, const mxArray *prhs[]) {
     /* variable declarations here */
     Eigen::VectorXd c;
-    Eigen::MatrixXd x, L;
+    Eigen::MatrixXd x;
+		Eigen::SparseMatrix<double> L;
 
   	std::vector<Eigen::VectorXi, Eigen::aligned_allocator<Eigen::VectorXi> > E;
+		
+		// std::cout << "1 mex polynomial_coefficients" << std::endl;
 
 		igl::matlab::parse_rhs_double(prhs+0, x);
-    igl::matlab::parse_rhs_double(prhs+1, L);
+		// std::cout << "2 mex polynomial_coefficients" << std::endl;
+    igl::matlab::parse_rhs(prhs+1, L);
+		// std::cout << "3 mex polynomial_coefficients" << std::endl;
 		parse_cell_index(prhs, 2, E);
+		// std::cout << "4 mex polynomial_coefficients" << std::endl;
 
 		sim::vem3dmesh_polynomial_coefficients(c, x, L, E);
 
+		// std::cout << "5 mex polynomial_coefficients" << std::endl;
+
 		igl::matlab::prepare_lhs_double(c, plhs);
+		// std::cout << "6 mex polynomial_coefficients" << std::endl;
 }

### 2021-01-20 matlab/mex/vem3dmesh_simulate_one_step.cpp

@@ -79,10 +79,10 @@ void mexFunction(int nlhs, mxArray *plhs[],
     /* variable declarations here */
     Eigen::VectorXd qdot_new;
 
-    Eigen::MatrixXd x_fixed, params, M, ME, L;
+    Eigen::MatrixXd x_fixed, params, M, ME;
     Eigen::VectorXd q, qdot, f_ext, volume, volumes; 
 
-		Eigen::SparseMatrix<double> P, J;
+		Eigen::SparseMatrix<double> P, J, L;
 
 		std::vector<Eigen::MatrixXd, Eigen::aligned_allocator<Eigen::MatrixXd> > dF_dc;
 		std::vector<Eigen::SparseMatrix<double>, Eigen::aligned_allocator<Eigen::SparseMatrix<double>>> dF_dc_S;
@@ -110,7 +110,7 @@ void mexFunction(int nlhs, mxArray *plhs[],
 		// std::cout << "3" << std::endl;
 		igl::matlab::parse_rhs_double(prhs+10, M);
 		igl::matlab::parse_rhs_double(prhs+11, ME);
-		igl::matlab::parse_rhs_double(prhs+12, L);
+		igl::matlab::parse_rhs(prhs+12, L);
 		igl::matlab::parse_rhs(prhs+13, P);
 		igl::matlab::parse_rhs(prhs+14, J);
 		// std::cout << "4" << std::endl;

### 2021-01-20 matlab/vem_simulate_nurbs_fmincon.m

@@ -131,7 +131,17 @@ function vem_simulate_nurbs_fmincon(parts, varargin)
     ii=1;
     for t=0:config.dt:30
         tic
+        
+        % Preparing input for stiffness matrix mex function.
+        b = [];
+        for i=1:numel(E)
+            b = [b (x(:,E{i}))];
+        end
+        b = b(:);
 
+        % Solve for polynomial coefficients (projection operators).
+        c = L * b;
+        
         % Solve for polynomial coefficients (projection operators).
         c = vem3dmesh_polynomial_coefficients(x, L, E);
         
@@ -185,21 +195,21 @@ function [e, g, H] = energy(qdot_new)
     q_new = q + config.dt*qdot_new;
     x_new = reshape(P'*J*q_new,3,[]) + x_fixed;
     dt = config.dt;
-    
+
     % Solve for polynomial coefficients (projection operators).
     c = vem3dmesh_polynomial_coefficients(x_new, L, E);
-    
+
     neohookean_e =  vem3dmesh_neohookean_q(c, vol, params, dF_dc, dF_dc_S, d);
-     
+
     e = 0.5*qdot_new'*J'*PMP*J*qdot_new - qdot_new'*J'*PMP*J*qdot + ...
         config.k_stability * x_new(:)' * ME * x_new(:) + ...
         0.5 * neohookean_e - ...
-        qdot_new'*J'*f_gravity; 
+        qdot_new'*J'*f_gravity;
 
     if nargout > 1
         g_neohookean = vem3dmesh_neohookean_dq(x_new, c, vol, params, dF_dc, dF_dc_S, ME, L, ...
                                        k, n, d, size(x0_coms,2), config.k_stability);
-                               
+                    
         g = J' * (PMP*J*(qdot_new - qdot) + ...
             + dt*P*g_neohookean + ...
             - f_gravity);

### 2021-01-20 matlab/vem_simulate_nurbs_newtons.m

@@ -1,4 +1,4 @@
-function vem_simulate_nurbs_fmincon_param(parts, varargin)
+function vem_simulate_nurbs_newtons(parts, varargin)
     % Simulation parameter parsing
     p = inputParser;
     addParameter(p, 'dt', 0.01);                % timestep

### 2021-01-20 paper/sections/results.tex
figures
fix image dimensions

@@
 \dave{we should show the nurbs model in rhino/fusion for every example, maybe an exploded view as well}
+
+\dave{use as many different meshes as possible for figures (i.e try not to have the same mesh show up in more than one figure if possible).}
 
@@
 
-%different material parameters / material models 
+%different surface reps
 \begin{figure}
   \includegraphics[width=\columnwidth]{example-image-a}
-  \caption{Big image simulating the rocket with all different combinations of things}
-  \label{fig:materials}
+  \caption{show subd simulation if we have it. If it works really well we'll just mix subds and nurbs throughout the paper. }
+  \label{fig:subd}
+\end{figure}
+
+%editing example
+\begin{figure}
+  \includegraphics[width=\columnwidth]{example-image-a}
+  \caption{show input to simulation, show output then show output loaded in rhino/fusion 360 }
+  \label{fig:edit}
 \end{figure}
@@
 \begin{figure}
-  \includegraphics[width=\textwidth]{example-image-a}
+  \includegraphics[width=\columnwidth]{example-image-a}
   \caption{Avocado to show heterogenous materials}

### 2021-01-20 matlab/vem_simulate_nurbs_newtons.m
update comments

@@ -135,14 +135,7 @@ function vem_simulate_nurbs_newtons(parts, varargin)
         % Solve for polynomial coefficients (projection operators).
         c = vem3dmesh_polynomial_coefficients(x, L, E);
         
-%         energy = @(qdot_new) vem3dmesh_energy_matlab(qdot_new, q, qdot, f_gravity, x_fixed, ...
-%                                               vol, params, dF_dc, dF_dc_S, w_I, E, ...
-%                                               M, ME, L, P, J, ...
-%                                               k, n, d, size(x0_coms,2), config.k_stability, config.dt);
-%     
-%         % solve for the velocity of the next time step
-%         qdot = fmincon(energy, 0*qdot, [], [], [],[], [],[], [], options);
-
+        % simulate one timestep using newton's method
         qdot = vem3dmesh_simulate_one_step(q, qdot, f_gravity, x_fixed, ...
                                               vol, params, dF_dc, dF_dc_S, w_I, E, ...
                                               M, ME, L, P, J, ...

### 2021-01-20 paper/sections/methods.tex
typo
VEM note
dave note

@@
      }
-     \caption{SEM Algorithm}
+     \caption{Shape Matching Element Algorithm}
     \end{algorithm}
@@
 In matrix form we have
+\dave{need to explain that system can be degenerate fix is to select a constant term}
 \begin{equation}
@@
 Like in the original Shape Matching \cite{10.1145/1073204.1073216}, we solve for the coefficients by minimizing the error in a least squares sense, resulting in
+\dave{no b just use J q, here's a good place to explain sampling the surface to build big J.}
 \begin{equation}
@@
 
+\dave{Shape Matching is VEM brief explanation here. Point of departure, one polynmoial isn't going to cut it for a whole object.}
 \subsection{Shape Matching Element Method}

### 2021-01-20 paper/sections/methods.tex
no more main pdf
note

@@
 
+\dave{weights are missing, just matching to points leaves us vulnerable to biasing towards clusters of points}
 The remaining step in defining a single element is to describe how to solve for the polynomial coefficients.
@@
 
+\dave{this is good but we shouldn't imply we solve the system above since we do Hierarchical ordering}
 \paragraph*{Hierarchical fitting}

### 2021-01-20 paper/sections/methods.tex
methods nurbs description

@@
 
-\dave{explain NURBS but not D-NURBS because that's Terzopolous's specific algorithm just explain the NURBS parameterization}
+Our algorithm acts on objects composed of multiple NURBS surfaces~\reffig{NURBS}.
 
-A point on a NURBS surface is typically expressed in the following form
+\begin{figure}
+    \includegraphics[width=\columnwidth]{example-image-a}
+    \caption{Figure showing simple NURBS patch with variables labelled}
+    \label{fig:NURBS}
+\end{figure}
+
+The three-dimensional position, $\defX$, of any point on a NURBS surface can be written as 
 \begin{equation}
 \label{eqn:nurbs_srf}
-    \mathbf{x}(u,v) = \frac{\sum_{i=1}^{n}\sum_{j=1}^{m}  \mathbf{p}_{i,j} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
-    {\sum_{i=1}^{n}\sum_{j=1}^{m} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
-    \text{,}
+\defX\left(u,v\right) = \sum_{i=1}^n\sum_{j=1}^n \phi_{ij}\left(u,v\right)\vc{q}_{ij},
+    %\mathbf{x}(u,v) = \frac{\sum_{i=1}^{n}\sum_{j=1}^{m}  \mathbf{p}_{i,j} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
+    %{\sum_{i=1}^{n}\sum_{j=1}^{m} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
+    %\text{,}
 \end{equation}
-where where have a total of $nm$ control points $\mathbf{p}$ and weights $w$.
-Using the B-spline basis functions and the weighted control points, this formula gives a map from our parametric coordinates $u,v$ to their corresponding position on the surface in $\mathbb{R}^{3m}$.
-$B_{i,k}(u)$ is the B-spline basis function for the $i-th$ control point with degree $k-1$.
+where $u$ and $v$ $\in \real$ are the coordinates of the 2D parametric domain, $n$ and $m$ are the number of control points in the $u$ and $v$ directions, $\mathbf{q}_{ij}\in\real^3$ are the control points and $\phi_{ij}\left(u,v\right)\in\real$ are the NURBS
+basis functions given by 
+\begin{equation*}
+    \phi_{ij}\left(u,v\right) = \frac{w_{i,j}B_{ik}(u)B_{jl}(v)}{\sum_{r=1}^{n}\sum_{s=1}^{m} w_{ij} B_{ik}(u)B_{jl}(v)},
+\end{equation*} where $B_{ik}$ (resp $B_{jl}$) is the $i^{th}$ ($j^{th}$) b-spline basis of order $k$ ($l$).
 
-We would like our generalized coordinates to be the control points of the model so that the dynamics operates directly on the original representation.
-Therefore, we would like to rewrite equation \ref{eqn:nurbs_srf} in the form
+Exploiting linearity with respect to the control points  we can express this mapping as 
 \begin{equation}
-    \mathbf{x}(u,v) = \mathbf{J}(u,v)\mathbf{q}
-    \text{,}
-\end{equation}
-where $\mathbf{q} = \left[ \mathbf{p}_{1,1}^T, \dots, \mathbf{p}_{i,j}^T, \dots, \mathbf{p}_{n,m}^T \right]^T \in \mathbb{R}^{3nm}$ is the set of $nm$ control points \myworries{in the overview I just write $n$ control points, but here i'm writing $nm$} for a single NURBS surface arranged as a column vector.
-$\mathbf{J} \in  \mathbb{R}^{3 x 3nm}$ is the NURBS jacobian that maps $(u,v)$ coordinates to world positions in $\mathbb{R}^{3}$.
-
-The NURBS jacobian, $\mathbf{J}$, is a matrix composed of horizontally concatenated $3x3$ blocks where the $(i,j)$-th block is $\frac{\partial \mathbf{x}}{\partial \mathbf{p}_{i,j}}$ for the $(i,j)$-th control point.
+    \defX\left(u,v\right) = \underbrace{\begin{pmatrix} \phi_{11}\left(u,v\right)\ident & \dots & \phi_{nn}\left(u,v\right)\ident \end{pmatrix}}_{\Jnurbs\left(u,v\right)}
+    \underbrace{\begin{pmatrix} \vc{q}_{11} \\ \vdots \\ \vc{q}_{nn}\end{pmatrix}}_{\vc{q}},
+\end{equation} where $\Jnurbs\left(u,v\right)$ is the NURBS Jacobian, $\ident$ is the $3\times3$ identity matrix and $\vc{q}$ is the vector of generalized coordinates~\dave{cite Lanczos}.
+The Shape Matching Element Method will follow from this kinematic description.
+% The NURBS jacobian, $\mathbf{J}$, is a matrix composed of horizontally concatenated $3x3$ blocks where the $(i,j)$-th block is $\frac{\partial \mathbf{x}}{\partial \mathbf{p}_{i,j}}$ for the $(i,j)$-th control point.
 $\frac{\partial \mathbf{x}}{\partial \mathbf{p}_{i,j}}$ is a diagonal $3x3$ matrix where each diagonal entry $N_{i,j}(\mathbf{u})$ takes the form
 
-\begin{equation}
-\label{eqn:jacobian_diagonal}
-    N_{i,j}(u,v)
-    = \frac{w_{i,j} B_{i,k}(u)B_{j,l}(v)}{\sum_{i=1}^{n}\sum_{j=1}^{m} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
-    \text{.}
-\end{equation}
-Thus the NURBS jacobian for a single $(u,v)$ pair is written as
-\begin{equation}
-\label{eqn:uv_jacobian}
-    \mathbf{J}(u,v) =
-    \left[ \frac{\partial \mathbf{x}}{\partial \mathbf{p}_{1,1}} \dots
-           \frac{\partial \mathbf{x}}{\partial \mathbf{p}_{i,j}} \dots 
-           \frac{\partial \mathbf{x}}{\partial \mathbf{p}_{n,m}}
-    \right]
-    \text{.}
-\end{equation}
+% \begin{equation}
+% \label{eqn:jacobian_diagonal}
+%     N_{i,j}(u,v)
+%     = \frac{w_{i,j} B_{i,k}(u)B_{j,l}(v)}{\sum_{i=1}^{n}\sum_{j=1}^{m} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
+%     \text{.}
+% \end{equation}
+% Thus the NURBS jacobian for a single $(u,v)$ pair is written as
+% \begin{equation}
+% \label{eqn:uv_jacobian}
+%     \mathbf{J}(u,v) =
+%     \left[ \frac{\partial \mathbf{x}}{\partial \mathbf{p}_{1,1}} \dots
+%            \frac{\partial \mathbf{x}}{\partial \mathbf{p}_{i,j}} \dots 
+%            \frac{\partial \mathbf{x}}{\partial \mathbf{p}_{n,m}}
+%     \right]
+%     \text{.}
+% \end{equation}
 
-With this equation, the full set of world positions of the NURBS model may be written as $\mathbf{x} = \mathbf{J}\mathbf{q} \in \mathbb{R}^{3N}$ where $N$ is the total number of $(u,v)$ coordinates.
+% With this equation, the full set of world positions of the NURBS model may be written as $\mathbf{x} = \mathbf{J}\mathbf{q} \in \mathbb{R}^{3N}$ where $N$ is the total number of $(u,v)$ coordinates.
 

### 2021-01-20 paper/sections/methods.tex
working on shape matching bit

@@
 The input to the Shape Matching Element Method (SEM)  is a NURBS \emph{boundary representation} of a volumetric object, along with a set of physical parameters 
-(density, constitutive model, model parameters etc). The output is an elastodynamic simulation. 
+(density, constitutive model, model parameters etc). 
+The boundary representation is composed of one or more (not necessarily explicit;y connected) NURBS surface primitives that we will call \emph{parts}.
+The output is an elastodynamic simulation. 
 SEM consists of preprocessing and runtime simulation phases~\refalg{sem} and at no point do we need to generate a volumetric mesh of the input. 
@@
 \subsection{Shape Matching}
+
+We begin by describing shape matching to a single NURBS part. 
+Given a two different configurations of the same part (specified by the control points), shape matching computes a polynomial deformation which best aligns them.
+
+Let $\vc{q^0}$ be the initial control points of the part, provided by the modeller and Let $\vc{q}$ be a vector of modified control point values.
+We can compute the undeformed (or reference) position of any point on our part using $\refX\left(u,v\right) = \Jnurbs\left(u,v\right)\mathbf{q^0}$ and the corresponding
+deformed (world space) position as $\defX\left(u,v\right) = \Jnurbs\left(u,v\right)\mathbf{q}$.
+
+We can define a polynomial deformation from $\refX = \begin{pmatrix} X & Y & Z \end{pmatrix}^T$ to $\defX$  as 
+\begin{equation}
+    \defX\left(\refX\right) = 
+    \underbrace{\begin{pmatrix}
+    \mathbf{p}^T(\refX-\bar{\refX}) & \mathbf{0} & \mathbf{0} & 1 & 0 & 0\\
+    \mathbf{0} & \mathbf{p}^T(\refX-\bar{\refX}) & \mathbf{0} & 0 & 1 & 0\\
+    \mathbf{0} & \mathbf{0} & \mathbf{p}^T(\refX-\bar{\refX}) & 0 & 0 & 1\\
+    \end{pmatrix}}_{\Pmatx}
+    \Pcox,
+\end{equation} where $\mathbf{p}^T\left(\refX\right)$ is the polynomial basis vector \newline (eg $\begin{pmatrix} X & Y & X\cdot Y & X\cdot X & Y\cdot Y \end{pmatrix}$ ),
+$\bar{\refX}$ is an a priori chosen origin around which to deform, $\Pco$ are the coefficients for the 
+non-constant part of the polynomial and $\tr$ is the translation of the deformation origin. 
+We can compute the shape matching deformation by solving
+\begin{equation}
+    \begin{pmatrix} 
+        \Pco\\
+        \tr
+    \end{pmatrix}^* = \arg\min_{\Pco,\tr} \int_{u,v} \left\lvert\left\lvert\Pmatx\Pcox - \refX\right\rvert\right\rvert^2_2 du\cdot dv,
+\end{equation} where we have surpressed the dependence of $\refX$ on $u$ and $v$ for the sake of brevity.
+
+Discretizing this integral and solving
+
+\dave{Ty how do we pick u,vs and weights for this formula?}
+
+%%%%% OLD STUFF BELOW HERE %%%%%%%
 \dave{Shape Matching to a single NURBS part}

### 2021-01-20 paper/sections/methods.tex
shape matching initial part done

@@
 \begin{equation}
-    \begin{pmatrix} 
-        \Pco\\
-        \tr
-    \end{pmatrix}^* = \arg\min_{\Pco,\tr} \int_{u,v} \left\lvert\left\lvert\Pmatx\Pcox - \refX\right\rvert\right\rvert^2_2 du\cdot dv,
-\end{equation} where we have surpressed the dependence of $\refX$ on $u$ and $v$ for the sake of brevity.
+   \Pcox^* = \arg\min_{\Pco,\tr} \int \left\lvert\left\lvert\Pmatx\Pcox - \Jnurbs\vc{q}\right\rvert\right\rvert^2_2 dudv,
+\end{equation} where we have suppressed the dependence of $\refX$ and $\Jnurbs$ on $u$ and $v$ for the sake of brevity.
 
-Discretizing this integral and solving
+We discretize the shape matching energy and minimize by solving
+\begin{equation}
+\samp{\Pmat}^T\wt\samp{\Pmat}\Pcox = \samp{\Pmat}^T\wt\samp{\Jnurbs}\vc{q}.
+\end{equation} Here $\samp{\Pmat}$ (resp. $\samp{\Jnurbs}$) is a $3s \times 3k$ (resp. $3s \times 3nm$) matrix where $s$ is the number of quadrature points, 
+$k$ is the number of  terms in the shape matching polynomial and $3nm$ is the number of NURBS control points. 
+$\samp{\Pmat}$ (resp. $\samp{\Jnurbs}$) is constructed by stacking the evaluations of $\Pmat(X\left(u, v\right))$ (resp. $\Jnurbs\left(u, v\right)$) at each
+quadrature point.
+$\wt$ is a diagonal, $3s \times 3s$ matrix of integration weights.
 

### 2021-01-20 paper/sections/methods.tex
relationship to VEM clarified

@@
 %\end{itemize}
-
+\dave{cite jacobson stretchable twistable bones for constant term}
 \section{Overview}
@@
 non-constant part of the polynomial and $\tr$ is the translation of the deformation origin. 
-We can compute the shape matching deformation by solving
+We can compute the shape matching deformation by solving~\cite{10.1145/1073204.1073216}
 \begin{equation}
@@
 \samp{\Pmat}^T\wt\samp{\Pmat}\Pcox = \samp{\Pmat}^T\wt\samp{\Jnurbs}\vc{q}.
+\label{eq:shapematch}
 \end{equation} Here $\samp{\Pmat}$ (resp. $\samp{\Jnurbs}$) is a $3s \times 3k$ (resp. $3s \times 3nm$) matrix where $s$ is the number of quadrature points, 
@@
 
-\dave{Ty how do we pick u,vs and weights for this formula?}
+\paragraph*{Relationship to the Virtual Element Method}
+For a non-planar NURBS part, and given a sufficient number of quadrature points, ~\refeq{shapematch} will be well-posed allowing us to write: $\Pcox^* = \Pi\vc{q}$ where
+\begin{equation*}
+    \Pi = \left(\samp{\Pmat}^T\wt\samp{\Pmat}\right)^{-1}\samp{\Pmat}^T\wt\samp{\Jnurbs},
+\end{equation*} is a projection from the space of control points onto the space of polynomials. 
+This projection serves an identical purpose (and is mathematically equivalent up to choice of metric) to the projection operator used in VEM~\cite{10.1142/S021820251440003X}.
+VEM uses a single polynomial projection per polyhedral element to solve differential equations on complex volumetric meshes.
+We interpret shape matching as converting our NURBS part into a single Virtual Element suitable for elastodynamics simulation.
+Unfortunately a single polynomial deformation space will not be suitably expressive for even moderately complex models. 
+Previous work~\cite{10.1145/1073204.1073216,10.1145/1275808.1276480} used additional lattices or partitions of the reference space to 
+increase the number of polynomials in the shape matching. Next we outline our meshless Shape Element Method which works directly on the NURBS
+geometry itself.
 
-%%%%% OLD STUFF BELOW HERE %%%%%%%
-\dave{Shape Matching to a single NURBS part}
-In typical FEM in graphics, we make use of tetrahedral or triangular elements with a piecewise-linear function space.
-In such a case our functions may be evaluated as a weighted combination of the nodal values of an element: $f(\mathbf x)=\sum_{i=0}^n f_i \phi_i(\mathbf{x})$ where $n$ is the number of nodal values and $\phi_i$ is the i-th barycentric coordinates.
-Thus in this case we have the simple deformation map $\mathbf{x(X)} = \sum_{i=0}^n \mathbf{x}_i \phi_i(\mathbf{X})$.
-However, our method contrasts this typical approach in that our elements are more general and take the form of any arbitrary boundary primitive, such as NURBS, and also that the function space in non-linear.
-
-In our algorithm, a single shape matching element is represented by a NURBS surface, a center of mass, and polynomial describing its current deformation.
-We construct these polynomials via a shape matching strategy.
-With the addition of the center of mass, we describe the shapes deformation about this point and thus the monomial basis in the case of a linear polynomial is $\left[ \mathbf 1^T (\mathbf{X - \bar{X}})^T \right]$ where $\mathbf 1^T$ is a constant $1$ column vector and $\mathbf{\bar{X}}$ is the center of mass for the element.
-To simplify the following formulas, we collect the non-constant terms of the monomial basis into a single vector $\mathbf{m}(\mathbf{X}-\mathbf{\bar{X}})$.
-For example if $\mathbf{X} = (X,Y,Z)$, $\mathbf{\bar{X}} = (\bar X, \bar Y, \bar Z)$, and $\mathbf{q}=\mathbf{X}-\mathbf{\bar{X}}$ with a quadratic polynomial we would have
-
-\begin{figure}
-    \includegraphics[width=\columnwidth]{example-image-a}
-    \caption{Figure showing shape matching to NURBS}
-    \label{fig:shapematching}
-\end{figure}
-
-\begin{equation}
-m(\mathbf{q)} = (
-					q_1, q_2, q_3,
-                   	q_1^2, q_2^2, q_3^3,
-                   	q_1 q_2, q_2 q_3
-                ) \in \mathbb{R}^k
-                \text{,}
-\end{equation}
-where $k$ is the size of the monomial basis. Next we arrange this basis into a matrix of the form
-\begin{equation}
-\mathbf{M}(\mathbf{q}) = \left[ \begin{array}{ccc}
-\mathbf{m}(\mathbf{q}) &  &  \\
- & \mathbf{m}(\mathbf{q})  &  \\
- &  & \mathbf{m}(\mathbf{q})  \\
-\end{array} \right] \in \mathbb{R}^{3x3k}
-\text{.}
-\end{equation}
-With this definition of the monomial basis, we may reconstruct the deformed position $\mathbf{x}$ for some material point $\mathbf{X}$ with
-\begin{equation}
-\mathbf{x}(\mathbf{X}) = \left[\mathbf{M}(\mathbf{X-\bar{X}}) \quad \mathbf{I} \right]
-\left[ \begin{array}{c}
-\mathbf{c} \\
-\mathbf{t} \\
-\end{array} \right]
-\text{,}
-\end{equation}
-where $\mathbf{c} \in \mathbb{R}^{3k}$ is the set of polynomial coefficients for the non-constant terms, $\mathbf{t} \in \mathbb{R}^3$ is the constant term coefficients, and $\mathbf{I} \in \mathbb{R}^{3x3}$ is an identity matrix.
-The reason for separating out the constant term becomes more evident when we examine case where we have multiple elements.
-
-\dave{here explain sampling on NURBS surface to get material points}
-The above formulation describes developing the map from a single $(u,v)$ coordinate to its corresponding position on the surface.
-In our simulation we represent the entire boundary of the NURBS model, so we require samples along each surface.
-We emphasize that we only require an initial selection of $(u,v)$ coordinates across each surface and the dynamics of the simulation modifies the set of control points $\mathbf{q}$ to modify the world space maps $\mathbf{x}(u,v)=\mathbf{J}(u,v)\mathbf{q}$.
-
-Let's say for a NURBS surface on the model the DOF column vector is $\mathbf{q}_i$, which is the subset of control points in the generalized coordinates for the $i$-th surface.
-For a given surface with $m$ pairs of $\mathbf{u}=(u,v)$ coordinates sampled in parametric space, we write the jacobian for the $i$-th surface as \myworries{(should I use something like $\hat{\mathbf{J}}$ instead?)}
-\begin{equation}
-\label{eqn:surface_jacobian}
-    \mathbf{J}_i =
-    \left[ \mathbf{J}(\mathbf{u})_1^T \dots \mathbf{J}(\mathbf{u})_m^T \right]^T
-    \text{.}
-\end{equation}
-Then if $\mathbf{x}_i$ is the set of $m$ world space positions for the $i$-th surface, we may write $\mathbf{x}_i = \mathbf{J}_i \mathbf{q}_i$.
-Finally, given the full set of generalized coordinates $\mathbf{q} = \left[ \mathbf{q}_i^T, \dots, \mathbf{q}_n^T \right]^T$ with $n$ surfaces, we may write jacobian $\mathbf{J}$ for the entire model as a block diagonal matrix:
-\begin{equation}
-\mathbf{J} = \left[ \begin{array}{ccccc}
-\mathbf{J}_1 &  &  &  &  \\
- & \ddots &  &  &  \\
- &  & \mathbf{J}_i & &  \\
- &  &  & \ddots &  \\
- &  &  &  & \mathbf{J}_n \\
-\end{array} \right]
-\text{.}
-\end{equation}
-
-\dave{weights are missing, just matching to points leaves us vulnerable to biasing towards clusters of points}
-The remaining step in defining a single element is to describe how to solve for the polynomial coefficients.
-For a single element, we have $n$ material points $(\mathbf{X}_1 \dots \mathbf{X}_n)$ and their corresponding deformed positions $(\mathbf{x}^*_1 \dots \mathbf{x}^*_n)$.
-As these elements are defined on a boundary, these are both known throughout the simulation.
-Therefore we may form a fitting problem to solve for the polynomial coefficients that minimize the error $\sum_i^n ||\mathbf{x}_i^* - \mathbf{X}_i ||^2$.
-We minimize this error in a least squares sense where we first form the system
-\begin{equation}
-\left[\begin{array}{c}
-\mathbf{M}(\mathbf{X_1-\bar{X}}) \quad \mathbf{I} \\
-\vdots \\
-\mathbf{M}(\mathbf{X_i-\bar{X}}) \quad \mathbf{I} \\
-\vdots \\
-\mathbf{M}(\mathbf{X_n-\bar{X}}) \quad \mathbf{I} \\
-\end{array} \right]
-\left[ \begin{array}{c}
-\mathbf{c} \\
-\mathbf{t} \\
-\end{array} \right] = 
-\left[ \begin{array}{c}
-\mathbf{x}_1^* \\
-\vdots \\
-\mathbf{x}_i^* \\
-\vdots \\
-\mathbf{x}_n^*
-\end{array} \right]
-\text{,}
-\end{equation}
-and then in the more compact form
-\begin{equation}
-\label{eqn:fitting_one}
-\left[
-\mathbf{A} \quad \mathbf{S}
-\right]
-\left[ \begin{array}{c}
-\mathbf{c} \\
-\mathbf{t} \\
-\end{array} \right] = \mathbf{b}
-\text{,}
-\end{equation}
-where $\mathbf{A}$ is the block containing the monomial matrices of the system, and $\mathbf{S}$ is the block of identity matrices.
-Next, to reduce the degrees of freedom of this system we let the deformed center of mass (the constant term coefficients), $t$, be equal to the mean of the deformed positions influenced by this center of mass.
-In the case of one element we have $t = \frac{1}{n}\sum_i^n x_i^*$.
-We describe the details on center of mass selection in section \myworries{insert here}.
-In matrix form we have
-\dave{need to explain that system can be degenerate fix is to select a constant term}
-\begin{equation}
-\label{eqn:t_solution}
-t = 
-\underbrace{
-\left[\begin{array}{ccccccc}
-\frac{1}{n} & & \cdots \frac{1}{n} & & \\
-& \frac{1}{n} & \cdots & \frac{1}{n} & \\
-& & \frac{1}{n} \cdots & & \frac{1}{n} \\
-\end{array} \right]}_{T} 
-\mathbf{b}
-\text{.}
-\end{equation} (\myworries{spacing on this looks bad....}). Finally, this leaves us with the system
-\begin{equation}
-\mathbf{A}\mathbf{c}=(\mathbf{I}-\mathbf{ST})\mathbf{b}
-\text{.}
-\end{equation}
-Like in the original Shape Matching \cite{10.1145/1073204.1073216}, we solve for the coefficients by minimizing the error in a least squares sense, resulting in
-\dave{no b just use J q, here's a good place to explain sampling the surface to build big J.}
-\begin{equation}
-\label{eqn:c_one_element}
-\begin{split}
-\mathbf{c} & = (\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T(\mathbf{I-ST})\mathbf{b} \\
-           & = \mathbf{L}\mathbf{b}
-\end{split}
-\text{.}
-\end{equation}
-This $\mathbf{L}$ matrix thus gives us our file ingredient in the element, coefficients defining the polynomial.
-We emphasize that $\mathbf{L}$ is constant throughout the simulation, therefore as $\mathbf{b}$ varies in the simulation, $\mathbf{Lb}$ gives us a new set of polynomial coefficients as $\mathbf{b}$ changes.
-
-\dave{Shape Matching is VEM brief explanation here. Point of departure, one polynmoial isn't going to cut it for a whole object.}
 \subsection{Shape Matching Element Method}

### 2021-01-21 matlab/vem_error_matrix.m
update vem_sim_nurbs to use new coms (still need to do others)

@@ -3,7 +3,6 @@ function ME = vem_error_matrix(Y, Y_S, L, d)
     M1 = zeros(size(L,1),size(L,1));
     M2 = zeros(size(L,1),size(L,2));
     M3 = eye(size(L,2));
-    tic
     for i=1:m
         M1 = M1 + Y_S{i}'*(Y{i}'*Y{i})*Y_S{i};
         M2(:,d*(i-1)+1:d*i) = Y_S{i}'*Y{i}';

### 2021-01-21 matlab/vem_simulate_nurbs.m

@@ -16,7 +16,7 @@ function vem_simulate_nurbs(parts, varargin)
     addParameter(p, 'sample_interior', 1);
     addParameter(p, 'distance_cutoff', 20);
     addParameter(p, 'enable_secondary_rays', true);
-    addParameter(p, 'fitting_mode', 'global');
+    addParameter(p, 'fitting_mode', 'hierarchical');
     addParameter(p, 'plot_points', false);
     addParameter(p, 'plot_com', true);
     addParameter(p, 'com_threshold', 100);
@@ -37,17 +37,6 @@ function vem_simulate_nurbs(parts, varargin)
 
     % Assembles global generalized coordinates
     [J, ~, q, E, x0] = nurbs_assemble_coords(parts);
-
-    % Generating centers of mass. Temporary method!
-    [x0_coms, com_cluster, com_map] = generate_com(parts, x0, E, ...
-        config.com_threshold, n);
-    
-    if config.plot_com
-        com_plt = plot3(x0_coms(1,:),x0_coms(2,:),x0_coms(3,:), ...
-                        '.','Color','g','MarkerSize',20);
-        hold on;
-    end
-    %%%%%%%%%%%%%%%%%%%%%%%%
     
     % Initial deformed positions and velocities
     x = x0;
@@ -82,16 +71,24 @@ function vem_simulate_nurbs(parts, varargin)
   	f_gravity = repmat([0 0 config.rho * config.gravity], size(x0,2),1)';
     f_gravity = config.dt*P*f_gravity(:);
 
-    % Shape Matrices
-    L = compute_shape_matrices(x0, x0_coms, com_map, E, ...
-        com_cluster, config.order, config.fitting_mode);
-
     % Compute Shape weights
     [w, w_I] = nurbs_blending_weights(parts, V', config.distance_cutoff, ...
         'Enable_Secondary_Rays', config.enable_secondary_rays);
     [w0, w0_I] = nurbs_blending_weights(parts, x0', config.distance_cutoff, ...
         'Enable_Secondary_Rays', config.enable_secondary_rays);
-                             
+    
+    % Generate centers of mass.
+    [x0_coms, com_cluster, com_map] = generate_com(x0, E, w, n);
+    if config.plot_com
+    com_plt = plot3(x0_coms(1,:),x0_coms(2,:),x0_coms(3,:), ...
+                    '.','Color','g','MarkerSize',20);
+    hold on;
+    end
[…]

### 2021-01-21 cpp/src/d2psi_neohookean_dF2.cpp
fix bugs in newton's method

@@ -196,8 +196,8 @@ typename DefoType::Scalar F1_1, F1_2, F1_3, F2_1, F2_2, F2_3,F3_1, F3_2, F3_3;
         F3_1 = F(2,0);
         F3_2 = F(2,1);
         F3_3 = F(2,2);
-		/*
-    ddw(0,0) = params1*(pow(F2_2*F3_3-F2_3*F3_2,2.0)*1.0/pow(F1_1*F2_2*F3_3-F1_1*F2_3*F3_2-F1_2*F2_1*F3_3+F1_2*F2_3*F3_1+F1_3*F2_1*F3_2-F1_3*F2_2*F3_1,2.0)*2.0+2.0)+params2*pow(F2_2*F3_3-F2_3*F3_2,2.0)*2.0;
+		
+        ddw(0,0) = params1*(pow(F2_2*F3_3-F2_3*F3_2,2.0)*1.0/pow(F1_1*F2_2*F3_3-F1_1*F2_3*F3_2-F1_2*F2_1*F3_3+F1_2*F2_3*F3_1+F1_3*F2_1*F3_2-F1_3*F2_2*F3_1,2.0)*2.0+2.0)+params2*pow(F2_2*F3_3-F2_3*F3_2,2.0)*2.0;
        ddw(0,1) = params2*(F1_2*F3_3-F1_3*F3_2)*(F2_2*F3_3-F2_3*F3_2)*-2.0-params1*(F1_2*F3_3-F1_3*F3_2)*(F2_2*F3_3-F2_3*F3_2)*1.0/pow(F1_1*F2_2*F3_3-F1_1*F2_3*F3_2-F1_2*F2_1*F3_3+F1_2*F2_3*F3_1+F1_3*F2_1*F3_2-F1_3*F2_2*F3_1,2.0)*2.0;
        ddw(0, 2) = params2*(F1_2*F2_3-F1_3*F2_2)*(F2_2*F3_3-F2_3*F3_2)*2.0+params1*(F1_2*F2_3-F1_3*F2_2)*(F2_2*F3_3-F2_3*F3_2)*1.0/pow(F1_1*F2_2*F3_3-F1_1*F2_3*F3_2-F1_2*F2_1*F3_3+F1_2*F2_3*F3_1+F1_3*F2_1*F3_2-F1_3*F2_2*F3_1,2.0)*2.0;
[…]

### 2021-01-21 cpp/src/vem3dmesh_simulate_one_step.cpp

@@ -86,15 +86,18 @@ void sim::vem3dmesh_simulate_one_step(Eigen::VectorXx<DerivedRet> &qdot_new,
 		Eigen::MatrixXd x_new;
 		Eigen::VectorXd c;
 		get_x_new_and_c(x_new, c, qdot_new);
+		// std::cout << "pass x_new: " << x_new.block(0,0,3,3) << std::endl;								
+		// std::cout << "pass c: " << c.segment(0, 6) << std::endl;								
+
 		// compute vem neohookean gradient
 		Eigen::VectorXd g_neohookean;
 		// std::cout << "start vem3dmesh_neohookean_dq" << std::endl;	
 		vem3dmesh_neohookean_dq(g_neohookean, x_new, c, volume, params, dF_dc, dF_dc_S, ME, L, 
 														k, n, d, x0_coms_size, k_stability);
-		// std::cout << "pass vem3dmesh_neohookean_dq" << std::endl;	
+		// std::cout << "pass vem3dmesh_neohookean_dq: " << std::endl << g_neohookean.segment(0,6) << std::endl;	
 		g = JtPMPtJ * (qdot_new - qdot) + J.transpose() * (dt*P*g_neohookean - f_ext); 	
 
-		// std::cout << "pass grad" << std::endl;								
+		// std::cout << "pass grad: " << g.segment(0, 6) << std::endl;								
 	};
 	auto hessian = [&](Eigen::MatrixXd &H, Eigen::VectorXd &qdot_new) {
 		// std::cout << "start hessian" << std::endl;			
@@ -104,10 +107,11 @@ void sim::vem3dmesh_simulate_one_step(Eigen::VectorXx<DerivedRet> &qdot_new,
 
 		Eigen::MatrixXd K;
 		vem3dmesh_neohookean_dq2(K, c, volume, params, dF_dc, W_I, k, n, x0_coms_size);
-		K = L.transpose() * (-K) * L;
+		K = L.transpose() * (K) * L;
+
 		H = JtPMPtJ + J.transpose() * P * K * P.transpose() * J * dt * dt;
 
-		// std::cout << "pass hessian" << std::endl;
+		// std::cout << "pass hessian: " << H.block(0,0,6,6) << std::endl;
 	};
 
 	Eigen::VectorXd tmp_g(qdot.rows());

### 2021-01-21 examples/linear_subD_torus.m

@@ -1,0 +1,25 @@
+function linear_subD_torus
+function pinned_ids = pin_function(x)
+    verts_to_pin = 2; 
+    [~,I] = mink(x(1,:),verts_to_pin);
+    pinned_ids = I(1:verts_to_pin);
+end
+
+iges_file = 'torus_subd.igs';
+
+part = nurbs_from_iges(iges_file);
+
+options.order = 1;
+options.gravity = -10;
+% options.rho = .2;
+options.pin_function = @pin_function;
+% options.lambda = 100;
+% options.mu = 1500;
+% options.sample_interior = 0; % only sample on boundary
+% options.distance_cutoff = 1;
+options.enable_secondary_rays = false;
+options.save_obj = true;
+axis equal
+vem_simulate_nurbs(part, options);
+
+end

### 2021-01-21 examples/quadratic_beam.m

@@ -34,6 +34,6 @@ options.enable_secondary_rays = 0;
 options.fitting_mode = 'hierarchical';
 options.save_output = 0;
 
-vem_simulate_nurbs(part, options);
+vem_simulate_nurbs_newtons(part, options);
 
 end

### 2021-01-21 examples/quadratic_rounded_cube.m

@@ -2,26 +2,23 @@ function quadratic_rounded_cube
 
 function pinned_ids = pin_function(x)
     % pinning corner points of the cube
-    x_min = 40;
-    y_min = 55;
+    x_min = 0.4;
+    y_min = 0.4;
     pinned_ids = find(x(1,:) > x_min & x(2,:) > y_min);
 end
 
-iges_file = 'rounded_cube.iges';
+iges_file = 'rounded_cube.igs';
 part = nurbs_from_iges(iges_file);
-
-YM = 25; %in Pascals
-pr =  0.15;
+YM = 1e4; %in Pascals
+pr =  0.45;
 [lambda, mu] = emu_to_lame(YM, pr);
-
 options.order = 2;
-options.rho = 1e-5;
-options.gravity = -100;
+options.rho = 1e-1;
+options.gravity = -10;
 options.pin_function = @pin_function;
-options.gravity = -100;
 options.lambda = lambda;
 options.mu = mu;
 
-vem_simulate_nurbs(part, options);
+vem_simulate_nurbs_newtons(part, options);
 
 end

### 2021-01-21 matlab/symbolic/neohookean_func.m

@@ -16,7 +16,9 @@ psi=C*(I3-3)+D*(J-1)^2;
 
 g = gradient(psi,F(:));
 H = hessian(psi,F(:));
-ccode(H)
+% ccode(psi)
+ccode(g)
+% ccode(H)
 % matlabFunction(g, 'vars', {F,C,D},'File','SNH_dF','Comments','Version: 1.0')
 % matlabFunction(H, 'vars', {F,C,D},'File','SNH_dF2','Comments','Version: 1.0')
 end

### 2021-01-21 paper/sections/methods.tex
one change

@@
 \begin{figure}
-    \includegraphics[width=\columnwidth]{example-image-a}
-    \caption{Figure showing simple NURBS patch with variables labelled}
+    \includegraphics[width=\columnwidth]{figures/nurbs_patch}
+    \caption{A cubic NURBS patch with 16 control points.}
     \label{fig:NURBS}
@@
 
+\dave{Ty: need some text to explain how we pick points on the nurbs surface.}
 We discretize the shape matching energy and minimize by solving

### 2021-01-22 examples/linear_rocket_with_collision.m
fixed gravity force

@@ -13,17 +13,27 @@ pr = 0.32;
 [lambda, mu] = emu_to_lame(YM, pr);
 
 options.order = 1;
-options.gravity = -1e3;
+options.gravity = -20;
 options.rho = 8e3;
 options.pin_function = @pin_function;
 options.lambda = lambda;
 options.mu = mu;
-options.distance_cutoff = 10;
 
-options.save_obj = true;
+options.distance_cutoff = 10; % use 5 for a single com;
+
+options.save_obj = false;
 
 options.k_stability = 1e9;
 options.collision_ratio = 1e7;
+
+options.x_samples = 5;
+options.y_samples = 9;
+options.z_samples = 15; % we want more samples along vertical axis
+
+% If you want centers of mass on the legs of the rocket, you'll need to
+% make sure points are sampled on the legs. You can fidget with the above
+% numbers to try to change it, if you need.
+% options.plot_points=true; % Enable to see all the equadrature points.
 
 options.collision_with_other = false;
 options.self_collision = false;

### 2021-01-22 figures/raycasting_integrate_volume.m

@@ -26,11 +26,19 @@ view(-15,30);
 nsamples = 20;
 x_vals = 1:nsamples;
 errors = zeros(size(x_vals));
-
 true_vol = 11774.6564; % mm^3 for the trident 
 % true_vol = 1.17697114 ; % rounded cube
 % true_vol = 5; % mm^3 for the beam 
 % true_vol = 267.11
+
+i=2;
+[V, ~] = raycast_quadrature(parts, [i i], 5);
+
+V_plt = plot3(V(1,:),V(2,:),V(3,:),'.','Color','m','MarkerSize',20);
+V([2 3],:) = V([3 2],:);
+F = 1:size(V,2);
+writePLY('samples.ply', V', F', 'ascii');
+
 for i = 1:nsamples
    [V, vols] = raycast_quadrature(parts, [i i], 5);
    e = abs(sum(vols) - true_vol) / true_vol;
@@ -41,7 +49,6 @@ for i = 1:nsamples
       V_plt.XData=V(1,:);
       V_plt.YData=V(2,:);
       V_plt.ZData=V(3,:);
-      
    end
 end
 
@@ -51,4 +58,5 @@ plot(x_vals,errors);
 title('Error as the number of samples increase in each direction');
 ylabel('Relative Absolute Error');
 xlabel('# Samples');
+xlim([1 20])
 end

### 2021-01-22 matlab/vem_simulate_nurbs.m

@@ -71,8 +71,8 @@ function vem_simulate_nurbs(parts, varargin)
     params = repmat(params,size(V,2),1);
         
     % Gravity force vector.
-  	f_gravity = repmat([0 0 config.rho * config.gravity], size(x0,2),1)';
-    f_gravity = config.dt*P*f_gravity(:);
+  	%f_gravity = repmat([0 0 config.rho * config.gravity], size(x0,2),1)';
+    %f_gravity = config.dt*P*f_gravity(:);
 
     % Compute Shape weights
     [w, w_I] = nurbs_blending_weights(parts, V', config.distance_cutoff, ...
@@ -140,6 +140,8 @@ function vem_simulate_nurbs(parts, varargin)
 
         % Force vector
         dV_dq = zeros(d*(k*n + size(x0_coms,2)),1);
+        
+        dg_dc = zeros(d*(k*n + size(x0_coms,2)),1);
 
         % Computing force dV/dq for each point.
         % TODO: move this to C++ :)
@@ -153,8 +155,13 @@ function vem_simulate_nurbs(parts, varargin)
             % Force vector
             dV_dF = neohookean_tet_dF(F, params(i,1), params(i,2));
             dV_dq = dV_dq +  dF_dc_S{i}' * dF_dc{i}' * dV_dF * vol(i);
+            
+            grav = [0 0 config.gravity];
+            dg_dc = dg_dc + (config.rho*vol(i)*grav*Y{i}*Y_S{i})';
         end
         dV_dq = L' * dV_dq;
+        
+        f_gravity = config.dt*P*(L' * dg_dc);
         
         % Error correction force
         f_error = - 2 * ME * x(:);

### 2021-01-22 paper/sections/methods.tex
slowly working my way through the mulitple projections section

@@
 %\end{itemize}
-\dave{cite jacobson stretchable twistable bones for constant term}
 \section{Overview}
@@
 The Shape Matching Element Method will follow from this kinematic description.
-% The NURBS jacobian, $\mathbf{J}$, is a matrix composed of horizontally concatenated $3x3$ blocks where the $(i,j)$-th block is $\frac{\partial \mathbf{x}}{\partial \mathbf{p}_{i,j}}$ for the $(i,j)$-th control point.
-$\frac{\partial \mathbf{x}}{\partial \mathbf{p}_{i,j}}$ is a diagonal $3x3$ matrix where each diagonal entry $N_{i,j}(\mathbf{u})$ takes the form
-
-% \begin{equation}
-% \label{eqn:jacobian_diagonal}
-%     N_{i,j}(u,v)
-%     = \frac{w_{i,j} B_{i,k}(u)B_{j,l}(v)}{\sum_{i=1}^{n}\sum_{j=1}^{m} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
-%     \text{.}
-% \end{equation}
-% Thus the NURBS jacobian for a single $(u,v)$ pair is written as
-% \begin{equation}
-% \label{eqn:uv_jacobian}
-%     \mathbf{J}(u,v) =
-%     \left[ \frac{\partial \mathbf{x}}{\partial \mathbf{p}_{1,1}} \dots
-%            \frac{\partial \mathbf{x}}{\partial \mathbf{p}_{i,j}} \dots 
-%            \frac{\partial \mathbf{x}}{\partial \mathbf{p}_{n,m}}
-%     \right]
-%     \text{.}
-% \end{equation}
-
-% With this equation, the full set of world positions of the NURBS model may be written as $\mathbf{x} = \mathbf{J}\mathbf{q} \in \mathbb{R}^{3N}$ where $N$ is the total number of $(u,v)$ coordinates.
 
@@
 \subsection{Shape Matching}
+\label{sec:shapematching}
 
@@
     \Pcox,
+    \label{eq:projection}
 \end{equation} where $\mathbf{p}^T\left(\refX\right)$ is the polynomial basis vector \newline (eg $\begin{pmatrix} X & Y & X\cdot Y & X\cdot X & Y\cdot Y \end{pmatrix}$ ),
@@
 $k$ is the number of  terms in the shape matching polynomial and $3nm$ is the number of NURBS control points. 
-$\samp{\Pmat}$ (resp. $\samp{\Jnurbs}$) is constructed by stacking the evaluations of $\Pmat(X\left(u, v\right))$ (resp. $\Jnurbs\left(u, v\right)$) at each
+$\samp{\Pmat}$ (resp. $\samp{\Jnurbs}$) is constructed by stacking the evaluations of $\Pmatxuv$ (resp. $\Jnurbs\left(u, v\right)$) at each
 quadrature point.
@@
 \subsection{Shape Matching Element Method}
-\dave{Multiple Parts}
-\dave{each polynomial/NURBS tuple is a shape matching element}
-\dave{Shape Matching Element assembles global kinematics using blendeded formultion}
+ Each input object is composed of a set of boundary representations, $\Prts$, one for each part $\prt\in\Prts$. 
+ Our kinematic mapping is a blended construction (as is done for skinning~\cite{10.1145/2614028.2615427}) which we write as
+
+ \begin{equation}
+    \defX\left(\refX\right) = \sum_{\prt_i\in\Prts} \bld_i\Pmatxi\Pi_i\vc{q}_i, 
+ \end{equation} where $w_i$ $\Pmatx_i$, $\Pi_i$ and $\vc{q}_i$ are the blending weights, polynomial basis, projection operator and control points for the $i^{th}$
+ part respectively.
+
+ \subsubsection*{Multiple Part Projection}
+As in ~\refsec{shapematching} we construct each $\Pi_i$ via shape matching. 
+It is tempting to use an identical polynomial basis matrix, $\Pmatx$ for all parts, $\prt_i$, but recall that constructing $\Pmatx$ requires the selection of a deformation
+origin $\bar{\refX}$~(\refeq{projection}). 
+Using the same origin for all polynomials limits the expressivity of the kinematic model~\cite{STBS:2011}.
+Instead we choose a number of points in $\refX$ ($|\Prts| > |\XBs| > 0$, where $\XBs$ is the set of points)  to act as deformation origins~(\refsec{origins}).
+For each part we now construct $\Pmatxi$ using the associated deformation origin $\bar{\refX}_j\in \XBs$, where each origin can
+be shared between multiple parts.
+
+Attempting to use $\Pmatx_i$ in~\refeq{shapematch} on a per-part basis is problematic as it can become ill-posed, especially for planar boundary representations. 
+To alleviate this problem we invoke the \emph{hierarchical ordering principle}, which argues that the behavior of a system is typically 
+dominated by lower order effects~\cite{li2006regularities}.
+
+ \dave{heirarchical ordering to handle degenerecies}
+
+\subsection{Meshless Blending Weight Computation}
+
+\subsection{Choosing Deformation Centers}
+\label{sec:origins}
+
+%%%% OLD STUFF BELOW HERE  %%%%%
 In a typical simulation we will have multiple shape matching elements where the boundary of each element is defined by a different NURBS surface.

### 2021-01-22 matlab/vem_simulate_nurbs.m
add lamppost example

@@ -22,6 +22,7 @@ function vem_simulate_nurbs(parts, varargin)
     addParameter(p, 'x_samples', 5);
     addParameter(p, 'y_samples', 9);
     addParameter(p, 'z_samples', 9);
+    addParameter(p, 'f_external', [0 0 0]);
 
     parse(p,varargin{:});
     config = p.Results;
@@ -142,6 +143,8 @@ function vem_simulate_nurbs(parts, varargin)
         dV_dq = zeros(d*(k*n + size(x0_coms,2)),1);
         
         dg_dc = zeros(d*(k*n + size(x0_coms,2)),1);
+        
+        df_ext_dc = zeros(d*(k*n + size(x0_coms,2)),1);
 
         % Computing force dV/dq for each point.
         % TODO: move this to C++ :)
@@ -158,10 +161,14 @@ function vem_simulate_nurbs(parts, varargin)
             
             grav = [0 0 config.gravity];
             dg_dc = dg_dc + (config.rho*vol(i)*grav*Y{i}*Y_S{i})';
+            
+            df_ext_dc = df_ext_dc + (vol(i)*config.f_external*Y{i}*Y_S{i})';
         end
         dV_dq = L' * dV_dq;
         
         f_gravity = config.dt*P*(L' * dg_dc);
+        
+        f_external = config.dt*P*(L' * df_ext_dc);
         
         % Error correction force
         f_error = - 2 * ME * x(:);
@@ -174,7 +181,7 @@ function vem_simulate_nurbs(parts, varargin)
         % Note: I believe i'm forgetting the error matrix stiffness matrix
         %       but I don't wanna break things so I haven't added it yet.
         lhs = J' * (P*(M - config.dt*config.dt*K)*P') * J;
-        rhs = J' * (P*M*P'*J*qdot + f_internal + f_gravity + f_error);
+        rhs = J' * (P*M*P'*J*qdot + f_internal + f_gravity + f_error + f_external);
         qdot = lhs \ rhs;
 
         % Update position

### 2021-01-22 cpp/include/parse_matlab_cells.h
added mexed version of external constant force

@@ -1,0 +1,59 @@
+#ifndef SIM_PARSE_MATLAB_CELLS
+#define SIM_PARSE_MATLAB_CELLS
+
+#include <Eigen/Dense>
+#include <EigenTypes.h>
+#include <igl/matlab/parse_rhs.h>
+#include <igl/matlab/prepare_lhs.h>
+
+namespace sim {
+
+	// Parse cells composed of dense matrices
+	template <typename DerivedV>
+	void parse_cell(const mxArray *prhs[], int index, EigenMatrixList<DerivedV>& V) {
+		const mxArray *cell = prhs[index];
+		const mwSize *dims = mxGetDimensions(prhs[index]);
+		const mxArray *cellElement;
+		mwIndex jcell;
+		for (jcell = 0; jcell < dims[0]; jcell++) {
+			cellElement = mxGetCell(cell, jcell);
+			Eigen::MatrixXx<DerivedV> element;
+			igl::matlab::parse_rhs_double(&cellElement, element);
+
+			V.emplace_back(element);
+		}
+	}
+
+	// Parse cells composed of dense vectors
+	template <typename DerivedV>
+	void parse_cell_index(const mxArray *prhs[], int index, EigenVectorList<DerivedV>& V) {
+		const mxArray *cell = prhs[index];
+		const mwSize *dims = mxGetDimensions(prhs[index]);
+		const mxArray *cellElement;
+		mwIndex jcell;
+		for (jcell = 0; jcell < dims[0]; jcell++) {
+			cellElement = mxGetCell(cell, jcell);
+			Eigen::VectorXx<DerivedV> element;
+			igl::matlab::parse_rhs_index(&cellElement, element);
+			V.emplace_back(element);
+		}
+	}
+
+	// Parse cells composed of sparse matrices
+	template <typename DerivedV>
+	void parse_cell_sparse(const mxArray *prhs[], int index, EigenSparseMatrixList<DerivedV>& V) {
+		const mxArray *cell = prhs[index];
+		const mwSize *dims = mxGetDimensions(prhs[index]);
+		const mxArray *cellElement;
+		mwIndex jcell;
+		for (jcell = 0; jcell < dims[0]; jcell++) {
+			cellElement = mxGetCell(cell, jcell);
+			Eigen::SparseMatrix<DerivedV> element;
+			igl::matlab::parse_rhs(&cellElement, element);
+			V.emplace_back(element);
+		}
+	}
+
+}
+
+#endif // SIM_PARSE_MATLAB_CELLS

### 2021-01-22 matlab/mex/vem_ext_force.cpp

@@ -1,0 +1,39 @@
+#include <mex.h>
+
+//matlab hooks from libigl (taken directly from gptoolbox)
+#include <igl/C_STR.h>
+#include <igl/matlab/mexErrMsgTxt.h>
+#undef assert
+#define assert( isOK ) ( (isOK) ? (void)0 : (void) mexErrMsgTxt(C_STR(__FILE__<<":"<<__LINE__<<": failed assertion `"<<#isOK<<"'"<<std::endl) ) )
+
+#include <igl/matlab/MexStream.h>
+#include <igl/matlab/parse_rhs.h>
+#include <igl/matlab/prepare_lhs.h>
+#include <igl/matlab/validate_arg.h>
+#include <igl/list_to_matrix.h>
+
+#include <parse_matlab_cells.h>
+
+#include <vem_ext_force.h>
+
+/* The gateway function */
+void mexFunction(int nlhs, mxArray *plhs[],
+                 int nrhs, const mxArray *prhs[]) {
+    /* variable declarations here */
+    Eigen::VectorXd g;
+
+    Eigen::VectorXd ext_force, mass;
+
+	EigenMatrixList<double> Y;
+	EigenSparseMatrixList<double> Y_S;
+
+	igl::matlab::parse_rhs_double(prhs+0, ext_force);
+	igl::matlab::parse_rhs_double(prhs+1, mass);
+	sim::parse_cell(prhs, 2, Y);
+	sim::parse_cell_sparse(prhs, 3, Y_S);
+	
+	//sim::vem3dmesh_neohookean_dq(g, x, c, volumes, params, dF_dc, dF_dc_S, ME, L, k, n, d, x0_coms_size, k_stability);
+	sim::vem_ext_force(g, ext_force, mass, Y, Y_S);
+
+	igl::matlab::prepare_lhs_double(g, plhs);
+}

### 2021-01-22 matlab/mex/vem_ext_force.m

@@ -1,0 +1,13 @@
+% VEM_EXT_FORCE Compute external forces for a specificed constant force
+% vector
+%
+% Inputs:
+%   ext_f   3 by 1 force vector
+%   mass    M by 1 list of per point masses
+%   Y       M by 3kn gradient of deformed positions with respect to their
+%                    polynomial coefficients.
+%
+% Outputs:
+%    g  3kn by 1 force vector for the set of coefficients where k is the
+%                size of the monomial basis and n is the number of parts.
+%

### 2021-01-22 matlab/vem_simulate_nurbs.m

@@ -70,10 +70,6 @@ function vem_simulate_nurbs(parts, varargin)
     params = [config.mu * 0.5, config.lambda * 0.5];
     params = repmat(params,size(V,2),1);
         
-    % Gravity force vector.
-  	%f_gravity = repmat([0 0 config.rho * config.gravity], size(x0,2),1)';
-    %f_gravity = config.dt*P*f_gravity(:);
-
     % Compute Shape weights
     [w, w_I] = nurbs_blending_weights(parts, V', config.distance_cutoff, ...
         'Enable_Secondary_Rays', config.enable_secondary_rays);
@@ -109,6 +105,10 @@ function vem_simulate_nurbs(parts, varargin)
     % projection operator (c are polynomial coefficients)
     [dF_dc, dF_dc_S] = vem_dF_dc(V, x0_coms, w, w_I, com_map, config.order, k);
 
+    % Gravity force vector.
+    dg_dc = vem_ext_force([0 0 config.gravity]', config.rho*vol, Y, Y_S);
+    f_gravity = config.dt*P*(L' * dg_dc);
+
     % Compute mass matrices
     ME = vem_error_matrix(Y0, Y0_S, L, d);
     M = vem_mass_matrix(Y, Y_S, L, config.rho .* vol);
@@ -141,8 +141,6 @@ function vem_simulate_nurbs(parts, varargin)
         % Force vector
         dV_dq = zeros(d*(k*n + size(x0_coms,2)),1);
         
-        dg_dc = zeros(d*(k*n + size(x0_coms,2)),1);
-
         % Computing force dV/dq for each point.
         % TODO: move this to C++ :)
         for i = 1:m
@@ -155,13 +153,8 @@ function vem_simulate_nurbs(parts, varargin)
             % Force vector
             dV_dF = neohookean_tet_dF(F, params(i,1), params(i,2));
             dV_dq = dV_dq +  dF_dc_S{i}' * dF_dc{i}' * dV_dF * vol(i);
-            
-            grav = [0 0 config.gravity];
-            dg_dc = dg_dc + (config.rho*vol(i)*grav*Y{i}*Y_S{i})';
-        end
+        end        
         dV_dq = L' * dV_dq;
-        
-        f_gravity = config.dt*P*(L' * dg_dc);
         
         % Error correction force
         f_error = - 2 * ME * x(:);

### 2021-01-22 matlab/vem_simulate_nurbs_fmincon.m

@@ -68,10 +68,6 @@ function vem_simulate_nurbs_fmincon(parts, varargin)
     % Lame parameters concatenated.
     params = [config.mu * 0.5, config.lambda * 0.5];
     params = repmat(params,size(V,2),1);
-        
-    % Gravity force vector.
-  	f_gravity = repmat([0 0 config.rho*config.gravity], size(x0,2),1)';
-    f_gravity = config.dt*P*f_gravity(:);
 
     % Compute Shape weights
     [w, w_I] = nurbs_blending_weights(parts, V', config.distance_cutoff, ...
@@ -108,6 +104,10 @@ function vem_simulate_nurbs_fmincon(parts, varargin)
     % projection operator (c are polynomial coefficients)
     [dF_dc, dF_dc_S] = vem_dF_dc(V, x0_coms, w, w_I, com_map, config.order, k);
 
+    % Gravity force vector.
+    dg_dc = vem_ext_force([0 0 config.gravity]', config.rho*vol, Y, Y_S);
+    f_gravity = config.dt*P*(L' * dg_dc);
+    
     % Compute mass matrices
     ME = vem_error_matrix(Y0, Y0_S, L, d);
     M = vem_mass_matrix(Y, Y_S, L, config.rho .* vol);

### 2021-01-22 matlab/vem_simulate_nurbs_newtons.m

@@ -59,7 +59,6 @@ function vem_simulate_nurbs_newtons(parts, varargin)
         V=x0;
         vol=ones(size(V,2),1);
     end
-    m = size(V,2);  % number of quadrature points
     
     if config.plot_points
         V_plot=plot3(V(1,:),V(2,:),V(3,:),'.','Color','m','MarkerSize',20);
@@ -68,10 +67,6 @@ function vem_simulate_nurbs_newtons(parts, varargin)
     % Lame parameters concatenated.
     params = [config.mu * 0.5, config.lambda * 0.5];
     params = repmat(params,size(V,2),1);
-        
-    % Gravity force vector.
-  	f_gravity = repmat([0 0 config.rho*config.gravity], size(x0,2),1)';
-    f_gravity = config.dt*P*f_gravity(:);
 
     % Compute Shape weights
     [w, w_I] = nurbs_blending_weights(parts, V', config.distance_cutoff, ...
@@ -107,6 +102,10 @@ function vem_simulate_nurbs_newtons(parts, varargin)
     % Computing each gradient of deformation gradient with respect to
     % projection operator (c are polynomial coefficients)
     [dF_dc, dF_dc_S] = vem_dF_dc(V, x0_coms, w, w_I, com_map, config.order, k);
+
+    % Gravity force vector.
+    dg_dc = vem_ext_force([0 0 config.gravity]', config.rho*vol, Y, Y_S);
+    f_gravity = config.dt*P*(L' * dg_dc);
 
     % Compute mass matrices
     ME = vem_error_matrix(Y0, Y0_S, L, d);

### 2021-01-22 matlab/vem_simulate_nurbs_with_collision.m

@@ -75,14 +75,7 @@ function vem_simulate_nurbs_with_collision(parts, varargin)
     % Lame parameters concatenated.
     params = [config.mu * 0.5, config.lambda * 0.5];
     params = repmat(params,size(V,2),1);
-        
-    % Gravity force vector.
-  	% f_gravity = repmat([0 0 config.rho*config.gravity], size(x0,2),1)';
-    % f_gravity = config.dt*P*f_gravity(:);
-    
-    % Undeformed Center of mass
-    x0_com = mean(x0,2);
-    
+
     % Compute Shape weights
     [w, w_I] = nurbs_blending_weights(parts, V', config.distance_cutoff, ...
         'Enable_Secondary_Rays', config.enable_secondary_rays);
@@ -118,6 +111,10 @@ function vem_simulate_nurbs_with_collision(parts, varargin)
     % projection operator (c are polynomial coefficients)
     [dF_dc, dF_dc_S] = vem_dF_dc(V, x0_coms, w, w_I, com_map, config.order, k);
 
+    % Gravity force vector.
+    dg_dc = vem_ext_force([0 0 config.gravity]', config.rho*vol, Y, Y_S);
+    f_gravity = config.dt*P*(L' * dg_dc);
+    
     % Compute mass matrices
     ME = vem_error_matrix(Y0, Y0_S, L, d);
     M = vem_mass_matrix(Y, Y_S, L, config.rho .* vol);
@@ -184,9 +181,7 @@ function vem_simulate_nurbs_with_collision(parts, varargin)
         
         % Force vector
         dV_dq = zeros(d*(k*n + size(x0_coms,2)),1);
-        
-        dg_dc = zeros(d*(k*n + size(x0_coms,2)),1);
-        
+                
         % Computing force dV/dq for each point.
         % TODO: move this to C++ :)
         for i = 1:m
@@ -200,12 +195,8 @@ function vem_simulate_nurbs_with_collision(parts, varargin)
             dV_dF = neohookean_tet_dF(F, params(i,1), params(i,2));
             dV_dq = dV_dq +  dF_dc_S{i}' * dF_dc{i}' * dV_dF * vol(i);
             
-            grav = [0 0 config.gravity];
-            dg_dc = dg_dc + (config.rho*vol(i)*grav*Y{i}*Y_S{i})';
         end
         dV_dq = L' * dV_dq;
-        
-        f_gravity = config.dt*P*(L' * dg_dc);
         
         % Error correction force
         f_error = - 2 * ME * x(:);

### 2021-01-22 vem_nurbs_with_collision.m

@@ -1,303 +1,0 @@
-function vem_nurbs
-    % Simulation parameters
-    dt = 0.01;          % timestep
-    %     C = 0.5 * 10000;     % Lame parameter 1
-    %     D = 0.5 * 150000;    % Lame parameter 2
-    %     gravity = -100;      % gravity force (direction is -z direction)
-    %     k_error = 100000;   % stiffness for stability term
-    %     order = 1;          % (1 or 2) linear or quadratic deformation
-    %     rho = .5;           % per point density (currently constant)
-    C = 0.5 * 5000;     % Lame parameter 1
-    D = 0.5 * 150000;   % Lame parameter 2
-    gravity = 0;        % no gravity
-    k_error = 100000;   % stiffness for stability term
-    order = 2;          % (1 or 2) linear or quadratic deformation
-    rho = .1;           % per point density (currently constant)
-    save_output = 0;    % (0 or 1) whether to output images of simulation
-    save_obj = 0;       % (0 or 1) whether to output obj files
-    d = 3;              % dimension (2 or 3)
-
-    % The number of elements in the monomial basis.
-    k = basis_size(d, order);
-    
-    % Read in NURBs 
-    fig=figure(1);
-    clf;
-    
-    % Read in file, generate coordinates and trimesh.
-    iges_file = 'rounded_cube.iges';
-%     iges_file = 'mug.iges';
-%     iges_file = 'rocket_with_nose.iges';
-%     iges_file = 'cylinder2.iges';
-%     iges_file = 'wrench_shrink.igs';
-%     iges_file = 'mug3.iges';
-
-    parts=nurbs_from_iges(iges_file);
-    parts=nurbs_plot(parts);
-    
-    % Assembles global generalized coordinates
-    [J, hires_J, q, E, x0] = nurbs_assemble_coords(parts);
-
-    % Initial deformed positions and velocities
-    x = x0;
-    qdot=zeros(size(q));
-    
-    % Setup pinned vertices constraint matrix
-    [~,I] = mink(x0(1,:),20);
-    pin_I = I(1:2);
-    P = fixed_point_constraint_matrix(x0',sort(pin_I)');
-    
-    % Plot all vertices
-    X_plot=plot3(x(1,pin_I),x(2,pin_I),x(3,pin_I),'.','Color','red','MarkerSize',20);
-    hold on;
-    
-    % Raycasting quadrature as described nowhere yet :)
[…]

### 2021-01-22 paper/sections/methods.tex
add section labels

@@
 
-Attempting to use $\Pmatx_i$ in~\refeq{shapematch} on a per-part basis is problematic as it can become ill-posed, especially for planar boundary representations. 
+Attempting to use $\Pmatxi$ in~\refeq{shapematch} on a per-part basis is problematic as it can become ill-posed, especially for planar boundary representations. 
 To alleviate this problem we invoke the \emph{hierarchical ordering principle}, which argues that the behavior of a system is typically 
@@
 
- \dave{heirarchical ordering to handle degenerecies}
-
 \subsection{Meshless Blending Weight Computation}
+\label{sec:weights}
 

### 2021-01-22 paper/sections/methods.tex
finally multipart projection

@@
 dominated by lower order effects~\cite{li2006regularities}. 
+Mechanically, this principle suggests that the amount of motion modelled by increasingly high-order polynomials is decreasing. 
+This implies that we can achieve acceptable shape matching results by performing shape matching hierarchically -- starting with the constant term and then 
+fitting the remaining deformation with progressively higher-order deformations.
+
+For a given part, $\prt_i$, with associated origin $\bar{\refX}_j$ we first estimate the constant term, $\tr_i$ as the centroid of all parts associated with $\bar{X}_j$.
+We next fit the linear part of the deformation by performing linear shape matching to $\prt_i$. 
+This can be repeated for polynomials of arbitrary high order, for instance here we apply this approach to quadratic shape matching:
+
+\begin{equation*}
+    \begin{split}
+        \tr_i  & = \sum_{\prt_k \in \bar{\refX}_j} \mathbbm{1}^T\wt_k \samp{\Jnurbs}_k\vc{q}_k \\
+        \vc{c}_i^1 & = \left(\samp{\Pmat^1}^T\wt_i\samp{\Pmat^1}\right)^{-1}\samp{\Pmat^1}^T\wt_i\left(\samp{\Jnurbs}_i\vc{q}_i-\mathbbm{1}\tr_i\right)\\
+        \vc{c}_i^2 & = \left(\samp{\Pmat^2}^T\wt_i\samp{\Pmat^2}\right)^{-1}\samp{\Pmat^2}^T\wt_i\left(\samp{\Jnurbs}_i\vc{q}_i-\samp{\Pmat^1}\vc{c}_i^1 - \mathbbm{1}\tr_i\right)
+    \end{split},
+    \label{eq:hfit} 
+\end{equation*} where $\Pmat^l$ contains only the monomial basis of order $l$,  $\samp{\Pmat^1}$ is the stacked evaluation of this matrix at $n$ quadrature points and $\vc{c}^l_i$ are the corresponding coefficients.
+Finally, $\wt_i$ is the diagonal integration weight matrix and $\mathbbm{1}\in\real^{3n\times3}$ is a matrix of stacked $3\times3$ identity matrices -- one for each quadrature point.
+
+Coupling via the deformation origins ensures our method will reproduce rigid and linear deformations in regions around these points, while higher order 
+terms help compensate for more local deformations.
+
+\dave{should have done Chebyshev shape matching to ensure higher order terms are orthogonal which would gaurantee exact reconstruction}
 

### 2021-01-23 paper/sections/methods.tex
performance table
multipart projection finally done, yeesh

@@
 
-\begin{figure}
-    \includegraphics[width=\columnwidth]{example-image-a}
-    \caption{Figure showing input model NURBS parts and shape matching to each part}
-    \label{fig:shapematching}
-\end{figure}
-
 \subsection{Shape Matching}
@@
 
+\begin{figure}[h]
+    \includegraphics[width=\columnwidth]{figures/projection_operator_solve}
+    \caption{A simple, four part model with two deformation origins and the corresponding matrix equation for the quadratic coefficients.}
+    \label{fig:multiparts}
+\end{figure}
+
 Attempting to use $\Pmatxi$ in~\refeq{shapematch} on a per-part basis is problematic as it can become ill-posed, especially for planar boundary representations. 
@@
 
-\begin{equation*}
+\begin{equation}
     \begin{split}
-        \tr_i  & = \sum_{\prt_k \in \bar{\refX}_j} \mathbbm{1}^T\wt_k \samp{\Jnurbs}_k\vc{q}_k \\
-        \vc{c}_i^1 & = \left(\samp{\Pmat^1}^T\wt_i\samp{\Pmat^1}\right)^{-1}\samp{\Pmat^1}^T\wt_i\left(\samp{\Jnurbs}_i\vc{q}_i-\mathbbm{1}\tr_i\right)\\
-        \vc{c}_i^2 & = \left(\samp{\Pmat^2}^T\wt_i\samp{\Pmat^2}\right)^{-1}\samp{\Pmat^2}^T\wt_i\left(\samp{\Jnurbs}_i\vc{q}_i-\samp{\Pmat^1}\vc{c}_i^1 - \mathbbm{1}\tr_i\right)
+        \tr_i  & = \sum_{\prt_k \in \bar{\refX}_j} \underbrace{\mathbbm{1}^T\wt_k \samp{\Jnurbs}_k}_{B_{ik}}\vc{q}_k \\
+        \vc{c}_i^1 & = \underbrace{\left(\samp{\Pmat^1}^T\wt_i\samp{\Pmat^1}\right)}_{A^1_i}^{-1}\bigg(\underbrace{\samp{\Pmat^1}^T\wt_i\samp{\Jnurbs}_i}_{B^1_i}\vc{q}_i-\underbrace{\samp{\Pmat^1}^T\wt_i\mathbbm{1}}_{T^1_i}\tr_i\bigg)\\
+        \vc{c}_i^2 & = \underbrace{\left(\samp{\Pmat^2}^T\wt_i\samp{\Pmat^2}\right)}_{A^2_i}^{-1}\bigg(\underbrace{\samp{\Pmat^2}^T\wt_i\samp{\Jnurbs}_i}_{B^2_i}\vc{q}_i-\underbrace{\samp{\Pmat^2}^T\wt_i\samp{\Pmat^1}}_{D^2_i}\vc{c}_i^1 - \underbrace{\samp{\Pmat^2}^T\wt_i\mathbbm{1}}_{T^2_i}\tr_i\bigg)
     \end{split},
     \label{eq:hfit} 
-\end{equation*} where $\Pmat^l$ contains only the monomial basis of order $l$,  $\samp{\Pmat^1}$ is the stacked evaluation of this matrix at $n$ quadrature points and $\vc{c}^l_i$ are the corresponding coefficients.
+\end{equation} where $\Pmat^l$ contains only the monomial basis of order $l$,  $\samp{\Pmat^1}$ is the stacked evaluation of this matrix at $n$ quadrature points and $\vc{c}^l_i$ are the corresponding coefficients.
 Finally, $\wt_i$ is the diagonal integration weight matrix and $\mathbbm{1}\in\real^{3n\times3}$ is a matrix of stacked $3\times3$ identity matrices -- one for each quadrature point.
 
+%%single projector for every object
+When assembled for all parts,~\refeq{hfit}, yields a block upper triangular matrix $U$ and a spare matrix $B$ ~(\reffig{multiparts}) which can be used to efficiently to compute the projection 
+operator for the entire object $\Pi = U^{-1}B\vc{q}$. 
+The structure of $U$ and $B$ ensures that $\Pi$ only couples objects which share deformation centers, which implies a sparse $\Pi$. 
+The per-part projection operators $\Pi_i$ correspond to blocks of rows of $\Pi$. 
 Coupling via the deformation origins ensures our method will reproduce rigid and linear deformations in regions around these points, while higher order 
@@
 %%%% OLD STUFF BELOW HERE  %%%%%
-In a typical simulation we will have multiple shape matching elements where the boundary of each element is defined by a different NURBS surface.
-Therefore for the entire model each shape has a boundary, a polynomial describing its deformation, as well as an associated center of mass.
-These centers of mass may not be distinct, meaning that a single center of mass may be shared among multiple surfaces.
-This sharing of center of masses is typical and in some cases required (\myworries{reference center of mass section that describes how for planar surfaces we need to share COM to get a COM out of plane}).
-This leads to a new definition of the center of mass where it is the mean of all the boundary points associated with it.
-
-In defining the shape matching fitting problem, we will form a system similar to equation \ref{eqn:fitting_one}.
-In this case where we have $n$ shape elements our set of unknown coefficients is $\hat{\mathbf{c}} = \left[\mathbf{c}_1^T \dots \mathbf{c}_i^T \dots \mathbf{c}_n^T \right]^T \in \mathbb{R}^{3kn}$ where $\mathbf{c}_i$ corresponds to the unknown polynomial coefficients for the $i$-th shape element.
-The set of boundary positions is similarly written as $\hat{\mathbf{b}}=\left[\mathbf{b}_1^T \dots \mathbf{b}_i^T \dots \mathbf{b}_n^T \right]^T \in \mathbb{R}^{3N}$ where $N$ is the total number of boundary points across all elements.
-If we have $m$ centers of mass we write the set of polynomial constant terms as $\hat{\mathbf{t}} = \left[\mathbf{t}_1^T \dots \mathbf{t}_i^T \dots \mathbf{t}_m^T \right]^T \in \mathbb{R}^{3m}$.
-With these terms, we write the polynomial fitting system for the entire model as
-\begin{equation}
-\begin{pmatrix}
- \mathbf{A}_1 &      &			  &			&				&\mathbf{S}_1\\  
- &            \ddots &            &  		&				&\vdots\\ 
- &            &      \mathbf{A}_i &   		&				&\mathbf{S}_i\\
- &            &      &            \ddots	&				&\vdots\\
- &            &      &            &			\mathbf{A}_n	&\mathbf{S}_n
-\end{pmatrix}
-\begin{bmatrix}
- \hat{\mathbf{c}}  \\ 
- \hat{\mathbf{t}}   
-\end{bmatrix}
-=
-\hat{\mathbf{b}}
-\text{,}
-\end{equation}
-where $\mathbf{A}_i$ is the monomial basis matrix for the $i$-th shape element, and $\mathbf{S}_i$ is a selection matrix that extracts the $j$-th center of mass which is associated with the $i$-th shape.
-We simplify this as
-\begin{equation}
-\begin{pmatrix}
-\hat{\mathbf{A}} & \hat{\mathbf{S}}
-\end{pmatrix}
-\begin{bmatrix}
- \hat{\mathbf{c}}  \\ 
- \hat{\mathbf{t}}   
-\end{bmatrix}
-=
-\hat{\mathbf{b}}
-\text{.}
-\end{equation}
-Like in the single element case, we set the $\hat{\mathbf{t}}$ to be equal to the mean of a set of boundary positions, but in this case each $\mathbf{t}_j$ is equal to the mean of only the boundary positions associated with the $j$-th center of mass.
-The definition of the values of $\hat{\mathbf{t}}$ is similar to \ref{eqn:t_solution}:
-\begin{equation}
-\hat{\mathbf{t}} =
-\underbrace{
-\begin{bmatrix}
-\mathbf{T}_1 \\ 
-\vdots       \\
-\mathbf{T_m}
-\end{bmatrix}}_{\hat{\mathbf{T}}}
-\text{,}
-\end{equation}
-
-where the entries each $\mathbf{T}_i \in \mathbb{R}^{3N}$ take a similar form, but are only nonzero for columns corresponding to associated boundary vertices.
-
-Finally, to solve for the full set of non-constant term coefficients, $\hat{\mathbf{c}}$, we solve in the same manner as equation \ref{eqn:c_one_element} where we have
-\begin{equation}
-\label{eqn:c_multiple_elements}
-\begin{split}
-\hat{\mathbf{c}} & = (\hat{\mathbf{A}}^T\hat{\mathbf{A}})^{-1}\hat{\mathbf{A}}^T(\mathbf{I-\hat{S}\hat{T})\hat{\mathbf{b}}} \\
-           & = \mathbf{L}\hat{\mathbf{b}}
-\end{split}
-\text{.}
-\end{equation}
-
-\dave{this is good but we shouldn't imply we solve the system above since we do Hierarchical ordering}
-\paragraph*{Hierarchical fitting}
-A problem may emerge in the shape matching process where the system may be underdetermined.
-To account for this we solve for the coefficients in a hierarchical manner where we initially solve for the linear terms and at each step solve the coefficients for the next higher order of deformation.
-The produces a set of smaller systems which are guaranteed to be determined \myworries{probably need to justify this more}.
-The intuition behind this can be described by the \textit{hierarchical ordering principle}, which argues that the behavior of a system is typically dominated by lower order effects \myworries{cite something here.}.
-To describe our fitting strategy, we begin with the following shape matching system for a single element
-\begin{equation}
-\label{eqn:fitting_one_copy}
-\left[
-\mathbf{A} \quad \mathbf{S}
-\right]
-\left[ \begin{array}{c}
-\mathbf{c} \\
-\mathbf{t} \\
-\end{array} \right] = \mathbf{b}
-\text{.}
-\end{equation}
-We will show the hierarchical fitting on a polynomial with a set of coefficients for the linear terms, $\mathbf{c}_{\text{lin}}$, and the coefficients for the quadratic terms $\mathbf{c}_{\text{quad}}$.
-Similarly, we also have $\mathbf{A}_{\text{lin}}$ and $\mathbf{A}_{\text{quad}}$ which are the blocks of $\mathbf{A}$ corresponding to the columns for the linear and quadratic monomial terms, respectively.
-First we construct a system for the linear terms:
-\begin{equation}
-\label{eqn:fitting_linear}
-\left[
-\mathbf{A}_{\text{lin}} \quad \mathbf{S}
-\right]
-\left[ \begin{array}{c}
-\mathbf{c}_{\text{lin}} \\
-\mathbf{t} \\
-\end{array} \right] = \mathbf{b}
-\text{,}
-\end{equation}
-which leads to the following expression for the linear coefficients
-\begin{equation}
-\label{eqn:c_linear}
-\begin{split}
-\mathbf{c}_{\text{lin}} & = (\mathbf{A}_{\text{lin}}^T\mathbf{A}_{\text{lin}})^{-1}\mathbf{A}_{\text{lin}}^T(\mathbf{I-ST})\mathbf{b} \\
-           & = \mathbf{L}_{\text{lin}}\mathbf{b}
-\end{split}
-\text{.}
-\end{equation}
-where $\mathbf{T}$ is defined in equation \ref{eqn:t_solution}. For the quadratic system, we subtract linear term estimate of the solution and solve for the coefficients that fit this residual:
-\begin{equation}
-\label{eqn:fitting_quadratic}
-\left[
-\mathbf{A}_{\text{quad}} \quad \mathbf{S}
-\right]
-\left[ \begin{array}{c}
-\mathbf{c}_{\text{quad}} \\
-\mathbf{t} \\
-\end{array} \right] = \mathbf{b} - \mathbf{A}_{\text{lin}}\mathbf{c}_{\text{lin}}
-\text{.}
-\end{equation}
-Using the solution for the linear coefficients in \ref{eqn:c_linear}, we arrive at the solution for the quadratic coefficients:
-\begin{equation}
-\label{eqn:c_quad}
-\begin{split}
-\mathbf{c}_{\text{quad}} & = (\mathbf{A}_{\text{quad}}^T\mathbf{A}_{\text{quad}})^{-1}\mathbf{A}_{\text{quad}}^T((\mathbf{I-ST})\mathbf{b} - \mathbf{A}_{\text{lin}}\mathbf{c}_{\text{lin}}) \\
-		   & = (\mathbf{A}_{\text{quad}}^T\mathbf{A}_{\text{quad}})^{-1}\mathbf{A}_{\text{quad}}^T(\mathbf{I-ST} - \mathbf{A}_{\text{lin}}\mathbf{L}_{\text{lin}})\mathbf{b} \\
-           & = \mathbf{L}_{\text{quad}}\mathbf{b}
-\end{split}
-\text{.}
-\end{equation}
-The final form for the full $\mathbf{L}$ matrix is thus given by
-\begin{equation}
-\mathbf{L} =
-\begin{bmatrix}
-\mathbf{L}_{\text{lin}} \\
-\mathbf{L}_{\text{quad}} \\
-\mathbf{T}
-\end{bmatrix}
-\end{equation}
-
-The recursive nature of this process permits a more general description for arbitrary orders, which we write in algorithm \myworries{TODO}.
-
 \subsection{Blending Weights}

### 2021-01-23 paper/sections/methods.tex
blah

@@
     \defX\left(\refX\right) = \sum_{\prt_i\in\Prts} \bld_i\Pmatxi\Pi_i\vc{q}_i, 
+    \label{eq:kinematics}
  \end{equation} where $w_i$ $\Pmatx_i$, $\Pi_i$ and $\vc{q}_i$ are the blending weights, polynomial basis, projection operator and control points for the $i^{th}$
@@
 
+\subsection{Choosing Deformation Centers}
+\label{sec:origins}
+\dave{Ty: try to follow this formula when writing 
+(1) what is the purpose of this part of the method ?
+(2) what are the criteria for success
+(3) what is our assumed input 
+(4) what is the desired output
+(5) what's the algorith,
+(6) add proofs/diagrams/figures that show we satisfy (2). }
+
 \subsection{Meshless Blending Weight Computation}
 \label{sec:weights}
-
-\subsection{Choosing Deformation Centers}
-\label{sec:origins}
-
-%%%% OLD STUFF BELOW HERE  %%%%%
-\subsection{Blending Weights}
 \dave{Desireable Properties}
@@
 \dave{maybe but weight figure here so reviewers immediately know that it looks pretty good}
+\dave{need to mention Monte Carlo Geometry processing but can't use it because our parts can overlap and that's a tricky boundary condition}
+
 All that remains for the deformation map is to compute blending weights.
@@
 
-\subsection{Multiple Part Projection}
-\dave{explain why you need to share constant terms (fit polynomial around a point in space), explain clustering and how to pick, forward reference to quadrature point}
 \subsection{Quadrature Points}

### 2021-01-23 paper/sections/results.tex
performance table
add CPU column to table
blah

@@
 %
-\section{Results}
+\section{Results and Discussion}
 
 \dave{we should show the nurbs model in rhino/fusion for every example, maybe an exploded view as well}
+\dave{mention that not so many standard NURBS models so we modelled lots of these ourselves .. show's power of approach that we can model and sim all these examples.}
+\dave{use as many different meshes as possible for figures (i.e try not to have the same mesh show up in more than one figure if possible).}
 
-\dave{use as many different meshes as possible for figures (i.e try not to have the same mesh show up in more than one figure if possible).}
+\begin{table*}[h]
+  \caption{Performance of the Shape Matching Element Method on all Examples.}
+  \label{tbl:compare}
+  \begin{center}
+  \begin{tabular}{l c c c c c c c c}
+   \textbf{Example} & \textbf{CPU} & $|\vc{q}|$ &  \textbf{Sample Model} & \textbf{Quadrature} & \textbf{Weights} & \textbf{Build $\Pi$} & \textbf{Time Step} \\
+   \hline 
+   \rowcolor[HTML]{DAE8FC} 
+   \textbf{Cantilever}  & ? & ? & ? & ? & ? & ? & ? & ? \\
+   \textbf{Rocket}      & ? & ? & ? & ? & ? & ? & ? & ? \\
+   \hline
+  \end{tabular}
+  \end{center}
+
+  \end{table*}
 

### 2021-01-23 paper/sections/methods.tex
Added raycasting and initial patch test figure
working on blending function section

@@
 
-All that remains for the deformation map is to compute blending weights.
+Our kinematic map~\refeq{kinematics} requires blending weights in order to homogenize our per-part representation. 
+Specifying weights manually would be antithetical to our goal of providing a seamless translation from surface representation to simulation.
+Automatic  weight computation is a well-studied topic (for examples ~\citet{10.1145/2766952,BBW:2011,0.1145/2010324.1964968}) but typically requires solving an optimization problem
+that itself relies on a volumetric discretization. 
+This can be avoided by using stochastic methods~\cite{Sawhney:2020:MCG} but do not yet support the jump boundary conditions we require for our problem.
+
+%All that remains for the deformation map is to compute blending weights.
 Our choice of the weighting function is guided by the simple idea that that for a given material point its weight should be highest to whichever surfaces it is closest.
@@
 \begin{figure}
-    \includegraphics[width=\columnwidth]{example-image-a}
+    \includegraphics[width=\columnwidth]{figures/raycasting_quadrature}
     \caption{Figure showing how raycasting quadrature works (top), convergence plot for volume integration (bottom}

### 2021-01-23 paper/sections/methods.tex
more blending weights

@@
 Specifying weights manually would be antithetical to our goal of providing a seamless translation from surface representation to simulation.
-Automatic  weight computation is a well-studied topic (for examples ~\citet{10.1145/2766952,BBW:2011,0.1145/2010324.1964968}) but typically requires solving an optimization problem
+Automatic  weight computation is a well-studied topic (for examples ~\citet{10.1145/2766952,BBW:2011} and ~\citet{10.1145/2010324.1964968}) but typically requires solving an optimization problem
 that itself relies on a volumetric discretization. 
 This can be avoided by using stochastic methods~\cite{Sawhney:2020:MCG} but do not yet support the jump boundary conditions we require for our problem.
+
+Ideally our shape functions would both partition unity and obey the Kronecker delta property (the $i^th$ blending function is one at $i^th$ part and zero on all other parts).
+These properties ensure that our simulation can properly represent motions up to the order of the polynomial basis and that the solution stored at each part is the actual deformed position.
+Additionally we would like our weights to be sparse both for performance reasons (leads to sparse matrices) and for expressivity (leads to locality of deformation).
+
+While a partition of unity and sparsity are straightforward to enforce, the overlapping and intersecting parts~(\reffig{badmodels}) often encountered in non-engineering models lead
+to a contradiction -- how can the blending functions be both one and zero at the points of overlap ? 
+Our solution is to allow weights to include bounded discontinuities at these points and it is these discontinuous boundary conditions that are difficult to enforce 
+with previous methods. 
+\dave{ideally need a didactic figure showing this}
 

### 2021-01-23 paper/sections/methods.tex
blending weights section done

--- 2021-01-23 paper/sections/methods.tex
@@
 \dave{Ty: try to follow this formula when writing 
+\dave{maybe this should come after weight computation ?}
 (1) what is the purpose of this part of the method ?
@@
 \label{sec:weights}
-\dave{Desireable Properties}
-\dave{Construction via post normalization}
-\dave{only need blending weights at quadrature points and surface samples so given sparse set of quadrature points compute blending weights using raycasting}
-\dave{things to explain (1) raycasting works for overlapping geometric, seperated geometry (2) weights glue objects together}
-\dave{maybe but weight figure here so reviewers immediately know that it looks pretty good}
-\dave{need to mention Monte Carlo Geometry processing but can't use it because our parts can overlap and that's a tricky boundary condition}
+% \dave{Desireable Properties}
+% \dave{Construction via post normalization}
+% \dave{only need blending weights at quadrature points and surface samples so given sparse set of quadrature points compute blending weights using raycasting}
+% \dave{things to explain (1) raycasting works for overlapping geometric, seperated geometry (2) weights glue objects together}
+% \dave{maybe but weight figure here so reviewers immediately know that it looks pretty good}
+% \dave{need to mention Monte Carlo Geometry processing but can't use it because our parts can overlap and that's a tricky boundary condition}
 
-All that remains for the deformation map is to compute blending weights.
-Our choice of the weighting function is guided by the simple idea that that for a given material point its weight should be highest to whichever surfaces it is closest.
-This idea lends itself to a simple distance-based weighting method.
-The first step in this process is to compute \textit{distance weights}.
-For a given point $\mathbf{X}$ and a set of $n$ we wish to compute a set of distance weights to each surface $\mathbf{\theta}(\mathbf{X}) = \left(\theta_1, \dots, \theta_i, \dots, \theta_n \right)$.
+Our kinematic map~(\refeq{kinematics}) requires blending weights in order to homogenize our per-part representation. 
+Specifying weights manually would be antithetical to our goal of providing a seamless translation from surface representation to physics-based animation.
+Automatic  weight computation is a well-studied topic (see ~\citet{10.1145/2766952,BBW:2011} and ~\citet{10.1145/2010324.1964968}) but typically requires solving 
+an optimization problem that itself relies on a volumetric discretization. 
+This can be avoided by using stochastic methods~\cite{Sawhney:2020:MCG} but these do not yet support the boundary conditions we require for our problem.
 
-For the $i$-th surface, we first find the distance, $d_{\text{primary}}$, from the point $\mathbf{X}$ to the closest point, $\mathbf{p}$, on the surface.
-Next we wish to satisfy the condition that the distance weight is highest on the surface and it decays towards zero as it approaches other surfaces.
-To achieve this, we shoot a ray from $\mathbf{p}$ towards $\mathbf{X}$ and check for intersections between this ray and the other $n-1$ surfaces.
-If this ray hits another surface, we let $d_{\text{total}}$ be the distance from $\mathbf{p}$ to the closest ray hit $\mathbf{q}$.
-These two distances enable us to write a simple linear function to compute the \textit{distance weight}:
+Ideally our shape functions would both partition unity and obey the Kronecker delta property (the $i^th$ blending function is one at $i^th$ part and zero on all other parts).
+These properties ensure that our simulation can properly represent translation and that the solution stored at each part is the actual deformed position.
+Additionally we would like our weights to be sparse both for performance reasons (leads to sparse matrices) and for expressivity (leads to locality of deformation).
+
+While partition of unity and sparsity are straightforward to enforce, the overlapping and intersecting parts~(\reffig{badmodels}) often encountered in non-engineering models leads
+to a contradiction -- how can the blending functions be both one and zero at the points of overlap ? 
+Our solution is to allow weights to include bounded discontinuities at these points and it is these discontinuous boundary conditions that are difficult to enforce 
+with previous methods. 
+\dave{ideally need a didactic figure showing this}
+
+Given a set of reference space points $\st{X}$ inside our model, we want to evaluate $\bld_i\left(\refX_j\right)$, where $\bld_i$ is the weight function associated with 
+the $i^{th}$ part, $\prt_i$, and $\refX_j\in\st{X}$.
+
+For each $\refX_j$ we find the nearest point on $\prt_i$, $\clst$.
+We cast a ray from $\clst$, towards $\refX_j$ and find the intersection with the nearest part, $\rhit$.
+
+\dave{Ty: How do we deal with non-planar parts ? What if the ray cast hits the same part that the closest point is on?}
+
+Next we compute two distances, $d_{\text{primary}}$, the distance from $\mathbf{X}_j$ to $\clst$ and $d_{\text{total}}$ the distance from  $\clst$ to $\rhit$.
+Using these two distances we define a linear \emph{distance weight} using the well-known ReLU function:
 \begin{equation}
-\theta_i = \max (1.0 - \frac{d_{\text{primary}}}{\min (d_{\text{total}}, D)}, 0.0)
+\theta_i = \max (1.0 - \frac{d_{\text{primary}}}{\min (d_{\text{total}}, \alpha)}, 0.0)
 \text{,}
 \end{equation}
-where $D$ is a fixed cutoff distance parameter.
-We see that this distance weight is the result of interpolating along this ray from $\mathbf{p}$ to $\mathbf{q}$ (or a fixed point that depends on $D$).
-Importantly, this gives a result where the distance weight to surface $i$ is equal to $1$ if it lies on $i$ and $0$ if it lies on another surface.
-More elaborate schemes may be developed for computing distances, such as using the Boundary Element Method, but we find this simple linear function to perform well for our purposes.
+where $\alpha$ is a distance cutoff parameter. 
+This distance weight~\reffig{plot_distance_weight} is the result of interpolating along this ray from $\clst$ to $\rhit$ (or a fixed point that depends on $\alpha$). 
+These weights obey the Kronecker delta property for all parts but does not partition unity.
 
-\begin{figure}
-    \includegraphics[width=\columnwidth]{example-image-a}
-    \caption{Figure showing how weight calculation works (left) ray casting plus result (right) after correcting for partition of unity}
-    \label{fig:weightcompute}
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=3.33in]{figures/plot_distance_weight.pdf}
+    \caption{We plot the distance weight with respect to the primary and total ray length. Here the cutoff distance is 50.
+    }
+    \label{fig:plot_distance_weight}
+    \vspace{-5pt}
   \end{figure}
 
-Now that we have a set of distances weights $\mathbf{theta}$, we seek to form a final set of \textit{blending weights} $\mathbf{w}(\mathbf{X}) =  \left( w_1, \dots, w_i, \dots, w_n \right)$ to blend the polynomials of the $n$ shape elements.
-In constructing these blending weights, we require that they adhere to a set of conditions.
-We require that (1) each weight is non-negative, (2) the sum of the weights is equal to 1 (\textit{partition of unity}), (3) and that if a point lies on a surface $i$, it will have a weight of 1 to $i$ and consequently zero to the others.
-To produce these blending weights we form a quadratic program that satisfies these constraints:
+%\begin{figure}
+%    \includegraphics[width=\columnwidth]{example-image-a}
+%    \caption{Figure showing how weight calculation works (left) ray casting plus result (right) after correcting for partition of unity}
+%    \label{fig:weightcompute}
+%\end{figure}
+
+Next we use these distance weights to compute our final blending weights.
+At each $\refX_j$ we want to compute $\mathbf{\bld}(\mathbf{X}_j) =  \left( \bld_1(\mathbf{X}_j), \dots, \bld_n(\mathbf{X}_j) \right)$, where 
+$\sum_{i=1}^n\bld_i(\mathbf{X}_j) = 1$, and each $\bld_i \geq 0$, which we enforce by solving a local optimization problem:
+
 \begin{equation}
 \begin{aligned}
-\mathbf{w}(\mathbf{X}) = \min_{\mathbf{w}} \quad & \mathbf{w}^T \Theta(\mathbf{X}) \mathbf{w}    \\
-\textrm{s.t.} \quad & 0 \leq w_i \leq 1                     \\
+\mathbf{w}(\mathbf{X}_j) =  \arg\min_{\mathbf{w}} \quad & \mathbf{w}^T \Theta(\mathbf{X}_j) \mathbf{w}    \\
+\textrm{st} \quad & 0 \leq w_i \leq 1 \mbox{ }\forall i                    \\
                   &   \sum_i^n w_i = 1                      \\
@@
 
-where
-\begin{equation}
-     \Theta(\mathbf{X}) = \text{diag}\left( \frac{1}{\theta_1},\frac{1}{\theta_2},\dots,\frac{1}{\theta_n}\right)
-\end{equation}
-We see that through the use of linear constraints we satisfy conditions (1) and (2).
-Satisfying the third condition (3) is achieved through using the distance weights.
-Intuitively, we see that for some small $\theta_i$, $\frac{1}{\theta_i}$ will be very large, encouraging the quadratic program to assign a small weight for $w_i$.
-Conversely, we see the opposite effect for large values of $\theta_i$.
+where $\Theta(\mathbf{X}) = \text{diag}\left( \frac{1}{\theta_1},\frac{1}{\theta_2},\dots,\frac{1}{\theta_n}\right)$.
+
+This post-process ensures that our blending weights partition unity and are non-negative~(\reffig{distance_weight_puft}). 
+For any point on a part that is non-intersecting, the weights will satisfy the Kronecker delta property since the inverse distance weights imply that only
+a single theta will have a non-infinite weight. 
+For points that lie on more than one part, the weights associated with the intersecting parts will be equal and all others will be zero
+In other words, the deformed positions of these intersecting points will be an average, which is a plausible solution and ensures continuity.
+The blending weights immediately jump back to one and zero for non-intersecting points on each part. 
+Finally, the distance cutoff $\alpha$ allows us to control the sparsity of the blending weights.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=3.33in]{figures/distance_weight_puft.pdf}
+    \caption{Our blending weights decay smoothly from 1.0 (yellow) to 0.0 (green) when moving away from its closest surface.
+Here we visualize the distribution of the distance weights (with cutoff distance 5.0) corresponding to each part.
+    }
+    \label{fig:distance_weight_puft}
+    \vspace{-5pt}
+  \end{figure} 
+  %
+  %
+
+Using raycasting to compute blending weights imbues our method with similar advantages to \citet{Sawhney:2020:MCG} (allowing us to 
+handle intersecting geometry, geometry with gaps and to perform constructive solid geometry operations on the fly),
+but allows us to use more general weighting functions and apply the appropriate behavior for intersecting parts.
+The price we pay for these advantages is that our weights are not C0 continuous but rather Lipchitz smooth.
+Perturbations in the boundaries of parts can cause bounded jumps in our distance weights. 
+However these jumps occupy an infinitesimal percentage of the volume of our object are easy to avoid when integrating
+physical quantities, leaving our simulation unaffected.
 

### 2021-01-23 paper/sections/results.tex
Added raycasting and initial patch test figure
added initial mug figure. needs work
blending weights section done

@@
 \dave{we should show the nurbs model in rhino/fusion for every example, maybe an exploded view as well}
+
 \dave{mention that not so many standard NURBS models so we modelled lots of these ourselves .. show's power of approach that we can model and sim all these examples.}
+
 \dave{use as many different meshes as possible for figures (i.e try not to have the same mesh show up in more than one figure if possible).}
+
+\dave{we should report which models we made ourselves}
 
@@
 % weights are good
-\begin{figure}
-  \centering
-  \includegraphics[width=3.33in]{figures/distance_weight_puft.pdf}
-  \caption{Our distance weights decay smoothly from 1.0 (yellow) to 0.0 (green) when moving away from its closest surface.
+% \begin{figure}
+%   \centering
+%   \includegraphics[width=3.33in]{figures/distance_weight_puft.pdf}
+%   \caption{Our distance weights decay smoothly from 1.0 (yellow) to 0.0 (green) when moving away from its closest surface.
 Here we visualize the distribution of the distance weights (with cutoff distance 5.0) corresponding to each part.
-  }
-  \label{fig:distance_weight_puft}
-  \vspace{-5pt}
-\end{figure} 
-%
-%
-\begin{figure}
-  \centering
-  \includegraphics[width=3.33in]{figures/plot_distance_weight.pdf}
-  \caption{We plot the distance weight with respect to the primary and total ray length. Here the cutoff distance is 50.
-  }
-  \label{fig:plot_distance_weight}
-  \vspace{-5pt}
-\end{figure}
+%   }
+%   \label{fig:distance_weight_puft}
+%   \vspace{-5pt}
+% \end{figure} 
+% %
+% %
+% \begin{figure}
+%   \centering
+%   \includegraphics[width=3.33in]{figures/plot_distance_weight.pdf}
+%   \caption{We plot the distance weight with respect to the primary and total ray length. Here the cutoff distance is 50.
+%   }
+%   \label{fig:plot_distance_weight}
+%   \vspace{-5pt}
+% \end{figure}
 
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{example-image-a}
+  \includegraphics[width=\columnwidth]{figures/patch_test.pdf}
   \caption{2D examples (left) rigidly transform boundary and show that solving the static problem gives us rigid motion on the interior (right) isotropically scale boundary and show solving static problem gives us constant strain on the interior}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{example-image-a}
+  \includegraphics[width=\columnwidth]{figures/mug_overlap}
   \caption{Coffee mug with overlapping handle (top) row of models with increasing overlap (middle) single weight image showing behaviour in the overlapping region, (bottom) simulation result for each one}

### 2021-01-23 paper/sections/methods.tex
paper in proper format

@@
 
-\subsection{Quadrature Points}
-\dave{quadrature via raycasting}
+\subsection{Meshless Integration}
+
+% intege
+
+%%%% OLD STUFF DOWN HERE %%%%
 The last step before we may discuss the dynamic simulation is to describe our strategy for generating material points $\mathbf{X}$ in the undeformed model.

### 2021-01-23 paper/sections/methods.tex
integration section

@@
 
-% intege
+We integrate physical quantities over our undeformed object using raycasting based quadrature~\cite{KHOSRAVIFARD201030}.
 
-%%%% OLD STUFF DOWN HERE %%%%
-The last step before we may discuss the dynamic simulation is to describe our strategy for generating material points $\mathbf{X}$ in the undeformed model.
-Our volume integration strategy is based on \cite{KHOSRAVIFARD201030} which we briefly describe in the following section
+\dave{Ty: a bit more here, show the transformation of the integral via divergence theorem then show the final discrete version. Clearly define the quadrature rules.}
 
-This paper addresses the integration of volume integrals of the form $\mathbf{I} = \int_\Omega f(\mathbf{x}) d\Omega$ where $\mathbf{x} = (x_1,x_2,x_3)$.
-The authors show that through application of the Green-Gauss theorem, this integral may be rewritten as 2D integral over a $yz$ plane where at each position along the plane we evaluate a line integral along the $x$-axis.
-To evaluate these line integrals, a raycasting procedure is performed to find a set of intersection along the line.
-These intersection define integration domains in which we sample 3D integration points.
-As a result of integrating over the 2D plane and raycasting at each 2D point, we produce an integration rule over the entire domain that can converge to the true integral value.
-
-In our approach we modify the raycasting procedure by allowing integration over intersecting NURBS surfaces.
-It is common in NURBS modeling to produce models in which many surfaces overlap.
-To account for this we augment the raycasting intersection step to use an approach common in Constructive Solid Geometry (CSG).
-For the set of intersection intervals, we instead take the union over all intervals to yield a new set of intervals handling intersecting solids.
-We find that this extension serves as a simple and practical approach to support inexact models (\myworries{not sure what to right here}).
-We note that this solution is not perfect and making this intersection step more robust serves as future work.
-
+We make a minor modifiction this procedure to handle common geometric pathologies encountered in NURBS modeling, such as overlapping surfaces by performing 
+CSG unions while raycasting\dave{find a citation}. 
+The output of this raycasting procedure is a set of quadrature points $\st{X}$ and weights $\st{v}$ that we use for integration and also as input to our 
+deformation origin extraction~\refsec{origins} and blending weights computation~\refsec{weights}
+\reffig{raycasting} shows the convergence of this quadrature approach when used to compute the volume of a NURBS model.
 \begin{figure}

### 2021-01-23 paper/sections/methods.tex
first pass done

@@
 
-\section{Methods}
+%\section{Methods}
+\section{Geometric Model}
 
@@
 
-\subsection{Shape Matching}
+\section{Shape Matching}
 \label{sec:shapematching}
@@
 
-\subsection{Shape Matching Element Method}
+\section{Shape Matching Element Method}
  Each input object is composed of a set of boundary representations, $\Prts$, one for each part $\prt\in\Prts$. 
@@
 
- \subsubsection*{Multiple Part Projection}
+ \subsection*{Multiple Part Projection}
 As in ~\refsec{shapematching} we construct each $\Pi_i$ via shape matching. 
@@
 
+\section{Physics-Based Animation}
+
 \subsection{Mass Matrix}
@@
 The second line is a result of the fact that all terms except for $\mathbf{Y(X)}$ are spatially constant, allowing us to move these outside the volume integral, which enables fast assembly of the mass matrix, $\mathbf{M}$.
+
 \subsection{Error Energy}
+\dave{We should cite VEM literature and point out this is a standard thing they do to make the algorithm consistent and convergent}
 Over the course of the simulation we may see the true boundary values deviate from their positions estimated by the polynomial.
@@
 \subsection{Deformation Gradient}
+\dave{gotta deal with this weight gradient}
 At this point we have shown how to fit a polynomial describing the deformation of each surface.
@@
 \subsection{External Forces}
+\dave{Jacobian transpose}
+
 \subsection{Time Integration}
@@
 
+\dave{Some extra notes on how we handle collisions, add a reference, just say we do this from here.}
 %%% ------- OLD STUFF BELOW --------

### 2021-01-24 paper/sections/conclusion.tex
conclusion draft text

\section{Future Work and Conclusions}

We have presented the Shape Matching Element Method (SEM), the first completely meshless approach to direct simulation of NURBS models.
Our approach is unique in its ability to infer volumetric shape from surface only models,
including ones with intersecting geometry, between parts and other defects common to non-engineering models.
Our method allows simulation of such input directly, with a variety of contitutive models and time integration schemes
and is compatibale with standard methods for collision resolution. 
We believe that SEM is a signficant improvement over standard physics-based animtion pipelines for NURBS models. 
As evidence of the efficacy of our approach, the authors submit that many of the examples in this paper were constructed ourselves
(since the standard graphics menagerie is not available as NURBS models). 
Modelling, cleaning, meshing and simulating would've been a burdensome experience without SEM's
ability to leap directly from (often hastily) constructed models to physics-based animations.

But SEM as it's presented here is only the beginning and we believe there are many exciting areas of future work to explore.
We are very excited to couple SEM to machine learning approaches for design, parameter estimation and Real2Sim applications.
One of the cumbersome elements in using finite element simulation for such problems is the need for robust, differentiable volumetric meshing.
While there has been some work on this~\dave{cite} it is far from a solved problem.
SEM removes this bottleneck entirely, providing a direct mapping from geometric input to physics-driven output. 
We believe this will both enable simpler and more robust algorithms for  physics-based ML and 
allow the application of such algorithms to a much broader class of problems.

SEM itself has much room for addition.
First, while we focus on NURBS surfaces here, the only part of SEM that is NURBS specific is the shape matching operation.
We believe there is potential to allow mixed models (models which include polygonal meshes, particles, subdivision surfaces and NURBS)
by extending the range of shape matching operations used by the algorithm. The shape matching operation itself could be improved to be material aware 
(to better handle heterogenous materials) or to be robust to noisy data (allowing direct simulation of scanned data). The Finite Element method took over 
40 years to take over the world of simulation. 
We see this work as the first step in a similar journey for SEM

### 2021-01-24 examples/quadratic_trimmed_mug_with_collision.m
update patch test script, added jello mug example script

@@ -1,0 +1,49 @@
+function quadratic_trimmed_mug_with_collision
+
+function pinned_ids = pin_function(x)
+  pinned_ids = [];
+end
+
+% iges_file = 'mug_complex4_bad.igs';
+iges_file = 'mug_complex4.igs';
+
+part = nurbs_from_iges(iges_file);
+
+% YM = 2e3; %in Pascals
+% pr =  0.25;
+% YM = 1e5;
+YM = 1e4;
+pr = 0.47;
+[lambda, mu] = emu_to_lame(YM, pr);
+
+options.dt=0.01;
+options.rho = 1e3;
+
+options.order = 2;
+options.pin_function = @pin_function;
+options.gravity = -10;
+options.enable_secondary_rays = true;
+options.lambda = lambda;
+options.mu = mu;
+options.sample_interior = 1;
+
+options.plot_points=1;
+options.x_samples=3; %2 on regular 3 on bad
+% options.y_samples=5;
+% options.y_samples=5;
+options.z_samples=11;
+
+options.distance_cutoff=0.1;
+options.distance_cutoff=0.2;
+options.save_obj = 1;
+% options.k_stability = 1e7;
+
+% options.collision_ratio = 4e-1;
+options.collision_ratio = 1;
+options.collision_with_other = false;
+options.self_collision = false;
+options.collision_with_plane = true;
+options.collision_plane_z = -0.4;
+vem_simulate_nurbs_with_collision(part, options);
+  
+end

### 2021-01-24 matlab/deprecated/silly_experiments/patch_test_2d.m

@@ -5,7 +5,7 @@ function patch_test_2d
     save_output = 0;    % (0 or 1) whether to output images of simulation
     
     figure(1);
-    clf;
+    clf; axis equal;
     % The number of elements in the monomial basis.
     % -- example: Draft equation 1 is second order with k == 5
     k = basis_size(d, order);
@@ -13,6 +13,7 @@ function patch_test_2d
     % Load a basic square mesh
     [V,F,UV,TF] = readOBJ('models/plane3.obj');
     F0=F;
+    V0=V;
     z = V(:,3);
     V = V(:,1:2);
     
@@ -63,13 +64,17 @@ function patch_test_2d
     
     x = x0;
     % Linear transform
-    x(1,:) = 2 * x(1,:);
+%     x(1,:) = 0.8 * x(1,:);
 %     x(2,:) = (3/4) * x(2,:);
-    
+%         x(2,:) = 1.5 * x(2,:);
+
     % Rigid transform
-%     theta = 45;
-%     R = [cosd(theta) -sind(theta); sind(theta) cosd(theta)];
-%     x = R * x;
+    theta = 0;
+    R = [cosd(theta) -sind(theta); sind(theta) cosd(theta)];
+    R(1,2) = 0.5;
+%     R(2,1) = 2;
+
+    x = R * x;
     
     E_lines2 = cell(numel(E),1);
     cm=copper(numel(E));
@@ -87,7 +92,93 @@ function patch_test_2d
     b = b(:);
     c = L * b;
     
-    w = compute_projected_weights(x0, E, V');
+    function [a,w] = weights(X)
+        dE = (V0(F(:,2),:) - V0(F(:,1),:));
+        norm2_dE = sum(dE.*dE, 2);
+
+        dVx = X(:,1)' - V0((F(:,1)),1);
+        dVy = X(:,2)' - V0((F(:,1)),2);
+
+        dEdotdV = (dE(:,1).*dVx +  dE(:,2).*dVy)./norm2_dE;
+
+        %bound constraints
+        dEdotdV = min(max(dEdotdV, 0),1);
+
+        a = zeros(size(X,1), size(F,1));
+        w = zeros(size(X,1), size(F,1));
+
+        %%% THIS BIT DOES THE RAYTRACING WEIGHTS %%%%%
+        for jj=1:size(F,1)
+
+            %nearest point on i^th edge of j^th point (column)
+            VE_X = (dEdotdV(jj,:).*dE(jj,1) + V0(F(jj,1),1))';
[…]

### 2021-01-24 examples/quadratic_rocket_soft_with_collision.m
add collisions with sphere

@@ -1,0 +1,41 @@
+function linear_rocket_soft_with_collision
+
+function pinned_ids = pin_function(x)
+    pinned_ids = [];
+end
+
+iges_file = 'rocket.iges';
+
+part = nurbs_from_iges(iges_file);
+
+YM = 7e6; %in Pascals
+pr = 0.4;
+[lambda, mu] = emu_to_lame(YM, pr);
+
+options.order = 2;
+options.gravity = -9.8;
+options.rho = 1.27e3;
+options.pin_function = @pin_function;
+options.lambda = lambda;
+options.mu = mu;
+options.distance_cutoff = 10;
+
+options.save_obj = true;
+options.save_obj_path = 'output/obj_rocket_soft_quadratic/';
+
+options.k_stability = YM*1e5;
+options.collision_ratio = 1e5;
+
+options.x_samples = 5;
+options.y_samples = 9;
+options.z_samples = 15; % we want more samples along vertical axis
+
+options.collision_with_other = false;
+options.self_collision = false;
+options.collision_with_plane = true;
+options.collision_plane_z = -20.0;
+options.initial_velocity = [0 0 -10];
+
+vem_simulate_nurbs_with_collision(part, options);
+
+end

### 2021-01-24 examples/linear_lamppost.m
commit before merge

@@ -6,16 +6,19 @@ function linear_lamppost
 % end
 
 function pinned_ids = pin_function(x)
-    verts_to_pin = 20; 
+    verts_to_pin = 10; 
     [~,I] = mink(x(3,:),verts_to_pin);
     pinned_ids = I(1:verts_to_pin);
+    
+    pinned_ids_2 = find(x(3,:) < 5 & x(3,:) > 4);
+    pinned_ids = [pinned_ids pinned_ids_2];
 end
 
 iges_file = 'lamppost.iges';
 
 part = nurbs_from_iges(iges_file);
 
-YM = 1e5; %in Pascals
+YM = 1e8; %in Pascals
 pr = 0.4;
 % YM = 2e11; %in Pascals
 % pr = 0.32;
@@ -24,17 +27,21 @@ pr = 0.4;
 % options.dt = 0.1;
 options.order = 1;
 options.gravity = 0;
-options.rho = 1.27e3;
+options.rho = 3e3;
+% options.rho = 1.27e3;
+% options.rho = 8e3;
 options.pin_function = @pin_function;
 options.lambda = lambda;
 options.mu = mu;
 options.distance_cutoff = 1.0;
-options.k_stability = YM * 1e2;
-options.plot_points = 1;
-options.x_samples = 10;
-% options.y_samples = 2;
+options.k_stability = YM * 1e-2;
+% options.plot_points = 1;
+options.x_samples = 5;
+options.y_samples = 5;
 options.z_samples = 15;
+% options.initial_velocity = [-20 0 0];
 options.f_external = [-1e4 0 0];
+options.f_external_time = 2.0;
 
 % vem_simulate_nurbs_newtons(part, options);
 vem_simulate_nurbs(part, options);

### 2021-01-24 matlab/vem_simulate_nurbs.m

@@ -19,10 +19,12 @@ function vem_simulate_nurbs(parts, varargin)
     addParameter(p, 'fitting_mode', 'hierarchical');
     addParameter(p, 'plot_points', false);
     addParameter(p, 'plot_com', true);
+    addParameter(p, 'initial_velocity', [0 0 0]);
     addParameter(p, 'x_samples', 5);
     addParameter(p, 'y_samples', 9);
     addParameter(p, 'z_samples', 9);
     addParameter(p, 'f_external', [0 0 0]);
+    addParameter(p, 'f_external_time', 1000);
 
     parse(p,varargin{:});
     config = p.Results;
@@ -43,7 +45,8 @@ function vem_simulate_nurbs(parts, varargin)
     
     % Initial deformed positions and velocities
     x = x0;
-    qdot=zeros(size(q));
+%     qdot=zeros(size(q));
+     qdot = reshape(repmat(config.initial_velocity, size(q,1)/3, 1)', [], 1);
     
     % Setup pinned vertices constraint matrix
     pin_I = config.pin_function(x0);
@@ -168,7 +171,11 @@ function vem_simulate_nurbs(parts, varargin)
         
         f_gravity = config.dt*P*(L' * dg_dc);
         
-        f_external = config.dt*P*(L' * df_ext_dc);
+        if t >= config.f_external_time
+          f_external = zeros(size(f_gravity,1), 1);
+        else
+          f_external = config.dt*P*(L' * df_ext_dc);
+        end
         
         % Error correction force
         f_error = - 2 * ME * x(:);

### 2021-01-24 matlab/vem_simulate_nurbs_with_collision.m

@@ -27,6 +27,7 @@ function vem_simulate_nurbs_with_collision(parts, varargin)
     addParameter(p, 'x_samples', 5);
     addParameter(p, 'y_samples', 9);
     addParameter(p, 'z_samples', 9);
+    addParameter(p, 'save_obj_path', 'output/obj/');
 
     parse(p,varargin{:});
     config = p.Results;
@@ -161,9 +162,13 @@ function vem_simulate_nurbs_with_collision(parts, varargin)
     end
     
     collide_plane_vid = [];
+    
+    if config.save_obj
+      mkdir(config.save_obj_path);
+    end
 
     ii=1;
-    for t=0:config.dt:30
+    for t=0:config.dt:40
         tic
         
         % Preparing input for stiffness matrix mex function.
@@ -256,7 +261,7 @@ function vem_simulate_nurbs_with_collision(parts, varargin)
         rhs = J' * (P*M*P'*J*qdot + f_internal + f_gravity + f_error) + f_collision;
         qdot = lhs \ rhs;
         
-        if config.collision_with_plane && size(collide_plane_vid, 1) > 5 && ii > 1200
+        if config.collision_with_plane && size(collide_plane_vid, 1) > 5 && ii >= 2500
           qdot = reshape(qdot, 3, [])';
           qdot(:, [1 2]) = 0 * qdot(:, [1 2]);
           qdot = reshape(qdot', [], 1); 
@@ -292,9 +297,9 @@ function vem_simulate_nurbs_with_collision(parts, varargin)
             V_plot.ZData = V(3,:);
         end
         drawnow
-        
-        if config.save_obj
-            obj_fn = "output/obj/part_" + int2str(ii) + ".obj";
+         
+        if config.save_obj && ii >= 1200
+            obj_fn = config.save_obj_path + "part_" + int2str(ii) + ".obj";
             nurbs_write_obj(q,parts,obj_fn,ii);
         end
         

### 2021-01-24 examples/quadratic_castle_with_collision.m
update

@@ -67,6 +67,6 @@ options.collision_sphere_c = [-10 10 85;
 %                               -0.2 0.2 1.5;
 %                               -0.2 -0.2 1.5];
 
-vem_simulate_nurbs_with_collision_new(part, options);
+vem_simulate_nurbs_with_collision(part, options);
 
 end

### 2021-01-24 paper/sections/results.tex
conclusion draft text
results draft

@@
 
+\reftbl{perf} shows the size of all our examples along with performance statistics. 
+Note that the individual parts of all models are \emph{not connected}.
+No continuity constraints or constructive solid geometry operations have been applied, rather, the SEM approach implicitly couples the boundary parts together
+to allow for seamless volumetric simulation. 
+Models were created using Autodesk Fusion 360 ~\dave{cite} and Rhinocerous 3D 7\dave{cite}.
+
 \begin{table*}[h]
   \caption{Performance of the Shape Matching Element Method on all Examples.}
-  \label{tbl:compare}
+  \label{tbl:perf}
   \begin{center}
@@
 %kinematic model is expressive
-\begin{figure}
-  \includegraphics[width=\columnwidth]{example-image-a}
-  \caption{Take an intesting geometry (maybe that geometry processing cacture), deform it, do our shape matching and then reconstruct the deformed pose.
+% \begin{figure}
+%   \includegraphics[width=\columnwidth]{example-image-a}
+%   \caption{Take an intesting geometry (maybe that geometry processing cacture), deform it, do our shape matching and then reconstruct the deformed pose.
 Do this for a bunch of poses (show's kinematic model is good).}
-  \label{fig:deform}
-\end{figure}
+%   \label{fig:deform}
+% \end{figure}
 
@@
 %correctness
+We first validated the physical plausibility of our method using a small, 2D, patch test (\reffig{patchtest}). 
+Here the test object is a square made of four edges (not joined at the corners) simulated using linear VEM with a single deformation center.
+We apply a battery of boundary conditions and resolve the deformation of the element by minimizing the elastic potential.
+We note that SEM is able to represent rigid motions, as well as shearing and anisotropic stretching. 
+This implies that, with sparse weights, SEM can resolve these motions locally, leading to physically plausible simulation results.
+
 \begin{figure}
@@
 
+Next we check the convergence of SEM with respect to linear, tetrahedral finite elements. \reffig{convergence} we compare identical cantileavered beams, simulated using 
+a Neohookean constitutive model (Young's Modulus \dave{??}, Poisson's Ratio \dave{??}) and implicit time integration with a timestep of $0.01$s. 
+We use SEM with quadratic polynomials for this test and observe that our SEM simulation, made of 6 independent NURBS patches show's good agreement with FEM.
 \begin{figure}
@@
 
+We also demonstrate that our raycasting weights computation creates shape-aware output. Figure~\reffig{independence} shows that manipulating
+parts that are nearby, but separate  behave in an appropriately independent fashion. 
 \begin{figure}
@@
 % relaxed modeling  constraints
+Here we demonstrate SEM's ability to work with poor models. \reffig{badmodels} shows simulations of two jelly coffee mugs. 
+The top row shows the result of a careless modeler who has deeply embedded the handle of the mug in the body in order to attach it, creating a large area of overlap.
+The bottom models has neglible overlap. 
+In both cases SEM produces a plausible physically-based animation, without requiring additional model clean-up.
 \begin{figure}
@@
 %different material parameters / material models 
+We derive the equations of motion for SEM using a general elastic potential and so SEM supports arbitrary constitutive models. \reffig{materials} shows a 
+rocket ship simulated with both high stiffness and low stiffness materials. 
 \begin{figure}
@@
 
+%multiple materials
+SEM also supports objects made up of multiple materials, as the simulation of this tire~(\reffig{tire}) illustrates.
+\begin{figure}
+  \includegraphics[width=\columnwidth]{example-image-a}
+  \caption{Tire to show heterogenous materials}
+  \label{fig:tire}
+\end{figure}
+
 %different surface reps
+Though we focus on NURBS surface representations, they only NURBS specific construction in the SEM method is the Jacobian\dave{ref}.
+Many other boundary representations admit a linear representation and are thus directly consumable by our method.
+\reffig{subd} shows an example of applying our SEM method to a subdivision model by making such a modification.
 \begin{figure}
@@
 %editing example
+We show that the output of the SEM simulation is itself edible in a NURBS modelling program~(\reffig{edit}).
 \begin{figure}
@@
 
-%multiple materials
-\begin{figure}
-  \includegraphics[width=\columnwidth]{example-image-a}
-  \caption{Avocado to show heterogenous materials}
-  \label{fig:avocado}
-\end{figure}
-
 %show stoppers
+Finally we show the ability of SEM to handle complex simulations involving deformation, contact and friction~(\reffig{staypuft} and \reffig{f1})).
 \begin{figure*}[htp]

### 2021-01-24 paper/sections/methods.tex
Initial dump of choosing deformation centers section. Cover your eyes!

@@
 (density, constitutive model, model parameters etc). 
-The boundary representation is composed of one or more (not necessarily explicit;y connected) NURBS surface primitives that we will call \emph{parts}.
+The boundary representation is composed of one or more (not necessarily explicitly connected) NURBS surface primitives that we will call \emph{parts}.
 The output is an elastodynamic simulation. 
@@
     \label{alg:sem}
-    \SetAlgoLined
-    \KwResult{Write here the result }
-     initialization\;
-     \While{While condition}{
-      instructions\;
-      \eIf{condition}{
-       instructions1\;
-       instructions2\;
-       }{
-       instructions3\;
-      }
-     }
+	\caption{Shape Matching Element Algorithm}\label{sem_algorithm}
+	\begin{algorithmic}[1]
+		\Procedure{Preprocess}{$model$}
+		\State{// Get initial DOF and sample boundary}
+		\State $\mathbf{q^0} \gets \text{GetControlPoints}(model)$ \Comment{Section 4}
+		\State $\mathbf{J}, \mathbf{W} \gets \text{SampleNURBS}(model)$ \Comment{Section 4}
+		\State $\mathbf{x^0} \gets \mathbf{Jq^0}$
+		\State{$\mathbf{q} \gets \mathbf{q^0}$}
+      \Statex
+
+    	\State{// Generate material points inside the model}
+		\State $\mathbf{X} \gets \text{RaycastQuadrature}(model)$ \Comment{Section 4.5}
+		\Statex
+
+      \State{// Weights for quadrature and boundary points}
+     	\State $\mathbf{w} \gets \text{BlendingWeights}(model, \mathbf{X}, d_{\text{max}})$    \Comment{Section 4.3}
+      	\State $\mathbf{w^*} \gets \text{BlendingWeights}(model, \mathbf{x^0}, d_{\text{max}})$   
+		\Statex
+
+		\State{// Generate centers of mass}
+		\State $\bar{\mathbf{X}} \gets \text{GenerateCOM}(\mathbf{X}, \mathbf{w})$  \Comment{Section 4.}
+		\Statex
+
+		\State{// Create projection operator}
+		\State $\Pi \gets \text{ShapeMatching}(\mathbf{q^0},\mathbf{J}, \mathbf{W},\bar{\mathbf{X}})$     	\Comment{Section 4.2}
+		\Statex
+
+		\State{// Form weighted monomial basis and mass matrices}
+		\State $\mathbf{Y} \gets \text{WeightedMonomials}(\mathbf{X}, \mathbf{\bar{X}}, \mathbf{w})$   \Comment{Eqn 32}     	
+		\State $\mathbf{Y^*} \gets \text{WeightedMonomials}(\mathbf{x^*}, \mathbf{\bar{X}}, \mathbf{w})$
+    	\State $\mathbf{M} \gets \text{MassMatrix}(\mathbf{Y}, \Pi)$			\Comment{Section 4.6}
+     	\State $\mathbf{M_E} \gets \text{ErrorMatrix}(\mathbf{Y^*}, \Pi)$	\Comment{Section 4.7}
+     	\Statex
+     	\myworries{todo: add volume}
+     	\While {$simulating$}
+     		\State{$\Pcox \gets \Pi\mathbf{q}$}
+     		\ForAll{$i \gets 1$ to $|\mathbf{X}|$}
+     			\State{$\mathbf{F}_i \gets \frac{\partial \mathbf{F}}{\partial c} c$}
+				\State{$\frac{\partial V}{\partial c} \gets \frac{\partial V}{\partial c} +
+				\frac{\partial \mathbf{F}}{\partial c}^T \frac{\partial \psi}{\partial \mathbf{F}}(\mathbf{F}_i)$}
+
+				\State{$\frac{\partial^2 V}{\partial c^2} \gets \frac{\partial^2 V}{\partial c^2} +
+				 \frac{\partial \mathbf{F}}{\partial c}^T 
+				 \frac{\partial^2 \psi}{\partial \mathbf{F}^2}(\mathbf{F}_i)
+				 \frac{\partial \mathbf{F}}{\partial c}
+				 $}
+  			\EndFor
+  			\State{$\mathbf{K} \gets \Pi^T \frac{\partial^2 V}{\partial c^2} \Pi$}
+  			\State{$f_{\text{internal}} \gets \Pi^T \frac{\partial V}{\partial c}$}
+  			\State{$f_{\text{error}} \gets \mathbf{M_E}\mathbf{q}$}
+  			\State{$f_{\text{total}} \gets f_{\text{internal}} + f_{\text{error}} + f_{\text{ext}}$}
+  			\State{$\dot{\mathbf{q}} \gets \text{TimeIntegration}
+  			(f_{\text{total}},\mathbf{K}, \dot{\mathbf{q}}, \mathbf{J} , \mathbf{M}, \mathbf{M_E})$}
+  			\State{$\mathbf{q} \gets \mathbf{q} + h\dot{\mathbf{q}}$}
+     	\EndWhile
+	\EndProcedure
+\end{algorithmic}
 \caption{Shape Matching Element Algorithm}
@@
 This can be repeated for polynomials of arbitrary high order, for instance here we apply this approach to quadratic shape matching:
-
+\ty{is the -1 above the A terms intentional? unsure if this is a common thing or an error}
 \begin{equation}
@@
 (6) add proofs/diagrams/figures that show we satisfy (2). }
+To ensure kinematic expressiveness, it is essential that multiple origins of deformation are permitted.
+Each part, $\prt_i$, must be associated with a single deformation origin, $\bar{\refX}_j$.
+Each origin is computed as the centroid of the set of parts with which it is associated.
+While we constraint a part to be paired with only one origin, an origin may be associated any number of parts.
+Next, we require that each deformation origin be associated with a minimum of two part so that we avoid having an origin lie on the surface of a part.\ty{Should we mention periodic surfaces?
+You can have a periodic nurbs sphere consisting of a single surface, which we couldn't handle with this constraint (easy fix, but not sure if worth mentioning)} Furthermore, we seek that the sparsity of the deformation origins mirrors that of the blending weights.
+Intuitively, this means that regions in which the weights are high for a subset of parts should deform about an origin centered in this region.
+
+From the set of blending weights we seek to construct a set of candidate deformation origins and then produce a set of deformation origins, $\XBs$, adhering to the outlined requirements.
+These requirements admit a graph structure, which we exploit by forming a simple network flow problem that produces a solution satisfying the requirements.
+The pairwise relationships between parts and deformation origins can be realized as a bipartite network flow problem.
+Flow between a part and origin is constrained to binary values and a flow value on edge between a part and origin candidate indicate that the part is paired with this deformation origin.
+We organize the problem so that the source node is connected to each candidate origin node, and the sink node is connected with each part.
+We place a capacity constraint of $1$ on the sink's edge, which enforces the constraint that each part is paired with one origin.
 
@@
 \dave{Ty: How do we deal with non-planar parts ? What if the ray cast hits the same part that the closest point is on?}
+\ty{Right now this doesn't happen as we only test intersection against other parts}
 

### 2021-01-24 paper/sections/methods.tex
added network flow figure

@@
 
+\ty{Update notation, split this into two figures}
 \begin{algorithm}[h]
@@
 
+\begin{figure}[h]
+    \includegraphics[width=\columnwidth]{figures/origin_network}
+    \caption{Example structure for the network flow selection of deformation origins with an example shape on the right. Red arrows indicate each parts origin association.}
+    \label{fig:multiparts}
+\end{figure}
+
 From the set of blending weights we seek to construct a set of candidate deformation origins and then produce a set of deformation origins, $\XBs$, adhering to the outlined requirements.
-These requirements admit a graph structure, which we exploit by forming a simple network flow problem that produces a solution satisfying the requirements.
+These requirements admit a graph structure ~(\reffig{multiparts}), which we exploit by forming a simple network flow problem that produces a solution satisfying the requirements.
 The pairwise relationships between parts and deformation origins can be realized as a bipartite network flow problem.

### 2021-01-24 paper/sections/methods.tex
some more words, free of charge

@@
 You can have a periodic nurbs sphere consisting of a single surface, which we couldn't handle with this constraint (easy fix, but not sure if worth mentioning)} Furthermore, we seek that the sparsity of the deformation origins mirrors that of the blending weights.
-Intuitively, this means that regions in which the weights are high for a subset of parts should deform about an origin centered in this region.
+Intuitively, this means that for a region in which weights are high for a subset of parts, the origin of deformation should be centered in this region.
 
@@
     \includegraphics[width=\columnwidth]{figures/origin_network}
-    \caption{Example structure for the network flow selection of deformation origins with an example shape on the right. Red arrows indicate each parts origin association.}
+    \caption{Example structure for the network flow selection of deformation origins with a shape consisting of three parts. Red arrows indicate each parts origin association.}
     \label{fig:multiparts}
@@
 These requirements admit a graph structure ~(\reffig{multiparts}), which we exploit by forming a simple network flow problem that produces a solution satisfying the requirements.
-The pairwise relationships between parts and deformation origins can be realized as a bipartite network flow problem.
+The relationships between parts and deformation origins can be realized as a bipartite min-cost-flow problem.
+
+Consider a directed bipartite graph $G=(V = A \cup B,E)$ where each node in $A$ corresponds to a deformation origin candidate and each node in $B$ represents one part.
+$E$ is the set of edges connecting nodes in $A$ to $B$ where each origin has an each to every part with which it attempts to associate.
+We augment $G$ to form $G'=(V' = A \cup B \cup \{s,t\}, E')$ where $V'$  is the new set of nodes now containing the source, $s$, and sink, $t$, nodes.
+$E'$ is the augmented set of nodes connecting $s$ to each candidate in $A$ and connecting each part in $B$ to the sink $t$.
+Each edge $(i,j) \in E'$ has an associated flow $f_{ij}$ and capacity constraint $u_{ij}$, and each edge in $E'$ also has a cost $c_{ij}$.
+The principal constraint for this network flow problem is the conservation constraint, meaning that the net flow for all nodes in $V$ is zero.
+
+First, we constrain all flow values to be integral and place a capacity of $1$ on the edges from $B$ to $t$.
+This has the effect of enforcing that any node in $B$ receives flow from a single edge, and as a consequence each part may only be associated with a single deformation origin.
 Flow between a part and origin is constrained to binary values and a flow value on edge between a part and origin candidate indicate that the part is paired with this deformation origin.

### 2021-01-24 paper/sections/methods.tex
nearly done with first draft of deformation origin part... how does stephen king write so much

@@
 \dave{Ty: need some text to explain how we pick points on the nurbs surface.}
+\ty{Need to think a little more on how to write this.
+The main idea is that each surface has a knot vector and each Bspline covers a span of those knots.
+Therefore to guarantee we represent all control points for a surface, we uniformly sample $d+1$ points in each span, where $d$ is the NURBS surface}
 We discretize the shape matching energy and minimize by solving
@@
 \dave{should have done Chebyshev shape matching to ensure higher order terms are orthogonal which would gaurantee exact reconstruction}
-
-\subsection{Choosing Deformation Centers}
-\label{sec:origins}
-\dave{Ty: try to follow this formula when writing 
-\dave{maybe this should come after weight computation ?}
-(1) what is the purpose of this part of the method ?
-(2) what are the criteria for success
-(3) what is our assumed input 
-(4) what is the desired output
-(5) what's the algorith,
-(6) add proofs/diagrams/figures that show we satisfy (2). }
-To ensure kinematic expressiveness, it is essential that multiple origins of deformation are permitted.
-Each part, $\prt_i$, must be associated with a single deformation origin, $\bar{\refX}_j$.
-Each origin is computed as the centroid of the set of parts with which it is associated.
-While we constraint a part to be paired with only one origin, an origin may be associated any number of parts.
-Next, we require that each deformation origin be associated with a minimum of two part so that we avoid having an origin lie on the surface of a part.\ty{Should we mention periodic surfaces?
-You can have a periodic nurbs sphere consisting of a single surface, which we couldn't handle with this constraint (easy fix, but not sure if worth mentioning)} Furthermore, we seek that the sparsity of the deformation origins mirrors that of the blending weights.
-Intuitively, this means that for a region in which weights are high for a subset of parts, the origin of deformation should be centered in this region.
-
-\begin{figure}[h]
-    \includegraphics[width=\columnwidth]{figures/origin_network}
-    \caption{Example structure for the network flow selection of deformation origins with a shape consisting of three parts. Red arrows indicate each parts origin association.}
-    \label{fig:multiparts}
-\end{figure}
-
-From the set of blending weights we seek to construct a set of candidate deformation origins and then produce a set of deformation origins, $\XBs$, adhering to the outlined requirements.
-These requirements admit a graph structure ~(\reffig{multiparts}), which we exploit by forming a simple network flow problem that produces a solution satisfying the requirements.
-The relationships between parts and deformation origins can be realized as a bipartite min-cost-flow problem.
-
-Consider a directed bipartite graph $G=(V = A \cup B,E)$ where each node in $A$ corresponds to a deformation origin candidate and each node in $B$ represents one part.
-$E$ is the set of edges connecting nodes in $A$ to $B$ where each origin has an each to every part with which it attempts to associate.
-We augment $G$ to form $G'=(V' = A \cup B \cup \{s,t\}, E')$ where $V'$  is the new set of nodes now containing the source, $s$, and sink, $t$, nodes.
-$E'$ is the augmented set of nodes connecting $s$ to each candidate in $A$ and connecting each part in $B$ to the sink $t$.
-Each edge $(i,j) \in E'$ has an associated flow $f_{ij}$ and capacity constraint $u_{ij}$, and each edge in $E'$ also has a cost $c_{ij}$.
-The principal constraint for this network flow problem is the conservation constraint, meaning that the net flow for all nodes in $V$ is zero.
-
-First, we constrain all flow values to be integral and place a capacity of $1$ on the edges from $B$ to $t$.
-This has the effect of enforcing that any node in $B$ receives flow from a single edge, and as a consequence each part may only be associated with a single deformation origin.
-Flow between a part and origin is constrained to binary values and a flow value on edge between a part and origin candidate indicate that the part is paired with this deformation origin.
-We organize the problem so that the source node is connected to each candidate origin node, and the sink node is connected with each part.
-We place a capacity constraint of $1$ on the sink's edge, which enforces the constraint that each part is paired with one origin.
 
@@
 physical quantities, leaving our simulation unaffected.
+
+\subsection{Choosing Deformation Centers}
+\label{sec:origins}
+\dave{Ty: try to follow this formula when writing 
+\dave{maybe this should come after weight computation ?}
+(1) what is the purpose of this part of the method ?
+(2) what are the criteria for success
+(3) what is our assumed input 
+(4) what is the desired output
+(5) what's the algorith,
+(6) add proofs/diagrams/figures that show we satisfy (2). }
+To ensure kinematic expressiveness, it is essential that multiple origins of deformation are permitted.
+Each part, $\prt_i$, must be associated with a single deformation origin, $\bar{\refX}_j$.
+Each origin is computed as the centroid of the set of parts with which it is associated.
+While we constrain a part to be paired with only one deformation origin, an origin may be associated any number of parts.
+Next, we require that each deformation origin be associated with a minimum of two parts, which avoids the case of an origin laying on the surface of a part.\ty{Should we mention periodic surfaces?
+You can have a periodic nurbs sphere consisting of a single surface, which we couldn't handle with this constraint (easy fix, but not sure if worth mentioning)} Furthermore, we seek that the sparsity of the deformation origins mirrors that of the blending weights.
+Intuitively, this means that for a region in which weights are high for a subset of parts, the origin of deformation should be centered in this region.
+
+\begin{figure}[h]
+    \includegraphics[width=\columnwidth]{figures/origin_network}
+    \caption{Example structure for the network flow selection of deformation origins with a shape consisting of three parts. Red arrows indicate each parts origin association.}
+    \label{fig:network_flow}
+\end{figure}
+
+From the set of blending weights we seek to construct a set of candidate deformation origins and then produce a set of deformation origins, $\XBs$, adhering to the outlined requirements.
+This set of candidate origins is identified by examining the unique combinations of weight assignments for the quadrature points.
+For each $\refX_j$, we form a set $\mathbf{c}\left(\refX_j\right)$ with $c_i\left(\refX_j\right)= 1-\delta_{\bld_i\left(\refX_j\right),0}$ where $\delta_{x,0}$ is the Kronecker delta.
+This produces a binary code for each quadrature point that identifies the parts to which this point has nonzero weights.
+We use each unique binary code as a deformation origin candidate.
+
+With the set of requirements and candidate origins, we see this problem may be organized in a graph structure ~(\reffig{network_flow}) where the relationships between parts and deformation origins can be realized as a bipartite graph.
+We exploit this structure by forming a min-cost-flow problem that produces a solution satisfying the requirements.
+
+Consider a directed bipartite graph $G=(V = A \cup B,E)$ where each node in $A$ corresponds to a deformation origin candidate and each node in $B$ represents a part.
+$E$ is the set of edges connecting nodes in $A$ to $B$ where each origin has an edge to every part with which it attempts to associate.
+We augment $G$ to form $G'=(V' = A \cup B \cup \{s,t\}, E')$ where $V'$  is the new set of nodes now containing the source, $s$, and sink, $t$, nodes.
+$E'$ is the augmented set of nodes connecting $s$ to each candidate in $A$ and connecting each part in $B$ to the sink $t$.
+Each edge $(i,j) \in E'$ has an associated flow $f_{ij}$ and capacity constraint $u_{ij}$, and each edge in $E'$ also has a cost $c_{ij}$.
+The principal constraint for this network flow problem is the conservation constraint, meaning that the net flow for all nodes in $V$ is zero.
+
+First, we constrain all flow values to be integral and place a capacity of $1$ on the edges from $B$ to $t$.
+This has the effect of enforcing that any part node in $B$ receives flow from a single edge, and as a consequence each part may only be associated with a single deformation origin.
+With the specified network and constraints, we solve this problem with an integer constrained linear program.
+We choose $\XBs$ to be all nodes in $A$ that have an outgoing flow of at least $1$.
 

### 2021-01-24 paper/sections/methods.tex
finished rough draft of the network flow stuff. its rough alright

@@
 This set of candidate origins is identified by examining the unique combinations of weight assignments for the quadrature points.
-For each $\refX_j$, we form a set $\mathbf{c}\left(\refX_j\right)$ with $c_i\left(\refX_j\right)= 1-\delta_{\bld_i\left(\refX_j\right),0}$ where $\delta_{x,0}$ is the Kronecker delta.
+For each $\refX_j$, we form a set $\mathbf{b}\left(\refX_j\right)$ with $b_i\left(\refX_j\right)= 1-\delta_{\bld_i\left(\refX_j\right),0}$ where $\delta_{x,0}$ is the Kronecker delta.
 This produces a binary code for each quadrature point that identifies the parts to which this point has nonzero weights.
@@
 $E'$ is the augmented set of nodes connecting $s$ to each candidate in $A$ and connecting each part in $B$ to the sink $t$.
-Each edge $(i,j) \in E'$ has an associated flow $f_{ij}$ and capacity constraint $u_{ij}$, and each edge in $E'$ also has a cost $c_{ij}$.
+Each edge $(i,j) \in E'$ has an associated flow $f_{ij}$ and capacity constraint $u_{ij}$, and each edge in $E$ also has a cost $c_{ij}$.
 The principal constraint for this network flow problem is the conservation constraint, meaning that the net flow for all nodes in $V$ is zero.
 
-First, we constrain all flow values to be integral and place a capacity of $1$ on the edges from $B$ to $t$.
+Next, we constrain all flow values to be integral and place a capacity of $1$ on the edges from $B$ to $t$.
 This has the effect of enforcing that any part node in $B$ receives flow from a single edge, and as a consequence each part may only be associated with a single deformation origin.
-With the specified network and constraints, we solve this problem with an integer constrained linear program.
-We choose $\XBs$ to be all nodes in $A$ that have an outgoing flow of at least $1$.
+The last step is to specify the costs on the edges, $E$, between the origin candidates and the parts.
+Costs are assigned in such a way that for each unique binary code, $\mathbf{b}$, we sum the set of weights for each quadrature point associated with this code.
+Since for each non-zero $b_i$, we have an edge connecting the node corresponding to $\mathbf{b}$ to the $i$-th part in $A$, we take the $i$-th sum to be the cost for this edge.
+\ty{Need to think how to rewrite this}.
+The resulting sum has a nonzero value With the specified network and constraints, we solve this problem with an integer constrained linear program.
+We choose $\XBs$ to be all the deformation origin candidates in $A$ that have an outgoing flow of at least $1$.
 

### 2021-01-24 examples/quadratic_castle_with_collision.m
update rocket figure and script for the bouncy castle

@@ -8,7 +8,7 @@ function pinned_ids = pin_function(x)
 end
 
 % iges_file = 'lamppost.iges';
-iges_file = 'castle_simple.iges';
+iges_file = 'castle_simple_smaller_flip.iges';
 
 % To avoid singular nurbs jacobian with excessive pinning, I up the sample
 % density to ensure we have enough unpinned samples.
@@ -18,19 +18,20 @@ part = nurbs_from_iges(iges_file, sample_density);
 
 %material properties
 % YM = 5e3; %in Pascals
-YM = 5e4; %in Pascals
+YM = 1e5; %in Pascals
 pr = 0.45;
 [lambda, mu] = emu_to_lame(YM, pr);
 
+options.dt = 0.01;
 options.order = 2;
 options.pin_function = @pin_function;
 options.gravity = -10;
 options.lambda = lambda;
 options.mu = mu;
-options.rho = 1.27e3;
+options.rho = 2e3;
 % options.distance_cutoff = 1.5;
 % options.distance_cutoff = 0.9;
-options.distance_cutoff = 1;
+options.distance_cutoff = 0.1;
 options.k_stability = 1e5;
 options.enable_secondary_rays = 1;
 options.fitting_mode = 'hierarchical';
@@ -41,31 +42,61 @@ options.save_output = 0;
 % options.z_samples = 10;
 % vem_simulate_nurbs_newtons(part, options);
 
+options.save_obj = true;
+options.save_obj_path = 'output/obj_castle/';
+
+options.k_stability = YM*1e5;
+
 % options.initial_velocity = [0 0 -10];
 options.collision_with_other = false;
 options.collision_other_position = [0 0 -10];
 options.collision_with_other_sim = false;
 options.self_collision = false;
-options.collision_with_plane = false;
+options.collision_with_plane = true;
+options.collision_plane_z = 0.684;
 options.collision_with_sphere = true;
-options.collision_ratio = 5e4;
+options.collision_ratio = 1.2e5;
 
-options.collision_sphere_r = 2.0;
-options.collision_sphere_c = [-10 10 85;
-                              -10 30 85;
-                              -10 50 85;
-                              -30 10 85;
-                              -30 30 85;
-                              -30 50 85;
-                              -50 10 85;
-                              -50 30 85;
-                              -50 50 85];
+options.collision_sphere_r = 0.0175;
+options.collision_sphere_rho = 2e3;
[…]

### 2021-01-24 matlab/vem_simulate_nurbs_with_collision.m

@@ -29,6 +29,7 @@ function vem_simulate_nurbs_with_collision_new(parts, varargin)
     addParameter(p, 'collision_sphere_c', []);           % center of the sphere
     addParameter(p, 'collision_sphere_r', 0.05);         % radius of the sphere
     addParameter(p, 'collision_sphere_rho', 2e3);        % density of the sphere
+    addParameter(p, 'collision_sphere_initial_veloity', [0 0 0]);        % density of the sphere
     addParameter(p, 'initial_velocity', [0 0 0]);        % initial velocity of the nurbs model
     addParameter(p, 'x_samples', 5);
     addParameter(p, 'y_samples', 9);
@@ -149,6 +150,7 @@ function vem_simulate_nurbs_with_collision_new(parts, varargin)
     [verts,faces] = triangulate_iges(parts);
     collision_ratio = config.collision_ratio;
     face_normal = normals(verts, faces);
+    writeOBJ('./test.obj', verts, faces, [], zeros(size(faces,1),3), face_normal, zeros(size(faces,1),3));
     
     if config.collision_with_other
 %       collide_iges_file = 'rounded_cube.iges';
@@ -171,11 +173,12 @@ function vem_simulate_nurbs_with_collision_new(parts, varargin)
     % draw the plane if needed
     if config.collision_with_plane
       if config.collision_plane_z > min(verts(:,3))
-        config.collision_plane_z = min(verts(:,3)) - 20.0;
+%         config.collision_plane_z = min(verts(:,3)) - 20.0;
+        config.collision_plane_z = min(verts(:,3));
       end
       hold on
-      x_range = 2 * (max(verts(:,1))-min(verts(:,1)));
-      y_range = 2 * (max(verts(:,2))-min(verts(:,2)));
+      x_range = 1 * (max(verts(:,1))-min(verts(:,1)));
+      y_range = 1 * (max(verts(:,2))-min(verts(:,2)));
       p = patch('XData', x_range*[-1 -1 1 1],'YData', y_range*[-1 1 1 -1], ...
                 'ZData', config.collision_plane_z * [1 1 1 1], ...
                 'FaceColor',[0.5 0.5 0.5], 'FaceAlpha', 0.5, 'FaceLighting', 'gouraud');
@@ -195,19 +198,20 @@ function vem_simulate_nurbs_with_collision_new(parts, varargin)
       sphere_r = config.collision_sphere_r;
       sphere_V = sphere_r * sphere_V;
       sphere_vol = (4/3) * pi * sphere_r^3;
[…]

### 2021-01-25 matlab/vem_simulate_nurbs_with_collision.m
updated the castle and chicken example, the castle doesn't look quite right now

@@ -345,7 +345,11 @@ function vem_simulate_nurbs_with_collision_new(parts, varargin)
                               (f_collision_sphere_i + f_collision_other_sphere_i + f_collision_sphere_plane_i + sphere_m * [0 0 config.gravity] - sphere_v(si, :)) / sphere_m;
             % hacky air friction
             if size(IF,1) == 0 && sphere_v(si, 3) > 0
-              sphere_v(si, 3) = 1.0 * sphere_v(si, 3);
+              if ii > 500
+                sphere_v(si, 3) = 0.1 * sphere_v(si, 3);
+              else
+                sphere_v(si, 3) = 0.9 * sphere_v(si, 3);
+              end
             elseif size(IF,1) > 0 && sphere_v(si, 3) < 0
               sphere_v(si, 3) = 0;
             end
@@ -361,6 +365,11 @@ function vem_simulate_nurbs_with_collision_new(parts, varargin)
           end
           hold off;
         end
+        
+        fprintf("f_internal = %d\n", sqrt(sum(f_internal.^2)));
+        fprintf("f_gravity = %d\n", sqrt(sum(f_gravity.^2)));
+        fprintf("f_error = %d\n", sqrt(sum(f_error.^2)));
+        fprintf("f_collision = %d\n", sqrt(sum(f_collision.^2)));
 
         f_collision = hires_J' * f_collision;
 
@@ -369,7 +378,7 @@ function vem_simulate_nurbs_with_collision_new(parts, varargin)
         rhs = J' * (P*M*P'*J*qdot + f_internal + f_gravity + f_error + f_external) + f_collision;
         qdot = lhs \ rhs;
         
-        if config.collision_with_plane && size(collide_plane_vid, 1) > 5 && ii >= 2500
+        if config.collision_with_plane && size(collide_plane_vid, 1) > 5 && ii >= 300
           qdot = reshape(qdot, 3, [])';
           qdot(:, [1 2]) = 0 * qdot(:, [1 2]);
           qdot = reshape(qdot', [], 1); 

### 2021-01-25 paper/sections/methods.tex
fix floating -1

@@
 This can be repeated for polynomials of arbitrary high order, for instance here we apply this approach to quadratic shape matching:
-\ty{is the -1 above the A terms intentional? unsure if this is a common thing or an error}
+
 \begin{equation}
@@
         \tr_i  & = \sum_{\prt_k \in \bar{\refX}_j} \underbrace{\mathbbm{1}^T\wt_k \samp{\Jnurbs}_k}_{B_{ik}}\vc{q}_k \\
-        \vc{c}_i^1 & = \underbrace{\left(\samp{\Pmat^1}^T\wt_i\samp{\Pmat^1}\right)}_{A^1_i}^{-1}\bigg(\underbrace{\samp{\Pmat^1}^T\wt_i\samp{\Jnurbs}_i}_{B^1_i}\vc{q}_i-\underbrace{\samp{\Pmat^1}^T\wt_i\mathbbm{1}}_{T^1_i}\tr_i\bigg)\\
-        \vc{c}_i^2 & = \underbrace{\left(\samp{\Pmat^2}^T\wt_i\samp{\Pmat^2}\right)}_{A^2_i}^{-1}\bigg(\underbrace{\samp{\Pmat^2}^T\wt_i\samp{\Jnurbs}_i}_{B^2_i}\vc{q}_i-\underbrace{\samp{\Pmat^2}^T\wt_i\samp{\Pmat^1}}_{D^2_i}\vc{c}_i^1 - \underbrace{\samp{\Pmat^2}^T\wt_i\mathbbm{1}}_{T^2_i}\tr_i\bigg)
+        \vc{c}_i^1 & = \smash{\underbrace{\left(\samp{\Pmat^1}^T\wt_i\samp{\Pmat^1}\right)}_{A^1_i}}^{-1}\bigg(\underbrace{\samp{\Pmat^1}^T\wt_i\samp{\Jnurbs}_i}_{B^1_i}\vc{q}_i-\underbrace{\samp{\Pmat^1}^T\wt_i\mathbbm{1}}_{T^1_i}\tr_i\bigg)\\
+        \vc{c}_i^2 & = \smash{\underbrace{\left(\samp{\Pmat^2}^T\wt_i\samp{\Pmat^2}\right)}_{A^2_i}}^{-1}\bigg(\underbrace{\samp{\Pmat^2}^T\wt_i\samp{\Jnurbs}_i}_{B^2_i}\vc{q}_i-\underbrace{\samp{\Pmat^2}^T\wt_i\samp{\Pmat^1}}_{D^2_i}\vc{c}_i^1 - \underbrace{\samp{\Pmat^2}^T\wt_i\mathbbm{1}}_{T^2_i}\tr_i\bigg)
     \end{split},
@@
 Specifying weights manually would be antithetical to our goal of providing a seamless translation from surface representation to physics-based animation.
-Automatic  weight computation is a well-studied topic (see ~\citet{10.1145/2766952,BBW:2011} and ~\citet{10.1145/2010324.1964968}) but typically requires solving 
+Automatic  weight computation is well-studied (see ~\citet{10.1145/2766952,BBW:2011} and ~\citet{10.1145/2010324.1964968}) but typically requires solving 
 an optimization problem that itself relies on a volumetric discretization. 
@@
 
-Ideally our shape functions would both partition unity and obey the Kronecker delta property (the $i^th$ blending function is one at $i^th$ part and zero on all other parts).
+Ideally our shape functions would both partition unity and obey the Kronecker delta property (the $i^{th}$ blending function is one at $i^{th}$ part and zero on all other parts).
 These properties ensure that our simulation can properly represent translation and that the solution stored at each part is the actual deformed position.
@@
 to a contradiction -- how can the blending functions be both one and zero at the points of overlap ? 
-Our solution is to allow weights to include bounded discontinuities at these points and it is these discontinuous boundary conditions that are difficult to enforce 
+Our solution is to allow weights to include bounded discontinuities at these points and it is these discontinuities that are difficult to enforce 
 with previous methods. 
@@
 \ty{Right now this doesn't happen as we only test intersection against other parts}
-
-Next we compute two distances, $d_{\text{primary}}$, the distance from $\mathbf{X}_j$ to $\clst$ and $d_{\text{total}}$ the distance from  $\clst$ to $\rhit$.
+\dave{can't that mean the ray can just miss everything ? If staypuft's body was just a single ellipse the wieghts inside should all be one, in this case they would be very odd.}
+We compute two distances, $d_{\text{primary}}$, the distance from $\mathbf{X}_j$ to $\clst$ and $d_{\text{total}}$ the distance from  $\clst$ to $\rhit$.
 Using these two distances we define a linear \emph{distance weight} using the well-known ReLU function:
@@
 where $\alpha$ is a distance cutoff parameter. 
-This distance weight~\reffig{plot_distance_weight} is the result of interpolating along this ray from $\clst$ to $\rhit$ (or a fixed point that depends on $\alpha$). 
-These weights obey the Kronecker delta property for all parts but does not partition unity.
+This distance weight~(\reffig{plot_distance_weight}) is the result of interpolating along the ray from $\clst$ to $\rhit$ (or a fixed point that depends on $\alpha$). 
+These weights obey the Kronecker delta property for all parts but do not partition unity which we repair using a post-normalization step.
 
@@
 
-Next we use these distance weights to compute our final blending weights.
 At each $\refX_j$ we want to compute $\mathbf{\bld}(\mathbf{X}_j) =  \left( \bld_1(\mathbf{X}_j), \dots, \bld_n(\mathbf{X}_j) \right)$, where 
@@
 For any point on a part that is non-intersecting, the weights will satisfy the Kronecker delta property since the inverse distance weights imply that only
-a single theta will have a non-infinite weight. 
+a single $\Theta_i$ will have a non-infinite weight. 
 For points that lie on more than one part, the weights associated with the intersecting parts will be equal and all others will be zero
-In other words, the deformed positions of these intersecting points will be an average, which is a plausible solution and ensures continuity.
-The blending weights immediately jump back to one and zero for non-intersecting points on each part. 
+This implies that the deformed positions of these intersecting points will be an average, which is a plausible solution and ensures continuity.
 Finally, the distance cutoff $\alpha$ allows us to control the sparsity of the blending weights.
@@
 but allows us to use more general weighting functions and apply the appropriate behavior for intersecting parts.
-The price we pay for these advantages is that our weights are not C0 continuous but rather Lipchitz smooth.
+The price we pay is that our weights are not C0 continuous but rather Lipchitz smooth.
 Perturbations in the boundaries of parts can cause bounded jumps in our distance weights. 
-However these jumps occupy an infinitesimal percentage of the volume of our object are easy to avoid when integrating
+However, these jumps occupy an infinitesimal percentage of the volume of our object are easy to avoid when integrating
 physical quantities, leaving our simulation unaffected.
 
-\subsection{Choosing Deformation Centers}
+\subsection{Choosing Deformation Origins}
 \label{sec:origins}
-\dave{Ty: try to follow this formula when writing 
-\dave{maybe this should come after weight computation ?}
-(1) what is the purpose of this part of the method ?
-(2) what are the criteria for success
-(3) what is our assumed input 
-(4) what is the desired output
-(5) what's the algorith,
-(6) add proofs/diagrams/figures that show we satisfy (2). }
+% \dave{Ty: try to follow this formula when writing 
+% \dave{maybe this should come after weight computation ?}
+% (1) what is the purpose of this part of the method ?
+% (2) what are the criteria for success
+% (3) what is our assumed input 
+% (4) what is the desired output
+% (5) what's the algorith,
+% (6) add proofs/diagrams/figures that show we satisfy (2). }
 To ensure kinematic expressiveness, it is essential that multiple origins of deformation are permitted.

### 2021-01-25 paper/sections/methods.tex
weight figure now wrapfig

@@
 
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=3.33in]{figures/plot_distance_weight.pdf}
-    \caption{We plot the distance weight with respect to the primary and total ray length. Here the cutoff distance is 50.
-    }
-    \label{fig:plot_distance_weight}
-    \vspace{-5pt}
-  \end{figure}
+% \begin{figure}[h]
+%     \centering
+%     \includegraphics[width=3.33in]{figures/plot_distance_weight.pdf}
+%     \caption{We plot the distance weight with respect to the primary and total ray length. Here the cutoff distance is 50.
+%     }
+%     \label{fig:plot_distance_weight}
+%     \vspace{-5pt}
+%   \end{figure}
 
@@
 where $\Theta(\mathbf{X}) = \text{diag}\left( \frac{1}{\theta_1},\frac{1}{\theta_2},\dots,\frac{1}{\theta_n}\right)$.
+
+\begin{wrapfigure}[11]{l}{0.31\columnwidth}
+    \begin{center}
+            %\vspace{-0.7cm}
+            %\hspace{-0.1cm}
+            \includegraphics[width=0.3\columnwidth]{figures/plot_distance_weight.pdf}
+
+    \end{center}
+    \label{fig:plot_distance_weight}
+    \caption{Distance weight as a function of primary and total ray length with cutoff distance 50.}
+  \end{wrapfigure}
 

### 2021-01-25 paper/sections/methods.tex
deformation origins

@@
 % (6) add proofs/diagrams/figures that show we satisfy (2). }
-To ensure kinematic expressiveness, it is essential that multiple origins of deformation are permitted.
 Each part, $\prt_i$, must be associated with a single deformation origin, $\bar{\refX}_j$. 
@@
 While we constrain a part to be paired with only one deformation origin, an origin may be associated any number of parts. 
-Next, we require that each deformation origin be associated with a minimum of two parts, which avoids the case of an origin laying on the surface of a part.\ty{Should we mention periodic surfaces?
-You can have a periodic nurbs sphere consisting of a single surface, which we couldn't handle with this constraint (easy fix, but not sure if worth mentioning)} Furthermore, we seek that the sparsity of the deformation origins mirrors that of the blending weights.
-Intuitively, this means that for a region in which weights are high for a subset of parts, the origin of deformation should be centered in this region.
+Further, we require that each deformation origin be associated with a minimum of two parts, which avoids the case of an origin laying on the surface of a part.\ty{Should we mention periodic surfaces?
+You can have a periodic nurbs sphere consisting of a single surface, which we couldn't handle with this constraint (easy fix, but not sure if worth mentioning)}\dave{yes}
+Finally, to avoid unduly coupling the parts and reducing the expressivity of the simulation, we seek a sparse association of origins 
+which mirrors the pointwise sparsity of our blending weights,
+
+%of the blending weights. Intuitively, this means that for a region in which weights are high for a subset of parts, the origin of deformation should be centered in this region.
 
@@
     \includegraphics[width=\columnwidth]{figures/origin_network}
-    \caption{Example structure for the network flow selection of deformation origins with a shape consisting of three parts. Red arrows indicate each parts origin association.}
+    \caption{Example structure for the network flow selection of deformation origins with a shape consisting of three parts.
+Red arrows indicate each parts origin association.
+\dave{make it clear edges of triangle are sep.
+parts}}
     \label{fig:network_flow} 
@@
 
-From the set of blending weights we seek to construct a set of candidate deformation origins and then produce a set of deformation origins, $\XBs$, adhering to the outlined requirements.
-This set of candidate origins is identified by examining the unique combinations of weight assignments for the quadrature points.
-For each $\refX_j$, we form a set $\mathbf{b}\left(\refX_j\right)$ with $b_i\left(\refX_j\right)= 1-\delta_{\bld_i\left(\refX_j\right),0}$ where $\delta_{x,0}$ is the Kronecker delta.
-This produces a binary code for each quadrature point that identifies the parts to which this point has nonzero weights.
-We use each unique binary code as a deformation origin candidate.
+Candidate deformation origins are selected by clustering quadrature points~(\refsec{quadrature}).
+We group quadrature points by the combination of parts influencing each point (non-zero blending weight~(\refsec{weights})).
+Each group now represents a unique part combination and the centroid of these parts becomes a deformation origin candidate.
+To assign deformation origins to parts we use an integer network flow approach~\dave{cite}~(\reffig{origin_network}).
 
-With the set of requirements and candidate origins, we see this problem may be organized in a graph structure ~(\reffig{network_flow}) where the relationships between parts and deformation origins can be realized as a bipartite graph.
-We exploit this structure by forming a min-cost-flow problem that produces a solution satisfying the requirements.
-
-Consider a directed bipartite graph $G=(V = A \cup B,E)$ where each node in $A$ corresponds to a deformation origin candidate and each node in $B$ represents a part.
+Consider a directed bipartite graph $G=(V = A \cup B,E)$ where each node in $A$ corresponds to a deformation origin candidate and each node in $B$ represents to a part. 
 $E$ is the set of edges connecting nodes in $A$ to $B$ where each origin has an edge to every part with which it attempts to associate. 
-We augment $G$ to form $G'=(V' = A \cup B \cup \{s,t\}, E')$ where $V'$  is the new set of nodes now containing the source, $s$, and sink, $t$, nodes.
+We augment $G$ to form $G'=(V' = A \cup B \cup \{s,t\}, E')$ where $V'$  is the new set of nodes now containing the source, $s$, and sink, $t$.
 $E'$ is the augmented set of nodes connecting $s$ to each candidate in $A$ and connecting each part in $B$ to the sink $t$.
 Each edge $(i,j) \in E'$ has an associated flow $f_{ij}$ and capacity constraint $u_{ij}$, and each edge in $E$ also has a cost $c_{ij}$.
-The principal constraint for this network flow problem is the conservation constraint, meaning that the net flow for all nodes in $V$ is zero.
+The principal constraint for this network flow problem is the conservation constraint, meaning that the net flow for all nodes in $V$ is zero.\dave{B is really overloaded}
+Each edge from from $B$ to $t$ is assigned a capacity of $1$, which enforces that any part node in $B$ receives flow from a single edge, 
+and as a consequence each part may only be associated with a single deformation origin. 
+The last step is to specify the costs on the edges, $E$, between the origin candidates and the parts.
 
-Next, we constrain all flow values to be integral and place a capacity of $1$ on the edges from $B$ to $t$.
-This has the effect of enforcing that any part node in $B$ receives flow from a single edge, and as a consequence each part may only be associated with a single deformation origin.
-The last step is to specify the costs on the edges, $E$, between the origin candidates and the parts.
-Costs are assigned in such a way that for each unique binary code, $\mathbf{b}$, we sum the set of weights for each quadrature point associated with this code.
-Since for each non-zero $b_i$, we have an edge connecting the node corresponding to $\mathbf{b}$ to the $i$-th part in $A$, we take the $i$-th sum to be the cost for this edge.
-\ty{Need to think how to rewrite this}.
-The resulting sum has a nonzero value With the specified network and constraints, we solve this problem with an integer constrained linear program.
-We choose $\XBs$ to be all the deformation origin candidates in $A$ that have an outgoing flow of at least $1$.
+Given a quadrature point group, and an edge from the associated deformation origin to $\prt_i$, we set the edge capacity to be the sum, 
+over all quadrature points in the group, of the blending weights associated with $\prt_i$.
+This capacity causes more heavily ``connected'' origin candidates to be selected. 
+We solve this problem with an integer constrained linear program\dave{cite solver} and choose $\XBs$ to be all the deformation origin candidates in $A$ 
+that have an outgoing flow of at least $1$. 
+This approach chooses deformation origins that satisfy our stated criteria.
 
 \subsection{Meshless Integration}
-
+\label{sec:quadrature}
 We integrate physical quantities over our undeformed object using raycasting based quadrature~\cite{KHOSRAVIFARD201030}.

### 2021-01-25 paper/sections/methods.tex
origin section

@@
 \dave{make it clear edges of triangle are sep.
-parts}}
+parts, add capacities and costs}}
     \label{fig:network_flow} 
@@
 
-Candidate deformation origins are selected by clustering quadrature points~(\refsec{quadrature}).
+Candidate deformation origins are selected by clustering quadrature points~(themselves computed in \refsec{quadrature}).
 We group quadrature points by the combination of parts influencing each point (non-zero blending weight~(\refsec{weights})).
-Each group now represents a unique part combination and the centroid of these parts becomes a deformation origin candidate.
+Each group now represents a unique part combination and the centroids of these groups become deformation origin candidates.
 To assign deformation origins to parts we use an integer network flow approach~\dave{cite}~(\reffig{origin_network}).
 
-Consider a directed bipartite graph $G=(V = A \cup B,E)$ where each node in $A$ corresponds to a deformation origin candidate and each node in $B$ represents to a part. 
+Consider a directed bipartite graph $G=(V = A \cup B,E)$ where each node in $A$ corresponds to a deformation origin candidate and each node in $B$ represents a part. 
 $E$ is the set of edges connecting nodes in $A$ to $B$ where each origin has an edge to every part with which it attempts to associate. 
@@
 The principal constraint for this network flow problem is the conservation constraint, meaning that the net flow for all nodes in $V$ is zero.\dave{B is really overloaded}
-Each edge from from $B$ to $t$ is assigned a capacity of $1$, which enforces that any part node in $B$ receives flow from a single edge, 
+Each edge from $B$ to $t$ is assigned a capacity of $1$, which enforces that any part node in $B$ receives flow from a single edge, 
 and as a consequence each part may only be associated with a single deformation origin. 
@@
 
-Given a quadrature point group, and an edge from the associated deformation origin to $\prt_i$, we set the edge capacity to be the sum, 
+Given a quadrature point group, and an edge from the associated deformation origin to $\prt_i$, we set the edge cost to be the sum, 
 over all quadrature points in the group, of the blending weights associated with $\prt_i$.
-This capacity causes more heavily ``connected'' origin candidates to be selected. 
-We solve this problem with an integer constrained linear program\dave{cite solver} and choose $\XBs$ to be all the deformation origin candidates in $A$ 
+This costs causes more heavily ``connected'' origin candidates to be penalized, encouraging sparsity during origin selection. 
+We solve this problem with an integer constrained linear program\dave{cite solver, add equation with formulation} and choose $\XBs$ to be all the deformation origin candidates in $A$ 
 that have an outgoing flow of at least $1$.
-This approach chooses deformation origins that satisfy our stated criteria.
 

### 2021-01-25 paper/sections/methods.tex
fix
one subsection to go

@@
 over all quadrature points in the group, of the blending weights associated with $\prt_i$.
-This costs causes more heavily ``connected'' origin candidates to be penalized, encouraging sparsity during origin selection. 
+This costs penalizes heavily ``connected'' origin candidates, encouraging sparsity during origin selection. 
 We solve this problem with an integer constrained linear program\dave{cite solver, add equation with formulation} and choose $\XBs$ to be all the deformation origin candidates in $A$ 
@@
 
+Having defined the individual components of the SEM method, we now briefly detail how we apply it to the problem of 
+physics-based animation, specifically elastodynamics. 
+The advantage of having a unified kinematic description~(\refeq{kinematics}) is that we can directly apply Lagrangian Mechanics~\dave{cite Lanczos again} to arrive at 
+the pertinent equations of motion:
+\begin{equation*}
+    \MM\ddot{\vc{q}} = \fo\left(\vc{q}\right) + \fext,
+\end{equation*} where $\MM\in\real^{n\times n}$ is the mass matrix, $\vc{q}\in\real^{n}$ is the stacked vector of surface control points, $\fo\in\real^{n}$ is the 
+vector of internal forces and $\fext\in\real^{n}$ are external forces such due to gravity or contact. 
+We use penalty springs~\cite{10.1145/1964921.1964932} to resolve contact, and the approach of \citet{10.1145/566654.566623} for approximating frictional effects. 
+Because SEM yields equations of motion in the standard form, it is compatible with any standard integration scheme, though in this submission we limit ourselves to
+Implicit Euler.
+
+\subsection{Generalized Velocity}
+
+We can rewrite \refeq{kinematics} in matrix form as
+
+\begin{equation}
+    \defX\left(\refX\right) = \shapef\left(\refX\right)\bldmat\Pi\vc{q},
+\end{equation} where $\shapef\in\real^{3\times n} = \begin{bmatrix}P_1\left(\refX\right) \dots P_m\left(\refX\right)\end{bmatrix}$, $m$ is number of 
+parts in the object, $\bldmat$ is the diagonal matrix of blending weights and $\Pi$ is the global projection operator.
+
+The velocity of a point on the object is then  
+\begin{equation}
+        \vc{v}\left(\refX\right) = \underbrace{\shapef\left(\refX\right)\bldmat\Pi}_{\Jkin}\dot{\vc{q}},
+    \end{equation} where $\Jkin\left(\refX\right)$ is the kinematic Jacobian and $\dot{\vc{q}}$ are the generalized velocities of the system.
+
 \subsection{Mass Matrix}
-The kinetic energy for a model may be expressed as
+
+We derive the mass matrix for our system from the definition of kinetic energy:
 \begin{equation}
-T = \frac{1}{2}\int_{ \Omega} \rho \dot{\mathbf{x}}^T\dot{\mathbf{x}} d\Omega
-\text{,}
+    T = \dot{\vc{q}}^T\underbrace{\Pi^T\int_{\Omega}\shapef\left(\refX\right)^Tshapef\left(\refX\right)d\Omega\Pi}_{\MM}\dot{\vc{q}},
 \end{equation}
-where $\Omega$ is the 3D integration domain, $\rho$ is the density, and $\dot{\mathbf{x}}$ is the velocity for some point in the domain.
+% \begin{equation}
+% T = \frac{1}{2}\int_{ \Omega} \rho \dot{\mathbf{x}}^T\dot{\mathbf{x}} d\Omega
+% \text{,}
+% \end{equation}
+% where $\Omega$ is the 3D integration domain, $\rho$ is the density, and $\dot{\mathbf{x}}$ is the velocity for some point in the domain.
 Using the deformation map defined in \ref{eqn:compact_x_map}, we may rewrite this kinetic energy as
-\begin{equation}
-\begin{split}
-T & = \frac{1}{2}\int_{ \Omega} \rho \dot{\mathbf{q}}^T \mathbf{J}^T\mathbf{Y(X)}^T\mathbf{Y(X)}\ d\Omega \\
-  & = \frac{1}{2} \dot{\mathbf{q}}^T \mathbf{J}^T \left[ \int_{ \Omega} \rho \mathbf{Y(X)}^T\mathbf{Y(X)} d\Omega \right] \mathbf{J}\dot{\mathbf{q}} \\
-  & = \frac{1}{2} \dot{\mathbf{q}}^T \mathbf{M} \dot{\mathbf{q}}
-\end{split}
-\text{.}
-\end{equation}
-The second line is a result of the fact that all terms except for $\mathbf{Y(X)}$ are spatially constant, allowing us to move these outside the volume integral, which enables fast assembly of the mass matrix, $\mathbf{M}$.
-
-\subsection{Error Energy}
-\dave{We should cite VEM literature and point out this is a standard thing they do to make the algorithm consistent and convergent}
-Over the course of the simulation we may see the true boundary values deviate from their positions estimated by the polynomial.
-This is most notable in cases where the degree of the polynomial on a NURBS is higher than the degree of the B-spline surface.
-To address this we augment our kinetic energy with a term to account for this error.
-The error for a boundary point $\mathbf{x}_i^*$ that is known to lie on the limit surface of the NURBS can be written as
-\begin{equation}
-\begin{split}
-\epsilon_i & = ||\mathbf{x}(\mathbf{X}_i) - \mathbf{x}_i^*|| \\
-		   & = ||(\mathbf{Y}(\mathbf{X}_i)\mathbf{L} - \mathbf{S_i})\mathbf{\hat{b}}||
-\end{split}
-\text{,}
-\end{equation}
-where $\mathbf{S_i}$ extracts $\mathbf{x}_i^*$ from $\mathbf{\hat{b}}$.
-We modify this error further to be expressed in terms of the control points.
-We note that $\mathbf{x}^*(u,v)$ yields the position on the NURBS surface whereas $\mathbf{x(X)}$ is the deformed value estimated by the polynomial fitting.
-Furthermore $\mathbf{X}$ may be written as a function of the parametric coordinates $\mathbf{X}(u,v)$ as its position on the surface is produced using the initial set of control points representing the undeformed model.
-This leads to an energy potential over the NURBS surface in parametric space
-\begin{equation}
-T_E = \iint (\mathbf{x}(\mathbf{X}(u,v)) - \mathbf{x}^*(u,v))^T(\mathbf{x}(\mathbf{X}(u,v)) - \mathbf{x}^*(u,v)) du dv
-\text{.}
-\end{equation}
-The discrete form of this written in terms of the control points may be written as
-\begin{equation}
-\begin{split}
-T_E & = \sum_i^N  \dot{\mathbf{q}}^T\mathbf{J}^T(\mathbf{Y}(\mathbf{X}_i)\mathbf{L} - \mathbf{S_i})^T(\mathbf{Y}(\mathbf{X}_i)\mathbf{L} - \mathbf{S_i})\mathbf{J}\dot{\mathbf{q}} \\
-  & = \dot{\mathbf{q}}^T\mathbf{J}^T \left[ \sum_i^N  (\mathbf{Y}(\mathbf{X}_i)\mathbf{L} - \mathbf{S_i})^T(\mathbf{Y}(\mathbf{X}_i)\mathbf{L} - \mathbf{S_i}) \right] \mathbf{J}\dot{\mathbf{q}} \\
-  & = \dot{\mathbf{q}}^T \mathbf{M}_E \dot{\mathbf{q}}
-\end{split}
-\text{,}
-\end{equation}
-where $N$ is the total number of boundary points and $\mathbf{M}_E$ is our \textit{error mass matrix}.
-% \subsection{Generalized Forces}
-% We write our potential energy as 
 % \begin{equation}
-%     V = \int_\Omega \psi(\mathbf{F(X)}) d\Omega
+% \begin{split}
+% T & = \frac{1}{2}\int_{ \Omega} \rho \dot{\mathbf{q}}^T \mathbf{J}^T\mathbf{Y(X)}^T\mathbf{Y(X)}\ d\Omega \\
+%   & = \frac{1}{2} \dot{\mathbf{q}}^T \mathbf{J}^T \left[ \int_{ \Omega} \rho \mathbf{Y(X)}^T\mathbf{Y(X)} d\Omega \right] \mathbf{J}\dot{\mathbf{q}} \\
+%   & = \frac{1}{2} \dot{\mathbf{q}}^T \mathbf{M} \dot{\mathbf{q}}
+% \end{split}
 % \text{.}
 % \end{equation}
-% To evaluate this energy (as well as its derivatives), we require a definition for the deformation gradient, $\mathbf{F(X)}$:
-% \begin{equation}
-%         \mathbf{F(X)} = \frac{\partial \mathbf{x}}{\partial \mathbf{X}}
-%                = \frac{\partial \mathbf{Y(X)}}{\partial \mathbf{X}}\mathbf{L}\hat{\mathbf{b}}
-% \text{.}
-% \end{equation}
-% In the expression for the deformed position $\mathbf{x(X)}$ we see that the dependence on $\mathbf{X}$ only exists in $\mathbf{Y(X)}$, making $\mathbf{L}$ and $\hat{\mathbf{b}}$ constant terms when compute $\mathbf{F}$.
-Full details on how to compute $\mathbf{F}$ and its derivative $\frac{\partial \mathbf{F}}{\partial \mathbf{q}}$ may be found in the appendix.
-(\myworries{I just wrote this because I'm too tired to right the d etails right now :p}.
-Now that we have an expression for the deformation gradient, we discretize our potential energy as follows
-% \begin{equation}
-%     V =  \sum_i^m \psi(\mathbf{F(X}_i) A_i
-% \text{,}
-% \end{equation}
-% where $m$ is the number of integration points and $A_i$ is the volume for the $i$-th integration point.
+% The second line is a result of the fact that all terms except for $\mathbf{Y(X)}$ are spatially constant, allowing us to move these outside the volume integral, which enables fast assembly of the mass matrix, $\mathbf{M}$.
 
-% Like in the case of the kinetic energy, we also introduce an error force potential, which takes the simple form
-% \begin{equation}
-% V_E = \mathbf{q}^T \mathbf{M}_E \mathbf{q}
-% \end{equation}
 \subsection{Deformation Gradient}
-\dave{gotta deal with this weight gradient}
-At this point we have shown how to fit a polynomial describing the deformation of each surface.
+% \dave{gotta deal with this weight gradient}
+% At this point we have shown how to fit a polynomial describing the deformation of each surface.
 With these polynomial coefficients coupled with their associated centers of mass, we can construct an estimate of some arbitrary material point $\mathbf{X}$ using the deformation described by the surface.
@@
 Thus we arrive at a definition for the deformed position, $\mathbf{x}$, of a material point $\mathbf{X}$ in the undeformed space:
-\begin{equation}
-\label{eqn:deformation_map}
-	\mathbf{x}(\mathbf{X}) = \sum_i^n w_i(\mathbf{X}) \left[ \mathbf{M}(\mathbf{X-\bar{X}}_j)\mathbf{c}_i + \mathbf{t}_j \right]
-	\text{,}
-\end{equation}
-where we have $n$ surfaces, and $w_i$ and $\mathbf{c}_i$ are the weights and the coefficients for the $i$-th surface, respectively.
+% \begin{equation}
+% \label{eqn:deformation_map}
+% 	\mathbf{x}(\mathbf{X}) = \sum_i^n w_i(\mathbf{X}) \left[ \mathbf{M}(\mathbf{X-\bar{X}}_j)\mathbf{c}_i + \mathbf{t}_j \right]
+% 	\text{,}
+% \end{equation}
+% where we have $n$ surfaces, and $w_i$ and $\mathbf{c}_i$ are the weights and the coefficients for the $i$-th surface, respectively.
 The terms $\mathbf{\bar{X}}_j$ and $\mathbf{t}_j$ the undeformed and deformed centers of mass, respectively, that are associated with the $i$-th surface.
 
+\subsection{Constitutive Models}
+Our algorithm supports arbitrary elastic constitutive models. 
+In this submission we focus on hyperelastic models~\cite{10.1145/2343483.2343501} wherein the potential energy stored by the object, per unit volume,
+is given by the strain energy density function $\psi\left(\defo\right)$, and its internal forces and stiffness are the negative gradient and hessian respectively. 
+Evaluating and applying constitutive models of this form requires the ability to (1) evaluate the deformation gradient over the reference domain of the model and 
+(2) integrate the resulting constitutive properties. 
+SEM provides both features and thus support for such models.
+
 \subsection{External Forces}
-\dave{Jacobian transpose}
+We compute the force of gravity and apply external forces in the standard way --
+$\vc{f}_g = \int_{\Omega} \rho\vc{g} d\Omega$, where $\vc{g}$ is the acceleration due to gravity acting on the object. 
+We leverage our meshless quadrature~(\refsec{quadrature}) for integration.
 
-\subsection{Time Integration}
-With our definition of the deformation map as well as a strategy for integrating the volume, we now have the necessary ingredient to perform dynamic simulation.
-We follow the Variational Mechanics approach that defines a pair of kinetic and potential energies, and we apply the Euler-Lagrange equation to develop our equations of motion.
+External point forces are applied using the Jacobian transpose technique which specifies that the generalized force, $\fext$, 
+that results from a point load, applied at $\refX$ is given by $\fext=\Jkin^T\vc{f}\left(\refX\right)$, where $\vc{f}$ is the point load itself.
+Using the Jacobian transpose allows us to easily apply all manner of external forces including collision and frictional forces.
 
-Before we define these energies we first would like to simplify the form of the deformation map presented in equation \ref{eqn:deformation_map}. We may write $\mathbf{x}(\mathbf{X})$ in matrix form as
+\subsection{Error Energy}
+Like all VEM-type methods, ours requires an error energy to ensure its consistency~\cite{10.1142/S021820251440003X}.
+A natural choice is to adapt the error energy specified by \citet{10.1145/1073204.1073216} to penalize deviations between the pose prescribed by  $\vc{q}$ and 
+that prescribed by polynomial basis.
+
+The total error at the boundary of the object can be written as
 \begin{equation}
-\mathbf{x}(\mathbf{X}) = \underbrace{\left[w_1\mathbf{M}(\mathbf{X - \bar{X}}_j) \dots w_n\mathbf{M}(\mathbf{X - \bar{X}}_j) \middle| \sum_i^{n_1}w_1^i\mathbf{I} \dots \sum_i^{n_m}w_m^i\mathbf{I} \right]}_{\mathbf{Y}(\mathbf{X})}
-\begin{bmatrix}
-\mathbf{\hat{c}} \\
-\mathbf{\hat{t}}
-\end{bmatrix}
-\text{,}
-\end{equation}
-\myworries{I'm not sure how to write this matrix form for x(X), especially on finding the best notation for the center of mass indexing. So this above equation is temporary.}
-where we have $n$ shape elements, $m$ centers of mass. We simplify the above equation and use the following form
-\begin{equation}
-\label{eqn:x_simple_form}
-\mathbf{x}(\mathbf{X}) = \mathbf{Y}(\mathbf{X})\mathbf{L}\hat{\mathbf{b}}
-\text{,}
-\end{equation}
-where we make use of equation \ref{eqn:c_multiple_elements} to rewrite this in terms of the boundary points $\mathbf{\hat{b}}$.
-As one final step, we rewrite this again in terms of the control points of our NURBS surfaces:
-\begin{equation}
-\label{eqn:compact_x_map}
-\mathbf{x}(\mathbf{X}) = \mathbf{Y}(\mathbf{X})\mathbf{LJq}
-\text{,}
-\end{equation}
-where we use the fact that $\mathbf{\hat{b}} = \mathbf{Jq}$.
-\myworries{not a fan of this, perhaps we change $\hat{b}$ to something else...} This gives us an expression for the deformed position in terms of our generalized coordinates, $\mathbf{q}$.
+\epsilon = \int_{\partial\Omega}||\Gamma\left(\refX\right)\vc{q} - \Jnurbs\left(\refX\right)\vc{q}||_2^2 d\refX \\
+\end{equation}, which is a quadratic.
 
-\dave{discrete Equations of motion with quadratured integrals, reference/details on how we handle collisions}
+We augment our physical system with the  quadratic penalty energy $\gamma\epsilon$ and use it to derive additional force and Hessian terms.
+In practice we find that choosing $\gamma $ to be the Young's Modulus of the elastic solid provides good behavior in all cases.
+If the polynomial basis can exactly match the deformation of the boundary representation, this error naturally elides to zero.
 
-We have now provided all the necessary ingredient to produce equations of motion for deformable bodies, which we derive via Euler-Lagrange equation. We begin with the Lagrangian
-\begin{equation}
-    L=(T+T_E)-(V + V_E)
-\end{equation}
-Then from the Euler-Lagrange equation we have:
-\begin{equation}
-    \frac{d}{dt} \frac{\partial L}{\partial \dot{\mathbf{p}}} + \frac{\partial L}{\partial {\mathbf{p}}} = 0
-\end{equation}
-\myworries{todo: not totally sure what to write here since we're experimenting with both linearly implicit backward euler as well as Newton's method soon}
-
-\begin{equation}
-   \frac{\partial L}{\partial \mathbf{q}} = \sum_i^m \frac{\partial V_i}{\partial \mathbf{q}} = \sum_i^m \frac{\partial \psi(\mathbf{F_i)}}{\partial \mathbf{F}}\frac{\partial \mathbf{F_i}}{\partial \mathbf{q}}A_i
-\end{equation}
-\
-
-\dave{Some extra notes on how we handle collisions, add a reference, just say we do this from here.}
+In the next section we show a plethora of physics-based animations created using our approach.
 %%% ------- OLD STUFF BELOW --------

### 2021-01-25 paper/sections/results.tex
update rocket figure and script for the bouncy castle
Add tire figure
results
more figures
one subsection to go

@@
 
+\dave{rendering we convert to triangle meshes, just didn't have a good path tracer lying around}
+
 \begin{table*}[h]
-  \caption{Performance of the Shape Matching Element Method on all Examples.}
+  \caption{Performance of the Shape Matching Element Method on all Examples. \dave{need table that also reports material params/constitutive model or add it here}}
   \label{tbl:perf}
@@
 rocket ship simulated with both high stiffness and low stiffness materials. 
-\begin{figure}
-  \includegraphics[width=\columnwidth]{example-image-a}
+\begin{figure*}
+  \includegraphics[width=\textwidth]{figures/rocket.pdf}
   \caption{Big image simulating the rocket with all different combinations of things}
   \label{fig:materials}
-\end{figure}
+\end{figure*}
 
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{example-image-a}
+  \includegraphics[width=\columnwidth]{figures/tire}
   \caption{Tire to show heterogenous materials}
@@
 \begin{figure*}[htp]
-  \includegraphics[width=\textwidth,height=3in]{example-image-a}
-  \caption{sequence of frames from F1 car simulation}
+  \includegraphics[width=\textwidth,height=3in]{figures/starship.pdf}
+  \caption{The Space-V starship makes a graceful landing on the red planet. }
   \label{fig:f1}

### 2021-01-26 paper/sections/conclusion.tex
more figures
note

@@
 \section{Future Work and Conclusions}
-
+\dave{mention that constraint from complimentary dynamics could be used to replace error energy}
 We have presented the Shape Matching Element Method (SEM), the first completely meshless approach to direct simulation of NURBS models.
@@
 We see this work as the first step in a similar journey for SEM
+
+To encourage future research on SEM the authors will release our SEM implementation under a permissive license as well as all models created for this submission.

### 2021-01-26 paper/sections/conclusion.tex
update conclusion

--- 2021-01-24 paper/sections/conclusion.tex
@@
 \section{Future Work and Conclusions}
-
 We have presented the Shape Matching Element Method (SEM), the first completely meshless approach to direct simulation of NURBS models.
@@
 by extending the range of shape matching operations used by the algorithm. The shape matching operation itself could be improved to be material aware 
-(to better handle heterogenous materials) or to be robust to noisy data (allowing direct simulation of scanned data). The Finite Element method took over 
-40 years to take over the world of simulation. 
-We see this work as the first step in a similar journey for SEM
+(to better handle heterogenous materials) or to be robust to noisy data (allowing direct simulation of scanned data). 
+It is also interesting to consider the relationship between our error term and the orthogonality constraint proposed by \citet{Zhang:CompDynamics:2020}. 
+Their approach could be used to remove this parameter from the approach entirely.
+
+The Finite Element method took over  40 years to mature to its current state and to become the preeminent tool in physically-based animation.
+We hope that this is the beginning of a similar, exciting journey for SEM. 
+Motivated by this sentiment, and to encourage future research on SEM, the authors will release our SEM implementation 
+under a permissive license as well as all models created for this submission.

### 2021-01-26 paper/main.tex
intro draft
done intro draft
starting methods. draft of method overview section
paper in proper format
change title
title change again
crummy abstract

@@
-\documentclass[sigconf]{acmart}
+\documentclass[acmtog, anonymous,review, nonacm, balance = false]{acmart}
 %%
@@
     \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
+
+\input{sections/pre.tex}
 
@@
 
-\title{Shape Matching Element Method}
+\title{The Shape Matching Element Method: Direct Animation of Curved Surface Models}
 
@@
 
+\begin{abstract}
+We introduce a new method for direct physics-based animation of volumetric curved models.  
+The Shape Matching Element Method is a completely meshless approach which is robust to gaps and overlaps in geometry, 
+is compatible with standard constitutive models and time  integration schemes, supports contact and frictional interactions 
+as well as direct editing of simulation output. 
+We demonstrate the efficacy of our algorithm by producing compelling physics-based animations from a variety of curved input models.
+\end{abstract}
+
 \maketitle
@@
 \input{sections/related}
+\input{sections/methods}
+
 \input{sections/results}
@@
 
-Dump of some equations
-\begin{align*}
-    \mathbf{M(X)} = (X_0 - \bar{X}_0, X_1 - \bar{X}_1)\\
-    \bar{\mathbf{x}} = \frac{1}{n}\sum_i^n x_i \\
-    \bar{\mathbf{X}} = \frac{1}{n}\sum_i^n X_i \\
-\end{align*}
-
 \end{document}

### 2021-01-26 examples/quadratic_puft_with_collision.m
added the puft example

@@ -1,0 +1,56 @@
+function quadratic_puft_with_collision
+
+function pinned_ids = pin_function(x)
+  pinned_ids = [];
+end
+ 
+iges_file = 'puft_full_simpler.iges'; % simplified version
+
+part = nurbs_from_iges(iges_file);
+
+% YM = 2e3; %in Pascals
+% pr =  0.25;
+% YM = 1e5;
+YM = 5e5;
+pr = 0.47;
+[lambda, mu] = emu_to_lame(YM, pr);
+
+options.dt=0.01;
+options.rho = 1.27e3;
+
+options.order = 2;
+options.pin_function = @pin_function;
+options.gravity = 0;
+options.enable_secondary_rays = true;
+options.lambda = lambda;
+options.mu = mu;
+% options.sample_interior = 1;
+
+options.plot_points=0;
+options.x_samples=3; %2 on regular 3 on bad
+options.y_samples=8;
+options.z_samples=15;
+% options.z_samples=11;
+
+% New cutoff heuristic thing. The second argument is the "sparsity" which
+% is a 0 to 1 value for how local you want deformations to be.
+% So 0.9 yields a cutoff distance value that is very local.
+cutoff_distance=cutoff_heuristic(part, 0.9);
+
+options.distance_cutoff = cutoff_distance;
+% options.distance_cutoff=0.2;
+options.save_obj = true;
+options.save_obj_path = 'output/puft_with_puft/';
+options.k_stability = 1e10;
+
+% options.collision_ratio = 4e-1;
+options.collision_ratio = 5e4;
+options.collision_with_other = true;
+options.collision_other_position = [20 0 0];
+options.self_collision = false;
+options.collision_with_plane = false;
+options.collision_plane_z = -20;
+options.initial_velocity = [10 0 0];
+vem_simulate_nurbs_with_collision(part, options);
+  
+end

### 2021-01-26 matlab/vem_simulate_nurbs_with_collision.m

@@ -135,6 +135,14 @@ function vem_simulate_nurbs_with_collision_new(parts, varargin)
     % Optional external force vector
     dext_dc = vem_ext_force(config.f_external', config.rho*vol, Y, Y_S);
     f_external = config.dt*P*(L' * dext_dc); 
+    
+    % Add these lines before mass matrix computation
+    bottom_ids = find(V(3,:) < 0.1); % set this number to top z value of the bottom area
+    YM = 1e10; pr = 0.4; % make it stiff than towers
+    [lambda, mu] = emu_to_lame(YM, pr);
+    new_lame = repmat([mu * 0.5  lambda * 0.5], numel(bottom_ids),1);
+    params(bottom_ids,:) = new_lame;
+    config.rho(bottom_ids) = 6e3;   % you can make it denser too
 
     % Compute mass matrices
     ME = vem_error_matrix(Y0, Y0_S, L, d);
@@ -153,14 +161,16 @@ function vem_simulate_nurbs_with_collision_new(parts, varargin)
     writeOBJ('./test.obj', verts, faces, [], zeros(size(faces,1),3), face_normal, zeros(size(faces,1),3));
     
     if config.collision_with_other
-%       collide_iges_file = 'rounded_cube.iges';
-%       collide_parts = nurbs_from_iges(collide_iges_file);
-%       [collide_verts, collide_faces] = triangulate_iges(collide_parts);
-      [collide_verts, collide_faces] = readOBJ('./models/bowl.obj');
-      collide_verts = 5 * collide_verts * (max(verts(:,1))-min(verts(:,1))) / (max(collide_verts(:,1))-min(collide_verts(:,1)));
+      collide_iges_file = 'puft_full_simpler.iges';
+      collide_parts = nurbs_from_iges(collide_iges_file);
+      [collide_verts, collide_faces] = triangulate_iges(collide_parts);
+%       [collide_verts, collide_faces] = readOBJ('./models/fandisk.obj');
+%       collide_verts = 5 * collide_verts * (max(verts(:,1))-min(verts(:,1))) / (max(collide_verts(:,1))-min(collide_verts(:,1)));
       if size(config.collision_other_position, 1) == 0 
         collide_verts = collide_verts + [0 0 max(verts(:,3)) + 10.0];
       else
+        tmp = [-1  0  0; 0 -1  0; 0  0  1];
+        collide_verts = (tmp * collide_verts')';
         collide_verts = collide_verts + config.collision_other_position;
[…]

### 2021-01-26 paper/sections/conclusion.tex

@@
 and is compatibale with standard methods for collision resolution. 
-We believe that SEM is a signficant improvement over standard physics-based animtion pipelines for NURBS models. 
+We believe that SEM is a significant improvement over standard physics-based animation pipelines for NURBS models. 
 As evidence of the efficacy of our approach, the authors submit that many of the examples in this paper were constructed ourselves
@@
 
-But SEM as it's presented here is only the beginning and we believe there are many exciting areas of future work to explore.
+SEM as it's presented here is only the beginning and we believe there are many exciting areas of future work to explore.
 We are very excited to couple SEM to machine learning approaches for design, parameter estimation and Real2Sim applications.
-One of the cumbersome elements in using finite element simulation for such problems is the need for robust, differentiable volumetric meshing.
-While there has been some work on this~\dave{cite} it is far from a solved problem.
+One of the cumbersome elements in using finite element simulation for such problems is the need for robust, differentiable volumetric meshing~\cite{gao2020deftet}.
 SEM removes this bottleneck entirely, providing a direct mapping from geometric input to physics-driven output. 
-We believe this will both enable simpler and more robust algorithms for  physics-based ML and 
+We believe SEM will enable simpler and more robust algorithms for  physics-based ML and 
 allow the application of such algorithms to a much broader class of problems.
@@
 (to better handle heterogenous materials) or to be robust to noisy data (allowing direct simulation of scanned data). 
-It is also interesting to consider the relationship between our error term and the orthogonality constraint proposed by \citet{Zhang:CompDynamics:2020}. 
+It is also interesting to consider the relationship between our error term and the orthogonality constraint proposed by~\citet{Zhang:CompDynamics:2020}. 
 Their approach could be used to remove this parameter from the approach entirely.

### 2021-01-26 paper/sections/conclusion.tex
notes

@@
 \section{Future Work and Conclusions}
+\dave{mention we focus on animation ... lots of work to validate this for engineering applications}
 We have presented the Shape Matching Element Method (SEM), the first completely meshless approach to direct simulation of NURBS models.

### 2021-01-26 paper/sections/intro.tex
fixed typos in intro

@@
 approriate physical parameters.  
-Because we explicitly use the NURBS boundary representation in the simulation its is straightforward to handle
-Dirichilet, Neumann boundary conditions and to apply contact resolution. 
-Crucially because we broadly target animation not necessarily simulation for engineering or manufacturing we don't require that models
+Because we explicitly use the NURBS boundary representation in the simulation it is straightforward to handle
+Dirichilet and Neumann boundary conditions and to apply contact resolution. 
+Crucially because we broadly target animation not necessarily simulation for engineering or manufacturing, we don't require that models
 satisfy the rigourous geometric requirements common for these applications

### 2021-01-26 paper/sections/intro.tex
fixed typos in intro

@@
 approriate physical parameters.  
-Because we explicitly use the NURBS boundary representation in the simulation it is straightforward to handle
+Because we explicitly use the NURBS boundary representation in the simulation, it is straightforward to handle
 Dirichilet and Neumann boundary conditions and to apply contact resolution. 

### 2021-01-26 paper/sections/methods.tex
methods draft done, paper draft done :)

@@
  Each input object is composed of a set of boundary representations, $\Prts$, one for each part $\prt\in\Prts$. 
- Our kinematic mapping is a blended construction (as is done for skinning~\cite{10.1145/2614028.2615427}) which we write as
+ Our kinematic mapping is a blended construction~\cite{10.1145/2614028.2615427,blendedFORKS2016} which we write as
 
@@
 \begin{equation}
-    T = \dot{\vc{q}}^T\underbrace{\Pi^T\int_{\Omega}\shapef\left(\refX\right)^Tshapef\left(\refX\right)d\Omega\Pi}_{\MM}\dot{\vc{q}},
+    T = \dot{\vc{q}}^T\underbrace{\Pi^T\int_{\Omega}\bldmat\left(\refX\right)^T\shapef\left(\refX\right)^T\shapef\left(\refX\right)\bldmat\left(\refX\right)d\Omega\Pi}_{\MM}\dot{\vc{q}},
 \end{equation}
@@
 \subsection{Deformation Gradient}
+The deformation at a point $\refX$ in the reference space is given by
+\begin{equation}
+    \frac{\partial\defX}{\partial \refX} = \sum_{i=1}^m\left(\bld_i\frac{\partial \Pmatxi}{\partial X}\Pi_i\vc{q}_i + \frac{\partial \bld_i}{\partial X}\Pmatxi\Pi_i\vc{q}\right), 
+\end{equation}
+
+For physics-based animation, we find that it is sufficient to assume the blending weights are constant in the region around quadrature points, allowing us to discard the second term.
 % \dave{gotta deal with this weight gradient}

### 2021-01-26 paper/sections/related.tex
figures
fixed typos and added an empty figure

@@
 %% Set the scene
-Geometric modeling is a necessary precursor to physics-based animation, however differing geometric representations for modelling and simulation often require
+Geometric modeling is a necessary precursor to physics-based animation, however differing geometric representations for modelling and simulation often requires
 time-consuming, complex geometry processing pipelines. 
@@
 Volumetric simulations relied on volumetric NURBS~\cite{10.1007/978-3-642-03596-8_2} but were limited to a narrow class of geometries.
-Finite volume methods are applicable to more general geometries~\cite{HEINRICH20121645,10.1002/nme.2311} but require a volumetric mesh be generated.
+Finite volume methods are applicable to more general geometries~\cite{HEINRICH20121645,10.1002/nme.2311} but require a volumetric mesh to be generated.
 Modern approaches are constructed around the extended finite element method which enriches the standard finite element basis with discontinuous basis functions

### 2021-01-26 paper/sections/methods.tex
fixed typos

@@
 
-We make a minor modifiction this procedure to handle common geometric pathologies encountered in NURBS modeling, such as overlapping surfaces by performing 
+We make a minor modifiction this procedure to handle common geometric pathologies encountered in NURBS modelling, such as overlapping surfaces by performing 
 CSG unions while raycasting\dave{find a citation}. 

### 2021-01-26 paper/sections/related.tex
fixed typos

@@
 %% Set the scene
-Geometric modeling is a necessary precursor to physics-based animation, however differing geometric representations for modelling and simulation often requires
+Geometric modelling is a necessary precursor to physics-based animation, however differing geometric representations for modelling and simulation often requires
 time-consuming, complex geometry processing pipelines. 
@@
 Finally these approaches require additional mechanisms to ensure that simulation results lie inside the shape space of the input model's primitives,
-increasing complexity.
+which increases the complexity.
 
 \emph{Shape Matching} is a meshless approach to physics-based animation~\cite{10.1145/1073204.1073216,10.1145/2019406.2019438} and geometry processing~\cite{10.1111/j.1467-8659.2012.03171.x}
-built around shape registration. The algorithm has been extended from volumetric triangle mesh input to cloth~\cite{10.2312/PE/vriphys/vriphys08/039-046} to particles~\cite{10.1145/1964921.1964987}
+built around shape registration. The algorithm has been extended from volumetric triangle mesh input to cloth~\cite{10.2312/PE/vriphys/vriphys08/039-046}, to particles~\cite{10.1145/1964921.1964987}
 and even to visual geometry for video games~\cite{10.1145/2994258.2994260}. 
 Shape Matching is fast~\cite{10.1145/1275808.1276480,10.5555/1632592.1632606} and meshless, but state-of-the-art methods require 
-additional modeling input to position simulation primitives~\cite{10.1145/1964921.1964987} or limit simulation primitives to be 
+additional modelling input to position simulation primitives~\cite{10.1145/1964921.1964987} or limit simulation primitives to be 
 collections of convex polytopes~\cite{10.1145/2994258.2994260}.
@@
 
-To alleviate these restrictions we can to turn to other meshless methods popular in computer graphics and 
+To alleviate these restrictions we turn to other meshless methods popular in computer graphics and 
 engineering~\cite{10.1145/1028523.1028542,https://doi.org/10.1002/fld.1650200824,10.1145/1778765.1778776,10.1145/2010324.1964968, 10.1145/1944846.1944855}.
-These support more advanced constitutive models and integration schemes but often require background integration meshes and lose the direct connection with
+These methods support more advanced constitutive models and integration schemes but often require background integration meshes and lose the direct connection with
 modelling geometry. 
@@
 Our method is entirely meshless (requiring no volumetric meshes or cutcells) and is compatible with standard constitutive models and time integrators.
-It guarantees that simulation output is directly consumable by the input modeling software and can ingest models which include large gaps,
+It guarantees that simulation output is directly consumable by the input modelling software and can ingest models which include large gaps,
 intersections and disconnected primitives without additional user input. 

### 2021-01-27 paper/sections/conclusion.tex
updated results

@@
 Our approach is unique in its ability to infer volumetric shape from surface only models,
-including ones with intersecting geometry, between parts and other defects common to non-engineering models.
-Our method allows simulation of such input directly, with a variety of contitutive models and time integration schemes
-and is compatibale with standard methods for collision resolution. 
+including ones with intersecting geometry between parts and other defects common to non-engineering models.
+Our method allows simulation of such input directly with a variety of constitutive models and time integration schemes,
+and is compatible with standard methods for collision resolution. 
 We believe that SEM is a significant improvement over standard physics-based animation pipelines for NURBS models. 

### 2021-01-27 paper/sections/intro.tex
updated results

--- 2021-01-26 paper/sections/intro.tex
@@
 approriate physical parameters.  
-Because we explicitly use the NURBS boundary representation in the simulation it is straightforward to handle
-Dirichilet and Neumann boundary conditions and to apply contact resolution. 
+Because we explicitly use the NURBS boundary representation in the simulation, it is straightforward to handle
+Dirichlet and Neumann boundary conditions and to apply contact resolution. 
 Crucially because we broadly target animation not necessarily simulation for engineering or manufacturing, we don't require that models

### 2021-01-27 paper/sections/results.tex
Added grumpy figure, not super happy with how it turned out
figure settings
fixed typos and added an empty figure
fixed typos
updated results

@@
 
+\begin{figure*}
+  \includegraphics[width=\textwidth]{figures/chicken.pdf}
+  \caption{Simulation of a chicken model with gaps between surfaces}
+  \label{fig:chicken}
+\end{figure*}
+
 \reftbl{perf} shows the size of all our examples along with performance statistics. 
@@
 
-\dave{rendering we convert to triangle meshes, just didn't have a good path tracer lying around}
+For raycasting operations and rendering, we triangulate the NURBS surfaces, which only represents an implementation simplification.
+Our algorithm is not dependent on triangulating the boundary, and an ideally one would use a path tracer.
+
+\ty{Would be nice to mention TNURBS, which would give us a good reason to include the wrench (which i really wanna do).
+I doubt we have time to do a 3 point bending test, but a simple elastic wrench simulation would be easy to cook up}
+For handling trimmed NURBS surfaces, we augment our sampling of the surfaces by using raycasting integration in 2D.
+Trimmed NURBS may be defined by a boundary curve in the parametric space, so we raycast to find intersection intervals in the UV space on which quadrature points are generated.
+For rendering Trimmed NURBS, we discretize the the boundary curve as a polyline and use triangle ~\ty{cite} to form a boundary conforming triangulation.
+The vertices of this triangulation serve as fixed set of high resolution UV samples for constructing a render mesh.
 
@@
    \textbf{Cantilever}      & ? & ? & ? & ? & ? & ? & ? & ? \\
-   \textbf{Rocket}      & ? & ? & ? & ? & ? & ? & ? & ? \\
+   \textbf{Coffee mug}      & ? & ? & ? & ? & ? & ? & ? & ? \\
+   \textbf{Rocket}          & ? & 648 & ? & ? & ? & ? & ? & ? \\
+   \textbf{Castle}          & ? & ? & ? & ? & ? & ? & ? & ? \\
+   \textbf{Chicken}         & ? & ? & ? & ? & ? & ? & ? & ? \\
+   \textbf{Starship}        & ? & ? & ? & ? & ? & ? & ? & ? \\
+   \textbf{Tire}            & ? & ? & ? & ? & ? & ? & ? & ? \\
    \hline
@@
 
-\begin{figure}
+\begin{figure}[h]
   \includegraphics[width=\columnwidth]{figures/patch_test.pdf}
-  \caption{2D examples (left) rigidly transform boundary and show that solving the static problem gives us rigid motion on the interior (right) isotropically scale boundary and show solving static problem gives us constant strain on the interior}
+  \caption{2D examples (left) undeformed model defined by four surface elements on the boundary (right) applying affine transformations to the boundary and show that solving the static problem gives rigid motion for rotation and constant strain for shearing and stretching}
   \label{fig:patchtest}
@@
 
-Next we check the convergence of SEM with respect to linear, tetrahedral finite elements. \reffig{convergence} we compare identical cantileavered beams, simulated using 
-a Neohookean constitutive model (Young's Modulus \dave{??}, Poisson's Ratio \dave{??}) and implicit time integration with a timestep of $0.01$s. 
-We use SEM with quadratic polynomials for this test and observe that our SEM simulation, made of 6 independent NURBS patches show's good agreement with FEM.
-\begin{figure}
-  \includegraphics[width=\columnwidth]{example-image-a}
-  \caption{(left) high res FEM cantileavered bar (right) several simulations using more and more surface patches plus center of mass. Need to show beam is just  unattached patches}
+Next we check the convergence of SEM with respect to linear, tetrahedral finite elements. \reffig{convergence} we compare identical cantilevered beams, simulated using 
+a Neohookean constitutive model (Young's Modulus $0.001$ GPa, Poisson's Ratio 0.45) and implicit time integration with a timestep of $0.01$s. 
+We use SEM with quadratic polynomials for this test and observe that our SEM simulation, made of 24 independent NURBS patches shows good agreement with FEM.
+Each subdivision of the NURBS beam enables more complicated kinematics, but very few surface elements are needed to produce compelling results.
+
+\begin{figure}[h]
+  \includegraphics[width=\columnwidth]{figures/beams.pdf}
+  \caption{(left) high res FEM cantilevered bar (right) several simulations using more and more surface.}
   \label{fig:convergence}
@@
 
-We also demonstrate that our raycasting weights computation creates shape-aware output. Figure~\reffig{independence} shows that manipulating
-parts that are nearby, but separate  behave in an appropriately independent fashion. 
-\begin{figure}
-  \includegraphics[width=\columnwidth]{example-image-a}
+We also demonstrate that our raycasting weight computation creates shape-aware output. Figure~\reffig{independence} shows that manipulating
+parts that are nearby but separate will behave in an appropriately independent fashion. 
+\begin{figure}[h]
+  \includegraphics[width=\columnwidth]{figures/independent_movement}
   \caption{X shape and show arms of X move independently}
@@
 
-% relaxed modeling  constraints
+% relaxed modelling  constraints
 Here we demonstrate SEM's ability to work with poor models. \reffig{badmodels} shows simulations of two jelly coffee mugs. 
-The top row shows the result of a careless modeler who has deeply embedded the handle of the mug in the body in order to attach it, creating a large area of overlap.
-The bottom models has neglible overlap. 
+The top row shows a model with negligible overlap, whereas the bottom row shows the result of a careless modeller who has deeply embedded the handle of the mug in the body in order to attach it, creating a large area of overlap.
 In both cases SEM produces a plausible physically-based animation, without requiring additional model clean-up.
-\begin{figure}
+\begin{figure}[h]
   \includegraphics[width=\columnwidth]{figures/mug_overlap}
@@
 \end{figure}
+Similarly, in \reffig{chicken} we show a simulation with gaps between surfaces.
+Despite the lack of explicit connectivity, the blending weights have the effect of implicitly enforcing connections at these seams.
 
@@
 rocket ship simulated with both high stiffness and low stiffness materials. 
-\begin{figure*}
+\begin{figure*}[htp]
   \includegraphics[width=\textwidth]{figures/rocket.pdf}
   \caption{Big image simulating the rocket with all different combinations of things}
-  \label{fig:materials}
+  \label{fig:rocket}
 \end{figure*}
@@
 SEM also supports objects made up of multiple materials, as the simulation of this tire~(\reffig{tire}) illustrates.
-\begin{figure}
+\begin{figure}[h]
   \includegraphics[width=\columnwidth]{figures/tire}
@@
 %different surface reps
-Though we focus on NURBS surface representations, they only NURBS specific construction in the SEM method is the Jacobian\dave{ref}.
+Though we focus on NURBS surface representations, the only NURBS specific construction in the SEM method is the Jacobian\dave{ref}.
 Many other boundary representations admit a linear representation and are thus directly consumable by our method.
 \reffig{subd} shows an example of applying our SEM method to a subdivision model by making such a modification.
-\begin{figure}
+\begin{figure}[h]
   \includegraphics[width=\columnwidth]{example-image-a}
@@
 
+% different time integration scheme
+\begin{figure}[h]
+  \includegraphics[width=\columnwidth]{example-image-a}
+  \caption{show mug simulation using newton's method, linearly implicit Euler (and maybe RK4). }
+  \label{fig:time_integration}
+\end{figure}
+
 %editing example
 We show that the output of the SEM simulation is itself edible in a NURBS modelling program~(\reffig{edit}).
-\begin{figure}
+\begin{figure}[h]
   \includegraphics[width=\columnwidth]{example-image-a}
@@
 %show stoppers
-Finally we show the ability of SEM to handle complex simulations involving deformation, contact and friction~(\reffig{staypuft} and \reffig{f1})).
+Finally we show the ability of SEM to handle complex simulations involving deformation, contact and friction~(\reffig{staypuft} and \reffig{f1}).
 \begin{figure*}[htp]

### 2021-01-27 paper/sections/methods.tex
updated algorithm figure

@@
 
-\ty{Update notation, split this into two figures}
 \begin{algorithm}[h]
@@
 	\begin{algorithmic}[1]
-		\Procedure{Preprocess}{$model$}
-		\State{// Get initial DOF and sample boundary}
-		\State $\mathbf{q^0} \gets \text{GetControlPoints}(model)$ \Comment{Section 4}
-		\State $\mathbf{J}, \mathbf{W} \gets \text{SampleNURBS}(model)$ \Comment{Section 4}
-		\State $\mathbf{x^0} \gets \mathbf{Jq^0}$
-		\State{$\mathbf{q} \gets \mathbf{q^0}$}
+		\Procedure{PrepareSim}{$model$}
+		\State{// Get initial DOF and sample boundary  \Comment{Section \ref{sec:geometric_model}}}
+		\State $\vc{q} \gets \text{GetControlPoints}(model)$
+		\State $\samp{\Jnurbs}, \wt \gets \text{SampleNURBS}(model)$
+		\State $\vc{x^0} \gets \wt \samp{\Jnurbs} \vc{q}$
         \Statex
@@
     	\State{// Generate material points inside the model}
-		\State $\mathbf{X} \gets \text{RaycastQuadrature}(model)$ \Comment{Section 4.5}
+		\State $\st{X} \gets \text{RaycastQuadrature}(model)$ \Comment{Section \ref{sec:quadrature}}
 		\Statex
 
-      \State{// Weights for quadrature and boundary points}
-     	\State $\mathbf{w} \gets \text{BlendingWeights}(model, \mathbf{X}, d_{\text{max}})$    \Comment{Section 4.3}
-      	\State $\mathbf{w^*} \gets \text{BlendingWeights}(model, \mathbf{x^0}, d_{\text{max}})$   
+		\State{// Set of blending weights for each $\refX \in \st{X}$} 
+     	\State $\st{W} \gets \text{BlendingWeights}(model, \st{X}, \alpha)$    \Comment{Section \ref{sec:weights}}
 		\Statex
-
-		\State{// Generate centers of mass}
-		\State $\bar{\mathbf{X}} \gets \text{GenerateCOM}(\mathbf{X}, \mathbf{w})$  \Comment{Section 4.}
+		\State $\XBs \gets \text{DeformationOrigins}(\st{X}, \st{w})$  \Comment{Section \ref{sec:origins}}
 		\Statex
@@
 		\State{// Create projection operator}
-		\State $\Pi \gets \text{ShapeMatching}(\mathbf{q^0},\mathbf{J}, \mathbf{W},\bar{\mathbf{X}})$     	\Comment{Section 4.2}
+		\State $\Pi \gets \text{ShapeMatching}(\vc{q}, \samp{\Jnurbs}, \wt, \XBs)$ \Comment{Section \ref{sec:shapematching}}
 		\Statex
 
-		\State{// Form weighted monomial basis and mass matrices}
-		\State $\mathbf{Y} \gets \text{WeightedMonomials}(\mathbf{X}, \mathbf{\bar{X}}, \mathbf{w})$   \Comment{Eqn 32}     	
-		\State $\mathbf{Y^*} \gets \text{WeightedMonomials}(\mathbf{x^*}, \mathbf{\bar{X}}, \mathbf{w})$
-    	\State $\mathbf{M} \gets \text{MassMatrix}(\mathbf{Y}, \Pi)$			\Comment{Section 4.6}
-     	\State $\mathbf{M_E} \gets \text{ErrorMatrix}(\mathbf{Y^*}, \Pi)$	\Comment{Section 4.7}
-     	\Statex
-     	\myworries{todo: add volume}
-     	\While {$simulating$}
-     		\State{$\Pcox \gets \Pi\mathbf{q}$}
-     		\ForAll{$i \gets 1$ to $|\mathbf{X}|$}
-     			\State{$\mathbf{F}_i \gets \frac{\partial \mathbf{F}}{\partial c} c$}
-				\State{$\frac{\partial V}{\partial c} \gets \frac{\partial V}{\partial c} +
-				\frac{\partial \mathbf{F}}{\partial c}^T \frac{\partial \psi}{\partial \mathbf{F}}(\mathbf{F}_i)$}
-
-				\State{$\frac{\partial^2 V}{\partial c^2} \gets \frac{\partial^2 V}{\partial c^2} +
-				 \frac{\partial \mathbf{F}}{\partial c}^T 
-				 \frac{\partial^2 \psi}{\partial \mathbf{F}^2}(\mathbf{F}_i)
-				 \frac{\partial \mathbf{F}}{\partial c}
-				 $}
-  			\EndFor
-  			\State{$\mathbf{K} \gets \Pi^T \frac{\partial^2 V}{\partial c^2} \Pi$}
-  			\State{$f_{\text{internal}} \gets \Pi^T \frac{\partial V}{\partial c}$}
-  			\State{$f_{\text{error}} \gets \mathbf{M_E}\mathbf{q}$}
-  			\State{$f_{\text{total}} \gets f_{\text{internal}} + f_{\text{error}} + f_{\text{ext}}$}
-  			\State{$\dot{\mathbf{q}} \gets \text{TimeIntegration}
-  			(f_{\text{total}},\mathbf{K}, \dot{\mathbf{q}}, \mathbf{J} , \mathbf{M}, \mathbf{M_E})$}
-  			\State{$\mathbf{q} \gets \mathbf{q} + h\dot{\mathbf{q}}$}
-     	\EndWhile
+		\State{// Form mass matrix and error hessian, respectively}
+    	\State $\MM \gets \text{MassMatrix}(\st{X}, \XBs, \st{W},\Pi)$    \Comment{Section \ref{ssec:mass_matrix}}
+     	\State $\ME \gets \text{ErrorMatrix}(\vc{x^0}, \XBs, \st{W},\Pi)$  \Comment{Section \ref{ssec:error_energy}}
+%     	\myworries{todo: add volume}
+%     	\While {$simulating$}
+%     		\State{$\Pcox \gets \Pi\mathbf{q}$}
+%     		\ForAll{$i \gets 1$ to $|\st{X}|$}
+%     			\State{$\mathbf{F}_i \gets \frac{\partial \mathbf{F}}{\partial c} c$}
+%				\State{$\frac{\partial V}{\partial c} \gets \frac{\partial V}{\partial c} +
+%				\frac{\partial \mathbf{F}}{\partial c}^T \frac{\partial \psi}{\partial \mathbf{F}}(\mathbf{F}_i)$}
+%				
+%				\State{$\frac{\partial^2 V}{\partial c^2} \gets \frac{\partial^2 V}{\partial c^2} +
+%				 \frac{\partial \mathbf{F}}{\partial c}^T 
+%				 \frac{\partial^2 \psi}{\partial \mathbf{F}^2}(\mathbf{F}_i)
+%				 \frac{\partial \mathbf{F}}{\partial c}
+%				 $}
+%  			\EndFor
+%  			\State{$K \gets \Pi^T \frac{\partial^2 V}{\partial c^2} \Pi$}
+%  			\State{$f_{\text{internal}} \gets \Pi^T \frac{\partial V}{\partial c}$}
+%  			\State{$f_{\text{error}} \gets \ME \vc{q}$}
+%  			\State{$f_{\text{total}} \gets f_{\text{internal}} + f_{\text{error}} + f_{\text{ext}}$}
+%  			\State{$\dot{\mathbf{q}} \gets \text{TimeIntegration}
+%  			(f_{\text{total}},\mathbf{K}, \dot{\mathbf{q}}, \mathbf{J} , \mathbf{M}, \mathbf{M_E})$}
+%  			\State{$\mathbf{q} \gets \mathbf{q} + h\dot{\mathbf{q}}$}
+%     	\EndWhile
 	\EndProcedure
@@
 \section{Geometric Model}
+\label{sec:geometric_model}
 
@@
 \section{Physics-Based Animation}
-
 Having defined the individual components of the SEM method, we now briefly detail how we apply it to the problem of 
@@
 \subsection{Mass Matrix}
+\label{ssec:mass_matrix}
 
@@
 \subsection{Error Energy}
+\label{ssec:error_energy}
 Like all VEM-type methods, ours requires an error energy to ensure its consistency~\cite{10.1142/S021820251440003X}.

### 2021-01-27 paper/sections/results.tex
took a pass on the result section

@@
   \includegraphics[width=\textwidth]{figures/chicken.pdf}
-  \caption{Simulation of a chicken model with gaps between surfaces}
+  \caption{Simulation of a chicken model with gaps between surfaces.}
   \label{fig:chicken}
@@
 \reftbl{perf} shows the size of all our examples along with performance statistics. 
-Note that the individual parts of all models are \emph{not connected}.
-No continuity constraints or constructive solid geometry operations have been applied, rather, the SEM approach implicitly couples the boundary parts together
-to allow for seamless volumetric simulation. 
-Models were created using Autodesk Fusion 360 ~\dave{cite} and Rhinocerous 3D 7\dave{cite}.
+Note that the individual parts of all models are \emph{not connected}, and no continuity constraints or constructive solid geometry operations have been applied. 
+Rather, the SEM approach implicitly couples the boundary parts together to allow for seamless volumetric simulation. 
+Models were created using Autodesk Fusion 360 \cite{AutodeskFusion360} and Rhinocerous 3D 7 \cite{mcneel-rhinoceros}.
 
 For raycasting operations and rendering, we triangulate the NURBS surfaces, which only represents an implementation simplification.
-Our algorithm is not dependent on triangulating the boundary, and an ideally one would use a path tracer.
+Our algorithm is not dependent on triangulating the boundary, and an ideal alternative would be using a path tracer.
 
@@
 Trimmed NURBS may be defined by a boundary curve in the parametric space, so we raycast to find intersection intervals in the UV space on which quadrature points are generated.
-For rendering Trimmed NURBS, we discretize the the boundary curve as a polyline and use triangle ~\ty{cite} to form a boundary conforming triangulation.
+For rendering Trimmed NURBS, we discretize the the boundary curve as a polyline and use triangle \cite{shewchuk96b, SHEWCHUK200221} to form a boundary conforming triangulation.
 The vertices of this triangulation serve as fixed set of high resolution UV samples for constructing a render mesh.
@@
 %correctness
-We first validated the physical plausibility of our method using a small, 2D, patch test (\reffig{patchtest}). 
+We first validate the physical plausibility of our method using a small 2D patch test (see \reffig{patchtest}). 
 Here the test object is a square made of four edges (not joined at the corners) simulated using linear VEM with a single deformation center.
@@
   \includegraphics[width=\columnwidth]{figures/patch_test.pdf}
-  \caption{2D examples (left) undeformed model defined by four surface elements on the boundary (right) applying affine transformations to the boundary and show that solving the static problem gives rigid motion for rotation and constant strain for shearing and stretching}
+  \caption{2D patch test.
+By applying affine transformations to the boundary of the original undeformed model (left), we show that solving the static problem gives rigid motion for rotation and constant strain for shearing and stretching (right).}
+
+  % \caption{2D patch test.
+Left: the original undeformed model defined by four surface elements on the boundary.
+Right:  applying affine transformations to the boundary and show that solving the static problem gives rigid motion for rotation and constant strain for shearing and stretching}
   \label{fig:patchtest}
@@
 
-Next we check the convergence of SEM with respect to linear, tetrahedral finite elements. \reffig{convergence} we compare identical cantilevered beams, simulated using 
-a Neohookean constitutive model (Young's Modulus $0.001$ GPa, Poisson's Ratio 0.45) and implicit time integration with a timestep of $0.01$s. 
-We use SEM with quadratic polynomials for this test and observe that our SEM simulation, made of 24 independent NURBS patches shows good agreement with FEM.
+We further demonstrate the convergence of our SEM with respect to linear tetrahedral finite elements when increasing the number of patches in use.
+In \reffig{convergence}, we compare identical cantilevered beams, simulated using a Neohookean constitutive model (Young's Modulus $=$ $0.001$ GPa, Poisson's Ratio $=$ 0.45) and implicit time integration with a timestep of $0.01$s.
+We use SEM with quadratic polynomials for this test and observe that our SEM simulation, made of 24 independent NURBS patches, shows good agreement with FEM.
 Each subdivision of the NURBS beam enables more complicated kinematics, but very few surface elements are needed to produce compelling results.
@@
   \includegraphics[width=\columnwidth]{figures/beams.pdf}
-  \caption{(left) high res FEM cantilevered bar (right) several simulations using more and more surface.}
+  \caption{Our SEM simulation is able to converges to high resolution FEM simulation result (left) as we increase the number of surfaces in the model (right). }
+  % \caption{(left) high res FEM cantilevered bar (right) several simulations using more and more surface.}
   \label{fig:convergence}
@@
 
-We also demonstrate that our raycasting weight computation creates shape-aware output. Figure~\reffig{independence} shows that manipulating
-parts that are nearby but separate will behave in an appropriately independent fashion. 
+We also show that our raycasting weight computation is able to create shape-aware output. \reffig{independence} shows that manipulating
+parts that are nearby but separated will behave in an appropriately independent fashion. 
 \begin{figure}[h]
   \includegraphics[width=\columnwidth]{figures/independent_movement}
-  \caption{X shape and show arms of X move independently}
+  \caption{Our raycasting weight computation produces share-aware blending weights. Here the locality of our blending weights allows the two legs of the Gumby model to move independently. }
   \label{fig:independence}
@@
 % relaxed modelling constraints
-Here we demonstrate SEM's ability to work with poor models. \reffig{badmodels} shows simulations of two jelly coffee mugs. 
+By virtue of its meshless nature, our SEM is robust to a wide range of challenging models with large gaps, intersections and disconnected primitives.
+\reffig{badmodels} shows simulations of two jelly coffee mugs. 
 The top row shows a model with negligible overlap, whereas the bottom row shows the result of a careless modeller who has deeply embedded the handle of the mug in the body in order to attach it, creating a large area of overlap.
-In both cases SEM produces a plausible physically-based animation, without requiring additional model clean-up.
+In both cases SEM produces a plausible physically-based animation without requiring additional model clean-up.
 \begin{figure}[h]
   \includegraphics[width=\columnwidth]{figures/mug_overlap}
-  \caption{Coffee mug with overlapping handle (top) row of models with increasing overlap (middle) single weight image showing behaviour in the overlapping region, (bottom) simulation result for each one}
+  \caption{Our SEM is robust to overlapping regions and self-intersections.
+Here the simulation result of the coffee mug with overlapping handle (bottom) still matches its corresponding counterpart without negligible overlaps (top).
+}
+  % \caption{Coffee mug with overlapping handle.
+(top) row of models with increasing overlap (middle) single weight image showing behaviour in the overlapping region, (bottom) simulation result for each one}
   \label{fig:badmodels}
@@
 %different material parameters / material models 
-We derive the equations of motion for SEM using a general elastic potential and so SEM supports arbitrary constitutive models. \reffig{materials} shows a 
-rocket ship simulated with both high stiffness and low stiffness materials. 
+Leveraging the fact that the equations of motion for our SEM are derived using a general elastic potential, our SEM is theoretically guaranteed to support arbitrary constitutive models.
+% We derive the equations of motion for SEM using a general elastic potential and thus SEM supports arbitrary constitutive models. 
+\reffig{rocket} shows a rocket ship simulated with both high stiffness and low stiffness materials. 
+%multiple materials
+Oue SEM also supports the simulation of objects made up of heterogenous materials (see \reffig{tire}).
 \begin{figure*}[htp]
   \includegraphics[width=\textwidth]{figures/rocket.pdf}
-  \caption{Big image simulating the rocket with all different combinations of things}
+  \caption{Our SEM is able to handle models with open surfaces. Here we show a rocket (with open surfaces at the bottom) lands on the ground under gravity. }
   \label{fig:rocket}
@@
 
-%multiple materials
-SEM also supports objects made up of multiple materials, as the simulation of this tire~(\reffig{tire}) illustrates.
 \begin{figure}[h]
   \includegraphics[width=\columnwidth]{figures/tire}
-  \caption{Tire to show heterogenous materials}
+  \caption{Our SEM is directly applicable to the simulation of objects with heterogenous materials.}
   \label{fig:tire}
@@
 
-%different surface reps
-Though we focus on NURBS surface representations, the only NURBS specific construction in the SEM method is the Jacobian\dave{ref}.
-Many other boundary representations admit a linear representation and are thus directly consumable by our method.
-\reffig{subd} shows an example of applying our SEM method to a subdivision model by making such a modification.
-\begin{figure}[h]
-  \includegraphics[width=\columnwidth]{example-image-a}
-  \caption{show subd simulation if we have it. If it works really well we'll just mix subds and nurbs throughout the paper. }
-  \label{fig:subd}
-\end{figure}
+% %different surface reps
+% Though we focus on NURBS surface representations, the only NURBS specific construction in the SEM method is the Jacobian \refeq{velocity_jacobian}.
+% Many other boundary representations admit a linear representation and are thus directly consumable by our method.
+% \reffig{subd} shows an example of applying our SEM method to a subdivision model by making such a modification.
+% \begin{figure}[h]
+%   \includegraphics[width=\columnwidth]{example-image-a}
+%   \caption{show subd simulation if we have it. If it works really well we'll just mix subds and nurbs throughout the paper. }
+%   \label{fig:subd}
+% \end{figure}
 
@@
   \includegraphics[width=\columnwidth]{example-image-a}
-  \caption{show mug simulation using newton's method, linearly implicit Euler (and maybe RK4). }
+  \caption{show mug simulation using newton's method, linearly implicit Euler \Honglin{Maybe I can do that quickly}. }
   \label{fig:time_integration}
@@
 %editing example
-We show that the output of the SEM simulation is itself edible in a NURBS modelling program~(\reffig{edit}).
+For downstream applications, the output of our SEM simulation is itself editable in a NURBS modelling program (see \reffig{edit}). 
 \begin{figure}[h]
@@
 %show stoppers
-Finally we show the ability of SEM to handle complex simulations involving deformation, contact and friction~(\reffig{staypuft} and \reffig{f1}).
+Finally we show the ability of SEM to handle complex simulations involving deformation, contact and friction.
+In \reffig{staypuft} and \reffig{starship}, we show two staypuft characters collide with each other in space and a Space-V starship lands on Mars in a graceful way.
 \begin{figure*}[htp]
@@
   \caption{The Space-V starship makes a graceful landing on the red planet. }
-  \label{fig:f1}
+  \label{fig:starship}
 \end{figure*}

### 2021-01-27 paper/sections/methods.tex

@@
 The output is an elastodynamic simulation. 
-SEM consists of preprocessing and runtime simulation phases~\refalg{sem} and at no point do we need to generate a volumetric mesh of the input. 
+SEM consists of preprocessing and runtime simulation phases (~\refalg{sem_preprocess}, ~\refalg{sem_simulation}) and at no point do we need to generate a volumetric mesh of the input. 
 In the \emph{preprocessing} stage we use raycasting to find quadrature points and weights for volumetric integration, as well as to compute part blending parameters. 
@@
 \begin{algorithm}[h]
-    \label{alg:sem}
-	\caption{Shape Matching Element Algorithm}\label{sem_algorithm}
 	\begin{algorithmic}[1]
-		\Procedure{PrepareSim}{$model$}
+		\Procedure{Preprocess}{$model$}
 		\State{// Get initial DOF and sample boundary  \Comment{Section \ref{sec:geometric_model}}}
@@
      	\State $\ME \gets \text{ErrorMatrix}(\vc{x^0}, \XBs, \st{W},\Pi)$  \Comment{Section \ref{ssec:error_energy}}
-%     	\myworries{todo: add volume}
-%     	\While {$simulating$}
-%     		\State{$\Pcox \gets \Pi\mathbf{q}$}
-%     		\ForAll{$i \gets 1$ to $|\st{X}|$}
-%     			\State{$\mathbf{F}_i \gets \frac{\partial \mathbf{F}}{\partial c} c$}
-%				\State{$\frac{\partial V}{\partial c} \gets \frac{\partial V}{\partial c} +
-%				\frac{\partial \mathbf{F}}{\partial c}^T \frac{\partial \psi}{\partial \mathbf{F}}(\mathbf{F}_i)$}
-%				
-%				\State{$\frac{\partial^2 V}{\partial c^2} \gets \frac{\partial^2 V}{\partial c^2} +
-%				 \frac{\partial \mathbf{F}}{\partial c}^T 
-%				 \frac{\partial^2 \psi}{\partial \mathbf{F}^2}(\mathbf{F}_i)
-%				 \frac{\partial \mathbf{F}}{\partial c}
-%				 $}
-%  			\EndFor
-%  			\State{$K \gets \Pi^T \frac{\partial^2 V}{\partial c^2} \Pi$}
-%  			\State{$f_{\text{internal}} \gets \Pi^T \frac{\partial V}{\partial c}$}
-%  			\State{$f_{\text{error}} \gets \ME \vc{q}$}
-%  			\State{$f_{\text{total}} \gets f_{\text{internal}} + f_{\text{error}} + f_{\text{ext}}$}
-%  			\State{$\dot{\mathbf{q}} \gets \text{TimeIntegration}
-%  			(f_{\text{total}},\mathbf{K}, \dot{\mathbf{q}}, \mathbf{J} , \mathbf{M}, \mathbf{M_E})$}
-%  			\State{$\mathbf{q} \gets \mathbf{q} + h\dot{\mathbf{q}}$}
-%     	\EndWhile
 	\EndProcedure
 \end{algorithmic}
-\caption{Shape Matching Element Algorithm}
+\caption{Shape Matching Element Preprocessing}
+\label{alg:sem_preprocess}
 \end{algorithm}
@@
 \label{eqn:nurbs_srf}
-\defX\left(u,v\right) = \sum_{i=1}^n\sum_{j=1}^n \phi_{ij}\left(u,v\right)\vc{q}_{ij},
+\defX\left(u,v\right) = \sum_{i=1}^n\sum_{j=1}^m \phi_{ij}\left(u,v\right)\vc{q}_{ij},
     %\mathbf{x}(u,v) = \frac{\sum_{i=1}^{n}\sum_{j=1}^{m}  \mathbf{p}_{i,j} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
@@
 \end{equation}
-where $u$ and $v$ $\in \real$ are the coordinates of the 2D parametric domain, $n$ and $m$ are the number of control points in the $u$ and $v$ directions, $\mathbf{q}_{ij}\in\real^3$ are the control points and $\phi_{ij}\left(u,v\right)\in\real$ are the NURBS
+where $u$ and $v$ $\in \real$ are the coordinates in the 2D parametric domain, $n$ and $m$ are the number of control points in the $u$ and $v$ directions, $\mathbf{q}_{ij}\in\real^3$ are the control points and $\phi_{ij}\left(u,v\right)\in\real$ are the NURBS
 basis functions given by 
@@
     \phi_{ij}\left(u,v\right) = \frac{w_{i,j}B_{ik}(u)B_{jl}(v)}{\sum_{r=1}^{n}\sum_{s=1}^{m} w_{ij} B_{ik}(u)B_{jl}(v)},
-\end{equation*} where $B_{ik}$ (resp $B_{jl}$) is the $i^{th}$ ($j^{th}$) b-spline basis of order $k$ ($l$).
+\end{equation*} where $B_{ik}$ (resp. $B_{jl}$) is the $i^{th}$ ($j^{th}$) B-spline basis of order $k$ ($l$).
 
-Exploiting linearity with respect to the control points  we can express this mapping as 
+Exploiting linearity with respect to the control points, we can express this mapping as
 \begin{equation}
     \defX\left(u,v\right) = \underbrace{\begin{pmatrix} \phi_{11}\left(u,v\right)\ident & \dots & \phi_{nn}\left(u,v\right)\ident \end{pmatrix}}_{\Jnurbs\left(u,v\right)}
-    \underbrace{\begin{pmatrix} \vc{q}_{11} \\ \vdots \\ \vc{q}_{nn}\end{pmatrix}}_{\vc{q}},
+    \underbrace{\begin{pmatrix} \vc{q}_{11} \\ \vdots \\ \vc{q}_{nm}\end{pmatrix}}_{\vc{q}},
 \end{equation} where $\Jnurbs\left(u,v\right)$ is the NURBS Jacobian, $\ident$ is the $3\times3$ identity matrix and $\vc{q}$ is the vector of generalized coordinates~\dave{cite Lanczos}.
@@
 
+\begin{algorithm}[h]
+	\begin{algorithmic}[1]
+		\Procedure{Preprocess}{$model$}
+     	\myworries{todo: add volume}
+     	\While {$simulating$}
+     		\State{$\Pcox \gets \Pi\mathbf{q}$}
+     		\ForAll{$i \gets 1$ to $|\st{X}|$}
+     			\State{$\mathbf{F}_i \gets \frac{\partial \mathbf{F}}{\partial c} c$}
+				\State{$\frac{\partial V}{\partial c} \gets \frac{\partial V}{\partial c} +
+				\frac{\partial \mathbf{F}}{\partial c}^T \frac{\partial \psi}{\partial \mathbf{F}}(\mathbf{F}_i)$}
+
+				\State{$\frac{\partial^2 V}{\partial c^2} \gets \frac{\partial^2 V}{\partial c^2} +
+				 \frac{\partial \mathbf{F}}{\partial c}^T 
+				 \frac{\partial^2 \psi}{\partial \mathbf{F}^2}(\mathbf{F}_i)
+				 \frac{\partial \mathbf{F}}{\partial c}
+				 $}
+  			\EndFor
+  			\State{$K \gets \Pi^T \frac{\partial^2 V}{\partial c^2} \Pi$}
+  			\State{$f_{\text{internal}} \gets \Pi^T \frac{\partial V}{\partial c}$}
+  			\State{$f_{\text{error}} \gets \ME \vc{q}$}
+  			\State{$f_{\text{total}} \gets f_{\text{internal}} + f_{\text{error}} + f_{\text{ext}}$}
+  			\State{$\dot{\mathbf{q}} \gets \text{TimeIntegration}
+  			(f_{\text{total}},\mathbf{K}, \dot{\mathbf{q}}, \mathbf{J} , \mathbf{M}, \mathbf{M_E})$}
+  			\State{$\mathbf{q} \gets \mathbf{q} + h\dot{\mathbf{q}}$}
+     	\EndWhile
+	\EndProcedure
+\end{algorithmic}
+\caption{Shape Matching Element Simulation Loop}
+\label{alg:sem_simulation}
+\end{algorithm}
+
 \subsection{Generalized Velocity}

### 2021-01-27 paper/sections/methods.tex
section 5 first pass done

@@
 We begin by describing shape matching to a single NURBS part. 
-Given a two different configurations of the same part (specified by the control points), shape matching computes a polynomial deformation which best aligns them.
+Given two different configurations of the same part (specified by the control points), shape matching computes a polynomial deformation that best aligns them.
 
-Let $\vc{q^0}$ be the initial control points of the part, provided by the modeller and Let $\vc{q}$ be a vector of modified control point values.
+Let $\vc{q^0}$ be the initial control points of the part, provided by the modeller, and $\vc{q}$ be a vector of modified control point values.
 We can compute the undeformed (or reference) position of any point on our part using $\refX\left(u,v\right) = \Jnurbs\left(u,v\right)\mathbf{q^0}$ and the corresponding
@@
 
-\dave{Ty: need some text to explain how we pick points on the nurbs surface.}
-\ty{Need to think a little more on how to write this.
-The main idea is that each surface has a knot vector and each Bspline covers a span of those knots.
-Therefore to guarantee we represent all control points for a surface, we uniformly sample $d+1$ points in each span, where $d$ is the NURBS surface}
 We discretize the shape matching energy and minimize by solving
@@
 \label{eq:shapematch}
-\end{equation} Here $\samp{\Pmat}$ (resp. $\samp{\Jnurbs}$) is a $3s \times 3k$ (resp. $3s \times 3nm$) matrix where $s$ is the number of quadrature points, 
+\end{equation} Here $\samp{\Pmat}$ (resp. $\samp{\Jnurbs}$) is a $3s \times 3(k+1)$ (resp. $3s \times 3nm$) matrix where $s$ is the number of quadrature points, 
 $k$ is the number of  terms in the shape matching polynomial and $3nm$ is the number of NURBS control points. 
@@
 $\wt$ is a diagonal, $3s \times 3s$ matrix of integration weights.
+
+Quadrature points are selected so that $\samp{\Jnurbs}$ is of sufficient rank required to represent all control points.
+We achieve this by sampling over each knot span defined by the  B-splines, and find it effective to use the order of the NURBS as the number of quadrature points sampled in each knot span.
+\ty{the number of samples may be different if order in the u direction and v differ, but I figured it was best to keep this simple.}
 
@@
 We interpret shape matching as converting our NURBS part into a single Virtual Element suitable for elastodynamics simulation.
-Unfortunately a single polynomial deformation space will not be suitably expressive for even moderately complex models. 
+Unfortunately, a single polynomial deformation space will not be suitably expressive for even moderately complex models.
 Previous work~\cite{10.1145/1073204.1073216,10.1145/1275808.1276480} used additional lattices or partitions of the reference space to 
-increase the number of polynomials in the shape matching. Next we outline our meshless Shape Element Method which works directly on the NURBS
+increase the number of polynomials in the shape matching. Next we outline our meshless Shape Matching Element Method which works directly on the NURBS
 geometry itself.
@@
  \begin{equation}
-    \defX\left(\refX\right) = \sum_{\prt_i\in\Prts} \bld_i\Pmatxi\Pi_i\vc{q}_i, 
+    \defX\left(\refX\right) = \sum_{\prt_i\in\Prts} \bldxi\Pmatxi\Pi_i\vc{q}_i, 
     \label{eq:kinematics}
- \end{equation} where $w_i$ $\Pmatx_i$, $\Pi_i$ and $\vc{q}_i$ are the blending weights, polynomial basis, projection operator and control points for the $i^{th}$
+ \end{equation} where $\bldxi$, $\Pmatx_i$, $\Pi_i$ and $\vc{q}_i$ are the blending weights, polynomial basis, projection operator and control points for the $i^{th}$
  part respectively.

### 2021-01-27 paper/sections/conclusion.tex
conclusion

--- 2021-01-26 paper/sections/conclusion.tex
@@
 \section{Future Work and Conclusions}
-We have presented the Shape Matching Element Method (SEM), the first completely meshless approach to direct simulation of NURBS models.
-Our approach is unique in its ability to infer volumetric shape from surface only models,
-including ones with intersecting geometry, between parts and other defects common to non-engineering models.
-Our method allows simulation of such input directly, with a variety of contitutive models and time integration schemes
-and is compatibale with standard methods for collision resolution. 
+We have presented the Shape Matching Element Method (SEM), the first completely meshless approach to direct simulation of curved surface models, made from NURBS primitives.
+Our approach is unique in its ability to infer volumetric shape from surface only input,
+including input with intersecting geometry between parts and other defects common to non-engineering models.
 We believe that SEM is a significant improvement over standard physics-based animation pipelines for NURBS models. 
@@
 Their approach could be used to remove this parameter from the approach entirely.
+Finally, in this paper we have focused on applying SEM to physics-based animation, there is a significant amount of additional work needed to 
+extend SEM reliably into engineering applications.
 

### 2021-01-27 paper/sections/conclusion.tex
intro done

@@
 including input with intersecting geometry between parts and other defects common to non-engineering models.
-We believe that SEM is a significant improvement over standard physics-based animation pipelines for NURBS models. 
-As evidence of the efficacy of our approach, the authors submit that many of the examples in this paper were constructed ourselves
+We believe that SEM is a significant improvement over standard physics-based animation pipelines. 
+As evidence of this, the authors submit that many of the examples in this paper were constructed ourselves
 (since the standard graphics menagerie is not available as NURBS models). 
@@
 
-SEM as it's presented here is only the beginning and we believe there are many exciting areas of future work to explore.
+SEM, as it's presented here, is in its infancy and we believe there are many exciting areas of future work to explore.
 We are very excited to couple SEM to machine learning approaches for design, parameter estimation and Real2Sim applications.
@@
 
-SEM itself has much room for addition. 
+SEM itself has much room for improvement. 
 First, while we focus on NURBS surfaces here, the only part of SEM that is NURBS specific is the shape matching operation.
 We believe there is potential to allow mixed models (models which include polygonal meshes, particles, subdivision surfaces and NURBS)
-by extending the range of shape matching operations used by the algorithm. The shape matching operation itself could be improved to be material aware 
+by extending the range of shape matching operations used by the algorithm. The shape matching operation itself could be improved to be material-aware 
 (to better handle heterogenous materials) or to be robust to noisy data (allowing direct simulation of scanned data). 
 It is also interesting to consider the relationship between our error term and the orthogonality constraint proposed by~\citet{Zhang:CompDynamics:2020}. 
-Their approach could be used to remove this parameter from the approach entirely.
+Their approach could be used to remove this parameter from SEM entirely.
 Finally, in this paper we have focused on applying SEM to physics-based animation, there is a significant amount of additional work needed to 

### 2021-01-27 paper/sections/methods.tex
methods 1 done

@@
 
-At runtime we use standard time integration schemes to time step the system (we show linearly implicit, fully implicit, runge-kutta four). 
-These require us to evaluate the energy, gradient and Hessian (implicit only) which we do at the previously computed quadrature points. 
-For collisions we use penalty forces and apply them as external forces during integration.
+At runtime we use standard time integration schemes to time step the system. 
+This requires us to evaluate the energy, gradient and Hessian (implicit only) which we do at the previously computed quadrature points. 
+For collisions we use penalty forces and apply them as external forces during integration (though other methods could be used).
 
@@
 		\State{// Create projection operator}
-		\State $\Pi \gets \text{ShapeMatching}(\vc{q}, \samp{\Jnurbs}, \wt, \XBs)$ \Comment{Section \ref{sec:shapematching}}
+		\State $\Pi \gets \text{ShapeMatching}(\vc{q}, \samp{\Jnurbs}, \st{\wt}, \XBs)$ \Comment{Section \ref{sec:shapematching}}
 		\Statex
@@
 
-Our algorithm acts on objects composed of multiple NURBS surfaces~\reffig{NURBS}.
-
-\begin{figure}
+Our algorithm acts on objects composed of multiple NURBS surfaces~(\reffig{NURBS}).
+\begin{figure}[h]
     \includegraphics[width=\columnwidth]{figures/nurbs_patch}
@@
 \end{figure}
-
 The three-dimensional position, $\defX$, of any point on a NURBS surface can be written as 
@@
 \begin{equation*}
-    \phi_{ij}\left(u,v\right) = \frac{w_{i,j}B_{ik}(u)B_{jl}(v)}{\sum_{r=1}^{n}\sum_{s=1}^{m} w_{ij} B_{ik}(u)B_{jl}(v)},
-\end{equation*} where $B_{ik}$ (resp. $B_{jl}$) is the $i^{th}$ ($j^{th}$) B-spline basis of order $k$ ($l$).
+    \phi_{ij}\left(u,v\right) = \frac{\omega_{i,j}\beta_{ik}(u)\beta_{jl}(v)}{\sum_{r=1}^{n}\sum_{s=1}^{m} \omega_{ij} \beta_{ik}(u)\beta_{jl}(v)},
+\end{equation*} where $\beta_{ik}$ (resp. $\beta_{jl}$) is the $i^{th}$ ($j^{th}$) B-spline basis of order $k$ ($l$).
 
@@
 \begin{equation}
-    \defX\left(u,v\right) = \underbrace{\begin{pmatrix} \phi_{11}\left(u,v\right)\ident & \dots & \phi_{nn}\left(u,v\right)\ident \end{pmatrix}}_{\Jnurbs\left(u,v\right)}
+    \defX\left(u,v\right) = \underbrace{\begin{pmatrix} \phi_{11}\left(u,v\right)\ident & \dots & \phi_{nm}\left(u,v\right)\ident \end{pmatrix}}_{\Jnurbs\left(u,v\right)}
     \underbrace{\begin{pmatrix} \vc{q}_{11} \\ \vdots \\ \vc{q}_{nm}\end{pmatrix}}_{\vc{q}},
-\end{equation} where $\Jnurbs\left(u,v\right)$ is the NURBS Jacobian, $\ident$ is the $3\times3$ identity matrix and $\vc{q}$ is the vector of generalized coordinates~\dave{cite Lanczos}.
+\end{equation} where $\Jnurbs\left(u,v\right)$ is the NURBS Jacobian, $\ident$ is the $3\times3$ identity matrix and $\vc{q}$ is the vector of generalized coordinates~\cite{lanczos2012variational}
 The Shape Matching Element Method will follow from this kinematic description.
@@
 
-Let $\vc{q^0}$ be the initial control points of the part, provided by the modeller, and $\vc{q}$ be a vector of modified control point values.
+Let $\vc{q^0}$ be the initial control points of the part, provided by the modeler, and $\vc{q}$ be a vector of modified control point values.
 We can compute the undeformed (or reference) position of any point on our part using $\refX\left(u,v\right) = \Jnurbs\left(u,v\right)\mathbf{q^0}$ and the corresponding
@@
     \defX\left(\refX\right) = 
-    \underbrace{\begin{pmatrix}
+    \underbrace{\begin{bmatrix}
     \mathbf{p}^T(\refX-\bar{\refX}) & \mathbf{0} & \mathbf{0} & 1 & 0 & 0\\
@@
     \mathbf{0} & \mathbf{0} & \mathbf{p}^T(\refX-\bar{\refX}) & 0 & 0 & 1\\
-    \end{pmatrix}}_{\Pmatx}
+    \end{bmatrix}}_{\Pmatx}
     \Pcox,
     \label{eq:projection}
-\end{equation} where $\mathbf{p}^T\left(\refX\right)$ is the polynomial basis vector \newline (eg $\begin{pmatrix} X & Y & X\cdot Y & X\cdot X & Y\cdot Y \end{pmatrix}$ ),
+\end{equation} where $\mathbf{p}^T\left(\refX\right)$ is the polynomial basis vector \newline (eg $\begin{bmatrix} X & Y & Z & X\cdot Y & X\cdot Z & Y\cdot Z& X\cdot X & Y\cdot Y & Z\cdot Z \end{bmatrix}$ ),
 $\bar{\refX}$ is an a priori chosen origin around which to deform, $\Pco$ are the coefficients for the 
@@
 Previous work~\cite{10.1145/1073204.1073216,10.1145/1275808.1276480} used additional lattices or partitions of the reference space to 
-increase the number of polynomials in the shape matching. Next we outline our meshless Shape Matching Element Method which works directly on the NURBS
+increase the number of polynomials in the shape matching. In contrast, our meshless Shape Matching Element Method which works directly on the NURBS
 geometry itself.
@@
     \label{eq:kinematics}
- \end{equation} where $\bldxi$, $\Pmatx_i$, $\Pi_i$ and $\vc{q}_i$ are the blending weights, polynomial basis, projection operator and control points for the $i^{th}$
+ \end{equation} where $\bldxi$, $\Pmatxi$, $\Pi_i$ and $\vc{q}_i$ are the blending weights, polynomial basis, projection operator and control points for the $i^{th}$
  part respectively.

### 2021-01-27 paper/sections/methods.tex
all reviewed up to 6.2

--- 2021-01-27 paper/sections/methods.tex
@@
  part respectively.  
+ \ty{I'm wondering if we should add a line emphasizing that while we linearly blend, the resulting deformation map is a nonlinear map unique to each X. But maybe this obvious}
 
@@
 Using the same origin for all polynomials limits the expressivity of the kinematic model~\cite{STBS:2011}.
-Instead we choose a number of points in $\refX$ ($|\Prts| > |\XBs| > 0$, where $\XBs$ is the set of points)  to act as deformation origins~(\refsec{origins}).
-For each part we now construct $\Pmatxi$ using the associated deformation origin $\bar{\refX}_j\in \XBs$, where each origin can
+Instead we choose a set of points $\XBs$ ($0 < |\XBs| < |\Prts|$) to act as deformation origins~(\refsec{origins}).
+For each part we now construct $\Pmatxi$ using its associated deformation origin $\bar{\refX}_j\in \XBs$, where each origin can
 be shared between multiple parts.
@@
     \includegraphics[width=\columnwidth]{figures/projection_operator_solve}
-    \caption{A simple, four part model with two deformation origins and the corresponding matrix equation for the quadratic coefficients.}
+    \caption{A simple, four part model with two deformation origins and the corresponding matrix equation for the quadratic coefficients.
+\ty{these $T^1$ and $T^2$ blocks in the U matrix need to be swapped I think} }
     \label{fig:multiparts}
@@
 
-For a given part, $\prt_i$, with associated origin $\bar{\refX}_j$ we first estimate the constant term, $\tr_i$ as the centroid of all parts associated with $\bar{X}_j$.
-We next fit the linear part of the deformation by performing linear shape matching to $\prt_i$. 
+For a given part, $\prt_i$, with associated origin $\bar{\refX}_j$ we first estimate the constant term, $\tr_i$ as the centroid of all parts associated with $\bar{\refX}_j$.
+Next we fit the linear part of the deformation by performing linear shape matching to $\prt_i$. 
 This can be repeated for polynomials of arbitrary high order, for instance here we apply this approach to quadratic shape matching:
@@
 Finally, $\wt_i$ is the diagonal integration weight matrix and $\mathbbm{1}\in\real^{3n\times3}$ is a matrix of stacked $3\times3$ identity matrices -- one for each quadrature point.
+\ty{it looks like we are missing the divisor for t?
+This looks like it's computing the sum and $\mathbbm{1}$ should be defined as $1/N$ where $N$ is the total number of points in all associated parts}
 
 %%single projector for every object
-When assembled for all parts,~\refeq{hfit}, yields a block upper triangular matrix $U$ and a spare matrix $B$ ~(\reffig{multiparts}) which can be used to efficiently to compute the projection 
+When assembled for all parts,~\refeq{hfit}, yields a block upper triangular matrix $U$ and a sparse matrix $B$ ~(\reffig{multiparts}) which can be used to efficiently to compute the projection 
 operator for the entire object $\Pi = U^{-1}B\vc{q}$. 
@@
 
-\dave{Ty: How do we deal with non-planar parts ? What if the ray cast hits the same part that the closest point is on?}
-\ty{Right now this doesn't happen as we only test intersection against other parts}
-\dave{can't that mean the ray can just miss everything ? If staypuft's body was just a single ellipse the wieghts inside should all be one, in this case they would be very odd.}
-We compute two distances, $d_{\text{primary}}$, the distance from $\mathbf{X}_j$ to $\clst$ and $d_{\text{total}}$ the distance from  $\clst$ to $\rhit$.
 Using these two distances we define a linear \emph{distance weight} using the well-known ReLU function:
@@
 These weights obey the Kronecker delta property for all parts but do not partition unity which we repair using a post-normalization step.
+For non-planar surfaces, $\rhit$ may lie on the same part as $\clst$.
+For points inside closed periodic surfaces, it is critical that the weights equal one in this case, so we set $\theta_i = 1$.
 
@@
 Finally, the distance cutoff $\alpha$ allows us to control the sparsity of the blending weights.
+We employ a simple heuristic to select $\alpha$ by searching for a $\alpha$ that produces a set of blending weights meeting a desired sparsity target $T \in \left( 0,1\right)$, which is defined as the ratio of non-zeros to total number of weights.
+This yields a unitless, easy to tune parameter.
 
@@
 but allows us to use more general weighting functions and apply the appropriate behavior for intersecting parts.
-The price we pay is that our weights are not C0 continuous but rather Lipchitz smooth.
+The price we pay is that our weights are not C0 continuous but rather Lipschitz smooth.
 Perturbations in the boundaries of parts can cause bounded jumps in our distance weights. 
-However, these jumps occupy an infinitesimal percentage of the volume of our object are easy to avoid when integrating
+However, these jumps occupy an infinitesimal percentage of the volume of our object, and are easy to avoid when integrating
 physical quantities, leaving our simulation unaffected.

### 2021-01-27 paper/sections/intro.tex

@@
 NURBS primitives were the first geometric representation used for physics-based animation, yet,
-despite over three decades of research, animation of NURBS objects remains a challenge.
+despite over three decades of research, animation of curved models such as those built with NURBS, remains a challenge.
 
@@
 Unfortunately IGA approaches for volumetric objects still require background volumetric structures, typically
-regular grids that make satisfaction of boundary conditions difficult (which makes collision resolution difficult)
+regular grids that make satisfaction of boundary conditions difficult (which makes collision resolution hard)
 or more complicated cut-cell grids which introduce non-trivial root finding problems into the mix. 
@@
 criteria such as that the mesh is manufacturable. 
-These inputs are much cleaner than those produced by a typical modeller.
+These inputs are much cleaner than those produced by a typical animation modeller.
 
@@
 
-Our approach is an extension of the recently developed Virtual Element Method (VEM) for solving PDEs on domains tiled with arbitrary polygons.
-We establish a connection between VEM and the well-known shape matching simulation algorithm  which enables us 
-to derive equations of motion for an arbitrary NURBS model using Lagrangian Mechanics. 
-Importantly, we show how to replace volumetric data structures for integration with ray casting approaches which enables
-our meshless approach to isogeometric elastodynamic physically-based animation of volumetric structures.
+Our approach draws inspiration from the recently developed Virtual Element Method (VEM) for solving partial differential equations on domains tiled with arbitrary polygons.
+We establish a connection between VEM and the well-known shape matching simulation algorithm. This allows us 
+to derive equations of motion for an arbitrary curved surface model, made of NURBS, via Lagrangian Mechanics. 
+Importantly, we show how to replace volumetric data structures for integration and blending weight computation with ray casting approaches which enables
+our meshless approach to elastodynamic physically-based animation.
 

### 2021-01-27 paper/sections/related.tex

@@
 %% Set the scene
-Geometric modelling is a necessary precursor to physics-based animation, however differing geometric representations for modelling and simulation often requires
+Geometric modeling is a necessary precursor to physics-based animation, however connecting differing geometric representations for modelling and simulation often requires
 time-consuming, complex geometry processing pipelines. 
@@
 to enable improved boundary handling~\cite{https://doi.org/10.1002/nme.3120,10.1007/s00466-013-0854-7,https://doi.org/10.1002/nme.4852,SAFDARI2016373,10.1145/3355089.3356576}. 
-These methods typically start with an easy-to-generate structured volumetric mesh (tetrahedral or hexahedral), ``cutting'' the NURBS geomtric model against
+These methods typically start with an easy-to-generate structured volumetric mesh (tetrahedral or hexahedral), ``cutting'' the NURBS geometric model against
 it to enable boundary handling (such a mesh is called a \emph{cutcell} mesh). 
@@
 engineering~\cite{10.1145/1028523.1028542,https://doi.org/10.1002/fld.1650200824,10.1145/1778765.1778776,10.1145/2010324.1964968, 10.1145/1944846.1944855}.
-These methods support more advanced constitutive models and integration schemes but often require background integration meshes and lose the direct connection with
+These methods support more advanced constitutive models and integration schemes but often require background integration meshes, limit themselves to low order deformation functions 
+and lose the direct connection with
 modelling geometry. 
@@
 VEM typically assumes that the boundary of the the problem domain is described by a piecewise linear complex
-which makes its standard formulation incompatible with our NURBS based input geometry.
+which makes its standard formulation incompatible with our curved input geometry.
 
@@
 intersections and disconnected primitives without additional user input. 
-These features mean that our algorithm is the first truly meshless approach for the direct simulation of NURBS models for physics-based animation.
+These features mean that our algorithm is the first truly meshless approach for the direct simulation of curved surface, NURBS models for physics-based animation.
 

### 2021-01-27 paper/sections/results.tex

@@
 
-\dave{we should show the nurbs model in rhino/fusion for every example, maybe an exploded view as well}
+\begin{table*}[h]
+  \caption{Performance of, and parameters for the Shape Matching Element Method on all examples. All timings are reported in seconds, physical  parameters are
+  reported with appropriate units. $\rho$ is the applied density, \textbf{E} is the  Young's Modulus and $\mu$ is the Poisson's ratio.
+  }
+  \label{tbl:perf}
+  \begin{center}
+  \begin{tabular}{l c c c c c c c c c c}
+   \textbf{Example} $|\vc{q}|$ & \textbf{Material}&  $\rho$(kg/m$^3$) & \textbf{E}(GPa) & $\mu$ &\textbf{Sample Model}(s) & \textbf{Quadrature}(s)& \textbf{Weights}(s)& \textbf{Build $\Pi$}(s)& \textbf{Step}(s)\\
+   \hline 
+   \rowcolor[HTML]{DAE8FC} 
+   \textbf{Cantilever}      & ? & ? & ? & ? & ? & ? & ? & ? & ? & ? \\
+   \textbf{Coffee mug}      & ?& ? & ? & ? & ? & ? & ? & ? & ? & ?  \\
+   \textbf{Rocket}          & ?& ? & 648 & ? & ? & ? & ? & ? & ? & ?\\
+   \textbf{Castle}          & ?& ? & ? & ? & ? & ? & ? & ? & ? & ?\\
+   \textbf{Chicken}        & ? & ? & ? & ? & ? & ? & ? & ? & ? & ?\\
+   \textbf{Starship}       & ? & ? & ? & ? & ? & ? & ? & ? & ? & ?\\
+   \textbf{Tire}           & ? & ? & ? & ? & ? & ? & ? & ? & ? & ?\\
+   \hline
+  \end{tabular}
+  \end{center}
 
-\dave{mention that not so many standard NURBS models so we modelled lots of these ourselves .. show's power of approach that we can model and sim all these examples.}
+  \end{table*}
 
-\dave{use as many different meshes as possible for figures (i.e try not to have the same mesh show up in more than one figure if possible).}
+%\dave{we should show the nurbs model in rhino/fusion for every example, maybe an exploded view as well}
 
-\dave{we should report which models we made ourselves}
+%\dave{mention that not so many standard NURBS models so we modelled lots of these ourselves .. show's power of approach that we can model and sim all these examples.}
 
-\begin{figure*}
-  \includegraphics[width=\textwidth]{figures/chicken.pdf}
-  \caption{Simulation of a chicken model with gaps between surfaces.}
-  \label{fig:chicken}
-\end{figure*}
+%\dave{use as many different meshes as possible for figures (i.e try not to have the same mesh show up in more than one figure if possible).}
 
-\reftbl{perf} shows the size of all our examples along with performance statistics. 
+%\dave{we should report which models we made ourselves}
+We implemented SEM in a combination of MATLAB and C++ using both GPToolbox~\cite{gptoolbox} and Bartels~\cite{bartels} for geometry processing and constitutive models 
+respectively. Benchmarks were performed using a  ~\dave{CPU} with \dave{xx}MB of RAM.
+
+\reftbl{perf} shows the size of all our examples along with performance statistics and relevant parameters.
 Note that the individual parts of all models are \emph{not connected}, and no continuity constraints or constructive solid geometry operations have been applied. 
@@
 
-For raycasting operations and rendering, we triangulate the NURBS surfaces, which only represents an implementation simplification.
+For raycasting operations and rendering we rely on Embree~\cite{10.1145/2601097.2601199} and triangulate the NURBS surfaces, which only represents an implementation simplification. 
 Our algorithm is not dependent on triangulating the boundary, and an ideal alternative would be using a path tracer.
 
-\ty{Would be nice to mention TNURBS, which would give us a good reason to include the wrench (which i really wanna do).
+\nicetohave{Would be nice to mention TNURBS, which would give us a good reason to include the wrench (which i really wanna do).
 I doubt we have time to do a 3 point bending test, but a simple elastic wrench simulation would be easy to cook up}
@@
 The vertices of this triangulation serve as fixed set of high resolution UV samples for constructing a render mesh.
-
-\begin{table*}[h]
-  \caption{Performance of the Shape Matching Element Method on all Examples. \dave{need table that also reports material params/constitutive model or add it here}}
-  \label{tbl:perf}
-  \begin{center}
-  \begin{tabular}{l c c c c c c c c}
-   \textbf{Example} & \textbf{CPU} & $|\vc{q}|$ &  \textbf{Sample Model} & \textbf{Quadrature} & \textbf{Weights} & \textbf{Build $\Pi$} & \textbf{Time Step} \\
-   \hline 
-   \rowcolor[HTML]{DAE8FC} 
-   \textbf{Cantilever}      & ? & ? & ? & ? & ? & ? & ? & ? \\
-   \textbf{Coffee mug}      & ? & ? & ? & ? & ? & ? & ? & ? \\
-   \textbf{Rocket}          & ? & 648 & ? & ? & ? & ? & ? & ? \\
-   \textbf{Castle}          & ? & ? & ? & ? & ? & ? & ? & ? \\
-   \textbf{Chicken}         & ? & ? & ? & ? & ? & ? & ? & ? \\
-   \textbf{Starship}        & ? & ? & ? & ? & ? & ? & ? & ? \\
-   \textbf{Tire}            & ? & ? & ? & ? & ? & ? & ? & ? \\
-   \hline
-  \end{tabular}
-  \end{center}
-
-  \end{table*}
 
@@
 %correctness
-We first validate the physical plausibility of our method using a small 2D patch test (see \reffig{patchtest}). 
-Here the test object is a square made of four edges (not joined at the corners) simulated using linear VEM with a single deformation center.
-We apply a battery of boundary conditions and resolve the deformation of the element by minimizing the elastic potential.
-We note that SEM is able to represent rigid motions, as well as shearing and anisotropic stretching. 
-This implies that, with sparse weights, SEM can resolve these motions locally, leading to physically plausible simulation results.
-
 \begin{figure}[h]
@@
 By applying affine transformations to the boundary of the original undeformed model (left), we show that solving the static problem gives rigid motion for rotation and constant strain for shearing and stretching (right).}
-
   % \caption{2D patch test.
@@
 
-We further demonstrate the convergence of our SEM with respect to linear tetrahedral finite elements when increasing the number of patches in use.
-In \reffig{convergence}, we compare identical cantilevered beams, simulated using a Neohookean constitutive model (Young's Modulus $=$ $0.001$ GPa, Poisson's Ratio $=$ 0.45) and implicit time integration with a timestep of $0.01$s.
+We first validate the physical plausibility of our method using a small 2D patch test~(\reffig{patchtest}). 
+Here the test object is a square made of four edges (not joined at the corners) simulated using linear polynomials with a single deformation center.
+We apply a battery of boundary conditions and resolve the deformation of the element by minimizing the elastic potential.
+We note that SEM is able to represent rigid motions, as well as shearing and anisotropic stretching. 
+This implies that, with sparse weights, SEM can resolve these motions locally, leading to physically plausible simulation results.
+
+We further demonstrate \emph{qualitative} convergence of SEM with respect to linear tetrahedral finite elements when increasing the number of patches in use.
+In \reffig{convergence}, we compare identical cantilevered beams, simulated using a Neohookean constitutive model (Young's Modulus $=$ $0.001$ GPa, Poisson's Ratio $=$ $0.45$) and implicit time integration with a timestep of $0.01$s.
 We use SEM with quadratic polynomials for this test and observe that our SEM simulation, made of 24 independent NURBS patches, shows good agreement with FEM.
-Each subdivision of the NURBS beam enables more complicated kinematics, but very few surface elements are needed to produce compelling results.
+Each subdivision of the NURBS beam enables more complicated kinematics,
+but very few surface elements are needed to produce compelling results.
 
@@
   \includegraphics[width=\columnwidth]{figures/beams.pdf}
-  \caption{Our SEM simulation is able to converges to high resolution FEM simulation result (left) as we increase the number of surfaces in the model (right). }
-  % \caption{(left) high res FEM cantilevered bar (right) several simulations using more and more surface.}
+  \caption{Our SEM simulation is able to qualitatively converge to a high resolution FEM simulation result (left) as we increase the number of surfaces in the model (right). }
   \label{fig:convergence}
@@
 We also show that our raycasting weight computation is able to create shape-aware output. \reffig{independence} shows that manipulating
-parts that are nearby but separated will behave in an appropriately independent fashion. 
+parts that are nearby but separated will behave in an appropriately independent fashion. Our grumpy model is capable of extending 
+its leg without causing unrealistic deformations in the plant foot.
+
 \begin{figure}[h]
   \includegraphics[width=\columnwidth]{figures/independent_movement}
-  \caption{Our raycasting weight computation produces share-aware blending weights. Here the locality of our blending weights allows the two legs of the Gumby model to move independently. }
+  \caption{Our raycasting weight computation produces share-aware blending weights. Here the locality of our blending weights allows the two legs of the grumpy model to move independently. }
   \label{fig:independence}
@@
 % relaxed modelling constraints
-By virtue of its meshless nature, our SEM is robust to a wide range of challenging models with large gaps, intersections and disconnected primitives.
-\reffig{badmodels} shows simulations of two jelly coffee mugs. 
+By virtue of its meshless nature, SEM is robust to a wide range of challenging models with large gaps and disconnected primitives.
+\reffig{fig:chicken} shows frames from a simulation of a rubber chicken. Note that the chicken model itself features large gaps between the individual NURBS parts. 
+Despite the lack of explicit connectivity, the SEM blending weights have the effect of implicitly enforcing connectivity at these seams.
+
+\begin{figure*}[htp]
+  \includegraphics[width=\textwidth]{figures/chicken.pdf}
+  \caption{Simulation of a chicken model with gaps between surfaces.}
+  \label{fig:chicken}
+\end{figure*}
+
+SEM is also robust to intersections in modelling input. \reffig{badmodels} shows simulations of two jelly coffee mugs. 
 The top row shows a model with negligible overlap, whereas the bottom row shows the result of a careless modeller who has deeply embedded the handle of the mug in the body in order to attach it, creating a large area of overlap.
 In both cases SEM produces a plausible physically-based animation without requiring additional model clean-up.
-\begin{figure}[h]
+
+\begin{figure}[H]
   \includegraphics[width=\columnwidth]{figures/mug_overlap}
@@
 \end{figure}
-Similarly, in \reffig{chicken} we show a simulation with gaps between surfaces.
-Despite the lack of explicit connectivity, the blending weights have the effect of implicitly enforcing connections at these seams.
 
 %different material parameters / material models 
-Leveraging the fact that the equations of motion for our SEM are derived using a general elastic potential, our SEM is theoretically guaranteed to support arbitrary constitutive models.
+Since the equations of motion for SEM are derived using a general elastic potential, it is theoretically capable of supporting arbitrary elastic constitutive models.
 % We derive the equations of motion for SEM using a general elastic potential and thus SEM supports arbitrary constitutive models. 
-\reffig{rocket} shows a rocket ship simulated with both high stiffness and low stiffness materials. 
+\reffig{rocket} shows a rocket ship simulated with both high stiffness (steel) and low stiffness materials (rubber). 
+In both cases intuitive and visually pleasing results are created wherein the trajectories correctly reflect the desired material characteristics.  
 %multiple materials
-Oue SEM also supports the simulation of objects made up of heterogenous materials (see \reffig{tire}).
 \begin{figure*}[htp]
   \includegraphics[width=\textwidth]{figures/rocket.pdf}
-  \caption{Our SEM is able to handle models with open surfaces. Here we show a rocket (with open surfaces at the bottom) lands on the ground under gravity. }
+  \caption{SEM is able to produce animation using a wide variety of material parameters. Here we simulate both a steel and rubber rocket ship. 
+  To make this even more challenging, the rocket model is an open surface (at the bottom), but a plausible animation is still generated.}
   \label{fig:rocket}
 \end{figure*}
+
+SEM also supports the simulation of objects made up of heterogenous materials. By specifying different material properties inside nested 
+parts we are able to simulate a drag racing wheel with a metal hub and soft rubber tread~(\reffig{tire}).
 
@@
 % different time integration scheme
+\nicetohave{Integration Figure}
+%\begin{figure}[h]
+%  \includegraphics[width=\columnwidth]{example-image-a}
+%  \caption{show mug simulation using newton's method, linearly implicit Euler \Honglin{Maybe I can do that quickly}. }
+%  \label{fig:time_integration}
+%\end{figure}
+
+%editing example
+For downstream applications, the output of  SEM simulation is itself editable in a NURBS modelling program (see \reffig{edit}). 
+This can be useful for visual effects pipelines, allowing artists to easily post process animations using the same tools in which they were created.
 \begin{figure}[h]
   \includegraphics[width=\columnwidth]{example-image-a}
-  \caption{show mug simulation using newton's method, linearly implicit Euler \Honglin{Maybe I can do that quickly}. }
-  \label{fig:time_integration}
-\end{figure}
-
-%editing example
-For downstream applications, the output of our SEM simulation is itself editable in a NURBS modelling program (see \reffig{edit}). 
-\begin{figure}[h]
-  \includegraphics[width=\columnwidth]{example-image-a}
-  \caption{show input to simulation, show output then show output loaded in rhino/fusion 360 }
+  \caption{Here we show the output of an SEM animation loaded into Rhinoceros 3D 7 for additional editing.}
   \label{fig:edit}
@@
 %show stoppers
-Finally we show the ability of SEM to handle complex simulations involving deformation, contact and friction.
-In \reffig{staypuft} and \reffig{starship}, we show two staypuft characters collide with each other in space and a Space-V starship lands on Mars in a graceful way.
+Finally we show a few additional examples, highlighting the ability of SEM to handle complex simulations involving deformation, contact and friction.
+In \reffig{staypuft} we show two astronauts in jaunty space suits collide in zero gravity while in \reffig{starship} 
+we show an animation of the Space-V starship landing on Mars in a graceful way.
+
 \begin{figure*}[htp]
   \includegraphics[width=\textwidth,height=3in]{example-image-a}
-  \caption{sequence of frames from staypuft simulation}
+  \caption{Two astronauts, in jaunty spacesuits collide while spacewalking in zero-g.}
   \label{fig:staypuft}

### 2021-01-27 paper/sections/methods.tex
methods

--- 2021-01-27 paper/sections/methods.tex
@@
 
+\nicetohave{now I do want that shape matching figure showing constant, linear and quadratic matching}
+
 \paragraph*{Relationship to the Virtual Element Method}
@@
 
-\begin{figure}[h]
-    \includegraphics[width=\columnwidth]{figures/projection_operator_solve}
-    \caption{A simple, four part model with two deformation origins and the corresponding matrix equation for the quadratic coefficients.}
-    \label{fig:multiparts}
-\end{figure}
-
 Attempting to use $\Pmatxi$ in~\refeq{shapematch} on a per-part basis is problematic as it can become ill-posed, especially for planar boundary representations. 
@@
 dominated by lower order effects~\cite{li2006regularities}. 
-Mechanically, this principle suggests that the amount of motion modelled by increasingly high-order polynomials is decreasing. 
+Mechanically, this principle suggests that the amount of motion modelled by increasingly high-order polynomials is, itself, decreasing. 
 This implies that we can achieve acceptable shape matching results by performing shape matching hierarchically -- starting with the constant term and then 
-fitting the remaining deformation with progressively higher-order deformations.
+fitting the remaining deformation with progressively higher-order expressions.
 
@@
     \label{eq:hfit} 
-\end{equation} where $\Pmat^l$ contains only the monomial basis of order $l$,  $\samp{\Pmat^1}$ is the stacked evaluation of this matrix at $n$ quadrature points and $\vc{c}^l_i$ are the corresponding coefficients.
-Finally, $\wt_i$ is the diagonal integration weight matrix and $\mathbbm{1}\in\real^{3n\times3}$ is a matrix of stacked $3\times3$ identity matrices -- one for each quadrature point.
+\end{equation} where $\Pmat^l$ contains only the monomial basis of order $l$,  $\samp{\Pmat^l}$ is the stacked evaluation of this matrix at the $s$ quadrature points and $\vc{c}^l_i$ are the corresponding coefficients.
+Finally, $\wt_i$ is the diagonal integration weight matrix and $\mathbbm{1}\in\real^{3s\times3}$ is a matrix of stacked $3\times3$ identity matrices -- one for each quadrature point.
+
+\begin{figure}[h]
+    \includegraphics[width=\columnwidth]{figures/projection_operator_solve}
+    \caption{A simple, four part model with two deformation origins and the corresponding matrix equation for the quadratic coefficients.}
+    \label{fig:multiparts}
+\end{figure}
 

### 2021-01-27 paper/sections/results.tex
table half-finished, will finish that next morning

@@
   \begin{tabular}{l c c c c c c c c c c}
-   \textbf{Example} $|\vc{q}|$ & \textbf{Material}&  $\rho$(kg/m$^3$) & \textbf{E}(GPa) & $\mu$ &\textbf{Sample Model}(s) & \textbf{Quadrature}(s)& \textbf{Weights}(s)& \textbf{Build $\Pi$}(s)& \textbf{Step}(s)\\
+   \textbf{Example} & $|\vc{q}|$ & \textbf{Material}&  $\rho$(kg/m$^3$) & \textbf{E}(Pa) & $\mu$ &\textbf{Sample Model}(s) & \textbf{Quadrature}(s)& \textbf{Weights}(s)& \textbf{Build $\Pi$}(s)& \textbf{Step}(s)\\
    \hline 
    \rowcolor[HTML]{DAE8FC} 
-   \textbf{Cantilever}      & ? & ? & ? & ? & ? & ? & ? & ? & ? & ? \\
-   \textbf{Coffee mug}      & ?& ? & ? & ? & ? & ? & ? & ? & ? & ?  \\
-   \textbf{Rocket}          & ?& ? & 648 & ? & ? & ? & ? & ? & ? & ?\\
-   \textbf{Castle}          & ?& ? & ? & ? & ? & ? & ? & ? & ? & ?\\
-   \textbf{Chicken}        & ? & ? & ? & ? & ? & ? & ? & ? & ? & ?\\
-   \textbf{Starship}       & ? & ? & ? & ? & ? & ? & ? & ? & ? & ?\\
-   \textbf{Tire}           & ? & ? & ? & ? & ? & ? & ? & ? & ? & ?\\
+   \textbf{Cantilever (soft)}       & ? & Jelly & $1 \times 10^3$ & $1 \times 10^6$& 0.45 & ? & ? & ? & ? & ? \\
+   \textbf{Cantilever (stiff)}      & ? & Rubber & $1 \times 10^3$ & $5 \times 10^6$ & 0.45 & ? & ? & ? & ? & ? \\
+   \textbf{Coffee mug}              & ? & Jelly & $1 \times 10^3$ & $1 \times 10^4$ & 0.47 & ? & ? & ? & ? & ?  \\
+   \textbf{Rocket (soft)}           & 648 & Rubber & $1.27 \times 10^3$ & $7 \times 10^6$ & 0.4 & ? & ? & ? & ? & ? \\
+   \textbf{Rocket (stiff)}          & 648 & Steel & $8 \times 10^3$ & $2 \times 10^{11}$ & 0.32 & ? & ? & ? & ? & ? \\
+   \textbf{Castle}                  & ? & Jelly & $1.27 \times 10^3$ & $2 \times 10^3$ & 0.45 & ? & ? & ? & ? & ? \\
+   \textbf{Chicken}                 & ? & Rubber & $1.27 \times 10^3$ & $1 \times 10^4$ & 0.47 & ? & ? & ? & ? & ? \\
+   \textbf{Starship}                & ? & Steel & $8 \times 10^3$ & $2 \times 10^{11}$ & 0.32 & ? & ? & ? & ? & ? \\
+   \textbf{Tire}                    & ? & Rubber & $2 \times 10^3$ & $1 \times 10^6$ & 0.47 & ? & ? & ? & ? & ? & ?\\
    \hline
@@
 We implemented SEM in a combination of MATLAB and C++ using both GPToolbox~\cite{gptoolbox} and Bartels~\cite{bartels} for geometry processing and constitutive models 
-respectively. Benchmarks were performed using a  ~\dave{CPU} with \dave{xx}MB of RAM.
+respectively. Benchmarks were performed using a MacBook Pro with an Intel i5 2.3GHz processor, 16GB of RAM and an Intel Iris Plus Graphics 655 GPU.
 

### 2021-01-27 paper/sections/methods.tex

--- 2021-01-27 paper/sections/methods.tex
@@
 
-\begin{figure}[h]
-    \includegraphics[width=\columnwidth]{figures/projection_operator_solve}
-    \caption{A simple, four part model with two deformation origins and the corresponding matrix equation for the quadratic coefficients.
-\ty{these $T^1$ and $T^2$ blocks in the U matrix need to be swapped I think} }
-    \label{fig:multiparts}
-\end{figure}
-
 Attempting to use $\Pmatxi$ in~\refeq{shapematch} on a per-part basis is problematic as it can become ill-posed, especially for planar boundary representations. 
@@
     \label{eq:hfit} 
-\end{equation} where $\Pmat^l$ contains only the monomial basis of order $l$,  $\samp{\Pmat^1}$ is the stacked evaluation of this matrix at $n$ quadrature points and $\vc{c}^l_i$ are the corresponding coefficients.
-Finally, $\wt_i$ is the diagonal integration weight matrix and $\mathbbm{1}\in\real^{3n\times3}$ is a matrix of stacked $3\times3$ identity matrices -- one for each quadrature point.
-\ty{it looks like we are missing the divisor for t?
-This looks like it's computing the sum and $\mathbbm{1}$ should be defined as $1/N$ where $N$ is the total number of points in all associated parts}
+\end{equation} where $\Pmat^l$ contains only the monomial basis of order $l$,  $\samp{\Pmat^l}$ is the stacked evaluation of this matrix at the $s$ quadrature points and $\vc{c}^l_i$ are the corresponding coefficients.
+Finally, $\wt_i$ is the diagonal integration weight matrix and $\mathbbm{1}\in\real^{3s\times3}$ is a matrix of stacked $3\times3$ identity matrices -- one for each quadrature point.
+
+\begin{figure}[h]
+    \includegraphics[width=\columnwidth]{figures/projection_operator_solve}
+    \caption{A simple, four part model with two deformation origins and the corresponding matrix equation for the quadratic coefficients.}
+    \label{fig:multiparts}
+\end{figure}
 
@@
 the $i^{th}$ part, $\prt_i$, and $\refX_j\in\st{X}$. 
-
 For each $\refX_j$ we find the nearest point on $\prt_i$, $\clst$.
@@
     \end{center}
-    \label{fig:plot_distance_weight}
-    \caption{Distance weight as a function of primary and total ray length with cutoff distance 50.}
+    \caption{ \label{fig:plot_distance_weight} Distance weight as a function of primary and total ray length with cutoff distance 50.}
   \end{wrapfigure}
@@
 While we constrain a part to be paired with only one deformation origin, an origin may be associated any number of parts. 
-Further, we require that each deformation origin be associated with a minimum of two parts, which avoids the case of an origin laying on the surface of a part.\ty{Should we mention periodic surfaces?
+Further, we require that each deformation origin be associated with a minimum of two parts, which avoids the case of an origin laying on the surface of a part.%\ty{Should we mention periodic surfaces?
 You can have a periodic nurbs sphere consisting of a single surface, which we couldn't handle with this constraint (easy fix, but not sure if worth mentioning)}\dave{yes}

### 2021-01-27 paper/sections/results.tex
fix table
fit table onto page

@@
   \begin{tabular}{l c c c c c c c c c c}
-   \textbf{Example} & $|\vc{q}|$ & \textbf{Material}&  $\rho$(kg/m$^3$) & \textbf{E}(Pa) & $\mu$ &\textbf{Sample Model}(s) & \textbf{Quadrature}(s)& \textbf{Weights}(s)& \textbf{Build $\Pi$}(s)& \textbf{Step}(s)\\
+   \textbf{Example} & $|\vc{q}|$ & \textbf{Material}&  $\rho$(kg/m$^3$) & \textbf{E}(Pa) & $\mu$ &\textbf{Sample}(s) & \textbf{Quadrature}(s)& \textbf{Weights}(s)& \textbf{Build $\Pi$}(s)& \textbf{Step}(s)\\
    \hline 
@@
    \textbf{Starship}                & ? & Steel & $8 \times 10^3$ & $2 \times 10^{11}$ & 0.32 & ? & ? & ? & ? & ? \\
-   \textbf{Tire}                    & ? & Rubber & $2 \times 10^3$ & $1 \times 10^6$ & 0.47 & ? & ? & ? & ? & ? & ?\\
+   \textbf{Tire}                    & ? & Rubber & $2 \times 10^3$ & $1 \times 10^6$ & 0.47 & ? & ? & ? & ? & ? \\
    \hline

### 2021-01-27 paper/sections/methods.tex
add 6.2 notes

--- 2021-01-27 paper/sections/methods.tex
@@
 % (6) add proofs/diagrams/figures that show we satisfy (2). }
-Each part, $\prt_i$, must be associated with a single deformation origin, $\bar{\refX}_j$. 
-Each origin is computed as the centroid of the set of parts with which it is associated. 
+Each part, $\prt_i$, must be associated with a single deformation origin, $\bar{\refX}_j$, and each origin is computed as the centroid of the set of parts with which it is associated. 
 While we constrain a part to be paired with only one deformation origin, an origin may be associated any number of parts. 
-Further, we require that each deformation origin be associated with a minimum of two parts, which avoids the case of an origin laying on the surface of a part.\ty{Should we mention periodic surfaces?
-You can have a periodic nurbs sphere consisting of a single surface, which we couldn't handle with this constraint (easy fix, but not sure if worth mentioning)}\dave{yes}
-Finally, to avoid unduly coupling the parts and reducing the expressivity of the simulation, we seek a sparse association of origins 
-which mirrors the pointwise sparsity of our blending weights,
+Further, we require that each deformation origin be associated with a minimum of two parts, which avoids the case of an origin laying on the surface of a part.
+One exception to this rule is periodic NURBS surfaces, where we allow for a deformation origin associated with only this part.
+Finally, to avoid unduly coupling the parts and limiting the expressivity of the simulation, we seek a sparse association of origins which mirrors the pointwise sparsity of our blending weights.
 
@@
 Each group now represents a unique part combination and the centroids of these groups become deformation origin candidates.
-To assign deformation origins to parts we use an integer network flow approach~\dave{cite}~(\reffig{origin_network}).
+To assign deformation origins to parts we use an integer network flow approach~\dave{cite}~(\reffig{network_flow}).
 
@@
 $E$ is the set of edges connecting nodes in $A$ to $B$ where each origin has an edge to every part with which it attempts to associate. 
-We augment $G$ to form $G'=(V' = A \cup B \cup \{s,t\}, E')$ where $V'$  is the new set of nodes now containing the source, $s$, and sink, $t$. 
+We augment $G$ to form $G'=(V' = A \cup B \cup \{s,t\}, E')$ where $V'$  is the new set of nodes now containing a source, $s$, and sink, $t$. 
 $E'$ is the augmented set of nodes connecting $s$ to each candidate in $A$ and connecting each part in $B$ to the sink $t$. 
@@
 and as a consequence each part may only be associated with a single deformation origin. 
-The last step is to specify the costs on the edges, $E$, between the origin candidates and the parts.
+The last step is to specify the costs on the edges, $E$, between the deformation origin candidates and the parts.
 
 Given a quadrature point group, and an edge from the associated deformation origin to $\prt_i$, we set the edge cost to be the sum, 
-over all quadrature points in the group, of the blending weights associated with $\prt_i$.
+over all quadrature points in the group of the blending weights associated with $\prt_i$.
 This costs penalizes heavily ``connected'' origin candidates, encouraging sparsity during origin selection. 
-We solve this problem with an integer constrained linear program\dave{cite solver, add equation with formulation} and choose $\XBs$ to be all the deformation origin candidates in $A$ 
-that have an outgoing flow of at least $1$.
+We solve this problem with the following integer constrained linear program
+\begin{equation}
+\begin{aligned}
+\mathbf{f} =  \arg\min_{\mathbf{f}} \quad & \mathbf{c}^T \mathbf{f}    \\
+\textrm{st} \quad & A_{\text{cons}} \mathbf{f} = 0  \\
+                  & f_i \in \integer \mbox{ }\forall i	\\
+                  & 0 \leq f_i \leq u_i
+\end{aligned}
+\end{equation}
+where $\mathbf{f}$ and $\mathbf{c}$ are the edge flow and cost values, respectively, assembled into vectors, and $A_{\text{cons}}$ forms the set of constraints satisfying the flow conservation constraint.
+\ty{we use the default in intlinprog, which should be dual-simplex} We choose $\XBs$ to be all the deformation origin candidates in $A$ that have an outgoing flow of at least $1$.
 

### 2021-01-27 paper/sections/methods.tex

@@
 While we constrain a part to be paired with only one deformation origin, an origin may be associated any number of parts. 
+%<<<<<<< HEAD
 Further, we require that each deformation origin be associated with a minimum of two parts, which avoids the case of an origin laying on the surface of a part.
@@
 Finally, to avoid unduly coupling the parts and limiting the expressivity of the simulation, we seek a sparse association of origins which mirrors the pointwise sparsity of our blending weights.
+%=======
+%Further, we require that each deformation origin be associated with a minimum of two parts, which avoids the case of an origin laying on the surface of a part.%\ty{Should we mention periodic surfaces?
+You can have a periodic nurbs sphere consisting of a single surface, which we couldn't handle with this constraint (easy fix, but not sure if worth mentioning)}\dave{yes}
+%Finally, to avoid unduly coupling the parts and reducing the expressivity of the simulation, we seek a sparse association of origins 
+%which mirrors the pointwise sparsity of our blending weights, 
+%>>>>>>> 198c892300fe3ce994edc867d772724e8c90c47f
 

### 2021-01-27 paper/sections/results.tex
updated timing table

--- 2021-01-27 paper/sections/results.tex
@@
   \caption{Performance of, and parameters for the Shape Matching Element Method on all examples. All timings are reported in seconds, physical  parameters are
-  reported with appropriate units. $\rho$ is the applied density, \textbf{E} is the  Young's Modulus and $\mu$ is the Poisson's ratio.
+  reported with appropriate units.
+$\rho$ is the applied density, \textbf{E} is the  Young's Modulus and $\mu$ is the Poisson's ratio.
+Here the time per step doesn't include the time for collision handling.
   }
@@
   \begin{tabular}{l c c c c c c c c c c}
-   \textbf{Example} $|\vc{q}|$ & \textbf{Material}&  $\rho$(kg/m$^3$) & \textbf{E}(GPa) & $\mu$ &\textbf{Sample Model}(s) & \textbf{Quadrature}(s)& \textbf{Weights}(s)& \textbf{Build $\Pi$}(s)& \textbf{Step}(s)\\
+   \textbf{Example} & $|\vc{q}|$ & \textbf{Material}&  $\rho$(kg/m$^3$) & \textbf{E}(Pa) & $\mu$ &\textbf{Sample Model}(s) & \textbf{Quadrature}(s)& \textbf{Weights}(s) & \textbf{Build $\Pi$}(s)& \textbf{Step}(s)\\
    \hline 
    \rowcolor[HTML]{DAE8FC} 
-   \textbf{Cantilever}      & ? & ? & ? & ? & ? & ? & ? & ? & ? & ? \\
-   \textbf{Coffee mug}      & ?& ? & ? & ? & ? & ? & ? & ? & ? & ?  \\
-   \textbf{Rocket}          & ?& ? & 648 & ? & ? & ? & ? & ? & ? & ?\\
-   \textbf{Castle}          & ?& ? & ? & ? & ? & ? & ? & ? & ? & ?\\
-   \textbf{Chicken}        & ? & ? & ? & ? & ? & ? & ? & ? & ? & ?\\
-   \textbf{Starship}       & ? & ? & ? & ? & ? & ? & ? & ? & ? & ?\\
-   \textbf{Tire}           & ? & ? & ? & ? & ? & ? & ? & ? & ? & ?\\
+   \textbf{Cantilever (soft)}       & 564 & Jelly & $1 \times 10^3$ & $1 \times 10^6$& 0.45 & $2.53 \times 10^{-2}$ & $4.45 \times 10^{-3}$ & $4.78 \times 10^{-1}$ & $1.86 \times 10^{-1}$ & $1.20 \times 10^{0}$ \\
+   \textbf{Cantilever (stiff)}      & 564 & Rubber & $1 \times 10^3$ & $5 \times 10^6$ & 0.45 & $2.48 \times 10^{-2}$ & $6.37 \times 10^{-3}$ & $5.10 \times 10^{-1}$ & $1.68 \times 10^{-1}$ & $1.19 \times 10^{0}$ \\
+   \textbf{Coffee mug}              & 1800 & Jelly & $1 \times 10^3$ & $1 \times 10^4$ & 0.47 & $4.18 \times 10^{-1}$ & $2.87 \times 10^{-2}$ & $3.77 \times 10^{0}$ & $2.08 \times 10^{0}$ & $1.19 \times 10^{1}$  \\
+   \textbf{Rocket (soft)}           & 648 & Rubber & $1.27 \times 10^3$ & $7 \times 10^6$ & 0.4 & $2.83 \times 10^{-2}$ & $5.62 \times 10^{-3}$ & $5.58 \times 10^{-1}$ & $6.83 \times 10^{-2}$ & $1.26 \times 10^{0}$ \\
+   \textbf{Rocket (stiff)}          & 648 & Steel & $8 \times 10^3$ & $2 \times 10^{11}$ & 0.32 & $2.56 \times 10^{-2}$ & $5.97 \times 10^{-3}$ & $4.59 \times 10^{-1}$ & $3.48 \times 10^{-2}$ & $9.99 \times 10^{-2}$ \\
+   \textbf{Castle}                  & 1872 & Jelly & $1.27 \times 10^3$ & $2 \times 10^3$ & 0.45 & $4.84 \times 10^{-1}$ & $1.44 \times 10^{-2}$ & $1.52 \times 10^{0}$ & $5.78 \times 10^{-1}$ & $8.02 \times 10^{0}$ \\
+   \textbf{Chicken}                 & 1773 & Rubber & $1.27 \times 10^3$ & $1 \times 10^4$ & 0.47 & $9.19 \times 10^{-1}$ & $2.31 \times 10^{-2}$ & $3.38 \times 10^{0}$ & $1.91 \times 10^{0}$ & $1.08 \times 10^{1}$ \\
+   \textbf{Starship}                & 750 & Steel & $8 \times 10^3$ & $2 \times 10^{11}$ & 0.32 & $6.78 \times 10^{-2}$ & $1.52 \times 10^{-2}$ & $8.50 \times 10^{-1}$ & $9.52 \times 10^{-2}$ & $6.95 \times 10^{-1}$ \\
+   \textbf{Astronauts}              & ? & ? & ? & ? & ? & ? & ? & ? & ? & ? \\
+   \textbf{Tire}                    & ? & Rubber/Steel & $2 \times 10^3$ / $3 \times 10^3$ & $1 \times 10^6$ / $7 \times 10^9$ & 0.47 / 0.35 & ? & ? & ? & ? & ? \\
    \hline

### 2021-01-27 paper/sections/methods.tex

@@
 While we constrain a part to be paired with only one deformation origin, an origin may be associated any number of parts.
-%<<<<<<< HEAD
-Further, we require that each deformation origin be associated with a minimum of two parts, which avoids the case of an origin laying on the surface of a part.
-One exception to this rule is periodic NURBS surfaces, where we allow for a deformation origin associated with only this part.
-Finally, to avoid unduly coupling the parts and limiting the expressivity of the simulation, we seek a sparse association of origins which mirrors the pointwise sparsity of our blending weights.
-%=======
-%Further, we require that each deformation origin be associated with a minimum of two parts, which avoids the case of an origin laying on the surface of a part.%\ty{Should we mention periodic surfaces?
-You can have a periodic nurbs sphere consisting of a single surface, which we couldn't handle with this constraint (easy fix, but not sure if worth mentioning)}\dave{yes}
-%Finally, to avoid unduly coupling the parts and reducing the expressivity of the simulation, we seek a sparse association of origins 
-%which mirrors the pointwise sparsity of our blending weights, 
-%>>>>>>> 198c892300fe3ce994edc867d772724e8c90c47f
 
@@
 Each group now represents a unique part combination and the centroids of these groups become deformation origin candidates.
-To assign deformation origins to parts we use an integer network flow approach~\dave{cite}~(\reffig{network_flow}).
+To assign deformation origins to parts we use an integer network flow approach~(\reffig{network_flow}).
 
@@
 Each edge $(i,j) \in E'$ has an associated flow $f_{ij}$ and capacity constraint $u_{ij}$, and each edge in $E$ also has a cost $c_{ij}$. 
-The principal constraint for this network flow problem is the conservation constraint, meaning that the net flow for all nodes in $V$ is zero.\dave{B is really overloaded}
+The principal constraint for this network flow problem is the conservation constraint, meaning that the net flow for all nodes in $V$ is zero.
 Each edge from $B$ to $t$ is assigned a capacity of $1$, which enforces that any part node in $B$ receives flow from a single edge, 
@@
 over all quadrature points in the group of the blending weights associated with $\prt_i$.
-This costs penalizes heavily ``connected'' origin candidates, encouraging sparsity during origin selection. 
-We solve this problem with the following integer constrained linear program
+This cost penalizes heavily ``connected'' origin candidates, encouraging sparsity during origin selection. 
+This amounts to minimizing the following integer linear program:
 \begin{equation}
@@
 where $\mathbf{f}$ and $\mathbf{c}$ are the edge flow and cost values, respectively, assembled into vectors, and $A_{\text{cons}}$ forms the set of constraints satisfying the flow conservation constraint.
-\ty{we use the default in intlinprog, which should be dual-simplex} We choose $\XBs$ to be all the deformation origin candidates in $A$ that have an outgoing flow of at least $1$.
+We solve this using the dual simplex solver in MATLAB.
+%\ty{we use the default in intlinprog, which should be dual-simplex} We choose $\XBs$ to be all the deformation origin candidates in $A$ that have an outgoing flow of at least $1$.
 

### 2021-01-27 paper/sections/methods.tex
12 pages

@@
 physics-based animation, specifically elastodynamics. 
-The advantage of having a unified kinematic description~(\refeq{kinematics}) is that we can directly apply Lagrangian Mechanics~\dave{cite Lanczos again} to arrive at 
+The advantage of having a unified kinematic description~(\refeq{kinematics}) is that we can directly apply Lagrangian Mechanics~\cite{lanczos2012variational} to arrive at 
 the pertinent equations of motion:
 \begin{equation*}
-    \MM\ddot{\vc{q}} = \fo\left(\vc{q}\right) + \fext,
-\end{equation*} where $\MM\in\real^{n\times n}$ is the mass matrix, $\vc{q}\in\real^{n}$ is the stacked vector of surface control points, $\fo\in\real^{n}$ is the 
-vector of internal forces and $\fext\in\real^{n}$ are external forces such due to gravity or contact. 
+    \MM\ddot{\vc{q}} = \fo\left(\vc{q}\right) + \vc{f}_g + \fext,
+\end{equation*} where $\MM\in\real^{n\times n}$ is the mass matrix, $\ddot{\vc{q}}\in\real^{n}$ are the generalized accelerations, $\vc{q}\in\real^{n}$ is the stacked vector of surface control points, $\fo\in\real^{n}$ is the
+vector of internal forces, $\vc{f}_g\in\real^{n}$ are body forces such as gravity and $\fext\in\real^{n}$ are external forces such as those due to contact. 
 We use penalty springs~\cite{10.1145/1964921.1964932} to resolve contact, and the approach of \citet{10.1145/566654.566623} for approximating frictional effects. 
@@
 \begin{equation}
-    T = \dot{\vc{q}}^T\underbrace{\Pi^T\int_{\Omega}\bldmat\left(\refX\right)^T\shapef\left(\refX\right)^T\shapef\left(\refX\right)\bldmat\left(\refX\right)d\Omega\Pi}_{\MM}\dot{\vc{q}},
+    T = \frac{1}{2}\dot{\vc{q}}^T\underbrace{\Pi^T\int_{\Omega}\bldmat\left(\refX\right)^T\shapef\left(\refX\right)^T\shapef\left(\refX\right)\bldmat\left(\refX\right)d\Omega\Pi}_{\MM}\dot{\vc{q}},
 \end{equation}
@@
 \subsection{External Forces}
-We compute the force of gravity and apply external forces in the standard way --
-$\vc{f}_g = \int_{\Omega} \rho\vc{g} d\Omega$, where $\vc{g}$ is the acceleration due to gravity acting on the object. 
+We compute the force of gravity and apply external forces in the standard way.
+Gravity is defined as $\vc{f}_g = \int_{\Omega} \rho\vc{g} d\Omega$, where $\vc{g}$ is the acceleration due to gravity acting on the object. 
 We leverage our meshless quadrature~(\refsec{quadrature}) for integration.
@@
 \begin{equation}
-\epsilon = \int_{\partial\Omega}||\Gamma\left(\refX\right)\vc{q} - \Jnurbs\left(\refX\right)\vc{q}||_2^2 d\refX \\
-\end{equation}, which is a quadratic.
+\epsilon = \int_{\partial\Omega}||\Gamma\left(\refX\right)\vc{q} - \Jnurbs\left(\refX\right)\vc{q}||_2^2 d\refX, \\
+\end{equation} which is a quadratic.
 

### 2021-01-27 paper/sections/results.tex
12 pages

--- 2021-01-27 paper/sections/results.tex
@@
   \begin{tabular}{l c c c c c c c c c c}
-   \textbf{Example} & $|\vc{q}|$ & \textbf{Material}&  $\rho$(kg/m$^3$) & \textbf{E}(Pa) & $\mu$ &\textbf{Sample}(s) & \textbf{Quadrature}(s)& \textbf{Weights}(s)& \textbf{Build $\Pi$}(s)& \textbf{Step}(s)\\
+   \textbf{Example} & $|\vc{q}|$ & \textbf{Material}&  $\rho$(kg/m$^3$) & \textbf{E}(Pa) & $\mu$ &\textbf{Sample}(s) & \textbf{Quad}(s)& \textbf{Weights}(s) & \textbf{Build $\Pi$}(s)& \textbf{Step}(s)\\
    \hline 
    \rowcolor[HTML]{DAE8FC} 
-   \textbf{Cantilever (soft)}       & ? & Jelly & $1 \times 10^3$ & $1 \times 10^6$& 0.45 & ? & ? & ? & ? & ? \\
-   \textbf{Cantilever (stiff)}      & ? & Rubber & $1 \times 10^3$ & $5 \times 10^6$ & 0.45 & ? & ? & ? & ? & ? \\
-   \textbf{Coffee mug}              & ? & Jelly & $1 \times 10^3$ & $1 \times 10^4$ & 0.47 & ? & ? & ? & ? & ?  \\
-   \textbf{Rocket (soft)}           & 648 & Rubber & $1.27 \times 10^3$ & $7 \times 10^6$ & 0.4 & ? & ? & ? & ? & ? \\
-   \textbf{Rocket (stiff)}          & 648 & Steel & $8 \times 10^3$ & $2 \times 10^{11}$ & 0.32 & ? & ? & ? & ? & ? \\
-   \textbf{Castle}                  & ? & Jelly & $1.27 \times 10^3$ & $2 \times 10^3$ & 0.45 & ? & ? & ? & ? & ? \\
-   \textbf{Chicken}                 & ? & Rubber & $1.27 \times 10^3$ & $1 \times 10^4$ & 0.47 & ? & ? & ? & ? & ? \\
-   \textbf{Starship}                & ? & Steel & $8 \times 10^3$ & $2 \times 10^{11}$ & 0.32 & ? & ? & ? & ? & ? \\
-   \textbf{Tire}                    & ? & Rubber & $2 \times 10^3$ & $1 \times 10^6$ & 0.47 & ? & ? & ? & ? & ? \\
+   \textbf{Cantilever (soft)}       & 564 & Jelly & $1 e^3$ & $1e10^6$& 0.45 & $2.53 e^{-2}$ & $4.45 e^{-3}$ & $4.78 e^{-1}$ & $1.86 e^{-1}$ & $1.20 e^{0}$ \\
+   \textbf{Cantilever (stiff)}      & 564 & Rubber & $1 e^3$ & $5 e^6$ & 0.45 & $2.48 e^{-2}$ & $6.37 e^{-3}$ & $5.10 e^{-1}$ & $1.68 e^{-1}$ & $1.19 e^{0}$ \\
+   \textbf{Coffee mug}              & 1800 & Jelly & $1 e^3$ & $1 e^4$ & 0.47 & $4.18 e^{-1}$ & $2.87 e^{-2}$ & $3.77 e^{0}$ & $2.08 e^{0}$ & $1.19 e^{1}$  \\
+   \textbf{Rocket (soft)}           & 648 & Rubber & $1.27 e^3$ & $7 e^6$ & 0.4 & $2.83 e^{-2}$ & $5.62 e^{-3}$ & $5.58 e^{-1}$ & $6.83 e^{-2}$ & $1.26 e^{0}$ \\
+   \textbf{Rocket (stiff)}          & 648 & Steel & $8 e^3$ & $2 e^{11}$ & 0.32 & $2.56 e^{-2}$ & $5.97 e^{-3}$ & $4.59 e^{-1}$ & $3.48 e^{-2}$ & $9.99 e^{-2}$ \\
+   \textbf{Castle}                  & 1872 & Jelly & $1.27 e^3$ & $2 e^3$ & 0.45 & $4.84 e^{-1}$ & $1.44 e^{-2}$ & $1.52 e^{0}$ & $5.78 e^{-1}$ & $8.02 e^{0}$ \\
+   \textbf{Chicken}                 & 1773 & Rubber & $1.27 e^3$ & $1 e^4$ & 0.47 & $9.19 e^{-1}$ & $2.31 e^{-2}$ & $3.38 e^{0}$ & $1.91 e^{0}$ & $1.08 e^{1}$ \\
+   \textbf{Starship}                & 750 & Steel & $8 e^3$ & $2 e^{11}$ & 0.32 & $6.78 e^{-2}$ & $1.52 e^{-2}$ & $8.50 e^{-1}$ & $9.52 e^{-2}$ & $6.95 e^{-1}$ \\
+   \textbf{Astronauts}              & ? & ? & ? & ? & ? & ? & ? & ? & ? & ? \\
+   \textbf{Tire}                    & ? & Rubber/Steel & $2 e^3$ / $3 e^3$ & $1 e^6$ / $7 e^9$ & 0.47 / 0.35 & ? & ? & ? & ? & ? \\
    \hline
@@
 %\dave{we should report which models we made ourselves}
-We implemented SEM in a combination of MATLAB and C++ using both GPToolbox~\cite{gptoolbox} and Bartels~\cite{bartels} for geometry processing and constitutive models 
+We implemented SEM using a combination of MATLAB and C++ using both GPToolbox~\cite{gptoolbox} and Bartels~\cite{bartels} for geometry processing and constitutive models 
 respectively. Benchmarks were performed using a MacBook Pro with an Intel i5 2.3GHz processor, 16GB of RAM and an Intel Iris Plus Graphics 655 GPU.
@@
 
-For raycasting operations and rendering we rely on Embree~\cite{10.1145/2601097.2601199} and triangulate the NURBS surfaces, which only represents an implementation simplification. 
-Our algorithm is not dependent on triangulating the boundary, and an ideal alternative would be using a path tracer.
+For raycasting operations and rendering we rely on Embree~\cite{10.1145/2601097.2601199} and Blender~\cite{blender}, and triangulate the NURBS surfaces, which only represents an implementation simplification.
+Our algorithm is not dependent on triangulating the boundary, and an ideal alternative would be using a path tracer that directly accepts NURBS geometry.
 
@@
 By virtue of its meshless nature, SEM is robust to a wide range of challenging models with large gaps and disconnected primitives.
-\reffig{fig:chicken} shows frames from a simulation of a rubber chicken. Note that the chicken model itself features large gaps between the individual NURBS parts. 
+\reffig{chicken} shows frames from a simulation of a rubber chicken. Note that the chicken model itself features large gaps between the individual NURBS parts. 
 Despite the lack of explicit connectivity, the SEM blending weights have the effect of implicitly enforcing connectivity at these seams.
@@
 
-\begin{figure}[H]
+\begin{figure}[h]
   \includegraphics[width=\columnwidth]{figures/mug_overlap}

### 2021-01-27 paper/sections/intro.tex
cite terzopolous

@@
 are used for computer-aided design (CAD), computational fabrication and computer animation. 
-NURBS primitives were the first geometric representation used for physics-based animation, yet,
+NURBS primitives were the first geometric representation used for physics-based animation~\cite{10.1145/37402.37427}, yet,
 despite over three decades of research, animation of curved models such as those built with NURBS, remains a challenge.

### 2021-01-27 matlab/nurbs_utils/nurbs_plot.m
added quadrature explanation

@@ -1,6 +1,8 @@
 function nurbs = nurbs_plot(nurbs)
     cm=jet(numel(nurbs));
     for ii = 1:numel(nurbs)
+        %         plt = patch('Faces',nurbs{ii}.hires_T,'Vertices',nurbs{ii}.hires_x0', ...
+
         plt = patch('Faces',nurbs{ii}.T,'Vertices',nurbs{ii}.x0', ...
             'FaceAlpha',0.5,'EdgeColor','none','FaceColor',cm(ii,:));
         hold on;

### 2021-01-27 matlab/nurbs_utils/nurbs_write_obj.m

@@ -10,10 +10,10 @@ function nurbs_write_obj(q, parts, path, ii)
         xi = squeeze(sum(parts{i}.hires_J .* qi,1));
 
         faces_i = parts{i}.hires_T + size(verts,1);
-        normals_i = nurbs_normals(parts{i}.srf.nurbs, parts{i}.hires_UV, parts{i}.p);
+%         normals_i = nurbs_normals(parts{i}.srf.nurbs, parts{i}.hires_UV, parts{i}.p);
         faces=[faces; faces_i];
         verts=[verts; xi'];
-        normals=[normals; normals_i];
+%         normals=[normals; normals_i];
         
         % obj_fn = "output/obj/part_" + i + "_" + int2str(ii) + ".obj";
         % writeOBJ(obj_fn, fvc.vertices, fvc.faces);
@@ -21,6 +21,6 @@ function nurbs_write_obj(q, parts, path, ii)
     end
     
     % you better have gptoolbox, son
-    writeOBJ(path, verts, faces, [], zeros(size(faces,1),3), normals, zeros(size(faces,1),3));
+    writeOBJ(path, verts, faces);%;, [], zeros(size(faces,1),3), normals, zeros(size(faces,1),3));
 end
 

### 2021-01-27 paper/sections/results.tex
finished the timing table
puft

@@
 Finally we show a few additional examples, highlighting the ability of SEM to handle complex simulations involving deformation, contact and friction.
-In \reffig{staypuft} we show two astronauts in jaunty space suits collide in zero gravity while in \reffig{starship} 
-we show an animation of the Space-V starship landing on Mars in a graceful way.
+In \reffig{staypuft} we show two astronauts in jaunty space suits collide in zero gravity. 
+SEM gracefully allows bounded discontinuities in simulation output to allow for rich motion. 
+This is similar to Discontinous Galerkin approaches but in SEM this behavior is emerges from the kinematic model, rather than from additional flux terms~\cite{kaufmann2009flexible}.
+
+In \reffig{starship} we show an animation of the Space-V starship landing on Mars in a graceful way.
 
 \begin{figure*}[htp]
-  \includegraphics[width=\textwidth,height=3in]{example-image-a}
-  \caption{Two astronauts, in jaunty spacesuits collide while spacewalking in zero-g.}
+  \includegraphics[width=\textwidth]{figures/astronauts.pdf}
+  \caption{Two astronauts, in jaunty spacesuits collide while spacewalking in zero-g.
+SEM correctly handles severely disconnected parts by correctly estimating elastic energy even in discontinuous regions.}
   \label{fig:staypuft}

### 2021-01-27 paper/sections/methods.tex

@@
 \dave{Ty: a bit more here, show the transformation of the integral via divergence theorem then show the final discrete version. Clearly define the quadrature rules.}
+The key idea behind raycasting quadrature is to transform a domain integral of the form $I=\int_{\Omega} f(X,Y,Z) d\Omega$ to an easier form that enables us to select quadrature points in an entirely meshless manner.
+The first transformation is to convert the domain of the integral to an \textit{auxiliary domain}, $\Omega_R$ on which integration is easy to compute.
+In our case we take this domain to be an axis-aligned bounding box around our model, and rewrite integration as $I=\int_{\Omega_R} g(X,Y,Z) d\Omega_R$ where $g$ evaluates to zero outside of $\Omega$, but equals $f$ inside $\Omega$.
+Next we apply the Divergence theorem \ty{the paper calls it Green-Gauss theorem, but these are same, right} to rewrite our integral as
+\newcommand{\Zeta}{\mathrm{Z}}
+\begin{equation}
+\int_{\Omega_R} g(X,Y,Z) d\Omega_R = \int_{\Gamma_R}\left(\int_{X_1}^{X_2} g(\alpha, Y, Z) d\alpha)\right)\mathbf{n} d\Gamma.
+\end{equation}
+where $\Gamma_R$ is the boundary of the bounding box with normals $\mathbf{n}$ and $X_1$ and $X_2$ represent the minimum and maximum values of the box along the $x$-axis.
+With this form, all we require to evaluate the outer integral is to integrate over the $y$-$z$ plane located at $X_1$.
+With this simple form, the normal term may be excluded as it always equals $1$ in the case of an axis-aligned bounding box.
+Finally, raycasting is used to evaluate the line integral along the x-axis.
+From here, we find all the intersection intervals where the ray is inside $\Omega$, and select quadrature points uniformly along this interval.
+Thus, this transformation from the original volume integral two and integral over a $y$-$z$ plane plus axis-aligned rays enables us to integrate the volumes of our NURBS models in a completely meshless approach.
 
-We make a minor modifiction this procedure to handle common geometric pathologies encountered in NURBS modelling, such as overlapping surfaces by performing 
+We make a minor modification this procedure to handle common geometric pathologies encountered in NURBS modelling, such as overlapping surfaces by performing 
 CSG unions while raycasting\dave{find a citation}. 

### 2021-01-27 paper/sections/methods.tex
updated simulation loop algo
papes

@@
 We make a minor modification this procedure to handle common geometric pathologies encountered in NURBS modelling, such as overlapping surfaces by performing 
-CSG unions while raycasting\dave{find a citation}. 
+CSG unions while raycasting~\cite{ROTH1982109}. 
 The output of this raycasting procedure is a set of quadrature points $\st{X}$ and weights $\st{v}$ that we use for integration and also as input to our 

### 2021-01-27 paper/sections/methods.tex
63 pass

@@
 In our case we take this domain to be an axis-aligned bounding box around our model, and rewrite integration as $I=\int_{\Omega_R} g(X,Y,Z) d\Omega_R$ where $g$ evaluates to zero outside of $\Omega$, but equals $f$ inside $\Omega$.
-Next we apply the Divergence theorem \ty{the paper calls it Green-Gauss theorem, but these are same, right} to rewrite our integral as
+Next we apply the Divergence theorem to rewrite our integral as
 \newcommand{\Zeta}{\mathrm{Z}}
@@
 With this form, all we require to evaluate the outer integral is to integrate over the $y$-$z$ plane located at $X_1$.
-With this simple form, the normal term may be excluded as it always equals $1$ in the case of an axis-aligned bounding box.
+With this simple form, the normal term may be excluded as it always equals $1$ \dave{isn;t it a vector $(1,0,0)$ or something like that ?
+}in the case of an axis-aligned bounding box.
 Finally, raycasting is used to evaluate the line integral along the x-axis.
@@
 \reffig{raycasting} shows the convergence of this quadrature approach when used to compute the volume of a NURBS model. 
+In principle we could apply detail-aware sampling methods~\cite{DBLP:journals/tog/WangFP12} to improve the robustness of raycasting quadrature even further,
+but we found it unnecessary for the examples in this paper.
 \begin{figure}
@@
 	\begin{algorithmic}[1]
-		\Procedure{Preprocess}{$model$}
-     	\myworries{todo: add volume}
+		\Procedure{Simulate}{$model$}
+     	 %\MM\ddot{\vc{q}} = \fo\left(\vc{q}\right) + \vc{f}_g + \fext,
      	\While {$simulating$}
-     		\State{$\Pcox \gets \Pi\mathbf{q}$}
+     		\State{$\Pcox \gets \Pi\vc{q}$}
      		\ForAll{$i \gets 1$ to $|\st{X}|$}
-     			\State{$\mathbf{F}_i \gets \frac{\partial \mathbf{F}}{\partial c} c$}
+     			\State{$\mathbf{F}_i \gets \frac{\partial \mathbf{F}}{\partial c} c \quad$ // deformation gradient}
 				\State{$\frac{\partial V}{\partial c} \gets \frac{\partial V}{\partial c} +
-				\frac{\partial \mathbf{F}}{\partial c}^T \frac{\partial \psi}{\partial \mathbf{F}}(\mathbf{F}_i)$}
+				\frac{\partial \mathbf{F}}{\partial c}^T \frac{\partial \psi}{\partial \mathbf{F}}(\mathbf{F}_i) \cdot \text{volume}_i$}
 
@@
 				 \frac{\partial \mathbf{F}}{\partial c}
-				 $}
+				 \cdot \text{volume}_i$}
   			\EndFor
   			\State{$K \gets \Pi^T \frac{\partial^2 V}{\partial c^2} \Pi$}
-  			\State{$f_{\text{internal}} \gets \Pi^T \frac{\partial V}{\partial c}$}
-  			\State{$f_{\text{error}} \gets \ME \vc{q}$}
-  			\State{$f_{\text{total}} \gets f_{\text{internal}} + f_{\text{error}} + f_{\text{ext}}$}
-  			\State{$\dot{\mathbf{q}} \gets \text{TimeIntegration}
-  			(f_{\text{total}},\mathbf{K}, \dot{\mathbf{q}}, \mathbf{J} , \mathbf{M}, \mathbf{M_E})$}
-  			\State{$\mathbf{q} \gets \mathbf{q} + h\dot{\mathbf{q}}$}
+  			\State{$\fo\left(\vc{q}\right) \gets \Pi^T \frac{\partial V}{\partial c} + \ME \vc{q}$}
+  			\State{$\vc{f}_{\text{total}} \gets \fo\left(\vc{q}\right) + \vc{f}_g + \fext$}
+  			\State{$\dot{\vc{q}} \gets \text{TimeIntegration}
+  			(K, \vc{f}_{\text{total}}, \dot{\vc{q}}, \samp{\Jnurbs} , \MM, \ME)$}
+  			\State{$\vc{q} \gets \vc{q} + h\dot{\vc{q}} \quad$ // $h$ is timestep size}
      	\EndWhile

### 2021-01-27 paper/sections/results.tex
update table comments

@@
 \begin{table*}[h]
-  \caption{Performance of, and parameters for the Shape Matching Element Method on all examples. All timings are reported in seconds, physical  parameters are
-  reported with appropriate units.
-$\rho$ is the applied density, \textbf{E} is the  Young's Modulus and $\mu$ is the Poisson's ratio.
-Here the time per step doesn't include the time for collision handling.
+  \caption{Performance of, and parameters for the Shape Matching Element Method on all examples. All wall-clock timings are reported in seconds, physical  parameters are
+  reported with appropriate units. $\rho$ is the applied density, \textbf{E} is the  Young's Modulus and $\mu$ is the Poisson's ratio.
+  \textbf{Sample} is the time taken to sample and construct $\samp{\Jnurbs}$ and $\samp{\Pmatx}$, \textbf{Quad} is the time taken to generate quadrature points via raycasting,
+  \textbf{Weights} is the time to compute blending weights, \textbf{Build $\Pi$} is the time to build the projection operator and  \textbf{Step} is the average time required to step the simulation (not including collision detection).
   }
@@
    \textbf{Castle}                  & 1872 & Jelly & $1.27 e^3$ & $2 e^3$ & 0.45 & $4.84 e^{-1}$ & $1.44 e^{-2}$ & $1.52 e^{0}$ & $5.78 e^{-1}$ & $8.02 e^{0}$ \\
-   \textbf{Chicken}                 & 1773 & Rubber & $1.27 e^3$ & $1 e^4$ & 0.47 & $9.19 e^{-1}$ & $2.31 e^{-2}$ & $3.38 e^{0}$ & $1.91 e^{0}$ & $1.08 e^{1}$ \\
+   \textbf{Chicken}                 & 1773 & Jelly & $1.27 e^3$ & $1 e^4$ & 0.47 & $9.19 e^{-1}$ & $2.31 e^{-2}$ & $3.38 e^{0}$ & $1.91 e^{0}$ & $1.08 e^{1}$ \\
    \textbf{Starship}                & 750 & Steel & $8 e^3$ & $2 e^{11}$ & 0.32 & $6.78 e^{-2}$ & $1.52 e^{-2}$ & $8.50 e^{-1}$ & $9.52 e^{-2}$ & $6.95 e^{-1}$ \\
-   \textbf{Astronauts}              & ? & ? & ? & ? & ? & ? & ? & ? & ? & ? \\
-   \textbf{Tire}                    & ? & Rubber/Steel & $2 e^3$ / $3 e^3$ & $1 e^6$ / $7 e^9$ & 0.47 / 0.35 & ? & ? & ? & ? & ? \\
+   \textbf{Astronaut (soft)}       & 3609 & Rubber &  $1.27 e^3$  & $1 e^6$  & 0.47 & $2.47 e^{0}$ & $6.71 e^{-2}$ & $1.13 e^{1}$ & $1.09 e^{1}$ & $1.41 e^{1}$ \\
+   \textbf{Tire}                    & 1404 & Rubber/Steel & $2 e^3$/$3 e^3$ & $1 e^6$/$7 e^9$ & 0.47 / 0.35 & $1.74 e^{-1}$ & $1.32 e^{-2}$ & $1.82 e^{0}$ & $6.79 e^{-1}$ & $1.28 e^{1}$ \\
    \hline

### 2021-01-27 paper/sections/results.tex
typo

@@
 \begin{figure}[h]
-  \includegraphics[width=\columnwidth]{example-image-a}
-  \caption{Here we show the output of an SEM animation loaded into Rhinoceros 3D 7 for additional editing.}
+  \includegraphics[width=\columnwidth]{figures/export_iges.pdf}
+  \caption{Here we show the output of an SEM animation loaded into Rhinoceros 3D 7 for additional editing.\ty{note: initial version of figure. Currently remaking it :)}}
   \label{fig:edit}
@@
 SEM gracefully allows bounded discontinuities in simulation output to allow for rich motion. 
-This is similar to Discontinous Galerkin approaches but in SEM this behavior is emerges from the kinematic model, rather than from additional flux terms~\cite{kaufmann2009flexible}.
+This is similar to Discontinous Galerkin approaches but in SEM this behavior emerges from the kinematic model, rather than from additional flux terms~\cite{kaufmann2009flexible}.
 

### 2021-01-27 paper/sections/results.tex
mentioned collision and fixed typo

@@
 Our algorithm is not dependent on triangulating the boundary, and an ideal alternative would be using a path tracer that directly accepts NURBS geometry.
+For collision detection and handling, we use the triangulation to detect inter-penetrations at the beginning of every iteration, and attempt to move the collided vertex out of its collision surface using a spring-like penalty force, similar to \cite{10.1145/3355089.3356486, 10.1145/2010324.1964932}.
 
@@
 SEM gracefully allows bounded discontinuities in simulation output to allow for rich motion. 
-This is similar to Discontinous Galerkin approaches but in SEM this behavior emerges from the kinematic model, rather than from additional flux terms~\cite{kaufmann2009flexible}.
+This is similar to Discontinous Galerkin approaches but in SEM this behavior is emerges from the kinematic model, rather than from additional flux terms~\cite{kaufmann2009flexible}.
 

### 2021-01-27 paper/sections/methods.tex
mentioned collision and fixed typo
smalledits

@@
     \begin{split}
-        \tr_i  & = \sum_{\prt_k \in \bar{\refX}_j} \underbrace{\mathbbm{1}^T\wt_k \samp{\Jnurbs}_k}_{B_{ik}}\vc{q}_k \\
+\tr_i  & = \frac{1}{\sum_{\prt_k \in \bar{\refX}_j} \Tr \wt_k}
+        \sum_{\prt_k \in \bar{\refX}_j} \underbrace{\mathbbm{1}^T\wt_k \samp{\Jnurbs}_k}_{B_{ik}}\vc{q}_k \\   
         \vc{c}_i^1 & = \smash{\underbrace{\left(\samp{\Pmat^1}^T\wt_i\samp{\Pmat^1}\right)}_{A^1_i}}^{-1}\bigg(\underbrace{\samp{\Pmat^1}^T\wt_i\samp{\Jnurbs}_i}_{B^1_i}\vc{q}_i-\underbrace{\samp{\Pmat^1}^T\wt_i\mathbbm{1}}_{T^1_i}\tr_i\bigg)\\
@@
 \mathbf{w}(\mathbf{X}_j) =  \arg\min_{\mathbf{w}} \quad & \mathbf{w}^T \Theta(\mathbf{X}_j) \mathbf{w}    \\
-\textrm{st} \quad & 0 \leq w_i \leq 1 \mbox{ }\forall i                    \\
+\textrm{s.t.} \quad & 0 \leq w_i \leq 1 \mbox{ }\forall i                    \\
                   &   \sum_i^n w_i = 1                      \\
@@
 Finally, the distance cutoff $\alpha$ allows us to control the sparsity of the blending weights.
-We employ a simple heuristic to select $\alpha$ by searching for a $\alpha$ that produces a set of blending weights meeting a desired sparsity target $T \in \left( 0,1\right)$, which is defined as the ratio of non-zeros to total number of weights.
+We employ a simple heuristic to select $\alpha$ by searching for one that produces a set of blending weights meeting a desired sparsity target $T \in \left( 0,1\right)$, which is defined as the ratio of non-zeros to total number of weights.
 This yields a unitless, easy to tune parameter.
@@
 \mathbf{f} =  \arg\min_{\mathbf{f}} \quad & \mathbf{c}^T \mathbf{f}    \\
-\textrm{st} \quad & A_{\text{cons}} \mathbf{f} = 0  \\
-                  & f_i \in \integer \mbox{ }\forall i	\\
+\textrm{s.t.} \quad & A_{\text{cons}} \mathbf{f} = 0  \\
+                  & f_i \in \integer \mbox{ }, \forall i	\\
                   & 0 \leq f_i \leq u_i

### 2021-01-27 paper/sections/methods.tex
done

@@
     \begin{split}
-\tr_i  & = \frac{1}{\sum_{\prt_k \in \bar{\refX}_j} \Tr \wt_k}
-        \sum_{\prt_k \in \bar{\refX}_j} \underbrace{\mathbbm{1}^T\wt_k \samp{\Jnurbs}_k}_{B_{ik}}\vc{q}_k \\   
+\tr_i  & = \sum_{\prt_k \in \bar{\refX}_j} \underbrace{\frac{1}{\sigma_k}\mathbbm{1}^T\wt_k \samp{\Jnurbs}_k}_{B_{ik}}\vc{q}_k \\   
         \vc{c}_i^1 & = \smash{\underbrace{\left(\samp{\Pmat^1}^T\wt_i\samp{\Pmat^1}\right)}_{A^1_i}}^{-1}\bigg(\underbrace{\samp{\Pmat^1}^T\wt_i\samp{\Jnurbs}_i}_{B^1_i}\vc{q}_i-\underbrace{\samp{\Pmat^1}^T\wt_i\mathbbm{1}}_{T^1_i}\tr_i\bigg)\\
@@
     \label{eq:hfit} 
-\end{equation} where $\Pmat^l$ contains only the monomial basis of order $l$,  $\samp{\Pmat^l}$ is the stacked evaluation of this matrix at the $s$ quadrature points and $\vc{c}^l_i$ are the corresponding coefficients.
+\end{equation} where $\sigma_k=\sum_{\prt_k \in \bar{\refX}_j} \Tr \wt_k$ is the normalization constant for the $k^{th}$ part, $\Pmat^l$ contains only the monomial basis of order $l$,  $\samp{\Pmat^l}$ is the stacked evaluation of this matrix at the $s$ quadrature points and $\vc{c}^l_i$ are the corresponding coefficients.
 Finally, $\wt_i$ is the diagonal integration weight matrix and $\mathbbm{1}\in\real^{3s\times3}$ is a matrix of stacked $3\times3$ identity matrices -- one for each quadrature point.

### 2021-01-28 paper/sections/intro.tex
first pass over intro

@@
 For volumetric objects, this process often involves identifying and discretizing the interior of the modelled object, 
-typically either as a tetrahedral or hexahedral mesh.%
+typically either as a tetrahedral or hexahedral mesh. 
 This procedure is both expensive and difficult, especially if the surface model is constructed from higher-order boundary
@@
 Isogeometric Analysis (IGA) is a physics simulation methodology that uses the control variables of the NURBS model
-as the degrees-of-freedom (DOFS) of the simulation itself. 
+as the degrees-of-freedom (DOFs) of the simulation itself. 
 Unfortunately IGA approaches for volumetric objects still require background volumetric structures, typically
@@
 
-We present the first truly meshless (no volumetric discretization generated) algorithm for direct, nonlinear elastodynamic simulation of NURBs models. 
+We present the first truly meshless (no volumetric discretization required) algorithm for direct, nonlinear elastodynamic simulation of NURBs models. 
 Our nonlinear elastodynamic simulation scheme requires only a boundary description of the object (we do not strictly require a solid model) and
@@
 Dirichlet and Neumann boundary conditions and to apply contact resolution. 
-Crucially because we broadly target animation not necessarily simulation for engineering or manufacturing, we don't require that models
-satisfy the rigourous geometric requirements common for these applications
+Crucially, because we broadly target animation not necessarily simulation for engineering or manufacturing, we don't require that models
+satisfy the rigourous geometric requirements common for these applications.
 
 Our approach draws inspiration from the recently developed Virtual Element Method (VEM) for solving partial differential equations on domains tiled with arbitrary polygons.
-We establish a connection between VEM and the well-known shape matching simulation algorithm. This allows us 
+We establish a connection between VEM and the well-known shape matching simulation algorithm~\cite{10.1145/1073204.1073216,10.1145/2019406.2019438}. This allows us 
 to derive equations of motion for an arbitrary curved surface model, made of NURBS, via Lagrangian Mechanics. 

### 2021-01-28 paper/sections/related.tex
fixed typos on related works

@@
 
-For many physically-based animation tasks, it would be desirable to bypass volumetric meshing entirely and directly simulate the geometric model.
-An ideal approach would avoid meshing of any kind (no volumetric meshes or cutcells), support continuum mechanics type constitutive models 
+For many physics-based animation tasks, it would be desirable to bypass volumetric meshing entirely and directly simulate the geometric model.
+An ideal approach would avoid meshing of any kind (no volumetric meshes or cut-cells), support continuum mechanics-type constitutive models 
 and energies that have become standard in physics-based animation pipelines, be compatible with a wide range of time integration schemes and
@@
 \emph{Isogeometric Analysis}~\cite{10.1002/9780470749081.ch7} endeavors to perform simulation directly on the NURBS output from 
-Computer-Aided Design (CAD) software. Initial attempts used NURBS surfaces to represent the mid-surface of thin objects~\cite{10.1145/176579.176580}.
+Computer-Aided Design (CAD) software. Initial attempts used NURBS surfaces to represent the medial surface of thin objects~\cite{10.1145/176579.176580}.
 Volumetric simulations relied on volumetric NURBS~\cite{10.1007/978-3-642-03596-8_2} but were limited to a narrow class of geometries.
@@
 Modern approaches are constructed around the extended finite element method which enriches the standard finite element basis with discontinuous basis functions
-to enable improved boundary handling~\cite{https://doi.org/10.1002/nme.3120,10.1007/s00466-013-0854-7,https://doi.org/10.1002/nme.4852,SAFDARI2016373,10.1145/3355089.3356576}. 
+to improve boundary handling~\cite{https://doi.org/10.1002/nme.3120,10.1007/s00466-013-0854-7,https://doi.org/10.1002/nme.4852,SAFDARI2016373,10.1145/3355089.3356576}. 
 These methods typically start with an easy-to-generate structured volumetric mesh (tetrahedral or hexahedral), ``cutting'' the NURBS geometric model against
-it to enable boundary handling (such a mesh is called a \emph{cutcell} mesh). 
+it to enable boundary handling (such a mesh is called a \emph{cut-cell} mesh). 
 Like volumetric meshing, this cutting operation can be difficult and our ideal method would avoid it. 
@@
 intersections and disconnected primitives without additional user input. 
-These features mean that our algorithm is the first truly meshless approach for the direct simulation of curved surface, NURBS models for physics-based animation.
+These features mean that our algorithm is the first truly meshless approach for the direct simulation of curved surface, NURBS models in physics-based animation.
 

### 2021-01-28 paper/sections/methods.tex
fix typos until equ 7 in the method section

@@
 The input to the Shape Matching Element Method (SEM)  is a NURBS \emph{boundary representation} of a volumetric object, along with a set of physical parameters 
-(density, constitutive model, model parameters etc). 
+(density, constitutive model, model parameters, etc). 
 The boundary representation is composed of one or more (not necessarily explicitly connected) NURBS surface primitives that we will call \emph{parts}.
@@
 In the \emph{preprocessing} stage we use raycasting to find quadrature points and weights for volumetric integration, as well as to compute part blending parameters. 
-Finally we construct local shape matching operators and construct the mass matrix for our problem.
+Finally we construct local shape matching operators and the mass matrix for our problem.
 
@@
 		\State $\vc{q} \gets \text{GetControlPoints}(model)$
-		\State $\samp{\Jnurbs}, \wt \gets \text{SampleNURBS}(model)$
+		\State $\samp{\Jnurbs}, \wt \gets \text{SampleNURBS}(model)$  \Comment{Section \ref{sec:shapematching}}
 		\State $\vc{x^0} \gets \wt \samp{\Jnurbs} \vc{q}$
@@
 \begin{equation*}
-    \phi_{ij}\left(u,v\right) = \frac{\omega_{i,j}\beta_{ik}(u)\beta_{jl}(v)}{\sum_{r=1}^{n}\sum_{s=1}^{m} \omega_{ij} \beta_{ik}(u)\beta_{jl}(v)},
-\end{equation*} where $\beta_{ik}$ (resp. $\beta_{jl}$) is the $i^{th}$ ($j^{th}$) B-spline basis of order $k$ ($l$).
+    \phi_{ij}\left(u,v\right) = \frac{\omega_{i,j}\beta_{ik}(u)\beta_{jl}(v)}{\sum_{r=1}^{n}\sum_{s=1}^{m} \omega_{i,j} \beta_{ik}(u)\beta_{jl}(v)},
+\end{equation*} where $\beta_{ik}$ (resp. $\beta_{jl}$) is the $i^{th}$ ($j^{th}$) B-spline basis of order $k$ ($l$) and $\omega_{i,j}$ is the NURBS weight.
 
@@
     \label{eq:projection}
-\end{equation} where $\mathbf{p}^T\left(\refX\right)$ is the polynomial basis vector \newline (eg $\begin{bmatrix} X & Y & Z & X\cdot Y & X\cdot Z & Y\cdot Z& X\cdot X & Y\cdot Y & Z\cdot Z \end{bmatrix}$ ),
+\end{equation} where $\mathbf{p}^T\left(\refX\right)$ is the polynomial basis vector \newline (e.g., $\begin{bmatrix} X & Y & Z & X\cdot Y & X\cdot Z & Y\cdot Z& X\cdot X & Y\cdot Y & Z\cdot Z \end{bmatrix}$ ),
 $\bar{\refX}$ is an a priori chosen origin around which to deform, $\Pco$ are the coefficients for the 
@@
 \end{equation} Here $\samp{\Pmat}$ (resp. $\samp{\Jnurbs}$) is a $3s \times 3(k+1)$ (resp. $3s \times 3nm$) matrix where $s$ is the number of quadrature points, 
-$k$ is the number of  terms in the shape matching polynomial and $3nm$ is the number of NURBS control points. 
+$k$ is the number of  terms in the shape matching polynomial and $nm$ is the number of NURBS control points. 
 $\samp{\Pmat}$ (resp. $\samp{\Jnurbs}$) is constructed by stacking the evaluations of $\Pmatxuv$ (resp. $\Jnurbs\left(u, v\right)$) at each
@@
 \paragraph*{Relationship to the Virtual Element Method}
-For a non-planar NURBS part, and given a sufficient number of quadrature points, ~\refeq{shapematch} will be well-posed allowing us to write: $\Pcox^* = \Pi\vc{q}$ where
+For a non-planar NURBS part, a sufficient number of quadrature points will make ~\refeq{shapematch} will be well-posed, allowing us to write: $\Pcox^* = \Pi\vc{q}$ where
 \begin{equation*}
@@
 Previous work~\cite{10.1145/1073204.1073216,10.1145/1275808.1276480} used additional lattices or partitions of the reference space to 
-increase the number of polynomials in the shape matching. In contrast, our meshless Shape Matching Element Method which works directly on the NURBS
+increase the number of polynomials in the shape matching. In contrast, our meshless Shape Matching Element Method works directly on the NURBS
 geometry itself.
@@
         \vc{c}_i^1 & = \smash{\underbrace{\left(\samp{\Pmat^1}^T\wt_i\samp{\Pmat^1}\right)}_{A^1_i}}^{-1}\bigg(\underbrace{\samp{\Pmat^1}^T\wt_i\samp{\Jnurbs}_i}_{B^1_i}\vc{q}_i-\underbrace{\samp{\Pmat^1}^T\wt_i\mathbbm{1}}_{T^1_i}\tr_i\bigg)\\
-        \vc{c}_i^2 & = \smash{\underbrace{\left(\samp{\Pmat^2}^T\wt_i\samp{\Pmat^2}\right)}_{A^2_i}}^{-1}\bigg(\underbrace{\samp{\Pmat^2}^T\wt_i\samp{\Jnurbs}_i}_{B^2_i}\vc{q}_i-\underbrace{\samp{\Pmat^2}^T\wt_i\samp{\Pmat^1}}_{D^2_i}\vc{c}_i^1 - \underbrace{\samp{\Pmat^2}^T\wt_i\mathbbm{1}}_{T^2_i}\tr_i\bigg)
-    \end{split},
+        \vc{c}_i^2 & = \smash{\underbrace{\left(\samp{\Pmat^2}^T\wt_i\samp{\Pmat^2}\right)}_{A^2_i}}^{-1}\bigg(\underbrace{\samp{\Pmat^2}^T\wt_i\samp{\Jnurbs}_i}_{B^2_i}\vc{q}_i-\underbrace{\samp{\Pmat^2}^T\wt_i\samp{\Pmat^1}}_{D^2_i}\vc{c}_i^1 - \underbrace{\samp{\Pmat^2}^T\wt_i\mathbbm{1}}_{T^2_i}\tr_i\bigg),
+    \end{split}
     \label{eq:hfit} 

### 2021-01-28 paper/main.tex

@@
 
+\acmSubmissionID{230}
 \newcommand\myworries[1]{\textcolor{red}{#1}}
@@
 \begin{teaserfigure}
-  \includegraphics[width=\textwidth]{castle.png}
-  \caption{tmp}
-  \Description{}
+  \includegraphics[width=\textwidth]{figures/teaser_castle.pdf}
+  \caption{Using the shape matching element method we can directly simulate this NURBS surface model of a bouncy castle as a volumetric elastic object without the need for volumetric meshing of any kind.}
   \label{fig:teaser}
@@
 \begin{abstract}
-We introduce a new method for direct physics-based animation of volumetric curved models.  
-The Shape Matching Element Method is a completely meshless approach which is robust to gaps and overlaps in geometry, 
-is compatible with standard constitutive models and time  integration schemes, supports contact and frictional interactions 
-as well as direct editing of simulation output. 
+We introduce a new method for direct physics-based animation of volumetric curved models, represented using NURBS surfaces.  
+Our technical contribution is the Shape Matching Element Method (SEM).
+SEM is a completely meshless algorithm, the first to simultaneously 
+be robust to gaps and overlaps in geometry, 
+be compatible with standard constitutive models and time integration schemes, support contact and frictional interactions  
+and to preserve feature correspondence during simulation which enables editable simulated output. 
 We demonstrate the efficacy of our algorithm by producing compelling physics-based animations from a variety of curved input models.

### 2021-01-28 paper/sections/methods.tex
first pass over the method

@@
 Specifying weights manually would be antithetical to our goal of providing a seamless translation from surface representation to physics-based animation.
-Automatic  weight computation is well-studied (see ~\citet{10.1145/2766952,BBW:2011} and ~\citet{10.1145/2010324.1964968}) but typically requires solving 
+Automatic  weight computation is well-studied (see ~\citet{10.1145/2766952}, ~\citet{BBW:2011} and ~\citet{10.1145/2010324.1964968}), but typically requires solving 
 an optimization problem that itself relies on a volumetric discretization. 
@@
 While partition of unity and sparsity are straightforward to enforce, the overlapping and intersecting parts~(\reffig{badmodels}) often encountered in non-engineering models leads
-to a contradiction -- how can the blending functions be both one and zero at the points of overlap ? 
+to a contradiction --- how can the blending functions be both one and zero at the points of overlap? 
 Our solution is to allow weights to include bounded discontinuities at these points and it is these discontinuities that are difficult to enforce 
@@
 
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=3.33in]{figures/plot_distance_weight.pdf}
+    \caption{Distance weight as a function of primary and total ray length with cutoff distance 50.}
+    % \caption{We plot the distance weight with respect to the primary and total ray length. Here the cutoff distance is 50.
+    % }
+    \label{fig:plot_distance_weight}
+    \vspace{-5pt}
+\end{figure}
+
 Using these two distances we define a linear \emph{distance weight} using the well-known ReLU function:
 \begin{equation}
-\theta_i = \max (1.0 - \frac{d_{\text{primary}}}{\min (d_{\text{total}}, \alpha)}, 0.0)
+\theta_i = \max \left(1.0 - \frac{d_{\text{primary}}}{\min (d_{\text{total}}, \alpha)}, 0.0\right)
 \text{,}
@@
 For points inside closed periodic surfaces, it is critical that the weights equal one in this case, so we set $\theta_i = 1$.
-
-% \begin{figure}[h]
-%     \centering
-%     \includegraphics[width=3.33in]{figures/plot_distance_weight.pdf}
-%     \caption{We plot the distance weight with respect to the primary and total ray length. Here the cutoff distance is 50.
-%     }
-%     \label{fig:plot_distance_weight}
-%     \vspace{-5pt}
-%   \end{figure}
 
@@
 
-\begin{wrapfigure}[11]{l}{0.31\columnwidth}
-    \begin{center}
-            %\vspace{-0.7cm}
-            %\hspace{-0.1cm}
-            \includegraphics[width=0.3\columnwidth]{figures/plot_distance_weight.pdf}
+% \begin{wrapfigure}[11]{l}{0.31\columnwidth}
+%     \begin{center}
+%             %\vspace{-0.7cm}
+%             %\hspace{-0.1cm}
+%             \includegraphics[width=0.3\columnwidth]{figures/plot_distance_weight.pdf}
 
-    \end{center}
-    \caption{ \label{fig:plot_distance_weight} Distance weight as a function of primary and total ray length with cutoff distance 50.}
-  \end{wrapfigure}
+%     \end{center}
+%     \caption{ \label{fig:plot_distance_weight} Distance weight as a function of primary and total ray length with cutoff distance 50.}
+%   \end{wrapfigure}
 
@@
 but allows us to use more general weighting functions and apply the appropriate behavior for intersecting parts.
-The price we pay is that our weights are not C0 continuous but rather Lipschitz smooth.
+The price we pay is that our weights are not $C^0$ continuous but rather Lipschitz smooth.
 Perturbations in the boundaries of parts can cause bounded jumps in our distance weights. 
@@
 From here, we find all the intersection intervals where the ray is inside $\Omega$, and select quadrature points uniformly along this interval.
-Thus, this transformation from the original volume integral two and integral over a $y$-$z$ plane plus axis-aligned rays enables us to integrate the volumes of our NURBS models in a completely meshless approach.
+Thus, this transformation from the original volume integral into an integral over a $y$-$z$ plane plus axis-aligned rays enables us to integrate the volumes of our NURBS models in a completely meshless approach.
 
-We make a minor modification this procedure to handle common geometric pathologies encountered in NURBS modelling, such as overlapping surfaces by performing 
+We make a minor modification to this procedure to handle common geometric pathologies encountered in NURBS modelling, such as overlapping surfaces, by performing 
 CSG unions while raycasting~\cite{ROTH1982109}. 
 The output of this raycasting procedure is a set of quadrature points $\st{X}$ and weights $\st{v}$ that we use for integration and also as input to our 
-deformation origin extraction~\refsec{origins} and blending weights computation~\refsec{weights}
+deformation origin extraction (~\refsec{origins}) and blending weights computation (~\refsec{weights})
 \reffig{raycasting} shows the convergence of this quadrature approach when used to compute the volume of a NURBS model. 
@@
     \includegraphics[width=\columnwidth]{figures/raycasting_quadrature}
-    \caption{Figure showing how raycasting quadrature works (top), convergence plot for volume integration (bottom}
+    \caption{When integrating physical quantities over the volumetric domain of an object, our raycasting quadrature is robust and converges to the groundtruth result as we increase the number of rays in use.
+}
+    % \caption{Figure showing how raycasting quadrature works (top), convergence plot for volume integration (bottom}
     \label{fig:raycasting}
@@
     \MM\ddot{\vc{q}} = \fo\left(\vc{q}\right) + \vc{f}_g + \fext,
-\end{equation*} where $\MM\in\real^{n\times n}$ is the mass matrix, $\ddot{\vc{q}}\in\real^{n}$ are the generalized accelerations, $\vc{q}\in\real^{n}$ is the stacked vector of surface control points, $\fo\in\real^{n}$ is the
+\end{equation*} where $\MM\in\real^{n\times n}$ is the mass matrix, $\vc{q}\in\real^{n}$ is the stacked vector of surface control points, $\ddot{\vc{q}}\in\real^{n}$ is the generalized accelerations, $\fo\in\real^{n}$ is the
 vector of internal forces, $\vc{f}_g\in\real^{n}$ are body forces such as gravity and $\fext\in\real^{n}$ are external forces such as those due to contact. 
@@
 \begin{equation}
-    \defX\left(\refX\right) = \shapef\left(\refX\right)\bldmat\Pi\vc{q},
+    \defX\left(\refX\right) = \underbrace{\shapef\left(\refX\right)\bldmat\Pi}_{\Jkin\left(\refX\right)}\vc{q},
 \end{equation} where $\shapef\in\real^{3\times n} = \begin{bmatrix}P_1\left(\refX\right) \dots P_m\left(\refX\right)\end{bmatrix}$, $m$ is number of 
-parts in the object, $\bldmat$ is the diagonal matrix of blending weights and $\Pi$ is the global projection operator.
+parts in the object, $\bldmat$ is the diagonal matrix of blending weights, $\Pi$ is the global projection operator and $\Jkin\left(\refX\right)$ is the kinematic Jacobian.
 
@@
 \begin{equation}
-        \vc{v}\left(\refX\right) = \underbrace{\shapef\left(\refX\right)\bldmat\Pi}_{\Jkin}\dot{\vc{q}},
+        \vc{v}\left(\refX\right) = \Jkin\left(\refX\right)\dot{\vc{q}},
     \end{equation} where $\Jkin\left(\refX\right)$ is the kinematic Jacobian and $\dot{\vc{q}}$ are the generalized velocities of the system.
@@
 In this submission we focus on hyperelastic models~\cite{10.1145/2343483.2343501} wherein the potential energy stored by the object, per unit volume,
-is given by the strain energy density function $\psi\left(\defo\right)$, and its internal forces and stiffness are the negative gradient and hessian respectively. 
+is given by the strain energy density function $\psi\left(\defo\right)$, and its internal forces and stiffness are the negative gradient and Hessian respectively. 
 Evaluating and applying constitutive models of this form requires the ability to (1) evaluate the deformation gradient over the reference domain of the model and 

### 2021-01-28 paper/sections/methods.tex
update distance weight plot

@@
     \centering
-    \includegraphics[width=3.33in]{figures/plot_distance_weight.pdf}
-    \caption{Distance weight as a function of primary and total ray length with cutoff distance 50.}
+    \includegraphics[width=3.33in]{figures/plot_distance_weight_both.pdf}
+    \caption{Distance weight as a function of primary and total ray length with cutoff distance 50 (left). Here we use different colors to denote different parts (right).}
     % \caption{We plot the distance weight with respect to the primary and total ray length. Here the cutoff distance is 50.
@@
 \end{equation}
-where $\alpha$ is a distance cutoff parameter. 
+where $\alpha$ is a distance cutoff parameter, $d_{\text{primary}}$ is the primary ray length and $d_{\text{total}}$ is the total ray length. 
 This distance weight~(\reffig{plot_distance_weight}) is the result of interpolating along the ray from $\clst$ to $\rhit$ (or a fixed point that depends on $\alpha$). 

### 2021-01-28 paper/sections/conclusion.tex
first pass on results and conclusion

@@
 Their approach could be used to remove this parameter from SEM entirely.
-Finally, in this paper we have focused on applying SEM to physics-based animation, there is a significant amount of additional work needed to 
+Finally, in this paper we have focused on applying SEM to physics-based animation, and there is a significant amount of additional work needed to 
 extend SEM reliably into engineering applications.
 
-The Finite Element method took over 40 years to mature to its current state and to become the preeminent tool in physically-based animation.
+The Finite Element method took over 40 years to mature to its current state and to become the preeminent tool in physics-based animation.
 We hope that this is the beginning of a similar, exciting journey for SEM. 

### 2021-01-28 paper/sections/results.tex
mentioned collision and fixed typo
first pass on results and conclusion

--- 2021-01-27 paper/sections/results.tex
@@
    \rowcolor[HTML]{DAE8FC} 
-   \textbf{Cantilever (soft)}       & 564 & Jelly & $1 e^3$ & $1e10^6$& 0.45 & $2.53 e^{-2}$ & $4.45 e^{-3}$ & $4.78 e^{-1}$ & $1.86 e^{-1}$ & $1.20 e^{0}$ \\
+   \textbf{Cantilever (soft)}       & 564 & Jelly & $1 e^3$ & $1e^6$& 0.45 & $2.53 e^{-2}$ & $4.45 e^{-3}$ & $4.78 e^{-1}$ & $1.86 e^{-1}$ & $1.20 e^{0}$ \\
    \textbf{Cantilever (stiff)}      & 564 & Rubber & $1 e^3$ & $5 e^6$ & 0.45 & $2.48 e^{-2}$ & $6.37 e^{-3}$ & $5.10 e^{-1}$ & $1.68 e^{-1}$ & $1.19 e^{0}$ \\
@@
 For raycasting operations and rendering we rely on Embree~\cite{10.1145/2601097.2601199} and Blender~\cite{blender}, and triangulate the NURBS surfaces, which only represents an implementation simplification.
-Our algorithm is not dependent on triangulating the boundary, and an ideal alternative would be using a path tracer that directly accepts NURBS geometry.
+Our core algorithm is not dependent on triangulating the boundary, and an ideal alternative would be using a path tracer that directly accepts NURBS geometry.
+For collision detection and handling, for simplicity we use the triangulation to detect inter-penetrations at each iteration, and attempt to move the collided vertex out of its collision surface using a spring-like penalty force, similar to \cite{10.1145/3355089.3356486, 10.1145/2010324.1964932}.
+An ideal alternative would be adopting the implicit representation of the surfaces for collision detection and resolution \cite{10.1145/1516522.1516523, 10.1145/3306346.3323010}.
 
@@
   \includegraphics[width=\textwidth,height=3in]{figures/starship.pdf}
-  \caption{The Space-V starship makes a graceful landing on the red planet. }
+  \caption{Our SEM is robust to stiff materials, complicated geometry and non-trivial collisions. Here we show the Space-V starship makes a graceful landing on the rugged surface of Mars. }
   \label{fig:starship}

### 2021-01-28 paper/sections/intro.tex
fixed typo in the intro

@@
 Dirichlet and Neumann boundary conditions and to apply contact resolution. 
-Crucially, because we broadly target animation not necessarily simulation for engineering or manufacturing, we don't require that models
+Crucially, because we broadly target animation and not necessarily simulation for engineering or manufacturing, we don't require that models
 satisfy the rigourous geometric requirements common for these applications.

### 2021-01-28 paper/sections/methods.tex
update distance weight plot caption
fixed typo in equation

@@
 When assembled for all parts,~\refeq{hfit}, yields a block upper triangular matrix $U$ and a sparse matrix $B$ ~(\reffig{multiparts}) which can be used to efficiently to compute the projection 
-operator for the entire object $\Pi = U^{-1}B\vc{q}$. 
+operator for the entire object $\Pi = U^{-1}B$. 
 The structure of $U$ and $B$ ensures that $\Pi$ only couples objects which share deformation centers, which implies a sparse $\Pi$. 
@@
     \includegraphics[width=3.33in]{figures/plot_distance_weight_both.pdf}
-    \caption{Distance weight as a function of primary and total ray length with cutoff distance 50 (left). Here we use different colors to denote different parts (right).}
+    \caption{Distance weight as a function of primary ray length $d_{\text{primary}}$ and total ray length $d_{\text{total}}$ with cutoff distance 50 (left).
+Here we use different colors to denote different parts (right).}
     % \caption{We plot the distance weight with respect to the primary and total ray length. Here the cutoff distance is 50.
@@
 \begin{equation}
-\int_{\Omega_R} g(X,Y,Z) d\Omega_R = \int_{\Gamma_R}\left(\int_{X_1}^{X_2} g(\alpha, Y, Z) d\alpha)\right)\mathbf{n} d\Gamma.
+\int_{\Omega_R} g(X,Y,Z) d\Omega_R = \int_{\Gamma_R}\left(\int_{X_1}^{X_2} g(\alpha, Y, Z) d\alpha)\right)\mathbf{n} d\Gamma_R.
 \end{equation}

### 2021-01-28 paper/sections/methods.tex
took a pass on otman's comments

@@
 %%single projector for every object
-When assembled for all parts,~\refeq{hfit}, yields a block upper triangular matrix $U$ and a sparse matrix $B$ ~(\reffig{multiparts}) which can be used to efficiently to compute the projection 
+When assembled for all parts,~\refeq{hfit}, yields a block upper triangular matrix $U$ and a sparse matrix $B$ ~(\reffig{multiparts}) which can be used to efficiently compute the projection 
 operator for the entire object $\Pi = U^{-1}B$. 
@@
     \defX\left(\refX\right) = \underbrace{\shapef\left(\refX\right)\bldmat\Pi}_{\Jkin\left(\refX\right)}\vc{q},
-\end{equation} where $\shapef\in\real^{3\times n} = \begin{bmatrix}P_1\left(\refX\right) \dots P_m\left(\refX\right)\end{bmatrix}$, $m$ is number of 
+\end{equation} where $\shapef\in\real^{3\times n} = \begin{bmatrix}P_1\left(\refX\right) \dots P_m\left(\refX\right)\end{bmatrix}$, $m$ is the number of 
 parts in the object, $\bldmat$ is the diagonal matrix of blending weights, $\Pi$ is the global projection operator and $\Jkin\left(\refX\right)$ is the kinematic Jacobian.

### 2021-01-28 paper/sections/intro.tex
referenced teaser and changed cutcell to cut-cell

@@
 We present the first truly meshless (no volumetric discretization required) algorithm for direct, nonlinear elastodynamic simulation of NURBs models. 
-Our nonlinear elastodynamic simulation scheme requires only a boundary description of the object (we do not strictly require a solid model) and
+Our nonlinear elastodynamic simulation scheme requires only a boundary description of the object (we do not strictly require a solid model, see \reffig{teaser}) and
 approriate physical parameters.  

### 2021-01-28 paper/sections/related.tex
referenced teaser and changed cutcell to cut-cell

@@
 Like volumetric meshing, this cutting operation can be difficult and our ideal method would avoid it. 
-Some cutcell algorithms assume engineering/manufacturing quality input, which puts tight requirements on input models~\cite{10.1145/3355089.3356576}.
+Some cut-cell algorithms assume engineering/manufacturing quality input, which puts tight requirements on input models~\cite{10.1145/3355089.3356576}.
 Finally these approaches require additional mechanisms to ensure that simulation results lie inside the shape space of the input model's primitives,
@@
 modelling geometry. 
-Like cutcell-based, Isogeometric Analysis approaches, additional constraints must be added to ensure that simulation output can be represented by the input model.
+Like cut-cell-based, Isogeometric Analysis approaches, additional constraints must be added to ensure that simulation output can be represented by the input model.
 Given this, we conclude that there is no existing algorithm that meets our desiderata for success.
@@
 algorithms for skinning~\cite{skinningcourse:2014}.
-Our method is entirely meshless (requiring no volumetric meshes or cutcells) and is compatible with standard constitutive models and time integrators.
+Our method is entirely meshless (requiring no volumetric meshes or cut-cells) and is compatible with standard constitutive models and time integrators.
 It guarantees that simulation output is directly consumable by the input modelling software and can ingest models which include large gaps,

### 2021-01-28 paper/sections/results.tex
took a pass on otman's comments
referenced teaser and changed cutcell to cut-cell

@@
 Models were created using Autodesk Fusion 360 \cite{AutodeskFusion360} and Rhinocerous 3D 7 \cite{mcneel-rhinoceros}. 
+Please see the supplemental video for more animation results.
 
@@
   \includegraphics[width=\columnwidth]{figures/independent_movement}
-  \caption{Our raycasting weight computation produces share-aware blending weights. Here the locality of our blending weights allows the two legs of the grumpy model to move independently. }
+  \caption{Our raycasting weight computation produces shape-aware blending weights. Here the locality of our blending weights allows the two legs of the grumpy model to move independently. }
   \label{fig:independence}
@@
 
+In \reffig{teaser} we directly simulate the NURBS surface model of a bouncy castle under a periodic wind force. 
 In \reffig{starship} we show an animation of the Space-V starship landing on Mars in a graceful way.
+Our SEM successfully handles the stiff materials, complicated geometry and non-trivial collisions in the scene.
 

### 2021-01-28 paper/sections/results.tex
sort table

@@
    \rowcolor[HTML]{DAE8FC} 
-   \textbf{Cantilever (soft)}       & 564 & Jelly & $1 e^3$ & $1e^6$& 0.45 & $2.53 e^{-2}$ & $4.45 e^{-3}$ & $4.78 e^{-1}$ & $1.86 e^{-1}$ & $1.20 e^{0}$ \\
-   \textbf{Cantilever (stiff)}      & 564 & Rubber & $1 e^3$ & $5 e^6$ & 0.45 & $2.48 e^{-2}$ & $6.37 e^{-3}$ & $5.10 e^{-1}$ & $1.68 e^{-1}$ & $1.19 e^{0}$ \\
-   \textbf{Coffee mug}              & 1800 & Jelly & $1 e^3$ & $1 e^4$ & 0.47 & $4.18 e^{-1}$ & $2.87 e^{-2}$ & $3.77 e^{0}$ & $2.08 e^{0}$ & $1.19 e^{1}$  \\
+   %\textbf{Cantilever (soft)}       & 564 & Jelly & $1 e^3$ & $1e^6$& 0.45 & $2.53 e^{-2}$ & $4.45 e^{-3}$ & $4.78 e^{-1}$ & $1.86 e^{-1}$ & $1.20 e^{0}$ \\
+   %\textbf{Cantilever (stiff)}      & 564 & Rubber & $1 e^3$ & $5 e^6$ & 0.45 & $2.48 e^{-2}$ & $6.37 e^{-3}$ & $5.10 e^{-1}$ & $1.68 e^{-1}$ & $1.19 e^{0}$ \\
+   \rowcolor[HTML]{DAE8FC} 
    \textbf{Rocket (soft)}           & 648 & Rubber & $1.27 e^3$ & $7 e^6$ & 0.4 & $2.83 e^{-2}$ & $5.62 e^{-3}$ & $5.58 e^{-1}$ & $6.83 e^{-2}$ & $1.26 e^{0}$ \\
    \textbf{Rocket (stiff)}          & 648 & Steel & $8 e^3$ & $2 e^{11}$ & 0.32 & $2.56 e^{-2}$ & $5.97 e^{-3}$ & $4.59 e^{-1}$ & $3.48 e^{-2}$ & $9.99 e^{-2}$ \\
+   \rowcolor[HTML]{DAE8FC} 
+   \textbf{Starship}                & 750 & Steel & $8 e^3$ & $2 e^{11}$ & 0.32 & $6.78 e^{-2}$ & $1.52 e^{-2}$ & $8.50 e^{-1}$ & $9.52 e^{-2}$ & $6.95 e^{-1}$ \\
+   \textbf{Tire}                    & 1404 & Rubber/Steel & $2 e^3$/$3 e^3$ & $1 e^6$/$7 e^9$ & 0.47 / 0.35 & $1.74 e^{-1}$ & $1.32 e^{-2}$ & $1.82 e^{0}$ & $6.79 e^{-1}$ & $1.28 e^{1}$ \\
+   \rowcolor[HTML]{DAE8FC} 
+   \textbf{Chicken}                 & 1773 & Jelly & $1.27 e^3$ & $1 e^4$ & 0.47 & $9.19 e^{-1}$ & $2.31 e^{-2}$ & $3.38 e^{0}$ & $1.91 e^{0}$ & $1.08 e^{1}$ \\
+   \textbf{Coffee mug}              & 1800 & Jelly & $1 e^3$ & $1 e^4$ & 0.47 & $4.18 e^{-1}$ & $2.87 e^{-2}$ & $3.77 e^{0}$ & $2.08 e^{0}$ & $1.19 e^{1}$  \\   
+   \rowcolor[HTML]{DAE8FC} 
    \textbf{Castle}                  & 1872 & Jelly & $1.27 e^3$ & $2 e^3$ & 0.45 & $4.84 e^{-1}$ & $1.44 e^{-2}$ & $1.52 e^{0}$ & $5.78 e^{-1}$ & $8.02 e^{0}$ \\
-   \textbf{Chicken}                 & 1773 & Jelly & $1.27 e^3$ & $1 e^4$ & 0.47 & $9.19 e^{-1}$ & $2.31 e^{-2}$ & $3.38 e^{0}$ & $1.91 e^{0}$ & $1.08 e^{1}$ \\
-   \textbf{Starship}                & 750 & Steel & $8 e^3$ & $2 e^{11}$ & 0.32 & $6.78 e^{-2}$ & $1.52 e^{-2}$ & $8.50 e^{-1}$ & $9.52 e^{-2}$ & $6.95 e^{-1}$ \\
    \textbf{Astronaut (soft)}       & 3609 & Rubber &  $1.27 e^3$  & $1 e^6$  & 0.47 & $2.47 e^{0}$ & $6.71 e^{-2}$ & $1.13 e^{1}$ & $1.09 e^{1}$ & $1.41 e^{1}$ \\
-   \textbf{Tire}                    & 1404 & Rubber/Steel & $2 e^3$/$3 e^3$ & $1 e^6$/$7 e^9$ & 0.47 / 0.35 & $1.74 e^{-1}$ & $1.32 e^{-2}$ & $1.82 e^{0}$ & $6.79 e^{-1}$ & $1.28 e^{1}$ \\
    \hline
@@
 We further demonstrate \emph{qualitative} convergence of SEM with respect to linear tetrahedral finite elements when increasing the number of patches in use.
-In \reffig{convergence}, we compare identical cantilevered beams, simulated using a Neohookean constitutive model (Young's Modulus $=$ $0.001$ GPa, Poisson's Ratio $=$ $0.45$) and implicit time integration with a timestep of $0.01$s.
+In \reffig{convergence}, we compare the static configurations of identical cantilevered beams, simulated using a Neohookean constitutive model (Young's Modulus $=$ $0.001$ GPa, Poisson's Ratio $=$ $0.45$).
 We use SEM with quadratic polynomials for this test and observe that our SEM simulation, made of 24 independent NURBS patches, shows good agreement with FEM.

### 2021-01-28 paper/sections/results.tex
edits

@@
 In \reffig{starship} we show an animation of the Space-V starship landing on Mars in a graceful way. 
-Our SEM successfully handles the stiff materials, complicated geometry and non-trivial collisions in the scene.
+This example shows the benefit of NURBS modeling, a relatively small number of primitives can represent a complex shape.
+SEM successfully handles the stiff materials, complicated geometry and non-trivial collisions in this scene.
 

### 2021-02-03 cpp/include/vem3dmesh_neohookean_dq2.h
Added mexed version of mass matrix and error hessian

@@ -7,7 +7,6 @@
 //#include <vem3d_neohookean_dq2.h>
 #include <d2psi_neohookean_dF2.h> // TODO remove and replace with neohookean dq2
 
-
 namespace sim {
 
 template<typename DerivedRet, typename DerivedC, typename DerivedVol, typename DerivedParam>

### 2021-02-03 cpp/include/vem_error_matrix.h

@@ -1,0 +1,20 @@
+#ifndef SIM_VEM_ERROR_MATRIX_H
+#define SIM_VEM_ERROR_MATRIX_H
+
+#include <Eigen/Dense>
+#include <EigenTypes.h>
+
+namespace sim {
+
+	template<typename DerivedRet, typename DerivedY>
+	void vem_error_matrix(Eigen::MatrixXx<DerivedRet> &g,
+		const EigenMatrixList<DerivedY>& Y,
+		const Eigen::SparseMatrix<double> & L,
+		const EigenVectorList<int>& W_I,
+		const EigenVectorList<int>& C_I,
+		int d, int k, int n);
+}
+
+#include <../src/vem_error_matrix.cpp>
+
+#endif // SIM_VEM_ERROR_MATRIX_H

### 2021-02-03 cpp/include/vem_mass_matrix.h

@@ -1,0 +1,21 @@
+#ifndef SIM_VEM_MASS_MATRIX_H
+#define SIM_VEM_MASS_MATRIX_H
+
+#include <Eigen/Dense>
+#include <EigenTypes.h>
+
+namespace sim {
+
+template<typename DerivedRet, typename DerivedY, typename DerivedM>
+void vem_mass_matrix(Eigen::MatrixXx<DerivedRet> &g,
+	const EigenMatrixList<DerivedY>& Y,
+	const Eigen::SparseMatrix<double> & L,
+	const Eigen::VectorXx<DerivedM>& mass,
+	const EigenVectorList<int>& W_I,
+	const EigenVectorList<int>& C_I,
+	int d, int k, int n);
+}
+
+#include <../src/vem_mass_matrix.cpp>
+
+#endif // SIM_VEM_MASS_MATRIX_H

### 2021-02-03 cpp/src/vem3dmesh_neohookean_dq2.cpp

@@ -6,6 +6,8 @@
 #include <omp.h>
 #endif
 
+#include <chrono>
+
 template<typename DerivedRet, typename DerivedC, typename DerivedVol, typename DerivedParam>
 void sim::vem3dmesh_neohookean_dq2(Eigen::MatrixXx<DerivedRet> &g,
 	const Eigen::VectorXx<DerivedC> &c,
@@ -18,11 +20,19 @@ void sim::vem3dmesh_neohookean_dq2(Eigen::MatrixXx<DerivedRet> &g,
 	g.resize(3 * (k*n + m), 3 * (k*n + m));
 	g.setZero();
 
+	double time_0 = 0;
+	double time_1 = 0;
+	double time_2 = 0;
+	double time_3 = 0;
+	using namespace std::chrono;
+
 	for (int i = 0; i < dF_dc.size(); ++i) {
 		//std::cout << "i: " << i << std::endl;
 		Eigen::MatrixXx<DerivedParam> CD = params.row(i);
 
 		Eigen::Matrix<DerivedC, 9, 9> d2psi_dF2;
+
+		//high_resolution_clock::time_point t0 = high_resolution_clock::now();
 
 		// TODO move this to separate file
 		Eigen::Matrix<DerivedC, 9, 1> F_flat; 
@@ -31,33 +41,20 @@ void sim::vem3dmesh_neohookean_dq2(Eigen::MatrixXx<DerivedRet> &g,
 			int idx = W_I[i](j);
 			F_flat += dF_dc[i].block(0,3*k*j, 9,3*k) * c.segment(3 * k*idx, 3 * k);
 		}
-		//std::cout << " did we get fflat? " << F_flat << std::endl;
 		Eigen::Matrix3d F = unflatten<3, 3>(F_flat);
+
+		//high_resolution_clock::time_point t1 = high_resolution_clock::now();
 
 		//std::cout << "F: " << F << std::endl;
 		d2psi_neohookean_dF2(d2psi_dF2, F, CD);
 		//std::cout << "d2psi: " << i << ": \n" << d2psi_dF2 << std::endl;
 
+		//high_resolution_clock::time_point t2 = high_resolution_clock::now();
+
 		Eigen::MatrixXx<DerivedRet> tmp = dF_dc[i].transpose() * d2psi_dF2 * dF_dc[i] * volume(i);
-		if (i == 0) {
[…]

### 2021-02-03 cpp/src/vem_error_matrix.cpp

@@ -1,0 +1,74 @@
+#ifdef VEM_USE_OPENMP
+#include <omp.h>
+#endif
+
+template<typename DerivedRet, typename DerivedY>
+void sim::vem_error_matrix(Eigen::MatrixXx<DerivedRet> &g,
+	const EigenMatrixList<DerivedY>& Y,
+	const Eigen::SparseMatrix<double> & L,
+	const EigenVectorList<int>& W_I,
+	const EigenVectorList<int>& C_I,
+	int d, int k, int n) {
+
+	std::cout << "ysize: " << Y.size() << std::endl;
+	int sz = L.cols();
+
+	Eigen::MatrixXd tmp1;
+	Eigen::MatrixXd tmp2;
+	tmp1.resize(L.rows(), L.rows());
+	tmp2.resize(L.rows(), L.cols());
+	tmp1.setZero();
+	tmp2.setZero();
+	
+	int dk = d * k;
+	int dkn = dk * n;
+
+	auto assembly_idx = [=](int idx, int dkm, int i) {
+		int ret;
+		if (idx < dkm) {
+			ret = W_I[i](idx / dk) * dk + idx % dk;
+		}
+		else {
+			ret = dkn + C_I[i]((idx - dkm) / d) * d + idx % d;
+		}
+		return ret;
+	};
+
+	for (int i = 0; i < Y.size(); ++i) {
+		Eigen::MatrixXd tmp = Y[i].transpose() * Y[i];
+
+		int dkm = dk * W_I[i].rows(); // number of associated parts
+
+		// Assembly
+		for (int jj = 0; jj < tmp.cols(); ++jj) {
+			for (int ii = 0; ii < tmp.rows(); ++ii) {
+				int r = assembly_idx(ii, dkm, i);
+				int c = assembly_idx(jj, dkm, i);
+				tmp1.coeffRef(r, c) += tmp(ii, jj);
+			}
+		}
+
+		for (int dim = 0; dim < d; ++dim) {
+			// Copy weighted polynomial basis matrix
+			for (int ii = 0; ii < W_I[i].rows(); ++ii) {
+				for (int jj = 0; jj < k; ++jj) {
+					int r = W_I[i](ii)*dk + dim * k + jj;
+					int c = i * d + dim;
+					// Same for each dim, so only copy from one row of Y
+					tmp2.coeffRef(r, c) = Y[i](0, ii*dk + jj);
+				}
+			}
+
[…]

### 2021-02-03 cpp/src/vem_mass_matrix.cpp

@@ -1,0 +1,47 @@
+#ifdef VEM_USE_OPENMP
+#include <omp.h>
+#endif
+
+template<typename DerivedRet, typename DerivedY, typename DerivedM>
+void sim::vem_mass_matrix(Eigen::MatrixXx<DerivedRet> &g,
+	const EigenMatrixList<DerivedY>& Y,
+	const Eigen::SparseMatrix<double> & L,
+	const Eigen::VectorXx<DerivedM>& mass,
+	const EigenVectorList<int>& W_I,
+	const EigenVectorList<int>& C_I,
+	int d, int k, int n) {
+
+	int sz = L.rows();
+	g.resize(sz,sz);
+	g.setZero();
+
+	int dk = d * k;
+	int dkn = dk * n;
+	
+	auto assembly_idx = [=](int idx, int dkm, int i) {
+		int ret;
+		if (idx < dkm) {
+			ret = W_I[i](idx / dk) * dk + idx % dk;
+		}
+		else {
+			ret = dkn + C_I[i]((idx - dkm) / d) * d + idx % d;
+		}
+		return ret;
+	};
+
+	for (int i = 0; i < Y.size(); ++i) {
+		Eigen::MatrixXd tmp = Y[i].transpose() * Y[i] * mass(i);
+
+		int dkm = dk * W_I[i].rows(); // number of associated parts
+
+		// Assembly
+		for (int jj = 0; jj < tmp.cols(); ++jj) {
+			for (int ii = 0; ii < tmp.rows(); ++ii) {
+				int r = assembly_idx(ii, dkm, i);
+				int c = assembly_idx(jj, dkm, i);
+				g.coeffRef(r, c) += tmp(ii, jj);
+			}
+		}
+	}
+	g = L.transpose() * g * L;
+}

### 2021-02-03 examples/quadratic_rounded_cube.m

@@ -13,12 +13,13 @@ YM = 1e6; %in Pascals
 pr =  0.45;
 [lambda, mu] = emu_to_lame(YM, pr);
 options.order = 2;
-% options.rho = 1e-1;
 options.rho = 1e3;
 options.gravity = -10;
 options.pin_function = @pin_function;
 options.lambda = lambda;
 options.mu = mu;
+options.k_stability=YM;
+options.distance_cutoff=cutoff_heuristic(part, 0.9);
 
 % vem_simulate_nurbs_newtons(part, options);
 vem_simulate_nurbs(part, options);

### 2021-02-03 matlab/mex/vem_error_matrix.cpp

@@ -1,0 +1,40 @@
+#include <mex.h>
+
+//matlab hooks from libigl (taken directly from gptoolbox)
+#include <igl/C_STR.h>
+#include <igl/matlab/mexErrMsgTxt.h>
+#undef assert
+#define assert( isOK ) ( (isOK) ? (void)0 : (void) mexErrMsgTxt(C_STR(__FILE__<<":"<<__LINE__<<": failed assertion `"<<#isOK<<"'"<<std::endl) ) )
+
+#include <igl/matlab/MexStream.h>
+#include <igl/matlab/parse_rhs.h>
+#include <igl/matlab/prepare_lhs.h>
+#include <igl/matlab/validate_arg.h>
+#include <igl/list_to_matrix.h>
+
+#include <parse_matlab_cells.h>
+
+#include <vem_error_matrix.h>
+
+/* The gateway function */
+void mexFunction(int nlhs, mxArray *plhs[],
+	int nrhs, const mxArray *prhs[]) {
+	/* variable declarations here */
+	Eigen::MatrixXd g;
+
+	Eigen::SparseMatrix<double> L;
+	EigenMatrixList<double> Y;
+	EigenVectorList<int> W_I, C_I;
+
+	sim::parse_cell(prhs, 0, Y);
+	igl::matlab::parse_rhs(prhs + 1, L);
+	sim::parse_cell_index(prhs, 2, W_I);
+	sim::parse_cell_index(prhs, 3, C_I);
+	int d = (int)*mxGetPr(prhs[4]);
+	int k = (int)*mxGetPr(prhs[5]);
+	int n = (int)*mxGetPr(prhs[6]);
+
+	sim::vem_error_matrix(g, Y, L, W_I, C_I, d, k, n);
+
+	igl::matlab::prepare_lhs_double(g, plhs);
+}

### 2021-02-03 matlab/mex/vem_mass_matrix.cpp

@@ -1,0 +1,42 @@
+#include <mex.h>
+
+//matlab hooks from libigl (taken directly from gptoolbox)
+#include <igl/C_STR.h>
+#include <igl/matlab/mexErrMsgTxt.h>
+#undef assert
+#define assert( isOK ) ( (isOK) ? (void)0 : (void) mexErrMsgTxt(C_STR(__FILE__<<":"<<__LINE__<<": failed assertion `"<<#isOK<<"'"<<std::endl) ) )
+
+#include <igl/matlab/MexStream.h>
+#include <igl/matlab/parse_rhs.h>
+#include <igl/matlab/prepare_lhs.h>
+#include <igl/matlab/validate_arg.h>
+#include <igl/list_to_matrix.h>
+
+#include <parse_matlab_cells.h>
+
+#include <vem_mass_matrix.h>
+
+/* The gateway function */
+void mexFunction(int nlhs, mxArray *plhs[],
+                 int nrhs, const mxArray *prhs[]) {
+    /* variable declarations here */
+    Eigen::MatrixXd g;
+
+    Eigen::VectorXd mass;
+	Eigen::SparseMatrix<double> L;
+	EigenMatrixList<double> Y;
+	EigenVectorList<int> W_I, C_I;
+
+	sim::parse_cell(prhs, 0, Y);
+	igl::matlab::parse_rhs(prhs+1, L);
+	igl::matlab::parse_rhs_double(prhs + 2, mass);
+	sim::parse_cell_index(prhs, 3, W_I);
+	sim::parse_cell_index(prhs, 4, C_I);
+	int d = (int)*mxGetPr(prhs[5]);
+	int k = (int)*mxGetPr(prhs[6]);
+	int n = (int)*mxGetPr(prhs[7]);
+
+	
+	sim::vem_mass_matrix(g, Y, L, mass, W_I, C_I, d, k, n);
+	igl::matlab::prepare_lhs_double(g, plhs);
+}

### 2021-02-03 matlab/vem_dx_dc.m

@@ -1,9 +1,10 @@
-function [Y,S] = vem_dx_dc(x, x_coms, w, W_I, com_map, order, k)
+function [Y,S,C_I] = vem_dx_dc(x, x_coms, w, W_I, com_map, order, k)
     d = size(x,1);
     n = size(x,2);
     
     Y = cell(n,1);
     S = cell(n,1);
+    C_I = cell(n,1);
     
     for ii=1:n
         m = numel(W_I{ii});
@@ -38,6 +39,7 @@ function [Y,S] = vem_dx_dc(x, x_coms, w, W_I, com_map, order, k)
         end
         Y{ii} = Yi;
         S{ii} = sparse(S_i);
+        C_I{ii} = C;
     end
 end
 

### 2021-02-03 matlab/vem_simulate_nurbs.m

@@ -92,8 +92,8 @@ function vem_simulate_nurbs(parts, varargin)
         com_cluster, config.order, config.fitting_mode);
 
     % Build Monomial bases for all quadrature points
-    [Y,Y_S] = vem_dx_dc(V, x0_coms, w, w_I, com_map, config.order, k);
-    [Y0,Y0_S] = vem_dx_dc(x0, x0_coms, w0, w0_I, com_map, config.order, k);
+    [Y,Y_S,C_I] = vem_dx_dc(V, x0_coms, w, w_I, com_map, config.order, k);
+    [Y0,Y0_S,C0_I] = vem_dx_dc(x0, x0_coms, w0, w0_I, com_map, config.order, k);
 
     % Fixed x values.
     x_fixed = zeros(size(x0));
@@ -117,8 +117,9 @@ function vem_simulate_nurbs(parts, varargin)
     f_external = config.dt*P*(L' * dext_dc); 
 
     % Compute mass matrices
-    ME = vem_error_matrix(Y0, Y0_S, L, d);
-    M = vem_mass_matrix(Y, Y_S, L, config.rho .* vol);
+    ME = vem_error_matrix(Y0, L, w0_I, C0_I, d, k, n);
+    M = vem_mass_matrix(Y, L, config.rho .* vol, w_I, C_I, d, k, n);
+    
     M = (M + config.k_stability*ME); % sparse?
 
     ii=1;

### 2021-02-03 matlab/vem_simulate_nurbs_with_collision.m

@@ -85,14 +85,7 @@ function vem_simulate_nurbs_with_collision_new(parts, varargin)
     % Lame parameters concatenated.
     params = [config.mu * 0.5, config.lambda * 0.5];
     params = repmat(params,size(V,2),1);
-        
-    % Gravity force vector.
-  	% f_gravity = repmat([0 0 config.rho*config.gravity], size(x0,2),1)';
-    % f_gravity = config.dt*P*f_gravity(:);
-    
-    % Undeformed Center of mass
-    x0_com = mean(x0,2);
-    
+
     % Compute Shape weights
     [w, w_I] = nurbs_blending_weights(parts, V', config.distance_cutoff, ...
         'Enable_Secondary_Rays', config.enable_secondary_rays);
@@ -112,8 +105,8 @@ function vem_simulate_nurbs_with_collision_new(parts, varargin)
         com_cluster, config.order, config.fitting_mode);
     
     % Build Monomial bases for all quadrature points
-    [Y,Y_S] = vem_dx_dc(V, x0_coms, w, w_I, com_map, config.order, k);
-    [Y0,Y0_S] = vem_dx_dc(x0, x0_coms, w0, w0_I, com_map, config.order, k);
+    [Y,Y_S,com_I] = vem_dx_dc(V, x0_coms, w, w_I, com_map, config.order, k);
+    [Y0,Y0_S,C0_I] = vem_dx_dc(x0, x0_coms, w0, w0_I, com_map, config.order, k);
     
     % Fixed x values.
     x_fixed = zeros(size(x0));
@@ -145,8 +138,9 @@ function vem_simulate_nurbs_with_collision_new(parts, varargin)
     config.rho(bottom_ids) = 6e3;   % you can make it denser too
 
     % Compute mass matrices
-    ME = vem_error_matrix(Y0, Y0_S, L, d);
-    M = vem_mass_matrix(Y, Y_S, L, config.rho .* vol);
+    ME = vem_error_matrix(Y0, L, w0_I, C0_I, d, k, n);
+    M = vem_mass_matrix(Y, L, config.rho .* vol, w_I, com_I, d, k, n);
+
     M = (M + config.k_stability*ME); % sparse?
     % Save & load these matrices for large models to save time.
     % save('saveM.mat','M');
@@ -158,7 +152,7 @@ function vem_simulate_nurbs_with_collision_new(parts, varargin)
     [verts,faces] = triangulate_iges(parts);
     collision_ratio = config.collision_ratio;
     face_normal = normals(verts, faces);
[…]

### 2021-04-06 paper/sections/results.tex
add lamppost timing and reorder the timing tables
specify the type of examples for all animations
add nnz and sparsity plot; clean deprecated figures

@@
 \begin{table*}[h]
+  \setlength{\tabcolsep}{3pt}
   \caption{Performance of, and parameters for the Shape Matching Element Method on all examples. All wall-clock timings are reported in seconds, physical parameters are
@@
   \textbf{Sample} is the time taken to sample and construct $\samp{\Jnurbs}$ and $\samp{\Pmatx}$, \textbf{Quad} is the time taken to generate quadrature points via raycasting,
-  \textbf{Weights} is the time to compute blending weights, \textbf{Build $\Pi$} is the time to build the projection operator and  \textbf{Step} is the average time required to step the simulation (not including collision detection).
+  \textbf{Weights} is the time to compute blending weights, \textbf{Build $\Pi$} is the time to build the projection operator, \textbf{Step} is the average time required to step the simulation (not including collision detection)
+  \revise{and \textbf{Nnz} is the ratio of nonzeros to the total number of elements in the stiffness matrix. }
+  \revise{We use linearly implicit time integration on all examples and thus the number of step per frame is 1 in the simulation.}
   }
@@
   \begin{center}
-  \begin{tabular}{l c c c c c c c c c c}
-   \textbf{Example} & $|\vc{q}|$ & \textbf{Material}&  $\rho$(kg/m$^3$) & \textbf{E}(Pa) & $\mu$ &\textbf{Sample}(s) & \textbf{Quad}(s)& \textbf{Weights}(s) & \textbf{Build $\Pi$}(s)& \textbf{Step}(s)\\
+  \begin{tabular}{l c c c c c c c c c c c}
+   \textbf{Example} & $|\vc{q}|$ & \textbf{Material}&  $\rho$(kg/m$^3$) & \textbf{E}(Pa) & $\mu$ &\textbf{Sample}(s) & \textbf{Quad}(s)& \textbf{Weights}(s) & \textbf{Build $\Pi$}(s) & \textbf{Step}(s) & \revise{\textbf{Nnz}(\%)} \\
    \hline 
@@
    \rowcolor[HTML]{DAE8FC} 
-   \textbf{Rocket (soft)}           & 648 & Rubber & $1.27 e^3$ & $7 e^6$ & 0.4 & $2.83 e^{-2}$ & $5.62 e^{-3}$ & $5.58 e^{-1}$ & $6.83 e^{-2}$ & $1.26 e^{0}$ \\
-   \textbf{Rocket (stiff)}          & 648 & Steel & $8 e^3$ & $2 e^{11}$ & 0.32 & $2.56 e^{-2}$ & $5.97 e^{-3}$ & $4.59 e^{-1}$ & $3.48 e^{-2}$ & $9.99 e^{-2}$ \\
+   \textbf{Rocket (soft)}           & 648 & Rubber & $1.27 e^3$ & $7 e^6$ & 0.40 & $2.83 e^{-2}$ & $5.62 e^{-3}$ & $5.58 e^{-1}$ & $6.83 e^{-2}$ & $1.26 e^{0}$ & \revise{100} \\
+   \textbf{Rocket (stiff)}          & 648 & Steel & $8 e^3$ & $2 e^{11}$ & 0.32 & $2.56 e^{-2}$ & $5.97 e^{-3}$ & $4.59 e^{-1}$ & $3.48 e^{-2}$ & $9.99 e^{-2}$ & \revise{100} \\
    \rowcolor[HTML]{DAE8FC} 
-   \textbf{Starship}                & 750 & Steel & $8 e^3$ & $2 e^{11}$ & 0.32 & $6.78 e^{-2}$ & $1.52 e^{-2}$ & $8.50 e^{-1}$ & $9.52 e^{-2}$ & $6.95 e^{-1}$ \\
-   \textbf{Tire}                    & 1404 & Rubber/Steel & $2 e^3$/$3 e^3$ & $1 e^6$/$7 e^9$ & 0.47 / 0.35 & $1.74 e^{-1}$ & $1.32 e^{-2}$ & $1.82 e^{0}$ & $6.79 e^{-1}$ & $1.28 e^{1}$ \\
+   \textbf{Starship}                & 750 & Steel & $8 e^3$ & $2 e^{11}$ & 0.32 & $6.78 e^{-2}$ & $1.52 e^{-2}$ & $8.50 e^{-1}$ & $9.52 e^{-2}$ & $6.95 e^{-1}$ & \revise{100} \\
+   \textbf{Lamppost}                & 942 & Rubber & $3 e^3$ & $1 e^6$ & 0.40 & $1.28 e^{-1}$ & $1.27 e^{-2}$ & $1.89 e^{0}$ & $1.07 e^{0}$ & $2.22 e^{0}$ & \revise{56.38} \\
    \rowcolor[HTML]{DAE8FC} 
-   \textbf{Chicken}                 & 1773 & Jelly & $1.27 e^3$ & $1 e^4$ & 0.47 & $9.19 e^{-1}$ & $2.31 e^{-2}$ & $3.38 e^{0}$ & $1.91 e^{0}$ & $1.08 e^{1}$ \\
-   \textbf{Coffee mug}              & 1800 & Jelly & $1 e^3$ & $1 e^4$ & 0.47 & $4.18 e^{-1}$ & $2.87 e^{-2}$ & $3.77 e^{0}$ & $2.08 e^{0}$ & $1.19 e^{1}$  \\   
+   \textbf{Tire}                    & 1404 & Rubber/Steel & $2 e^3$/$3 e^3$ & $1 e^6$/$7 e^9$ & 0.47/0.35 & $1.74 e^{-1}$ & $1.32 e^{-2}$ & $1.82 e^{0}$ & $6.79 e^{-1}$ & $1.28 e^{1}$ & \revise{92.5} \\
+   \textbf{Chicken}                 & 1773 & Jelly & $1.27 e^3$ & $1 e^4$ & 0.47 & $9.19 e^{-1}$ & $2.31 e^{-2}$ & $3.38 e^{0}$ & $1.91 e^{0}$ & $1.08 e^{1}$ & \revise{100} \\
    \rowcolor[HTML]{DAE8FC} 
-   \textbf{Castle}                  & 1872 & Jelly & $1.27 e^3$ & $2 e^3$ & 0.45 & $4.84 e^{-1}$ & $1.44 e^{-2}$ & $1.52 e^{0}$ & $5.78 e^{-1}$ & $8.02 e^{0}$ \\
-   \textbf{Astronaut (soft)}       & 3609 & Rubber &  $1.27 e^3$  & $1 e^6$  & 0.47 & $2.47 e^{0}$ & $6.71 e^{-2}$ & $1.13 e^{1}$ & $1.09 e^{1}$ & $1.41 e^{1}$ \\
+   \textbf{Coffee mug (soft)}       & 1800 & Jelly & $1 e^3$ & $1 e^4$ & 0.47 & $4.18 e^{-1}$ & $2.87 e^{-2}$ & $3.77 e^{0}$ & $2.08 e^{0}$ & $1.19 e^{1}$ & \revise{100} \\ 
+   \textbf{Coffee mug (stiff)}      & 1800 & Rubber & $1 e^3$ & $4 e^7$ & 0.40 & $4.15 e^{-2}$ & $1.17 e^{-2}$ & $6.86 e^{-1}$ & $1.09 e^{-1}$ & $5.00 e^{-1}$  & \revise{81.42} \\ 
+   \rowcolor[HTML]{DAE8FC} 
+   \textbf{Castle}                  & 1872 & Jelly & $1.27 e^3$ & $2 e^3$ & 0.45 & $4.84 e^{-1}$ & $1.44 e^{-2}$ & $1.52 e^{0}$ & $5.78 e^{-1}$ & $8.02 e^{0}$ & \revise{100} \\
+   \textbf{Grumpy}                  & 2010 & Jelly & $1 e^3$ & $1 e^4$ & 0.40 & $2.92 e^{-1}$ & $1.23 e^{-2}$ & $2.29 e^{0}$ & $2.89 e^{0}$ & $6.02 e^{0}$ & \revise{84.15} \\    
+   \rowcolor[HTML]{DAE8FC} 
+   \textbf{Astronaut (soft)}        & 3609 & Rubber &  $1.27 e^3$  & $1 e^6$  & 0.47 & $2.47 e^{0}$ & $6.71 e^{-2}$ & $1.13 e^{1}$ & $1.09 e^{1}$ & $1.41 e^{1}$ & \revise{49.04} \\
+   \revise{\textbf{Squid}}          & \revise{4347} & \revise{Jelly} & \revise{ $1 e^3$ } & \revise{$2 e^4$ } & \revise{0.40} & \revise{$4.48 e^{-1}$} & \revise{$2.47 e^{-2}$} & \revise{$1.54 e^{1}$} & \revise{$4.48 e^{1}$} & \revise{$6.73 e^{0}$} & \revise{25.47} \\
+   \rowcolor[HTML]{DAE8FC} 
    \hline
@@
   \end{table*}
+
+% \begin{table*}[h]
+%   \caption{Performance of, and parameters for the Shape Matching Element Method on all examples. All wall-clock timings are reported in seconds, physical parameters are
+%   reported with appropriate units. $\rho$ is the applied density, \textbf{E} is the  Young's Modulus and $\mu$ is the Poisson's ratio.
+%   \textbf{Sample} is the time taken to sample and construct $\samp{\Jnurbs}$ and $\samp{\Pmatx}$, \textbf{Quad} is the time taken to generate quadrature points via raycasting,
+%   \textbf{Weights} is the time to compute blending weights, \textbf{Build $\Pi$} is the time to build the projection operator and  \textbf{Step} is the average time required to step the simulation (not including collision detection).
+%   \revise{We use linearly implicit time integration on all examples and thus the number of step per frame is 1 in the simulation.}
+%   }
+%   \label{tbl:perf}
+%   \begin{center}
+%   \begin{tabular}{l c c c c c c c c c c}
+%    \textbf{Example} & $|\vc{q}|$ & \textbf{Material}&  $\rho$(kg/m$^3$) & \textbf{E}(Pa) & $\mu$ &\textbf{Sample}(s) & \textbf{Quad}(s)& \textbf{Weights}(s) & \textbf{Build $\Pi$}(s)& \textbf{Step}(s) \\
+%    \hline 
+%    \rowcolor[HTML]{DAE8FC}
+%    %\textbf{Cantilever (soft)}       & 564 & Jelly & $1 e^3$ & $1e^6$& 0.45 & $2.53 e^{-2}$ & $4.45 e^{-3}$ & $4.78 e^{-1}$ & $1.86 e^{-1}$ & $1.20 e^{0}$ \\
+%    %\textbf{Cantilever (stiff)}      & 564 & Rubber & $1 e^3$ & $5 e^6$ & 0.45 & $2.48 e^{-2}$ & $6.37 e^{-3}$ & $5.10 e^{-1}$ & $1.68 e^{-1}$ & $1.19 e^{0}$ \\
+%    \rowcolor[HTML]{DAE8FC} 
+%    \textbf{Rocket (soft)}           & 648 & Rubber & $1.27 e^3$ & $7 e^6$ & 0.40 & $2.83 e^{-2}$ & $5.62 e^{-3}$ & $5.58 e^{-1}$ & $6.83 e^{-2}$ & $1.26 e^{0}$ \\
+%    \textbf{Rocket (stiff)}          & 648 & Steel & $8 e^3$ & $2 e^{11}$ & 0.32 & $2.56 e^{-2}$ & $5.97 e^{-3}$ & $4.59 e^{-1}$ & $3.48 e^{-2}$ & $9.99 e^{-2}$ \\
+%    \rowcolor[HTML]{DAE8FC} 
+%    \textbf{Starship}                & 750 & Steel & $8 e^3$ & $2 e^{11}$ & 0.32 & $6.78 e^{-2}$ & $1.52 e^{-2}$ & $8.50 e^{-1}$ & $9.52 e^{-2}$ & $6.95 e^{-1}$ \\
+%    \textbf{Lamppost}                & 942 & Rubber & $3 e^3$ & $1 e^6$ & 0.40 & $1.28 e^{-1}$ & $1.27 e^{-2}$ & $1.89 e^{0}$ & $1.07 e^{0}$ & $2.22 e^{0}$ \\
+%    \rowcolor[HTML]{DAE8FC} 
+%    \textbf{Tire}                    & 1404 & Rubber/Steel & $2 e^3$/$3 e^3$ & $1 e^6$/$7 e^9$ & 0.47/0.35 & $1.74 e^{-1}$ & $1.32 e^{-2}$ & $1.82 e^{0}$ & $6.79 e^{-1}$ & $1.28 e^{1}$ \\
+%    \textbf{Chicken}                 & 1773 & Jelly & $1.27 e^3$ & $1 e^4$ & 0.47 & $9.19 e^{-1}$ & $2.31 e^{-2}$ & $3.38 e^{0}$ & $1.91 e^{0}$ & $1.08 e^{1}$ \\
+%    \rowcolor[HTML]{DAE8FC} 
+%    \textbf{Coffee mug (soft)}       & 1800 & Jelly & $1 e^3$ & $1 e^4$ & 0.47 & $4.18 e^{-1}$ & $2.87 e^{-2}$ & $3.77 e^{0}$ & $2.08 e^{0}$ & $1.19 e^{1}$  \\ 
+%    \textbf{Coffee mug (stiff)}      & 1800 & Rubber & $1 e^3$ & $4 e^7$ & 0.40 & $4.15 e^{-2}$ & $1.17 e^{-2}$ & $6.86 e^{-1}$ & $1.09 e^{-1}$ & $5.00 e^{-1}$  \\ 
+%    \rowcolor[HTML]{DAE8FC} 
+%    \textbf{Castle}                  & 1872 & Jelly & $1.27 e^3$ & $2 e^3$ & 0.45 & $4.84 e^{-1}$ & $1.44 e^{-2}$ & $1.52 e^{0}$ & $5.78 e^{-1}$ & $8.02 e^{0}$ \\
+%    \textbf{Grumpy}                  & 2010 & Jelly & $1 e^3$ & $1 e^4$ & 0.40 & $2.92 e^{-1}$ & $1.23 e^{-2}$ & $2.29 e^{0}$ & $2.89 e^{0}$ & $6.02 e^{0}$  \\    
+%    \rowcolor[HTML]{DAE8FC} 
+%    \textbf{Astronaut (soft)}        & 3609 & Rubber &  $1.27 e^3$  & $1 e^6$  & 0.47 & $2.47 e^{0}$ & $6.71 e^{-2}$ & $1.13 e^{1}$ & $1.09 e^{1}$ & $1.41 e^{1}$ \\
+%    \textbf{Squid}                   & 4347 & Rubber &  $1 e^3$  & $2 e^4$  & 0.40 & $4.48 e^{-1}$ & $2.47 e^{-2}$ & $1.54 e^{1}$ & $4.48 e^{1}$ & $6.73 e^{0}$ \\
+%    \rowcolor[HTML]{DAE8FC} 
+%    \hline
+%   \end{tabular}
+%   \end{center}
+
+%   \end{table*}
 
@@
 \reftbl{perf} shows the size of all our examples along with performance statistics and relevant parameters. 
+\revise{We also plot the sparsity patterns of several stiffness matrices in our examples in \reffig{stiffness_sparsity}. }
 Note that the individual parts of all models are \emph{not connected}, and no continuity constraints or constructive solid geometry operations have been applied. 
 Rather, the SEM approach implicitly couples the boundary parts together to allow for seamless volumetric simulation. 
+We use quadratic deformation for all the animation results, except the rocket (stiff) and starship which use linear deformation.
 Models were created using Autodesk Fusion 360 \cite{AutodeskFusion360} and Rhinocerous 3D 7 \cite{mcneel-rhinoceros}. 
@@
   \label{fig:patchtest}
+\end{figure}
+
+\begin{figure}[h]
+  \includegraphics[width=\columnwidth]{figures/stiffness_sparsity.pdf}
+  \caption{\revise{Our stiffness matrices can be sparse, while the corresponding matrix has to be dense in the traditional Boundary Element Method \cite{ArtDefo1999}. 
+  Sparsity of our stiffness matrix is primarily determined by weight sparsity (number of non-zero weights at a given quadrature point) and polynomial order. }}
+  \label{fig:stiffness_sparsity}
 \end{figure}

### 2021-04-07 paper/sections/results.tex
added figures

@@
 
+\begin{figure}[h]
+  \includegraphics[width=\columnwidth]{figures/scalability.pdf}
+  \caption{\revise{As a more extreme example of the shape-aware property of our algorithm, we simulate a falling octopus, and observe the 8 limbs moving independently.}}
+  \label{fig:scalability}
+\end{figure}
+
 %% Results that relate to preprocessing
@@
   Sparsity of our stiffness matrix is primarily determined by weight sparsity (number of non-zero weights at a given quadrature point) and polynomial order. }}
-  \label{fig:stiffness_sparsity}
 \end{figure}
@@
 
-We further demonstrate \emph{qualitative} convergence of SEM with respect to linear tetrahedral finite elements when increasing the number of patches in use.
+We demonstrate \emph{qualitative} convergence of SEM with respect to linear tetrahedral finite elements when increasing the number of patches in use.
 In \reffig{convergence}, we compare the static configurations of identical cantilevered beams, simulated using a Neohookean constitutive model (Young's Modulus $=$ $0.001$ GPa, Poisson's Ratio $=$ $0.45$).
+\revise{We further perform basic \emph{quantitative} comparison against FEM by measuring the maximum deflection of the beam as we increase the number of patches along the beam.
+In \reffig{plot_beam_deflection}, the maximum beam deflection of the SEM beams is reported as a percentage of the FEM beam's maximum deflection.
+The parameters of the experiment match that of the previous qualitative test.}
+
+%\begin{wrapfigure}{l}{0.5\columnwidth}
+%    	\includegraphics{figures/deflection.pdf}
+%	\caption{\revise{Te maximum beam deflection, as a percentage of the FEM beam's maximum deflection, is shown to converge to the FEM result as we add more parts along the beam.}}
+%	  \label{fig:plot_beam_deflection}
+%\end{wrapfigure}
+\begin{figure}[h]
+  \includegraphics[width=\columnwidth]{figures/deflection_full.pdf}
+	\caption{\revise{Te maximum beam deflection, as a percentage of the FEM beam's maximum deflection, is shown to converge to the FEM result as we add more parts along the beam.}}
+  \label{fig:plot_beam_deflection}
+\end{figure}
+
 We use SEM with quadratic polynomials for this test and observe that our SEM simulation, made of 24 independent NURBS patches, shows good agreement with FEM.
@@
 
+\begin{figure}[h]
+  \includegraphics[width=\columnwidth]{figures/octopus.pdf}
+  \caption{\revise{As a more extreme example of the shape-aware property of our algorithm, we simulate a falling octopus, and observe the 8 limbs moving independently.
+Self-collision is not accounted for here.}}
+  \label{fig:octopus}
+\end{figure}
+
+\begin{figure}[h]
+  \includegraphics[width=\columnwidth]{figures/beam_twist.pdf}
+  \caption{\revise{With the left end of the beam held fixed, we show the result of rotating the right end of the 22 part beam up to 360 degrees.}}
+  \label{fig:twist}
+\end{figure}
+
 In \reffig{teaser} we directly simulate the NURBS surface model of a bouncy castle under a periodic wind force. 

### 2021-04-07 paper/sections/intro.tex
fixed typos

@@
 
-NURBS (Non-uniform Rational B Splines) are a popular higher-order modelling primitive which 
+NURBS (Non-uniform Rational \revise{B-Splines}) are a popular higher-order modelling primitive which 
 are used for computer-aided design (CAD), computational fabrication and computer animation. 

### 2021-04-07 paper/sections/methods.tex
fixed typos

@@
 basis functions given by 
+\revise{
 \begin{equation*}
-    \phi_{ij}\left(u,v\right) = \frac{\omega_{i,j}\beta_{ik}(u)\beta_{jl}(v)}{\sum_{r=1}^{n}\sum_{s=1}^{m} \omega_{i,j} \beta_{ik}(u)\beta_{jl}(v)},
-\end{equation*} where $\beta_{ik}$ (resp. $\beta_{jl}$) is the $i^{th}$ ($j^{th}$) B-spline basis of order $k$ ($l$) and $\omega_{i,j}$ is the NURBS weight.
+    \phi_{ij}\left(u,v\right) = \frac{\omega_{i,j}\beta_{ik}(u)\beta_{jl}(v)}{\sum_{r=1}^{n}\sum_{s=1}^{m} \omega_{r,s} \beta_{rk}(u)\beta_{sl}(v)},
+\end{equation*}} where $\beta_{ik}$ (resp. $\beta_{jl}$) is the $i^{th}$ ($j^{th}$) B-spline basis of order $k$ ($l$) and $\omega_{i,j}$ is the NURBS weight.
 
@@
 We can define a polynomial deformation from $\refX = \begin{pmatrix} X & Y & Z \end{pmatrix}^T$ to $\defX$  as 
+\revise{
 \begin{equation}
@@
     \underbrace{\begin{bmatrix}
-    \mathbf{p}^T(\refX-\bar{\refX}) & \mathbf{0} & \mathbf{0} & 1 & 0 & 0\\
-    \mathbf{0} & \mathbf{p}^T(\refX-\bar{\refX}) & \mathbf{0} & 0 & 1 & 0\\
-    \mathbf{0} & \mathbf{0} & \mathbf{p}^T(\refX-\bar{\refX}) & 0 & 0 & 1\\
+    \mathbf{p}(\refX-\bar{\refX})^T & \mathbf{0} & \mathbf{0} & 1 & 0 & 0\\
+    \mathbf{0} & \mathbf{p}(\refX-\bar{\refX})^T & \mathbf{0} & 0 & 1 & 0\\
+    \mathbf{0} & \mathbf{0} & \mathbf{p}(\refX-\bar{\refX})^T & 0 & 0 & 1\\
     \end{bmatrix}}_{\Pmatx}
@@
     \label{eq:projection}
-\end{equation} where $\mathbf{p}^T\left(\refX\right)$ is the polynomial basis vector \newline (e.g., $\begin{bmatrix} X & Y & Z & X\cdot Y & X\cdot Z & Y\cdot Z& X\cdot X & Y\cdot Y & Z\cdot Z \end{bmatrix}$ ),
+\end{equation} where $\mathbf{p}\left(\refX\right)^T$} is the polynomial basis vector \newline (e.g., $\begin{bmatrix} X & Y & Z & X\cdot Y & X\cdot Z & Y\cdot Z& X\cdot X & Y\cdot Y & Z\cdot Z \end{bmatrix}$ ),
 $\bar{\refX}$ is an a priori chosen origin around which to deform, $\Pco$ are the coefficients for the 
@@
 \paragraph*{Relationship to the Virtual Element Method}
-For a non-planar NURBS part, a sufficient number of quadrature points will make ~\refeq{shapematch} will be well-posed, allowing us to write: $\Pcox^* = \Pi\vc{q}$ where
+For a non-planar NURBS part, a sufficient number of quadrature points \revise{will make ~\refeq{shapematch} well-posed}, allowing us to write: $\Pcox^* = \Pi\vc{q}$ where
 \begin{equation*}
@@
 \newcommand{\Zeta}{\mathrm{Z}}
+
 \begin{equation}
-\int_{\Omega_R} g(X,Y,Z) d\Omega_R = \int_{\Gamma_R}\left(\int_{X_1}^{X_2} g(\alpha, Y, Z) d\alpha)\right)\mathbf{n} d\Gamma_R.
+\int_{\Omega_R} g(X,Y,Z) d\Omega_R = \int_{\Gamma_R}\left(\int_{X_1}^{X_2} \revise{g(\alpha, Y, Z) d\alpha}\right)\mathbf{n} d\Gamma_R.
 \end{equation}
+
 where $\Gamma_R$ is the boundary of the bounding box with normals $\mathbf{n}$ and $X_1$ and $X_2$ represent the minimum and maximum values of the box along the $x$-axis.
@@
     \defX\left(\refX\right) = \underbrace{\shapef\left(\refX\right)\bldmat\Pi}_{\Jkin\left(\refX\right)}\vc{q},
-\end{equation} where $\shapef\in\real^{3\times n} = \begin{bmatrix}P_1\left(\refX\right) \dots P_m\left(\refX\right)\end{bmatrix}$, $m$ is the number of 
-parts in the object, $\bldmat$ is the diagonal matrix of blending weights, $\Pi$ is the global projection operator and $\Jkin\left(\refX\right)$ is the kinematic Jacobian.
+\end{equation} where $\shapef\in\real^{3\times n} = \begin{bmatrix}P_1\left(\refX\right) \dots P_m\left(\refX\right)\end{bmatrix}$, \revise{$n$ is of size $3(km + |\XBs|)$ with $k$ being the number of terms in each polynomial, $m$ is the number of parts, and $|\XBs|$ is the number of deformation origins.
+\ty{previously said:$m$ is the number of
+parts in the object}} $\bldmat$ is the diagonal matrix of blending weights, $\Pi$ is the global projection operator and $\Jkin\left(\refX\right)$ is the kinematic Jacobian.
 

### 2021-04-07 paper/sections/results.tex
accidentally deleted fig label
fixed typos
update results and more typo fixes

@@
   \caption{Performance of, and parameters for the Shape Matching Element Method on all examples. All wall-clock timings are reported in seconds, physical parameters are
-  reported with appropriate units. $\rho$ is the applied density, \textbf{E} is the  Young's Modulus and $\mu$ is the Poisson's ratio.
+  reported with appropriate units. $\rho$ is the applied density, \textbf{E} is the  Young's Modulus and \revise{$\nu$} is the Poisson's ratio.
   \textbf{Sample} is the time taken to sample and construct $\samp{\Jnurbs}$ and $\samp{\Pmatx}$, \textbf{Quad} is the time taken to generate quadrature points via raycasting,
@@
   \begin{tabular}{l c c c c c c c c c c c}
-   \textbf{Example} & $|\vc{q}|$ & \textbf{Material}&  $\rho$(kg/m$^3$) & \textbf{E}(Pa) & $\mu$ &\textbf{Sample}(s) & \textbf{Quad}(s)& \textbf{Weights}(s) & \textbf{Build $\Pi$}(s) & \textbf{Step}(s) & \revise{\textbf{Nnz}(\%)} \\
+   \textbf{Example} & $|\vc{q}|$ & \textbf{Material}&  $\rho$(kg/m$^3$) & \textbf{E}(Pa) & \revise{$\nu$} &\textbf{Sample}(s) & \textbf{Quad}(s)& \textbf{Weights}(s) & \textbf{Build $\Pi$}(s) & \textbf{Step}(s) & \revise{\textbf{Nnz}(\%)} \\
    \hline 
@@
 We use quadratic deformation for all the animation results, except the rocket (stiff) and starship which use linear deformation.
-Models were created using Autodesk Fusion 360 \cite{AutodeskFusion360} and Rhinocerous 3D 7 \cite{mcneel-rhinoceros}. 
+Models were created using Autodesk Fusion 360 \cite{AutodeskFusion360} and \revise{Rhinoceros} 3D 7 \cite{mcneel-rhinoceros}. 
 Please see the supplemental video for more animation results.
@@
 The vertices of this triangulation serve as fixed set of high resolution UV samples for constructing a render mesh.
-
-\begin{figure}[h]
-  \includegraphics[width=\columnwidth]{figures/scalability.pdf}
-  \caption{\revise{As a more extreme example of the shape-aware property of our algorithm, we simulate a falling octopus, and observe the 8 limbs moving independently.}}
-  \label{fig:scalability}
-\end{figure}
 
@@
   Sparsity of our stiffness matrix is primarily determined by weight sparsity (number of non-zero weights at a given quadrature point) and polynomial order. }}
+  \label{fig:stiffness_sparsity}
 \end{figure}
@@
 
-We use SEM with quadratic polynomials for this test and observe that our SEM simulation, made of 24 independent NURBS patches, shows good agreement with FEM.
+We use SEM with quadratic polynomials for this test and observe that our SEM simulation, made of \revise{22} independent NURBS patches, shows good agreement with FEM.
 Each subdivision of the NURBS beam enables more complicated kinematics,
@@
   \label{fig:convergence}
+\end{figure}
+
+\revise{Additionally, we perform a twisting beam experiment on the 22 part beam model.
+Despite the relatively small number of NURBS patches, we see in \reffig{beam_twist} that the quadratic polynonmials permit complex deformation.}
+
+\begin{figure}[h]
+  \includegraphics[width=\columnwidth]{figures/beam_twist.pdf}
+  \caption{\revise{With the left end of the beam held fixed, we show the result of rotating the right end of the 22 part beam up to 360 degrees.}}
+  \label{fig:beam_twist}
+\end{figure}
+
+\revise{We perform a basic scalability test shown in \reffig{scalability} by iteratively adding more degrees of freedom to a model and measuring average runtime of a simulation step, and report the ratio of number of nonzeros in the stiffness matrix.
+We begin with a unit cube composed of 6 parts and on each iteration we extend the cube, adding 4 more parts.
+The raycasting quadrature is updated on each
+iteration so that a constant number of samples (45 in this case) lie in each cube. This experiment was performed using a AMD Ryzen 2600 with 16GB of RAM}
+
+\begin{figure}[h]
+  \includegraphics[width=\columnwidth]{figures/scalability.pdf}
+  \caption{\revise{As a more extreme example of the shape-aware property of our algorithm, we simulate a falling octopus, and observe the 8 limbs moving independently.}}
+  \label{fig:scalability}
 \end{figure}
@@
 This is similar to Discontinous Galerkin approaches but in SEM this behavior emerges from the kinematic model, rather than from additional flux terms~\cite{kaufmann2009flexible}.
+\revise{In \reffig{octopus} we demonstrate the shape-aware nature of our algorithm on a  stable Neo-Hookean ~\cite{SNHkim} simulation of an octopus, and observe independent motion for each limb.}
 
@@
   \label{fig:octopus}
-\end{figure}
-
-\begin{figure}[h]
-  \includegraphics[width=\columnwidth]{figures/beam_twist.pdf}
-  \caption{\revise{With the left end of the beam held fixed, we show the result of rotating the right end of the 22 part beam up to 360 degrees.}}
-  \label{fig:twist}
 \end{figure}

### 2021-04-07 examples/beam_twist.m
update beam example as well as scalability and deflection experiments

@@ -5,10 +5,7 @@ part = nurbs_from_iges(iges_file);
 
 % cutoff_distance=cutoff_heuristic(part, 0.9,40)
 
-
 %material properties
-% YM = 5e3; %in Pascals
-YM = 5e6; %in Pascals
 YM = 1e6; %in Pascals
 pr =  0.45;
 [lambda, mu] = emu_to_lame(YM, pr);
@@ -20,26 +17,15 @@ options.gravity = 0;
 options.lambda = lambda;
 options.mu = mu;
 options.rho = 1e3;
-% options.distance_cutoff = 1;
-% options.distance_cutoff = 0.6;
 options.distance_cutoff = 0.51;
-% options.k_stability = 1e6;
 options.k_stability = 2e6;
 options.enable_secondary_rays = 1;
 options.fitting_mode = 'hierarchical';
 options.save_output = 0;
-options.save_obj = 1;
+options.save_obj = 0;
 % options.plot_points = 1;
-options.x_samples = 25;
-options.y_samples = 3;
-options.z_samples = 3;
-% options.x_samples = 50;
-% options.y_samples = 1;
-% options.z_samples = 1;
 options.x_samples = 15;
 options.y_samples = 3;
 options.z_samples = 3;
-% vem_simulate_nurbs_newtons(part, options);
 simulate_twist(part, options);
-
 end

### 2021-04-07 figures/deflection_test.m

@@ -1,0 +1,52 @@
+function deflection_test
+
+    
+% Read in cube mesh 
+files={'b1.igs', 'b2.igs', 'b3.igs', 'b4.igs', 'b5.igs', 'b6.igs','b7.igs','b8.igs','b9.igs','b10.igs'};%, ...
+      % 'b7.igs', 'b8.igs'};%, 'b9.igs', 'b10.igs'};%, 'b20.igs'};
+n = numel(files);
+dofs=zeros(n,1);
+nparts=zeros(n,1);
+def=zeros(n,1);
+
+YM = 1e6; %in Pascals
+pr =  0.45;
+[lambda, mu] = emu_to_lame(YM, pr);
+
+options.dt = 0.1;
+options.order = 2;
+options.pin_function = @pin_function;
+options.gravity = -10;
+options.lambda = lambda;
+options.mu = mu;
+options.rho = 1e3;
+options.distance_cutoff = 0.501;
+% options.distance_cutoff = 1;
+options.k_stability = 1e7;
+options.enable_secondary_rays = 1;
+options.fitting_mode = 'hierarchical';
+options.save_output = 0;
+options.save_obj = 0;
+options.plot_com=1;
+% options.plot_points = 1;
+options.x_samples = 100;
+options.y_samples = 3;
+options.z_samples = 3;
+
+iters=20;
+for i=1:n
+    parts = nurbs_from_iges(files{i});
+    dofs(i) = numel(parts{1}.p)*numel(parts);
+    [def(i),deflections] = deflection_test_step(parts,options);
+    nparts(i) = numel(parts);
+    figure(3);
+    plot(1:iters, deflections);
+    hold on;
+end
+figure(4);
+plot(nparts, abs(def ./ -4.1579),'LineWidth',2);
+xlabel('DOFs');
+ylabel('m');
+title('Max Deflection');
+% yline(-4.1579,'Color','r');
+end

### 2021-04-07 figures/deflection_test_step.m

@@ -1,10 +1,10 @@
-function [pcnt,time1,time2]=scalability_test_step(parts,varargin)
+function [max_def,deflections]=deflection_test_step(parts,varargin)
     % Simulation parameter parsing
     p = inputParser;
     addParameter(p, 'dt', 0.01);                % timestep
     addParameter(p, 'lambda', 0.5 * 1700);    	% Lame parameter 1
     addParameter(p, 'mu', 0.5 * 15000);        	% Lame parameter 2
-    addParameter(p, 'gravity', -10);          	% gravity force (direction is -z direction)
+    addParameter(p, 'gravity', -300);          	% gravity force (direction is -z direction)
     addParameter(p, 'k_stability', 1e5);       	% stiffness for stability term
     addParameter(p, 'order', 1);              	% (1 or 2) linear or quadratic deformation
     addParameter(p, 'rho', 1);                 	% per point density (currently constant)
@@ -14,7 +14,7 @@ function [pcnt,time1,time2]=scalability_test_step(parts,varargin)
     addParameter(p, 'save_resultion', 20);    	% the amount of subdivision for the output obj
     addParameter(p, 'pin_function', @(x) 1);
     addParameter(p, 'sample_interior', 1);
-    addParameter(p, 'distance_cutoff', 0.51);
+    addParameter(p, 'distance_cutoff', 20);
     addParameter(p, 'enable_secondary_rays', true);
     addParameter(p, 'fitting_mode', 'hierarchical');
     addParameter(p, 'plot_points', false);
@@ -29,13 +29,15 @@ function [pcnt,time1,time2]=scalability_test_step(parts,varargin)
     
     parse(p,varargin{:});
     config = p.Results;
-    pcnt=0;
-    tic
-
+    
     d = 3;              % dimension (2 or 3)
     n = numel(parts);	% number of shapes
     k = basis_size(d, config.order);
     
+    fig=figure(2);
+    clf;
+    parts=nurbs_plot(parts);
+
     % Assembles global generalized coordinates
     [J, ~, q, E, x0, S] = nurbs_assemble_coords(parts);
     
@@ -43,7 +45,18 @@ function [pcnt,time1,time2]=scalability_test_step(parts,varargin)
     x = x0;
[…]

### 2021-04-07 figures/scalability_test.m

@@ -6,22 +6,42 @@ fig=figure(2);
 iges_file = 'cube.iges';
 parts = nurbs_from_iges(iges_file);
 
+
+YM = 1e6; %in Pascals
+pr =  0.45;
+[lambda, mu] = emu_to_lame(YM, pr);
+
+
 n=50;
 nnz=zeros(n,1);
-times=zeros(n,1);
+dofs=zeros(n,1);
+time1=zeros(n,1);
+time2=zeros(n,1);
+substeps=1;
 for i=1:n
     clf;
     parts=nurbs_plot(parts);
-    
     options.x_samples = 5*i;
     options.y_samples = 3;
     options.z_samples = 3;
-    options.distance_cutoff = 0.51;
+    options.gravity = -10;
+    options.lambda = lambda;
+    options.mu = mu;
+    options.rho = 1e3;
+    options.distance_cutoff = 0.501;
 
-    [nnz(i), times(i)] =scalability_test_step(parts,options)
+    options.order = 2; % shouldn't matter, right?
+    options.distance_cutoff = 0.51;
+    dofs(i) = numel(parts{1}.p)*numel(parts);
+    for j=1:substeps
+        [nnz(i), t1, t2] = scalability_test_step(parts,options);
+        time1(i) = time1(i) + t1/substeps;
+        time2(i) = time2(i) + t2/substeps;
+    end
     drawnow
-    fn=sprintf('output/img/scalability_%03d.png',i);
-    saveas(fig,fn);
+
+%     fn=sprintf('output/img/scalability_%03d.png',i);
+%     saveas(fig,fn);
             
     % Push end farther, and duplicate 4 parts
     parts{1}.p(1:3:end) = parts{1}.p(1:3:end) + 1;
@@ -37,13 +57,25 @@ for i=1:n
     end
 end
 figure(3);
-plot(1:n, nnz,'LineWidth',2);
-xlabel('# cubes');
+plot(dofs, nnz,'LineWidth',2);
+xlabel('DOFs');
 ylabel('%');
 title('Percent nnz');
 figure(4);
-plot(1:n, times,'LineWidth',2);
-xlabel('# cubes');
+% plot(1:n, times,'LineWidth',2);
+% area(dofs, [time1 time2],'LineWidth',2, 'EdgeAlpha',0.0);
+plot(dofs, time1,'LineWidth',2);
+xlabel('DOFs');
+ylabel('time (s)');
+title('Total time in seconds');
[…]

### 2021-04-07 matlab/simulate_twist.m

@@ -136,10 +136,12 @@ function simulate_twist(parts, varargin)
     f_external = config.dt*P*(L' * dext_dc); 
 
     % Compute mass matrices
-    ME = vem_error_matrix(Y0, L, w0_I, C0_I, d, k, n);
-    M = vem_mass_matrix(Y, L, config.rho .* vol, w_I, C_I, d, k, n);
-    
-    M = (M + config.k_stability*ME); % sparse?
+%     ME = vem_error_matrix(Y0, L, w0_I, C0_I, d, k, n);
+%     M = vem_mass_matrix(Y, L, config.rho .* vol, w_I, C_I, d, k, n);
+    % make sure matlab/deprecated is on your path to use these versions
+    M = vem_mass_matrix_matlab(Y, Y_S, L, config.rho .* vol);
+    ME = vem_error_matrix_matlab(Y0, Y0_S, L, d);
+    M = (M + config.k_stability*ME);
 
     ii=1;
     img_ii=1;
@@ -210,12 +212,12 @@ function simulate_twist(parts, varargin)
 %         plt.Vertices=x';
 %         plt_AO=apply_ambient_occlusion(plt,'AddLights',false,'AO',plt_AO);
 
-%         for i=1:numel(parts)
-%             x_sz = size(parts{i}.x0,2);
-%             xi = x(:,x_idx+1:x_idx+x_sz);
-%             parts{i}.plt.Vertices =xi';
-%             x_idx = x_idx+x_sz;
-%         end
+        for i=1:numel(parts)
+            x_sz = size(parts{i}.x0,2);
+            xi = x(:,x_idx+1:x_idx+x_sz);
+            parts{i}.plt.Vertices =xi';
+            x_idx = x_idx+x_sz;
+        end
 %         
 %         if config.plot_com
 %             x_coms = c(d*k*n + 1:end); % extract centers of mass
@@ -229,7 +231,7 @@ function simulate_twist(parts, varargin)
 %             V_plot.YData = V(2,:);
 %             V_plot.ZData = V(3,:);
 %         end
-%         drawnow
+        drawnow
         
         if config.save_obj && mod(ii,4)==0
             obj_fn = config.save_obj_path + "part_" + int2str(obj_ii) + ".obj";

### 2021-04-07 figures/scalability_test.m
blah blah blah

@@ -73,9 +73,9 @@ plot(dofs, time2,'LineWidth',2);
 xlabel('DOFs');
 ylabel('time (s)');
 title('Total time in seconds');
-figure(6)
-area(dofs, [time1 time2],'LineWidth',2, 'EdgeAlpha',0.0);
-xlabel('DOFs');
-ylabel('time (s)');
-title('Total time in seconds');
+% figure(6)
+% area(dofs, [time1 time2],'LineWidth',2, 'EdgeAlpha',0.0);
+% xlabel('DOFs');
+% ylabel('time (s)');
+% title('Total time in seconds');
 end

### 2021-04-08 paper/sections/conclusion.tex
update captions and figure descriptions

@@
 (to better handle heterogenous materials) or to be robust to noisy data (allowing direct simulation of scanned data).
+\revise{Additionally, more effort is required to ensure that the shape matching fit is optimal, which we do not provide guarantees for with our hierarchical shape matching approach.
+In future work we intend to explore using robust solvers or orthogonal bases for the fitting step.} \ty{more can be said here I suppose.
+I tried pinv() on the beam, which surprisingly worked, but it's just unpredictable since it should still fail on my very basic 2D box example.
+Haven't done much experimentation beyond this though, so I just write future work}
 It is also interesting to consider the relationship between our error term and the orthogonality constraint proposed by~\citet{Zhang:CompDynamics:2020}. 

### 2021-04-08 paper/sections/results.tex
update captions and figure descriptions

@@
   \includegraphics[width=\columnwidth]{figures/deflection_full.pdf}
-	\caption{\revise{Te maximum beam deflection, as a percentage of the FEM beam's maximum deflection, is shown to converge to the FEM result as we add more parts along the beam.}}
+	\caption{\revise{Te maximum beam deflection, as a percentage of the FEM beam's maximum deflection, is shown to converge to the FEM result as we subdivide the beam.\ty{currently a very blank figure, which I dont mind, but could consider adding inset beam pictures to give an idea of how this experiment works}}}
   \label{fig:plot_beam_deflection}
@@
 We begin with a unit cube composed of 6 parts and on each iteration we extend the cube, adding 4 more parts.
-The raycasting quadrature is updated on each
-iteration so that a constant number of samples (45 in this case) lie in each cube. This experiment was performed using a AMD Ryzen 2600 with 16GB of RAM}
+The raycasting quadrature is updated on each iteration so that a constant number of samples (45 in this case) lie in each cube.
+This represents a best case scenario in which sparsity is maximized as the number of degrees of freedom increases.
+As performance is tied to sparsity of the stiffness matrix, we see reasonable growth in the reported times, despite minimal effort on optimizing performance.
+This experiment was performed using a AMD Ryzen 2600 with 16GB of RAM}
 
@@
   \includegraphics[width=\columnwidth]{figures/scalability.pdf}
-  \caption{\revise{As a more extreme example of the shape-aware property of our algorithm, we simulate a falling octopus, and observe the 8 limbs moving independently.}}
+  \caption{\revise{Here we progressively add more degrees of freedom to a beam by iteratively inserting 4 part sections, extending the beam in the $x$-axis direction at each step.
+On the left we report the number of non-zeros as a percentage of the total number of elements in the stiffness matrix (nnz$\%$)}.
+The right plot shows the average runtime of a simulation step as the number of degrees of freedom increases.}
   \label{fig:scalability}

### 2021-04-08 paper/sections/results.tex
small discussion on discontinuities

@@
 This is similar to Discontinous Galerkin approaches but in SEM this behavior emerges from the kinematic model, rather than from additional flux terms~\cite{kaufmann2009flexible}.
+\revise{Our soft enforcement of part connectivity at boundaries is a result of weighting function attributing equal weight values to equidistant parts.
+This equal weighting induces elastic forces that attempt to enforce continuity.
+In our examples this only introduces minor artifacts affecting the rendering, but this may be addressed using a separate rendering mesh or by using a representation with explicit connectivity.}
 \revise{In \reffig{octopus} we demonstrate the shape-aware nature of our algorithm on a  stable Neo-Hookean ~\cite{SNHkim} simulation of an octopus, and observe independent motion for each limb.}

### 2021-04-08 paper/sections/related.tex
oh boy that was a rough paragraph...

@@
 which increases the complexity.
+
+\revise{
+\emph{Embedded Methods} are a popular technique in computer graphics in which complex geometry is embedded in a coarse mesh that guides the simulation.
+The early work of ~\cite{10.5555/1009379.1009573} proposed a method for volumetric simulation of surface meshes by embedding the surface in a background cubic mesh may be seen as an alternative strategy to ours.
+However, producing an embedded mesh introduces a number of challenges since the coarse mesh's topology must mirror that of the embedded surface, which necessitates complex hierarchical methods like that of ~\cite{10.1145/1531326.1531358}.
+Also like in IGA methods, many methods rely on cut cells to conform to the boundary, requiring complicated geometric operations.
+Furthermore when the DOFs do not lie on the boundary, complex methods such as Nitsche's Method ~\cite{Nitsche1971berEV} are required for handling Dirichlet boundary conditions.
+By making use of an entirely meshless pipeline and operating directly on the input curves, we avoid much of these difficulties in embedded methods.
+}
 

### 2021-04-08 paper/sections/results.tex
add beam twist result
update scalability time plot

@@
    \textbf{Starship}                & 750 & Steel & $8 e^3$ & $2 e^{11}$ & 0.32 & $6.78 e^{-2}$ & $1.52 e^{-2}$ & $8.50 e^{-1}$ & $9.52 e^{-2}$ & $6.95 e^{-1}$ & \revise{100} \\
+   \revise{\textbf{Beam (twist)}}   & \revise{756} & \revise{Rubber} & \revise{$1 e^3$} & \revise{$1 e^6$} & \revise{0.45} & \revise{$1.97 e^{-1}$} & \revise{$2.25 e^{-2}$} & \revise{$1.35 e^{0}$} & \revise{$5.16 e^{-1}$} & \revise{$3.95 e^{0}$} & \revise{46.94} \\
+   \rowcolor[HTML]{DAE8FC} 
    \textbf{Lamppost}                & 942 & Rubber & $3 e^3$ & $1 e^6$ & 0.40 & $1.28 e^{-1}$ & $1.27 e^{-2}$ & $1.89 e^{0}$ & $1.07 e^{0}$ & $2.22 e^{0}$ & \revise{56.38} \\
+   \textbf{Tire}                    & 1404 & Rubber/Steel & $2 e^3$/$3 e^3$ & $1 e^6$/$7 e^9$ & 0.47/0.35 & $1.74 e^{-1}$ & $1.32 e^{-2}$ & $1.82 e^{0}$ & $6.79 e^{-1}$ & $1.28 e^{1}$ & \revise{92.5} \\
    \rowcolor[HTML]{DAE8FC} 
-   \textbf{Tire}                    & 1404 & Rubber/Steel & $2 e^3$/$3 e^3$ & $1 e^6$/$7 e^9$ & 0.47/0.35 & $1.74 e^{-1}$ & $1.32 e^{-2}$ & $1.82 e^{0}$ & $6.79 e^{-1}$ & $1.28 e^{1}$ & \revise{92.5} \\
    \textbf{Chicken}                 & 1773 & Jelly & $1.27 e^3$ & $1 e^4$ & 0.47 & $9.19 e^{-1}$ & $2.31 e^{-2}$ & $3.38 e^{0}$ & $1.91 e^{0}$ & $1.08 e^{1}$ & \revise{100} \\
+   \textbf{Coffee mug (soft)}       & 1800 & Jelly & $1 e^3$ & $1 e^4$ & 0.47 & $4.18 e^{-1}$ & $2.87 e^{-2}$ & $3.77 e^{0}$ & $2.08 e^{0}$ & $1.19 e^{1}$ & \revise{100} \\ 
    \rowcolor[HTML]{DAE8FC} 
-   \textbf{Coffee mug (soft)}       & 1800 & Jelly & $1 e^3$ & $1 e^4$ & 0.47 & $4.18 e^{-1}$ & $2.87 e^{-2}$ & $3.77 e^{0}$ & $2.08 e^{0}$ & $1.19 e^{1}$ & \revise{100} \\ 
    \textbf{Coffee mug (stiff)}      & 1800 & Rubber & $1 e^3$ & $4 e^7$ & 0.40 & $4.15 e^{-2}$ & $1.17 e^{-2}$ & $6.86 e^{-1}$ & $1.09 e^{-1}$ & $5.00 e^{-1}$  & \revise{81.42} \\ 
+   \textbf{Castle}                  & 1872 & Jelly & $1.27 e^3$ & $2 e^3$ & 0.45 & $4.84 e^{-1}$ & $1.44 e^{-2}$ & $1.52 e^{0}$ & $5.78 e^{-1}$ & $8.02 e^{0}$ & \revise{100} \\
    \rowcolor[HTML]{DAE8FC} 
-   \textbf{Castle}                  & 1872 & Jelly & $1.27 e^3$ & $2 e^3$ & 0.45 & $4.84 e^{-1}$ & $1.44 e^{-2}$ & $1.52 e^{0}$ & $5.78 e^{-1}$ & $8.02 e^{0}$ & \revise{100} \\
    \textbf{Grumpy}                  & 2010 & Jelly & $1 e^3$ & $1 e^4$ & 0.40 & $2.92 e^{-1}$ & $1.23 e^{-2}$ & $2.29 e^{0}$ & $2.89 e^{0}$ & $6.02 e^{0}$ & \revise{84.15} \\    
+   \textbf{Astronaut (soft)}        & 3609 & Rubber &  $1.27 e^3$  & $1 e^6$  & 0.47 & $2.47 e^{0}$ & $6.71 e^{-2}$ & $1.13 e^{1}$ & $1.09 e^{1}$ & $1.41 e^{1}$ & \revise{49.04} \\
    \rowcolor[HTML]{DAE8FC} 
-   \textbf{Astronaut (soft)}        & 3609 & Rubber &  $1.27 e^3$  & $1 e^6$  & 0.47 & $2.47 e^{0}$ & $6.71 e^{-2}$ & $1.13 e^{1}$ & $1.09 e^{1}$ & $1.41 e^{1}$ & \revise{49.04} \\
    \revise{\textbf{Squid}}          & \revise{4347} & \revise{Jelly} & \revise{ $1 e^3$ } & \revise{$2 e^4$ } & \revise{0.40} & \revise{$4.48 e^{-1}$} & \revise{$2.47 e^{-2}$} & \revise{$1.54 e^{1}$} & \revise{$4.48 e^{1}$} & \revise{$6.73 e^{0}$} & \revise{25.47} \\
@@
 This represents a best case scenario in which sparsity is maximized as the number of degrees of freedom increases.
-As performance is tied to sparsity of the stiffness matrix, we see reasonable growth in the reported times, despite minimal effort on optimizing performance.
-This experiment was performed using a AMD Ryzen 2600 with 16GB of RAM}
+As performance is tied to sparsity of the stiffness matrix, we see reasonable growth in the reported times, despite minimal effort on optimizing performance.}
 
@@
   \caption{\revise{Here we progressively add more degrees of freedom to a beam by iteratively inserting 4 part sections, extending the beam in the $x$-axis direction at each step.
-On the left we report the number of non-zeros as a percentage of the total number of elements in the stiffness matrix (nnz$\%$)}.
-The right plot shows the average runtime of a simulation step as the number of degrees of freedom increases.}
+On the left we report the number of non-zeros as a percentage of the total number of elements in the stiffness matrix (nnz$\%$).
+The right plot shows the average runtime of a simulation step as the number of degrees of freedom increases.}}
   \label{fig:scalability}

### 2021-04-08 paper/sections/conclusion.tex
Dave Changes

@@
 (to better handle heterogenous materials) or to be robust to noisy data (allowing direct simulation of scanned data).
-\revise{Additionally, more effort is required to ensure that the shape matching fit is optimal, which we do not provide guarantees for with our hierarchical shape matching approach.
-In future work we intend to explore using robust solvers or orthogonal bases for the fitting step.} \ty{more can be said here I suppose.
+\revise{In future work we also intend to explore using robust solvers or orthogonal bases for the shape matching step itself.} \ty{more can be said here I suppose.
 I tried pinv() on the beam, which surprisingly worked, but it's just unpredictable since it should still fail on my very basic 2D box example.

### 2021-04-08 paper/sections/methods.tex
update results and more typo fixes
Dave Changes

@@
 
-\dave{should have done Chebyshev shape matching to ensure higher order terms are orthogonal which would gaurantee exact reconstruction}
+\revise{Note that hierarchical is not guaranteed to find the optimal polynomial fit and potentially risks overfitting to low-order terms.
+        However we find that it provides qualitatively resonable results and avoids much of the dense matrix arithmatic and singular value decompositions 
+        of alternative methods such as the Moore-Penorse inverse or QR factorization.}
 
@@
 \text{,}
+\label{eq:RELU}
 \end{equation}
@@
 but allows us to use more general weighting functions and apply the appropriate behavior for intersecting parts.
-The price we pay is that our weights are not $C^0$ continuous but rather Lipschitz smooth.
-Perturbations in the boundaries of parts can cause bounded jumps in our distance weights. 
+The price we pay is that our weights are not $C^0$ continuous but rather contain bounded jumps. 
 However, these jumps occupy an infinitesimal percentage of the volume of our object, and are easy to avoid when integrating
@@
 %\ty{we use the default in intlinprog, which should be dual-simplex} We choose $\XBs$ to be all the deformation origin candidates in $A$ that have an outgoing flow of at least $1$.
+
+\revise{Note that this approach yields compactly supported weight functions due to the cut-off parameter~(Eq.~8). 
+Depending on this value, we obtain a range of system matrix sparsity patterns -- from fully dense to block sparse~(Fig.~9). 
+Importantly, unlike other boundary-only approaches, our method is not guaranteed to produce dense system matrices.}
 
@@
 \begin{equation}
-    T = \frac{1}{2}\dot{\vc{q}}^T\underbrace{\Pi^T\int_{\Omega}\bldmat\left(\refX\right)^T\shapef\left(\refX\right)^T\shapef\left(\refX\right)\bldmat\left(\refX\right)d\Omega\Pi}_{\MM}\dot{\vc{q}},
+    T = \frac{1}{2}\dot{\vc{q}}^T\underbrace{\Pi^T\int_{\Omega} \revise{\rho}\bldmat\left(\refX\right)^T\shapef\left(\refX\right)^T\shapef\left(\refX\right)\bldmat\left(\refX\right)d\Omega\Pi}_{\MM}\dot{\vc{q}}.
 \end{equation}
+
 % \begin{equation}

### 2021-04-08 paper/sections/related.tex
Dave Changes

--- 2021-01-28 paper/sections/related.tex
@@
 which increases the complexity.
+
+\revise{
+\emph{Embedded Methods} attempt to sidestep many of these issues by enclosing complex surface geometry inside a simulation coarse mesh~\cite{10.5555/1009379.1009573}. 
+However, producing an appropriate mesh embedding introduces a number of challenges since the coarse mesh's connectivity must mirror that of the embedded surface.
+This necessitates the use of complex hierarchical methods like that of~\citet{10.1145/1531326.1531358}, which themselves require the user to correctly intuit an appropriate ultimate grid resolution.
+Alternately one can introduce cut cell~\cite{Tao:Mandoline:2019} or extended Finite Element~\cite{10.1145/3355089.3356576} approaches which bring with them complicated geometric operations. 
+Furthermore when the DOFs do not lie on the boundary, methods such as Nitsche's Method ~\cite{Nitsche1971berEV} are required for handling Dirichlet boundary conditions. 
+Our method inherits shape connectivity directly from the boundary input, avoiding these difficulties entirely.
+}
 

### 2021-04-08 paper/sections/results.tex
Dave Changes

--- 2021-04-08 paper/sections/results.tex
@@
 
-\revise{We perform a basic scalability test shown in \reffig{scalability} by iteratively adding more degrees of freedom to a model and measuring average runtime of a simulation step, and report the ratio of number of nonzeros in the stiffness matrix.
-We begin with a unit cube composed of 6 parts and on each iteration we extend the cube, adding 4 more parts.
-The raycasting quadrature is updated on each iteration so that a constant number of samples (45 in this case) lie in each cube.
-This represents a best case scenario in which sparsity is maximized as the number of degrees of freedom increases.
-As performance is tied to sparsity of the stiffness matrix, we see reasonable growth in the reported times, despite minimal effort on optimizing performance.
+\revise{\reffig{scalability} shows a scalability study in which a  unit cube, composed of 6 parts, is extended by appending 4 more parts at a time.
+Note that our algorithm scales similarly to standard high-order finite element approaches due to the compact support of our weight functions. 
 This experiment was performed using a AMD Ryzen 2600 with 16GB of RAM}

### 2021-04-08 paper/sections/results.tex

@@
 
-We also show that our raycasting weight computation is able to create shape-aware output. \reffig{independence} shows that manipulating
+We also show that our raycasting weight computation is able to create shape-aware output. \reffig{independence} and \reffig{octopus} show that manipulating
 parts that are nearby but separated will behave in an appropriately independent fashion. Our grumpy model is capable of extending 
-its leg without causing unrealistic deformations in the plant foot.
+its leg without causing unrealistic deformations in the plant foot, and our octopus model shows independent motion of all eight limbs upon contact with the ground.
 
@@
   \label{fig:independence}
+\end{figure}
+
+\begin{figure}[h]
+  \includegraphics[width=\columnwidth]{figures/octopus.pdf}
+  \caption{\revise{We simulate a falling octopus, and observe the 8 limbs moving independently. Self-collision is not accounted for here.}}
+  \label{fig:octopus}
 \end{figure}
@@
 In our examples this only introduces minor artifacts affecting the rendering, but this may be addressed using a separate rendering mesh or by using a representation with explicit connectivity.}
-\revise{In \reffig{octopus} we demonstrate the shape-aware nature of our algorithm on a  stable Neo-Hookean ~\cite{SNHkim} simulation of an octopus, and observe independent motion for each limb.}
-
-\begin{figure}[h]
-  \includegraphics[width=\columnwidth]{figures/octopus.pdf}
-  \caption{\revise{As a more extreme example of the shape-aware property of our algorithm, we simulate a falling octopus, and observe the 8 limbs moving independently.
-Self-collision is not accounted for here.}}
-  \label{fig:octopus}
-\end{figure}
 

### 2021-04-08 paper/sections/methods.tex
typos

@@
 
-\revise{Note that hierarchical is not guaranteed to find the optimal polynomial fit and potentially risks overfitting to low-order terms.
-        However we find that it provides qualitatively resonable results and avoids much of the dense matrix arithmatic and singular value decompositions 
-        of alternative methods such as the Moore-Penorse inverse or QR factorization.}
+\revise{Note that hierarchical shape matching is not guaranteed to find the optimal polynomial fit and potentially risks overfitting to low-order terms.
+        However we find that it provides qualitatively reasonable results and avoids much of the dense matrix arithmetic and singular value decompositions 
+        of alternative methods such as the Moore-Penrose inverse or QR factorization.}
 

### 2021-04-08 paper/sections/results.tex
typos

@@
 In \reffig{convergence}, we compare the static configurations of identical cantilevered beams, simulated using a Neohookean constitutive model (Young's Modulus $=$ $0.001$ GPa, Poisson's Ratio $=$ $0.45$).
-\revise{We further perform basic \emph{quantitative} comparison against FEM by measuring the maximum deflection of the beam as we increase the number of patches along the beam.
+\revise{We further perform a basic \emph{quantitative} comparison against FEM by measuring the maximum deflection of the beam as we increase the number of patches along the beam.
 In \reffig{plot_beam_deflection}, the maximum beam deflection of the SEM beams is reported as a percentage of the FEM beam's maximum deflection.
@@
 \revise{\reffig{scalability} shows a scalability study in which a  unit cube, composed of 6 parts, is extended by appending 4 more parts at a time.
-Note that our algorithm scales similarly to standard high-order finite element approaches due to the compact support of our weight functions. 
-This experiment was performed using a AMD Ryzen 2600 with 16GB of RAM}
+Note that our algorithm scales similarly to standard high-order finite element approaches due to the compact support of our weight functions.}
 
@@
 This is similar to Discontinous Galerkin approaches but in SEM this behavior emerges from the kinematic model, rather than from additional flux terms~\cite{kaufmann2009flexible}.
-\revise{Our soft enforcement of part connectivity at boundaries is a result of weighting function attributing equal weight values to equidistant parts.
+\revise{Our soft enforcement of part connectivity at boundaries is a result of the weighting function attributing equal weight values to equidistant parts.
 This equal weighting induces elastic forces that attempt to enforce continuity.
-In our examples this only introduces minor artifacts affecting the rendering, but this may be addressed using a separate rendering mesh or by using a representation with explicit connectivity.}
+In our examples, these discontinuities only introduce minor artifacts affecting the rendering, but this may be addressed using a separate render mesh or by using a representation with explicit connectivity.}
 

### 2021-04-08 paper/sections/results.tex
beam stuff updated

@@
 In \reffig{convergence}, we compare the static configurations of identical cantilevered beams, simulated using a Neohookean constitutive model (Young's Modulus $=$ $0.001$ GPa, Poisson's Ratio $=$ $0.45$).
-\revise{We further perform a basic \emph{quantitative} comparison against FEM by measuring the maximum deflection of the beam as we increase the number of patches along the beam.
-In \reffig{plot_beam_deflection}, the maximum beam deflection of the SEM beams is reported as a percentage of the FEM beam's maximum deflection.
-The parameters of the experiment match that of the previous qualitative test.}
+\revise{We further perform a basic \emph{quantitative} comparison against FEM by measuring the deflection of the beam at resting position as we increase the number of patches along the beam.
+In \reffig{plot_beam_deflection}, the beam deflection of the SEM beams is reported as a percentage of the FEM beam's deflection.}
 
@@
 
-We use SEM with quadratic polynomials for this test and observe that our SEM simulation, made of \revise{22} independent NURBS patches, shows good agreement with FEM.
+We use SEM with quadratic polynomials for this test and observe that our SEM simulation, made of \revise{26} independent NURBS patches, shows good agreement with FEM.
 Each subdivision of the NURBS beam enables more complicated kinematics,
@@
 
-\revise{Additionally, we perform a twisting beam experiment on the 22 part beam model.
+\revise{Additionally, we perform a twisting beam experiment on a 22 part beam model.
 Despite the relatively small number of NURBS patches, we see in \reffig{beam_twist} that the quadratic polynonmials permit complex deformation.}
@@
   \includegraphics[width=\columnwidth]{figures/beam_twist.pdf}
-  \caption{\revise{With the left end of the beam held fixed, we show the result of rotating the right end of the 22 part beam up to 360 degrees.}}
+  \caption{\revise{With the left end of the beam held fixed, we show the result of rotating the right end of a 22 part beam up to 360 degrees.}}
   \label{fig:beam_twist}

### 2021-04-08 paper/sections/results.tex
small figure updates

@@
 In \reffig{convergence}, we compare the static configurations of identical cantilevered beams, simulated using a Neohookean constitutive model (Young's Modulus $=$ $0.001$ GPa, Poisson's Ratio $=$ $0.45$).
-\revise{We further perform a basic \emph{quantitative} comparison against FEM by measuring the deflection of the beam at resting position as we increase the number of patches along the beam.
-In \reffig{plot_beam_deflection}, the beam deflection of the SEM beams is reported as a percentage of the FEM beam's deflection.}
+\revise{We further perform a basic \emph{quantitative} comparison against FEM by measuring the deflection of the beam at its resting position as we increase the number of patches along the beam.
+In \reffig{plot_beam_deflection}, the deflection of the SEM beams is reported as a percentage of the FEM beam's deflection.}
 
@@
   \includegraphics[width=\columnwidth]{figures/deflection_full.pdf}
-	\caption{\revise{Te maximum beam deflection, as a percentage of the FEM beam's maximum deflection, is shown to converge to the FEM result as we subdivide the beam.\ty{currently a very blank figure, which I dont mind, but could consider adding inset beam pictures to give an idea of how this experiment works}}}
+	\caption{\revise{The SEM beam deflection is plotted as a percentage of the FEM beam's deflection when the beam is at rest.
+As we subdivide the SEM beam, its deflection converges to the FEM beam's deflection.
+\ty{currently a very blank figure, which I dont mind, but could consider adding inset beam pictures to give an idea of how this experiment works}}}
   \label{fig:plot_beam_deflection}
@@
   \caption{\revise{Here we progressively add more degrees of freedom to a beam by iteratively inserting 4 part sections, extending the beam in the $x$-axis direction at each step.
-On the left we report the number of non-zeros as a percentage of the total number of elements in the stiffness matrix (nnz$\%$).
+On the left we report the number of non-zeros as a percentage of the total number of elements in the stiffness matrix (Nnz$\%$).
 The right plot shows the average runtime of a simulation step as the number of degrees of freedom increases.}}

### 2021-04-09 paper/sections/results.tex
shtuff

@@
 
+\begin{figure}[h]
+  \includegraphics[width=\columnwidth]{figures/stiffness_sparsity.pdf}
+  \caption{\revise{Our stiffness matrices can be sparse, while the corresponding matrix has to be dense in the traditional Boundary Element Method \cite{ArtDefo1999}. 
+  Sparsity of our stiffness matrix is primarily determined by weight sparsity (number of non-zero weights at a given quadrature point) and polynomial order. }}
+  \label{fig:stiffness_sparsity}
+\end{figure}
+
 For raycasting operations and rendering we rely on Embree~\cite{10.1145/2601097.2601199} and Blender~\cite{blender}, and triangulate the NURBS surfaces, which only represents an implementation simplification.
@@
   \label{fig:patchtest}
-\end{figure}
-
-\begin{figure}[h]
-  \includegraphics[width=\columnwidth]{figures/stiffness_sparsity.pdf}
-  \caption{\revise{Our stiffness matrices can be sparse, while the corresponding matrix has to be dense in the traditional Boundary Element Method \cite{ArtDefo1999}. 
-  Sparsity of our stiffness matrix is primarily determined by weight sparsity (number of non-zero weights at a given quadrature point) and polynomial order. }}
-  \label{fig:stiffness_sparsity}
 \end{figure}
@@
 
-We also show that our raycasting weight computation is able to create shape-aware output. \reffig{independence} and \reffig{octopus} show that manipulating
+We also show that our raycasting weight computation is able to create shape-aware output. \reffig{independence} and \revise{\reffig{octopus} show that manipulating
 parts that are nearby but separated will behave in an appropriately independent fashion. Our grumpy model is capable of extending 
-its leg without causing unrealistic deformations in the plant foot, and our octopus model shows independent motion of all eight limbs upon contact with the ground.
+its leg without causing unrealistic deformations in the plant foot, and our octopus model shows independent motion of all eight limbs upon contact with the ground.}
 

### 2021-04-09 paper/sections/methods.tex
beam stuff updated
fix typo
typo

@@
 We achieve this by sampling over each knot span defined by the  B-splines, and find it effective to use the order of the NURBS as the number of quadrature points sampled in each knot span.
+\revise{Positions and weights of the quadrature points in each span are chosen using Gauss-Lobatto quadrature.}
 \ty{the number of samples may be different if order in the u direction and v differ, but I figured it was best to keep this simple.}
@@
 \revise{Note that this approach yields compactly supported weight functions due to the cut-off parameter~(Eq.~8). 
-Depending on this value, we obtain a range of system matrix sparsity patterns -- from fully dense to block sparse~(Fig.~9). 
+Depending on this value, we obtain a range of system matrix sparsity patterns -- from fully dense to block sparse~(\reffig{stiffness_sparsity}). 
 Importantly, unlike other boundary-only approaches, our method is not guaranteed to produce dense system matrices.}

### 2021-04-27 paper/main.tex
add authors to main tex
Dave Changes
fixing typos

@@
-\documentclass[acmtog, anonymous,review, nonacm, balance = false]{acmart}
+\documentclass[acmtog, anonymous=false, review=false, balance=false]{acmart}
 %%
@@
 \input{sections/pre.tex}
+
+\setcopyright{acmlicensed}\acmJournal{TOG}
+\acmYear{2021}\acmVolume{40}\acmNumber{4}\acmArticle{69}\acmMonth{8} \acmDOI{10.1145/3450626.3459772}
 
@@
 \title{The Shape Matching Element Method: Direct Animation of Curved Surface Models}
+
+\author{Ty Trusty}
+
+\email{trusty@cs.toronto.edu}
+%
+\author{Honglin Chen}
+
+\email{chl9797@cs.toronto.edu}
+%
+%
+\author{David I.W. Levin}
+\email{diwlevin@cs.toronto.edu}
+
+\affiliation{%
+ \institution{University of Toronto}
+ \streetaddress{40 St. George Street}
+ \city{Toronto}
+ \state{ON}
+ \country{Canada}
+ \postcode{M5S 2E4}
+}
+
+\begin{CCSXML}
+<ccs2012>
+   <concept>
+       <concept_id>10010147</concept_id>
+       <concept_desc>Computing methodologies</concept_desc>
+       <concept_significance>500</concept_significance>
+       </concept>
+   <concept>
+       <concept_id>10010147.10010371.10010352.10010379</concept_id>
+       <concept_desc>Computing methodologies~Physical simulation</concept_desc>
+       <concept_significance>500</concept_significance>
+       </concept>
+ </ccs2012>
+\end{CCSXML}
+
+\ccsdesc[500]{Computing methodologies}
+\ccsdesc[500]{Computing methodologies~Physical simulation}
+
+%
+% End generated code
+%
+
+\keywords{Physically-based animation, NURBS simulation, Meshless simulation, Lagrangian Mechanics}
 

### 2021-04-27 paper/sections/conclusion.tex
fixing typos

@@
 by extending the range of shape matching operations used by the algorithm. The shape matching operation itself could be improved to be material-aware 
-(to better handle heterogenous materials) or to be robust to noisy data (allowing direct simulation of scanned data).
+(to better handle heterogeneous materials) or to be robust to noisy data (allowing direct simulation of scanned data).
 \revise{In future work we also intend to explore using robust solvers or orthogonal bases for the shape matching step itself.} \ty{more can be said here I suppose.
@@
 It is also interesting to consider the relationship between our error term and the orthogonality constraint proposed by~\citet{Zhang:CompDynamics:2020}. 
-Their approach could be used to remove this parameter from SEM entirely.
+Their approach could be used to remove this term from SEM entirely.
 Finally, in this paper we have focused on applying SEM to physics-based animation, and there is a significant amount of additional work needed to 

### 2021-04-27 paper/sections/intro.tex
fixing typos

@@
 NURBS primitives were the first geometric representation used for physics-based animation~\cite{10.1145/37402.37427}, yet,
-despite over three decades of research, animation of curved models such as those built with NURBS, remains a challenge.
+despite over three decades of research, animation of curved models, such as those built with NURBS, remains a challenge.
 
@@
 Our nonlinear elastodynamic simulation scheme requires only a boundary description of the object (we do not strictly require a solid model, see \reffig{teaser}) and
-approriate physical parameters.  
+appropriate physical parameters.  
 Because we explicitly use the NURBS boundary representation in the simulation, it is straightforward to handle
@@
 Crucially, because we broadly target animation and not necessarily simulation for engineering or manufacturing, we don't require that models
-satisfy the rigourous geometric requirements common for these applications.
+satisfy the rigorous geometric requirements common for these applications.
 

### 2021-04-27 paper/sections/methods.tex
fixing typos

@@
 The input to the Shape Matching Element Method (SEM)  is a NURBS \emph{boundary representation} of a volumetric object, along with a set of physical parameters 
-(density, constitutive model, model parameters, etc). 
+(density, constitutive model, model parameters, etc.). 
 The boundary representation is composed of one or more (not necessarily explicitly connected) NURBS surface primitives that we will call \emph{parts}.
 The output is an elastodynamic simulation. 
-SEM consists of preprocessing and runtime simulation phases (~\refalg{sem_preprocess}, ~\refalg{sem_simulation}) and at no point do we need to generate a volumetric mesh of the input. 
+SEM consists of preprocessing and runtime simulation phases (\refalg{sem_preprocess}, ~\refalg{sem_simulation}) and at no point do we need to generate a volumetric mesh of the input. 
 In the \emph{preprocessing} stage we use raycasting to find quadrature points and weights for volumetric integration, as well as to compute part blending parameters. 
@@
 Specifying weights manually would be antithetical to our goal of providing a seamless translation from surface representation to physics-based animation.
-Automatic  weight computation is well-studied (see ~\citet{10.1145/2766952}, ~\citet{BBW:2011} and ~\citet{10.1145/2010324.1964968}), but typically requires solving 
+Automatic  weight computation is well-studied (see ~\cite{10.1145/2766952}, ~\cite{BBW:2011} and ~\cite{10.1145/2010324.1964968}), but typically requires solving 
 an optimization problem that itself relies on a volumetric discretization. 
@@
 \end{equation}
-
 where $\Theta(\mathbf{X}) = \text{diag}\left( \frac{1}{\theta_1},\frac{1}{\theta_2},\dots,\frac{1}{\theta_n}\right)$.
@@
 Each part, $\prt_i$, must be associated with a single deformation origin, $\bar{\refX}_j$, and each origin is computed as the centroid of the set of parts with which it is associated. 
-While we constrain a part to be paired with only one deformation origin, an origin may be associated any number of parts.
+While we constrain a part to be paired with only one deformation origin, an origin may be associated with any number of parts.
 
@@
     \caption{Example structure for the network flow selection of deformation origins with a shape consisting of three parts.
-Red arrows indicate each parts origin association.
+Red arrows indicate each part's origin association.
 \dave{make it clear edges of triangle are sep.
@@
 \begin{aligned}
-\mathbf{f} =  \arg\min_{\mathbf{f}} \quad & \mathbf{c}^T \mathbf{f}    \\
+\mathbf{f} =  \argmin_{\mathbf{f}} \quad & \mathbf{c}^T \mathbf{f}    \\
 \textrm{s.t.} \quad & A_{\text{cons}} \mathbf{f} = 0  \\
@@
 The output of this raycasting procedure is a set of quadrature points $\st{X}$ and weights $\st{v}$ that we use for integration and also as input to our 
-deformation origin extraction (~\refsec{origins}) and blending weights computation (~\refsec{weights})
+deformation origin extraction (\refsec{origins}) and blending weights computation (\refsec{weights})
 \reffig{raycasting} shows the convergence of this quadrature approach when used to compute the volume of a NURBS model. 

### 2021-04-27 paper/sections/results.tex
fixing typos

@@
   \setlength{\tabcolsep}{3pt}
-  \caption{Performance of, and parameters for the Shape Matching Element Method on all examples. All wall-clock timings are reported in seconds, physical parameters are
+  \caption{Performance and parameters for the Shape Matching Element Method on all examples. All wall-clock timings are reported in seconds, physical parameters are
   reported with appropriate units. $\rho$ is the applied density, \textbf{E} is the  Young's Modulus and \revise{$\nu$} is the Poisson's ratio.
@@
 \revise{Additionally, we perform a twisting beam experiment on a 22 part beam model.
-Despite the relatively small number of NURBS patches, we see in \reffig{beam_twist} that the quadratic polynonmials permit complex deformation.}
+Despite the relatively small number of NURBS patches, we see in \reffig{beam_twist} that the quadratic polynomials permit complex deformation.}
 
@@
 
-SEM also supports the simulation of objects made up of heterogenous materials. By specifying different material properties inside nested 
+SEM also supports the simulation of objects made up of heterogeneous materials. By specifying different material properties inside nested 
 parts we are able to simulate a drag racing wheel with a metal hub and soft rubber tread~(\reffig{tire}).
@@
 For downstream applications, the output of  SEM simulation is itself editable in a NURBS modelling program (see \reffig{edit}). 
-This can be useful for visual effects pipelines, allowing artists to easily post process animations using the same tools in which they were created.
+This can be useful for visual effects pipelines, allowing artists to easily post-process animations using the same tools in which they were created.
 \begin{figure}[h]
@@
 In \reffig{starship} we show an animation of the Space-V starship landing on Mars in a graceful way. 
-This example shows the benefit of NURBS modeling, a relatively small number of primitives can represent a complex shape.
+This example shows the benefit of NURBS modeling: a relatively small number of primitives can represent a complex shape.
 SEM successfully handles the stiff materials, complicated geometry and non-trivial collisions in this scene.

### 2021-04-27 paper/sections/results.tex
reviewer typos, removing revision colors from final version

@@
 In \reffig{convergence}, we compare the static configurations of identical cantilevered beams, simulated using a Neohookean constitutive model (Young's Modulus $=$ $0.001$ GPa, Poisson's Ratio $=$ $0.45$).
-\revise{We further perform a basic \emph{quantitative} comparison against FEM by measuring the deflection of the beam at its resting position as we increase the number of patches along the beam.
+\revise{We further perform a basic \emph{quantitative} comparison against FEM by measuring the deflection of the beam at equilibrium as we increase the number of patches along the beam.
 In \reffig{plot_beam_deflection}, the deflection of the SEM beams is reported as a percentage of the FEM beam's deflection.}
@@
   \includegraphics[width=\columnwidth]{figures/deflection_full.pdf}
-	\caption{\revise{The SEM beam deflection is plotted as a percentage of the FEM beam's deflection when the beam is at rest.
+	\caption{\revise{The SEM beam deflection is plotted as a percentage of the FEM beam's deflection when the beam is at equilibrium.
 As we subdivide the SEM beam, its deflection converges to the FEM beam's deflection.

### 2021-04-27 paper/main.tex
squeezing paper to fit 14 pages

@@
 \author{Ty Trusty}
-
 \email{trusty@cs.toronto.edu}
+\affiliation{%
+ \institution{University of Toronto}
+ \streetaddress{40 St. George Street}
+ \city{Toronto}
+ \state{ON}
+ \country{Canada}
+ \postcode{M5S 2E4}
+}
 %
 \author{Honglin Chen}
-
 \email{chl9797@cs.toronto.edu}
+\affiliation{%
+ \institution{University of Toronto}
+ \streetaddress{40 St. George Street}
+ \city{Toronto}
+ \state{ON}
+ \country{Canada}
+ \postcode{M5S 2E4}
+}
 %
@@
 \email{diwlevin@cs.toronto.edu}
-
 \affiliation{%
@@
 
+\begin{acks}
+This work is funded in part by NSERC Discovery (RGPIN-2017-05524), Connaught Fund (503114), CFI-JELF Fund, Accelerator (RGPAS-2017-507909), and the Canada Research Chairs Program.
+We thank Abhishek Madan, Otman Benchekroun, Sarah Kushner, Michael Tao, Nicholas Vining, and Christopher Batty for proofreading; anonymous reviewers for their helpful comments and suggestions.
+\end{acks}
+
 %%

### 2021-04-27 paper/sections/results.tex
squeezing paper to fit 14 pages

@@
 %editing example
-For downstream applications, the output of  SEM simulation is itself editable in a NURBS modelling program (see \reffig{edit}). 
-This can be useful for visual effects pipelines, allowing artists to easily post-process animations using the same tools in which they were created.
+For downstream applications, the output of SEM simulation is itself editable in a NURBS modeling program (see \reffig{edit}). 
+%This can be useful for visual effects pipelines, allowing artists to easily
+% post-process animations using the same tools in which they were created.
+This can allow artists to easily post-process animations using the same tools in which they were created.
 \begin{figure}[h]
@@
 \begin{figure*}[htp]
-  \includegraphics[width=\textwidth,height=3in]{figures/starship.pdf}
+  %\includegraphics[width=\textwidth,height=3in]{figures/starship.pdf}
+  \includegraphics[width=\textwidth]{figures/starship_short.pdf}
   \caption{Our SEM is robust to stiff materials, complicated geometry and non-trivial collisions. Here we show the Space-V starship makes a graceful landing on the rugged surface of Mars. }

### 2021-05-05 paper/sections/related.tex
fix sectioning

@@
 
-\subsection*{Contributions} 
+\subsection{Contributions} 
 In this paper we develop a new Shape Matching-based, Virtual Element Method which is directly compatible 
